### 2016-01-04 natural history of scavenging.tex
Scavenger Review Paper

\documentclass[a4paper,12pt]{article}
\usepackage{ecography}
\usepackage{lmodern}

\title{The natural history of scavenging in vertebrates}
\running{Scavenging in vertebrates}

\author{Adam Kane, \and Kevin Healy, \and Thomas Guillerme \and Graeme Ruxton, \and \& Andrew Jackson.}

\affiliations{
\item A. Kane (\url{adam.kane@ucc.ie}), University College Cork, Cooperage Building\, School of Biological 
Earth and Environmental Sciences\,Cork, Ireland.
\item K. Healy, T. Guillerme and A. Jackson, Trinity College Dublin, Department of Zoology;
School of Natural Sciences, Dublin Ireland.
\item G. Ruxton, School of Biology, Sir Harold Mitchell Building, Greenside Place, St Andrews, KY16 9TH,
United Kingdom
}

\nwords{9999}

\begin{document}

\maketitle

\begin{abstract}
  Scavengers existed in the past and they exist now. 
  Often under appreciated. 
  Three main habitat types considered: land, air and sea. 
  Different drivers in these areas. 
  Review looks at these 
\end{abstract}

\newpage

\section*{Introduction}
Scavenging is a widespread behaviour amonng vertebrates where most if not all carnivores act as facultative scavengers. 
It is recognised that scavengers have an important role in keeping energy flows at a higher trophic level in food webs than decomposers because they consume relatively more carrion \citep{devault2003scavenging}. 
Scavengers also provide useful ecosystem services by acting as barriers to the spread of disease by quickly consuming rotting carcasses which have often died from illness \citep{ogada2012dropping}.
Despite this, scavengers are a seriously understudied group \citep{sekercioglu2006increasing,selva2007nested,wilson2011scavenging}. 
\cite{devault2003scavenging} propose that this is due to both human disgust at carrion itself and the difficulty in determining if an ingested prey item was killed or scavenged. 
The latter point means that studying the natural history of this behaviour is particulalry problematic. Indeed, even data on the proportion of carrion in the diet of extant species are sorely lacking \citep{benbow2015introduction}. 
\\ One avenue to infer scavenging from palaeontological data can be achieved by determining if a prey item was simply too big for the carnivore to have tackled in cases where tooth marks are found \citep{pobiner2008paleoecological}. 
Comparative analysis can also allow us to ascertain which morphologies and physiologies are likely to have been found in scavenging species in the past \citep{ruxton2004obligate}. 
\\ Our review is divided up into three sections, namely the land, air and sea. 
These are then subdivided into three geological eras the Cenozoic, Mesozoic and Palaeozoic. 
Each of these environments has a distinct phyiscal character that affects how a species forages for food which has obvious relevance for an animal searching for carrion.  
However, there is some commonality to a scavenger's environment and the problems in finding food that one would encounter. 
Notably, the resource environment of a scavenger is a patchily distributed one, because it is difficult to predict both when and where a carcass is produced. 
As a result of this, any animal existing as a scavenger must maximise its detection capabilities and minimise its locomotory costs \citep{ruxton2004obligate}.
Exploring which groups are likely to have moved towards these traits and thus existed as sccavengers over palaentological time forms the basis of this work.

\section{Terrestrial Scavengers}
As noted above the ease with which natural selection pushes an animal towards the optima of cheap locomotion and large detection capabilities is dependent on the environment the species is in. 
Land-based scavengers can be thought of as existing in a 2-dimensional plane while foraging for carrion directly. 
The range at which they can detect carcasses is thus defined by the radius of their sensory organs, usually the visual and olfactory senses.

\subsection*{Cenozoic}
Among terrestrial African carnivores, hyenas, jackals, lions and leopards all take sizable proportions of carrion in their diet.
In the case of the spotted hyena (\textit{Crocuta crocuta}) it can be as high as 99\% \citep{benbow2015introduction}. 
Yet, no contemporary terrestrial vertebrate exists as an obligate scavenger. 
The selective pressures that push mammals and reptiles towards scavenging do not seem to undermine their ability to hunt, perhaps explaining the absence of obligate scavengers in these groups \citep{ruxton2004obligate}.
\\There is a long running debate on the tendency of human ancestors to scavenge. 
Some recent studies have found "that passive scavenging from abandoned larger felid kills could have been a high-yield, though potentially dangerous, foraging strategy for early hominins even without considering within-bone nutrients" \citep{pobiner2015new}.

Osteophagy is known across a range of terrestrial carnivores.
Some fat-rich mammalian bones have an energy density (6.7 kJ/g) comparable with that of muscle tissue, making skeletal remains an enticing resource \citep{brown1989study}. 
Hyenas have a bite force capable of breaking open bones and early hominins had the ability to craft tools for this purpose \citep{hone2010feeding,ARCM:ARCM12084}. 
In light of this, the skeletal remains of carrion may act as trove of food to carnivores who can access it.

\cite{ruxton2004obligate} in a theoretical study suggested that "a 1 tonne mammal or reptile, in an ecosystem yielding carrion at densities similar to the current Serengeti, could have met its energy requirements if it could detect carrion over a distance of the order of 400–500 m."

\subsection*{Mesozoic}
In a recent publication a modelling approach found that theropods between 27 kg and 1044 kg would have gained a significant energetic advantage over individuals at both the small and large extremes of theropod body mass through their scavenging efficiency. 
This humped pattern is the result of the disparity between the scaling of energetic cost which scales according to an exponent of 0.91 and energy input that scales according to a cubic polynomial that initially scales according to an exponent of 1.07 but plateaus after 1000kg. 
The polynomial behaviour of energy input is itself the result of the limitations imposed by gut capacity and the overall availability of carcasses after competition.

As we discussed for the case of Cenozoic carnivores, osteophagy could be extremely beneficial to a scavenger. 
In Mesozoic systems some extremely large theropod dinosaurs had a morphology which suggests an ability to process bone e.g. the robust skull and dentition of \textit{Tyrannosaurus rex}. 
There is direct evidence that \textit{T.rex} did this in the form of distinctive wear marks on its tooth apices \citep{farlow1994wear,schubert2005wear} and the presence of bone fragments in its coprolites \citep{chin1998king}. 
The animal also had an enormous bite force, with one estimate putting it at 57000 Newtons \citep{bates2012estimating}. 
This is noted as being powerful enough to break open skeletal material \citep{rayfield2001cranial}. 
Osteophagy may have been even more viable during this era because the body mass distribution of herbviores tended to be skewed towards larger sizes \citep{10.1371/journal.pone.0051925}. 
When we couple this with the fact that skeletal mass scales greater than linearly with body mass \citep{prange1979scaling} there would have been a lot of bones to consume in the environment provided an animal had the biology to process it.

\textit{Allosaurus} tooth marks on a hadrosaur in the Late Jurassic \citep{chure1997one}. 
Late Triassic scavenging on a prosauropod by basal carnivorous archosaurs \citep{hungerbuhler1998taphonomy}.

\subsection*{Palaeozoic}
Synapsids 
Sprawling gait

\section{Aerial Scavengers}
Species capable of flight have effectively added an extra spatial dimension, i.e. the vertical component, to their sensory environment.
This allows them to look down on a landscape where they are unencumbered by obstacles that would obstruct the view of a terrestrial scavenger.
Moreover, flight is a cheaper means of locomotion than running \citep{tucker1975energetic}. 
Thus, it appear that would-be scavengers have a distinct advantage by taking flight.

\subsection*{Cenozoic}
Birds are the dominant vertebrate fliers today and include the best known scavengers on Earth, the vultures. 
These birds consist of two convergent groups, old world and new world vultures and represent the only example of obligate vertebrate scavengers. 
They have a suite of adaptations that allow them to flourish as obligate scavengers.
Vultures extend the energetic advantage of flight further by engaging in soaring instead of flapping flight, which is even cheaper \citep{hedenstrom1993migration}.
Their efficiency is illustrated by cases of predators like bears and wolves benefiting by taking more carrion in their diet in areas bereft of vultures through competitive release \citep{devault2003scavenging}. 
In flight, birds possess a huge advantage over any terrestrial obligate scavenger. 
Flight affords them the ability to range over a much larger area and detect carrion from an elevated vantage point.
\cite{pennycuick1972soaring} conservatively estimated that a Gyps vulture could identify activity at a carcass 4 km away.

Many other bird species take a significant amount of carrion in their diet notably the eagles, storks and corvids. 
Although, none of them are obligate scavengers. 
By contrast with terrestrial species, the traits that render a bird adept at scavenging do undermine its ability to function as an effective predator \citep{devault2003scavenging}.

\subsection*{Mesozoic}
In ancient ecosystems, the volant pterosaurs have also been postulated as existing in a vulture-like niche \citep{witton2008reappraisal}. 
Certain clades of these animals could reach enormous sizes (e.g. Azhdarchids) and look to have engaged in soaring flight. 
However, the inflexibility of their necks and straight, rather than hooked jaw morphology argues against their existing as obligate scavengers \citep{witton2008reappraisal}. 
As yet, no one has ever attempted an energetics approach for this group which is likely due to the many uncertainties over their biology \citep{witton2010size}.

\subsection*{Palaeozoic}
The absence of flying vertebrates in the Palaeozoic may have permitted terrestrial forms to take in a higher proportion of carrion in their diet.

\section{Aquatic Scavengers}
The existence of an obligate scavenger in a marine setting also remains hypothetical \citep{britton1994marine,smith2003ecology,ruxton2004energetic,ruxton2005searching}. 
Carrion in this environment is produced by dead flesh and marine organisms when their carcasses descend to the sea floor. 
This low-light environment means animals detect resources through chemo- and mechanoreception \citep{ruxton2004energetic}. 
Detection distances are far lower than they would be in the air (< 100 m) as a result. 
However, water is a medium that is conducive to low-cost movement \citep{tucker1975energetic} and so may be able to support a small obligate scavenging fish \citep{ruxton2004energetic,ruxton2005searching}. 
Although, for the time being this remains conjectural.

\subsection*{Cenozoic}
A likely instance of scavenging between a 4-million-year-old white shark (\textit{Carcharodon}) and mysticete whale from Peru \citep{ehret2009caught}.
Bite marks on early Holocene Tursiops truncatus fossils from the North Sea indicate scavenging by rays (Chondrichthyes, Rajidae) \citep{van2009bite}. 
Possible scavenging on a juvenile fur seal from the Late Neogene \citep{boessenecker2011mammalian}.

\subsection*{Mesozoic}
Evidence of scavenging in a Cretaceous shark species \textit{Squalicorax} whereby remains of a mosasaur and a hadrosaur were discovered with embedded shark teeth \citep{schwimmer1997scavenging}. 
\subsection*{Palaeozoic}
Evolution of sharks, known scavengers. 
Evidence of vertebrate scavenging dates back to the early Permian approximately 300 MYA \citep{reisz2006articulated}.

\section*{Results}

\section*{Discussion}

\section*{Acknowledgments}

A lot of people are to thank here.

\newpage

\bibliography{bibfile}

\end{document}

### 2016-01-05 natural history of scavenging.tex
Some updates in the intro
Im still getting used to github so let me know if this makes sense

@@
 Scavengers also provide useful ecosystem services by acting as barriers to the spread of disease by quickly consuming rotting carcasses which have often died from illness \citep{ogada2012dropping}.
+%(Since we are intrested in scavanging in th paleo record ecosystem services might not be that relavent, although modulators of disease is still relavent) 
 Despite this, scavengers are a seriously understudied group \citep{sekercioglu2006increasing,selva2007nested,wilson2011scavenging}.  
 \cite{devault2003scavenging} propose that this is due to both human disgust at carrion itself and the difficulty in determining if an ingested prey item was killed or scavenged. 
-The latter point means that studying the natural history of this behaviour is particulalry problematic.
+The latter point means that studying the natural history of this behaviour is particularly problematic.
 Indeed, even data on the proportion of carrion in the diet of extant species are sorely lacking \citep{benbow2015introduction}.
+The limitations in studying extant scavenging behviour is much larger in extinct species and systems with the obvious lack of observational data available.
+This means indirect observations in the fossil record and other approaches such as energetics must be used to infer these behaviours.
+%
 \\ One avenue to infer scavenging from palaeontological data can be achieved by determining if a prey item was simply too big for the carnivore to have tackled in cases where tooth marks are found \citep{pobiner2008paleoecological}.
 Comparative analysis can also allow us to ascertain which morphologies and physiologies are likely to have been found in scavenging species in the past \citep{ruxton2004obligate}.
+The development of indirect measures of scavenging in paleontology can in turn be applied to current scavenging systems that also suffer from a lack of observational data.
+In this review we collate methods (could this be another way of structuring it, just an idea) and research form palaeontology relating to scavenging behaviour and show that ignore this literature would be a missed opportunity for understanding extant scavenging.
+
 \\ Our review is divided up into three sections, namely the land, air and sea. 
@@
 Exploring which groups are likely to have moved towards these traits and thus existed as sccavengers over palaentological time forms the basis of this work.
+%remember the journal is an ecological one at heart so I would always keep in mind about making it useful for ecologists over paelo people.
+In particular from the website "ECOGRAPHY publishes papers focused on broad spatial and temporal patterns, particularly studies of population and community ecology, macroecology, biogeography, and ecological conservation.
+Studies in ecological genetics and historical ecology are welcomed in the context of explaining contemporary ecological patterns".
 

### 2016-01-05 natural history of scavenging.tex
added ref

@@
 Earth and Environmental Sciences\,Cork, Ireland.
-\item K. Healy, T. Guillerme and A. Jackson, Trinity College Dublin, Department of Zoology;
+\item K. Healy and A. Jackson, Trinity College Dublin, Department of Zoology;
 School of Natural Sciences, Dublin Ireland.
+\item T. Guillerme, Imperial College London, Silwood Park Campus, Department of Life Sciences, Buckhurst Road, Ascot SL5 7PY, UK.
+%TG: hehehe!
 \item G. Ruxton, School of Biology, Sir Harold Mitchell Building, Greenside Place, St Andrews, KY16 9TH,
@@
 \section*{Introduction}
-Scavenging is a widespread behaviour amonng vertebrates where most if not all carnivores act as facultative scavengers. 
+%TG: it might be better to reorganise the intro as:
+%§1 - scavenging is understudied because it's hard to say who's a scavenger
+%§2 - it's even more problematic in the fossil record but people have tried some indirect approaches
+%§3 - good thing is that we can apply these indirect approaches to extant systems
+%§4 - in this review we discuss some stuff.
+%TG: or something along these lines, obviously flexible!
+
+%Scavenging is a widespread behaviour among vertebrates where most if not all carnivores act as facultative scavengers.
+%TG: or something more along the lines as:
+Most if not all carnivorous vertebrates are facultative scavenging behaviour to some extant.
 It is recognised that scavengers have an important role in keeping energy flows at a higher trophic level in food webs than decomposers because they consume relatively more carrion \citep{devault2003scavenging}.
@@
 Indeed, even data on the proportion of carrion in the diet of extant species are sorely lacking \citep{benbow2015introduction}.
-The limitations in studying extant scavenging behviour is much larger in extinct species and systems with the obvious lack of observational data available.
-This means indirect observations in the fossil record and other approaches such as energetics must be used to infer these behaviours.
-%
-\\ One avenue to infer scavenging from palaeontological data can be achieved by determining if a prey item was simply too big for the carnivore to have tackled in cases where tooth marks are found \citep{pobiner2008paleoecological}.
+
+%The limitations in studying extant scavenging behaviour is much larger in extinct species and systems with the obvious lack of observational data available.
+%TG: or more like this?
+The limitations in studying scavenging in extant species are even bigger in extinct species and past systems since the obvious lack of available direct behavioural data (CITE) %TG: bet you there's some paper for that
+This means indirect observations in the fossil record and other approaches such as energetics must be used to infer these behaviours. %TG: I'll squiz your Am Nat paper here!
+One avenue to infer scavenging from palaeontological data can be achieved by determining if a prey item was simply too big for the carnivore to have tackled in cases where tooth marks are found \citep{pobiner2008paleoecological}.
 Comparative analysis can also allow us to ascertain which morphologies and physiologies are likely to have been found in scavenging species in the past \citep{ruxton2004obligate}.
-The development of indirect measures of scavenging in paleontology can in turn be applied to current scavenging systems that also suffer from a lack of observational data.
+The development of indirect measures of scavenging in palaeontology can in turn be applied to current scavenging systems that also suffer from a lack of observational data.
+
+%§4
 In this review we collate methods (could this be another way of structuring it, just an idea) and research form palaeontology relating to scavenging behaviour and show that ignore this literature would be a missed opportunity for understanding extant scavenging.
+%TG: totally agree.
+I think we need to clearly know what this review is about: scavenging through time (like a story line of scavenging past and present - a bit boring if I can speak my mind) or how to study scavenging (through time or any other aspect - probably more interesting to more people I guess).
+%TG: additionally I find the divisions land/air/sea and Ceno/Meso/Paleozoic are a bit scholar.
+It might be better to just discuss the different techniques for investigating scavenging and include land/air/sea and Ceno/Meso/Paleo in there no?
+For example:
+%\section{Method 1: direct observation}
+%This can be totally doable in extant system for scavengers everywhere.
+%It is advantageous because it's direct and reliable observations.
+%But it has some limitations such as sending Adam to sit in the sun for ours and is hard to apply in deep sea environements or in any past ecosystems...
 
-\\ Our review is divided up into three sections, namely the land, air and sea. 
+Our review is divided up into three sections, namely the land, air and sea. 
 These are then subdivided into three geological eras the Cenozoic, Mesozoic and Palaeozoic. 
@@
 The range at which they can detect carcasses is thus defined by the radius of their sensory organs, usually the visual and olfactory senses.
+
+Large body size confers many advantages to a scavenger. 
+It grants an animal competitive benefits against other species at a carcass, it allows for some resistance to starvation because of greater fat stores and it often means the sensory organs are elevated giving a higher likelihood of discovering food \citep{farlow1994speculations}.
 

### 2016-01-06 natural history of scavenging.tex
Edits to intro

@@
 The range at which they can detect carcasses is thus defined by the radius of their sensory organs, usually the visual and olfactory senses.
-
-Large body size confers many advantages to a scavenger. 
-It grants an animal competitive benefits against other species at a carcass, it allows for some resistance to starvation because of greater fat stores and it often means the sensory organs are elevated giving a higher likelihood of discovering food \citep{farlow1994speculations}.
 

### 2016-02-25 natural history of scavenging.tex
added snake data and some bits and pieces from Benbow

--- 2016-01-05 natural history of scavenging.tex
@@
 
+"Indeed, in some ecosystems, vertebrates have been documented to assimilate as much as 90\% of the available carrion"\citep{benbow2015introduction}.
+
 %The limitations in studying extant scavenging behaviour is much larger in extinct species and systems with the obvious lack of observational data available.
@@
 Studies in ecological genetics and historical ecology are welcomed in the context of explaining contemporary ecological patterns".
+"Interkingdom Competition among Vertebrates, Invertebrates, and Microbes"\citep{benbow2015introduction}
+"For example, vertebrate scavengers appear to be disadvantaged when humidity and temperature favor microbial and invertebrate reproduction" \citep{benbow2015introduction}
 
@@
 \cite{ruxton2004obligate} in a theoretical study suggested that "a 1 tonne mammal or reptile, in an ecosystem yielding carrion at densities similar to the current Serengeti, could have met its energy requirements if it could detect carrion over a distance of the order of 400–500 m."
+
+A review by \cite{devault2002scavenging} found occurrences of scavenging spread across five families of snakes and states that this behaviour is "far more common than currently acknowledged."
+
+"\cite{sazima1990necrofagia} predict-ed  that  several  components  of  snake  be-havior contribute to scavenging propensity.
+They  postulated  that  snakes  that  rely  on chemosensory information to acquire prey might  forage  for  carrion  more  frequently than  those  whose  modes  of  prey  acquisi-tion are driven by visual cues.
+Additionally, \cite{sazima1990necrofagia} suggested that habitat and diet preferences might be indicators   of   scavenging   propensity   be-cause  they  influence  the  likelihood  of  a particular  species  encountering  carrion naturally.
+Based on these criteria, they pre-dicted that aquatic or semi-aquatic pisciv-orous  snakes  would  scavenge  more  fre-quently than other species.
+Moreover, \cite{sazima1990necrofagia}   suggested that  water  currents  that  induce  aggrega-tions of carrion heighten the probability of carrion detection by these snakes (see also \cite{savitzky1992laboratory}).
+Also,  chemical  gradients may  be  more  uniform  and  give  more  de-pendable  directional information in water \citep{sazima1990necrofagia}." \citep{devault2002scavenging}
+
+"Investigations of ophidian metabolic re-quirements unveil additional advantages to carrion  utilization.
+Snakes  exhibit  exceed-ingly  low  maintenance  metabolisms,  and most  can  survive  on  a  few  scant  feedings per year.
+It  is, therefore, possible for snakes to rely large-ly  on  infrequent,  less  energy-rich  meals.
+Carrion, which is by nature ephemeral and unpredictable, may represent such a food source  to  snakes.
+Scavenging  might  allow snakes to meet their low metabolic needs more  easily  and  without  the  costs  associ-ated with subduing prey." \citep{devault2002scavenging}
 
@@
 These birds consist of two convergent groups, old world and new world vultures and represent the only example of obligate vertebrate scavengers. 
-They have a suite of adaptations that allow them to flourish as obligate scavengers.
+They have a suite of adaptations that allow them to flourish in this niche.
 Vultures extend the energetic advantage of flight further by engaging in soaring instead of flapping flight, which is even cheaper \citep{hedenstrom1993migration}.
@@
 By contrast with terrestrial species, the traits that render a bird adept at scavenging do undermine its ability to function as an effective predator \citep{devault2003scavenging}.
+
+The ability to lower BMR at night is reported in some \textit{Gyps} vultures.
 
@@
 The existence of an obligate scavenger in a marine setting also remains hypothetical \citep{britton1994marine,smith2003ecology,ruxton2004energetic,ruxton2005searching}. 
-Carrion in this environment is produced by dead flesh and marine organisms when their carcasses descend to the sea floor. 
+Carrion in this environment is produced by dead flesh and other marine organisms when their carcasses descend to the sea floor. 
 This low-light environment means animals detect resources through chemo- and mechanoreception \citep{ruxton2004energetic}. 
@@
 Although, for the time being this remains conjectural.
+
+"although some benthic scavengers (e.g., hagfish: family Myxinidae) rely on necrophagy for a large portion of their diet and may indeed be obligate scavengers" \citep{benbow2015introduction}
 

### 2016-03-04 natural history of scavenging.tex
intro update

@@
 %TG: or something along these lines, obviously flexible!
+Historically, scavengers have not been viewed as the most charismatic animals.
+This may go some way to explaining the gap in our knowledge of the prevalence of this behaviour.
+Prof Sanborn Tenney writing in 1877 for The American Naturalist had this to say about one well known group, “Prominent among the mammalian scavengers are the hyenas, the ugliest in their general appearance of all the flesh eaters.” He contrasts these with “nobler kinds” of carnivores such as lions and tigers.
+Even aside from our own subjective biases, scavenging is a difficult behaviour to detect after the fact.
+Without catching a carnivore in the act of killing we are left to infer how the prey was killed.
+Some simple heuristics can inform us, for instance, in cases where the prey item was simply too large to have been killed by the ostensible predator.
+But clearly, a scavenger doesn’t only feed on animals too big for it to have hunted.
+This is significant, because as we are now beginning to realise, the extent of this behaviour is enormous among the carnivores such that, “in some ecosystems, vertebrates have been documented to assimilate as much as 90\% of the available carrion" \citep{benbow2015introduction}.
+Even Tenney’s noble big cats are now known to take in a significant portion of carrion in their diet where some lion populations get over 50\% of their meat from carcasses.
+The obvious lack of direct behavioural data compounds the difficulty of discerning scavenging among extinct forms.
+Indeed, a single species of dinosaur notwithstanding, a synthesis describing the natural history of scavengers is absent from the literature.
 
@@
 
-"\cite{sazima1990necrofagia} predict-ed  that  several  components  of  snake  be-havior contribute to scavenging propensity.
-They  postulated  that  snakes  that  rely  on chemosensory information to acquire prey might  forage  for  carrion  more  frequently than  those  whose  modes  of  prey  acquisi-tion are driven by visual cues.
+"\cite{sazima1990necrofagia} predicted that  several components  of  snake  be-havior contribute to scavenging propensity.
+They  postulated  that  snakes  that  rely  on chemosensory information to acquire prey might forage  for  carrion  more  frequently than  those  whose  modes  of prey acquisition are driven by visual cues.
 Additionally, \cite{sazima1990necrofagia} suggested that habitat and diet preferences might be indicators of scavenging   propensity   be-cause  they  influence  the  likelihood of a particular species encountering carrion naturally.

### 2016-03-07 natural history of scavenging.tex
Added quotes from Benbow

@@
 Studies in ecological genetics and historical ecology are welcomed in the context of explaining contemporary ecological patterns".
-"Interkingdom Competition among Vertebrates, Invertebrates, and Microbes"\citep{benbow2015introduction}
-"For example, vertebrate scavengers appear to be disadvantaged when humidity and temperature favor microbial and invertebrate reproduction" \citep{benbow2015introduction}
+
+%"Interkingdom Competition among Vertebrates, Invertebrates, and Microbes"\citep{benbow2015introduction}
+%"For example, vertebrate scavengers appear to be disadvantaged when humidity and temperature favor microbial and invertebrate reproduction" \citep{benbow2015introduction}
+
+%"However, in general, vertebrate scavenging represents the widest dispersal of nutrients and energy from carcasses as vertebrate movement scales away to the broader landscape" \citep{benbow2015introduction}
+
+ %"Increased kill rates by top predators may represent a little acknowledged marginal cost to the vertebrate community, directly attributable to scavenging activity.
+Given the impor -tance of top-down effects in many ecosystems, even a minor alteration to predation rates as driven by vertebrate scavengers may cause a significant flux in community composition."\citep{benbow2015introduction}
+
+ %"Cortés-Avizanda et al.
+(2009) found that the abundance of prey species (i.e., hares—Lepus spp.
+and squirrels—Sciurus  spp.
+in this case) decreased in sectors containing a carcass based on evidence from tracks in snow.
+An interesting hypothesis emerged, in which the scavengers that are recruited to a carcass may have temporarily played the dual role of increasing predator abundance near each carcass"\citep{benbow2015introduction}
+
+%"Historically, the prevalence of scavenging activities has been greatly underestimated.
+However, upon recognition that (1) in most ecosystems, a large number of animals die from causes other than predation and thus become available to scavengers; (2) most carcasses are scavenged by vertebrates before they are completely decomposed by arthropods and bacteria; and (3) almost all carnivorous animals are facultative scavengers, the importance of scavenging in food webs seems unsurprising" \citep{benbow2015introduction}
+
+%"This dispersion of carrion biomass by vertebrates is especially evident when carrion is initially concentrated spatially.
+For example, carcasses produced from fishing by-catch (Furness 2003), salmon (Salmonidae) die-offs (Hewson 1995), forest fires (Blanchard and Knight 1990), and single large carcasses (e.g., whales—Cetacea; Smith and Baco 2003) are often visited by multiple scavengers that range widely and therefore transport the nutrients from those carcasses over large distances."\citep{benbow2015introduction}
+
+%"Cross-habitat nutrient transport can produce a variety of important outcomes in recipient systems (e.g., Polis et al.
+2004), and scavengers can play a significant role in moving “ecologi -cal subsidies” between habitats.
+For example, the use of ocean-derived carrion by terrestrial mammals (Rose and Polis 1998) and birds (Schlacher et al.
+2013) is extensive and may strongly influence dynamics of coastal food webs."\citep{benbow2015introduction}
+
+% "Markandya et  al.
+(2008) estimated that the total costs to human health (including rabies cases from dog bites) that resulted from severe vulture declines totaled over $34 billion from 1993 to 2006.
+Also, Ogada et al.
+(2012) determined that the exclusion of vultures from large animal carcasses in Kenya resulted in a tripling of carcass decomposition times."\citep{benbow2015introduction}
+
+% "Changes in global climate are also expected to substantially alter the availability of carrion resources in marine ecosystems as surface production of organic material could decline by 50% or more (Smith et al.
+2008).
+Such a drastic reduction in productivity undoubtedly will impact a multitude of ecosystem processes, including the availability and distribution of carrion within marine ecosystems." \citep{benbow2015introduction}
 
@@
 
+%"Of all the mammalian carnivores in Africa, brown, striped, and spotted hyenas (hereafter referred to simply as hyenas) derive the largest portion of their diet from scavenging.
+This is not surprising, as they show many unique adaptations specifically for scavenging.
+They have a unique body posture and a “rocking-horse” gait (Eloff 1964; Tilson and Henschel 1986; Hofer and East 1993a; Frank 1996) that allows them to cover long distances in search of carrion and prey."\citep{benbow2015introduction}
+
 A review by \cite{devault2002scavenging} found occurrences of scavenging spread across five families of snakes and states that this behaviour is "far more common than currently acknowledged."
 
-"\cite{sazima1990necrofagia} predicted that  several components  of  snake  be-havior contribute to scavenging propensity.
+%"\cite{sazima1990necrofagia} predicted that  several components  of  snake  be-havior contribute to scavenging propensity.
 They  postulated  that  snakes  that  rely  on chemosensory information to acquire prey might forage  for  carrion  more  frequently than  those  whose  modes  of prey acquisition are driven by visual cues.
@@
 
-"Investigations of ophidian metabolic re-quirements unveil additional advantages to carrion  utilization.
+%"Investigations of ophidian metabolic re-quirements unveil additional advantages to carrion  utilization.
 Snakes  exhibit  exceed-ingly  low  maintenance  metabolisms,  and most  can  survive  on  a  few  scant  feedings per year.

### 2016-03-11 natural history of scavenging.tex
Added new section
Aerial written and terrestrial started

@@
 %TG: or something along these lines, obviously flexible!
-Historically, scavengers have not been viewed as the most charismatic animals.
+Historically, scavengers have not been viewed as the most charismatic of animals.
 This may go some way to explaining the gap in our knowledge of the prevalence of this behaviour.
-Prof Sanborn Tenney writing in 1877 for The American Naturalist had this to say about one well known group, “Prominent among the mammalian scavengers are the hyenas, the ugliest in their general appearance of all the flesh eaters.” He contrasts these with “nobler kinds” of carnivores such as lions and tigers.
+Consider Professor Sanborn Tenney writing in 1877 for The American Naturalist who had this to say about one well known group, “Prominent among the mammalian scavengers are the hyenas, the ugliest in their general appearance of all the flesh eaters.” He contrasts these with “nobler kinds” of carnivores such as lions and tigers.
 Even aside from our own subjective biases, scavenging is a difficult behaviour to detect after the fact.
 Without catching a carnivore in the act of killing we are left to infer how the prey was killed.
-Some simple heuristics can inform us, for instance, in cases where the prey item was simply too large to have been killed by the ostensible predator.
+Some simple heuristics can inform us, for instance, in cases where the prey item was simply too large to have been killed by the ostensible predator \citep{pobiner2008paleoecological}.
 But clearly, a scavenger doesn’t only feed on animals too big for it to have hunted.
-This is significant, because as we are now beginning to realise, the extent of this behaviour is enormous among the carnivores such that, “in some ecosystems, vertebrates have been documented to assimilate as much as 90\% of the available carrion" \citep{benbow2015introduction}.
-Even Tenney’s noble big cats are now known to take in a significant portion of carrion in their diet where some lion populations get over 50\% of their meat from carcasses.
 The obvious lack of direct behavioural data compounds the difficulty of discerning scavenging among extinct forms.
 Indeed, a single species of dinosaur notwithstanding, a synthesis describing the natural history of scavengers is absent from the literature.
+Fortunately, research in this area is on the rise.
+As a result we are now beginning to realise the extent of this behaviour such that, “in some ecosystems, vertebrates have been documented to assimilate as much as 90\% of the available carrion" \citep{benbow2015introduction}.
+Even Tenney’s noble big cats are now known to take in a significant portion of carrion in their diet where some lion populations get over 50\% of their meat from carcasses.
+A suite of methods have been used to discern the most suitable morphologies, physiologies and environments for a scavenging lifestyle to prosper.
+It is our aim in this review to employ these methods to gain an understanding of scavengers past and present.
+
+The chief hurdle to scavenging is finding a sufficient quantity of food, the occurence of which is difficult to predict in space and time.
+The idea of scrounging from predator kills is undermined from studies showing that in the  majority of ecosystems more animals die from disease and starvation than predation \citep{benbow2015introduction}.
+Thus, any animal existing as a scavenger must maximise its detection capabilities and minimise its locomotory costs \citep{ruxton2004obligate}.
+The habitat must also be productive enough to sustain an animal biomass that will eventually produce carcasses.
+
+\section{Aerial Scavengers}
+
+Vultures represent the best known scavengers on Earth.
+These birds consist of two convergent groups, old world and new world vultures and represent the only example of obligate vertebrate scavengers today.
+Given their unique position, they have been extensively studied to determine what adaptations they possess that allows them to so flourish in this niche.
+As such, we can begin by exploring the adapdations and the environments of vultures to draw comparisons with other scavenging species and \textit{their} environments.
+
+Species capable of flight have effectively added an extra spatial dimension, i.e.
+the vertical component, to their sensory environment over land animals.
+This allows them to look down on a landscape where they are unencumbered by obstacles that would obstruct the view of a terrestrial scavenger.
+Such an ability has obvious benefits in detecting carrion.
+Vultures are known to have impressive visual acuity with one estimate indicating Lappet-faced Vultures (\textit{Torgos tracheliotus}) are capable of detecting a 2 metre carcass over 10 km away \citep{spiegel2013factors}.
+We know that many birds exist as facultative scavengers; storks, eagles, corvids, are all known to take substantial quantities of carrion in their diet.
+And eagles in particular are known to have highly developed visual abilities.
+It follows from this that the evolution of flight allowed aerial animals to detect far more carrion than their terrestrial counterparts.
+
+Moreover, having a panoramic view means being able to gather a wealth of information from other foragers, be they conspecifics or other species.
+Again, returing to vultures, the genus \textit{Gyps} consists of highly social and colonially nesting species.
+These behaviours allow them forage far more efficiently because one bird can scrounge information on the location of food from another successful forager.
+
+Flight is also cheaper means of locomotion than running \citep{tucker1975energetic}.
+This advantage can be extended further in larger species by engaging in soaring instead of flapping flight, which is even cheaper still (~ twice BMR) \citep{hedenstrom1993migration}.
+The advantages this confers are clear from the information we have on the enormous foraging ranges of some seabirds and accipiters.
+Indeed, it would be pointless to have incredible detection abilites and not have a cost efficient movement to benefit from it.
+
+Avian flight originates in the Jurassic Period, conincident with the fossils of \textit{Archaeopteryx lithographica} so many of these benefits would have been realised from that point on for carnivorous birds.
+However, vertebrate flight is much older than this where pterosaurs predate bird origins by a considerable margin (the Triassic Period).
+Scavenging in this diverse group has been hypothesied many times before.
+Certain clades of these animals could reach enormous sizes (e.g.
+Azhdarchids with wingspans of 11 metres) and look to have engaged in soaring flight.
+Although \cite{witton2008reappraisal} argued that neck inflexibility and straight, rather than hooked jaw morphology points against their existing as \textit{obligate} scavengers, Azhdarchid terrestrial proficency indicates they would have been comfortable foraging on the ground.
+Indeed, extant Marabou Storks have a comparable morphology and are noted facultative scavengers so it is reasonable to believe that certain pterosaurs behaved similarly.
+
+Large body size confers substantial dominance benefits \citep{ruxton2004obligate}.
+Thus we would expect scavengers to have this trait selected for even in the case of weight-constrained fliers.
+Cinereous Vultures (\textit{Aegypius monachus}) and condors (\textit{Vultur gryphus}, \textit{Gymnogyps californianus}) all have body masses that can exceed 10 kg and represent some of the heaviest bird species capable of flight \citep{ferguson2001raptors,donazar2002effects}.
+And as we have noted the Azhdarchid pterosaurs were far bigger again, with estimated body masses of around 80 kg.
+
+The only other vertebrate group capable of powered flight are the bats where scavenging has not been recorded to our knowledge.
+Their visual acuity is famously poor and echolocation does not lend itself to discovering immobile carrion.
+Their small size and poor terrestrial ability would also count against them at a carcass.
+The bat fossil record is notoriously poor owing to their fragile skeletons so we are unable to determine if extinct species were more suited to this lifestyle.
+
+Vertebrate scavengers in general are responsible for the dispersal of nutrients.
+Consider the diversity of animals that can end up feeding at the carcass of an elephant.
+Here we have an incredibly dense and nutrient rich patch that ends up being distributed widely.
+Thus, in an ecological context the evolution of flight coupled with an ability of scavenge resulted in a world with a far more widely distributed nutrient landscape.
+In the absence of vertebrate scavengers, invertebrates and microorganisms would consume the carcass in-situ or at least distribute the constituent nutrients over a much shorter range.
+
+\section{Terrestrial Scavengers}
+Terrestrial scavengers can be thought of as existing in a two-dimensional plane while foraging for carrion directly. 
+They can detect carcasses at a range that is defined by the radius of their sensory organs, usually the visual and olfactory senses.
+As a consequence they have a much more restricted view of the landscape than do aerial foragers.
+No contemporary terrestrial vertebrate exists as an obligate scavenger but most if not all carnivorous vertebrates are facultative scavenging behaviour to some extent.
+Terrstrial scavenging in the mammals is probably best known in an African context where hyenas, jackals and lions all take sizable proportions of carrion in their diet.
+In the case of the spotted hyena (\textit{Crocuta crocuta}) it can be as high as 99\% \citep{benbow2015introduction}.
+We can again use this species as our efficient terrestrial scavenger to compare with other species.
+
+Similar to vultures they have well developed sensory organs, particularly in olfcation whereby they can detect a rotting carcass 2 km downwind.
+They have a characteristic "rocking horse gait"  which allows them to cover great distances.
+"Apart from being an important water-conservation strategy, nocturnal behavior may have evolved in the Hyaenidae as a means of reducing competition with the other dominant scavengers in African ecosystems, the vultures, which are exclusivly diurnal (Houston 1979)" Carnivore Behavior, Ecology, and Evolution By John L.
+Gittleman
 
@@
 
-\section{Aerial Scavengers}
+%\section{Aerial Scavengers}
 Species capable of flight have effectively added an extra spatial dimension, i.e. the vertical component, to their sensory environment.
@@
 
-\subsection*{Cenozoic}
+%\subsection*{Cenozoic}
 Birds are the dominant vertebrate fliers today and include the best known scavengers on Earth, the vultures. 
@@
 
-\subsection*{Mesozoic}
+%\subsection*{Mesozoic}
 In ancient ecosystems, the volant pterosaurs have also been postulated as existing in a vulture-like niche \citep{witton2008reappraisal}. 
@@
 
-\subsection*{Palaeozoic}
+%\subsection*{Palaeozoic}
 The absence of flying vertebrates in the Palaeozoic may have permitted terrestrial forms to take in a higher proportion of carrion in their diet.

### 2016-03-14 natural history of scavenging.tex
More sections
Wrote terrestrial and started aquatic

@@
 \usepackage{lmodern}
+
+ \renewcommand{\familydefault}{\sfdefault}
 
@@
 The habitat must also be productive enough to sustain an animal biomass that will eventually produce carcasses.
+It is recognised that scavengers keep energy flows at a higher trophic level in food webs than decomposers because they consume relatively more carrion \citep{devault2003scavenging}.
 
-\section{Aerial Scavengers}
+\section*{Aerial Scavengers}
 
 Vultures represent the best known scavengers on Earth.
-These birds consist of two convergent groups, old world and new world vultures and represent the only example of obligate vertebrate scavengers today.
+These birds consist of two convergent groups, from the old and the new world and represent the only example of obligate vertebrate scavengers today.
 Given their unique position, they have been extensively studied to determine what adaptations they possess that allows them to so flourish in this niche.
@@
 Flight is also cheaper means of locomotion than running \citep{tucker1975energetic}.
-This advantage can be extended further in larger species by engaging in soaring instead of flapping flight, which is even cheaper still (~ twice BMR) \citep{hedenstrom1993migration}.
+This advantage can be extended further in larger species by engaging in soaring instead of flapping flight, which is even cheaper still (approximately twice BMR) \citep{hedenstrom1993migration}.
 The advantages this confers are clear from the information we have on the enormous foraging ranges of some seabirds and accipiters.
-Indeed, it would be pointless to have incredible detection abilites and not have a cost efficient movement to benefit from it.
+Clearly, it would be pointless to have incredible detection abilites and not have a cost efficient movement to benefit from it.
 
 Avian flight originates in the Jurassic Period, conincident with the fossils of \textit{Archaeopteryx lithographica} so many of these benefits would have been realised from that point on for carnivorous birds.
-However, vertebrate flight is much older than this where pterosaurs predate bird origins by a considerable margin (the Triassic Period).
+However, vertebrate flight is much older than this where pterosaurs predate bird origins by a considerable margin in the Triassic Period.
 Scavenging in this diverse group has been hypothesied many times before.
@@
 Large body size confers substantial dominance benefits \citep{ruxton2004obligate}.
-Thus we would expect scavengers to have this trait selected for even in the case of weight-constrained fliers.
+Thus, we would expect scavengers to have this trait selected for even in the case of weight-constrained fliers.
 Cinereous Vultures (\textit{Aegypius monachus}) and condors (\textit{Vultur gryphus}, \textit{Gymnogyps californianus}) all have body masses that can exceed 10 kg and represent some of the heaviest bird species capable of flight \citep{ferguson2001raptors,donazar2002effects}.
@@
 Here we have an incredibly dense and nutrient rich patch that ends up being distributed widely.
-Thus, in an ecological context the evolution of flight coupled with an ability of scavenge resulted in a world with a far more widely distributed nutrient landscape.
+Thus, in an ecological context, the evolution of flight coupled with the ability to scavenge resulted in a world with a far more widely distributed nutrient landscape.
 In the absence of vertebrate scavengers, invertebrates and microorganisms would consume the carcass in-situ or at least distribute the constituent nutrients over a much shorter range.
 
-\section{Terrestrial Scavengers}
-Terrestrial scavengers can be thought of as existing in a two-dimensional plane while foraging for carrion directly. 
+\section*{Terrestrial Scavengers}
+A simplificaiton of terrestrial scavengers is one of them existing in a two-dimensional plane while foraging for carrion directly.
 They can detect carcasses at a range that is defined by the radius of their sensory organs, usually the visual and olfactory senses.
-As a consequence they have a much more restricted view of the landscape than do aerial foragers.
-No contemporary terrestrial vertebrate exists as an obligate scavenger but most if not all carnivorous vertebrates are facultative scavenging behaviour to some extent.
-Terrstrial scavenging in the mammals is probably best known in an African context where hyenas, jackals and lions all take sizable proportions of carrion in their diet.
-In the case of the spotted hyena (\textit{Crocuta crocuta}) it can be as high as 99\% \citep{benbow2015introduction}.
-We can again use this species as our efficient terrestrial scavenger to compare with other species.
+As a consequence, they have a much more restricted view of the landscape than do aerial foragers.
+No contemporary terrestrial vertebrate exists as an obligate scavenger but most if not all are facultative scavengers to some extent.
+\cite{ruxton2004obligate} offer a reason for this in that the traits that allow for vultures to exist as scavengers undermined their ability to hunt but that the same forces have not prevented mammals from doing so.
+The same authors in a theoretical study do concede that "a 1 tonne mammal or reptile, in an ecosystem yielding carrion at densities similar to the current Serengeti, could have met its energy requirements if it could detect carrion over a distance of the order of 400–500 m."\citep{ruxton2004obligate}.
 
-Similar to vultures they have well developed sensory organs, particularly in olfcation whereby they can detect a rotting carcass 2 km downwind.
-They have a characteristic "rocking horse gait"  which allows them to cover great distances.
-"Apart from being an important water-conservation strategy, nocturnal behavior may have evolved in the Hyaenidae as a means of reducing competition with the other dominant scavengers in African ecosystems, the vultures, which are exclusivly diurnal (Houston 1979)" Carnivore Behavior, Ecology, and Evolution By John L.
+Terrestrial scavenging in the mammals is probably best known in an African context where hyenas, jackals and lions all take sizable proportions of carrion in their diet.
+In the spotted hyena (\textit{Crocuta crocuta}), striped hyena (\textit{Hyaena hyaena}) and brown hyena (\textit{Hyaena brunnea}) it can be as high as 99\% \citep{benbow2015introduction}.
+Therefore, we can again use these species as our efficient terrestrial scavengers to compare with other forms.
+
+Similar to vultures they have well developed sensory organs, particularly in olfaction whereby they can detect a rotting carcass 2 km downwind.
+They have a characteristic "rocking horse gait"  which allows them to cover great distances efficiently.
+The bone crushing ability of hyenas reveals another useful scavenger trait.
+Since carrion is not dispatched directly, often the most easily accessible and choicest components of the carcass will be missing or, if present, will be fought over.
+Being able to extract nutrients from remnants gives the scavenger a great advantage.
+Osteophagy is known across a range of terrestrial carnivores.
+Some fat-rich mammalian bones have an energy density (6.7 kJ/g) comparable with that of muscle tissue, making skeletal remains an enticing resource \citep{brown1989study}.
+This ability reached its zenith among hyenas with the evolution of the 110 kg \textit{Pachycrocuta brevirostris} during the Pliocene \citep{palmqvist2011giant}.
+The ability to process bone means a carcass fed on by hyenas will be reduced to nothing, whereas the skeleton will remain in carrion attended by species restriced to feeding on the flesh.
+
+Many of these adapations to scavenging are found in the other major extant terrestrial mammalian carnviores, the bears, dogs and cats to a greater or lesser extent.
+Though the specific mix of features realised in hyenas suggest this is the model organism for terrestrial scavenging among mammals in the past.
+Indeed, the bone-crushing dogs that evolved during the Oligocene (subfamily Borophaginae) have been compared to hyenas in terms of their feeding ecology \citep{van2003chapter,martin2016pursuit}.
+Interestingly such comparisons have given insight into the feeding ecology of early hominins who, for instance, had the ability to craft tools for breaking open bones \citep{hone2010feeding,ARCM:ARCM12084}.
+
+By contrast, a successful reptilian scavenger requires a far different set of adapations.
+Modern forms are ectothermic, limiting their activity periods.
+This is exacerbated by the sprawling gait seen in lizards which results in Carrier's Constraint such that the animal can't move and breathe at the same time because the lateral movements impedes its lungs.
+This manifests itself in aspects such as maximum sustainable speed where an equivalent mammal has a six to seven fold increase \citep{ruben1995evolution}.
+A lower metabolism does give reptiles an advantage however, in that over the course of a year their food requirements can be 30 times smaller than an endotherm of equal size \citep{Nagy1621}.
+Any adaptations that reduce energetic costs are likely to be selected in scavengers.
+\cite{devault2002scavenging} suggest this is an avenue for scavenging in snakes because they "exhibit  exceedingly  low  maintenance  metabolisms,  and most  can  survive  on  a  few  scant  feedings per year.
+It  is, therefore, possible for snakes to rely largely  on  infrequent,  less  energy-rich  meals." In the same review the authors found occurrences of scavenging spread across five families of snakes and stated that this behaviour is "far more common than currently acknowledged."\citep{devault2002scavenging}.
+
+Unsurprisingly, given their enduring appeal, the prevalence of scavenging has been explored in the carnivorous, theropod dinosaurs.
+These animals ranged from the chicken-sized to the whale-sized all of which were bipedal.
+They are quite alien to anything we know today which restricts our ability to understand their ecology far more so than extinct mammals \citep{weishampel2004dinosauria}.
+Of relevance, are the questions that still persist about their metabolism \citep{grady2014evidence} and sensory perception.
+We do know that they walked with the erect gait of mammals or birds rather than the sprawling gait of lizards and that they were most likely facultative scavengers \citep{depalma2013physical}.
+Much work has focused on the existence of the behaviour in \textit{Tyrannosaurus rex} \citep{ruxton2003could,carbone2011intra} but a recent energetics study investigated the likely prevelance of scavenging across a range of body sizes.
+In it the authors demonstrated that species of intermediate body masses (approx.
+500 kg) would have gained the most benefit from scavenging.
+This was the result of gut capacity limitations and the effects of competition at the carcass.
+At the larger extreme this owes to the fact that gut capacity doesn't scale isometrically with body mass so the benefits of greater mass level off; there's only so much food an individual can consume at a single sitting.
+For the smaller species, larger competitors would have prevented their access to carrion.
+
+As we discussed for the case of Cenozoic carnivores, osteophagy could be extremely beneficial to a scavenger. 
+In Mesozoic systems some extremely large theropod dinosaurs had a morphology which suggests an ability to process bone e.g.
+the robust skull and dentition of \textit{T rex}.
+There is direct evidence that \textit{T.rex} did this in the form of distinctive wear marks on its tooth apices \citep{farlow1994wear,schubert2005wear} and the presence of bone fragments in its coprolites \citep{chin1998king}.
+The animal also had an enormous bite force, with one estimate putting it at 57000 Newtons \citep{bates2012estimating}.
+This is noted as being powerful enough to break open skeletal material \citep{rayfield2001cranial}.
+Osteophagy may have been even more viable during this era because the body mass distribution of herbviores tended to be skewed towards larger sizes \citep{10.1371/journal.pone.0051925}.
+When we couple this with the fact that skeletal mass scales greater than linearly with body mass \citep{prange1979scaling} there would have been a lot of bone material to consume in the environment provided an animal had the biology to process it  \citep{chure1997one}.
+
+Evidence of vertebrate scavenging dates back to the early Permian approximately 300 MYA \citep{reisz2006articulated}.
+
+\section*{Aquatic Scavengers} 
+An aquatic environment presents challenges for direct observational studies and so, similar to the approaches involving extinct species, much work has approached the question of scavenging propensity from an energetics perspective.
+The existence of an obligate scavenger in a marine setting is uncertain \citep{britton1994marine,smith2003ecology,ruxton2004energetic,ruxton2005searching}.
+Carrion in this environment is produced by marine organisms when their carcasses descend to the sea floor.
+In this low-light environment detection distances are far lower (< 100 m) than they would be in the air.
+As such, animals detect resources through chemo- and mechanoreception more so than through vision \citep{ruxton2004energetic}.
+However, water is a medium that is conducive to low-cost movement \citep{tucker1975energetic} and so may be able to support an obligate scavenging fish \citep{ruxton2004energetic,ruxton2005searching}.
+\cite{benbow2015introduction} do note that "some benthic scavengers (e.g., hagfish: family Myxinidae) rely on necrophagy for a large portion of their diet and may indeed be obligate scavengers".
+
+Extant aquatic snakes are deemed as having the most suitable physiology and environment for scavenging.
+A hypothesis put forth by \citep{sazima1990necrofagia} argued that chemical gradients in water would allow for a relatively easier detection of carrion.
+This gained some support from \cite{devault2002scavenging}, who found a preponderence of aquatic snake species in their review of this behaviour.
+
+The presence of occasional bounties of carrion in the form of whale falls has led some researchers to investigate if a scavenger could survive by seeking out these remains exclusively.
+\cite{ruxton2005searching} argued that although this is energetically feasible it's ecologically unlikely.
+Any animal that could seek out such whale carcasses is unlikely to have ignored other types of carrion.
+Although no aquatic species have ever exceeded the size of whales, some enormous animals have evolved in this environment before the evolution of whales, including \textit{Leedsichthys}, a bony fish from the Jurassic Period, that weighed in excess of 20 tonnes.
+Thus, the energetic feasiblity of a marine scavenger has a long history.
+
+As with the aerial and terrestrial enviornments we have evidence of facultative scavenging among extinct, aquatic species.
+For example, the remains of a mosasaur and a terrestrial hadrosaur were discovered with embedded teeth from a Cretaceous shark \textit{Squalicorax} \citep{schwimmer1997scavenging}.
+As well as a likely instance of scavenging between a 4-million-year-old white shark (\textit{Carcharodon}) and mysticete whale from Peru \citep{ehret2009caught}.
+
+ %\citep{sazima1990necrofagia} predicted that aquatic or semi-aquatic piscivorous snakes  would  scavenge  more  frequently than other species.
+Moreover, \cite{sazima1990necrofagia}  suggested that water currents that induce  aggregations of carrion heighten the probability of carrion detection by these snakes (see also \cite{savitzky1992laboratory}).
+Also,  chemical  gradients may  be  more  uniform  and  give  more  de-pendable  directional information in water \citep{sazima1990necrofagia}." \citep{devault2002scavenging}
+
+%"Apart from being an important water-conservation strategy, nocturnal behavior may have evolved in the Hyaenidae as a means of reducing competition with the other dominant scavengers in African ecosystems, the vultures, which are exclusivly diurnal (Houston 1979)" Carnivore Behavior, Ecology, and Evolution By John L.
 Gittleman
+
+% "Thus, on an annual basis, the FMR (and food requirements) of an endotherm are likely to be around 30 times greater than those for an equivalent-sized ectotherm"
+
+%"Maximal sustainable speeds of mammals are six- to sevenfold those of lizards of equal size,"
 
@@
 %TG: or something more along the lines as:
-Most if not all carnivorous vertebrates are facultative scavenging behaviour to some extant.
-It is recognised that scavengers have an important role in keeping energy flows at a higher trophic level in food webs than decomposers because they consume relatively more carrion \citep{devault2003scavenging}.
-Scavengers also provide useful ecosystem services by acting as barriers to the spread of disease by quickly consuming rotting carcasses which have often died from illness \citep{ogada2012dropping}.
+%Most if not all carnivorous vertebrates are facultative scavenging behaviour to some extant.
+%It is recognised that scavengers have an important role in keeping energy flows at a higher trophic level in food webs than decomposers because they consume relatively more carrion \citep{devault2003scavenging}.
+%Scavengers also provide useful ecosystem services by acting as barriers to the spread of disease by quickly consuming rotting carcasses which have often died from illness \citep{ogada2012dropping}.
 %(Since we are intrested in scavanging in th paleo record ecosystem services might not be that relavent, although modulators of disease is still relavent) 
-Despite this, scavengers are a seriously understudied group \citep{sekercioglu2006increasing,selva2007nested,wilson2011scavenging}.  
-\cite{devault2003scavenging} propose that this is due to both human disgust at carrion itself and the difficulty in determining if an ingested prey item was killed or scavenged. 
-The latter point means that studying the natural history of this behaviour is particularly problematic.
-Indeed, even data on the proportion of carrion in the diet of extant species are sorely lacking \citep{benbow2015introduction}.
+%Despite this, scavengers are a seriously understudied group \citep{sekercioglu2006increasing,selva2007nested,wilson2011scavenging}.  
+%\cite{devault2003scavenging} propose that this is due to both human disgust at carrion itself and the difficulty in determining if an ingested prey item was killed or scavenged. 
+%The latter point means that studying the natural history of this behaviour is particularly problematic.
+%Indeed, even data on the proportion of carrion in the diet of extant species are sorely lacking \citep{benbow2015introduction}.
 
-"Indeed, in some ecosystems, vertebrates have been documented to assimilate as much as 90\% of the available carrion"\citep{benbow2015introduction}.
+%"Indeed, in some ecosystems, vertebrates have been documented to assimilate as much as 90\% of the available carrion"\citep{benbow2015introduction}.
 
@@
 %TG: or more like this? 
-The limitations in studying scavenging in extant species are even bigger in extinct species and past systems since the obvious lack of available direct behavioural data (CITE) %TG: bet you there's some paper for that
-This means indirect observations in the fossil record and other approaches such as energetics must be used to infer these behaviours. %TG: I'll squiz your Am Nat paper here!
-One avenue to infer scavenging from palaeontological data can be achieved by determining if a prey item was simply too big for the carnivore to have tackled in cases where tooth marks are found \citep{pobiner2008paleoecological}.
-Comparative analysis can also allow us to ascertain which morphologies and physiologies are likely to have been found in scavenging species in the past \citep{ruxton2004obligate}.
-The development of indirect measures of scavenging in palaeontology can in turn be applied to current scavenging systems that also suffer from a lack of observational data.
+% The limitations in studying scavenging in extant species are even bigger in extinct species and past systems since the obvious lack of available direct behavioural data (CITE) %TG: bet you there's some paper for that
+% This means indirect observations in the fossil record and other approaches such as energetics must be used to infer these behaviours. %TG: I'll squiz your Am Nat paper here!
+% One avenue to infer scavenging from palaeontological data can be achieved by determining if a prey item was simply too big for the carnivore to have tackled in cases where tooth marks are found \citep{pobiner2008paleoecological}.
+% Comparative analysis can also allow us to ascertain which morphologies and physiologies are likely to have been found in scavenging species in the past \citep{ruxton2004obligate}.
+% The development of indirect measures of scavenging in palaeontology can in turn be applied to current scavenging systems that also suffer from a lack of observational data.
 
 %§4
-In this review we collate methods (could this be another way of structuring it, just an idea) and research form palaeontology relating to scavenging behaviour and show that ignore this literature would be a missed opportunity for understanding extant scavenging.
+% In this review we collate methods (could this be another way of structuring it, just an idea) and research form palaeontology relating to scavenging behaviour and show that ignore this literature would be a missed opportunity for understanding extant scavenging.
 %TG: totally agree.
@@
 
-Our review is divided up into three sections, namely the land, air and sea. 
-These are then subdivided into three geological eras the Cenozoic, Mesozoic and Palaeozoic. 
-Each of these environments has a distinct phyiscal character that affects how a species forages for food which has obvious relevance for an animal searching for carrion.  
-However, there is some commonality to a scavenger's environment and the problems in finding food that one would encounter. 
-Notably, the resource environment of a scavenger is a patchily distributed one, because it is difficult to predict both when and where a carcass is produced. 
-As a result of this, any animal existing as a scavenger must maximise its detection capabilities and minimise its locomotory costs \citep{ruxton2004obligate}.
-Exploring which groups are likely to have moved towards these traits and thus existed as sccavengers over palaentological time forms the basis of this work.
+%Our review is divided up into three sections, namely the land, air and sea. 
+%These are then subdivided into three geological eras the Cenozoic, Mesozoic and Palaeozoic. 
+%Each of these environments has a distinct phyiscal character that affects how a species forages for food which has obvious relevance for an animal searching for carrion.  
+%However, there is some commonality to a scavenger's environment and the problems in finding food that one would encounter. 
+%Notably, the resource environment of a scavenger is a patchily distributed one, because it is difficult to predict both when and where a carcass is produced. 
+%As a result of this, any animal existing as a scavenger must maximise its detection capabilities and minimise its locomotory costs \citep{ruxton2004obligate}.
+%Exploring which groups are likely to have moved towards these traits and thus existed as sccavengers over palaentological time forms the basis of this work.
 %remember the journal is an ecological one at heart so I would always keep in mind about making it useful for ecologists over paelo people.
@@
 
-\section{Terrestrial Scavengers}
-As noted above the ease with which natural selection pushes an animal towards the optima of cheap locomotion and large detection capabilities is dependent on the environment the species is in. 
-Land-based scavengers can be thought of as existing in a 2-dimensional plane while foraging for carrion directly. 
-The range at which they can detect carcasses is thus defined by the radius of their sensory organs, usually the visual and olfactory senses.
+%\subsection*{Cenozoic}
+%Among terrestrial African carnivores, hyenas, jackals, lions and leopards all take sizable proportions of carrion in their diet.
+%In the case of the spotted hyena (\textit{Crocuta crocuta}) it can be as high as 99\% \citep{benbow2015introduction}. 
+%Yet, no contemporary terrestrial vertebrate exists as an obligate scavenger. 
+%The selective pressures that push mammals and reptiles towards scavenging do not seem to undermine their ability to hunt, perhaps explaining the absence of obligate scavengers in these groups \citep{ruxton2004obligate}.
+%\\There is a long running debate on the tendency of human ancestors to scavenge. 
+%Some recent studies have found "that passive scavenging from abandoned larger felid kills could have been a high-yield, though potentially dangerous, foraging strategy for early hominins even without considering within-bone nutrients" \citep{pobiner2015new}.
 
-Large body size confers many advantages to a scavenger. 
-It grants an animal competitive benefits against other species at a carcass, it allows for some resistance to starvation because of greater fat stores and it often means the sensory organs are elevated giving a higher likelihood of discovering food \citep{farlow1994speculations}.
-
-\subsection*{Cenozoic}
-Among terrestrial African carnivores, hyenas, jackals, lions and leopards all take sizable proportions of carrion in their diet.
-In the case of the spotted hyena (\textit{Crocuta crocuta}) it can be as high as 99\% \citep{benbow2015introduction}. 
-Yet, no contemporary terrestrial vertebrate exists as an obligate scavenger. 
-The selective pressures that push mammals and reptiles towards scavenging do not seem to undermine their ability to hunt, perhaps explaining the absence of obligate scavengers in these groups \citep{ruxton2004obligate}.
-\\There is a long running debate on the tendency of human ancestors to scavenge. 
-Some recent studies have found "that passive scavenging from abandoned larger felid kills could have been a high-yield, though potentially dangerous, foraging strategy for early hominins even without considering within-bone nutrients" \citep{pobiner2015new}.
-
-Osteophagy is known across a range of terrestrial carnivores.
-Some fat-rich mammalian bones have an energy density (6.7 kJ/g) comparable with that of muscle tissue, making skeletal remains an enticing resource \citep{brown1989study}. 
-Hyenas have a bite force capable of breaking open bones and early hominins had the ability to craft tools for this purpose \citep{hone2010feeding,ARCM:ARCM12084}. 
-In light of this, the skeletal remains of carrion may act as trove of food to carnivores who can access it.
-
-\cite{ruxton2004obligate} in a theoretical study suggested that "a 1 tonne mammal or reptile, in an ecosystem yielding carrion at densities similar to the current Serengeti, could have met its energy requirements if it could detect carrion over a distance of the order of 400–500 m."
+%Osteophagy is known across a range of terrestrial carnivores.
+%Some fat-rich mammalian bones have an energy density (6.7 kJ/g) comparable with that of muscle tissue, making skeletal remains an enticing resource \citep{brown1989study}. 
+% Hyenas have a bite force capable of breaking open bones and 
+% In light of this, the skeletal remains of carrion may act as trove of food to carnivores who can access it.
 
@@
 
-A review by \cite{devault2002scavenging} found occurrences of scavenging spread across five families of snakes and states that this behaviour is "far more common than currently acknowledged."
+% A review by \cite{devault2002scavenging} found occurrences of scavenging spread across five families of snakes and states that this behaviour is "far more common than currently acknowledged."
 
@@
 
-\subsection*{Mesozoic}
-In a recent publication a modelling approach found that theropods between 27 kg and 1044 kg would have gained a significant energetic advantage over individuals at both the small and large extremes of theropod body mass through their scavenging efficiency.
-This humped pattern is the result of the disparity between the scaling of energetic cost which scales according to an exponent of 0.91 and energy input that scales according to a cubic polynomial that initially scales according to an exponent of 1.07 but plateaus after 1000kg.
-The polynomial behaviour of energy input is itself the result of the limitations imposed by gut capacity and the overall availability of carcasses after competition.
+%\textit{Allosaurus} tooth marks on a hadrosaur in the Late Jurassic. 
+% Late Triassic scavenging on a prosauropod by basal carnivorous archosaurs \citep{hungerbuhler1998taphonomy}.
 
-As we discussed for the case of Cenozoic carnivores, osteophagy could be extremely beneficial to a scavenger. 
-In Mesozoic systems some extremely large theropod dinosaurs had a morphology which suggests an ability to process bone e.g. the robust skull and dentition of \textit{Tyrannosaurus rex}. 
-There is direct evidence that \textit{T.rex} did this in the form of distinctive wear marks on its tooth apices \citep{farlow1994wear,schubert2005wear} and the presence of bone fragments in its coprolites \citep{chin1998king}.
-The animal also had an enormous bite force, with one estimate putting it at 57000 Newtons \citep{bates2012estimating}. 
-This is noted as being powerful enough to break open skeletal material \citep{rayfield2001cranial}. 
-Osteophagy may have been even more viable during this era because the body mass distribution of herbviores tended to be skewed towards larger sizes \citep{10.1371/journal.pone.0051925}. 
-When we couple this with the fact that skeletal mass scales greater than linearly with body mass \citep{prange1979scaling} there would have been a lot of bones to consume in the environment provided an animal had the biology to process it.
-
-\textit{Allosaurus} tooth marks on a hadrosaur in the Late Jurassic \citep{chure1997one}. 
-Late Triassic scavenging on a prosauropod by basal carnivorous archosaurs \citep{hungerbuhler1998taphonomy}.
-
-\subsection*{Palaeozoic}
-Synapsids 
-Sprawling gait
-
-%\section{Aerial Scavengers}
-Species capable of flight have effectively added an extra spatial dimension, i.e. the vertical component, to their sensory environment.
-This allows them to look down on a landscape where they are unencumbered by obstacles that would obstruct the view of a terrestrial scavenger.
-Moreover, flight is a cheaper means of locomotion than running \citep{tucker1975energetic}. 
-Thus, it appear that would-be scavengers have a distinct advantage by taking flight.
+%\subsection*{Palaeozoic}
+% The absence of flying vertebrates in the Palaeozoic may have permitted terrestrial forms to take in a higher proportion of carrion in their diet.
 
 %\subsection*{Cenozoic}
-Birds are the dominant vertebrate fliers today and include the best known scavengers on Earth, the vultures. 
-These birds consist of two convergent groups, old world and new world vultures and represent the only example of obligate vertebrate scavengers. 
-They have a suite of adaptations that allow them to flourish in this niche.
-Vultures extend the energetic advantage of flight further by engaging in soaring instead of flapping flight, which is even cheaper \citep{hedenstrom1993migration}.
-Their efficiency is illustrated by cases of predators like bears and wolves benefiting by taking more carrion in their diet in areas bereft of vultures through competitive release \citep{devault2003scavenging}.
-In flight, birds possess a huge advantage over any terrestrial obligate scavenger. 
-Flight affords them the ability to range over a much larger area and detect carrion from an elevated vantage point.
-\cite{pennycuick1972soaring} conservatively estimated that a Gyps vulture could identify activity at a carcass 4 km away.
-
-Many other bird species take a significant amount of carrion in their diet notably the eagles, storks and corvids. 
-Although, none of them are obligate scavengers. 
-By contrast with terrestrial species, the traits that render a bird adept at scavenging do undermine its ability to function as an effective predator \citep{devault2003scavenging}.
-
-The ability to lower BMR at night is reported in some \textit{Gyps} vultures.
+%A likely instance of scavenging between a 4-million-year-old white shark (\textit{Carcharodon}) and mysticete whale from Peru \citep{ehret2009caught}.
+%Bite marks on early Holocene Tursiops truncatus fossils from the North Sea indicate scavenging by rays (Chondrichthyes, Rajidae) \citep{van2009bite}. 
+%Possible scavenging on a juvenile fur seal from the Late Neogene \citep{boessenecker2011mammalian}.
 
 %\subsection*{Mesozoic}
-In ancient ecosystems, the volant pterosaurs have also been postulated as existing in a vulture-like niche \citep{witton2008reappraisal}. 
-Certain clades of these animals could reach enormous sizes (e.g. Azhdarchids) and look to have engaged in soaring flight. 
-However, the inflexibility of their necks and straight, rather than hooked jaw morphology argues against their existing as obligate scavengers \citep{witton2008reappraisal}. 
-As yet, no one has ever attempted an energetics approach for this group which is likely due to the many uncertainties over their biology \citep{witton2010size}.
 
 %\subsection*{Palaeozoic}
-The absence of flying vertebrates in the Palaeozoic may have permitted terrestrial forms to take in a higher proportion of carrion in their diet.
+%Evolution of sharks, known scavengers.
 
-\section{Aquatic Scavengers}
-The existence of an obligate scavenger in a marine setting also remains hypothetical \citep{britton1994marine,smith2003ecology,ruxton2004energetic,ruxton2005searching}. 
-Carrion in this environment is produced by dead flesh and other marine organisms when their carcasses descend to the sea floor. 
-This low-light environment means animals detect resources through chemo- and mechanoreception \citep{ruxton2004energetic}. 
-Detection distances are far lower than they would be in the air (< 100 m) as a result. 
-However, water is a medium that is conducive to low-cost movement \citep{tucker1975energetic} and so may be able to support a small obligate scavenging fish \citep{ruxton2004energetic,ruxton2005searching}.
-Although, for the time being this remains conjectural.
+%\section*{Results}
 
-"although some benthic scavengers (e.g., hagfish: family Myxinidae) rely on necrophagy for a large portion of their diet and may indeed be obligate scavengers" \citep{benbow2015introduction}
-
-\subsection*{Cenozoic}
-A likely instance of scavenging between a 4-million-year-old white shark (\textit{Carcharodon}) and mysticete whale from Peru \citep{ehret2009caught}.
-Bite marks on early Holocene Tursiops truncatus fossils from the North Sea indicate scavenging by rays (Chondrichthyes, Rajidae) \citep{van2009bite}. 
-Possible scavenging on a juvenile fur seal from the Late Neogene \citep{boessenecker2011mammalian}.
-
-\subsection*{Mesozoic}
-Evidence of scavenging in a Cretaceous shark species \textit{Squalicorax} whereby remains of a mosasaur and a hadrosaur were discovered with embedded shark teeth \citep{schwimmer1997scavenging}. 
-\subsection*{Palaeozoic}
-Evolution of sharks, known scavengers. 
-Evidence of vertebrate scavenging dates back to the early Permian approximately 300 MYA \citep{reisz2006articulated}.
-
-\section*{Results}
-
-\section*{Discussion}
+%\section*{Discussion}
 

### 2016-03-15 natural history of scavenging.tex
Cleaned up sections

@@
 
-\author{Adam Kane, \and Kevin Healy, \and Thomas Guillerme \and Graeme Ruxton, \and \& Andrew Jackson.}
+\author{Adam Kane, Kevin Healy, Thomas Guillerme, Graeme Ruxton, \& Andrew Jackson.}
 
@@
 This may go some way to explaining the gap in our knowledge of the prevalence of this behaviour.
-Consider Professor Sanborn Tenney writing in 1877 for The American Naturalist who had this to say about one well known group, “Prominent among the mammalian scavengers are the hyenas, the ugliest in their general appearance of all the flesh eaters.” He contrasts these with “nobler kinds” of carnivores such as lions and tigers.
+Consider Professor Sanborn Tenney writing in 1877 for The American Naturalist who had this to say about one well known group, “Prominent among the mammalian scavengers are the hyenas, the ugliest in their general appearance of all the flesh eaters.” He contrasts these with “nobler kinds” of carnivores such as lions and tigers \citep{tenney1877few}.
 Even aside from our own subjective biases, scavenging is a difficult behaviour to detect after the fact.
@@
 Indeed, a single species of dinosaur notwithstanding, a synthesis describing the natural history of scavengers is absent from the literature.
-Fortunately, research in this area is on the rise.
-As a result we are now beginning to realise the extent of this behaviour such that, “in some ecosystems, vertebrates have been documented to assimilate as much as 90\% of the available carrion" \citep{benbow2015introduction}.
+Fortunately, research on scavenging is on the rise \citep{manga2006vulture}.
+As a result, we are now beginning to realise the extent of this behaviour such that, “in some ecosystems, vertebrates have been documented to assimilate as much as 90\% of the available carrion" \citep{benbow2015introduction}.
 Even Tenney’s noble big cats are now known to take in a significant portion of carrion in their diet where some lion populations get over 50\% of their meat from carcasses.
-A suite of methods have been used to discern the most suitable morphologies, physiologies and environments for a scavenging lifestyle to prosper.
-It is our aim in this review to employ these methods to gain an understanding of scavengers past and present.
-
+By recognising the difficulty in directly observing scavenging, a suite of methods have been used to discern the most suitable morphologies, physiologies and environments for a scavenging lifestyle to prosper.
+Here we chart the natural history of vertebrate scavenging on air, land and in water by focussing on the dominant vertebrate groups.
+%It is our aim in this review to employ these methods to gain an understanding of scavengers past and present.
+\subsection*{The Difficulty of Scavenging}
 The chief hurdle to scavenging is finding a sufficient quantity of food, the occurence of which is difficult to predict in space and time.
+Once found the scavenger has to overcome the agents of decay produced by the action of microorganisms on the carcass.
 The idea of scrounging from predator kills is undermined from studies showing that in the  majority of ecosystems more animals die from disease and starvation than predation \citep{benbow2015introduction}.
@@
 The habitat must also be productive enough to sustain an animal biomass that will eventually produce carcasses.
-It is recognised that scavengers keep energy flows at a higher trophic level in food webs than decomposers because they consume relatively more carrion \citep{devault2003scavenging}.
 
@@
 And eagles in particular are known to have highly developed visual abilities.
-It follows from this that the evolution of flight allowed aerial animals to detect far more carrion than their terrestrial counterparts.
+It follows from this that the evolution of flight allowed aerial animals to detect far more carrion than their terrestrial counterparts \citep{AR:AR22815}.
 
-Moreover, having a panoramic view means being able to gather a wealth of information from other foragers, be they conspecifics or other species.
-Again, returing to vultures, the genus \textit{Gyps} consists of highly social and colonially nesting species.
-These behaviours allow them forage far more efficiently because one bird can scrounge information on the location of food from another successful forager.
+Moreover, having a panoramic view means being able to gather a wealth of information from other foragers, be they conspecifics or other species \citep{jackson2008effect}.
+Again, returing to vultures, the genus \textit{Gyps} consists of highly social and colonially nesting species \citep{fernandez2015density}.
+These behaviours allow them forage far more efficiently because one bird can scrounge information on the location of food from another successful forager \citep{KaneVul}.
+
+Aside from sight, three species within the new world family Cathartidae, (genus \textit{Cathartes}), have well developed olfactory systems \citep{AR:AR22815}.
+Among them are the Turkey Vultures \textit{Cathartes aura} which were able to locate 90\% of baits set out in a tropical forest \citep{houston1986olfaction}.
+This would be impossible for the visually reliant old world species.
+This point illustrates how the environment can impact search efficiency depending on the sensory system that's used.
 
@@
 This advantage can be extended further in larger species by engaging in soaring instead of flapping flight, which is even cheaper still (approximately twice BMR) \citep{hedenstrom1993migration}.
-The advantages this confers are clear from the information we have on the enormous foraging ranges of some seabirds and accipiters.
+The advantages this confers are clear from the information we have on the enormous foraging ranges of some seabirds and accipiters \citep{bamford2007ranging}.
 Clearly, it would be pointless to have incredible detection abilites and not have a cost efficient movement to benefit from it.
+Although, as with detection ability, the environment has a role to play here.
+Vultures and eagles tend to soar using thermals and if these air pockets don't form, say on a cloudy day, the bird is grounded \citep{mundy1992vultures}.
+In many habitats (e.g.
+the arctic) it is simply not possible for sufficiently powerful thermals to form and as a consequence large-bodied vultures cannot exist.
+The upshot of this is that terrestrial carnivores like bears and wolves take more carrion \citep{devault2003scavenging}.
 
@@
 Certain clades of these animals could reach enormous sizes (e.g.
-Azhdarchids with wingspans of 11 metres) and look to have engaged in soaring flight.
+Azhdarchids with wingspans of 11 metres) and look to have engaged in soaring flight \citep{witton2010size}.
 Although \cite{witton2008reappraisal} argued that neck inflexibility and straight, rather than hooked jaw morphology points against their existing as \textit{obligate} scavengers, Azhdarchid terrestrial proficency indicates they would have been comfortable foraging on the ground.
@@
 
-Vertebrate scavengers in general are responsible for the dispersal of nutrients.
-Consider the diversity of animals that can end up feeding at the carcass of an elephant.
-Here we have an incredibly dense and nutrient rich patch that ends up being distributed widely.
-Thus, in an ecological context, the evolution of flight coupled with the ability to scavenge resulted in a world with a far more widely distributed nutrient landscape.
-In the absence of vertebrate scavengers, invertebrates and microorganisms would consume the carcass in-situ or at least distribute the constituent nutrients over a much shorter range.
-
 \section*{Terrestrial Scavengers}
-A simplificaiton of terrestrial scavengers is one of them existing in a two-dimensional plane while foraging for carrion directly.
+A simplificaiton of terrestrial, vertebrate scavengers is one of them existing in a two-dimensional plane while foraging for carrion directly.
 They can detect carcasses at a range that is defined by the radius of their sensory organs, usually the visual and olfactory senses.
@@
 \cite{ruxton2004obligate} offer a reason for this in that the traits that allow for vultures to exist as scavengers undermined their ability to hunt but that the same forces have not prevented mammals from doing so.
-The same authors in a theoretical study do concede that "a 1 tonne mammal or reptile, in an ecosystem yielding carrion at densities similar to the current Serengeti, could have met its energy requirements if it could detect carrion over a distance of the order of 400–500 m."\citep{ruxton2004obligate}.
+The same authors calculated that "a 1 tonne mammal or reptile, in an ecosystem yielding carrion at densities similar to the current Serengeti, could have met its energy requirements if it could detect carrion over a distance of the order of 400–500 m."\citep{ruxton2004obligate}.
 
@@
 This ability reached its zenith among hyenas with the evolution of the 110 kg \textit{Pachycrocuta brevirostris} during the Pliocene \citep{palmqvist2011giant}.
-The ability to process bone means a carcass fed on by hyenas will be reduced to nothing, whereas the skeleton will remain in carrion attended by species restriced to feeding on the flesh.
+Some work on extinct sabretooths suggests they may have left a large amount of food for would-be scavengers because of their unique skull morphology.
+As a result the decline of Machairodontinae sabretooths has been offered as an explanation for the extinction of \textit{P brevirostris} \citep{palmqvist2011giant}.
+A major difficulty for terrestrial scavengers is competition with vultures.
+Noctural behaviour in the Hyaenidae in general has been put forth as an adaptation to reduce competition with these exclusively diurnal birds \citep{gittleman2013carnivore}.
 
-Many of these adapations to scavenging are found in the other major extant terrestrial mammalian carnviores, the bears, dogs and cats to a greater or lesser extent.
+The order Carnivora sees its origins in the Eocene where it split into the Caniforma and Feliforma.
+And many of the aforesaid adapations for scavenging are found in these other major terrestrial mammalian carnviores.
 Though the specific mix of features realised in hyenas suggest this is the model organism for terrestrial scavenging among mammals in the past.
 Indeed, the bone-crushing dogs that evolved during the Oligocene (subfamily Borophaginae) have been compared to hyenas in terms of their feeding ecology \citep{van2003chapter,martin2016pursuit}.
+
 Interestingly such comparisons have given insight into the feeding ecology of early hominins who, for instance, had the ability to craft tools for breaking open bones \citep{hone2010feeding,ARCM:ARCM12084}.
+The question of where our ancestors placed on the hunter-scavenger axis during the Plio-Pleistocene has been a matter of debate for years.
+A recent study investigating potential scavenging opportunities for hominins in Kenya found that, even when discounting bone material, there is a substantial amount of scavengeable meat left on predated remains; sufficient to sustain the requirements of an adult male \textit{Homo erectus} \citep{pobiner2015new}.
+In some historical hominin-inhabited areas there were a greater number of felids than hyenids.
+This is significant because hyenas will leave far less flesh on a carcass than a felid such as a sabretooth enabling contemperaneous hominins to benefit \citep{pobiner2015new}.
+The intelligence, resultant tool-use and cooperative nature of hominins means they could likely adapt to take on more or less carrion depending on their environment.
 
-By contrast, a successful reptilian scavenger requires a far different set of adapations.
+By contrast to mammals, a successful reptilian scavenger requires a far different set of adapations.
 Modern forms are ectothermic, limiting their activity periods.
-This is exacerbated by the sprawling gait seen in lizards which results in Carrier's Constraint such that the animal can't move and breathe at the same time because the lateral movements impedes its lungs.
+This is exacerbated by the sprawling gait seen in lizards which results in Carrier's Constraint such that the animal can't move and breathe at the same time because the lateral movements impedes its lungs \citep{carrier1987evolution}.
 This manifests itself in aspects such as maximum sustainable speed where an equivalent mammal has a six to seven fold increase \citep{ruben1995evolution}.
@@
 
-Evidence of vertebrate scavenging dates back to the early Permian approximately 300 MYA \citep{reisz2006articulated}.
+Of course, tetrapod terrestrial dominance predates the evolution of the dinosaurs.
+It is in the Permian, almost 300 millions years ago, that we have the earliest evidence of vertebrate scavenging where a temnospondyl amphibian fed on the carcass of \textit{Varanops}, a predatory synapsid of the time \citep{reisz2006articulated}.
+The absence of flying vertebrates in the Palaeozoic may have permitted terrestrial forms to take in a higher proportion of carrion in their diet.
+As noted earlier, a system similar to the Serengeti could support a terrestrial scavenger \citep{ruxton2004obligate}
 
@@
 An aquatic environment presents challenges for direct observational studies and so, similar to the approaches involving extinct species, much work has approached the question of scavenging propensity from an energetics perspective.
+Although as with the aerial and terrestrial enviornments we have evidence of facultative scavenging among aquatic species.
+For example, the remains of a mosasaur and a terrestrial hadrosaur were discovered with embedded teeth from a Cretaceous shark \textit{Squalicorax} \citep{schwimmer1997scavenging}.
+As well as a likely instance of scavenging between a 4-million-year-old white shark (\textit{Carcharodon}) and mysticete whale from Peru \citep{ehret2009caught}.
 The existence of an obligate scavenger in a marine setting is uncertain \citep{britton1994marine,smith2003ecology,ruxton2004energetic,ruxton2005searching}.
-Carrion in this environment is produced by marine organisms when their carcasses descend to the sea floor.
-In this low-light environment detection distances are far lower (< 100 m) than they would be in the air.
+Depending on the species a carcass in this environment either floats or descends to the sea floor.
+In this latter low-light environment detection distances are far lower (< 100 m) than they would be in the air.
 As such, animals detect resources through chemo- and mechanoreception more so than through vision \citep{ruxton2004energetic}.
@@
 Extant aquatic snakes are deemed as having the most suitable physiology and environment for scavenging.
-A hypothesis put forth by \citep{sazima1990necrofagia} argued that chemical gradients in water would allow for a relatively easier detection of carrion.
+A hypothesis put forth by \cite{sazima1990necrofagia} argued that chemical gradients in water would allow for a relatively easier detection of carrion.
 This gained some support from \cite{devault2002scavenging}, who found a preponderence of aquatic snake species in their review of this behaviour.
@@
 Thus, the energetic feasiblity of a marine scavenger has a long history.
+One point of interest is that of the whaling industry, which provided a bonanza of floating carcasses especially during the 20th century \citep{Whitehead415}.
+This meant Killer Whales \textit{Orcinus orca} could switch from hunting to scavenging, a switch made that much easier by the noise of the whaling vessels that would effectively ring the "dinner-bells" \citep{Whitehead415}.
+Early whales such as \textit{Basilosaurus} seem to fit into the same niche as Killer Whales and we have some evidence for scavenging in this group as well \citep{fahlke2012bite}.
 
-As with the aerial and terrestrial enviornments we have evidence of facultative scavenging among extinct, aquatic species.
-For example, the remains of a mosasaur and a terrestrial hadrosaur were discovered with embedded teeth from a Cretaceous shark \textit{Squalicorax} \citep{schwimmer1997scavenging}.
-As well as a likely instance of scavenging between a 4-million-year-old white shark (\textit{Carcharodon}) and mysticete whale from Peru \citep{ehret2009caught}.
+\section*{Ecological Role}
+It is recognised that scavengers keep energy flows at a higher trophic level in food webs than decomposers because they consume relatively more carrion \citep{devault2003scavenging}.
+They are also hugely important for the dispersal of nutrients \citep{benbow2015introduction}.
+Consider the diversity of animals that can end up feeding at the carcass of an elephant.
+Here we have an incredibly dense and nutrient rich patch that ends up being distributed widely.
+In the absence of vertebrate scavengers, invertebrates and microorganisms would consume the carcass in-situ or at least distribute the constituent nutrients over a much shorter range.
+This effect has been magnified as vertebrates evolved certain key traits that allowed them to range farther, namely an upright gait, an endothermic metabolism and of course, flight.
+To quantify this effect with a simple example we can turn  to some allometric relationships relating sustainable travelling speed to body mass.
+In the case of mammals and reptiles these are 1.15 * body mass (kg) \textsuperscript{0.12} and 0.23 * body mass (kg) \textsuperscript{0.12}
+respectively \citep{ruxton2004obligate}.
+We can insert these into a foraging radius model ((duration * speed)/2)/1000 for a 12 hour foraging day which shows that while a 10 kg reptile can range 6.5 km an equally sized mammal can range nearly 33 km \citep{Enstipp2006Energetics}.
+Thus, in an ecological context, the evolution of these steps coupled with the ability to scavenge resulted in a world with a far more widely distributed nutrient landscape.
 

### 2016-03-15 foraging range model.R
foraging range code

@@ -1,0 +1,8 @@
+## Sustainable travelling speed (v, m s-1) as a function of mass (M, kg)
+M<-10
+mSpeed<-1.15*M^0.12 ## mammals
+rSpeed<-0.23*M^0.12 ## reptiles
+
+duration <- 43200 # 12 hours 
+speed = c(mSpeed, rSpeed)
+((duration * speed)/2)/1000

### 2016-03-16 natural history of scavenging.tex
Added equations templates
Minor updates (mainly dates and a couple of comments)

@@
 \usepackage{lmodern}
+\usepackage{amsmath}
 
@@
 These birds consist of two convergent groups, from the old and the new world and represent the only example of obligate vertebrate scavengers today.
-Given their unique position, they have been extensively studied to determine what adaptations they possess that allows them to so flourish in this niche.
-As such, we can begin by exploring the adapdations and the environments of vultures to draw comparisons with other scavenging species and \textit{their} environments.
+Given their unique position,%TG: what do you mean by their position?
+Phylogenetic?
+Position on the "scavenging scale"?
+ they have been extensively studied to determine what adaptations they possess that allows them to so flourish in this niche.
+As such, we can begin by exploring the adaptations and the environments of vultures to draw comparisons with other scavenging species and \textit{their} environments.
 
@@
 Moreover, having a panoramic view means being able to gather a wealth of information from other foragers, be they conspecifics or other species \citep{jackson2008effect}.
-Again, returing to vultures, the genus \textit{Gyps} consists of highly social and colonially nesting species \citep{fernandez2015density}.
+Again, returning to vultures, the genus \textit{Gyps} consists of highly social and colonially nesting species \citep{fernandez2015density}.
 These behaviours allow them forage far more efficiently because one bird can scrounge information on the location of food from another successful forager \citep{KaneVul}.
@@
 
-Avian flight originates in the Jurassic Period, conincident with the fossils of \textit{Archaeopteryx lithographica} so many of these benefits would have been realised from that point on for carnivorous birds.
-However, vertebrate flight is much older than this where pterosaurs predate bird origins by a considerable margin in the Triassic Period.
+Avian flight originates in the Late Jurassic (163.5-145 Million years ago - Mya), conincident with the fossils of \textit{Archaeopteryx lithographica} so many of these benefits would have been realised from that point on for carnivorous birds.
+However, vertebrate flight is much older than this where pterosaurs predate bird origins by a considerable margin in the Late Triassic (235-201.3 Mya).
 Scavenging in this diverse group has been hypothesied many times before.
 Certain clades of these animals could reach enormous sizes (e.g.
-Azhdarchids with wingspans of 11 metres) and look to have engaged in soaring flight \citep{witton2010size}.
+Azhdarchids with wingspans of 11 metres)%CITE
+ and look to have engaged in soaring flight \citep{witton2010size}.
 Although \cite{witton2008reappraisal} argued that neck inflexibility and straight, rather than hooked jaw morphology points against their existing as \textit{obligate} scavengers, Azhdarchid terrestrial proficency indicates they would have been comfortable foraging on the ground.
-Indeed, extant Marabou Storks have a comparable morphology and are noted facultative scavengers so it is reasonable to believe that certain pterosaurs behaved similarly.
+Indeed, extant Marabou Storks (\textit{Leptoptilos crumenifer}) %TG: Check if that's the one
+have a comparable morphology and are noted facultative scavengers so it is reasonable to believe that certain pterosaurs behaved similarly.
 
@@
 And as we have noted the Azhdarchid pterosaurs were far bigger again, with estimated body masses of around 80 kg.
+%CITE
 
@@
 The bat fossil record is notoriously poor owing to their fragile skeletons so we are unable to determine if extinct species were more suited to this lifestyle.
+%TG: Two points here: one speculation for the non-scavenging behaviour could be because of diet.
+Big bats (which are better suited for scavenging, following previous paragraph) are essentially frugivores and therefore probably lack of adaptations for digesting the meat.
+Also carnivores bats are mainly the microbats that are insectivores.
+Maybe they scavenge on dead insects but here we're talking about scavenging on real meat right?
+%TG: Second point: bats fossil record is of course not so good but it's actually way better than birds!
+Most of the data we have is on teeth of course but it's probably enough for having a rough body mass and diet estimate!
+And that's much better than birds where you're stuck with just some fragments of long bones.
 
 \section*{Terrestrial Scavengers}
-A simplificaiton of terrestrial, vertebrate scavengers is one of them existing in a two-dimensional plane while foraging for carrion directly.
+A simplification of terrestrial, vertebrate scavengers is one of them existing in a two-dimensional plane while foraging for carrion directly.
 They can detect carcasses at a range that is defined by the radius of their sensory organs, usually the visual and olfactory senses.
@@
 Some work on extinct sabretooths suggests they may have left a large amount of food for would-be scavengers because of their unique skull morphology.
-As a result the decline of Machairodontinae sabretooths has been offered as an explanation for the extinction of \textit{P brevirostris} \citep{palmqvist2011giant}.
+As a result the decline of Machairodontinae sabretooths has been offered as an explanation for the extinction of \textit{P.
+brevirostris} \citep{palmqvist2011giant}.
 A major difficulty for terrestrial scavengers is competition with vultures.
@@
 
-The order Carnivora sees its origins in the Eocene where it split into the Caniforma and Feliforma.
+The order Carnivora sees its origins in the Middle Eocene (56-33.9 Mya) %TG: the middle Eocene date is actually for stem carnivora.
+It's rather important for paleao people but I think it will sound pedantic in this review
+where it split into the Caniforma and Feliforma.
 And many of the aforesaid adapations for scavenging are found in these other major terrestrial mammalian carnviores.
 Though the specific mix of features realised in hyenas suggest this is the model organism for terrestrial scavenging among mammals in the past.
-Indeed, the bone-crushing dogs that evolved during the Oligocene (subfamily Borophaginae) have been compared to hyenas in terms of their feeding ecology \citep{van2003chapter,martin2016pursuit}.
+Indeed, the bone-crushing dogs that evolved during the Oligocene (subfamily Borophaginae; 33.9-23.03 Mya) have been compared to hyenas in terms of their feeding ecology \citep{van2003chapter,martin2016pursuit}.
 
 Interestingly such comparisons have given insight into the feeding ecology of early hominins who, for instance, had the ability to craft tools for breaking open bones \citep{hone2010feeding,ARCM:ARCM12084}.
-The question of where our ancestors placed on the hunter-scavenger axis during the Plio-Pleistocene has been a matter of debate for years.
+The question of where our ancestors placed on the hunter-scavenger axis during the Plio-Pleistocene (3.6-0.126 Mya) has been a matter of debate for years.
 A recent study investigating potential scavenging opportunities for hominins in Kenya found that, even when discounting bone material, there is a substantial amount of scavengeable meat left on predated remains; sufficient to sustain the requirements of an adult male \textit{Homo erectus} \citep{pobiner2015new}.
 In some historical hominin-inhabited areas there were a greater number of felids than hyenids.
-This is significant because hyenas will leave far less flesh on a carcass than a felid such as a sabretooth enabling contemperaneous hominins to benefit \citep{pobiner2015new}.
-The intelligence, resultant tool-use and cooperative nature of hominins means they could likely adapt to take on more or less carrion depending on their environment.
+This is significant because hyenas will %TG: or "could"?
+This is hypothetical behaviour stuff!
+Maybe past hyenas where exclusive sushi eaters!
+ leave far less flesh on a carcass than a felid such as a sabretooth enabling contemperaneous hominins to benefit \citep{pobiner2015new}.
+The intelligence,%TG: Not sure about the intelligence thing: the sentence as the same meaning without and will be less anthropocentric sounding.
+ resultant tool-use and cooperative nature of hominins means they could likely adapt to take on more or less carrion depending on their environment.
 
@@
 We do know that they walked with the erect gait of mammals or birds rather than the sprawling gait of lizards and that they were most likely facultative scavengers \citep{depalma2013physical}.
-Much work has focused on the existence of the behaviour in \textit{Tyrannosaurus rex} \citep{ruxton2003could,carbone2011intra} but a recent energetics study investigated the likely prevelance of scavenging across a range of body sizes.
+Much work has focused on the existence of the behaviour in \textit{Tyrannosaurus rex} \citep{ruxton2003could,carbone2011intra} but a recent energetics study investigated the likely prevalence of scavenging across a range of body sizes.
 In it the authors demonstrated that species of intermediate body masses (approx.
@@
 In Mesozoic systems some extremely large theropod dinosaurs had a morphology which suggests an ability to process bone e.g.
-the robust skull and dentition of \textit{T rex}.
-There is direct evidence that \textit{T.rex} did this in the form of distinctive wear marks on its tooth apices \citep{farlow1994wear,schubert2005wear} and the presence of bone fragments in its coprolites \citep{chin1998king}.
+the robust skull and dentition of \textit{T.
+rex}.
+There is direct evidence that \textit{T.
+rex} did this in the form of distinctive wear marks on its tooth apices \citep{farlow1994wear,schubert2005wear} and the presence of bone fragments in its coprolites \citep{chin1998king}.
 The animal also had an enormous bite force, with one estimate putting it at 57000 Newtons \citep{bates2012estimating}.
@@
 An aquatic environment presents challenges for direct observational studies and so, similar to the approaches involving extinct species, much work has approached the question of scavenging propensity from an energetics perspective.
-Although as with the aerial and terrestrial enviornments we have evidence of facultative scavenging among aquatic species.
+Although as with the aerial and terrestrial environments we have evidence of facultative scavenging among aquatic species.
 For example, the remains of a mosasaur and a terrestrial hadrosaur were discovered with embedded teeth from a Cretaceous shark \textit{Squalicorax} \citep{schwimmer1997scavenging}.
-As well as a likely instance of scavenging between a 4-million-year-old white shark (\textit{Carcharodon}) and mysticete whale from Peru \citep{ehret2009caught}.
+As well as a likely instance of scavenging between a 4 million years old white shark (\textit{Carcharodon}) and mysticete whale from Peru \citep{ehret2009caught}.
 The existence of an obligate scavenger in a marine setting is uncertain \citep{britton1994marine,smith2003ecology,ruxton2004energetic,ruxton2005searching}.
@@
 Any animal that could seek out such whale carcasses is unlikely to have ignored other types of carrion.
-Although no aquatic species have ever exceeded the size of whales, some enormous animals have evolved in this environment before the evolution of whales, including \textit{Leedsichthys}, a bony fish from the Jurassic Period, that weighed in excess of 20 tonnes.
+Although no aquatic species have ever exceeded the size of whales, some enormous animals have evolved in this environment before the evolution of whales, including \textit{Leedsichthys}, a bony fish from the Middle Jurassic (174.1-163.5 Mya), that weighed in excess of 20 tonnes.
 Thus, the energetic feasiblity of a marine scavenger has a long history.
@@
 Thus, in an ecological context, the evolution of these steps coupled with the ability to scavenge resulted in a world with a far more widely distributed nutrient landscape.
+
+% %TG: LaTeX format template for the equations in the R file
+% \begin{equation}
+%   \text{Travelling speed} = \frac{duration \times speed}{2} \times 10^3
+% \end{equation}
+% where \textit{C} is a constant of $1.15$ for mammals and $0.23$ for squamates, %TG: if you meant squamates from reptiles
+% and speed being:
+% \begin{equation}
+%   speed = C \times M^{0.12}
+% \end{equation}
 

### 2016-03-16 natural history of scavenging.tex
section on temp

--- 2016-03-15 natural history of scavenging.tex
@@
 \begin{document}
+
+% %TG: LaTeX format template for the equations in the R file
+% \begin{equation}
+%     \text{Travelling speed} = \frac{duration \times speed}{2} \times 10^3
+% \end{equation}
+% where \textit{C} is a constant of $1.15$ for mammals and $0.23$ for squamates, %TG: if you meant squamates from reptiles
+% and speed being:
+% \begin{equation}
+%     speed = C \times M^{0.12}
+% \end{equation}
 
@@
 The chief hurdle to scavenging is finding a sufficient quantity of food, the occurence of which is difficult to predict in space and time.
-Once found the scavenger has to overcome the agents of decay produced by the action of microorganisms on the carcass.
-The idea of scrounging from predator kills is undermined from studies showing that in the  majority of ecosystems more animals die from disease and starvation than predation \citep{benbow2015introduction}.
+Once found the scavenger has to then overcome the agents of decay produced by the action of microorganisms on the carcass \citep{ruxton2014fruit}.
+The idea of scrounging from predator kills is undermined by studies showing that in the  majority of ecosystems more animals die from disease and starvation than predation \citep{benbow2015introduction}.
 Thus, any animal existing as a scavenger must maximise its detection capabilities and minimise its locomotory costs \citep{ruxton2004obligate}.
@@
 Given their unique position, they have been extensively studied to determine what adaptations they possess that allows them to so flourish in this niche.
+Underscoring their prowess are studies such as \cite{ogada2012effects} who note a three fold increase in carcass persistence times in the absence of vultures.
 As such, we can begin by exploring the adapdations and the environments of vultures to draw comparisons with other scavenging species and \textit{their} environments.
@@
 Aside from sight, three species within the new world family Cathartidae, (genus \textit{Cathartes}), have well developed olfactory systems \citep{AR:AR22815}.
-Among them are the Turkey Vultures \textit{Cathartes aura} which were able to locate 90\% of baits set out in a tropical forest \citep{houston1986olfaction}.
+Among them are the Turkey Vultures (\textit{Cathartes aura}) which were able to locate 90\% of baits set out in a tropical forest \citep{houston1986olfaction}.
 This would be impossible for the visually reliant old world species.
@@
 
-Flight is also cheaper means of locomotion than running \citep{tucker1975energetic}.
+Flight is also a cheaper means of locomotion than running \citep{tucker1975energetic}.
 This advantage can be extended further in larger species by engaging in soaring instead of flapping flight, which is even cheaper still (approximately twice BMR) \citep{hedenstrom1993migration}.
-The advantages this confers are clear from the information we have on the enormous foraging ranges of some seabirds and accipiters \citep{bamford2007ranging}.
+The benefits this confers are clear from the information we have on the enormous foraging ranges of some seabirds and accipiters \citep{bamford2007ranging}.
 Clearly, it would be pointless to have incredible detection abilites and not have a cost efficient movement to benefit from it.
@@
 
-Large body size confers substantial dominance benefits \citep{ruxton2004obligate}.
+Large body size confers substantial dominance and starvation resistance benefits \citep{ruxton2004obligate}.
 Thus, we would expect scavengers to have this trait selected for even in the case of weight-constrained fliers.
@@
 They have a characteristic "rocking horse gait"  which allows them to cover great distances efficiently.
+A major difficulty for terrestrial scavengers is competition with vultures.
+Noctural behaviour in the Hyaenidae in general has been put forth as an adaptation to reduce competition with these exclusively diurnal birds \citep{gittleman2013carnivore}.
 The bone crushing ability of hyenas reveals another useful scavenger trait.
@@
 As a result the decline of Machairodontinae sabretooths has been offered as an explanation for the extinction of \textit{P brevirostris} \citep{palmqvist2011giant}.
-A major difficulty for terrestrial scavengers is competition with vultures.
-Noctural behaviour in the Hyaenidae in general has been put forth as an adaptation to reduce competition with these exclusively diurnal birds \citep{gittleman2013carnivore}.
 
@@
 Though the specific mix of features realised in hyenas suggest this is the model organism for terrestrial scavenging among mammals in the past.
-Indeed, the bone-crushing dogs that evolved during the Oligocene (subfamily Borophaginae) have been compared to hyenas in terms of their feeding ecology \citep{van2003chapter,martin2016pursuit}.
+For example, the bone-crushing dogs that evolved during the Oligocene (subfamily Borophaginae) have been compared to hyenas in terms of their feeding ecology \citep{van2003chapter,martin2016pursuit}.
+Certainly, it is profitable to compare extant terrestrial species to their prehistoric forebears given the dominance of mammalian carnviores since the Eocene.
 
@@
 This is significant because hyenas will leave far less flesh on a carcass than a felid such as a sabretooth enabling contemperaneous hominins to benefit \citep{pobiner2015new}.
-The intelligence, resultant tool-use and cooperative nature of hominins means they could likely adapt to take on more or less carrion depending on their environment.
+The intelligence, resultant tool-use and cooperative nature of hominins meant they could likely adapt to take on more or less carrion depending on their environment.
 
@@
 Modern forms are ectothermic, limiting their activity periods.
-This is exacerbated by the sprawling gait seen in lizards which results in Carrier's Constraint such that the animal can't move and breathe at the same time because the lateral movements impedes its lungs \citep{carrier1987evolution}.
+This is exacerbated by the sprawling gait seen in many lizards which results in Carrier's Constraint such that the animal can't move and breathe at the same time because the lateral movements impedes its lungs \citep{carrier1987evolution}.
 This manifests itself in aspects such as maximum sustainable speed where an equivalent mammal has a six to seven fold increase \citep{ruben1995evolution}.
@@
 This was the result of gut capacity limitations and the effects of competition at the carcass.
-At the larger extreme this owes to the fact that gut capacity doesn't scale isometrically with body mass so the benefits of greater mass level off; there's only so much food an individual can consume at a single sitting.
+At the larger extreme this owes to the fact that gut capacity doesn't scale isometrically with body mass so the benefits of greater mass level off; there's only so much food an individual can consume at a single sitting \citep{calder1996size}.
 For the smaller species, larger competitors would have prevented their access to carrion.
@@
 The absence of flying vertebrates in the Palaeozoic may have permitted terrestrial forms to take in a higher proportion of carrion in their diet.
-As noted earlier, a system similar to the Serengeti could support a terrestrial scavenger \citep{ruxton2004obligate}
+As noted earlier, a system similar to the Serengeti in productivity could have supported a terrestrial scavenger \citep{ruxton2004obligate}.
 
@@
 Depending on the species a carcass in this environment either floats or descends to the sea floor.
-In this latter low-light environment detection distances are far lower (< 100 m) than they would be in the air.
+In the latter low-light environment, detection distances are far lower (< 100 m) than they would be in the air.
 As such, animals detect resources through chemo- and mechanoreception more so than through vision \citep{ruxton2004energetic}.
@@
 
- %\citep{sazima1990necrofagia} predicted that aquatic or semi-aquatic piscivorous snakes  would  scavenge  more  frequently than other species.
-Moreover, \cite{sazima1990necrofagia}  suggested that water currents that induce  aggregations of carrion heighten the probability of carrion detection by these snakes (see also \cite{savitzky1992laboratory}).
-Also,  chemical  gradients may  be  more  uniform  and  give  more  de-pendable  directional information in water \citep{sazima1990necrofagia}." \citep{devault2002scavenging}
+\section*{The Role of Climate}
+The geological record shows the Earth has undergone radical fluctuations in temperature.
+This will have had a significant bearing on the availability of carrion.
+A 10$^{\circ}$C increase in ambient temperature can double carcass decomposition rates \citep{benbow2015introduction}.
+Oceanic productivity, decomposition rates and habitat structure are all impacted by climactic conditions.
+\cite{benbow2015introduction} points to a series of studies showing how microbes and invertebrates benefit at higher temperatures to the detriment of vertebrate scavengers "above 20$^{\circ}$C vertebrates were able to detect and consume only 19\% of small-mammal carcasses, whereas at temperatures below 18$^{\circ}$C, vertebrates consumed 49\% of carcasses".
+Furthermore, "carcass decomposition occurs more rapidly in warm, moist habitats when compared with more xeric habitats."
+
+% Slight changes in temperature could have profound impacts on energy flow within ecosystems as carrion decomposition
+% rates roughly double for every 10°C increase in ambient temperature (Vass et al. 1992; Parmenter and
+% MacMahon 2009). Consequently, the availability of carrion to vertebrate scavengers could be reduced by
+%as much as 20%–40% over the next century based on projections from current climate change models
+% (Beasley et al. 2012b).
+
+%"Changes in global climate are also expected to substantially alter the availability of carrion resources in marine ecosystems as surface production of organic material could decline by 50\% or more (Smith et al.
+2008).
+Such a drastic reduction in productivity undoubtedly will impact a multitude of ecosystem processes, including the availability and distribution of carrion within marine ecosystems." \citep{benbow2015introduction}
+
+%"For example, vertebrate scavengers appear to be disadvantaged when humidity and temperature favor microbial and invertebrate reproduction (between 21C and 38C)" \citep{benbow2015introduction}
+
+%"In a study by DeVault et al.
+(2004b), it was found that at temperatures above 20C vertebrates were able to detect and consume only 19\% of small-mammal carcasses, whereas at temperatures below 18C, vertebrates consumed 49\% of carcasses" \citep{benbow2015introduction}
+
+%"For example, the available literature suggests that carcass decomposition occurs more rapidly in warm, moist habitats when compared with more xeric habitats" \citep{benbow2015introduction}
+
+%Of particular note is the increase in decomposition rate of carrion with temperature which could result in a seasonal reduction of carrion availability given predicted warming trends \citep{moleon2014inter}.
 

### 2016-03-18 natural history of scavenging.tex

@@
 \begin{document}
-
-% %TG: LaTeX format template for the equations in the R file
-% \begin{equation}
-%     \text{Travelling speed} = \frac{duration \times speed}{2} \times 10^3
-% \end{equation}
-% where \textit{C} is a constant of $1.15$ for mammals and $0.23$ for squamates, %TG: if you meant squamates from reptiles
-% and speed being:
-% \begin{equation}
-%     speed = C \times M^{0.12}
-% \end{equation}
 
@@
 As such, we can begin by exploring the adapdations and the environments of vultures to draw comparisons with other scavenging species and \textit{their} environments.
+%The chief hurdle to scavenging is finding a sufficient quantity of food, the occurence of which is difficult to predict in space and time.
+Once found the scavenger has to overcome the agents of decay produced by the action of microorganisms on the carcass.
+The idea of scrounging from predator kills is undermined from studies showing that in the majority of ecosystems more animals die from disease and starvation than predation \citep{benbow2015introduction}.
+Thus, any animal existing as a scavenger must maximise its detection capabilities and minimise its locomotory costs \citep{ruxton2004obligate}.
+The habitat must also be productive enough to sustain an animal biomass that will eventually produce carcasses.
+
+%\section*{Aerial Scavengers}
+
+%Vultures represent the best known scavengers on Earth.
+These birds consist of two convergent groups, from the old and the new world and represent the only example of obligate vertebrate scavengers today.
+Given their unique position,%TG: what do you mean by their position?
+Phylogenetic?
+Position on the "scavenging scale"?
+% they have been extensively studied to determine what adaptations they possess that allows them to so flourish in this niche.
+As such, we can begin by exploring the adaptations and the environments of vultures to draw comparisons with other scavenging species and \textit{their} environments.
 
@@
 And as we have noted the Azhdarchid pterosaurs were far bigger again, with estimated body masses of around 80 kg.
+%Large body size confers substantial dominance benefits \citep{ruxton2004obligate}.
+Thus, we would expect scavengers to have this trait selected for even in the case of weight-constrained fliers.
+Cinereous Vultures (\textit{Aegypius monachus}) and condors (\textit{Vultur gryphus}, \textit{Gymnogyps californianus}) all have body masses that can exceed 10 kg and represent some of the heaviest bird species capable of flight \citep{ferguson2001raptors,donazar2002effects}.
+And as we have noted the Azhdarchid pterosaurs were far bigger again, with estimated body masses of around 80 kg.
+%CITE
 
@@
 For the smaller species, larger competitors would have prevented their access to carrion.
+%Similar to vultures they have well developed sensory organs, particularly in olfaction whereby they can detect a rotting carcass 2 km downwind.
+They have a characteristic "rocking horse gait" which allows them to cover great distances efficiently.
+The bone crushing ability of hyenas reveals another useful scavenger trait.
+Since carrion is not dispatched directly, often the most easily accessible and choicest components of the carcass will be missing or, if present, will be fought over.
+Being able to extract nutrients from remnants gives the scavenger a great advantage.
+Osteophagy is known across a range of terrestrial carnivores.
+Some fat-rich mammalian bones have an energy density (6.7 kJ/g) comparable with that of muscle tissue, making skeletal remains an enticing resource \citep{brown1989study}.
+This ability reached its zenith among hyenas with the evolution of the 110 kg \textit{Pachycrocuta brevirostris} during the Pliocene \citep{palmqvist2011giant}.
+Some work on extinct sabretooths suggests they may have left a large amount of food for would-be scavengers because of their unique skull morphology.
+As a result the decline of Machairodontinae sabretooths has been offered as an explanation for the extinction of \textit{P.
+brevirostris} \citep{palmqvist2011giant}.
+A major difficulty for terrestrial scavengers is competition with vultures.
+Noctural behaviour in the Hyaenidae in general has been put forth as an adaptation to reduce competition with these exclusively diurnal birds \citep{gittleman2013carnivore}.
+
+%The order Carnivora sees its origins in the Middle Eocene (56-33.9 Mya) %TG: the middle Eocene date is actually for stem carnivora.
+It's rather important for paleao people but I think it will sound pedantic in this review
+%where it split into the Caniforma and Feliforma.
+And many of the aforesaid adapations for scavenging are found in these other major terrestrial mammalian carnviores.
+Though the specific mix of features realised in hyenas suggest this is the model organism for terrestrial scavenging among mammals in the past.
+Indeed, the bone-crushing dogs that evolved during the Oligocene (subfamily Borophaginae; 33.9-23.03 Mya) have been compared to hyenas in terms of their feeding ecology \citep{van2003chapter,martin2016pursuit}.
+
+%Interestingly such comparisons have given insight into the feeding ecology of early hominins who, for instance, had the ability to craft tools for breaking open bones \citep{hone2010feeding,ARCM:ARCM12084}.
+The question of where our ancestors placed on the hunter-scavenger axis during the Plio-Pleistocene (3.6-0.126 Mya) has been a matter of debate for years.
+A recent study investigating potential scavenging opportunities for hominins in Kenya found that, even when discounting bone material, there is a substantial amount of scavengeable meat left on predated remains; sufficient to sustain the requirements of an adult male \textit{Homo erectus} \citep{pobiner2015new}.
+In some historical hominin-inhabited areas there were a greater number of felids than hyenids.
+This is significant because hyenas will %TG: or "could"?
+This is hypothetical behaviour stuff!
+Maybe past hyenas where exclusive sushi eaters!
+ %leave far less flesh on a carcass than a felid such as a sabretooth enabling contemperaneous hominins to benefit \citep{pobiner2015new}.
+The intelligence,%TG: Not sure about the intelligence thing: the sentence as the same meaning without and will be less anthropocentric sounding.
+% resultant tool-use and cooperative nature of hominins means they could likely adapt to take on more or less carrion depending on their environment.
+
+%By contrast to mammals, a successful reptilian scavenger requires a far different set of adapations.
+Modern forms are ectothermic, limiting their activity periods.
+This is exacerbated by the sprawling gait seen in lizards which results in Carrier's Constraint such that the animal can't move and breathe at the same time because the lateral movements impedes its lungs \citep{carrier1987evolution}.
+This manifests itself in aspects such as maximum sustainable speed where an equivalent mammal has a six to seven fold increase \citep{ruben1995evolution}.
+A lower metabolism does give reptiles an advantage however, in that over the course of a year their food requirements can be 30 times smaller than an endotherm of equal size \citep{Nagy1621}.
+Any adaptations that reduce energetic costs are likely to be selected in scavengers.
+\cite{devault2002scavenging} suggest this is an avenue for scavenging in snakes because they "exhibit exceedingly low maintenance metabolisms, and most can survive on a few scant feedings per year.
+It is, therefore, possible for snakes to rely largely on infrequent, less energy-rich meals." In the same review the authors found occurrences of scavenging spread across five families of snakes and stated that this behaviour is "far more common than currently acknowledged."\citep{devault2002scavenging}.
+
+%Unsurprisingly, given their enduring appeal, the prevalence of scavenging has been explored in the carnivorous, theropod dinosaurs.
+These animals ranged from the chicken-sized to the whale-sized all of which were bipedal.
+They are quite alien to anything we know today which restricts our ability to understand their ecology far more so than extinct mammals \citep{weishampel2004dinosauria}.
+Of relevance, are the questions that still persist about their metabolism \citep{grady2014evidence} and sensory perception.
+We do know that they walked with the erect gait of mammals or birds rather than the sprawling gait of lizards and that they were most likely facultative scavengers \citep{depalma2013physical}.
+Much work has focused on the existence of the behaviour in \textit{Tyrannosaurus rex} \citep{ruxton2003could,carbone2011intra} but a recent energetics study investigated the likely prevalence of scavenging across a range of body sizes.
+In it the authors demonstrated that species of intermediate body masses (approx.
+500 kg) would have gained the most benefit from scavenging.
+This was the result of gut capacity limitations and the effects of competition at the carcass.
+At the larger extreme this owes to the fact that gut capacity doesn't scale isometrically with body mass so the benefits of greater mass level off; there's only so much food an individual can consume at a single sitting.
+For the smaller species, larger competitors would have prevented their access to carrion.
 
@@
 \section*{Aquatic Scavengers}
+
 An aquatic environment presents challenges for direct observational studies and so, similar to the approaches involving extinct species, much work has approached the question of scavenging propensity from an energetics perspective.
@@
 
+%An aquatic environment presents challenges for direct observational studies and so, similar to the approaches involving extinct species, much work has approached the question of scavenging propensity from an energetics perspective.
+Although as with the aerial and terrestrial environments we have evidence of facultative scavenging among aquatic species.
+For example, the remains of a mosasaur and a terrestrial hadrosaur were discovered with embedded teeth from a Cretaceous shark \textit{Squalicorax} \citep{schwimmer1997scavenging}.
+As well as a likely instance of scavenging between a 4 million years old white shark (\textit{Carcharodon}) and mysticete whale from Peru \citep{ehret2009caught}.
+%The existence of an obligate scavenger in a marine setting is uncertain \citep{britton1994marine,smith2003ecology,ruxton2004energetic,ruxton2005searching}.
+Depending on the species a carcass in this environment either floats or descends to the sea floor.
+In this latter low-light environment detection distances are far lower (< 100 m) than they would be in the air.
+As such, animals detect resources through chemo- and mechanoreception more so than through vision \citep{ruxton2004energetic}.
+However, water is a medium that is conducive to low-cost movement \citep{tucker1975energetic} and so may be able to support an obligate scavenging fish \citep{ruxton2004energetic,ruxton2005searching}.
+\cite{benbow2015introduction} do note that "some benthic scavengers (e.g., hagfish: family Myxinidae) rely on necrophagy for a large portion of their diet and may indeed be obligate scavengers".
+
 Extant aquatic snakes are deemed as having the most suitable physiology and environment for scavenging.
@@
 %Of particular note is the increase in decomposition rate of carrion with temperature which could result in a seasonal reduction of carrion availability given predicted warming trends \citep{moleon2014inter}.
+
+% %TG: LaTeX format template for the equations in the R file
+% \begin{equation}
+%   \text{Travelling speed} = \frac{duration \times speed}{2} \times 10^3
+% \end{equation}
+% where \textit{C} is a constant of $1.15$ for mammals and $0.23$ for squamates, %TG: if you meant squamates from reptiles
+% and speed being:
+% \begin{equation}
+%   speed = C \times M^{0.12}
+% \end{equation}
+
+ %\citep{sazima1990necrofagia} predicted that aquatic or semi-aquatic piscivorous snakes would scavenge more frequently than other species.
+Moreover, \cite{sazima1990necrofagia} suggested that water currents that induce aggregations of carrion heighten the probability of carrion detection by these snakes (see also \cite{savitzky1992laboratory}).
+Also, chemical gradients may be more uniform and give more de-pendable directional information in water \citep{sazima1990necrofagia}." \citep{devault2002scavenging}
 

### 2016-03-18 natural history of scavenging.tex
added refs

--- 2016-03-16 natural history of scavenging.tex
@@
 
+%TG: what do you mean by their position? Phylogenetic? Position on the "scavenging scale"?
+
 Species capable of flight have effectively added an extra spatial dimension, i.e.
@@
 
-Avian flight originates in the Jurassic Period, conincident with the fossils of \textit{Archaeopteryx lithographica} so many of these benefits would have been realised from that point on for carnivorous birds.
-However, vertebrate flight is much older than this where pterosaurs predate bird origins by a considerable margin in the Triassic Period.
+Avian flight originates in the Late Jurassic (163.5-145 Million years ago - Mya), conincident with the fossils of \textit{Archaeopteryx lithographica} so many of these benefits would have been realised from that point on for carnivorous birds.
+However, vertebrate flight is much older than this where pterosaurs predate bird origins by a considerable margin in the Late Triassic (235-201.3 Mya).
 Scavenging in this diverse group has been hypothesied many times before.
 Certain clades of these animals could reach enormous sizes (e.g.
-Azhdarchids with wingspans of 11 metres) and look to have engaged in soaring flight \citep{witton2010size}.
+Azhdarchids with wingspans of 11 metres \citep{witton2010size})%CITE
+ and look to have engaged in soaring flight \citep{witton2010size}.
 Although \cite{witton2008reappraisal} argued that neck inflexibility and straight, rather than hooked jaw morphology points against their existing as \textit{obligate} scavengers, Azhdarchid terrestrial proficency indicates they would have been comfortable foraging on the ground.
-Indeed, extant Marabou Storks have a comparable morphology and are noted facultative scavengers so it is reasonable to believe that certain pterosaurs behaved similarly.
+Indeed, extant Marabou Storks (\textit{Leptoptilos crumenifer}) %TG: Check if that's the one
+have a comparable morphology and are noted facultative scavengers so it is reasonable to believe that certain pterosaurs behaved similarly.
 
@@
 Cinereous Vultures (\textit{Aegypius monachus}) and condors (\textit{Vultur gryphus}, \textit{Gymnogyps californianus}) all have body masses that can exceed 10 kg and represent some of the heaviest bird species capable of flight \citep{ferguson2001raptors,donazar2002effects}.
-And as we have noted the Azhdarchid pterosaurs were far bigger again, with estimated body masses of around 80 kg.
+And as we have noted the Azhdarchid pterosaurs were far bigger again, with estimated body masses of over 200 kg \citep{witton2010size}.
+%CITE
 
 The only other vertebrate group capable of powered flight are the bats where scavenging has not been recorded to our knowledge.
-Their visual acuity is famously poor and echolocation does not lend itself to discovering immobile carrion.
-Their small size and poor terrestrial ability would also count against them at a carcass.
-The bat fossil record is notoriously poor owing to their fragile skeletons so we are unable to determine if extinct species were more suited to this lifestyle.
+Their visual acuity is famously poor and echolocation would not lend itself to discovering immobile carrion.
+Their small size and poor terrestrial ability would also count against them at a carcass \citep{riskin2006terrestrial}.
+The bat fossil record is notoriously poor owing to their fragile skeletons so we are unable to determine if extinct species were more suited to this lifestyle \citep{eiting2009global}.
+%TG: Two points here: one speculation for the non-scavenging behaviour could be because of diet.
+Big bats (which are better suited for scavenging, following previous paragraph) are essentially frugivores and therefore probably lack of adaptations for digesting the meat.
+Also carnivores bats are mainly the microbats that are insectivores.
+Maybe they scavenge on dead insects but here we're talking about scavenging on real meat right?
+%TG: Second point: bats fossil record is of course not so good but it's actually way better than birds!
+Most of the data we have is on teeth of course but it's probably enough for having a rough body mass and diet estimate!
+And that's much better than birds where you're stuck with just some fragments of long bones.
 
 \section*{Terrestrial Scavengers}
-A simplificaiton of terrestrial, vertebrate scavengers is one of them existing in a two-dimensional plane while foraging for carrion directly.
+A simplification of terrestrial, vertebrate scavengers is one of them existing in a two-dimensional plane while foraging for carrion directly.
 They can detect carcasses at a range that is defined by the radius of their sensory organs, usually the visual and olfactory senses.
@@
 
-The order Carnivora sees its origins in the Eocene where it split into the Caniforma and Feliforma.
+The order Carnivora sees its origins in the Middle Eocene (56-33.9 Mya) where it split into the Caniforma and Feliforma.
 And many of the aforesaid adapations for scavenging are found in these other major terrestrial mammalian carnviores.
@@
 In some historical hominin-inhabited areas there were a greater number of felids than hyenids.
-This is significant because hyenas will leave far less flesh on a carcass than a felid such as a sabretooth enabling contemperaneous hominins to benefit \citep{pobiner2015new}.
+This is significant because hyenas are likely to have left far less flesh on a carcass than a felid such as a sabretooth enabling contemperaneous hominins to benefit \citep{pobiner2015new}.
+%TG: or "could"? This is hypothetical behaviour stuff! Maybe past hyenas where exclusive sushi eaters!
+%TG: Not sure about the intelligence thing: the sentence as the same meaning without and will be less anthropocentric sounding.
 The intelligence, resultant tool-use and cooperative nature of hominins meant they could likely adapt to take on more or less carrion depending on their environment.
@@
 
-% Slight changes in temperature could have profound impacts on energy flow within ecosystems as carrion decomposition
-% rates roughly double for every 10°C increase in ambient temperature (Vass et al. 1992; Parmenter and
-% MacMahon 2009). Consequently, the availability of carrion to vertebrate scavengers could be reduced by
-%as much as 20%–40% over the next century based on projections from current climate change models
-% (Beasley et al. 2012b).
-
-%"Changes in global climate are also expected to substantially alter the availability of carrion resources in marine ecosystems as surface production of organic material could decline by 50\% or more (Smith et al.
-2008).
-Such a drastic reduction in productivity undoubtedly will impact a multitude of ecosystem processes, including the availability and distribution of carrion within marine ecosystems." \citep{benbow2015introduction}
-
-%"For example, vertebrate scavengers appear to be disadvantaged when humidity and temperature favor microbial and invertebrate reproduction (between 21C and 38C)" \citep{benbow2015introduction}
-
-%"In a study by DeVault et al.
-(2004b), it was found that at temperatures above 20C vertebrates were able to detect and consume only 19\% of small-mammal carcasses, whereas at temperatures below 18C, vertebrates consumed 49\% of carcasses" \citep{benbow2015introduction}
-
-%"For example, the available literature suggests that carcass decomposition occurs more rapidly in warm, moist habitats when compared with more xeric habitats" \citep{benbow2015introduction}
-
-%Of particular note is the increase in decomposition rate of carrion with temperature which could result in a seasonal reduction of carrion availability given predicted warming trends \citep{moleon2014inter}.
-
-%"Apart from being an important water-conservation strategy, nocturnal behavior may have evolved in the Hyaenidae as a means of reducing competition with the other dominant scavengers in African ecosystems, the vultures, which are exclusivly diurnal (Houston 1979)" Carnivore Behavior, Ecology, and Evolution By John L.
-Gittleman
-
-% "Thus, on an annual basis, the FMR (and food requirements) of an endotherm are likely to be around 30 times greater than those for an equivalent-sized ectotherm"
-
-%"Maximal sustainable speeds of mammals are six- to sevenfold those of lizards of equal size,"
+% %TG: LaTeX format template for the equations in the R file
+% \begin{equation}
+%   \text{Travelling speed} = \frac{duration \times speed}{2} \times 10^3
+% \end{equation}
+% where \textit{C} is a constant of $1.15$ for mammals and $0.23$ for squamates, %TG: if you meant squamates from reptiles
+% and speed being:
+% \begin{equation}
+%   speed = C \times M^{0.12}
+% \end{equation}
 
@@
 %(Since we are intrested in scavanging in th paleo record ecosystem services might not be that relavent, although modulators of disease is still relavent)
-%Despite this, scavengers are a seriously understudied group \citep{sekercioglu2006increasing,selva2007nested,wilson2011scavenging}.  
-%\cite{devault2003scavenging} propose that this is due to both human disgust at carrion itself and the difficulty in determining if an ingested prey item was killed or scavenged. 
-%The latter point means that studying the natural history of this behaviour is particularly problematic.
-%Indeed, even data on the proportion of carrion in the diet of extant species are sorely lacking \citep{benbow2015introduction}.
-
-%"Indeed, in some ecosystems, vertebrates have been documented to assimilate as much as 90\% of the available carrion"\citep{benbow2015introduction}.
 
@@
 
-%Our review is divided up into three sections, namely the land, air and sea. 
-%These are then subdivided into three geological eras the Cenozoic, Mesozoic and Palaeozoic. 
-%Each of these environments has a distinct phyiscal character that affects how a species forages for food which has obvious relevance for an animal searching for carrion.  
-%However, there is some commonality to a scavenger's environment and the problems in finding food that one would encounter. 
-%Notably, the resource environment of a scavenger is a patchily distributed one, because it is difficult to predict both when and where a carcass is produced. 
-%As a result of this, any animal existing as a scavenger must maximise its detection capabilities and minimise its locomotory costs \citep{ruxton2004obligate}.
-%Exploring which groups are likely to have moved towards these traits and thus existed as sccavengers over palaentological time forms the basis of this work.
 %remember the journal is an ecological one at heart so I would always keep in mind about making it useful for ecologists over paelo people.
@@
 Studies in ecological genetics and historical ecology are welcomed in the context of explaining contemporary ecological patterns".
-
-%"Interkingdom Competition among Vertebrates, Invertebrates, and Microbes"\citep{benbow2015introduction}
-%"For example, vertebrate scavengers appear to be disadvantaged when humidity and temperature favor microbial and invertebrate reproduction" \citep{benbow2015introduction}
 
@@
 
-% "Changes in global climate are also expected to substantially alter the availability of carrion resources in marine ecosystems as surface production of organic material could decline by 50% or more (Smith et al.
-2008).
-Such a drastic reduction in productivity undoubtedly will impact a multitude of ecosystem processes, including the availability and distribution of carrion within marine ecosystems." \citep{benbow2015introduction}
-
-%\subsection*{Cenozoic}
-%Among terrestrial African carnivores, hyenas, jackals, lions and leopards all take sizable proportions of carrion in their diet.
-%In the case of the spotted hyena (\textit{Crocuta crocuta}) it can be as high as 99\% \citep{benbow2015introduction}. 
-%Yet, no contemporary terrestrial vertebrate exists as an obligate scavenger. 
-%The selective pressures that push mammals and reptiles towards scavenging do not seem to undermine their ability to hunt, perhaps explaining the absence of obligate scavengers in these groups \citep{ruxton2004obligate}.
-%\\There is a long running debate on the tendency of human ancestors to scavenge. 
-%Some recent studies have found "that passive scavenging from abandoned larger felid kills could have been a high-yield, though potentially dangerous, foraging strategy for early hominins even without considering within-bone nutrients" \citep{pobiner2015new}.
-
-%Osteophagy is known across a range of terrestrial carnivores.
-%Some fat-rich mammalian bones have an energy density (6.7 kJ/g) comparable with that of muscle tissue, making skeletal remains an enticing resource \citep{brown1989study}. 
-% Hyenas have a bite force capable of breaking open bones and 
-% In light of this, the skeletal remains of carrion may act as trove of food to carnivores who can access it.
-
 %"Of all the mammalian carnivores in Africa, brown, striped, and spotted hyenas (hereafter referred to simply as hyenas) derive the largest portion of their diet from scavenging.
@@
 
-% A review by \cite{devault2002scavenging} found occurrences of scavenging spread across five families of snakes and states that this behaviour is "far more common than currently acknowledged."
-
-%"\cite{sazima1990necrofagia} predicted that  several components  of  snake  be-havior contribute to scavenging propensity.
-They  postulated  that  snakes  that  rely  on chemosensory information to acquire prey might forage  for  carrion  more  frequently than  those  whose  modes  of prey acquisition are driven by visual cues.
-Additionally, \cite{sazima1990necrofagia} suggested that habitat and diet preferences might be indicators of scavenging   propensity   be-cause  they  influence  the  likelihood of a particular species encountering carrion naturally.
-Based on these criteria, they pre-dicted that aquatic or semi-aquatic pisciv-orous snakes  would  scavenge  more  fre-quently than other species.
-Moreover, \cite{sazima1990necrofagia}  suggested that water currents that induce  aggrega-tions of carrion heighten the probability of carrion detection by these snakes (see also \cite{savitzky1992laboratory}).
-Also,  chemical  gradients may  be  more  uniform  and  give  more  de-pendable  directional information in water \citep{sazima1990necrofagia}." \citep{devault2002scavenging}
-
-%"Investigations of ophidian metabolic re-quirements unveil additional advantages to carrion  utilization.
-Snakes  exhibit  exceed-ingly  low  maintenance  metabolisms,  and most  can  survive  on  a  few  scant  feedings per year.
-It  is, therefore, possible for snakes to rely large-ly  on  infrequent,  less  energy-rich  meals.
-Carrion, which is by nature ephemeral and unpredictable, may represent such a food source  to  snakes.
-Scavenging  might  allow snakes to meet their low metabolic needs more  easily  and  without  the  costs  associ-ated with subduing prey." \citep{devault2002scavenging}
-
 %\textit{Allosaurus} tooth marks on a hadrosaur in the Late Jurassic. 
 % Late Triassic scavenging on a prosauropod by basal carnivorous archosaurs \citep{hungerbuhler1998taphonomy}.
-
-%\subsection*{Palaeozoic}
-% The absence of flying vertebrates in the Palaeozoic may have permitted terrestrial forms to take in a higher proportion of carrion in their diet.
-
-%\subsection*{Cenozoic}
 %A likely instance of scavenging between a 4-million-year-old white shark (\textit{Carcharodon}) and mysticete whale from Peru \citep{ehret2009caught}.
@@
 %Possible scavenging on a juvenile fur seal from the Late Neogene \citep{boessenecker2011mammalian}.
-
-%\subsection*{Mesozoic}
-
-%\subsection*{Palaeozoic}
-%Evolution of sharks, known scavengers.
-
-%\section*{Results}
-
-%\section*{Discussion}
 

### 2016-03-22 natural history of scavenging.tex
Added vultures age
added Graeme's ideas

@@
 Fortunately, research on scavenging is on the rise \citep{manga2006vulture}.
-As a result, we are now beginning to realise the extent of this behaviour such that, “in some ecosystems, vertebrates have been documented to assimilate as much as 90\% of the available carrion" \citep{benbow2015introduction}.
+As a result, we are now beginning to realise the extent of this behaviour such that, “in some ecosystems, vertebrates have been documented to assimilate as much as 90\% of the available carrion" \citep{beasley2015vertebrates}.
 Even Tenney’s noble big cats are now known to take in a significant portion of carrion in their diet where some lion populations get over 50\% of their meat from carcasses.
 By recognising the difficulty in directly observing scavenging, a suite of methods have been used to discern the most suitable morphologies, physiologies and environments for a scavenging lifestyle to prosper.
-Here we chart the natural history of vertebrate scavenging on air, land and in water by focussing on the dominant vertebrate groups.
+Here we chart the natural history of vertebrate scavenging on air, land and in water by focussing on the dominant vertebrate groups that existed at each time.
 %It is our aim in this review to employ these methods to gain an understanding of scavengers past and present.
@@
 The chief hurdle to scavenging is finding a sufficient quantity of food, the occurence of which is difficult to predict in space and time.
-Once found the scavenger has to then overcome the agents of decay produced by the action of microorganisms on the carcass \citep{ruxton2014fruit}.
+Once found, the scavenger has to then overcome the agents of decay produced by the action of microorganisms on the carcass \citep{ruxton2014fruit}.
 The idea of scrounging from predator kills is undermined by studies showing that in the  majority of ecosystems more animals die from disease and starvation than predation \citep{benbow2015introduction}.
@@
 \section*{Aerial Scavengers}
+Scavenging should be particularly attractive to avian predators compared to mammals.
+Solitary mammalian predators can kill prey up to the same body mass as themselves and sometimes an order of magnitude heavier (e.g.
+socially hunting lions \citep{owen2008predator}) .
+In contrast, birds of prey tend to kill prey smaller than themselves \citep{slagsvold2007prey}.
+This is likely due to their need to kill prey that they can fly away with, as well as the risk of injury being higher (which carries a higher mortality risk) for a bird than a mammal.
+Thus for birds, scavenging means they can exploit species that would otherwise be too big for them to kill.
 
@@
 These birds consist of two convergent groups, from the old and the new world and represent the only example of obligate vertebrate scavengers today.
+The families from which modern vultures arose, the Accipitridae and Cathartidae, appear during the Palaeocene (66 - 56 Mya) \citep{Jetz2012, Jarvis2014}.
 Given their unique position, they have been extensively studied to determine what adaptations they possess that allows them to so flourish in this niche.
@@
 This advantage can be extended further in larger species by engaging in soaring instead of flapping flight, which is even cheaper still (approximately twice BMR) \citep{hedenstrom1993migration}.
-The benefits this confers are clear from the information we have on the enormous foraging ranges of some seabirds and accipiters \citep{bamford2007ranging}.
+The benefits this confers are clear from the information we have on the enormous foraging ranges of many vultures \citep{spiegel2013factors}.
 Clearly, it would be pointless to have incredible detection abilites and not have a cost efficient movement to benefit from it.
@@
 Certain clades of these animals could reach enormous sizes (e.g.
-Azhdarchids with wingspans of 11 metres \citep{witton2010size})%CITE
- and look to have engaged in soaring flight \citep{witton2010size}.
+Azhdarchids with wingspans of 11 metres \citep{witton2010size}) and look to have engaged in soaring flight \citep{witton2010size}.
 Although \cite{witton2008reappraisal} argued that neck inflexibility and straight, rather than hooked jaw morphology points against their existing as \textit{obligate} scavengers, Azhdarchid terrestrial proficency indicates they would have been comfortable foraging on the ground.
@@
 As a consequence, they have a much more restricted view of the landscape than do aerial foragers.
-No contemporary terrestrial vertebrate exists as an obligate scavenger but most if not all are facultative scavengers to some extent.
+No contemporary terrestrial vertebrate exists as an obligate scavenger but most if not all are facultative scavengers to some extent \citep{beasley2015vertebrates}.
 \cite{ruxton2004obligate} offer a reason for this in that the traits that allow for vultures to exist as scavengers undermined their ability to hunt but that the same forces have not prevented mammals from doing so.
-The same authors calculated that "a 1 tonne mammal or reptile, in an ecosystem yielding carrion at densities similar to the current Serengeti, could have met its energy requirements if it could detect carrion over a distance of the order of 400–500 m."\citep{ruxton2004obligate}.
+Interestingly, the same authors calculated that "a 1 tonne mammal or reptile, in an ecosystem yielding carrion at densities similar to the current Serengeti, could have met its energy requirements if it could detect carrion over a distance of the order of 400–500 m."\citep{ruxton2004obligate}.
 
 Terrestrial scavenging in the mammals is probably best known in an African context where hyenas, jackals and lions all take sizable proportions of carrion in their diet.
-In the spotted hyena (\textit{Crocuta crocuta}), striped hyena (\textit{Hyaena hyaena}) and brown hyena (\textit{Hyaena brunnea}) it can be as high as 99\% \citep{benbow2015introduction}.
+In the spotted hyena (\textit{Crocuta crocuta}), striped hyena (\textit{Hyaena hyaena}) and brown hyena (\textit{Hyaena brunnea}) it can be over 90\% \citep{jones2015african}.
 Therefore, we can again use these species as our efficient terrestrial scavengers to compare with other forms.
 
-Similar to vultures they have well developed sensory organs, particularly in olfaction whereby they can detect a rotting carcass 2 km downwind.
-They have a characteristic "rocking horse gait"  which allows them to cover great distances efficiently.
+Similar to vultures they have well developed sensory organs, whereby they can smell a rotting carcass 4 km away and can hear the vocalisations of conspecifics at a distance of 10 km \citep{mills1989comparative}.
+They have a characteristic "rocking horse gait"  which allows them to cover great distances efficiently \citep{jones2015african}.
 A major difficulty for terrestrial scavengers is competition with vultures.
@@
 Being able to extract nutrients from remnants gives the scavenger a great advantage.
-Osteophagy is known across a range of terrestrial carnivores.
-Some fat-rich mammalian bones have an energy density (6.7 kJ/g) comparable with that of muscle tissue, making skeletal remains an enticing resource \citep{brown1989study}.
+Osteophagy is known across a range of terrestrial carnivores and given some fat-rich mammalian bones have an energy density (6.7 kJ/g) comparable with that of muscle tissue, it makes skeletal remains an enticing resource \citep{brown1989study}.
 This ability reached its zenith among hyenas with the evolution of the 110 kg \textit{Pachycrocuta brevirostris} during the Pliocene \citep{palmqvist2011giant}.
 Some work on extinct sabretooths suggests they may have left a large amount of food for would-be scavengers because of their unique skull morphology.
-As a result the decline of Machairodontinae sabretooths has been offered as an explanation for the extinction of \textit{P brevirostris} \citep{palmqvist2011giant}.
+As a result, the decline of Machairodontinae sabretooths has been offered as an explanation for the extinction of \textit{P brevirostris} \citep{palmqvist2011giant}.
 
@@
 
-Interestingly such comparisons have given insight into the feeding ecology of early hominins who, for instance, had the ability to craft tools for breaking open bones \citep{hone2010feeding,ARCM:ARCM12084}.
+Interestingly such comparisons have given insight into the feeding ecology of early hominins who, for instance, had the ability to craft tools for breaking open bones \citep{ARCM:ARCM12084}.
 The question of where our ancestors placed on the hunter-scavenger axis during the Plio-Pleistocene has been a matter of debate for years.
@@
 %TG: Not sure about the intelligence thing: the sentence as the same meaning without and will be less anthropocentric sounding.
-The intelligence, resultant tool-use and cooperative nature of hominins meant they could likely adapt to take on more or less carrion depending on their environment.
+The intelligence, resultant tool-use and cooperative nature of hominins meant they could likely adapt to take on more or less carrion depending on their environment \citep{moleon2014humans}.
 
@@
 It  is, therefore, possible for snakes to rely largely  on  infrequent,  less  energy-rich  meals." In the same review the authors found occurrences of scavenging spread across five families of snakes and stated that this behaviour is "far more common than currently acknowledged."\citep{devault2002scavenging}.
+The same reasoning can be applied to crocodlies and their allies \citep{forrest2003evidence}.
 
@@
 They are quite alien to anything we know today which restricts our ability to understand their ecology far more so than extinct mammals \citep{weishampel2004dinosauria}.
-Of relevance, are the questions that still persist about their metabolism \citep{grady2014evidence} and sensory perception.
-We do know that they walked with the erect gait of mammals or birds rather than the sprawling gait of lizards and that they were most likely facultative scavengers \citep{depalma2013physical}.
+Of relevance, are the questions that still persist about their metabolism \citep{grady2014evidence} and sensory perception \citep{farlow1994speculations}.
+We do know that they walked with the erect gait of mammals or birds rather than the sprawling gait of lizards and that they were most likely facultative scavengers \citep{weishampel2004dinosauria,depalma2013physical}.
 Much work has focused on the existence of the behaviour in \textit{Tyrannosaurus rex} \citep{ruxton2003could,carbone2011intra} but a recent energetics study investigated the likely prevelance of scavenging across a range of body sizes.
@@
 the robust skull and dentition of \textit{T.
-rex}.
+rex} \citep{hone2010feeding}.
 There is direct evidence that \textit{T.
@@
 As noted earlier, a system similar to the Serengeti in productivity could have supported a terrestrial scavenger \citep{ruxton2004obligate}.
+Indeed, in systems that were dominated by large ectothermic or mesothermic vertebrates
+the same primary productivity would have supported a greater biomass \citep{mcnab2009resources}.
+The upshot of this is there was a higher biomass of herbivores dying and offering scavenging opportunities.
+Predators were large bodied too compared to extant mammalian predators \citep{mcnab2009resources}, and so, especially if they were ectothermic, could last longer between meals rendering scavenging a more attractive behaviour relative to predation.
 
 \section*{Aquatic Scavengers} 
-
+Primary productivity is lower in almost all aquatic systems than terrestrial systems (except deserts) so as we go up the food chain the density of carcasses worth scavenging is going to be lower.
+But it is certainly known to occur in many aquatic vertebrates.
 An aquatic environment presents challenges for direct observational studies and so, similar to the approaches involving extinct species, much work has approached the question of scavenging propensity from an energetics perspective.
-Although as with the aerial and terrestrial enviornments we have evidence of facultative scavenging among aquatic species.
-For example, the remains of a mosasaur and a terrestrial hadrosaur were discovered with embedded teeth from a Cretaceous shark \textit{Squalicorax} \citep{schwimmer1997scavenging}.
+Although, as with the aerial and terrestrial enviornments we have evidence of facultative scavenging among aquatic species.
+For example, the remains of a mosasaur and a terrestrial hadrosaur were discovered with embedded teeth from a Cretaceous shark, \textit{Squalicorax} \citep{schwimmer1997scavenging}.
 As well as a likely instance of scavenging between a 4-million-year-old white shark (\textit{Carcharodon}) and mysticete whale from Peru \citep{ehret2009caught}.
 The existence of an obligate scavenger in a marine setting is uncertain \citep{britton1994marine,smith2003ecology,ruxton2004energetic,ruxton2005searching}.
-Depending on the species a carcass in this environment either floats or descends to the sea floor.
-In the latter low-light environment, detection distances are far lower (< 100 m) than they would be in the air.
+Depending on the species, a carcass in this environment either floats or descends to the sea floor \citep{Whitehead415}.
+In the latter low-light environment, visual detection distances are far lower (< 100 m) than they would be in the air.
 As such, animals detect resources through chemo- and mechanoreception more so than through vision \citep{ruxton2004energetic}.
 However, water is a medium that is conducive to low-cost movement \citep{tucker1975energetic} and so may be able to support an obligate scavenging fish \citep{ruxton2004energetic,ruxton2005searching}.
-\cite{benbow2015introduction} do note that "some benthic scavengers (e.g., hagfish: family Myxinidae) rely on necrophagy for a large portion of their diet and may indeed be obligate scavengers".
+\cite{beasley2015vertebrates} do note that "some benthic scavengers (e.g., hagfish: family Myxinidae) rely on necrophagy for a large portion of their diet and may indeed be obligate scavengers".
 
@@
 One point of interest is that of the whaling industry, which provided a bonanza of floating carcasses especially during the 20th century \citep{Whitehead415}.
-This meant Killer Whales \textit{Orcinus orca} could switch from hunting to scavenging, a switch made that much easier by the noise of the whaling vessels that would effectively ring the "dinner-bells" \citep{Whitehead415}.
+This meant Killer Whales (\textit{Orcinus orca}) could switch from hunting to scavenging, a switch made that much easier by the noise of the whaling vessels that would effectively ring the "dinner-bells" \citep{Whitehead415}.
 Early whales such as \textit{Basilosaurus} seem to fit into the same niche as Killer Whales and we have some evidence for scavenging in this group as well \citep{fahlke2012bite}.
+A final point to make is that vertebrates are relatively rarer in aquatic environments, because even large animals can get support from the buoyancy of the water without needing a backbone.
 
@@
 It is recognised that scavengers keep energy flows at a higher trophic level in food webs than decomposers because they consume relatively more carrion \citep{devault2003scavenging}.
-They are also hugely important for the dispersal of nutrients \citep{benbow2015introduction}.
+They are also hugely important for the dispersal of nutrients \citep{beasley2015vertebrates}.
 Consider the diversity of animals that can end up feeding at the carcass of an elephant.
@@
 The geological record shows the Earth has undergone radical fluctuations in temperature.
-This will have had a significant bearing on the availability of carrion.
-A 10$^{\circ}$C increase in ambient temperature can double carcass decomposition rates \citep{benbow2015introduction}.
-Oceanic productivity, decomposition rates and habitat structure are all impacted by climactic conditions.
-\cite{benbow2015introduction} points to a series of studies showing how microbes and invertebrates benefit at higher temperatures to the detriment of vertebrate scavengers "above 20$^{\circ}$C vertebrates were able to detect and consume only 19\% of small-mammal carcasses, whereas at temperatures below 18$^{\circ}$C, vertebrates consumed 49\% of carcasses".
-Furthermore, "carcass decomposition occurs more rapidly in warm, moist habitats when compared with more xeric habitats."
+This will have had a significant bearing on the availability and persistence of carrion.
+To illustrate the point, a 10$^{\circ}$C increase in ambient temperature can double carcass decomposition rates \citep{parmenter2009carrion} and geological evidence indicates that the Mesozoic Earth was at least 6 $^{\circ}$C warmer than now \citep{sellwood2006mesozoic}.
+In terms of specific habitats, it has been shown that decomposition is greater in warm and moist areas versus more xeric ones \citep{beasley2015vertebrates}.
+Moreover, oceanic productivity and habitat structure are all impacted by climactic conditions.
+The impacts these can have on scavengers have been empirically supported e.g.
+\cite{beasley2015vertebrates} who point to a series of studies showing how microbes and invertebrates benefit at higher temperatures to the detriment of vertebrate scavengers such that "above 20$^{\circ}$C vertebrates were able to detect and consume only 19\% of small-mammal carcasses, whereas at temperatures below 18$^{\circ}$C, vertebrates consumed 49\% of carcasses".
 

### 2016-03-22 natural history of scavenging.tex
tidied up affiliations

@@
 \affiliations{
-\item A. Kane (\url{adam.kane@ucc.ie}), University College Cork, Cooperage Building\, School of Biological 
-Earth and Environmental Sciences\,Cork, Ireland.
-\item K. Healy and A. Jackson, Trinity College Dublin, Department of Zoology;
-School of Natural Sciences, Dublin Ireland.
-\item T. Guillerme, Imperial College London, Silwood Park Campus, Department of Life Sciences, Buckhurst Road, Ascot SL5 7PY, UK.
-%TG: hehehe!
-\item G. Ruxton, School of Biology, Sir Harold Mitchell Building, Greenside Place, St Andrews, KY16 9TH,
-United Kingdom
+\item A. Kane (\url{adam.kane@ucc.ie}), University College Cork, Cooperage Building, School of Biological Earth and Environmental Sciences, Cork, Ireland.
+\item K. Healy and A. Jackson, Trinity College Dublin, Department of Zoology, School of Natural Sciences, Dublin Ireland.
+\item T. Guillerme, Imperial College London, Silwood Park Campus, Department of Life Sciences, Buckhurst Road, Ascot SL5 7PY, United Kingdom.
+\item G. Ruxton, School of Biology, Sir Harold Mitchell Building, Greenside Place, St Andrews, KY16 9TH, United Kingdom.
 }

### 2016-03-22 natural history of scavenging.tex
Added new title + comments + early tetrapods story

@@
 
-\title{The natural history of scavenging in vertebrates}
+\title{A recipe for scavenging and natural history}
 \running{Scavenging in vertebrates}
@@
 This may go some way to explaining the gap in our knowledge of the prevalence of this behaviour.
-Consider Professor Sanborn Tenney writing in 1877 for The American Naturalist who had this to say about one well known group, “Prominent among the mammalian scavengers are the hyenas, the ugliest in their general appearance of all the flesh eaters.” He contrasts these with “nobler kinds” of carnivores such as lions and tigers \citep{tenney1877few}.
+Consider Professor Sanborn Tenney writing in 1877 for The American Naturalist who had this to say about one well known group, ``Prominent among the mammalian scavengers are the hyenas, the ugliest in their general appearance of all the flesh eaters." He contrasts these with ``nobler kinds" of carnivores such as lions and tigers \citep{tenney1877few}.
 Even aside from our own subjective biases, scavenging is a difficult behaviour to detect after the fact.
@@
 Fortunately, research on scavenging is on the rise \citep{manga2006vulture}.
-As a result, we are now beginning to realise the extent of this behaviour such that, “in some ecosystems, vertebrates have been documented to assimilate as much as 90\% of the available carrion" \citep{beasley2015vertebrates}.
+As a result, we are now beginning to realise the extent of this behaviour such that, ``in some ecosystems, vertebrates have been documented to assimilate as much as 90\% of the available carrion" \citep{beasley2015vertebrates}.
 Even Tenney’s noble big cats are now known to take in a significant portion of carrion in their diet where some lion populations get over 50\% of their meat from carcasses.
@@
 %It is our aim in this review to employ these methods to gain an understanding of scavengers past and present.
-\subsection*{The Difficulty of Scavenging}
+\subsection*{The Difficulty of Scavenging} %KH: Make that the template for all the next sections
 The chief hurdle to scavenging is finding a sufficient quantity of food, the occurence of which is difficult to predict in space and time.
@@
 The habitat must also be productive enough to sustain an animal biomass that will eventually produce carcasses.
+%KH: need paragraph for how do we know something is (was) a scavenger
+%TG: Insert figure here: scavenging scale + four points: Locomotion/Detection/Consumption/Environment
 
@@
 \cite{ruxton2004obligate} offer a reason for this in that the traits that allow for vultures to exist as scavengers undermined their ability to hunt but that the same forces have not prevented mammals from doing so.
-Interestingly, the same authors calculated that "a 1 tonne mammal or reptile, in an ecosystem yielding carrion at densities similar to the current Serengeti, could have met its energy requirements if it could detect carrion over a distance of the order of 400–500 m."\citep{ruxton2004obligate}.
+Interestingly, the same authors calculated that ``a 1 tonne mammal or reptile, in an ecosystem yielding carrion at densities similar to the current Serengeti, could have met its energy requirements if it could detect carrion over a distance of the order of 400–500 m."\citep{ruxton2004obligate}.
 
@@
 Similar to vultures they have well developed sensory organs, whereby they can smell a rotting carcass 4 km away and can hear the vocalisations of conspecifics at a distance of 10 km \citep{mills1989comparative}.
-They have a characteristic "rocking horse gait"  which allows them to cover great distances efficiently \citep{jones2015african}.
+They have a characteristic ``rocking horse gait"  which allows them to cover great distances efficiently \citep{jones2015african}.
 A major difficulty for terrestrial scavengers is competition with vultures.
@@
 Any adaptations that reduce energetic costs are likely to be selected in scavengers.
-\cite{devault2002scavenging} suggest this is an avenue for scavenging in snakes because they "exhibit  exceedingly  low  maintenance  metabolisms,  and most  can  survive  on  a  few  scant  feedings per year.
-It  is, therefore, possible for snakes to rely largely  on  infrequent,  less  energy-rich  meals." In the same review the authors found occurrences of scavenging spread across five families of snakes and stated that this behaviour is "far more common than currently acknowledged."\citep{devault2002scavenging}.
+\cite{devault2002scavenging} suggest this is an avenue for scavenging in snakes because they ``exhibit  exceedingly  low  maintenance  metabolisms,  and most  can  survive  on  a  few  scant  feedings per year.
+It  is, therefore, possible for snakes to rely largely  on  infrequent,  less  energy-rich  meals." In the same review the authors found occurrences of scavenging spread across five families of snakes and stated that this behaviour is ``far more common than currently acknowledged."\citep{devault2002scavenging}.
 The same reasoning can be applied to crocodlies and their allies \citep{forrest2003evidence}.
@@
 
+%TG: To be modified and inserted somewhere:
+Scavenging behaviour might have evolved as soon as the first terrestrial tetrapods evolved.
+In fact, some of the earlier tetrapods tracks dating back to the early Middle Devonian (393.3 - 387.7 Mya) where found in intertidal environments \citep{Niedzwiedzki2009}.
+These environments are isolated from marine systems twice a day leaving potential carrions unexploited by marine vertebrates.
+\cite{Niedzwiedzki2009} suggests that these environments ``would thus have allowed marine ancestors of tetrapods gradually to acquire terrestrial competence while accessing a new and essentially untouched resource.''
+%TG: orginal sentence "The intertidal environment provides a ready food source of stranded marine animals on a twice-daily basis, in the immediate vicinity of the sea, and would thus have allowed marine ancestors of tetrapods gradually to acquire terrestrial competence while accessing a new and essentially untouched resource."
+
 \section*{Aquatic Scavengers} 
@@
 However, water is a medium that is conducive to low-cost movement \citep{tucker1975energetic} and so may be able to support an obligate scavenging fish \citep{ruxton2004energetic,ruxton2005searching}.
-\cite{beasley2015vertebrates} do note that "some benthic scavengers (e.g., hagfish: family Myxinidae) rely on necrophagy for a large portion of their diet and may indeed be obligate scavengers".
+\cite{beasley2015vertebrates} do note that ``some benthic scavengers (e.g., hagfish: family Myxinidae) rely on necrophagy for a large portion of their diet and may indeed be obligate scavengers".
 
@@
 One point of interest is that of the whaling industry, which provided a bonanza of floating carcasses especially during the 20th century \citep{Whitehead415}.
-This meant Killer Whales (\textit{Orcinus orca}) could switch from hunting to scavenging, a switch made that much easier by the noise of the whaling vessels that would effectively ring the "dinner-bells" \citep{Whitehead415}.
+This meant Killer Whales (\textit{Orcinus orca}) could switch from hunting to scavenging, a switch made that much easier by the noise of the whaling vessels that would effectively ring the ``dinner-bells" \citep{Whitehead415}.
 Early whales such as \textit{Basilosaurus} seem to fit into the same niche as Killer Whales and we have some evidence for scavenging in this group as well \citep{fahlke2012bite}.
@@
 
-\section*{Ecological Role}
-It is recognised that scavengers keep energy flows at a higher trophic level in food webs than decomposers because they consume relatively more carrion \citep{devault2003scavenging}.
+% Maybe drop this bit
+% \section*{Ecological Role}
+% It is recognised that scavengers keep energy flows at a higher trophic level in food webs than decomposers because they consume relatively more carrion \citep{devault2003scavenging}.
 They are also hugely important for the dispersal of nutrients \citep{beasley2015vertebrates}.
@@
 In the case of mammals and reptiles these are 1.15 * body mass (kg) \textsuperscript{0.12} and 0.23 * body mass (kg) \textsuperscript{0.12}
-respectively \citep{ruxton2004obligate}.
+% respectively \citep{ruxton2004obligate}.
 We can insert these into a foraging radius model ((duration * speed)/2)/1000 for a 12 hour foraging day which shows that while a 10 kg reptile can range 6.5 km an equally sized mammal can range nearly 33 km \citep{Enstipp2006Energetics}.
@@
 
-\section*{The Role of Climate}
+%\section*{The Role of Climate} %TG: Use it as a opening at the end of the paper (i.e. conclusion + finishing with this bit)
 The geological record shows the Earth has undergone radical fluctuations in temperature.
@@
 The impacts these can have on scavengers have been empirically supported e.g.
-\cite{beasley2015vertebrates} who point to a series of studies showing how microbes and invertebrates benefit at higher temperatures to the detriment of vertebrate scavengers such that "above 20$^{\circ}$C vertebrates were able to detect and consume only 19\% of small-mammal carcasses, whereas at temperatures below 18$^{\circ}$C, vertebrates consumed 49\% of carcasses".
+\cite{beasley2015vertebrates} who point to a series of studies showing how microbes and invertebrates benefit at higher temperatures to the detriment of vertebrate scavengers such that ``above 20$^{\circ}$C vertebrates were able to detect and consume only 19\% of small-mammal carcasses, whereas at temperatures below 18$^{\circ}$C, vertebrates consumed 49\% of carcasses".
+% Add bits on climate change/dectection rate change, etc...
 
@@
 %"Cross-habitat nutrient transport can produce a variety of important outcomes in recipient systems (e.g., Polis et al.
-2004), and scavengers can play a significant role in moving “ecologi -cal subsidies” between habitats.
+2004), and scavengers can play a significant role in moving ``ecologi -cal subsidies" between habitats.
 For example, the use of ocean-derived carrion by terrestrial mammals (Rose and Polis 1998) and birds (Schlacher et al.
@@
 This is not surprising, as they show many unique adaptations specifically for scavenging.
-They have a unique body posture and a “rocking-horse” gait (Eloff 1964; Tilson and Henschel 1986; Hofer and East 1993a; Frank 1996) that allows them to cover long distances in search of carrion and prey."\citep{benbow2015introduction}
+They have a unique body posture and a ``rocking-horse" gait (Eloff 1964; Tilson and Henschel 1986; Hofer and East 1993a; Frank 1996) that allows them to cover long distances in search of carrion and prey."\citep{benbow2015introduction}
 

### 2016-03-22 natural history of scavenging.tex
Single lined each sentence. Makes it look ugly but easier to track the edits.

@@
 This may go some way to explaining the gap in our knowledge of the prevalence of this behaviour.
-Consider Professor Sanborn Tenney writing in 1877 for The American Naturalist who had this to say about one well known group, ``Prominent among the mammalian scavengers are the hyenas, the ugliest in their general appearance of all the flesh eaters." He contrasts these with ``nobler kinds" of carnivores such as lions and tigers \citep{tenney1877few}.
+Consider Professor Sanborn Tenney writing in 1877 for The American Naturalist who had this to say about one well known group, ``Prominent among the mammalian scavengers are the hyenas, the ugliest in their general appearance of all the flesh eaters."
+He contrasts these with ``nobler kinds" of carnivores such as lions and tigers \citep{tenney1877few}.
 Even aside from our own subjective biases, scavenging is a difficult behaviour to detect after the fact.
@@
 %It is our aim in this review to employ these methods to gain an understanding of scavengers past and present.
+
 \subsection*{The Difficulty of Scavenging} %KH: Make that the template for all the next sections
@@
 The habitat must also be productive enough to sustain an animal biomass that will eventually produce carcasses.
+
 %KH: need paragraph for how do we know something is (was) a scavenger
@@
 Scavenging should be particularly attractive to avian predators compared to mammals.
-Solitary mammalian predators can kill prey up to the same body mass as themselves and sometimes an order of magnitude heavier (e.g.
-socially hunting lions \citep{owen2008predator}) .
+Solitary mammalian predators can kill prey up to the same body mass as themselves and sometimes an order of magnitude heavier (e.g. socially hunting lions \citep{owen2008predator}).
 In contrast, birds of prey tend to kill prey smaller than themselves \citep{slagsvold2007prey}.
@@
 
-Species capable of flight have effectively added an extra spatial dimension, i.e.
-the vertical component, to their sensory environment over land animals.
+Species capable of flight have effectively added an extra spatial dimension, i.e. the vertical component, to their sensory environment over land animals.
 This allows them to look down on a landscape where they are unencumbered by obstacles that would obstruct the view of a terrestrial scavenger.
@@
 Vultures and eagles tend to soar using thermals and if these air pockets don't form, say on a cloudy day, the bird is grounded \citep{mundy1992vultures}.
-In many habitats (e.g.
-the arctic) it is simply not possible for sufficiently powerful thermals to form and as a consequence large-bodied vultures cannot exist.
+In many habitats (e.g. the arctic) it is simply not possible for sufficiently powerful thermals to form and as a consequence large-bodied vultures cannot exist.
 The upshot of this is that terrestrial carnivores like bears and wolves take more carrion \citep{devault2003scavenging}.
@@
 Scavenging in this diverse group has been hypothesied many times before.
-Certain clades of these animals could reach enormous sizes (e.g.
-Azhdarchids with wingspans of 11 metres \citep{witton2010size}) and look to have engaged in soaring flight \citep{witton2010size}.
+Certain clades of these animals could reach enormous sizes (e.g. Azhdarchids with wingspans of 11 metres \citep{witton2010size}) and look to have engaged in soaring flight \citep{witton2010size}.
 Although \cite{witton2008reappraisal} argued that neck inflexibility and straight, rather than hooked jaw morphology points against their existing as \textit{obligate} scavengers, Azhdarchid terrestrial proficency indicates they would have been comfortable foraging on the ground.
@@
 And as we have noted the Azhdarchid pterosaurs were far bigger again, with estimated body masses of over 200 kg \citep{witton2010size}.
-%CITE
 
@@
 This is significant because hyenas are likely to have left far less flesh on a carcass than a felid such as a sabretooth enabling contemperaneous hominins to benefit \citep{pobiner2015new}.
+
 %TG: or "could"? This is hypothetical behaviour stuff! Maybe past hyenas where exclusive sushi eaters!
@@
 Much work has focused on the existence of the behaviour in \textit{Tyrannosaurus rex} \citep{ruxton2003could,carbone2011intra} but a recent energetics study investigated the likely prevelance of scavenging across a range of body sizes.
-In it the authors demonstrated that species of intermediate body masses (approx.
-500 kg) would have gained the most benefit from scavenging.
+In it the authors demonstrated that species of intermediate body masses (approximatively 500 kg) would have gained the most benefit from scavenging.
 This was the result of gut capacity limitations and the effects of competition at the carcass.
@@
 As we discussed for the case of Cenozoic carnivores, osteophagy could be extremely beneficial to a scavenger.
-In Mesozoic systems some extremely large theropod dinosaurs had a morphology which suggests an ability to process bone e.g.
-the robust skull and dentition of \textit{T.
-rex} \citep{hone2010feeding}.
+
+In Mesozoic systems some extremely large theropod dinosaurs had a morphology which suggests an ability to process bone e.g. the robust skull and dentition of \textit{T. rex} \citep{hone2010feeding}.
 There is direct evidence that \textit{T.
@@
 As well as a likely instance of scavenging between a 4-million-year-old white shark (\textit{Carcharodon}) and mysticete whale from Peru \citep{ehret2009caught}.
+
 The existence of an obligate scavenger in a marine setting is uncertain \citep{britton1994marine,smith2003ecology,ruxton2004energetic,ruxton2005searching}.
@@
 \cite{beasley2015vertebrates} who point to a series of studies showing how microbes and invertebrates benefit at higher temperatures to the detriment of vertebrate scavengers such that ``above 20$^{\circ}$C vertebrates were able to detect and consume only 19\% of small-mammal carcasses, whereas at temperatures below 18$^{\circ}$C, vertebrates consumed 49\% of carcasses".
+
 % Add bits on climate change/dectection rate change, etc...

### 2016-03-22 natural history of scavenging.tex
new frame for the paper
aerial and terrestrial sections changed to reflect this

@@
 
-\title{A recipe for scavenging and natural history}
+\title{A Recipe for Scavenging - the natural history of a behaviour}
 \running{Scavenging in vertebrates}
@@
 The chief hurdle to scavenging is finding a sufficient quantity of food, the occurence of which is difficult to predict in space and time.
-Once found, the scavenger has to then overcome the agents of decay produced by the action of microorganisms on the carcass \citep{ruxton2014fruit}.
-The idea of scrounging from predator kills is undermined by studies showing that in the  majority of ecosystems more animals die from disease and starvation than predation \citep{benbow2015introduction}.
-Thus, any animal existing as a scavenger must maximise its detection capabilities and minimise its locomotory costs \citep{ruxton2004obligate}.
+Thus, any animal existing as a scavenger must minimise its locomotory costs and maximise its detection capabilities \citep{ruxton2004obligate}.
+Once found, the scavenger has to process the carrion and overcome the agents of decay produced by the action of microorganisms on the carcass \citep{ruxton2014fruit}. 
 The habitat must also be productive enough to sustain an animal biomass that will eventually produce carcasses.
+We can draw on the image of a scavenger moving through its environment, searching for food and trying to process it efficiently as we explore the prevalence of this behaviour through time. 
+% The idea of scrounging from predator kills is undermined by studies showing that in the  majority of ecosystems more animals die from disease and starvation than predation \citep{benbow2015introduction}.
 
@@
 \section*{Aerial Scavengers}
-Scavenging should be particularly attractive to avian predators compared to mammals.
-Solitary mammalian predators can kill prey up to the same body mass as themselves and sometimes an order of magnitude heavier (e.g. socially hunting lions \citep{owen2008predator}).
-In contrast, birds of prey tend to kill prey smaller than themselves \citep{slagsvold2007prey}.
-This is likely due to their need to kill prey that they can fly away with, as well as the risk of injury being higher (which carries a higher mortality risk) for a bird than a mammal.
-Thus for birds, scavenging means they can exploit species that would otherwise be too big for them to kill.
-
 Vultures represent the best known scavengers on Earth.
@@
 
-%TG: what do you mean by their position? Phylogenetic? Position on the "scavenging scale"?
+\subsection*{Locomotion}
+Flight is a cheaper means of locomotion than running \citep{tucker1975energetic}.
+This advantage can be extended further in larger species by engaging in soaring instead of flapping flight, which is even cheaper still (approximately twice BMR) \citep{hedenstrom1993migration}.
+The benefits this confers are clear from the information we have on the enormous foraging ranges of many vultures \citep{spiegel2013factors}.
+Avian flight originates in the Late Jurassic (163.5-145 Million years ago - Mya), conincident with the fossils of \textit{Archaeopteryx lithographica} so many of these benefits would have been realised from that point on for carnivorous birds.
+However, vertebrate flight is much older than this where pterosaurs predate bird origins by a considerable margin in the Late Triassic (235-201.3 Mya).
+Scavenging in this diverse group has been hypothesied many times before.
+Certain clades of these animals could reach enormous sizes (e.g. Azhdarchids with wingspans of 11 metres \citep{witton2010size}) and look to have engaged in soaring flight \citep{witton2010size}.
 
+The only other vertebrate group capable of powered flight are the bats where scavenging has not been recorded to our knowledge.
+
+\subsection*{Detection}
 Species capable of flight have effectively added an extra spatial dimension, i.e. the vertical component, to their sensory environment over land animals.
@@
 Among them are the Turkey Vultures (\textit{Cathartes aura}) which were able to locate 90\% of baits set out in a tropical forest \citep{houston1986olfaction}.
-This would be impossible for the visually reliant old world species.
-This point illustrates how the environment can impact search efficiency depending on the sensory system that's used.
+% This would be impossible for the visually reliant old world species.
 
-Flight is also a cheaper means of locomotion than running \citep{tucker1975energetic}.
-This advantage can be extended further in larger species by engaging in soaring instead of flapping flight, which is even cheaper still (approximately twice BMR) \citep{hedenstrom1993migration}.
-The benefits this confers are clear from the information we have on the enormous foraging ranges of many vultures \citep{spiegel2013factors}.
-Clearly, it would be pointless to have incredible detection abilites and not have a cost efficient movement to benefit from it.
-Although, as with detection ability, the environment has a role to play here.
-Vultures and eagles tend to soar using thermals and if these air pockets don't form, say on a cloudy day, the bird is grounded \citep{mundy1992vultures}.
-In many habitats (e.g. the arctic) it is simply not possible for sufficiently powerful thermals to form and as a consequence large-bodied vultures cannot exist.
-The upshot of this is that terrestrial carnivores like bears and wolves take more carrion \citep{devault2003scavenging}.
+Clearly, it would be pointless to have incredible ranging abilites and not have the sensory architecture to benefit from it.
 
-Avian flight originates in the Late Jurassic (163.5-145 Million years ago - Mya), conincident with the fossils of \textit{Archaeopteryx lithographica} so many of these benefits would have been realised from that point on for carnivorous birds.
-However, vertebrate flight is much older than this where pterosaurs predate bird origins by a considerable margin in the Late Triassic (235-201.3 Mya).
-Scavenging in this diverse group has been hypothesied many times before.
-Certain clades of these animals could reach enormous sizes (e.g. Azhdarchids with wingspans of 11 metres \citep{witton2010size}) and look to have engaged in soaring flight \citep{witton2010size}.
-Although \cite{witton2008reappraisal} argued that neck inflexibility and straight, rather than hooked jaw morphology points against their existing as \textit{obligate} scavengers, Azhdarchid terrestrial proficency indicates they would have been comfortable foraging on the ground.
-Indeed, extant Marabou Storks (\textit{Leptoptilos crumenifer}) %TG: Check if that's the one
-have a comparable morphology and are noted facultative scavengers so it is reasonable to believe that certain pterosaurs behaved similarly.
+\subsection*{Processing}
+Scavenging should be particularly attractive to avian predators compared to mammals.
+Solitary mammalian predators can kill prey up to the same body mass as themselves and sometimes an order of magnitude heavier (e.g. socially hunting lions \citep{owen2008predator}).
+In contrast, birds of prey tend to kill prey smaller than themselves \citep{slagsvold2007prey}.
+This is likely due to their need to kill prey that they can fly away with, as well as the risk of injury being higher (which carries a higher mortality risk) for a bird than a mammal.
+Thus for birds, scavenging means they can exploit species that would otherwise be too big for them to kill.
 
-Large body size confers substantial dominance and starvation resistance benefits \citep{ruxton2004obligate}.
+Large body size confers substantial dominance and starvation-resistance benefits \citep{ruxton2004obligate}.
 Thus, we would expect scavengers to have this trait selected for even in the case of weight-constrained fliers.
@@
 
-The only other vertebrate group capable of powered flight are the bats where scavenging has not been recorded to our knowledge.
-Their visual acuity is famously poor and echolocation would not lend itself to discovering immobile carrion.
+Although \cite{witton2008reappraisal} argued that neck inflexibility and straight, rather than hooked jaw morphology points against pterosaurs existing as \textit{obligate} scavengers, Azhdarchid terrestrial proficency indicates they would have been comfortable foraging on the ground.
+Indeed, extant Marabou Storks (\textit{Leptoptilos crumenifer}) %TG: Check if that's the one
+have a comparable morphology and are noted facultative scavengers so it is reasonable to believe that certain pterosaurs behaved similarly.
+
+Returning to bats, their visual acuity is famously poor and echolocation would not lend itself to discovering immobile carrion.
 Their small size and poor terrestrial ability would also count against them at a carcass \citep{riskin2006terrestrial}.
@@
 %TG: Two points here: one speculation for the non-scavenging behaviour could be because of diet.
-Big bats (which are better suited for scavenging, following previous paragraph) are essentially frugivores and therefore probably lack of adaptations for digesting the meat.
-Also carnivores bats are mainly the microbats that are insectivores.
-Maybe they scavenge on dead insects but here we're talking about scavenging on real meat right?
+Big bats (which are better suited for scavenging, following our previous argument) are typically frugivores and therefore probably lack of adaptations for digesting the meat.
+Also carnivorous bats are mainly the microbats which are insectivorous.
+% Maybe they scavenge on dead insects but here we're talking about scavenging on real meat right?
 %TG: Second point: bats fossil record is of course not so good but it's actually way better than birds!
@@
 
+\subsection*{Environment}
+Vultures and eagles tend to soar using thermals and if these air pockets don't form, say on a cloudy day, the bird is grounded \citep{mundy1992vultures}.
+In many habitats (e.g. the arctic) it is simply not possible for sufficiently powerful thermals to form and as a consequence large-bodied vultures cannot exist.
+The upshot of this is that terrestrial carnivores like bears and wolves take more carrion \citep{devault2003scavenging}.
+The use of different sensory systems also illustrates the impact of the environment. 
+The relatively open savanna systems of Africa are well suited to a visually dependent vulture whereas more forested areas would select for species that have a well developed oflfactory system  \citep{houston1986olfaction}.
+
+% Although, as with detection ability, the environment has a role to play here.
+
+%TG: what do you mean by their position? Phylogenetic? Position on the "scavenging scale"?
+
+% This point illustrates how the environment can impact search efficiency depending on the sensory system that's used.
+
 \section*{Terrestrial Scavengers}
-A simplification of terrestrial, vertebrate scavengers is one of them existing in a two-dimensional plane while foraging for carrion directly.
-They can detect carcasses at a range that is defined by the radius of their sensory organs, usually the visual and olfactory senses.
-As a consequence, they have a much more restricted view of the landscape than do aerial foragers.
+Terrestrial scavenging in the mammals is probably best known in an African context where hyenas, jackals and lions all take sizable proportions of carrion in their diet.
+In the spotted hyena (\textit{Crocuta crocuta}), striped hyena (\textit{Hyaena hyaena}) and brown hyena (\textit{Hyaena brunnea}) it can be over 90\% \citep{jones2015african}.
+Therefore, we can again use these species as our efficient terrestrial scavengers to compare with other forms.
+
 No contemporary terrestrial vertebrate exists as an obligate scavenger but most if not all are facultative scavengers to some extent \citep{beasley2015vertebrates}.
@@
 
-Terrestrial scavenging in the mammals is probably best known in an African context where hyenas, jackals and lions all take sizable proportions of carrion in their diet.
-In the spotted hyena (\textit{Crocuta crocuta}), striped hyena (\textit{Hyaena hyaena}) and brown hyena (\textit{Hyaena brunnea}) it can be over 90\% \citep{jones2015african}.
-Therefore, we can again use these species as our efficient terrestrial scavengers to compare with other forms.
+\subsection*{Locomotion}
+As endotherms, mammals can sustain long bouts of energetically expensive activity. 
+By contrast, modern reptiles are ectothermic, limiting their activity periods.
+% a successful reptilian scavenger requires a far different set of adapations.
+This is exacerbated by the sprawling gait seen in many lizards which results in Carrier's Constraint such that the animal can't move and breathe at the same time because the lateral movements impedes its lungs \citep{carrier1987evolution}.
+This manifests itself in aspects such as maximum sustainable speed where an equivalent mammal has a six to seven fold increase \citep{ruben1995evolution}.
+To quantify this effect with a simple example we can turn to some allometric relationships relating sustainable travelling speed to body mass.
+In the case of mammals and reptiles these are 1.15 * body mass (kg) \textsuperscript{0.12} and 0.23 * body mass (kg) \textsuperscript{0.12}
+% respectively \citep{ruxton2004obligate}.
+We can insert these into a foraging radius model ((duration * speed)/2)/1000 for a 12 hour foraging day which shows that while a 10 kg reptile can range 6.5 km an equally sized mammal can range nearly 33 km \citep{Enstipp2006Energetics}.
+Hyenas extend this advantage with their characteristic ``rocking horse gait"  which allows them to cover great distances efficiently \citep{jones2015african}.
 
-Similar to vultures they have well developed sensory organs, whereby they can smell a rotting carcass 4 km away and can hear the vocalisations of conspecifics at a distance of 10 km \citep{mills1989comparative}.
-They have a characteristic ``rocking horse gait"  which allows them to cover great distances efficiently \citep{jones2015african}.
-A major difficulty for terrestrial scavengers is competition with vultures.
-Noctural behaviour in the Hyaenidae in general has been put forth as an adaptation to reduce competition with these exclusively diurnal birds \citep{gittleman2013carnivore}.
+Unsurprisingly, given their enduring appeal, the prevalence of scavenging has been explored in the carnivorous, theropod dinosaurs.
+These animals ranged from the chicken-sized to the whale-sized all of which were bipedal.
+They are quite alien to anything we know today which restricts our ability to understand their ecology far more so than extinct mammals \citep{weishampel2004dinosauria}.
+Of relevance, are the questions that still persist about their metabolism \citep{grady2014evidence}. 
+We do know that they walked with the erect gait of mammals or birds rather than the sprawling gait of lizards and that they were most likely facultative scavengers \citep{weishampel2004dinosauria,depalma2013physical}.
+
+Of course, tetrapod terrestrial dominance predates the evolution of the dinosaurs.
+It is in the Permian, almost 300 millions years ago, that we have the earliest evidence of vertebrate scavenging where a temnospondyl amphibian fed on the carcass of \textit{Varanops}, a predatory synapsid of the time \citep{reisz2006articulated}.
+
+\subsection*{Detection}
+A simplification of terrestrial, vertebrate scavengers in sensory terms is one of them existing in a two-dimensional plane while foraging for carrion directly.
+They can detect carcasses at a range that is defined by the radius of their sensory organs, usually the visual and olfactory senses.
+As a consequence, they have a much more restricted view of the landscape than do aerial foragers.
+Similar to vultures hyenas have well developed sensory organs, whereby they can smell a rotting carcass 4 km away and can hear the vocalisations of conspecifics at a distance of 10 km \citep{mills1989comparative}.
+Using the same approach \cite{spiegel2013factors} used to determine the visual acuity of vultures we found a Spotted hyena could resolve a 2 metre target at 1 km distance.
+
+\subsection*{Proceesing}
 The bone crushing ability of hyenas reveals another useful scavenger trait.
@@
 As a result, the decline of Machairodontinae sabretooths has been offered as an explanation for the extinction of \textit{P brevirostris} \citep{palmqvist2011giant}.
-
 The order Carnivora sees its origins in the Middle Eocene (56-33.9 Mya) where it split into the Caniforma and Feliforma.
@@
 
-By contrast to mammals, a successful reptilian scavenger requires a far different set of adapations.
-Modern forms are ectothermic, limiting their activity periods.
-This is exacerbated by the sprawling gait seen in many lizards which results in Carrier's Constraint such that the animal can't move and breathe at the same time because the lateral movements impedes its lungs \citep{carrier1987evolution}.
-This manifests itself in aspects such as maximum sustainable speed where an equivalent mammal has a six to seven fold increase \citep{ruben1995evolution}.
-A lower metabolism does give reptiles an advantage however, in that over the course of a year their food requirements can be 30 times smaller than an endotherm of equal size \citep{Nagy1621}.
+When it comes to reptiles they possees an advantage here, in that over the course of a year their food requirements can be 30 times smaller than an endotherm of equal size \citep{Nagy1621}.
 Any adaptations that reduce energetic costs are likely to be selected in scavengers.
@@
 
-Unsurprisingly, given their enduring appeal, the prevalence of scavenging has been explored in the carnivorous, theropod dinosaurs.
-These animals ranged from the chicken-sized to the whale-sized all of which were bipedal.
-They are quite alien to anything we know today which restricts our ability to understand their ecology far more so than extinct mammals \citep{weishampel2004dinosauria}.
-Of relevance, are the questions that still persist about their metabolism \citep{grady2014evidence} and sensory perception \citep{farlow1994speculations}.
-We do know that they walked with the erect gait of mammals or birds rather than the sprawling gait of lizards and that they were most likely facultative scavengers \citep{weishampel2004dinosauria,depalma2013physical}.
-Much work has focused on the existence of the behaviour in \textit{Tyrannosaurus rex} \citep{ruxton2003could,carbone2011intra} but a recent energetics study investigated the likely prevelance of scavenging across a range of body sizes.
+Much work has focused on the existence of scavenging in dinosaurs by using simple energetics approaches that typically focused on a single species namely \textit{Tyrannosaurus rex} \citep{ruxton2003could,carbone2011intra} but a recent modelling study investigated the likely prevelance of scavenging across a range of body sizes.
 In it the authors demonstrated that species of intermediate body masses (approximatively 500 kg) would have gained the most benefit from scavenging.
@@
 For the smaller species, larger competitors would have prevented their access to carrion.
-
-As we discussed for the case of Cenozoic carnivores, osteophagy could be extremely beneficial to a scavenger.
 
@@
 This is noted as being powerful enough to break open skeletal material \citep{rayfield2001cranial}.
-Osteophagy may have been even more viable during this era because the body mass distribution of herbviores tended to be skewed towards larger sizes \citep{10.1371/journal.pone.0051925}.
+
+\subsection*{Environment}
+A major difficulty for terrestrial scavengers is competition with vultures.
+Noctural behaviour in the Hyaenidae in general has been put forth as an adaptation to reduce competition with these exclusively diurnal birds \citep{gittleman2013carnivore}.
+
+Osteophagy may have been even more viable during the Mesozoic era because the body mass distribution of herbviores tended to be skewed towards larger sizes \citep{10.1371/journal.pone.0051925}.
 When we couple this with the fact that skeletal mass scales greater than linearly with body mass \citep{prange1979scaling} there would have been a lot of bone material to consume in the environment provided an animal had the biology to process it \citep{chure1997one}.
+As we discussed earlier, osteophagy could be extremely beneficial to a scavenger.
 
-Of course, tetrapod terrestrial dominance predates the evolution of the dinosaurs.
-It is in the Permian, almost 300 millions years ago, that we have the earliest evidence of vertebrate scavenging where a temnospondyl amphibian fed on the carcass of \textit{Varanops}, a predatory synapsid of the time \citep{reisz2006articulated}.
 The absence of flying vertebrates in the Palaeozoic may have permitted terrestrial forms to take in a higher proportion of carrion in their diet.
@@
 Scavenging behaviour might have evolved as soon as the first terrestrial tetrapods evolved.
-In fact, some of the earlier tetrapods tracks dating back to the early Middle Devonian (393.3 - 387.7 Mya) where found in intertidal environments \citep{Niedzwiedzki2009}.
-These environments are isolated from marine systems twice a day leaving potential carrions unexploited by marine vertebrates.
+In fact, some of the earlier tetrapods tracks dating back to the early Middle Devonian (393.3 - 387.7 Mya) were found in intertidal environments \citep{Niedzwiedzki2009}.
+These environments are isolated from marine systems twice a day leaving potential carrion unexploited by marine vertebrates.
 \cite{Niedzwiedzki2009} suggests that these environments ``would thus have allowed marine ancestors of tetrapods gradually to acquire terrestrial competence while accessing a new and essentially untouched resource.''
 %TG: orginal sentence "The intertidal environment provides a ready food source of stranded marine animals on a twice-daily basis, in the immediate vicinity of the sea, and would thus have allowed marine ancestors of tetrapods gradually to acquire terrestrial competence while accessing a new and essentially untouched resource."
+
+% and sensory perception \citep{farlow1994speculations}.
 

### 2016-03-23 natural history of scavenging.tex
new headings

@@
 By recognising the difficulty in directly observing scavenging, a suite of methods have been used to discern the most suitable morphologies, physiologies and environments for a scavenging lifestyle to prosper.
-Here we chart the natural history of vertebrate scavenging on air, land and in water by focussing on the dominant vertebrate groups that existed at each time.
+Here we chart the natural history of scavenging by looking at the potential for the behaviour in dominant vertebrate groups.
 %It is our aim in this review to employ these methods to gain an understanding of scavengers past and present.
@@
 
-\section*{Aerial Scavengers}
-Vultures represent the best known scavengers on Earth.
-These birds consist of two convergent groups, from the old and the new world and represent the only example of obligate vertebrate scavengers today.
-The families from which modern vultures arose, the Accipitridae and Cathartidae, appear during the Palaeocene (66 - 56 Mya) \citep{Jetz2012, Jarvis2014}.
-Given their unique position, they have been extensively studied to determine what adaptations they possess that allows them to so flourish in this niche.
-Underscoring their prowess are studies such as \cite{ogada2012effects} who note a three fold increase in carcass persistence times in the absence of vultures.
-As such, we can begin by exploring the adapdations and the environments of vultures to draw comparisons with other scavenging species and \textit{their} environments.
+% \section*{Aerial Scavengers}
+% Vultures represent 
+%These birds consist of two convergent groups, from the old and the new world and represent the only example of obligate vertebrate scavengers today.
+%The families from which modern vultures arose, the Accipitridae and Cathartidae, appear during the Palaeocene (66 - 56 Mya) \citep{Jetz2012, Jarvis2014}.
+%Given their unique position, they have been extensively studied to determine what adaptations they possess that allows them to so flourish in this niche.
+%Underscoring their prowess are studies such as \cite{ogada2012effects} who note a three fold increase in carcass persistence times in the absence of vultures.
+%As such, we can begin by exploring the adapdations and the environments of vultures to draw comparisons with other scavenging species and \textit{their} environments.
 
 \subsection*{Locomotion}
-Flight is a cheaper means of locomotion than running \citep{tucker1975energetic}.
-This advantage can be extended further in larger species by engaging in soaring instead of flapping flight, which is even cheaper still (approximately twice BMR) \citep{hedenstrom1993migration}.
-The benefits this confers are clear from the information we have on the enormous foraging ranges of many vultures \citep{spiegel2013factors}.
-Avian flight originates in the Late Jurassic (163.5-145 Million years ago - Mya), conincident with the fossils of \textit{Archaeopteryx lithographica} so many of these benefits would have been realised from that point on for carnivorous birds.
-However, vertebrate flight is much older than this where pterosaurs predate bird origins by a considerable margin in the Late Triassic (235-201.3 Mya).
-Scavenging in this diverse group has been hypothesied many times before.
-Certain clades of these animals could reach enormous sizes (e.g. Azhdarchids with wingspans of 11 metres \citep{witton2010size}) and look to have engaged in soaring flight \citep{witton2010size}.
+As endotherms, mammals can sustain long bouts of energetically expensive activity. 
+By contrast, modern reptiles are ectothermic, limiting their activity periods.
+% a successful reptilian scavenger requires a far different set of adapations.
+This is exacerbated by the sprawling gait seen in many lizards which results in Carrier's Constraint such that the animal can't move and breathe at the same time because the lateral movements impedes its lungs \citep{carrier1987evolution}.
+An effect which manifests itself in aspects such as maximum sustainable speed where an equivalent mammal has a six to seven fold increase \citep{ruben1995evolution}.
+To quantify this effect with a simple example we can turn to some allometric relationships relating sustainable travelling speed to body mass.
+In the case of mammals and reptiles these are 1.15 * body mass (kg) \textsuperscript{0.12} and 0.23 * body mass (kg) \textsuperscript{0.12} respectively \citep{ruxton2004obligate}.
+We can insert these into a foraging radius model ((duration * speed)/2)/1000 for a 12 hour foraging day which shows that while a 10 kg reptile can range 6.5 km an equally sized mammal can range nearly 33 km \citep{Enstipp2006Energetics}.
+
+Today, terrestrial scavenging in the mammals is probably best known in an African context where hyenas, jackals and lions all take sizable proportions of carrion in their diet.
+In the spotted hyena (\textit{Crocuta crocuta}), striped hyena (\textit{Hyaena hyaena}) and brown hyena (\textit{Hyaena brunnea}) it can be over 90\% \citep{jones2015african}.
+Yet, no contemporary terrestrial vertebrate exists as an obligate scavenger but most if not all are facultative to some extent \citep{beasley2015vertebrates}.
+Therefore, we can use hyenas as examples of efficient terrestrial scavengers to compare with other forms. 
+In terms of locomotion, they employ a characteristic ``rocking horse gait"  which allows them to cover great distances efficiently, loping at 10 km/hr \citep{mills1989comparative,jones2015african}. 
+The order Carnivora saw its origins in the Middle Eocene (56-33.9 Mya) where it split into the Caniforma and Feliforma.
+So we can trace efficient terrestrial movement by carnivores from this point on.
+Certainly, it is profitable to compare extant terrestrial species to their prehistoric forebears given the dominance of mammalian carnviores since the Eocene.
+
+Unsurprisingly, given their enduring appeal, the prevalence of scavenging has been explored in the carnivorous, theropod dinosaurs.
+Dinosaurs in general, were the dominant terrestrial forms for most of the Mesozoic Era. 
+where the theropods ranged from the chicken-sized to the whale-sized, all of which were bipedal.
+They are quite alien to anything we know today which restricts our ability to understand their ecology far more so than extinct mammals \citep{weishampel2004dinosauria}.
+Of relevance, are the questions that still persist about their metabolism with the latest evidence suggesting they were mesothermic i.e.
+intermediate to ecto- and endotherms \citep{grady2014evidence}.
+We do know that they walked with the erect gait of mammals or birds rather than the sprawling gait of lizards and that they were most likely facultative scavengers \citep{weishampel2004dinosauria,depalma2013physical}.
+Taken together, this implies dinosaurs had a foraging range that fell in between that of modern terrestrial mammals and reptiles.
+
+Of course, tetrapod terrestrial dominance predates the evolution of the dinosaurs.
+It is in the Permian, almost 300 millions years ago, that we have the earliest evidence of vertebrate scavenging where a temnospondyl amphibian fed on the carcass of \textit{Varanops}, a predatory synapsid of the time \citep{reisz2006articulated}.
+And it is with the evolution of endothermy in the therapsid-mammal lineage \citep{clarke2010temperature} that terrestrial vertebrates would have had the ability to range widely, a vital component in seeking out carrion.
+
+Scavenging behaviour might have evolved on land as soon as the first terrestrial tetrapods evolved.
+In fact, some of the earlier tetrapods tracks dating back to the early Middle Devonian (393.3 - 387.7 Mya) were found in intertidal environments \citep{Niedzwiedzki2009}.
+These environments are isolated from marine systems twice a day leaving potential carrion unexploited by marine vertebrates.
+\cite{Niedzwiedzki2009} suggests that these environments ``would thus have allowed marine ancestors of tetrapods gradually to acquire terrestrial competence while accessing a new and essentially untouched resource.''
+
+Flight is a cheaper means of locomotion than walking or running \citep{tucker1975energetic}. 
+Many extant birds exist as facultative scavengers; storks, eagles, corvids, are all known to take substantial quantities of carrion in their diet \citep{kendall2013alternative}.
+The advantage of flight can be extended further in larger species that engage in soaring instead of flapping flight, which is even cheaper still (approximately twice the basal metabolic rate) \citep{hedenstrom1993migration,spivey2014analysing}.
+The benefits this confers are clear from the information we have on the enormous foraging ranges of many vultures \citep{spiegel2013factors} and seabirds \citep{thaxter2012seabird}.
+In the former case we have the best known scavengers on Earth. 
+Vultures consist of two convergent groups, from the old and the new world and represent the only example of obligate vertebrate scavengers today.
+The families from which modern vultures arose, the Accipitridae and Cathartidae, appear during the Palaeocene (66 - 56 Mya) \citep{Jetz2012, Jarvis2014}.
+Avian flight is far older than this and originates in the Late Jurassic (163.5-145 Million years ago - Mya), conincident with the fossils of \textit{Archaeopteryx lithographica} so many of these benefits would have been realised from that point on for carnivorous birds.
+However, vertebrate flight is much older still where pterosaurs predate bird origins by a considerable margin in the Late Triassic (235-201.3 Mya).
+Scavenging in this diverse group has been hypothesied many times before \citep{witton2008reappraisal}.
+Certain clades of these animals could reach enormous sizes (e.g.
+Azhdarchids with wingspans of 11 metres \citep{witton2010size}) and, notably, look to have engaged in soaring flight \citep{witton2010size}.
 
 The only other vertebrate group capable of powered flight are the bats where scavenging has not been recorded to our knowledge.
+The bat fossil record is notoriously poor owing to their fragile skeletons so we are unable to determine if extinct species were more suited to this lifestyle \citep{eiting2009global}.
+Although it does not seem that flight is the main criterion precluding them from scavenging (see below).
+
+Aquatic scavengers have a locomotory benefit because water is a medium that is conducive to low-cost movement \citep{tucker1975energetic}. 
+In fact, the cost of swimming is lower than either running or flying \citep{williams1999evolution}. 
+This has led some researchers to argue for the feasability of a scavenging fish \citep{ruxton2004energetic,ruxton2005searching}. 
+As with the aerial and terrestrial enviornments we have evidence of facultative scavenging among extant and extinct aquatic species.
+For example, the remains of a mosasaur and a terrestrial hadrosaur were discovered with embedded teeth from a Cretaceous shark, \textit{Squalicorax} \citep{schwimmer1997scavenging}.
+As well as a likely instance of scavenging between a 4-million-year-old white shark (\textit{Carcharodon}) and mysticete whale from Peru \citep{ehret2009caught}.
+We might expect then that by combining an aquatic environment and an endothermic metabolism that marine mammals would prosper as scavengers. 
+We know fossil pinnipeds and cetaceans from 60 Mya have transitional features indicative of their trajectory to fully aquatic species \citep{williams1999evolution}.  
+But despite this movement away from land the energetic savings were negligible because the \textit{total} cost incurred by a swimming marine mammal is high \citep{williams1999evolution}. 
+As \cite{williams1999evolution} notes, ``Free-ranging animals must contend with the total energetic expenditure associated with supporting basic biological functions as well as with moving the body and appendages through the environment."
 
 \subsection*{Detection}
+A simplification of terrestrial, vertebrate scavengers in sensory terms is one of them existing in a two-dimensional plane while foraging for carrion directly.
+They can detect carcasses at a range that is defined by the radius of their sensory organs. %, usually the visual and olfactory senses.
+As a consequence, they have a much more restricted view of the landscape than do aerial foragers.
+Hyenas have well developed sensory organs, whereby they can smell a rotting carcass 4 km away and can hear the vocalisations of conspecifics at a distance of 10 km \citep{mills1989comparative}. 
+Using the approach of \cite{spiegel2013factors} we estimate a spotted hyena could resolve a 2 metre target at 1 km distance. 
+\cite{ruxton2004obligate} calculated that ``a 1 tonne mammal or reptile, in an ecosystem yielding carrion at densities similar to the current Serengeti, could have met its energy requirements if it could detect carrion over a distance of the order of 400–500 m.".
+
 Species capable of flight have effectively added an extra spatial dimension, i.e. the vertical component, to their sensory environment over land animals.
@@
 Vultures are known to have impressive visual acuity with one estimate indicating Lappet-faced Vultures (\textit{Torgos tracheliotus}) are capable of detecting a 2 metre carcass over 10 km away \citep{spiegel2013factors}.
-We know that many birds exist as facultative scavengers; storks, eagles, corvids, are all known to take substantial quantities of carrion in their diet.
-And eagles in particular are known to have highly developed visual abilities.
+Eagles too are known to have highly developed visual abilities \citep{reymond1985spatial}.
 It follows from this that the evolution of flight allowed aerial animals to detect far more carrion than their terrestrial counterparts \citep{AR:AR22815}.
@@
 Among them are the Turkey Vultures (\textit{Cathartes aura}) which were able to locate 90\% of baits set out in a tropical forest \citep{houston1986olfaction}.
+An atuned sense of smell is obviously useful in detecting decaying carrion.
+Clearly, it would be pointless to have incredible ranging abilites and not have the sensory architecture to benefit from it.
 % This would be impossible for the visually reliant old world species.
 
-Clearly, it would be pointless to have incredible ranging abilites and not have the sensory architecture to benefit from it.
+Returning to bats, their visual acuity is famously poor, and echolocation would not lend itself to discovering immobile carrion.
+Their small size and poor terrestrial ability would also count against them at a carcass \citep{riskin2006terrestrial}.
+
+Depending on the species, a carcass in water either floats or descends to the sea floor \citep{Whitehead415}.
+In the latter low-light environment, visual detection distances are far lower (< 100 m) than they would be in the air.
+As such, animals detect resources through chemo- and mechanoreception more so than through vision \citep{ruxton2004energetic}.
+Extant aquatic snakes are deemed as having the most suitable physiology for scavenging.
+A hypothesis put forth by \cite{sazima1990necrofagia} argued that chemical gradients in water would allow for a relatively easier detection of carrion.
+This gained some support from \cite{devault2002scavenging}, who found a preponderence of aquatic snake species in their review of this behaviour.
 
 \subsection*{Processing}
-Scavenging should be particularly attractive to avian predators compared to mammals.
-Solitary mammalian predators can kill prey up to the same body mass as themselves and sometimes an order of magnitude heavier (e.g. socially hunting lions \citep{owen2008predator}).
-In contrast, birds of prey tend to kill prey smaller than themselves \citep{slagsvold2007prey}.
-This is likely due to their need to kill prey that they can fly away with, as well as the risk of injury being higher (which carries a higher mortality risk) for a bird than a mammal.
-Thus for birds, scavenging means they can exploit species that would otherwise be too big for them to kill.
-
-Large body size confers substantial dominance and starvation-resistance benefits \citep{ruxton2004obligate}.
-Thus, we would expect scavengers to have this trait selected for even in the case of weight-constrained fliers.
-Cinereous Vultures (\textit{Aegypius monachus}) and condors (\textit{Vultur gryphus}, \textit{Gymnogyps californianus}) all have body masses that can exceed 10 kg and represent some of the heaviest bird species capable of flight \citep{ferguson2001raptors,donazar2002effects}.
-And as we have noted the Azhdarchid pterosaurs were far bigger again, with estimated body masses of over 200 kg \citep{witton2010size}.
-
-Although \cite{witton2008reappraisal} argued that neck inflexibility and straight, rather than hooked jaw morphology points against pterosaurs existing as \textit{obligate} scavengers, Azhdarchid terrestrial proficency indicates they would have been comfortable foraging on the ground.
-Indeed, extant Marabou Storks (\textit{Leptoptilos crumenifer}) %TG: Check if that's the one
-have a comparable morphology and are noted facultative scavengers so it is reasonable to believe that certain pterosaurs behaved similarly.
-
-Returning to bats, their visual acuity is famously poor and echolocation would not lend itself to discovering immobile carrion.
-Their small size and poor terrestrial ability would also count against them at a carcass \citep{riskin2006terrestrial}.
-The bat fossil record is notoriously poor owing to their fragile skeletons so we are unable to determine if extinct species were more suited to this lifestyle \citep{eiting2009global}.
-%TG: Two points here: one speculation for the non-scavenging behaviour could be because of diet.
-Big bats (which are better suited for scavenging, following our previous argument) are typically frugivores and therefore probably lack of adaptations for digesting the meat.
-Also carnivorous bats are mainly the microbats which are insectivorous.
-% Maybe they scavenge on dead insects but here we're talking about scavenging on real meat right?
-%TG: Second point: bats fossil record is of course not so good but it's actually way better than birds!
-Most of the data we have is on teeth of course but it's probably enough for having a rough body mass and diet estimate!
-And that's much better than birds where you're stuck with just some fragments of long bones.
-
-\subsection*{Environment}
-Vultures and eagles tend to soar using thermals and if these air pockets don't form, say on a cloudy day, the bird is grounded \citep{mundy1992vultures}.
-In many habitats (e.g. the arctic) it is simply not possible for sufficiently powerful thermals to form and as a consequence large-bodied vultures cannot exist.
-The upshot of this is that terrestrial carnivores like bears and wolves take more carrion \citep{devault2003scavenging}.
-The use of different sensory systems also illustrates the impact of the environment. 
-The relatively open savanna systems of Africa are well suited to a visually dependent vulture whereas more forested areas would select for species that have a well developed oflfactory system  \citep{houston1986olfaction}.
-
-% Although, as with detection ability, the environment has a role to play here.
-
-%TG: what do you mean by their position? Phylogenetic? Position on the "scavenging scale"?
-
-% This point illustrates how the environment can impact search efficiency depending on the sensory system that's used.
-
-\section*{Terrestrial Scavengers}
-Terrestrial scavenging in the mammals is probably best known in an African context where hyenas, jackals and lions all take sizable proportions of carrion in their diet.
-In the spotted hyena (\textit{Crocuta crocuta}), striped hyena (\textit{Hyaena hyaena}) and brown hyena (\textit{Hyaena brunnea}) it can be over 90\% \citep{jones2015african}.
-Therefore, we can again use these species as our efficient terrestrial scavengers to compare with other forms.
-
-No contemporary terrestrial vertebrate exists as an obligate scavenger but most if not all are facultative scavengers to some extent \citep{beasley2015vertebrates}.
-\cite{ruxton2004obligate} offer a reason for this in that the traits that allow for vultures to exist as scavengers undermined their ability to hunt but that the same forces have not prevented mammals from doing so.
-Interestingly, the same authors calculated that ``a 1 tonne mammal or reptile, in an ecosystem yielding carrion at densities similar to the current Serengeti, could have met its energy requirements if it could detect carrion over a distance of the order of 400–500 m."\citep{ruxton2004obligate}.
-
-\subsection*{Locomotion}
-As endotherms, mammals can sustain long bouts of energetically expensive activity. 
-By contrast, modern reptiles are ectothermic, limiting their activity periods.
-% a successful reptilian scavenger requires a far different set of adapations.
-This is exacerbated by the sprawling gait seen in many lizards which results in Carrier's Constraint such that the animal can't move and breathe at the same time because the lateral movements impedes its lungs \citep{carrier1987evolution}.
-This manifests itself in aspects such as maximum sustainable speed where an equivalent mammal has a six to seven fold increase \citep{ruben1995evolution}.
-To quantify this effect with a simple example we can turn to some allometric relationships relating sustainable travelling speed to body mass.
-In the case of mammals and reptiles these are 1.15 * body mass (kg) \textsuperscript{0.12} and 0.23 * body mass (kg) \textsuperscript{0.12}
-% respectively \citep{ruxton2004obligate}.
-We can insert these into a foraging radius model ((duration * speed)/2)/1000 for a 12 hour foraging day which shows that while a 10 kg reptile can range 6.5 km an equally sized mammal can range nearly 33 km \citep{Enstipp2006Energetics}.
-Hyenas extend this advantage with their characteristic ``rocking horse gait"  which allows them to cover great distances efficiently \citep{jones2015african}.
-
-Unsurprisingly, given their enduring appeal, the prevalence of scavenging has been explored in the carnivorous, theropod dinosaurs.
-These animals ranged from the chicken-sized to the whale-sized all of which were bipedal.
-They are quite alien to anything we know today which restricts our ability to understand their ecology far more so than extinct mammals \citep{weishampel2004dinosauria}.
-Of relevance, are the questions that still persist about their metabolism \citep{grady2014evidence}. 
-We do know that they walked with the erect gait of mammals or birds rather than the sprawling gait of lizards and that they were most likely facultative scavengers \citep{weishampel2004dinosauria,depalma2013physical}.
-
-Of course, tetrapod terrestrial dominance predates the evolution of the dinosaurs.
-It is in the Permian, almost 300 millions years ago, that we have the earliest evidence of vertebrate scavenging where a temnospondyl amphibian fed on the carcass of \textit{Varanops}, a predatory synapsid of the time \citep{reisz2006articulated}.
-
-\subsection*{Detection}
-A simplification of terrestrial, vertebrate scavengers in sensory terms is one of them existing in a two-dimensional plane while foraging for carrion directly.
-They can detect carcasses at a range that is defined by the radius of their sensory organs, usually the visual and olfactory senses.
-As a consequence, they have a much more restricted view of the landscape than do aerial foragers.
-Similar to vultures hyenas have well developed sensory organs, whereby they can smell a rotting carcass 4 km away and can hear the vocalisations of conspecifics at a distance of 10 km \citep{mills1989comparative}.
-Using the same approach \cite{spiegel2013factors} used to determine the visual acuity of vultures we found a Spotted hyena could resolve a 2 metre target at 1 km distance.
-
-\subsection*{Proceesing}
-The bone crushing ability of hyenas reveals another useful scavenger trait.
 Since carrion is not dispatched directly, often the most easily accessible and choicest components of the carcass will be missing or, if present, will be fought over.
 Being able to extract nutrients from remnants gives the scavenger a great advantage.
+Thus, the bone crushing ability of hyenas reveals another useful scavenger trait.
 Osteophagy is known across a range of terrestrial carnivores and given some fat-rich mammalian bones have an energy density (6.7 kJ/g) comparable with that of muscle tissue, it makes skeletal remains an enticing resource \citep{brown1989study}.
@@
 As a result, the decline of Machairodontinae sabretooths has been offered as an explanation for the extinction of \textit{P brevirostris} \citep{palmqvist2011giant}.
-The order Carnivora sees its origins in the Middle Eocene (56-33.9 Mya) where it split into the Caniforma and Feliforma.
-And many of the aforesaid adapations for scavenging are found in these other major terrestrial mammalian carnviores.
-Though the specific mix of features realised in hyenas suggest this is the model organism for terrestrial scavenging among mammals in the past.
-For example, the bone-crushing dogs that evolved during the Oligocene (subfamily Borophaginae) have been compared to hyenas in terms of their feeding ecology \citep{van2003chapter,martin2016pursuit}.
-Certainly, it is profitable to compare extant terrestrial species to their prehistoric forebears given the dominance of mammalian carnviores since the Eocene.
+% And many of the aforesaid adapations for scavenging are found in these other major terrestrial mammalian carnviores.
+% Though the specific mix of features realised in hyenas suggest this is the model organism for terrestrial scavenging among mammals in the past.
+% These traits have been used to infer scavenging lifestyles in extinct mammals. 
+The bone-crushing dogs that evolved during the Oligocene (subfamily Borophaginae) have been compared to hyenas in terms of their feeding ecology as well \citep{van2003chapter,martin2016pursuit}.
 
-Interestingly such comparisons have given insight into the feeding ecology of early hominins who, for instance, had the ability to craft tools for breaking open bones \citep{ARCM:ARCM12084}.
+Interestingly, such comparisons have given insight into the feeding ecology of early hominins who, for instance, had the ability to craft tools for breaking open bones \citep{ARCM:ARCM12084}.
 The question of where our ancestors placed on the hunter-scavenger axis during the Plio-Pleistocene has been a matter of debate for years.
@@
 The same reasoning can be applied to crocodlies and their allies \citep{forrest2003evidence}. 
+A sit and wait strategy is viable for an ectotherm.
 
@@
 
+Scavenging should be particularly attractive to avian predators compared to mammals.
+Solitary mammalian predators can kill prey up to the same body mass as themselves and sometimes an order of magnitude heavier (e.g. socially hunting lions \citep{owen2008predator}).
+In contrast, birds of prey tend to kill prey smaller than themselves \citep{slagsvold2007prey}.
+This is likely due to their need to kill prey that they can fly away with, as well as the risk of injury being higher (which carries a higher mortality risk) for a bird than a mammal.
+Thus for birds, scavenging means they can exploit species that would otherwise be too big for them to kill.
+
+Large body size confers substantial dominance and starvation-resistance benefits \citep{ruxton2004obligate}.
+Thus, we would expect scavengers to have this trait selected for even in the case of weight-constrained fliers.
+Cinereous Vultures (\textit{Aegypius monachus}) and condors (\textit{Vultur gryphus}, \textit{Gymnogyps californianus}) all have body masses that can exceed 10 kg and represent some of the heaviest bird species capable of flight \citep{ferguson2001raptors,donazar2002effects}.
+
+And as we have noted the Azhdarchid pterosaurs were far bigger again, with estimated body masses of over 200 kg \citep{witton2010size}.
+Although \cite{witton2008reappraisal} argued that neck inflexibility and straight, rather than hooked jaw morphology points against pterosaurs existing as \textit{obligate} scavengers, Azhdarchid terrestrial proficency indicates they would have been comfortable foraging on the ground.
+Indeed, extant Marabou Storks (\textit{Leptoptilos crumenifer}) %TG: Check if that's the one
+have a comparable morphology and are noted facultative scavengers so it is reasonable to believe that certain pterosaurs behaved similarly.
+
+The competitive ability of even the largest bird is radically diminished in their interactions with mammalian competitors however, and as such they tend to consume carrion rapidly. 
+\cite{houston1974role} observed a group of \textit{Gyps} vultures consuming all of the soft tissue from a 50 kg Grant’s gazelle \textit{Nanger granti} in eight minutes.
+
+The first bird lineages did not have beaks and were predominantly carnivorous \citep{martyniuk2012field}. 
+This implies that, among the earliest species, scavenging would have been a live opportunity cf. their descendants who evolved beaks, restricting their ability to consume flesh.
+
+Avian scavengers have to compete with microorganisms and so have evolved incredibly acidic stomachs that allow them to consume and process putrefied flesh with no ill effects \citep{houston1975digestive,roggenbuck2014microbiome}.
+This adapation is not restricted to vultures though, \cite{gremillet2012vultures} showed wandering albatrosses \textit{Diomedea exulans} had an average pH of 1.5, which enables them to consume fisheries discards.
+%This led the authors to term them 'vultures of the seas'.
+
+%TG: Two points here: one speculation for the non-scavenging behaviour could be because of diet.
+Big bats (which are better suited for scavenging, following our previous argument) are typically frugivores and therefore probably lack of adaptations for digesting  meat.
+While carnivorous bats are mainly found in the microbats which are insectivorous.
+% Maybe they scavenge on dead insects but here we're talking about scavenging on real meat right?
+%TG: Second point: bats fossil record is of course not so good but it's actually way better than birds!
+Most of the data we have is on teeth of course but it's probably enough for having a rough body mass and diet estimate!
+And that's much better than birds where you're stuck with just some fragments of long bones.
+
 \subsection*{Environment}
-A major difficulty for terrestrial scavengers is competition with vultures.
-Noctural behaviour in the Hyaenidae in general has been put forth as an adaptation to reduce competition with these exclusively diurnal birds \citep{gittleman2013carnivore}.
-
+Both the biotic and abiotic environment a would-be scavenger finds itself in can influence to degree to which it can depend on carrion. 
+As noted earlier, a system similar to the Serengeti in productivity could have supported a terrestrial scavenger \citep{ruxton2004obligate}.
+Indeed, in systems that were dominated by large ectothermic or mesothermic vertebrates 
+the same primary productivity would have supported a greater biomass \citep{mcnab2009resources}.
+The upshot of this is there was a higher biomass of herbivores dying and offering scavenging opportunities.
+Predators were large-bodied too compared to extant mammalian predators \citep{mcnab2009resources}, and so, especially if they were ectothermic, could last longer between meals rendering scavenging a more attractive behaviour relative to predation.
+The absence of flying vertebrates in the Palaeozoic may have permitted terrestrial forms to take in a higher proportion of carrion in their diet.
 Osteophagy may have been even more viable during the Mesozoic era because the body mass distribution of herbviores tended to be skewed towards larger sizes \citep{10.1371/journal.pone.0051925}.
@@
 
-The absence of flying vertebrates in the Palaeozoic may have permitted terrestrial forms to take in a higher proportion of carrion in their diet.
-As noted earlier, a system similar to the Serengeti in productivity could have supported a terrestrial scavenger \citep{ruxton2004obligate}.
-Indeed, in systems that were dominated by large ectothermic or mesothermic vertebrates 
-the same primary productivity would have supported a greater biomass \citep{mcnab2009resources}.
-The upshot of this is there was a higher biomass of herbivores dying and offering scavenging opportunities.
-Predators were large bodied too compared to extant mammalian predators \citep{mcnab2009resources}, and so, especially if they were ectothermic, could last longer between meals rendering scavenging a more attractive behaviour relative to predation.
+Vultures and eagles tend to soar using thermals and if these air pockets don't form, say on a cloudy day, the bird is grounded \citep{mundy1992vultures}.
+In many habitats (e.g. the arctic) it is simply not possible for sufficiently powerful thermals to form and as a consequence large-bodied vultures cannot exist.
+The upshot of this is that terrestrial carnivores like bears and wolves take more carrion \citep{devault2003scavenging}.
+The use of different sensory systems also illustrates the impact of the environment. 
+The relatively open savanna systems of Africa are well suited to a visually dependent vulture whereas more forested areas would select for species that have a well developed olfactory system  \citep{houston1986olfaction}.
 
-%TG: To be modified and inserted somewhere:
-Scavenging behaviour might have evolved as soon as the first terrestrial tetrapods evolved.
-In fact, some of the earlier tetrapods tracks dating back to the early Middle Devonian (393.3 - 387.7 Mya) were found in intertidal environments \citep{Niedzwiedzki2009}.
-These environments are isolated from marine systems twice a day leaving potential carrion unexploited by marine vertebrates.
-\cite{Niedzwiedzki2009} suggests that these environments ``would thus have allowed marine ancestors of tetrapods gradually to acquire terrestrial competence while accessing a new and essentially untouched resource.''
-%TG: orginal sentence "The intertidal environment provides a ready food source of stranded marine animals on a twice-daily basis, in the immediate vicinity of the sea, and would thus have allowed marine ancestors of tetrapods gradually to acquire terrestrial competence while accessing a new and essentially untouched resource."
+A major difficulty for terrestrial scavengers is competition with vultures.
+Noctural behaviour in the Hyaenidae in general has been put forth as an adaptation to reduce competition with these exclusively diurnal birds \citep{gittleman2013carnivore}.
 
-% and sensory perception \citep{farlow1994speculations}.
-
-\section*{Aquatic Scavengers} 
-Primary productivity is lower in almost all aquatic systems than terrestrial systems (except deserts) so as we go up the food chain the density of carcasses worth scavenging is going to be lower.
-But it is certainly known to occur in many aquatic vertebrates.
-An aquatic environment presents challenges for direct observational studies and so, similar to the approaches involving extinct species, much work has approached the question of scavenging propensity from an energetics perspective.
-Although, as with the aerial and terrestrial enviornments we have evidence of facultative scavenging among aquatic species.
-For example, the remains of a mosasaur and a terrestrial hadrosaur were discovered with embedded teeth from a Cretaceous shark, \textit{Squalicorax} \citep{schwimmer1997scavenging}.
-As well as a likely instance of scavenging between a 4-million-year-old white shark (\textit{Carcharodon}) and mysticete whale from Peru \citep{ehret2009caught}.
-
-The existence of an obligate scavenger in a marine setting is uncertain \citep{britton1994marine,smith2003ecology,ruxton2004energetic,ruxton2005searching}.
-Depending on the species, a carcass in this environment either floats or descends to the sea floor \citep{Whitehead415}.
-In the latter low-light environment, visual detection distances are far lower (< 100 m) than they would be in the air.
-As such, animals detect resources through chemo- and mechanoreception more so than through vision \citep{ruxton2004energetic}.
-However, water is a medium that is conducive to low-cost movement \citep{tucker1975energetic} and so may be able to support an obligate scavenging fish \citep{ruxton2004energetic,ruxton2005searching}.
-\cite{beasley2015vertebrates} do note that ``some benthic scavengers (e.g., hagfish: family Myxinidae) rely on necrophagy for a large portion of their diet and may indeed be obligate scavengers".
-
-Extant aquatic snakes are deemed as having the most suitable physiology and environment for scavenging.
-A hypothesis put forth by \cite{sazima1990necrofagia} argued that chemical gradients in water would allow for a relatively easier detection of carrion.
-This gained some support from \cite{devault2002scavenging}, who found a preponderence of aquatic snake species in their review of this behaviour.
+% Although, as with detection ability, the environment has a role to play here.
 
@@
 Early whales such as \textit{Basilosaurus} seem to fit into the same niche as Killer Whales and we have some evidence for scavenging in this group as well \citep{fahlke2012bite}.
-A final point to make is that vertebrates are relatively rarer in aquatic environments, because even large animals can get support from the buoyancy of the water without needing a backbone.
+
+%TG: what do you mean by their position? Phylogenetic? Position on the "scavenging scale"?
+
+% This point illustrates how the environment can impact search efficiency depending on the sensory system that's used.
+
+% \section*{Terrestrial Scavengers}
+
+%\cite{ruxton2004obligate} offer a reason for this in that the traits that allow for vultures to exist as scavengers undermined their ability to hunt but that the same forces have not prevented mammals from doing so.
+
+%\subsection*{Environment}
+
+%TG: To be modified and inserted somewhere:
+
+%TG: orginal sentence "The intertidal environment provides a ready food source of stranded marine animals on a twice-daily basis, in the immediate vicinity of the sea, and would thus have allowed marine ancestors of tetrapods gradually to acquire terrestrial competence while accessing a new and essentially untouched resource."
+
+% and sensory perception \citep{farlow1994speculations}.
+
+%\section*{Aquatic Scavengers} 
+%An aquatic environment presents challenges for direct observational studies and so, similar to the approaches involving extinct species, much work has approached the question of scavenging propensity from an energetics perspective.
+%But it is certainly known to occur in many aquatic vertebrates.
+
+%A point to note is that vertebrates are relatively rarer in aquatic environments, because even large animals can get support from the buoyancy of the water without needing a backbone.
+
+%\subsection*{Locomotion}
+
+%\subsection*{Detection}
+%The existence of an obligate scavenger in a marine setting is uncertain \citep{britton1994marine,smith2003ecology,ruxton2004energetic,ruxton2005searching}.
+%Depending on the species, a carcass in this environment either floats or descends to the sea floor \citep{Whitehead415}.
+%In the latter low-light environment, visual detection distances are far lower (< 100 m) than they would be in the air.
+%As such, animals detect resources through chemo- and mechanoreception more so than through vision \citep{ruxton2004energetic}.
+\cite{beasley2015vertebrates} do note that ``some benthic scavengers (e.g., hagfish: family Myxinidae) rely on necrophagy for a large portion of their diet and may indeed be obligate scavengers".
+
+%\subsection*{Processing}
+
+%\subsection*{Environment}
+
+%Primary productivity is lower in almost all aquatic systems than terrestrial systems (except deserts) so as we go up the food chain the density of carcasses worth scavenging is going to be lower.
 
@@
 % It is recognised that scavengers keep energy flows at a higher trophic level in food webs than decomposers because they consume relatively more carrion \citep{devault2003scavenging}.
-They are also hugely important for the dispersal of nutrients \citep{beasley2015vertebrates}.
-Consider the diversity of animals that can end up feeding at the carcass of an elephant.
-Here we have an incredibly dense and nutrient rich patch that ends up being distributed widely.
-In the absence of vertebrate scavengers, invertebrates and microorganisms would consume the carcass in-situ or at least distribute the constituent nutrients over a much shorter range.
-This effect has been magnified as vertebrates evolved certain key traits that allowed them to range farther, namely an upright gait, an endothermic metabolism and of course, flight.
-To quantify this effect with a simple example we can turn to some allometric relationships relating sustainable travelling speed to body mass.
-In the case of mammals and reptiles these are 1.15 * body mass (kg) \textsuperscript{0.12} and 0.23 * body mass (kg) \textsuperscript{0.12}
+%They are also hugely important for the dispersal of nutrients \citep{beasley2015vertebrates}.
+%Consider the diversity of animals that can end up feeding at the carcass of an elephant.
+%Here we have an incredibly dense and nutrient rich patch that ends up being distributed widely.
+%In the absence of vertebrate scavengers, invertebrates and microorganisms would consume the carcass in-situ or at least distribute the constituent nutrients over a much shorter range.
+%This effect has been magnified as vertebrates evolved certain key traits that allowed them to range farther, namely an upright gait, an endothermic metabolism and of course, flight.
+%To quantify this effect with a simple example we can turn to some allometric relationships relating sustainable travelling speed to body mass.
+%In the case of mammals and reptiles these are 1.15 * body mass (kg) \textsuperscript{0.12} and 0.23 * body mass (kg) \textsuperscript{0.12}
 % respectively \citep{ruxton2004obligate}.
-We can insert these into a foraging radius model ((duration * speed)/2)/1000 for a 12 hour foraging day which shows that while a 10 kg reptile can range 6.5 km an equally sized mammal can range nearly 33 km \citep{Enstipp2006Energetics}.
-Thus, in an ecological context, the evolution of these steps coupled with the ability to scavenge resulted in a world with a far more widely distributed nutrient landscape.
+%We can insert these into a foraging radius model ((duration * speed)/2)/1000 for a 12 hour foraging day which shows that while a 10 kg reptile can range 6.5 km an equally sized mammal can range nearly 33 km \citep{Enstipp2006Energetics}.
+%Thus, in an ecological context, the evolution of these steps coupled with the ability to scavenge resulted in a world with a far more widely distributed nutrient landscape.
 
 %\section*{The Role of Climate} %TG: Use it as a opening at the end of the paper (i.e. conclusion + finishing with this bit)
+\section*{Conclusion} 
 The geological record shows the Earth has undergone radical fluctuations in temperature.

### 2016-03-24 natural history of scavenging.tex
Edits on locomotion section

@@
 \usepackage{amsmath}
+\usepackage{xfrac}
 
@@
 This may go some way to explaining the gap in our knowledge of the prevalence of this behaviour.
-Consider Professor Sanborn Tenney writing in 1877 for The American Naturalist who had this to say about one well known group, ``Prominent among the mammalian scavengers are the hyenas, the ugliest in their general appearance of all the flesh eaters."
+Consider Professor Sanborn Tenney writing in 1877 for The American Naturalist who had this to say about one well known group, ``prominent among the mammalian scavengers are the hyenas, the ugliest in their general appearance of all the flesh eaters."
 He contrasts these with ``nobler kinds" of carnivores such as lions and tigers \citep{tenney1877few}.
@@
 To quantify this effect with a simple example we can turn to some allometric relationships relating sustainable travelling speed to body mass.
-In the case of mammals and reptiles these are 1.15 * body mass (kg) \textsuperscript{0.12} and 0.23 * body mass (kg) \textsuperscript{0.12} respectively \citep{ruxton2004obligate}.
-We can insert these into a foraging radius model ((duration * speed)/2)/1000 for a 12 hour foraging day which shows that while a 10 kg reptile can range 6.5 km an equally sized mammal can range nearly 33 km \citep{Enstipp2006Energetics}.
+In the case of mammals and reptiles these are 1.15 $\times$ body mass (kg) \textsuperscript{0.12} and 0.23 $\times$ body mass (kg) \textsuperscript{0.12} respectively \citep{ruxton2004obligate}.
+We can insert these into a foraging radius model $\sfrac{\frac{\text{duration} \times \text{speed}}{2}}{1000}$ for a 12 hour foraging day which shows that while a 10 kg reptile can range 6.5 km an equally sized mammal can range nearly 33 km \citep{Enstipp2006Energetics}.
 
@@
 In terms of locomotion, they employ a characteristic ``rocking horse gait"  which allows them to cover great distances efficiently, loping at 10 km/hr \citep{mills1989comparative,jones2015african}. 
-The order Carnivora saw its origins in the Middle Eocene (56-33.9 Mya) where it split into the Caniforma and Feliforma.
+The order Carnivora saw its origins in the Middle Eocene (56-33.9 Million years ago - Mya) where it split into the Caniforma and Feliforma.
+%TG: a phylo side note: the correct sentence would be the "The stem Carnivora saw its origins ...
+when it split into Caniforma and Feliforma." But I think we can leave it out because, god, it sounds pedantic!
 So we can trace efficient terrestrial movement by carnivores from this point on.
 Certainly, it is profitable to compare extant terrestrial species to their prehistoric forebears given the dominance of mammalian carnviores since the Eocene.
+%TG: ewgh...
+Thomas does not entirely agree.
+But then again, until now, reviewers didn't agreed with Thomas.
 
@@
 We do know that they walked with the erect gait of mammals or birds rather than the sprawling gait of lizards and that they were most likely facultative scavengers \citep{weishampel2004dinosauria,depalma2013physical}.
+%TG: and cite "Kane et al in press" no?
+I think these people did some kind of work on that matter.
 Taken together, this implies dinosaurs had a foraging range that fell in between that of modern terrestrial mammals and reptiles.
@@
 Of course, tetrapod terrestrial dominance predates the evolution of the dinosaurs.
+%TG: I will down that "dominance" term.
+Maybe change the sentence to "Of course the importance of terrestrial tetrapods in ecosystems predates the evolution of the dinosaurs."
 It is in the Permian, almost 300 millions years ago, that we have the earliest evidence of vertebrate scavenging where a temnospondyl amphibian fed on the carcass of \textit{Varanops}, a predatory synapsid of the time \citep{reisz2006articulated}.
@@
 
+%TG: I think it's missing a transition here: maybe something along "But moving on land as it's limitations..."
+
 Flight is a cheaper means of locomotion than walking or running \citep{tucker1975energetic}. 
@@
 Vultures consist of two convergent groups, from the old and the new world and represent the only example of obligate vertebrate scavengers today.
-The families from which modern vultures arose, the Accipitridae and Cathartidae, appear during the Palaeocene (66 - 56 Mya) \citep{Jetz2012, Jarvis2014}.
-Avian flight is far older than this and originates in the Late Jurassic (163.5-145 Million years ago - Mya), conincident with the fossils of \textit{Archaeopteryx lithographica} so many of these benefits would have been realised from that point on for carnivorous birds.
+The families from which modern vultures arose, the Accipitridae and Cathartidae, appear during the Palaeocene \citep[66 - 56 Mya; ][]{Jetz2012, Jarvis2014}.
+Avian flight is far older than this and originates in the Late Jurassic (163.5-145 Mya), conincident with the fossils of \textit{Archaeopteryx lithographica} so many of these benefits would have been realised from that point on for carnivorous birds.
 However, vertebrate flight is much older still where pterosaurs predate bird origins by a considerable margin in the Late Triassic (235-201.3 Mya).
 Scavenging in this diverse group has been hypothesied many times before \citep{witton2008reappraisal}.
-Certain clades of these animals could reach enormous sizes (e.g.
-Azhdarchids with wingspans of 11 metres \citep{witton2010size}) and, notably, look to have engaged in soaring flight \citep{witton2010size}.
+Certain clades of these animals could reach enormous sizes \citep[e.g.
+Azhdarchids with wingspans of 11 metres; ][]{witton2010size} and, notably, look to have engaged in soaring flight \citep{witton2010size}.
+%TG: this paragraph's a killer! Love it!
 
@@
 The bat fossil record is notoriously poor owing to their fragile skeletons so we are unable to determine if extinct species were more suited to this lifestyle \citep{eiting2009global}.
+%TG: again, I'm not sure it's notoriously poor!
+It might be poor for mammalogist standards but it's definitely a way better than birds fossil records for example!
 Although it does not seem that flight is the main criterion precluding them from scavenging (see below).
@@
 But despite this movement away from land the energetic savings were negligible because the \textit{total} cost incurred by a swimming marine mammal is high \citep{williams1999evolution}. 
-As \cite{williams1999evolution} notes, ``Free-ranging animals must contend with the total energetic expenditure associated with supporting basic biological functions as well as with moving the body and appendages through the environment."
+As \cite{williams1999evolution} notes, ``free-ranging animals must contend with the total energetic expenditure associated with supporting basic biological functions as well as with moving the body and appendages through the environment."
 

### 2016-03-24 natural history of scavenging.tex
Updates mainly on part 2

@@
 Hyenas have well developed sensory organs, whereby they can smell a rotting carcass 4 km away and can hear the vocalisations of conspecifics at a distance of 10 km \citep{mills1989comparative}. 
-Using the approach of \cite{spiegel2013factors} we estimate a spotted hyena could resolve a 2 metre target at 1 km distance. 
-\cite{ruxton2004obligate} calculated that ``a 1 tonne mammal or reptile, in an ecosystem yielding carrion at densities similar to the current Serengeti, could have met its energy requirements if it could detect carrion over a distance of the order of 400–500 m.".
+Using the approach of \cite{spiegel2013factors} we estimate a spotted hyena %TG: do we need a (\textit{Crocuta crocuta}) here?
+could resolve a 2 metre target at 1 km distance. 
+\cite{ruxton2004obligate} calculated that ``a 1 tonne mammal or reptile, in an ecosystem yielding carrion at densities similar to the current Serengeti, could have met its energy requirements if it could detect carrion over a distance of the order of 400–500 m".
 
@@
 Clearly, it would be pointless to have incredible ranging abilites and not have the sensory architecture to benefit from it.
+%TG: this sentence makes totally sense but starting it by "Clearly" kind of undermines it.
+To me it sounds like when in a presentation some spawn of Satan puts up a table and starts saying: "As you can see".
 % This would be impossible for the visually reliant old world species.
 
+%TG: I think it's missing some connection here.
+The "returning" makes it a bit jumpy.
+Maybe it could be better to include that directly in the paragraph above since it is making the point: "Some other volant groups, however, have not developed all these things".
 Returning to bats, their visual acuity is famously poor, and echolocation would not lend itself to discovering immobile carrion.
 Their small size and poor terrestrial ability would also count against them at a carcass \citep{riskin2006terrestrial}.
+
+%TG: Same, I feel it's missing a transition here.
 
@@
 Osteophagy is known across a range of terrestrial carnivores and given some fat-rich mammalian bones have an energy density (6.7 kJ/g) comparable with that of muscle tissue, it makes skeletal remains an enticing resource \citep{brown1989study}.
-This ability reached its zenith among hyenas with the evolution of the 110 kg \textit{Pachycrocuta brevirostris} during the Pliocene \citep{palmqvist2011giant}.
+This ability reached its zenith among hyenas with the evolution of the 110 kg \textit{Pachycrocuta brevirostris} during the Pliocene \citep[3.6 - 2.58 Mya; ][]{palmqvist2011giant}.
 Some work on extinct sabretooths suggests they may have left a large amount of food for would-be scavengers because of their unique skull morphology.
@@
 % These traits have been used to infer scavenging lifestyles in extinct mammals. 
-The bone-crushing dogs that evolved during the Oligocene (subfamily Borophaginae) have been compared to hyenas in terms of their feeding ecology as well \citep{van2003chapter,martin2016pursuit}.
+The bone-crushing dogs that evolved during the Oligocene (subfamily Borophaginae; 33.9 - 23.03 Mya) have been compared to hyenas in terms of their feeding ecology as well \citep{van2003chapter,martin2016pursuit}.
 

### 2016-03-29 natural history of scavenging.tex
Minor edits to locomotion part

@@
 Certainly, it is profitable to compare extant terrestrial species to their prehistoric forebears given the dominance of mammalian carnviores since the Eocene.
-%TG: ewgh...
-Thomas does not entirely agree.
-But then again, until now, reviewers didn't agreed with Thomas.
 
 Unsurprisingly, given their enduring appeal, the prevalence of scavenging has been explored in the carnivorous, theropod dinosaurs.
-Dinosaurs in general, were the dominant terrestrial forms for most of the Mesozoic Era. 
-where the theropods ranged from the chicken-sized to the whale-sized, all of which were bipedal.
+They were the dominant terrestrial forms for most of the Mesozoic Era (252.17 - 66 Mya) and ranged from the chicken-sized to the whale-sized, all of which were bipedal.
 They are quite alien to anything we know today which restricts our ability to understand their ecology far more so than extinct mammals \citep{weishampel2004dinosauria}.
@@
 Maybe change the sentence to "Of course the importance of terrestrial tetrapods in ecosystems predates the evolution of the dinosaurs."
-It is in the Permian, almost 300 millions years ago, that we have the earliest evidence of vertebrate scavenging where a temnospondyl amphibian fed on the carcass of \textit{Varanops}, a predatory synapsid of the time \citep{reisz2006articulated}.
+It is during the Permian (298.9 - 252.17 Mya) %, almost 300 millions years ago, %TG: Being pedantic here but Permian is just *not* 300 Mya...
+As a side note for the dates, I say we just use the geological epochs/periods/whatever since they are the actual precise dating of the fossil occurence (and they sound better).
+I suggest we just add the dates in brackets to give an indication.
+that we have the earliest evidence of vertebrate scavenging where a temnospondyl amphibian fed on the carcass of \textit{Varanops}, a predatory synapsid of the time \citep{reisz2006articulated}. 
 And it is with the evolution of endothermy in the therapsid-mammal lineage \citep{clarke2010temperature} that terrestrial vertebrates would have had the ability to range widely, a vital component in seeking out carrion.

### 2016-03-29 natural history of scavenging.tex
snakes as detoxifiers

@@
 %TG: I think it's missing a transition here: maybe something along "But moving on land as it's limitations..."
-
+But it is in the air that we find scavengers par excellence. 
 Flight is a cheaper means of locomotion than walking or running \citep{tucker1975energetic}. 
-Many extant birds exist as facultative scavengers; storks, eagles, corvids, are all known to take substantial quantities of carrion in their diet \citep{kendall2013alternative}.
+We know that many extant birds exist as facultative scavengers; storks, eagles, corvids, all take substantial quantities of carrion in their diet \citep{kendall2013alternative}.
 The advantage of flight can be extended further in larger species that engage in soaring instead of flapping flight, which is even cheaper still (approximately twice the basal metabolic rate) \citep{hedenstrom1993migration,spivey2014analysing}.
@@
 As such, animals detect resources through chemo- and mechanoreception more so than through vision \citep{ruxton2004energetic}.
-Extant aquatic snakes are deemed as having the most suitable physiology for scavenging.
+This is particularly relevant to extant aquatic snakes who are deemed as having the most suitable physiology for scavenging.
 A hypothesis put forth by \cite{sazima1990necrofagia} argued that chemical gradients in water would allow for a relatively easier detection of carrion.
@@
 This is significant because hyenas are likely to have left far less flesh on a carcass than a felid such as a sabretooth enabling contemperaneous hominins to benefit \citep{pobiner2015new}.
-
+The intelligence, resultant tool-use and cooperative nature of hominins meant they could likely adapt to take on more or less carrion depending on their environment \citep{moleon2014humans}.
 %TG: or "could"? This is hypothetical behaviour stuff! Maybe past hyenas where exclusive sushi eaters!
 %TG: Not sure about the intelligence thing: the sentence as the same meaning without and will be less anthropocentric sounding.
-The intelligence, resultant tool-use and cooperative nature of hominins meant they could likely adapt to take on more or less carrion depending on their environment \citep{moleon2014humans}.
 
@@
 
-Avian scavengers have to compete with microorganisms and so have evolved incredibly acidic stomachs that allow them to consume and process putrefied flesh with no ill effects \citep{houston1975digestive,roggenbuck2014microbiome}.
-This adapation is not restricted to vultures though, \cite{gremillet2012vultures} showed wandering albatrosses \textit{Diomedea exulans} had an average pH of 1.5, which enables them to consume fisheries discards.
-%This led the authors to term them 'vultures of the seas'.
+\cite{shivik2006vultures} points out that ``evolutionary pressures favor detection maximizers relative to toxification minimizers in competitive interactions for carcasses." 
+But the fact remains that overcoming microorganism toxins is still a beneficial adaptation to any scavenger. 
+Avian scavengers have evolved incredibly acidic stomachs that allow them to consume and process putrefied flesh with no ill effects \citep{houston1975digestive,roggenbuck2014microbiome}. 
+This adapation is not restricted to vultures though, \cite{gremillet2012vultures} showed wandering albatrosses (\textit{Diomedea exulans} so-called 'vultures of the seas') had an average pH of 1.5, which enables them to consume fisheries discards.
+Outside of the birds there is evidence of selection for 'toxification minimizers'.
+From our earlier arguments we know that ecthotherms are limited in their ability to find carrion as quickly as endotherms. 
+This implies later arrivers would benefit especially from well-developed detoxifying apparatus. 
+\cite{shivik2006vultures} suggests that ``specialized oral structures in snakes may have evolved under pressures associated with scavenging."
+Moreover, some authorities have charted an evolutionary course from basal fossorial snakes to modern terrestrial species by way of an obligate scavenger intermediate \citep{bauchot2006snakes}.
+
+%Being far less vagile than other vertebrate species,
+%snakes are expected to develop detoxification strategies to overcome
+%chemical defenses and make the best use of carrion.
+%There is additional evidence that  Evolutionary pressures are not limited to competition
+%for carcasses, of course, but evolutionarily, as snakes
+%developed from eyeless fossorial species and radiated into terrestrial
+%and arboreal predatory species (Rage 1994), an obligate
+%scavenger evolutionary intermediate was likely.
 
 %TG: Two points here: one speculation for the non-scavenging behaviour could be because of diet.
-Big bats (which are better suited for scavenging, following our previous argument) are typically frugivores and therefore probably lack of adaptations for digesting  meat.
+It is in the ability to process carrion that bats suffer.
+Big bats (which are better suited for scavenging, following our previous argument) are typically frugivores and therefore lack the adaptations for digesting  meat.
 While carnivorous bats are mainly found in the microbats which are insectivorous.
@@
 And that's much better than birds where you're stuck with just some fragments of long bones.
+
+% Durophagy amount aquatic species, turtles, sharks, trigger fish, crocs etc.
 

### 2016-03-30 natural history of scavenging.tex
Added carcasses on figure 1 + included it in text
updated figure with dino! (doesn't look that great though...
cleaned up sections

@@
 
+\begin{figure}[!htbp]
+\centering
+   \includegraphics[width=1\textwidth]{Summary_figure/Summary_figure.pdf}
+\caption{Factors influencing the proportion of scavenging in a vertebrates' diet. Blue lines indicates a reduction in the factor and red lines indicates an increase.}
+%TG: Probably needs some more explanations...
+\label{Summary_figure}
+\end{figure}
+
 %KH: need paragraph for how do we know something is (was) a scavenger
@@
 This is exacerbated by the sprawling gait seen in many lizards which results in Carrier's Constraint such that the animal can't move and breathe at the same time because the lateral movements impedes its lungs \citep{carrier1987evolution}.
-An effect which manifests itself in aspects such as maximum sustainable speed where an equivalent mammal has a six to seven fold increase \citep{ruben1995evolution}.
+This constraint manifests itself in aspects such as maximum sustainable speed where an equivalent mammal has a six to seven fold increase \citep{ruben1995evolution}.
 To quantify this effect with a simple example we can turn to some allometric relationships relating sustainable travelling speed to body mass.
@@
 We can insert these into a foraging radius model $\sfrac{\frac{\text{duration} \times \text{speed}}{2}}{1000}$ for a 12 hour foraging day which shows that while a 10 kg reptile can range 6.5 km an equally sized mammal can range nearly 33 km \citep{Enstipp2006Energetics}.
+For a foraging scavenger, this ability translates into a greater area searched for food.
 
@@
 In the spotted hyena (\textit{Crocuta crocuta}), striped hyena (\textit{Hyaena hyaena}) and brown hyena (\textit{Hyaena brunnea}) it can be over 90\% \citep{jones2015african}.
-Yet, no contemporary terrestrial vertebrate exists as an obligate scavenger but most if not all are facultative to some extent \citep{beasley2015vertebrates}.
-Therefore, we can use hyenas as examples of efficient terrestrial scavengers to compare with other forms. 
+And although no contemporary terrestrial vertebrate exists as an obligate scavenger most if not all are facultative to some extent \citep{beasley2015vertebrates}.
+The particular reliance of hyenas on carrion means we can use them as examples of efficient terrestrial scavengers to compare with other forms. 
 In terms of locomotion, they employ a characteristic ``rocking horse gait"  which allows them to cover great distances efficiently, loping at 10 km/hr \citep{mills1989comparative,jones2015african}. 
-The order Carnivora saw its origins in the Middle Eocene (56-33.9 Million years ago - Mya) where it split into the Caniforma and Feliforma.
-%TG: a phylo side note: the correct sentence would be the "The stem Carnivora saw its origins ...
+Such long-distance travel is apparent in African wild dogs (\textit{Lycaon pictus}) and many other canids \citep{pennycuick1995radius,janis2014forelimb}. 
+In contrast, big cats like leopards (\textit{Panthera pardus}) rely on ambush \citep{pennycuick1995radius}. 
+This allows us to make a broad distinction between the ambush strategies of cats and the pursuit/ pounce strategies of dogs, the latter being more suited to scavenging \citep{janis2014forelimb}. 
+We can (and have) use(d) these insights to compare extant terrestrial species to their prehistoric forebears given the dominance of mammalian carnviores since the Eocene 56-33.9 Million years ago (Mya) where the order split into the Caniforma and Feliforma \citep{van1987skeletal}.
+To take one example, \cite{anyonge1996locomotor} found \textit{Nimravides}, a genus of sabretooth cat from the Miocence (10.3 to 5.3 M ya) were likely to have been ambush predators which would have counted against them taking a lot of carrion.
+%The order Carnivora saw its origins in the Middle Eocene  %TG: a phylo side note: the correct sentence would be the "The stem Carnivora saw its origins ...
 when it split into Caniforma and Feliforma." But I think we can leave it out because, god, it sounds pedantic!
-So we can trace efficient terrestrial movement by carnivores from this point on.
-Certainly, it is profitable to compare extant terrestrial species to their prehistoric forebears given the dominance of mammalian carnviores since the Eocene.
+% So we can trace efficient terrestrial movement by carnivores from this point on.
 
@@
 that we have the earliest evidence of vertebrate scavenging where a temnospondyl amphibian fed on the carcass of \textit{Varanops}, a predatory synapsid of the time \citep{reisz2006articulated}. 
-And it is with the evolution of endothermy in the therapsid-mammal lineage \citep{clarke2010temperature} that terrestrial vertebrates would have had the ability to range widely, a vital component in seeking out carrion.
+And it is with the evolution of endothermy in the therapsid-mammal lineage \citep{clarke2010temperature} that terrestrial vertebrates would have gained the ability to range widely, a vital component in seeking out carrion.
 
-Scavenging behaviour might have evolved on land as soon as the first terrestrial tetrapods evolved.
+Scavenging behaviour may have evolved on land as soon as the first terrestrial tetrapods emerged.
 In fact, some of the earlier tetrapods tracks dating back to the early Middle Devonian (393.3 - 387.7 Mya) were found in intertidal environments \citep{Niedzwiedzki2009}.
 These environments are isolated from marine systems twice a day leaving potential carrion unexploited by marine vertebrates.
-\cite{Niedzwiedzki2009} suggests that these environments ``would thus have allowed marine ancestors of tetrapods gradually to acquire terrestrial competence while accessing a new and essentially untouched resource.''
+\cite{Niedzwiedzki2009} suggest that these environments ``would thus have allowed marine ancestors of tetrapods gradually to acquire terrestrial competence while accessing a new and essentially untouched resource.''
 
@@
 Flight is a cheaper means of locomotion than walking or running \citep{tucker1975energetic}. 
-We know that many extant birds exist as facultative scavengers; storks, eagles, corvids, all take substantial quantities of carrion in their diet \citep{kendall2013alternative}.
+We know that many extant birds exist as facultative scavengers because storks, raptors and corvids all take substantial quantities of carrion in their diet \citep{kendall2013alternative}.
 The advantage of flight can be extended further in larger species that engage in soaring instead of flapping flight, which is even cheaper still (approximately twice the basal metabolic rate) \citep{hedenstrom1993migration,spivey2014analysing}.
 The benefits this confers are clear from the information we have on the enormous foraging ranges of many vultures \citep{spiegel2013factors} and seabirds \citep{thaxter2012seabird}.
-In the former case we have the best known scavengers on Earth. 
-Vultures consist of two convergent groups, from the old and the new world and represent the only example of obligate vertebrate scavengers today.
+In vultures we have the best known scavengers on Earth. 
+These birds consist of two convergent groups, from the old and the new world where they represent the only example of obligate vertebrate scavengers today.
 The families from which modern vultures arose, the Accipitridae and Cathartidae, appear during the Palaeocene \citep[66 - 56 Mya; ][]{Jetz2012, Jarvis2014}.
-Avian flight is far older than this and originates in the Late Jurassic (163.5-145 Mya), conincident with the fossils of \textit{Archaeopteryx lithographica} so many of these benefits would have been realised from that point on for carnivorous birds.
-However, vertebrate flight is much older still where pterosaurs predate bird origins by a considerable margin in the Late Triassic (235-201.3 Mya).
+Yet, avian flight is far older than this and originates in the Late Jurassic (163.5-145 Mya), conincident with the fossils of \textit{Archaeopteryx lithographica} so many of these benefits would have been realised from that point on for carnivorous birds.
+And vertebrate flight is much older still where pterosaurs predate bird origins by a considerable margin in the Late Triassic (235-201.3 Mya).
 Scavenging in this diverse group has been hypothesied many times before \citep{witton2008reappraisal}.
@@
 This has led some researchers to argue for the feasability of a scavenging fish \citep{ruxton2004energetic,ruxton2005searching}. 
-As with the aerial and terrestrial enviornments we have evidence of facultative scavenging among extant and extinct aquatic species.
+As with the aerial and terrestrial enviornments we have evidence of facultative scavenging among extinct aquatic species.
 For example, the remains of a mosasaur and a terrestrial hadrosaur were discovered with embedded teeth from a Cretaceous shark, \textit{Squalicorax} \citep{schwimmer1997scavenging}.
 As well as a likely instance of scavenging between a 4-million-year-old white shark (\textit{Carcharodon}) and mysticete whale from Peru \citep{ehret2009caught}.
+Extant White sharks \textit{Carcharodon carcharias} too are known to feed on whale carcasses \citep{fallows2013white}.
 We might expect then that by combining an aquatic environment and an endothermic metabolism that marine mammals would prosper as scavengers.
@@
 But despite this movement away from land the energetic savings were negligible because the \textit{total} cost incurred by a swimming marine mammal is high \citep{williams1999evolution}. 
-As \cite{williams1999evolution} notes, ``free-ranging animals must contend with the total energetic expenditure associated with supporting basic biological functions as well as with moving the body and appendages through the environment."
+This is not to say that aquatic mammalian scavengers don't exist, only that their total energetic cost is similar to an equivalent terrestrial mammal. 
+%\cite{williams1999evolution} notes, ``free-ranging animals must contend with the total energetic expenditure associated with supporting basic biological functions as well as with moving the body and appendages through the environment."
 
 \subsection*{Detection}
+It would be pointless to have incredible ranging abilites and not have the sensory architecture to benefit from it.
+If we came at this from a position of complete ignorance we would predict scavengers to have well-developed senses and indeed this is what we find. 
 A simplification of terrestrial, vertebrate scavengers in sensory terms is one of them existing in a two-dimensional plane while foraging for carrion directly.
@@
 As a consequence, they have a much more restricted view of the landscape than do aerial foragers.
-Hyenas have well developed sensory organs, whereby they can smell a rotting carcass 4 km away and can hear the vocalisations of conspecifics at a distance of 10 km \citep{mills1989comparative}. 
-Using the approach of \cite{spiegel2013factors} we estimate a spotted hyena %TG: do we need a (\textit{Crocuta crocuta}) here?
-could resolve a 2 metre target at 1 km distance. 
-\cite{ruxton2004obligate} calculated that ``a 1 tonne mammal or reptile, in an ecosystem yielding carrion at densities similar to the current Serengeti, could have met its energy requirements if it could detect carrion over a distance of the order of 400–500 m".
+Hyenas make up for this in their ability to smell a rotting carcass 4 km away and to hear the vocalisations of conspecifics at a distance of 10 km \citep{mills1989comparative}. 
+% Using the approach of \cite{spiegel2013factors} we estimate a spotted hyena %TG: do we need a (\textit{Crocuta crocuta}) here?
+%could resolve a 2 metre target at 1 km distance. 
+While considering prehistoric habitats \cite{ruxton2004obligate} calculated that ``a 1 tonne mammal or reptile, in an ecosystem yielding carrion at densities similar to the current Serengeti, could have met its energy requirements if it could detect carrion over a distance of the order of 400–500 m".
+The senses of many extant (and in all probabilty extinct) carnivores meet this required distance, making scavenging feasbile for terrstrial species \citep{farlow1994speculations,mech2010wolves}.
 
@@
 Such an ability has obvious benefits in detecting carrion.
-Vultures are known to have impressive visual acuity with one estimate indicating Lappet-faced Vultures (\textit{Torgos tracheliotus}) are capable of detecting a 2 metre carcass over 10 km away \citep{spiegel2013factors}.
+Vultures are known to have impressive visual acuity, with one estimate indicating Lappet-faced Vultures (\textit{Torgos tracheliotus}) are capable of detecting a 2 metre carcass over 10 km away \citep{spiegel2013factors}.
 Eagles too are known to have highly developed visual abilities \citep{reymond1985spatial}.
@@
 
-Aside from sight, three species within the new world family Cathartidae, (genus \textit{Cathartes}), have well developed olfactory systems \citep{AR:AR22815}.
+We can contrast this ability to bats, whose visual acuity is famously poor. 
+It also appears that echolocation would not lend itself to discovering immobile carrion.
+Their small size and poor terrestrial ability would also count against them at a carcass \citep{riskin2006terrestrial}.
+
+Aside from sight, three species of vultures within the new world family Cathartidae, (genus \textit{Cathartes}), have well developed olfactory systems \citep{AR:AR22815}.
 Among them are the Turkey Vultures (\textit{Cathartes aura}) which were able to locate 90\% of baits set out in a tropical forest \citep{houston1986olfaction}.
-An atuned sense of smell is obviously useful in detecting decaying carrion.
-Clearly, it would be pointless to have incredible ranging abilites and not have the sensory architecture to benefit from it.
+An atuned sense of smell is obviously useful in detecting decaying carrion from the air.
  %TG: this sentence makes totally sense but starting it by "Clearly" kind of undermines it.
@@
 Maybe it could be better to include that directly in the paragraph above since it is making the point: "Some other volant groups, however, have not developed all these things".
-Returning to bats, their visual acuity is famously poor, and echolocation would not lend itself to discovering immobile carrion.
-Their small size and poor terrestrial ability would also count against them at a carcass \citep{riskin2006terrestrial}.
 
@@
 
-Depending on the species, a carcass in water either floats or descends to the sea floor \citep{Whitehead415}.
-In the latter low-light environment, visual detection distances are far lower (< 100 m) than they would be in the air.
+% Depending on the species, a carcass in water either floats or descends to the sea floor \citep{Whitehead415}.
+Water tends to be a low-light environment, so visual detection distances are far lower (< 100 m) than they would be in the air.
 As such, animals detect resources through chemo- and mechanoreception more so than through vision \citep{ruxton2004energetic}.
-This is particularly relevant to extant aquatic snakes who are deemed as having the most suitable physiology for scavenging.
-A hypothesis put forth by \cite{sazima1990necrofagia} argued that chemical gradients in water would allow for a relatively easier detection of carrion.
+This is particularly relevant to extant sharks and aquatic snakes who are deemed as having the most suitable physiology for scavenging.
+A hypothesis put forth by \cite{sazima1990necrofagia} argued that chemical gradients in water would allow for a relatively easier detection of carrion by snakes.
 This gained some support from \cite{devault2002scavenging}, who found a preponderence of aquatic snake species in their review of this behaviour.
+Smell seems to be the primary means of carcass detection in sharks as well. 
+\cite{fallows2013white} found that wind speed determined the number of sharks feeding at whale carcasses indicating they were dependent on detecting the odours from the decaying whales.
 
@@
 Since carrion is not dispatched directly, often the most easily accessible and choicest components of the carcass will be missing or, if present, will be fought over.
-Being able to extract nutrients from remnants gives the scavenger a great advantage.
+Being able to extract nutrients from remnants gives a scavenger a great advantage.
 Thus, the bone crushing ability of hyenas reveals another useful scavenger trait.
@@
 
-When it comes to reptiles they possees an advantage here, in that over the course of a year their food requirements can be 30 times smaller than an endotherm of equal size \citep{Nagy1621}.
-Any adaptations that reduce energetic costs are likely to be selected in scavengers.
-\cite{devault2002scavenging} suggest this is an avenue for scavenging in snakes because they ``exhibit  exceedingly  low  maintenance  metabolisms,  and most  can  survive  on  a  few  scant  feedings per year.
-It  is, therefore, possible for snakes to rely largely  on  infrequent,  less  energy-rich  meals." In the same review the authors found occurrences of scavenging spread across five families of snakes and stated that this behaviour is ``far more common than currently acknowledged."\citep{devault2002scavenging}.
-The same reasoning can be applied to crocodlies and their allies \citep{forrest2003evidence}. 
-A sit and wait strategy is viable for an ectotherm.
+In Mesozoic systems some extremely large theropod dinosaurs had a morphology indicative of an ability to process bone e.g. the robust skull and dentition of \textit{T. rex} \citep{hone2010feeding}.
+There is direct evidence that \textit{T.
+rex} did this in the form of distinctive wear marks on its tooth apices \citep{farlow1994wear,schubert2005wear} and the presence of bone fragments in its coprolites \citep{chin1998king}.
+The animal also had an enormous bite force, with one estimate putting it at 57000 Newtons \citep{bates2012estimating}.
+This is noted as being powerful enough to break open skeletal material \citep{rayfield2001cranial}.
 
-Much work has focused on the existence of scavenging in dinosaurs by using simple energetics approaches that typically focused on a single species namely \textit{Tyrannosaurus rex} \citep{ruxton2003could,carbone2011intra} but a recent modelling study investigated the likely prevelance of scavenging across a range of body sizes.
+Further, much work has focused on the existence of scavenging in dinosaurs by using simple energetics approaches that typically focused on a single species namely \textit{Tyrannosaurus rex} \citep{ruxton2003could,carbone2011intra} but a recent modelling study investigated the likely prevelance of scavenging across a range of body sizes.
 In it the authors demonstrated that species of intermediate body masses (approximatively 500 kg) would have gained the most benefit from scavenging.
@@
 
-In Mesozoic systems some extremely large theropod dinosaurs had a morphology which suggests an ability to process bone e.g. the robust skull and dentition of \textit{T. rex} \citep{hone2010feeding}.
-There is direct evidence that \textit{T.
-rex} did this in the form of distinctive wear marks on its tooth apices \citep{farlow1994wear,schubert2005wear} and the presence of bone fragments in its coprolites \citep{chin1998king}.
-The animal also had an enormous bite force, with one estimate putting it at 57000 Newtons \citep{bates2012estimating}.
-This is noted as being powerful enough to break open skeletal material \citep{rayfield2001cranial}.
+In addition to reducing locomotory costs we would expect adaptations that reduce energetic costs of maintenance to be selected for in scavengers because it would maximise the benefit derived from such a sporadic food source.
+Extant reptiles possess an advantage here, in that over the course of a year their food requirements can be 30 times smaller than an endotherm of equal size \citep{Nagy1621}.
+\cite{devault2002scavenging} suggest this is an avenue for scavenging in snakes because they ``exhibit  exceedingly  low  maintenance  metabolisms,  and most  can  survive  on  a  few  scant  feedings per year.
+It  is, therefore, possible for snakes to rely largely  on  infrequent,  less  energy-rich  meals." In the same review the authors found occurrences of scavenging spread across five families of snakes and stated that this behaviour is ``far more common than currently acknowledged."\citep{devault2002scavenging}.
+The same reasoning can be applied to crocodlies and their allies \citep{forrest2003evidence}. 
+A sit and wait strategy is viable for an ectotherm. 
+This low existence cost is also realised in many sharks who have coupled low locomotory costs with an ectothermic metabolism. 
+The upshot is that 30 kg of blubber can sustain a White shark for over six weeks \citep{carey1982temperature}.
 
@@
 In contrast, birds of prey tend to kill prey smaller than themselves \citep{slagsvold2007prey}.
-This is likely due to their need to kill prey that they can fly away with, as well as the risk of injury being higher (which carries a higher mortality risk) for a bird than a mammal.
-Thus for birds, scavenging means they can exploit species that would otherwise be too big for them to kill.
+This is likely due to their need to kill animals that they can fly away with, as well as the risk of injury being higher (which carries a higher mortality risk) for a bird than a mammal.
+Scavenging provides a means for birds to exploit species that would otherwise be too big for them to kill.
+% Thus, for birds, scavenging means they can exploit species that would otherwise be too big for them to kill.
 
@@
 The competitive ability of even the largest bird is radically diminished in their interactions with mammalian competitors however, and as such they tend to consume carrion rapidly. 
-\cite{houston1974role} observed a group of \textit{Gyps} vultures consuming all of the soft tissue from a 50 kg Grant’s gazelle \textit{Nanger granti} in eight minutes.
-
-The first bird lineages did not have beaks and were predominantly carnivorous \citep{martyniuk2012field}. 
-This implies that, among the earliest species, scavenging would have been a live opportunity cf. their descendants who evolved beaks, restricting their ability to consume flesh.
+\cite{houston1974role} observed a group of \textit{Gyps} vultures consuming all of the soft tissue from a 50 kg Grant’s gazelle (\textit{Nanger granti}) in eight minutes. 
+Their serrated tongues and hooked bills enabling them to achieve this feat \citep{houston1975digestive}. 
+Outside of raptors like vultures the specialised beaks of many modern bird lineages hinders their ability to eat meat. 
+ \cite{martyniuk2012field} notes that the first bird lineages did not have beaks and were predominantly carnivorous. 
+This implies that, among the earliest species, scavenging would have been a live opportunity cf. their descendants.
 
@@
 Both the biotic and abiotic environment a would-be scavenger finds itself in can influence to degree to which it can depend on carrion. 
-As noted earlier, a system similar to the Serengeti in productivity could have supported a terrestrial scavenger \citep{ruxton2004obligate}.
-Indeed, in systems that were dominated by large ectothermic or mesothermic vertebrates 
+As we noted before, a system similar to the Serengeti in productivity could have supported a terrestrial scavenger \citep{ruxton2004obligate}.
+Indeed, in systems that were dominated by large ectothermic or mesothermic vertebrates, 
 the same primary productivity would have supported a greater biomass \citep{mcnab2009resources}.
 The upshot of this is there was a higher biomass of herbivores dying and offering scavenging opportunities.
-Predators were large-bodied too compared to extant mammalian predators \citep{mcnab2009resources}, and so, especially if they were ectothermic, could last longer between meals rendering scavenging a more attractive behaviour relative to predation.
-The absence of flying vertebrates in the Palaeozoic may have permitted terrestrial forms to take in a higher proportion of carrion in their diet.
-Osteophagy may have been even more viable during the Mesozoic era because the body mass distribution of herbviores tended to be skewed towards larger sizes \citep{10.1371/journal.pone.0051925}.
+Predators were large-bodied too compared to extant mammalian predators \citep{mcnab2009resources}, and so, especially if they were ectothermic, could last longer between meals, rendering scavenging a more attractive behaviour relative to predation.
+Osteophagy may have been even more viable during the Mesozoic era because of this skewed body mass distribution of herbviores towards larger sizes \citep{10.1371/journal.pone.0051925}.
 When we couple this with the fact that skeletal mass scales greater than linearly with body mass \citep{prange1979scaling} there would have been a lot of bone material to consume in the environment provided an animal had the biology to process it \citep{chure1997one}.
-As we discussed earlier, osteophagy could be extremely beneficial to a scavenger.
+As we discussed earlier, this ability is often extremely beneficial to a scavenger.
 
@@
 The upshot of this is that terrestrial carnivores like bears and wolves take more carrion \citep{devault2003scavenging}.
+Certainly, a major difficulty for terrestrial scavengers is competition with vultures.
+Noctural behaviour in the hyaenidae in general has been put forth as an adaptation to reduce competition with these exclusively diurnal birds \citep{gittleman2013carnivore}.
+If we apply this line of reasoning over evolutionary time-scales, the absence of flying vertebrates in the Palaeozoic may have permitted terrestrial forms to take in a higher proportion of carrion in their diet.
+
 The use of different sensory systems also illustrates the impact of the environment. 
-The relatively open savanna systems of Africa are well suited to a visually dependent vulture whereas more forested areas would select for species that have a well developed olfactory system  \citep{houston1986olfaction}.
-
-A major difficulty for terrestrial scavengers is competition with vultures.
-Noctural behaviour in the Hyaenidae in general has been put forth as an adaptation to reduce competition with these exclusively diurnal birds \citep{gittleman2013carnivore}.
-
+The relatively open savanna systems of Africa are well suited to a visually dependent vulture whereas more forested areas would select for species that have a well-developed olfactory system  \citep{houston1986olfaction}.
+Again, a similar line of reasoning can be applied to aquatic species depending whether they forage near the well-lit surface or the dark benthos. 
 % Although, as with detection ability, the environment has a role to play here.
 
-The presence of occasional bounties of carrion in the form of whale falls has led some researchers to investigate if a scavenger could survive by seeking out these remains exclusively.
+Staying in the aquatic setting, the phenomenon of occasional bounties of carrion in the form of whale falls has led some researchers to investigate if a scavenger could survive by seeking out these remains exclusively.
 \cite{ruxton2005searching} argued that although this is energetically feasible it's ecologically unlikely.
-Any animal that could seek out such whale carcasses is unlikely to have ignored other types of carrion.
+Any animal that could find such whale carcasses is unlikely to have ignored other types of carrion.
 Although no aquatic species have ever exceeded the size of whales, some enormous animals have evolved in this environment before the evolution of whales, including \textit{Leedsichthys}, a bony fish from the Middle Jurassic (174.1-163.5 Mya), that weighed in excess of 20 tonnes.
-Thus, the energetic feasiblity of a marine scavenger has a long history.
+Thus, the energetic feasiblity of a marine scavenger that specialises on large carcasses has a long history.
 One point of interest is that of the whaling industry, which provided a bonanza of floating carcasses especially during the 20th century \citep{Whitehead415}.
@@
 Early whales such as \textit{Basilosaurus} seem to fit into the same niche as Killer Whales and we have some evidence for scavenging in this group as well \citep{fahlke2012bite}.
+
+Perhaps the greatest environmental driver of scavenging tendency is that of temperature. 
+The geological record shows the Earth has undergone radical fluctuations in temperature.
+This will have had a significant bearing on the availability and persistence of carrion.
+To illustrate the point, a 10$^{\circ}$C increase in ambient temperature can double carcass decomposition rates \citep{parmenter2009carrion} and geological evidence indicates that the Mesozoic Earth was at least 6 $^{\circ}$C warmer than now \citep{sellwood2006mesozoic}.
+In terms of specific habitats, it has been shown that decomposition is greater in warm and moist areas versus more xeric ones \citep{beasley2015vertebrates}.
+Moreover, oceanic productivity and habitat structure are all impacted by climactic conditions.
+The impacts these can have on scavengers have been empirically supported e.g.
+\cite{beasley2015vertebrates} who point to a series of studies showing how microbes and invertebrates benefit at higher temperatures to the detriment of vertebrate scavengers such that ``above 20$^{\circ}$C vertebrates were able to detect and consume only 19\% of small-mammal carcasses, whereas at temperatures below 18$^{\circ}$C, vertebrates consumed 49\% of carcasses".
+
+%\section*{The Role of Climate} %TG: Use it as a opening at the end of the paper (i.e. conclusion + finishing with this bit)
+\section*{Conclusion} 
+As is often the case in science, the present provides the key to the past.
+The animals of today, while often different (sometimes radically so) to their ancestors, allow us to make informed comparisons to extinct species.
+We have used this technique to give insight into the drivers of scavenging across terrestrial vertebrates through time.
+In common with any other forager be they grazer, browser or predator, scavengers past and present have had to balance their energetic costs with the gains of food. 
+% Add bits on climate change/dectection rate change, etc...
+
+\section*{Acknowledgments}
+
+A lot of people are to thank here.
+
+\newpage
+
+\bibliography{bibfile}
+
+\end{document}
 
@@
 %Thus, in an ecological context, the evolution of these steps coupled with the ability to scavenge resulted in a world with a far more widely distributed nutrient landscape.
-
-%\section*{The Role of Climate} %TG: Use it as a opening at the end of the paper (i.e. conclusion + finishing with this bit)
-\section*{Conclusion} 
-The geological record shows the Earth has undergone radical fluctuations in temperature.
-This will have had a significant bearing on the availability and persistence of carrion.
-To illustrate the point, a 10$^{\circ}$C increase in ambient temperature can double carcass decomposition rates \citep{parmenter2009carrion} and geological evidence indicates that the Mesozoic Earth was at least 6 $^{\circ}$C warmer than now \citep{sellwood2006mesozoic}.
-In terms of specific habitats, it has been shown that decomposition is greater in warm and moist areas versus more xeric ones \citep{beasley2015vertebrates}.
-Moreover, oceanic productivity and habitat structure are all impacted by climactic conditions.
-The impacts these can have on scavengers have been empirically supported e.g.
-\cite{beasley2015vertebrates} who point to a series of studies showing how microbes and invertebrates benefit at higher temperatures to the detriment of vertebrate scavengers such that ``above 20$^{\circ}$C vertebrates were able to detect and consume only 19\% of small-mammal carcasses, whereas at temperatures below 18$^{\circ}$C, vertebrates consumed 49\% of carcasses".
-
-% Add bits on climate change/dectection rate change, etc...
 
@@
 %Possible scavenging on a juvenile fur seal from the Late Neogene \citep{boessenecker2011mammalian}.
-
-\section*{Acknowledgments}
-
-A lot of people are to thank here.
-
-\newpage
-
-\bibliography{bibfile}
-
-\end{document}

### 2016-03-31 natural history of scavenging.tex
stegosaurus figure

@@
 \begin{abstract}
- Scavengers existed in the past and they exist now. 
- Often under appreciated. 
- Three main habitat types considered: land, air and sea. 
- Different drivers in these areas. 
- Review looks at these 
+Despite its prevalence cavenging is a difficult behaviour to observe in modern day carnivores.
+Yet, there are certain intrinsic and environmental features of a speices that push it towards a scavenging lifestlye.
+Chief among these are low-cost locomotion, high detection distances, effective carrion processing and a carrion producing environment.
 \end{abstract}
@@
 \centering
-   \includegraphics[width=1\textwidth]{Summary_figure/Summary_figure.pdf}
+   \includegraphics[width=1\textwidth]{Summary_figure/Summary_figure_Landscape.pdf}
 \caption{Factors influencing the proportion of scavenging in a vertebrates' diet. Blue lines indicates a reduction in the factor and red lines indicates an increase.}

### 2016-03-31 pandoc.sh
Added panic shell routine (probably needs renaming the .tex file without space for safety)

@@ -1,0 +1,1 @@
+pandoc -o natural_history_of_scavenging.docx natural\ history\ of\ scavenging.tex --bibliography bibfile.bib --filter pandoc-citeproc --csl apa.csl

### 2016-03-31 natural history of scavenging.tex
abstract and conclusion

--- 2016-03-30 natural history of scavenging.tex
@@
 \begin{abstract} 
- Scavengers existed in the past and they exist now. 
- Often under appreciated. 
- Three main habitat types considered: land, air and sea. 
- Different drivers in these areas. 
- Review looks at these 
+% Here, we chart the natural history of scavenging. 
+Despite its prevalence, scavenging is a difficult behaviour to observe in modern day carnivores and impossible to study directly in extinct species. 
+Yet, there are certain intrinsic and environmental features of a species that push it towards a scavenging lifestyle.
+Chief among these are low-cost locomotion, high detection distances, effective carcass processing and a carrion-producing habitat.
+We use these four components to highlight the morphologies and environments that would have been conducive to scavenging over geological time by focussing on dominant vertebrate groups of the land, sea and air.
+The result is a document on the natural history of scavenging, the first to our knowledge. 
+Our idea of a scale of scavenging can be applied to any species at any time to judge the importance of this behaviour in its diet. 
 \end{abstract}
@@
 Thus, any animal existing as a scavenger must minimise its locomotory costs and maximise its detection capabilities \citep{ruxton2004obligate}.
-Once found, the scavenger has to process the carrion and overcome the agents of decay produced by the action of microorganisms on the carcass \citep{ruxton2014fruit}. 
+Once found, the scavenger has to process the carrion and overcome the agents of decay produced by the action of microorganisms on the carcass in order to take in a suffienct amount of energy \citep{ruxton2014fruit}.
 The habitat must also be productive enough to sustain an animal biomass that will eventually produce carcasses.
@@
 
-% \section*{Aerial Scavengers}
-% Vultures represent 
-%These birds consist of two convergent groups, from the old and the new world and represent the only example of obligate vertebrate scavengers today.
-%The families from which modern vultures arose, the Accipitridae and Cathartidae, appear during the Palaeocene (66 - 56 Mya) \citep{Jetz2012, Jarvis2014}.
-%Given their unique position, they have been extensively studied to determine what adaptations they possess that allows them to so flourish in this niche.
-%Underscoring their prowess are studies such as \cite{ogada2012effects} who note a three fold increase in carcass persistence times in the absence of vultures.
-%As such, we can begin by exploring the adapdations and the environments of vultures to draw comparisons with other scavenging species and \textit{their} environments.
-
 \subsection*{Locomotion}
@@
 This is exacerbated by the sprawling gait seen in many lizards which results in Carrier's Constraint such that the animal can't move and breathe at the same time because the lateral movements impedes its lungs \citep{carrier1987evolution}.
-This constraint manifests itself in aspects such as maximum sustainable speed where an equivalent mammal has a six to seven fold increase \citep{ruben1995evolution}.
+These constraints manifests themselves in aspects such as maximum sustainable speed where an equivalent mammal has a six to seven fold increase \citep{ruben1995evolution}.
 To quantify this effect with a simple example we can turn to some allometric relationships relating sustainable travelling speed to body mass.
 In the case of mammals and reptiles these are 1.15 $\times$ body mass (kg) \textsuperscript{0.12} and 0.23 $\times$ body mass (kg) \textsuperscript{0.12} respectively \citep{ruxton2004obligate}.
-We can insert these into a foraging radius model $\sfrac{\frac{\text{duration} \times \text{speed}}{2}}{1000}$ for a 12 hour foraging day which shows that while a 10 kg reptile can range 6.5 km an equally sized mammal can range nearly 33 km \citep{Enstipp2006Energetics}.
+If we insert these into a foraging radius model $\sfrac{\frac{\text{duration} \times \text{speed}}{2}}{1000}$ for a 12 hour foraging day which shows that while a 10 kg reptile can range 6.5 km an equally sized mammal can range nearly 33 km \citep{Enstipp2006Energetics}.
 For a foraging scavenger, this ability translates into a greater area searched for food.
@@
 But it is in the air that we find scavengers par excellence. 
-Flight is a cheaper means of locomotion than walking or running \citep{tucker1975energetic}. 
+Flight is a cheaper means of locomotion than either walking or running \citep{tucker1975energetic}. 
 We know that many extant birds exist as facultative scavengers because storks, raptors and corvids all take substantial quantities of carrion in their diet \citep{kendall2013alternative}.
@@
 The families from which modern vultures arose, the Accipitridae and Cathartidae, appear during the Palaeocene \citep[66 - 56 Mya; ][]{Jetz2012, Jarvis2014}.
-Yet, avian flight is far older than this and originates in the Late Jurassic (163.5-145 Mya), conincident with the fossils of \textit{Archaeopteryx lithographica} so many of these benefits would have been realised from that point on for carnivorous birds.
+Yet, avian flight is far older than this and originates in the Late Jurassic (163.5-145 Mya), conincident with the fossils of \textit{Archaeopteryx lithographica}, so many of these benefits would have been realised from that point on for carnivorous birds.
 And vertebrate flight is much older still where pterosaurs predate bird origins by a considerable margin in the Late Triassic (235-201.3 Mya).
@@
 It would be pointless to have incredible ranging abilites and not have the sensory architecture to benefit from it.
-If we came at this from a position of complete ignorance we would predict scavengers to have well-developed senses and indeed this is what we find. 
+If we came at this from a position of complete ignorance we would predict scavengers to have well-developed senses and indeed, this is what we find. 
 A simplification of terrestrial, vertebrate scavengers in sensory terms is one of them existing in a two-dimensional plane while foraging for carrion directly.
@@
 
-Aside from sight, three species of vultures within the new world family Cathartidae, (genus \textit{Cathartes}), have well developed olfactory systems \citep{AR:AR22815}.
+Aside from sight, many birds have well developed olfactory systems \citep{AR:AR22815} including three species of vultures within the new world family Cathartidae, (genus \textit{Cathartes}).
+% three species of vultures within the new world family Cathartidae, (genus \textit{Cathartes}), have well developed olfactory systems \citep{AR:AR22815}.
 Among them are the Turkey Vultures (\textit{Cathartes aura}) which were able to locate 90\% of baits set out in a tropical forest \citep{houston1986olfaction}.
 An atuned sense of smell is obviously useful in detecting decaying carrion from the air.
- %TG: this sentence makes totally sense but starting it by "Clearly" kind of undermines it.
-To me it sounds like when in a presentation some spawn of Satan puts up a table and starts saying: "As you can see".
 % This would be impossible for the visually reliant old world species.
@@
 % Depending on the species, a carcass in water either floats or descends to the sea floor \citep{Whitehead415}.
-Water tends to be a low-light environment, so visual detection distances are far lower (< 100 m) than they would be in the air.
-As such, animals detect resources through chemo- and mechanoreception more so than through vision \citep{ruxton2004energetic}.
+Aquatic species have to contend with the low-light environment of water where visual detection distances are far lower (< 100 m) than they would be in the air.
+As such, animals here detect resources through chemo- and mechanoreception more so than through vision \citep{ruxton2004energetic}.
 This is particularly relevant to extant sharks and aquatic snakes who are deemed as having the most suitable physiology for scavenging.
@@
 In some historical hominin-inhabited areas there were a greater number of felids than hyenids.
-This is significant because hyenas are likely to have left far less flesh on a carcass than a felid such as a sabretooth enabling contemperaneous hominins to benefit \citep{pobiner2015new}.
+Again, this is significant because hyenas are likely to have left far less flesh on a carcass than a felid such as a sabretooth enabling contemperaneous hominins to benefit \citep{pobiner2015new}.
 The intelligence, resultant tool-use and cooperative nature of hominins meant they could likely adapt to take on more or less carrion depending on their environment \citep{moleon2014humans}.
@@
 
-Scavenging should be particularly attractive to avian predators compared to mammals.
+Scavenging should be particularly attractive to avian predators compared to mammals. % AK should this section go above to where we mention avian scavengers first? 
 Solitary mammalian predators can kill prey up to the same body mass as themselves and sometimes an order of magnitude heavier (e.g. socially hunting lions \citep{owen2008predator}).
@@
 Perhaps the greatest environmental driver of scavenging tendency is that of temperature. 
-The geological record shows the Earth has undergone radical fluctuations in temperature.
+The geological record shows the Earth has undergone radical fluctuations in temperature over time.
 This will have had a significant bearing on the availability and persistence of carrion.
@@
 In common with any other forager be they grazer, browser or predator, scavengers past and present have had to balance their energetic costs with the gains of food.
+Our four main axes: locomotion, detection, processing and environment can be used to create a scale of scavenging whereupon any species can be placed in order to establish the importance of carrion in it diet.
 % Add bits on climate change/dectection rate change, etc...

### 2016-03-31 natural history of scavenging.tex
Scavenging bats YAY!

@@
 \title{A Recipe for Scavenging - the natural history of a behaviour}
+%TG: nice title! I'd replace the hyphen with a comma though
 \running{Scavenging in vertebrates}
@@
 
-\affiliations{
+\affiliations\{
 \item A. Kane (\url{adam.kane@ucc.ie}), University College Cork, Cooperage Building, School of Biological Earth and Environmental Sciences, Cork, Ireland.
@@
 Indeed, a single species of dinosaur notwithstanding, a synthesis describing the natural history of scavengers is absent from the literature.
+%TG: I don't understand the dinosaur part of this sentence. Might be my non-native english but I thought it's worth noting for other non-english folk out there (even though I know they aren't many!).
 Fortunately, research on scavenging is on the rise \citep{manga2006vulture}.
@@
 \subsection*{Locomotion}
-As endotherms, mammals can sustain long bouts of energetically expensive activity. 
+As endotherms, mammals can sustain long bouts%TG: not sure about "bouts" get it with the context and I know it makes the whole review a bit more literary and thus nicer to read but it might exclude non-native english readers.
+ of energetically expensive activity. 
 By contrast, modern reptiles are ectothermic, limiting their activity periods.
@@
 In the spotted hyena (\textit{Crocuta crocuta}), striped hyena (\textit{Hyaena hyaena}) and brown hyena (\textit{Hyaena brunnea}) it can be over 90\% \citep{jones2015african}.
-And although no contemporary terrestrial vertebrate exists as an obligate scavenger most if not all are facultative to some extent \citep{beasley2015vertebrates}.
+And although no contemporary terrestrial vertebrate exists as an obligate scavenger most (if not all) %TG: or maybe no brackets... Up to you Adam.
+ are facultative to some extent \citep{beasley2015vertebrates}.
 The particular reliance of hyenas on carrion means we can use them as examples of efficient terrestrial scavengers to compare with other forms. 
@@
 This allows us to make a broad distinction between the ambush strategies of cats and the pursuit/ pounce strategies of dogs, the latter being more suited to scavenging \citep{janis2014forelimb}. 
-We can (and have) use(d) these insights to compare extant terrestrial species to their prehistoric forebears given the dominance of mammalian carnviores since the Eocene 56-33.9 Million years ago (Mya) where the order split into the Caniforma and Feliforma \citep{van1987skeletal}.
-To take one example, \cite{anyonge1996locomotor} found \textit{Nimravides}, a genus of sabretooth cat from the Miocence (10.3 to 5.3 M ya) were likely to have been ambush predators which would have counted against them taking a lot of carrion.
+We can (and have) use(d) these insights to compare extant terrestrial species to their prehistoric forebears given the dominance of mammalian carnviores since the Eocene (56-33.9 Million years ago; Mya) where the order split into the Caniforma and Feliforma \citep{van1987skeletal}.
+To take one example, \cite{anyonge1996locomotor} found that \textit{Nimravides}, a genus of sabretooth cat from the Miocence (10.3 to 5.3 Mya) were likely to have been ambush predators which would have counted against them taking a lot of carrion.
 %The order Carnivora saw its origins in the Middle Eocene  %TG: a phylo side note: the correct sentence would be the "The stem Carnivora saw its origins ...
@@
 % So we can trace efficient terrestrial movement by carnivores from this point on.
+
+%TG: missing some transition here. Maybe something short like "Further back in time" or "Previous to that"...
 
@@
 Maybe change the sentence to "Of course the importance of terrestrial tetrapods in ecosystems predates the evolution of the dinosaurs."
-It is during the Permian (298.9 - 252.17 Mya) %, almost 300 millions years ago, %TG: Being pedantic here but Permian is just *not* 300 Mya...
-As a side note for the dates, I say we just use the geological epochs/periods/whatever since they are the actual precise dating of the fossil occurence (and they sound better).
-I suggest we just add the dates in brackets to give an indication.
-that we have the earliest evidence of vertebrate scavenging where a temnospondyl amphibian fed on the carcass of \textit{Varanops}, a predatory synapsid of the time \citep{reisz2006articulated}. 
+It is during the Permian (298.9 - 252.17 Mya) that we have the earliest evidence of vertebrate scavenging where a temnospondyl amphibian fed on the carcass of \textit{Varanops}, a predatory synapsid of the time \citep{reisz2006articulated}.
 And it is with the evolution of endothermy in the therapsid-mammal lineage \citep{clarke2010temperature} that terrestrial vertebrates would have gained the ability to range widely, a vital component in seeking out carrion.
@@
 %TG: I think it's missing a transition here: maybe something along "But moving on land as it's limitations..."
-But it is in the air that we find scavengers par excellence. 
+But it is in the air that we find scavengers \textit{par excellence}. %TG: not sure how reviewers will go with French but I like it!
 Flight is a cheaper means of locomotion than either walking or running \citep{tucker1975energetic}. 
 We know that many extant birds exist as facultative scavengers because storks, raptors and corvids all take substantial quantities of carrion in their diet \citep{kendall2013alternative}.
-The advantage of flight can be extended further in larger species that engage in soaring instead of flapping flight, which is even cheaper still (approximately twice the basal metabolic rate) \citep{hedenstrom1993migration,spivey2014analysing}.
+The advantage of flight can be extended further in larger species that engage in soaring instead of flapping flight, which is even cheaper still \citep[pproximately twice the basal metabolic rate; ][]{hedenstrom1993migration,spivey2014analysing}.
 The benefits this confers are clear from the information we have on the enormous foraging ranges of many vultures \citep{spiegel2013factors} and seabirds \citep{thaxter2012seabird}.
@@
 These birds consist of two convergent groups, from the old and the new world where they represent the only example of obligate vertebrate scavengers today.
+%TG: Or "obligate vertebrate scavengers known to date." For a more emphasizing turn.
+
+%TG: I thought the paragraph was a bit long so here's a cut for "flying scavengers in the past"
 The families from which modern vultures arose, the Accipitridae and Cathartidae, appear during the Palaeocene \citep[66 - 56 Mya; ][]{Jetz2012, Jarvis2014}.
@@
 The only other vertebrate group capable of powered flight are the bats where scavenging has not been recorded to our knowledge.
-The bat fossil record is notoriously poor owing to their fragile skeletons so we are unable to determine if extinct species were more suited to this lifestyle \citep{eiting2009global}.
-%TG: again, I'm not sure it's notoriously poor!
-It might be poor for mammalogist standards but it's definitely a way better than birds fossil records for example!
-Although it does not seem that flight is the main criterion precluding them from scavenging (see below).
+However, some fossils evidences as early as in the Eocene (56 to 33.9 Mya) suggest the presence of possible scavenging bats.
+A big bat from the middle to late Eocene (56 to 33.9 Mya) bat with a robust cranio-mandibular morphology, \textit{Necromatis} (``death-eater''), could have intake carrion as a decent proportion of its diet \citep{Weithofer_Necromantis_1887,Hand_Necromantis_2012}.
+However, this example might be peculiar to this exceptional fossil since most other fossils seems less suited to this lifestyle \citep{eiting2009global}. 
+%Although it does not seem that flight is the main criterion precluding them from scavenging (see below).
+%TG: Actually, maybe this example fits better in the food processing part (and maybe change the silhouettes in the picture then)?
+%``The name Necromantis (``death-eater'') alludes to the diet that Weithofer assumed for this bat – carrion, but probably also small, live vertebrates, including mammals such as rodents and other bats.'' from the paper \cite{Hand_Necromantis_2012}.
 
@@
 In fact, the cost of swimming is lower than either running or flying \citep{williams1999evolution}. 
-This has led some researchers to argue for the feasability of a scavenging fish \citep{ruxton2004energetic,ruxton2005searching}. 
-As with the aerial and terrestrial enviornments we have evidence of facultative scavenging among extinct aquatic species.
+This has led some researchers to argue for the feasibility of a scavenging fish \citep{ruxton2004energetic,ruxton2005searching}. 
+As with the aerial and terrestrial environments we have evidence of facultative scavenging among extinct aquatic species.
 For example, the remains of a mosasaur and a terrestrial hadrosaur were discovered with embedded teeth from a Cretaceous shark, \textit{Squalicorax} \citep{schwimmer1997scavenging}.

### 2016-04-01 natural history of scavenging.tex
seabird details

@@
 
-\affiliations\{
-\item A. Kane (\url{adam.kane@ucc.ie}), University College Cork, Cooperage Building, School of Biological Earth and Environmental Sciences, Cork, Ireland.
-\item K. Healy and A. Jackson, Trinity College Dublin, Department of Zoology, School of Natural Sciences, Dublin Ireland.
-\item T. Guillerme, Imperial College London, Silwood Park Campus, Department of Life Sciences, Buckhurst Road, Ascot SL5 7PY, United Kingdom.
-\item G. Ruxton, School of Biology, Sir Harold Mitchell Building, Greenside Place, St Andrews, KY16 9TH, United Kingdom.
-}
+%\affiliations\{
+%\item A. Kane (\url{adam.kane@ucc.ie}), University College Cork, Cooperage Building, School of Biological Earth and Environmental Sciences, Cork, Ireland.
+%\item K. Healy and A. Jackson, Trinity College Dublin, Department of Zoology, School of Natural Sciences, Dublin Ireland.
+%\item T. Guillerme, Imperial College London, Silwood Park Campus, Department of Life Sciences, Buckhurst Road, Ascot SL5 7PY, United Kingdom.
+%\item G. Ruxton, School of Biology, Sir Harold Mitchell Building, Greenside Place, St Andrews, KY16 9TH, United Kingdom.
+%}
 
@@
 
-\maketitle
+% \maketitle
 
@@
 
-Of course, tetrapod terrestrial dominance predates the evolution of the dinosaurs.
+Of course, the importance of terrestrial tetrapods predates the evolution of the dinosaurs.
 %TG: I will down that "dominance" term.
-Maybe change the sentence to "Of course the importance of terrestrial tetrapods in ecosystems predates the evolution of the dinosaurs."
+Maybe change the sentence to "Of course the importance of terrestrial tetrapods in ecosystems predates the evolution of the dinosaurs." AK Done
 It is during the Permian (298.9 - 252.17 Mya) that we have the earliest evidence of vertebrate scavenging where a temnospondyl amphibian fed on the carcass of \textit{Varanops}, a predatory synapsid of the time \citep{reisz2006articulated}.
@@
 The benefits this confers are clear from the information we have on the enormous foraging ranges of many vultures \citep{spiegel2013factors} and seabirds \citep{thaxter2012seabird}.
-In vultures we have the best known scavengers on Earth. 
-These birds consist of two convergent groups, from the old and the new world where they represent the only example of obligate vertebrate scavengers today.
+The difficulty inherent in observing marine birds foraging at sea means our knowledge of their feeding habits is limited. 
+But it does seem that albatrosses, who can range many 100s of kilometres, take a substantial amount of carrion in their diet \citep{croxall1994dead}.
+This is typically in the form of squid, which float on the surface, allowing the birds to readily pluck their remains out of the water \citep{croxall1994dead}. 
+In vultures, who are more easily observed on land, we have the best known scavengers on Earth. 
+These birds consist of two convergent groups, from the old and the new world where they represent the only example of obligate vertebrate scavengers to date.
 %TG: Or "obligate vertebrate scavengers known to date." For a more emphasizing turn.
@@
 The only other vertebrate group capable of powered flight are the bats where scavenging has not been recorded to our knowledge.
-However, some fossils evidences as early as in the Eocene (56 to 33.9 Mya) suggest the presence of possible scavenging bats.
-A big bat from the middle to late Eocene (56 to 33.9 Mya) bat with a robust cranio-mandibular morphology, \textit{Necromatis} (``death-eater''), could have intake carrion as a decent proportion of its diet \citep{Weithofer_Necromantis_1887,Hand_Necromantis_2012}.
-However, this example might be peculiar to this exceptional fossil since most other fossils seems less suited to this lifestyle \citep{eiting2009global}. 
-%Although it does not seem that flight is the main criterion precluding them from scavenging (see below).
-%TG: Actually, maybe this example fits better in the food processing part (and maybe change the silhouettes in the picture then)?
-%``The name Necromantis (``death-eater'') alludes to the diet that Weithofer assumed for this bat – carrion, but probably also small, live vertebrates, including mammals such as rodents and other bats.'' from the paper \cite{Hand_Necromantis_2012}.
+However, some fossils evidence as early as in the Eocene (56 to 33.9 Mya) suggest the presence of possible scavenging bats.
+A large bat from the middle to late Eocene (56 to 33.9 Mya) with a robust cranio-mandibular morphology, \textit{Necromatis} (``death-eater''), could have intake carrion as a significant proportion of its diet \citep{Weithofer_Necromantis_1887,Hand_Necromantis_2012}.
+However, this example might be peculiar to this exceptional fossil given most other fossils represent animals less suited to the lifestyle \citep{eiting2009global}.
 
@@
 Thus, we would expect scavengers to have this trait selected for even in the case of weight-constrained fliers.
-Cinereous Vultures (\textit{Aegypius monachus}) and condors (\textit{Vultur gryphus}, \textit{Gymnogyps californianus}) all have body masses that can exceed 10 kg and represent some of the heaviest bird species capable of flight \citep{ferguson2001raptors,donazar2002effects}.
+Wandering Albatrosses (\textit{Diomedea exulans}), Cinereous Vultures (\textit{Aegypius monachus}) and condors (\textit{Vultur gryphus}, \textit{Gymnogyps californianus}) all have body masses that can exceed 10 kg and represent some of the heaviest bird species capable of flight \citep{weimerskirch1992reproductive,ferguson2001raptors,donazar2002effects}.
 
@@
 Indeed, extant Marabou Storks (\textit{Leptoptilos crumenifer}) %TG: Check if that's the one
-have a comparable morphology and are noted facultative scavengers so it is reasonable to believe that certain pterosaurs behaved similarly.
+have a comparable morphology and are noted facultative scavengers \citep{monadjem2012survival} so it is reasonable to believe that certain pterosaurs behaved similarly.
 

### 2016-04-04 natural history of scavenging.tex
Andrew's Headings

@@
 Historically, scavengers have not been viewed as the most charismatic of animals.
-This may go some way to explaining the gap in our knowledge of the prevalence of this behaviour.
-Consider Professor Sanborn Tenney writing in 1877 for The American Naturalist who had this to say about one well known group, ``prominent among the mammalian scavengers are the hyenas, the ugliest in their general appearance of all the flesh eaters."
+This may go some way to explaining the gap in our knowledge of the prevalence of this behaviour \citep{devault2003scavenging}.
+Professor Sanborn Tenney writing in 1877 for The American Naturalist had this to say about one well known group, ``prominent among the mammalian scavengers are the hyenas, the ugliest in their general appearance of all the flesh eaters."
 He contrasts these with ``nobler kinds" of carnivores such as lions and tigers \citep{tenney1877few}.
@@
 The obvious lack of direct behavioural data compounds the difficulty of discerning scavenging among extinct forms.
-Indeed, a single species of dinosaur notwithstanding, a synthesis describing the natural history of scavengers is absent from the literature.
+Indeed, a single species of dinosaur notwithstanding \citep{carbone2011intra}, a synthesis describing the natural history of scavengers is absent from the literature.
 %TG: I don't understand the dinosaur part of this sentence. Might be my non-native english but I thought it's worth noting for other non-english folk out there (even though I know they aren't many!).
@@
 As a result, we are now beginning to realise the extent of this behaviour such that, ``in some ecosystems, vertebrates have been documented to assimilate as much as 90\% of the available carrion" \citep{beasley2015vertebrates}.
-Even Tenney’s noble big cats are now known to take in a significant portion of carrion in their diet where some lion populations get over 50\% of their meat from carcasses.
-By recognising the difficulty in directly observing scavenging, a suite of methods have been used to discern the most suitable morphologies, physiologies and environments for a scavenging lifestyle to prosper.
-Here we chart the natural history of scavenging by looking at the potential for the behaviour in dominant vertebrate groups.
+This has profound implications for the trophic ecology of these systems and particuarly our models of them.
+Even Tenney’s noble big cats are now known to take in a significant portion of carrion in their diet where some lion populations get over 50\% of their meat from carcasses \citep{jones2015african}.
+By recognising the difficulty in directly observing scavenging, other methods have been turned to to discern the most suitable morphologies, physiologies and environments for a scavenging lifestyle to prosper.
+Here we chart the natural history of scavenging by assessing the potential for the behaviour in dominant vertebrate groups given their ecology and functional traits.
 %It is our aim in this review to employ these methods to gain an understanding of scavengers past and present.
@@
 Once found, the scavenger has to process the carrion and overcome the agents of decay produced by the action of microorganisms on the carcass in order to take in a suffienct amount of energy \citep{ruxton2014fruit}.
-The habitat must also be productive enough to sustain an animal biomass that will eventually produce carcasses.
-We can draw on the image of a scavenger moving through its environment, searching for food and trying to process it efficiently as we explore the prevalence of this behaviour through time. 
+The habitat must also be productive enough to sustain a pool of animal biomass that will eventually produce carcasses.
+These facets are essentially the key parameters found in the species functional response curves of optimal foraging theory, namely encounter rate, handling time and prey availability \citep{jeschke2002predator}.
+Here, we can draw on these aspects to envisage a scavenger moving through its environment, searching for food and trying to process it efficiently as we explore the prevalence of scavenging through time.
+Our aim is to produce a 'scale of scavenging' whereupon we can place any species, past or present, and assess the importance of scavenging in its diet. 
 % The idea of scrounging from predator kills is undermined by studies showing that in the  majority of ecosystems more animals die from disease and starvation than predation \citep{benbow2015introduction}.
@@
 %TG: Insert figure here: scavenging scale + four points: Locomotion/Detection/Consumption/Environment
+\section*{Encounter Rate}
 
 \subsection*{Locomotion}
-As endotherms, mammals can sustain long bouts%TG: not sure about "bouts" get it with the context and I know it makes the whole review a bit more literary and thus nicer to read but it might exclude non-native english readers.
- of energetically expensive activity. 
-By contrast, modern reptiles are ectothermic, limiting their activity periods.
+As endotherms, mammals can sustain long bouts %TG: not sure about "bouts" get it with the context and I know it makes the whole review a bit more literary and thus nicer to read but it might exclude non-native english readers.
+ of energetically expensive activity \citep{bennett1979endothermy}. 
+By contrast, modern reptiles are ectothermic, limiting their activity periods \citep{bennett1979endothermy}.
 % a successful reptilian scavenger requires a far different set of adapations.
-This is exacerbated by the sprawling gait seen in many lizards which results in Carrier's Constraint such that the animal can't move and breathe at the same time because the lateral movements impedes its lungs \citep{carrier1987evolution}.
-These constraints manifests themselves in aspects such as maximum sustainable speed where an equivalent mammal has a six to seven fold increase \citep{ruben1995evolution}.
+This is exacerbated by the sprawling gait seen in many lizards which results in the phenomenon known as Carrier's Constraint such that the animal can't move and breathe at the same time because the lateral movements impedes its lungs \citep{carrier1987evolution}.
+These constraints manifest in aspects such as maximum sustainable speed where an equivalent mammal enjoys a six to seven fold increase \citep{ruben1995evolution}.
 To quantify this effect with a simple example we can turn to some allometric relationships relating sustainable travelling speed to body mass.
 In the case of mammals and reptiles these are 1.15 $\times$ body mass (kg) \textsuperscript{0.12} and 0.23 $\times$ body mass (kg) \textsuperscript{0.12} respectively \citep{ruxton2004obligate}.
-If we insert these into a foraging radius model $\sfrac{\frac{\text{duration} \times \text{speed}}{2}}{1000}$ for a 12 hour foraging day which shows that while a 10 kg reptile can range 6.5 km an equally sized mammal can range nearly 33 km \citep{Enstipp2006Energetics}.
+If we insert these into a foraging radius model $\sfrac{\frac{\text{duration} \times \text{speed}}{2}}{1000}$ for a 12 hour foraging day which shows that while a 10 kg reptile can range 6.5 km, an equally sized mammal can range nearly 33 km \citep{Enstipp2006Energetics}.
 For a foraging scavenger, this ability translates into a greater area searched for food.
@@
 In the spotted hyena (\textit{Crocuta crocuta}), striped hyena (\textit{Hyaena hyaena}) and brown hyena (\textit{Hyaena brunnea}) it can be over 90\% \citep{jones2015african}.
-And although no contemporary terrestrial vertebrate exists as an obligate scavenger most (if not all) %TG: or maybe no brackets... Up to you Adam.
+And although no contemporary terrestrial vertebrate exists as an obligate scavenger, most, if not all, %TG: or maybe no brackets... Up to you Adam.
  are facultative to some extent \citep{beasley2015vertebrates}.
@@
 We can (and have) use(d) these insights to compare extant terrestrial species to their prehistoric forebears given the dominance of mammalian carnviores since the Eocene (56-33.9 Million years ago; Mya) where the order split into the Caniforma and Feliforma \citep{van1987skeletal}.
-To take one example, \cite{anyonge1996locomotor} found that \textit{Nimravides}, a genus of sabretooth cat from the Miocence (10.3 to 5.3 Mya) were likely to have been ambush predators which would have counted against them taking a lot of carrion.
+To take one example, \cite{anyonge1996locomotor} found that \textit{Nimravides}, a genus of sabretooth cat from the Miocence (10.3 to 5.3 Mya) were likely to have been ambush predators which would argue against them taking a lot of carrion.
 %The order Carnivora saw its origins in the Middle Eocene  %TG: a phylo side note: the correct sentence would be the "The stem Carnivora saw its origins ...
@@
 Unsurprisingly, given their enduring appeal, the prevalence of scavenging has been explored in the carnivorous, theropod dinosaurs.
-They were the dominant terrestrial forms for most of the Mesozoic Era (252.17 - 66 Mya) and ranged from the chicken-sized to the whale-sized, all of which were bipedal.
+They were the dominant terrestrial carnivores for most of the Mesozoic Era (252.17 - 66 Mya) and ranged from the chicken-sized to the whale-sized, all of which were bipedal.
 They are quite alien to anything we know today which restricts our ability to understand their ecology far more so than extinct mammals \citep{weishampel2004dinosauria}.
-Of relevance, are the questions that still persist about their metabolism with the latest evidence suggesting they were mesothermic i.e.
+Of relevance, are the questions that still persist about their metabolism, with the latest evidence suggesting they were mesothermic i.e.
 intermediate to ecto- and endotherms \citep{grady2014evidence}.
@@
 It is during the Permian (298.9 - 252.17 Mya) that we have the earliest evidence of vertebrate scavenging where a temnospondyl amphibian fed on the carcass of \textit{Varanops}, a predatory synapsid of the time \citep{reisz2006articulated}.
-And it is with the evolution of endothermy in the therapsid-mammal lineage \citep{clarke2010temperature} that terrestrial vertebrates would have gained the ability to range widely, a vital component in seeking out carrion.
+And it is with the evolution of endothermy in the therapsid-mammal lineage \citep{clarke2010temperature} that terrestrial vertebrates would have gained the ability to range more widely, a vital component in seeking out carrion.
 
@@
 But it is in the air that we find scavengers \textit{par excellence}. %TG: not sure how reviewers will go with French but I like it!
+Certainly, scavenging should be particularly attractive to avian predators compared to mammals. % Not sure about this paragraph 
+The latter can kill prey up to the same body mass as themselves and sometimes an order of magnitude heavier \citep[e.g. socially hunting lions; ][]{owen2008predator}.
+In contrast, birds of prey tend to kill prey smaller than themselves \citep{slagsvold2007prey} because of the greater cost of injury and need to carry off their food.
+% This is likely due to their need to kill animals that they can fly away with, as well as the risk of injury being higher (which carries a higher mortality risk) for a bird than a mammal.
+Scavenging provides a means for birds to exploit species that would otherwise be too big for them to kill.
+% Thus, for birds, scavenging means they can exploit species that would otherwise be too big for them to kill.
+
 Flight is a cheaper means of locomotion than either walking or running \citep{tucker1975energetic}. 
 We know that many extant birds exist as facultative scavengers because storks, raptors and corvids all take substantial quantities of carrion in their diet \citep{kendall2013alternative}.
-The advantage of flight can be extended further in larger species that engage in soaring instead of flapping flight, which is even cheaper still \citep[pproximately twice the basal metabolic rate; ][]{hedenstrom1993migration,spivey2014analysing}.
+The advantage of flight can be extended further in larger species that engage in soaring instead of flapping flight, which is even cheaper still for searching the environment \citep[approximately twice the basal metabolic rate; ][]{hedenstrom1993migration,spivey2014analysing}.
 The benefits this confers are clear from the information we have on the enormous foraging ranges of many vultures \citep{spiegel2013factors} and seabirds \citep{thaxter2012seabird}.
 The difficulty inherent in observing marine birds foraging at sea means our knowledge of their feeding habits is limited. 
-But it does seem that albatrosses, who can range many 100s of kilometres, take a substantial amount of carrion in their diet \citep{croxall1994dead}.
+But it does seem that albatrosses, who can range many hundreds of kilometres, take a substantial amount of carrion in their diet \citep{croxall1994dead}.
 This is typically in the form of squid, which float on the surface, allowing the birds to readily pluck their remains out of the water \citep{croxall1994dead}. 
@@
 %TG: I thought the paragraph was a bit long so here's a cut for "flying scavengers in the past"
-The families from which modern vultures arose, the Accipitridae and Cathartidae, appear during the Palaeocene \citep[66 - 56 Mya; ][]{Jetz2012, Jarvis2014}.
-Yet, avian flight is far older than this and originates in the Late Jurassic (163.5-145 Mya), conincident with the fossils of \textit{Archaeopteryx lithographica}, so many of these benefits would have been realised from that point on for carnivorous birds.
-And vertebrate flight is much older still where pterosaurs predate bird origins by a considerable margin in the Late Triassic (235-201.3 Mya).
+The two families from which modern vultures arose, the Accipitridae and Cathartidae, appear during the Palaeocene \citep[66 - 56 Mya; ][]{Jetz2012, Jarvis2014}.
+Yet, avian flight is far older than this and originates in the Late Jurassic (163.5-145 Mya), conincident with the fossils of \textit{Archaeopteryx lithographica}, so many of the benefits of flight would have been realised from that point on for carnivorous birds.
+Vertebrate flight is much older still, where pterosaurs predate bird origins by a considerable margin in the Late Triassic (235-201.3 Mya).
 Scavenging in this diverse group has been hypothesied many times before \citep{witton2008reappraisal}.
 Certain clades of these animals could reach enormous sizes \citep[e.g.
-Azhdarchids with wingspans of 11 metres; ][]{witton2010size} and, notably, look to have engaged in soaring flight \citep{witton2010size}.
+Azhdarchids with wingspans of 11 metres; ][]{witton2010size} and, notably, appear to have engaged in soaring flight \citep{witton2010size}.
 %TG: this paragraph's a killer! Love it!
@@
 The only other vertebrate group capable of powered flight are the bats where scavenging has not been recorded to our knowledge.
-However, some fossils evidence as early as in the Eocene (56 to 33.9 Mya) suggest the presence of possible scavenging bats.
+However, some fossil from the Eocene (56 to 33.9 Mya) suggest the presence of possible scavenging bats.
 A large bat from the middle to late Eocene (56 to 33.9 Mya) with a robust cranio-mandibular morphology, \textit{Necromatis} (``death-eater''), could have intake carrion as a significant proportion of its diet \citep{Weithofer_Necromantis_1887,Hand_Necromantis_2012}.
-However, this example might be peculiar to this exceptional fossil given most other fossils represent animals less suited to the lifestyle \citep{eiting2009global}.
+Although, this example might be peculiar to this exceptional fossil given most other remains represent animals less suited to the lifestyle \citep{eiting2009global}.
 
-Aquatic scavengers have a locomotory benefit because water is a medium that is conducive to low-cost movement \citep{tucker1975energetic}. 
+Similar to aerial species, aquatic scavengers have a locomotory benefit because water is a medium that is conducive to low-cost movement \citep{tucker1975energetic}. 
 In fact, the cost of swimming is lower than either running or flying \citep{williams1999evolution}. 
-This has led some researchers to argue for the feasibility of a scavenging fish \citep{ruxton2004energetic,ruxton2005searching}. 
-As with the aerial and terrestrial environments we have evidence of facultative scavenging among extinct aquatic species.
-For example, the remains of a mosasaur and a terrestrial hadrosaur were discovered with embedded teeth from a Cretaceous shark, \textit{Squalicorax} \citep{schwimmer1997scavenging}.
-As well as a likely instance of scavenging between a 4-million-year-old white shark (\textit{Carcharodon}) and mysticete whale from Peru \citep{ehret2009caught}.
-Extant White sharks \textit{Carcharodon carcharias} too are known to feed on whale carcasses \citep{fallows2013white}.
+This has led some researchers to argue for the feasibility of an obligate scavenging fish \citep{ruxton2004energetic,ruxton2005searching}. 
+As in aerial and terrestrial environments we have evidence of facultative scavenging among extinct aquatic species.
+For example, shark teeth have been found in the remains of dinosaurs, mososaurs and 4-million-year-old mysticete whales \citep{schwimmer1997scavenging,ehret2009caught}.  
+%the remains of a mosasaur and a terrestrial hadrosaur were discovered with embedded teeth from a Cretaceous shark, \textit{Squalicorax} \citep{schwimmer1997scavenging}.
+%As well as a likely instance of scavenging between a 4-million-year-old white shark (\textit{Carcharodon}) and  whale from Peru \citep{ehret2009caught}.
+Extant White sharks (\textit{Carcharodon carcharias}) too are known to feed on whale carcasses \citep{fallows2013white}.
+
 We might expect then that by combining an aquatic environment and an endothermic metabolism that marine mammals would prosper as scavengers. 
-We know fossil pinnipeds and cetaceans from 60 Mya have transitional features indicative of their trajectory to fully aquatic species \citep{williams1999evolution}.  
-But despite this movement away from land the energetic savings were negligible because the \textit{total} cost incurred by a swimming marine mammal is high \citep{williams1999evolution}. 
-This is not to say that aquatic mammalian scavengers don't exist, only that their total energetic cost is similar to an equivalent terrestrial mammal. 
+Fossil pinnipeds and cetaceans from 60 Mya have transitional features indicative of their trajectory to fully aquatic species \citep{williams1999evolution}.  
+But despite their movement away from land their energetic savings were negligible because the \textit{total} cost incurred by a swimming marine mammal is high \citep{williams1999evolution}. 
+Indeed, their total energetic cost is similar to an equivalent terrestrial mammal \citep{williams1999evolution}.
+This underscores the trade offs between the benefits of endothermy in terms of activity periods and the costs of maintaining such an expensive system.
+
 %\cite{williams1999evolution} notes, ``free-ranging animals must contend with the total energetic expenditure associated with supporting basic biological functions as well as with moving the body and appendages through the environment."
@@
 %could resolve a 2 metre target at 1 km distance. 
-While considering prehistoric habitats \cite{ruxton2004obligate} calculated that ``a 1 tonne mammal or reptile, in an ecosystem yielding carrion at densities similar to the current Serengeti, could have met its energy requirements if it could detect carrion over a distance of the order of 400–500 m".
-The senses of many extant (and in all probabilty extinct) carnivores meet this required distance, making scavenging feasbile for terrstrial species \citep{farlow1994speculations,mech2010wolves}.
+If we compare this to the energetics approach by \cite{ruxton2004obligate}, they calculated a terrestrial scavenger needs to be able to detect carrion at 500 m in order to survive, which is clearly within the ability of hyenas.
+% While considering prehistoric habitats \cite{ruxton2004obligate} calculated that ``a 1 tonne mammal or reptile, in an ecosystem yielding carrion at densities similar to the current Serengeti, could have met its energy requirements if it could detect carrion over a distance of the order of 400–500 m".
+Moreover, the senses of many extant (and in all probabilty extinct) carnivores meet this required distance, making scavenging feasible for terrestrial species \citep{farlow1994speculations,mech2010wolves}.
 
@@
 It follows from this that the evolution of flight allowed aerial animals to detect far more carrion than their terrestrial counterparts \citep{AR:AR22815}.
+We can contrast this ability with bats, whose visual acuity is famously poor. 
+It also appears that echolocation would not lend itself to discovering immobile carrion.
+Their small size and poor terrestrial ability would also count against them at a carcass \citep{riskin2006terrestrial}.
 
-Moreover, having a panoramic view means being able to gather a wealth of information from other foragers, be they conspecifics or other species \citep{jackson2008effect}.
+Having a panoramic view means being able to gather a wealth of information from other foragers, be they conspecifics or otherwise \citep{jackson2008effect}.
 Again, returning to vultures, the genus \textit{Gyps} consists of highly social and colonially nesting species \citep{fernandez2015density}.
 These behaviours allow them forage far more efficiently because one bird can scrounge information on the location of food from another successful forager \citep{KaneVul}.
-
-We can contrast this ability to bats, whose visual acuity is famously poor. 
-It also appears that echolocation would not lend itself to discovering immobile carrion.
-Their small size and poor terrestrial ability would also count against them at a carcass \citep{riskin2006terrestrial}.
 
@@
 % Depending on the species, a carcass in water either floats or descends to the sea floor \citep{Whitehead415}.
-Aquatic species have to contend with the low-light environment of water where visual detection distances are far lower (< 100 m) than they would be in the air.
+In contrast to the air, aquatic species have to contend with the low-light environment of water where visual detection distances are far lower (< 100 m) than they would be in the air.
 As such, animals here detect resources through chemo- and mechanoreception more so than through vision \citep{ruxton2004energetic}.
@@
 Smell seems to be the primary means of carcass detection in sharks as well. 
-\cite{fallows2013white} found that wind speed determined the number of sharks feeding at whale carcasses indicating they were dependent on detecting the odours from the decaying whales.
+\cite{fallows2013white} found that wind speed determined the number of sharks feeding at whale carcasses, indicating they were dependent on detecting the odours from the decaying whales.
 
-\subsection*{Processing}
+\section*{Handling Time}
 Since carrion is not dispatched directly, often the most easily accessible and choicest components of the carcass will be missing or, if present, will be fought over.
 Being able to extract nutrients from remnants gives a scavenger a great advantage.
-Thus, the bone crushing ability of hyenas reveals another useful scavenger trait.
+Thus, the bone crushing ability of hyenas and others reveals another useful scavenger trait.
 Osteophagy is known across a range of terrestrial carnivores and given some fat-rich mammalian bones have an energy density (6.7 kJ/g) comparable with that of muscle tissue, it makes skeletal remains an enticing resource \citep{brown1989study}.
 This ability reached its zenith among hyenas with the evolution of the 110 kg \textit{Pachycrocuta brevirostris} during the Pliocene \citep[3.6 - 2.58 Mya; ][]{palmqvist2011giant}.
-Some work on extinct sabretooths suggests they may have left a large amount of food for would-be scavengers because of their unique skull morphology.
-As a result, the decline of Machairodontinae sabretooths has been offered as an explanation for the extinction of \textit{P brevirostris} \citep{palmqvist2011giant}.
+Indeed, their extinction has been blamed on the decline of sabretooth cats (Machairodontinae), the unique skull morphology of the latter meant they would leave a large amount of food on a carcass for would-be scavengers \citep{palmqvist2011giant}.
+Earlier in the evolution of mammals, the bone-crushing dogs that evolved during the Oligocene (subfamily Borophaginae; 33.9 - 23.03 Mya) have been compared to hyenas in terms of their feeding ecology as well \citep{van2003chapter,martin2016pursuit}.
+
+% Some work on extinct sabretooths suggests they may have left a large amount of food for would-be scavengers because of their unique skull morphology.
+% As a result, the decline of Machairodontinae sabretooths has been offered as an explanation for the extinction of \textit{P brevirostris} \citep{palmqvist2011giant}.
 % And many of the aforesaid adapations for scavenging are found in these other major terrestrial mammalian carnviores.
@@
 % These traits have been used to infer scavenging lifestyles in extinct mammals.
-The bone-crushing dogs that evolved during the Oligocene (subfamily Borophaginae; 33.9 - 23.03 Mya) have been compared to hyenas in terms of their feeding ecology as well \citep{van2003chapter,martin2016pursuit}.
 
 Interestingly, such comparisons have given insight into the feeding ecology of early hominins who, for instance, had the ability to craft tools for breaking open bones \citep{ARCM:ARCM12084}.
-The question of where our ancestors placed on the hunter-scavenger axis during the Plio-Pleistocene has been a matter of debate for years.
+The question of where our ancestors placed on the hunter-scavenger axis during the Plio-Pleistocene has been a matter of debate for years \citep{dominguez2002hunting}.
 A recent study investigating potential scavenging opportunities for hominins in Kenya found that, even when discounting bone material, there is a substantial amount of scavengeable meat left on predated remains; sufficient to sustain the requirements of an adult male \textit{Homo erectus} \citep{pobiner2015new}.
@@
 In addition to reducing locomotory costs we would expect adaptations that reduce energetic costs of maintenance to be selected for in scavengers because it would maximise the benefit derived from such a sporadic food source.
-Extant reptiles possess an advantage here, in that over the course of a year their food requirements can be 30 times smaller than an endotherm of equal size \citep{Nagy1621}.
+Extant reptiles possess an advantage here, in that over the course of a year their food requirements can be 30 times lower than an endotherm of equal size \citep{Nagy1621}.
 \cite{devault2002scavenging} suggest this is an avenue for scavenging in snakes because they ``exhibit  exceedingly  low  maintenance  metabolisms,  and most  can  survive  on  a  few  scant  feedings per year.
@@
 
-Scavenging should be particularly attractive to avian predators compared to mammals. % AK should this section go above to where we mention avian scavengers first? 
-Solitary mammalian predators can kill prey up to the same body mass as themselves and sometimes an order of magnitude heavier (e.g. socially hunting lions \citep{owen2008predator}).
-In contrast, birds of prey tend to kill prey smaller than themselves \citep{slagsvold2007prey}.
-This is likely due to their need to kill animals that they can fly away with, as well as the risk of injury being higher (which carries a higher mortality risk) for a bird than a mammal.
-Scavenging provides a means for birds to exploit species that would otherwise be too big for them to kill.
-% Thus, for birds, scavenging means they can exploit species that would otherwise be too big for them to kill.
-
 Large body size confers substantial dominance and starvation-resistance benefits \citep{ruxton2004obligate}.
@@
 Although \cite{witton2008reappraisal} argued that neck inflexibility and straight, rather than hooked jaw morphology points against pterosaurs existing as \textit{obligate} scavengers, Azhdarchid terrestrial proficency indicates they would have been comfortable foraging on the ground.
-Indeed, extant Marabou Storks (\textit{Leptoptilos crumenifer}) %TG: Check if that's the one
+Indeed, extant Marabou Storks (\textit{Leptoptilos crumenifer}) 
 have a comparable morphology and are noted facultative scavengers \citep{monadjem2012survival} so it is reasonable to believe that certain pterosaurs behaved similarly.
@@
 Their serrated tongues and hooked bills enabling them to achieve this feat \citep{houston1975digestive}. 
-Outside of raptors like vultures the specialised beaks of many modern bird lineages hinders their ability to eat meat. 
- \cite{martyniuk2012field} notes that the first bird lineages did not have beaks and were predominantly carnivorous. 
-This implies that, among the earliest species, scavenging would have been a live opportunity cf. their descendants.
+Outside of raptors such as vultures, the specialised beaks of many modern bird lineages hinders their ability to eat meat which is in contrast to the first lineages that did not have this feature \citep{martyniuk2012field} .
+As \cite{martyniuk2012field} notes these early birds would thus have been predominantly carnivorous, which implies that scavenging would have been a live opportunity cf. their descendants.
 
-\cite{shivik2006vultures} points out that ``evolutionary pressures favor detection maximizers relative to toxification minimizers in competitive interactions for carcasses." 
-But the fact remains that overcoming microorganism toxins is still a beneficial adaptation to any scavenger. 
+Although the observation of \cite{shivik2006vultures} that ``evolutionary pressures favor detection maximizers relative to toxification minimizers in competitive interactions for carcasses." appears sound, the fact remains that overcoming microorganism toxins is still a beneficial adaptation to any scavenger.
 Avian scavengers have evolved incredibly acidic stomachs that allow them to consume and process putrefied flesh with no ill effects \citep{houston1975digestive,roggenbuck2014microbiome}. 
-This adapation is not restricted to vultures though, \cite{gremillet2012vultures} showed wandering albatrosses (\textit{Diomedea exulans} so-called 'vultures of the seas') had an average pH of 1.5, which enables them to consume fisheries discards.
-Outside of the birds there is evidence of selection for 'toxification minimizers'.
+This adapation is not restricted to vultures though, \cite{gremillet2012vultures} showed wandering albatrosses (\textit{Diomedea exulans}; so-called 'vultures of the seas') had an average pH of 1.5, which enables them to consume fisheries discards and squid carcasses.
+There is also evidence of selection for 'toxification minimizers' beyond birds among the ectotherms.
 From our earlier arguments we know that ecthotherms are limited in their ability to find carrion as quickly as endotherms. 
@@
 Big bats (which are better suited for scavenging, following our previous argument) are typically frugivores and therefore lack the adaptations for digesting  meat.
-While carnivorous bats are mainly found in the microbats which are insectivorous.
+While carnivorous bats are mainly found in the microbats which are insectivorous \citep{aguirre2003implications}.
 % Maybe they scavenge on dead insects but here we're talking about scavenging on real meat right?
@@
 
-\subsection*{Environment}
+\section*{Prey Availability}
 Both the biotic and abiotic environment a would-be scavenger finds itself in can influence to degree to which it can depend on carrion. 
-As we noted before, a system similar to the Serengeti in productivity could have supported a terrestrial scavenger \citep{ruxton2004obligate}.
+\cite{ruxton2004obligate} suggest a system similar to the Serengeti in productivity could have supported a mammalian or reptilian terrestrial scavenger \citep{ruxton2004obligate}.
 Indeed, in systems that were dominated by large ectothermic or mesothermic vertebrates, 
@@
 
-The use of different sensory systems also illustrates the impact of the environment. 
-The relatively open savanna systems of Africa are well suited to a visually dependent vulture whereas more forested areas would select for species that have a well-developed olfactory system  \citep{houston1986olfaction}.
-Again, a similar line of reasoning can be applied to aquatic species depending whether they forage near the well-lit surface or the dark benthos. 
+% The use of different sensory systems also illustrates the impact of the environment. 
+% The relatively open savanna systems of Africa are well suited to a visually dependent vulture whereas more forested areas would select for species that have a well-developed olfactory system  \citep{houston1986olfaction}.
+% Again, a similar line of reasoning can be applied to aquatic species depending whether they forage near the well-lit surface or the dark benthos. 
 % Although, as with detection ability, the environment has a role to play here.
@@
 As is often the case in science, the present provides the key to the past.
-The animals of today, while often different (sometimes radically so) to their ancestors, allow us to make informed comparisons to extinct species.
+The animals of today, while often different (sometimes radically so) to their ancestors, can be used to make informed comparisons to extinct species.
 We have used this technique to give insight into the drivers of scavenging across terrestrial vertebrates through time.
 In common with any other forager be they grazer, browser or predator, scavengers past and present have had to balance their energetic costs with the gains of food.
-Our four main axes: locomotion, detection, processing and environment can be used to create a scale of scavenging whereupon any species can be placed in order to establish the importance of carrion in it diet.
+The main factors we considered namely, encounter rate, handling time and prey availability can be used to create a scale of scavenging whereupon any species can be placed in order to establish the importance of carrion in it diet.
+We hope this approach will be useful in the effort to explore this most understudied of feeding ecologies. 
 % Add bits on climate change/dectection rate change, etc...

### 2016-04-05 natural history of scavenging KH edits.tex
Included extra doc KH edit with lots of changes

--- 2016-04-04 natural history of scavenging.tex
@@
 
+%the intro needs to lead towards a aim a bit more strongly
+
 \subsection*{The Difficulty of Scavenging} %KH: Make that the template for all the next sections
-The chief hurdle to scavenging is finding a sufficient quantity of food, the occurence of which is difficult to predict in space and time.
-Thus, any animal existing as a scavenger must minimise its locomotory costs and maximise its detection capabilities \citep{ruxton2004obligate}.
-Once found, the scavenger has to process the carrion and overcome the agents of decay produced by the action of microorganisms on the carcass in order to take in a suffienct amount of energy \citep{ruxton2014fruit}.
-The habitat must also be productive enough to sustain a pool of animal biomass that will eventually produce carcasses.
-These facets are essentially the key parameters found in the species functional response curves of optimal foraging theory, namely encounter rate, handling time and prey availability \citep{jeschke2002predator}.
-Here, we can draw on these aspects to envisage a scavenger moving through its environment, searching for food and trying to process it efficiently as we explore the prevalence of scavenging through time.
-Our aim is to produce a 'scale of scavenging' whereupon we can place any species, past or present, and assess the importance of scavenging in its diet. 
-% The idea of scrounging from predator kills is undermined by studies showing that in the  majority of ecosystems more animals die from disease and starvation than predation \citep{benbow2015introduction}.
+
+The chief hurdle of scavenging is finding a resource that is often difficult to predict in space and time.
+Though chance alone many species will avail of some level of opportunistic scavenging opportunities.
+However, species that rely on scavenging to sustain substantial portions of their diets must increase the probability of encountering a sufficient amount of carrion in order to meet some energetic demands.
+Once found, the scavenger must be able to out-compete any potential competitors and process a product under the processes of decay from toxin producing and disease causing microorganisms \citep{ruxton2014fruit}.
+Finally, the level of scavenging opportunities available will also depend on the density, size, and quality of carcases produced, all of which are dependent on complex ecosystem dynamics.
+These facets which are required for scavenging are essentially the key parameters found in the species functional response curves of optimal foraging theory, namely encounter rate, handling time and prey availability \citep{jeschke2002predator}.
+We compare the ability of scavenging across the vertabrates based on each of these facets and their relationship with scavenging.
+By framing scavenging in a framework of optimal foraging and identify the perquisite attributes and processes required for scavenging we produce a 'scale of scavenging' whereupon we can place any species, past or present, and assess the importance of scavenging in its diet.
 
@@
 \section*{Encounter Rate}
+All foraging processes depend on encounter rates between the consumer and the resource.
+In the simplest case, encounter rates can be thought of in terms of gas diffusion models were the movement of two agents depends on only on the relative speed of the agents.
+As carcasses are prone to being relatively stationary, the relative speed between a scavenger and carrion is dependent on the movement of the scavenger.
+As most vertebrates depended on more than simply bumping into resources like gas molecules and can actively detect resources through sensory abilities, scavenging will also depend on search rates which are dependent on both species physiology and the dimensionality of the environment (You guessed it Pawar et al 2012).
 
 \subsection*{Locomotion}
-As endotherms, mammals can sustain long bouts %TG: not sure about "bouts" get it with the context and I know it makes the whole review a bit more literary and thus nicer to read but it might exclude non-native english readers.
- of energetically expensive activity \citep{bennett1979endothermy}. 
-By contrast, modern reptiles are ectothermic, limiting their activity periods \citep{bennett1979endothermy}.
-% a successful reptilian scavenger requires a far different set of adapations.
+In contrast to the importance of locomotion in predation, locomotion relating to scavenging depends more on the ability to efficiency move over large areas.
+This ability generally requires an efficient transfer of metabolic energy into movement which may rely on both physiology (i.e.
+metabolism) and the medium of the environment in which they are moving (aerial, aquatic or terrestrial).
+
+Perhaps the most efficient form of locomotion in vertebrates is paradoxically found in flying species.
+Despite the energetic costs of flight the only known vertebrate obligate scavengers are the old and the new world vultures.
+While powered flight is energetically expensive (ref) it is the ability of these species to exploit air currents using large wingspans that allows them to soar while only approximately doubling their metabolic rate \citep{hedenstrom1993migration,spivey2014analysing}.
+By depending on thermal air flows these species can forage over enormous ranges \citep{spiegel2013factors}.
+This mode of locomotion is also exploited by sea birds, which use strong ocean winds to search large areas of the oceans \citep{thaxter2012seabird}.
+While many species of seabird are likely primarily predators, it seems that albatross, who can range many hundreds of kilometres, take a substantial amount of carrion in their diet \citep{croxall1994dead}.
+This is typically in the form of squid, which float on the surface, allowing the birds to readily pluck their remains out of the water \citep{croxall1994dead}.
+
+The use of soaring in extinct species as a means for scavenging is also likely.
+The two families from which modern vultures arose, the Accipitridae and Cathartidae, appear during the Palaeocene \citep[66 - 56 Mya; ][]{Jetz2012, Jarvis2014} and are  likley to have being soaring flyer's as demonstrated by their large wingspans (Is this true?).
+(Is there any other form of soaring?).
+However, soaring flight is likely to be far older than these groups with avian flight originating in the Late Jurassic (163.5-145 Mya) and pterosaurs in the Late Triassic (235-201.3 Mya).
+Scavenging in the diverse pterosaurs group has been hypothesied many times before \citep{witton2008reappraisal}.
+Certain clades of these animals could reach enormous sizes \citep[e.g.
+Azhdarchids with wingspans of 11 metres; ][]{witton2010size} and, notably, appear to have engaged in soaring flight \citep{witton2010size}.
+
+While soaring flight is perhaps the only viable means of locomotion that allows for an obligate scavenging life-style (ref Ruxton) powered flight is still an efficient means of locomotion.
+Flight is a cheaper means of locomotion than either walking or running \citep{tucker1975energetic} (This needs to be much clearer, needs units etc).
+We know that many extant birds exist as facultative scavengers because storks, raptors and corvids all take substantial quantities of carrion in their diet \citep{kendall2013alternative}.
+Similarly we would expect that extinct species would also gain in a similar fashion depending on the efficiency of flight.
+For example, early flying species such as archaeopteryx are predicted to be relatively inefficient (Is there anything on this as I imagine early flight was diverse).
+Similarly someithing abou t the effecieny of pterosuars flight.
+The importance of efficient flying over large areas may explain the lack of scavenging behavior in bats as they are general smaller and less efficient in comparison to larger birds (speculating here but I wouldnt be surprised), with only the textit{Necromatis} (``death-eater''), a large bat from the middle to late Eocene (56 to 33.9 Mya) with a robust cranio-mandibular morphology, a likely candidate for scavenging behavior amongst the x amount of known bat species \citep{Weithofer_Necromantis_1887,Hand_Necromantis_2012}.
+
+Similar to aerial species, aquatic scavengers have a locomotory benefit because water is a medium that is conducive to low-cost movement \citep{tucker1975energetic}.
+In fact, the cost of swimming is lower than either running or flying \citep{williams1999evolution} with only soaring flight likely to surpass it (I assume the reference is for powered flight, again this needs to be more precise).
+This has led some researchers to argue for the feasibility of an obligate scavenging fish \citep{ruxton2004energetic,ruxton2005searching}.
+Sharks perhaps best resemble the large soaring flyer's as they depend on large pectoral fins in order to maintain lift.
+Shark species show very large ranges (for example great white sharks) and are amongst the most lowest cost swimmers (I know the Greenland shark is the slowest swimmer).
+Faculative scavenging is seen in many marine groups including species of extant sharks such as White sharks (\textit{Carcharodon carcharias}) feeding on whale carcasses \citep{fallows2013white}, Greenland sharks feeding on seals (refs), sixgill sharks (the ref with the pigs) and also extinct species with shark teeth found in the remains of dinosaurs, mososaurs and 4-million-year-old mysticete whales \citep{schwimmer1997scavenging,ehret2009caught}.
+
+Other types of swimming in fish, other species propelled by fins (turtles and plesiosaurs etc)
+
+snakes - ie undulating
+
+endotermy - why it doesn't seem to work
+We might expect then that by combining an aquatic environment and an endothermic metabolism that marine mammals would prosper as scavengers.
+Fossil pinnipeds and cetaceans from 60 Mya have transitional features indicative of their trajectory to fully aquatic species \citep{williams1999evolution}.  
+But despite their movement away from land their energetic savings were negligible because the \textit{total} cost incurred by a swimming marine mammal is high \citep{williams1999evolution}. 
+Indeed, their total energetic cost is similar to an equivalent terrestrial mammal \citep{williams1999evolution}.
+This underscores the trade offs between the benefits of endothermy in terms of activity periods and the costs of maintaining such an expensive system.
+
+%\cite{williams1999evolution} notes, ``free-ranging animals must contend with the total energetic expenditure associated with supporting basic biological functions as well as with moving the body and appendages through the environment."
+
+Terrestrial environments are the most energetically costly in which to move, as a animal must trade off the cost of holding the body up and propelling itself forward.
+Species which do not support an upright position such as modern reptiles limit their ability to move for sustained periods \citep{bennett1979endothermy}.
 This is exacerbated by the sprawling gait seen in many lizards which results in the phenomenon known as Carrier's Constraint such that the animal can't move and breathe at the same time because the lateral movements impedes its lungs \citep{carrier1987evolution}.
-These constraints manifest in aspects such as maximum sustainable speed where an equivalent mammal enjoys a six to seven fold increase \citep{ruben1995evolution}.
+These constraints are likely to mean that many ectotermic terrestrial species, which cannot maintain an upright posture would depend on a relatively low level of scavenging (I'm not sure this is true though).
+This would also be true of extinct species.
+It is during the Permian (298.9 - 252.17 Mya) that we have the earliest evidence of vertebrate scavenging where a temnospondyl amphibian fed on the carcass of \textit{Varanops}, a predatory synapsid of the time \citep{reisz2006articulated}.
+Snakes are a bit different are more effecent at moving I think - laots of scavanging.
+Both these groups are more lilky to depend on low energetic costs in order to componsate for low ecnounter rates duee to thier inabilty to move.
+It is with the evolution of endothermy in the therapsid-mammal lineage \citep{clarke2010temperature} that terrestrial vertebrates would have gained the ability to range more widely, a vital component in seeking out carrion.
+
+Modern endothermic mammals can sustain longer periods of energetically expensive activity \citep{bennett1979endothermy} resulting in larger foraging ranges.
 To quantify this effect with a simple example we can turn to some allometric relationships relating sustainable travelling speed to body mass.
@@
 
+Something more a bipedal movement.
 Unsurprisingly, given their enduring appeal, the prevalence of scavenging has been explored in the carnivorous, theropod dinosaurs.
@@
 
-Of course, the importance of terrestrial tetrapods predates the evolution of the dinosaurs.
-%TG: I will down that "dominance" term.
-Maybe change the sentence to "Of course the importance of terrestrial tetrapods in ecosystems predates the evolution of the dinosaurs." AK Done
-It is during the Permian (298.9 - 252.17 Mya) that we have the earliest evidence of vertebrate scavenging where a temnospondyl amphibian fed on the carcass of \textit{Varanops}, a predatory synapsid of the time \citep{reisz2006articulated}.
-And it is with the evolution of endothermy in the therapsid-mammal lineage \citep{clarke2010temperature} that terrestrial vertebrates would have gained the ability to range more widely, a vital component in seeking out carrion.
-
-Scavenging behaviour may have evolved on land as soon as the first terrestrial tetrapods emerged.
-In fact, some of the earlier tetrapods tracks dating back to the early Middle Devonian (393.3 - 387.7 Mya) were found in intertidal environments \citep{Niedzwiedzki2009}.
-These environments are isolated from marine systems twice a day leaving potential carrion unexploited by marine vertebrates.
-\cite{Niedzwiedzki2009} suggest that these environments ``would thus have allowed marine ancestors of tetrapods gradually to acquire terrestrial competence while accessing a new and essentially untouched resource.''
-
-%TG: I think it's missing a transition here: maybe something along "But moving on land as it's limitations..."
-But it is in the air that we find scavengers \textit{par excellence}. %TG: not sure how reviewers will go with French but I like it!
-Certainly, scavenging should be particularly attractive to avian predators compared to mammals. % Not sure about this paragraph 
-The latter can kill prey up to the same body mass as themselves and sometimes an order of magnitude heavier \citep[e.g. socially hunting lions; ][]{owen2008predator}.
-In contrast, birds of prey tend to kill prey smaller than themselves \citep{slagsvold2007prey} because of the greater cost of injury and need to carry off their food.
-% This is likely due to their need to kill animals that they can fly away with, as well as the risk of injury being higher (which carries a higher mortality risk) for a bird than a mammal.
-Scavenging provides a means for birds to exploit species that would otherwise be too big for them to kill.
-% Thus, for birds, scavenging means they can exploit species that would otherwise be too big for them to kill.
-
-Flight is a cheaper means of locomotion than either walking or running \citep{tucker1975energetic}. 
-We know that many extant birds exist as facultative scavengers because storks, raptors and corvids all take substantial quantities of carrion in their diet \citep{kendall2013alternative}.
-The advantage of flight can be extended further in larger species that engage in soaring instead of flapping flight, which is even cheaper still for searching the environment \citep[approximately twice the basal metabolic rate; ][]{hedenstrom1993migration,spivey2014analysing}.
-The benefits this confers are clear from the information we have on the enormous foraging ranges of many vultures \citep{spiegel2013factors} and seabirds \citep{thaxter2012seabird}.
-The difficulty inherent in observing marine birds foraging at sea means our knowledge of their feeding habits is limited. 
-But it does seem that albatrosses, who can range many hundreds of kilometres, take a substantial amount of carrion in their diet \citep{croxall1994dead}.
-This is typically in the form of squid, which float on the surface, allowing the birds to readily pluck their remains out of the water \citep{croxall1994dead}. 
-In vultures, who are more easily observed on land, we have the best known scavengers on Earth. 
-These birds consist of two convergent groups, from the old and the new world where they represent the only example of obligate vertebrate scavengers to date.
-%TG: Or "obligate vertebrate scavengers known to date." For a more emphasizing turn.
-
-%TG: I thought the paragraph was a bit long so here's a cut for "flying scavengers in the past"
-The two families from which modern vultures arose, the Accipitridae and Cathartidae, appear during the Palaeocene \citep[66 - 56 Mya; ][]{Jetz2012, Jarvis2014}.
-Yet, avian flight is far older than this and originates in the Late Jurassic (163.5-145 Mya), conincident with the fossils of \textit{Archaeopteryx lithographica}, so many of the benefits of flight would have been realised from that point on for carnivorous birds.
-Vertebrate flight is much older still, where pterosaurs predate bird origins by a considerable margin in the Late Triassic (235-201.3 Mya).
-Scavenging in this diverse group has been hypothesied many times before \citep{witton2008reappraisal}.
-Certain clades of these animals could reach enormous sizes \citep[e.g.
-Azhdarchids with wingspans of 11 metres; ][]{witton2010size} and, notably, appear to have engaged in soaring flight \citep{witton2010size}.
-%TG: this paragraph's a killer! Love it!
-
-The only other vertebrate group capable of powered flight are the bats where scavenging has not been recorded to our knowledge.
-However, some fossil from the Eocene (56 to 33.9 Mya) suggest the presence of possible scavenging bats.
-A large bat from the middle to late Eocene (56 to 33.9 Mya) with a robust cranio-mandibular morphology, \textit{Necromatis} (``death-eater''), could have intake carrion as a significant proportion of its diet \citep{Weithofer_Necromantis_1887,Hand_Necromantis_2012}.
-Although, this example might be peculiar to this exceptional fossil given most other remains represent animals less suited to the lifestyle \citep{eiting2009global}.
-
-Similar to aerial species, aquatic scavengers have a locomotory benefit because water is a medium that is conducive to low-cost movement \citep{tucker1975energetic}. 
-In fact, the cost of swimming is lower than either running or flying \citep{williams1999evolution}. 
-This has led some researchers to argue for the feasibility of an obligate scavenging fish \citep{ruxton2004energetic,ruxton2005searching}. 
-As in aerial and terrestrial environments we have evidence of facultative scavenging among extinct aquatic species.
-For example, shark teeth have been found in the remains of dinosaurs, mososaurs and 4-million-year-old mysticete whales \citep{schwimmer1997scavenging,ehret2009caught}.  
-%the remains of a mosasaur and a terrestrial hadrosaur were discovered with embedded teeth from a Cretaceous shark, \textit{Squalicorax} \citep{schwimmer1997scavenging}.
-%As well as a likely instance of scavenging between a 4-million-year-old white shark (\textit{Carcharodon}) and  whale from Peru \citep{ehret2009caught}.
-Extant White sharks (\textit{Carcharodon carcharias}) too are known to feed on whale carcasses \citep{fallows2013white}.
-
-We might expect then that by combining an aquatic environment and an endothermic metabolism that marine mammals would prosper as scavengers. 
-Fossil pinnipeds and cetaceans from 60 Mya have transitional features indicative of their trajectory to fully aquatic species \citep{williams1999evolution}.  
-But despite their movement away from land their energetic savings were negligible because the \textit{total} cost incurred by a swimming marine mammal is high \citep{williams1999evolution}. 
-Indeed, their total energetic cost is similar to an equivalent terrestrial mammal \citep{williams1999evolution}.
-This underscores the trade offs between the benefits of endothermy in terms of activity periods and the costs of maintaining such an expensive system.
-
-%\cite{williams1999evolution} notes, ``free-ranging animals must contend with the total energetic expenditure associated with supporting basic biological functions as well as with moving the body and appendages through the environment."
+Something about humans (also the carnivours kangaroo which might be the most efficient mover.)
 
@@
 % These traits have been used to infer scavenging lifestyles in extinct mammals.
+
+%%%%%%%%%%this is moved from the locomotion section as it didn't seem to fit.
+Certainly, scavenging should be particularly attractive to avian predators compared to mammals. % Not sure about this paragraph 
+The latter can kill prey up to the same body mass as themselves and sometimes an order of magnitude heavier \citep[e.g. socially hunting lions; ][]{owen2008predator}.
+In contrast, birds of prey tend to kill prey smaller than themselves \citep{slagsvold2007prey} because of the greater cost of injury and need to carry off their food.
+% This is likely due to their need to kill animals that they can fly away with, as well as the risk of injury being higher (which carries a higher mortality risk) for a bird than a mammal.
+Scavenging provides a means for birds to exploit species that would otherwise be too big for them to kill.
+% Thus, for birds, scavenging means they can exploit species that would otherwise be too big for them to kill.
+%%%%%%%%%%this is moved from the locomotion section as it didn't seem to fit.
 

### 2016-04-06 natural history of scavenging KH edits.tex
Adam edits to Kev edits

@@
 \section*{Encounter Rate}
-All foraging processes depend on encounter rates between the consumer and the resource.
-In the simplest case, encounter rates can be thought of in terms of gas diffusion models were the movement of two agents depends on only on the relative speed of the agents.
-As carcasses are prone to being relatively stationary, the relative speed between a scavenger and carrion is dependent on the movement of the scavenger.
-As most vertebrates depended on more than simply bumping into resources like gas molecules and can actively detect resources through sensory abilities, scavenging will also depend on search rates which are dependent on both species physiology and the dimensionality of the environment (You guessed it Pawar et al 2012).
+All foraging processes depend on the encounter rate between consumer and resource. 
+In the simplest case, this rate can be thought of in terms of a gas diffusion model where the movement of two agents depends only on their relative speed.
+As carcasses are stationary, the relative speed between a scavenger and carrion is dependent on the movement of the scavenger.
+Vertebrates do more than simply bump into resources like gas molecules though, because they can actively detect resources through their sensory abilities. 
+Scavenging will also depend on search rates, which are dependent on both species physiology and the dimensionality of the environment (\citep{pawar2012dimensionality}).
 
 \subsection*{Locomotion}
-In contrast to the importance of locomotion in predation, locomotion relating to scavenging depends more on the ability to efficiency move over large areas.
-This ability generally requires an efficient transfer of metabolic energy into movement which may rely on both physiology (i.e.
-metabolism) and the medium of the environment in which they are moving (aerial, aquatic or terrestrial).
-
-Perhaps the most efficient form of locomotion in vertebrates is paradoxically found in flying species.
-Despite the energetic costs of flight the only known vertebrate obligate scavengers are the old and the new world vultures.
-While powered flight is energetically expensive (ref) it is the ability of these species to exploit air currents using large wingspans that allows them to soar while only approximately doubling their metabolic rate \citep{hedenstrom1993migration,spivey2014analysing}.
+In contrast to the importance of locomotion in predation, locomotion relating to scavenging depends more on the ability to efficiently move over large areas. 
+This ability generally requires an efficient transfer of metabolic energy into movement which relies on both physiology (i.e.
+metabolism) and the medium of the environment in which the animal is moving (i.e.
+aerial, aquatic or terrestrial).
+Perhaps the most efficient form of locomotion in vertebrates is, paradoxically, found in flying species. 
+Despite the energetic costs of flight, the only known vertebrate, obligate scavengers are the old and the new world vultures. 
+This is because, although powered flight is energetically expensive, species like vultures can exploit air currents using large wingspans which allows them to soar while only doubling their metabolic rate \citep{hedenstrom1993migration,spivey2014analysing}.
 By depending on thermal air flows these species can forage over enormous ranges \citep{spiegel2013factors}. 
-This mode of locomotion is also exploited by sea birds, which use strong ocean winds to search large areas of the oceans \citep{thaxter2012seabird}.
-While many species of seabird are likely primarily predators, it seems that albatross, who can range many hundreds of kilometres, take a substantial amount of carrion in their diet \citep{croxall1994dead}.
+This mode of locomotion is also exploited by sea birds, which use strong ocean winds to search large areas of the oceans \citep{norberg2012vertebrate,thaxter2012seabird}. 
+While many species of seabird are likely primarily predators, it seems that albatrosses, who can range many hundreds of kilometres, take a substantial amount of carrion in their diet \citep{croxall1994dead}.
 This is typically in the form of squid, which float on the surface, allowing the birds to readily pluck their remains out of the water \citep{croxall1994dead}.
 
-The use of soaring in extinct species as a means for scavenging is also likely.
-The two families from which modern vultures arose, the Accipitridae and Cathartidae, appear during the Palaeocene \citep[66 - 56 Mya; ][]{Jetz2012, Jarvis2014} and are  likley to have being soaring flyer's as demonstrated by their large wingspans (Is this true?).
+The two families from which modern vultures arose, the Accipitridae and Cathartidae, appear during the Palaeocene \citep[66 - 56 Mya; ][]{Jetz2012, Jarvis2014} % KH and are likley to have being soaring fliers as demonstrated by their large wingspans (Is this true?).
 (Is there any other form of soaring?).
-However, soaring flight is likely to be far older than these groups with avian flight originating in the Late Jurassic (163.5-145 Mya) and pterosaurs in the Late Triassic (235-201.3 Mya).
-Scavenging in the diverse pterosaurs group has been hypothesied many times before \citep{witton2008reappraisal}.
+AK I don't think this is true from what I've seen
+However, soaring flight is likely to be far older than this with avian flight originating in the Late Jurassic (163.5-145 Mya) and vertebrate flight in the Late Triassic (235-201.3 Mya) coincident with the pterosaurs.
+Indeed, scavenging in the diverse pterosaurs has been hypothesied many times before \citep{witton2008reappraisal}. 
 Certain clades of these animals could reach enormous sizes \citep[e.g.
 Azhdarchids with wingspans of 11 metres; ][]{witton2010size} and, notably, appear to have engaged in soaring flight \citep{witton2010size}.
+Thus, the use of soaring in extinct species as a means for scavenging is also likely.
 
-While soaring flight is perhaps the only viable means of locomotion that allows for an obligate scavenging life-style (ref Ruxton) powered flight is still an efficient means of locomotion.
-Flight is a cheaper means of locomotion than either walking or running \citep{tucker1975energetic} (This needs to be much clearer, needs units etc).
+While soaring is perhaps the only viable means of locomotion that allows for an obligate, scavenging life-style \citep{ruxton2004obligate} powered flight is still an efficient means of locomotion. 
+Avian flight is a cheaper means of locomotion than either walking or running \citep{tucker1975energetic} 
+Even taking account of maintenance costs this still bears out, where total cost of movement (J kg\textsuperscript{-1} m\textsuperscript{-1}) scales according to 5.2 $\times$ body mass (kg) \textsuperscript{-0.23} for fliers and 10.7 $\times$ body mass (kg) \textsuperscript{-0.32} for runners \citep{williams1999evolution}.
+
 We know that many extant birds exist as facultative scavengers because storks, raptors and corvids all take substantial quantities of carrion in their diet \citep{kendall2013alternative}. 
 Similarly we would expect that extinct species would also gain in a similar fashion depending on the efficiency of flight. 
-For example, early flying species such as archaeopteryx are predicted to be relatively inefficient (Is there anything on this as I imagine early flight was diverse).
-Similarly someithing abou t the effecieny of pterosuars flight.
-The importance of efficient flying over large areas may explain the lack of scavenging behavior in bats as they are general smaller and less efficient in comparison to larger birds (speculating here but I wouldnt be surprised), with only the textit{Necromatis} (``death-eater''), a large bat from the middle to late Eocene (56 to 33.9 Mya) with a robust cranio-mandibular morphology, a likely candidate for scavenging behavior amongst the x amount of known bat species \citep{Weithofer_Necromantis_1887,Hand_Necromantis_2012}.
+For example, early birds such as \textit{Archaeopteryx} are predicted to have been poor, relatively inefficient fliers \citep{nudds2010narrow}. 
+Similarly someithing about the effecieny of pterosuars flight. 
+The importance of efficient flying over large areas may explain the lack of scavenging behavior in bats as they are generally noctural, a time when they would receive no aid from convective air currents \citep{norberg2012vertebrate}.
+%smaller and less efficient in comparison to larger birds (speculating here but I wouldnt be surprised), %AK Don't think it is the case that bats are less efficient than birds in terms of flight 
+That said, \textit{Necromatis} (``death-eater''), a large bat from the middle to late Eocene (56 to 33.9 Mya) had a robust cranio-mandibular morphology, and is a likely candidate for scavenging behavior amongst the x amount of known bat species \citep{Weithofer_Necromantis_1887,Hand_Necromantis_2012}.
 
 Similar to aerial species, aquatic scavengers have a locomotory benefit because water is a medium that is conducive to low-cost movement \citep{tucker1975energetic}. 
-In fact, the cost of swimming is lower than either running or flying \citep{williams1999evolution} with only soaring flight likely to surpass it (I assume the reference is for powered flight, again this needs to be more precise).
+In fact, the total cost of movement (again in J kg\textsuperscript{-1} m\textsuperscript{-1}) in fish is lower than either running or flying where it scales according to 2.15 $\times$ body mass (kg) \textsuperscript{-0.25} \citep{williams1999evolution} with only \textit{soaring} flight likely to surpass it.
+%KH(I assume the reference is for powered flight, again this needs to be more precise).  
 This has led some researchers to argue for the feasibility of an obligate scavenging fish \citep{ruxton2004energetic,ruxton2005searching}. 
-Sharks perhaps best resemble the large soaring flyer's as they depend on large pectoral fins in order to maintain lift.
-Shark species show very large ranges (for example great white sharks) and are amongst the most lowest cost swimmers (I know the Greenland shark is the slowest swimmer).
-Faculative scavenging is seen in many marine groups including species of extant sharks such as White sharks (\textit{Carcharodon carcharias}) feeding on whale carcasses \citep{fallows2013white}, Greenland sharks feeding on seals (refs), sixgill sharks (the ref with the pigs) and also extinct species with shark teeth found in the remains of dinosaurs, mososaurs and 4-million-year-old mysticete whales \citep{schwimmer1997scavenging,ehret2009caught}.
+Sharks perhaps best resemble the large soaring fliers as they depend on large pectoral fins in order to maintain lift as they swim. 
+Many shark species show very large ranges (e.g. great white sharks (\textit{Carcharodon carcharias})) and have the lowest cost of swimming (I know the Greenland shark is the slowest swimmer). 
+Faculative scavenging is seen in many marine groups including species of extant sharks such as White sharks, that are known to feed on whale carcasses \citep{fallows2013white}, Greenland sharks feeding on seals \citep{watanabe2012slowest}, and sixgill sharks \citep{anderson2016impact}.
+There is evidence too of scavening in extinct species, where shark teeth have been found in the remains of dinosaurs, mososaurs and 4-million-year-old mysticete whales \citep{schwimmer1997scavenging,ehret2009caught}.
 
-Other types of swimming in fish, other species propelled by fins (turtles and plesiosaurs etc)
+% KH Other types of swimming in fish, other species propelled by fins (turtles and plesiosaurs etc)
 
-snakes - ie undulating
+% KH snakes - ie undulating
 
-endotermy - why it doesn't seem to work
-We might expect then that by combining an aquatic environment and an endothermic metabolism that marine mammals would prosper as scavengers.
+% KH endotermy - why it doesn't seem to work
+We might expect then that by combining an aquatic environment and an endothermic metabolism marine mammals would prosper as scavengers.
 Fossil pinnipeds and cetaceans from 60 Mya have transitional features indicative of their trajectory to fully aquatic species \citep{williams1999evolution}.  
 But despite their movement away from land their energetic savings were negligible because the \textit{total} cost incurred by a swimming marine mammal is high \citep{williams1999evolution}. 
-Indeed, their total energetic cost is similar to an equivalent terrestrial mammal \citep{williams1999evolution}.
+Indeed, the total energetic cost is similar to an equivalent terrestrial or aerial mammal \citep{williams1999evolution}.
 This underscores the trade offs between the benefits of endothermy in terms of activity periods and the costs of maintaining such an expensive system.
@@
 
-Terrestrial environments are the most energetically costly in which to move, as a animal must trade off the cost of holding the body up and propelling itself forward.
-Species which do not support an upright position such as modern reptiles limit their ability to move for sustained periods \citep{bennett1979endothermy}.
-This is exacerbated by the sprawling gait seen in many lizards which results in the phenomenon known as Carrier's Constraint such that the animal can't move and breathe at the same time because the lateral movements impedes its lungs \citep{carrier1987evolution}.
-These constraints are likely to mean that many ectotermic terrestrial species, which cannot maintain an upright posture would depend on a relatively low level of scavenging (I'm not sure this is true though).
+%% AK:  Need to be careful here because this whole idea of sprawling posture being worse off than erect posture is not that well supported in the literature. 
+%% Great Transformations in Vertebrate Evolution By Kenneth P. Dial, Neil Shubin, Elizabeth L. Brainerd talks about it at length 
+%% So I think this should stay pitched as an endo/ ectothermy thing rather than posture 
+Terrestrial environments are the most energetically costly in which to move, because an animal must trade off the cost of supporting its body off the ground while propelling itself forward. 
+The idea that the evolutionary transition in posture from the sprawling gait of reptiles to the erect posture of mammals conferred a huge advantage to the latter is often repeated e.g.
+benefits in terms of speed, efficiency, muscle effort and maneuverability \citep{sullivan2015posture}.
+Yet, as \cite{sullivan2015posture} states, most of the hypotheses in favour of this idea remain to be tested in the context of archosaur evolution. 
+Although, metabolic rate unquestionably impacts terrestrial species whereby ectotherms such as many modern reptiles, cannot move for sustained periods \citep{bennett1979endothermy}. 
+% Species which do not support an upright position such as modern reptiles limit their ability to move for sustained periods \citep{bennett1979endothermy}. 
+This is exacerbated by their sprawling gait which results in the phenomenon known as Carrier's constraint such that the animal can't move and breathe at the same time because the lateral movements impedes its lungs \citep{carrier1987evolution}.
+% These constraints are likely to mean that many ectotermic terrestrial species, which cannot maintain an upright posture would depend on a relatively low level of scavenging (I'm not sure this is true though).
 This would also be true of extinct species. 
 It is during the Permian (298.9 - 252.17 Mya) that we have the earliest evidence of vertebrate scavenging where a temnospondyl amphibian fed on the carcass of \textit{Varanops}, a predatory synapsid of the time \citep{reisz2006articulated}.
-Snakes are a bit different are more effecent at moving I think - laots of scavanging.
-Both these groups are more lilky to depend on low energetic costs in order to componsate for low ecnounter rates duee to thier inabilty to move.
+Snakes are a bit different are more effecent at moving I think - lots of scavanging. 
+Both these groups are more likely to depend on low energetic costs in order to componsate for low ecnounter rates duee to thier inabilty to move. 
 It is with the evolution of endothermy in the therapsid-mammal lineage \citep{clarke2010temperature} that terrestrial vertebrates would have gained the ability to range more widely, a vital component in seeking out carrion.
@@
 %TG: missing some transition here. Maybe something short like "Further back in time" or "Previous to that"...
+The evolution of bipedal movment freed up the forelimbs for other purposes e.g. wings, tool-use etc. 
+In terms of its cost \cite{alexander2004bipedal} shows that, in the case of humans, we are more economical than predicted while walking and less so while running.
 
-Something more a bipedal movement.
+%  KH Something more a bipedal movement.
 Unsurprisingly, given their enduring appeal, the prevalence of scavenging has been explored in the carnivorous, theropod dinosaurs.
@@
 
-Something about humans (also the carnivours kangaroo which might be the most efficient mover.)
+% KH Something about humans (also the carnivours kangaroo which might be the most efficient mover.)
 

### 2016-04-06 natural history of scavenging.tex
Edits up to handling time part

@@
 \section*{Introduction}
-%TG: it might be better to reorganise the intro as:
-%§1 - scavenging is understudied because it's hard to say who's a scavenger
-%§2 - it's even more problematic in the fossil record but people have tried some indirect approaches
-%§3 - good thing is that we can apply these indirect approaches to extant systems
-%§4 - in this review we discuss some stuff.
-%TG: or something along these lines, obviously flexible!
+
 Historically, scavengers have not been viewed as the most charismatic of animals.
@@
 
-\subsection*{The Difficulty of Scavenging} %KH: Make that the template for all the next sections
-The chief hurdle to scavenging is finding a sufficient quantity of food, the occurence of which is difficult to predict in space and time.
-Thus, any animal existing as a scavenger must minimise its locomotory costs and maximise its detection capabilities \citep{ruxton2004obligate}.
-Once found, the scavenger has to process the carrion and overcome the agents of decay produced by the action of microorganisms on the carcass in order to take in a suffienct amount of energy \citep{ruxton2014fruit}.
-The habitat must also be productive enough to sustain a pool of animal biomass that will eventually produce carcasses.
-These facets are essentially the key parameters found in the species functional response curves of optimal foraging theory, namely encounter rate, handling time and prey availability \citep{jeschke2002predator}.
-Here, we can draw on these aspects to envisage a scavenger moving through its environment, searching for food and trying to process it efficiently as we explore the prevalence of scavenging through time.
-Our aim is to produce a 'scale of scavenging' whereupon we can place any species, past or present, and assess the importance of scavenging in its diet. 
-% The idea of scrounging from predator kills is undermined by studies showing that in the  majority of ecosystems more animals die from disease and starvation than predation \citep{benbow2015introduction}.
+%the intro needs to lead towards a aim a bit more strongly
+
+\subsection*{The Difficulty of Scavenging}
+The chief hurdle of scavenging is finding a resource that is often difficult to predict in space and time.
+Through chance alone many species will avail of some level of opportunistic scavenging chances. %TG: Two times "opportunities", replaced by "chances"
+However, species that rely on scavenging to sustain substantial portions of their diets must increase the probability of encountering a sufficient amount of carrion in order to meet their energetic demands.
+%Once found, the scavenger must be able to out-compete any potential competitors and process a product under the processes of decay from toxin producing and disease causing microorganisms \citep{ruxton2014fruit}.
+%TG: make the following sentence more sense?
+Once found, the scavenger must be able to out-compete any potential competitors and process the carrion that is undergoing decay and thus contains toxins and diseases produced by micro-organisms \citep{ruxton2014fruit}.
+Finally, the level of scavenging opportunities available will also depend on the density, size, and quality of carcases produced, all of which are dependent on complex ecosystem dynamics.
+These facets which are required for scavenging are essentially the key parameters found in the species functional response curves of optimal foraging theory, namely encounter rate, handling time and prey availability \citep{jeschke2002predator}.
+%TG: Jaysus, this sentence doesn't read well!
+"Functional response curves of optimal foraging theory" can probably be phrased in a nicer/easier way...
+We compare the ability of scavenging across vertebrates focusing on each of these facets and their relationship with scavenging.
+By considering scavenging in a framework of optimal foraging and identifying the prerequisite attributes and processes required for scavenging we propose a ``scale of scavenging'' whereupon we can place any vertebrate species, past or present, and assess the importance of scavenging in its diet.
 
@@
 %KH: need paragraph for how do we know something is (was) a scavenger
-%TG: Insert figure here: scavenging scale + four points: Locomotion/Detection/Consumption/Environment
 \section*{Encounter Rate}
+All foraging processes depend on the encounter rate between the consumer and resources. 
+In the simplest case, this rate can be thought of in terms of a gas diffusion model where the movement of two agents depends only on their relative speed.
+As carcasses are stationary, the relative speed between a scavenger and carrion is only dependent on the movement of the scavenger. 
+Vertebrates do more than simply bump into resources like gas molecules though, because they can actively detect them through their sensory abilities. 
+Therefore, scavenging will also depend on search rates, which are dependent on both species physiology and the dimensionality of the environment \citep{pawar2012dimensionality}.
 
 \subsection*{Locomotion}
-As endotherms, mammals can sustain long bouts %TG: not sure about "bouts" get it with the context and I know it makes the whole review a bit more literary and thus nicer to read but it might exclude non-native english readers.
- of energetically expensive activity \citep{bennett1979endothermy}. 
-By contrast, modern reptiles are ectothermic, limiting their activity periods \citep{bennett1979endothermy}.
-% a successful reptilian scavenger requires a far different set of adapations.
-This is exacerbated by the sprawling gait seen in many lizards which results in the phenomenon known as Carrier's Constraint such that the animal can't move and breathe at the same time because the lateral movements impedes its lungs \citep{carrier1987evolution}.
-These constraints manifest in aspects such as maximum sustainable speed where an equivalent mammal enjoys a six to seven fold increase \citep{ruben1995evolution}.
+In contrast to the importance of locomotion in predation, locomotion relating to scavenging depends more on the ability to efficiently move over large areas. 
+This ability generally requires an efficient transfer of metabolic energy into movement which relies on both physiology (i.e.
+metabolism) and the medium of the environment in which the animal is moving (i.e.
+aerial, aquatic or terrestrial).
+Perhaps the most efficient form of locomotion in vertebrates is, paradoxically, found in flying species. 
+Despite the energetic costs of flight, the only known vertebrate obligate scavengers are the old and the new world vultures. 
+This is because, although powered flight is energetically expensive, species like vultures can exploit air currents using large wingspans which allows them to soar while only doubling their metabolic rate \citep{hedenstrom1993migration,spivey2014analysing}.
+%TG: I didn't get the meaning of this sentence, is soaring doubling the metabolic rate?
+or is it a word misplacement ("while only doubling their metabolic rate")?
+and if so, how does that make the point?
+By depending on thermal air flows these species can forage over vast ranges \citep{spiegel2013factors}. 
+This mode of locomotion is also exploited by sea birds, which use strong ocean winds to search large areas of the oceans \citep{norberg2012vertebrate,thaxter2012seabird}. 
+While many species of seabird are likely primarily predators, it seems that albatrosses, who can range many hundreds of kilometres, take a substantial amount of carrion in their diet \citep{croxall1994dead}.
+This is typically in the form of squid, which float on the surface, allowing the birds to readily pluck their remains out of the water \citep{croxall1994dead}.
+
+The two families from which modern vultures arose, the Accipitridae and Cathartidae, appear during the Palaeocene \citep[66 - 56 Million years ago (Mya); ][]{Jetz2012, Jarvis2014} % KH and are likley to have being soaring fliers as demonstrated by their large wingspans (Is this true?).
+(Is there any other form of soaring?).
+AK I don't think this is true from what I've seen
+However, soaring flight is likely to be far older than this with avian flight originating in the Late Jurassic (163.5-145 Mya) and vertebrate flight in the Late Triassic (235-201.3 Mya) coincident with the pterosaurs.
+Indeed, scavenging among pterosaurs has been hypothesised many times before \citep{witton2008reappraisal}. 
+Certain groups of these animals could reach enormous sizes \citep[e.g.
+Azhdarchids with wingspans of 11 metres; ][]{witton2010size} and, notably, appear to have engaged in soaring flight \citep{witton2010size}.
+Thus, the use of soaring in extinct species as a means for scavenging is also likely.
+
+While soaring is perhaps the only viable means of locomotion that allows for an obligate, scavenging life-style \citep{ruxton2004obligate} powered flight is still an efficient means of locomotion. 
+Avian flight is a cheaper means of locomotion than either walking or running \citep{tucker1975energetic} 
+Even taking account of maintenance costs this still bears out, where total cost of movement (J kg\textsuperscript{-1} m\textsuperscript{-1}) scales according to 5.2 $\times$ body mass (kg) \textsuperscript{-0.23} for fliers and 10.7 $\times$ body mass (kg) \textsuperscript{-0.32} for runners \citep{williams1999evolution}.
+
+We know that many extant birds exist as facultative scavengers because storks, raptors and corvids all take substantial quantities of carrion in their diet \citep{kendall2013alternative}. 
+Similarly we would expect that extinct species would also scavenge in a similar fashion depending on the efficiency of their flight. 
+For example, early birds such as \textit{Archaeopteryx} are predicted to have been poor, relatively inefficient fliers \citep{nudds2010narrow}. 
+%Similarly something about the efficiency of pterosaurs flight.
+TG: Was this a comment?
+By the way, about pterosaurs (though it might not be necessary to mention it), most of them where tiny little feckers like Archaeopteryx or even smaller.
+The importance of efficient flying over large areas may explain the lack of scavenging behaviour in bats as they are generally nocturnal, a time when they would receive no aid from convective air currents \citep{norberg2012vertebrate}.
+%smaller and less efficient in comparison to larger birds (speculating here but I wouldnt be surprised), %AK Don't think it is the case that bats are less efficient than birds in terms of flight 
+That said, \textit{Necromantis} (``death-eater''), a large bat from the middle to late Eocene (56 to 33.9 Mya) had a robust cranio-mandibular morphology, and is a likely candidate for scavenging behaviour \citep{Weithofer_Necromantis_1887,Hand_Necromantis_2012}.% amongst the x amount of known bat species \citep{Weithofer_Necromantis_1887,Hand_Necromantis_2012}.
+%TG: we can cut this number of bat species to avoid getting into stamp collecting debate.
++ there are effectively a two shit load units of them which makes the Necromantis example not really convincing.
+
+Similar to aerial species, aquatic scavengers have a locomotory benefit because water is a medium that is conducive to low-cost movement \citep{tucker1975energetic}. 
+In fact, the total cost of movement (again in J kg\textsuperscript{-1} m\textsuperscript{-1}) in fish is lower than either running or flying where it scales according to 2.15 $\times$ body mass (kg) \textsuperscript{-0.25} \citep{williams1999evolution} with only \textit{soaring} flight likely to surpass it.
+%KH(I assume the reference is for powered flight, again this needs to be more precise).  
+This has led some researchers to argue for the likelihood of an obligate scavenging fish \citep{ruxton2004energetic,ruxton2005searching}. 
+Sharks perhaps best resemble the large soaring fliers as they depend on large pectoral fins in order to maintain lift as they swim. 
+Many shark species show very large ranges (e.g.
+great white sharks (\textit{Carcharodon carcharias})) and have the lowest cost of swimming.%TG: I guess the following is a comment: (I know the Greenland shark is the slowest swimmer).
+Facultative scavenging is seen in many marine groups including species of extant sharks such as great white sharks \citep[known to feed on whale carcasses;][]{fallows2013white}, Greenland sharks \citep[feeding on seals;][]{watanabe2012slowest}, and sixgill sharks \citep{anderson2016impact}.
+There is evidence too of scavening in extinct species, where shark teeth have been found in the remains of dinosaurs, mosasaurs and mysticete whales from the Pliocene \citep[5.3 to 3.6 Mya;]{schwimmer1997scavenging,ehret2009caught}.
+
+% KH Other types of swimming in fish, other species propelled by fins (turtles and plesiosaurs etc)
+
+% KH snakes - ie undulating
+
+% KH endotermy - why it doesn't seem to work
+We might expect then that by combining an aquatic environment and an endothermic metabolism marine mammals would prosper as scavengers.
+Fossil pinnipeds and cetaceans from 60 Mya %TG: that early???
+have transitional features indicative of their trajectory to fully aquatic species \citep{williams1999evolution}.  
+But despite their movement away from land their energetic savings were negligible because the \textit{total} cost incurred by a swimming marine mammal is high \citep{williams1999evolution}. 
+Indeed, the total energetic cost is similar to an equivalent terrestrial or aerial mammal \citep{williams1999evolution}.
+This underscores the trade offs between the benefits of endothermy in terms of activity periods and the costs of maintaining such an expensive system.
+
+%\cite{williams1999evolution} notes, ``free-ranging animals must contend with the total energetic expenditure associated with supporting basic biological functions as well as with moving the body and appendages through the environment."
+
+%% AK:  Need to be careful here because this whole idea of sprawling posture being worse off than erect posture is not that well supported in the literature. 
+%% Great Transformations in Vertebrate Evolution By Kenneth P. Dial, Neil Shubin, Elizabeth L. Brainerd talks about it at length 
+%% So I think this should stay pitched as an endo/ ectothermy thing rather than posture 
+Terrestrial environments are the most energetically costly in which to move, because an animal must trade off the cost of supporting its body off the ground while propelling itself forward. 
+The idea that the evolutionary transition in posture from the sprawling gait of reptiles to the erect posture of mammals conferred a huge advantage to the latter is often mentioned \citep[e.g.
+benefits in terms of speed, efficiency, muscle effort and manoeuvrability; ]{sullivan2015posture}.
+Yet, as \cite{sullivan2015posture} states, most of the hypotheses in favour of this idea remain to be tested in the context of archosaur evolution. 
+Although, metabolic rate unquestionably impacts terrestrial species whereby ectotherms such as many modern reptiles, cannot move for sustained periods \citep{bennett1979endothermy}. 
+% Species which do not support an upright position such as modern reptiles limit their ability to move for sustained periods \citep{bennett1979endothermy}. 
+This is exacerbated by their sprawling gait which results in the phenomenon known as Carrier's constraint such that the animal can't move and breathe at the same time because the lateral movements impedes its lungs \citep{carrier1987evolution}.
+% These constraints are likely to mean that many ectotermic terrestrial species, which cannot maintain an upright posture would depend on a relatively low level of scavenging (I'm not sure this is true though).
+This would also be true of extinct species. 
+It is during the Permian (298.9 - 252.17 Mya) that we have the earliest evidence of vertebrate scavenging where a temnospondyl amphibian fed on the carcass of \textit{Varanops}, a predatory synapsid of the time \citep{reisz2006articulated}.
+%TG: I guess this is a comment too: Snakes are a bit different are more effecent at moving I think - lots of scavanging. 
+Both these groups are more likely to depend on low energetic costs in order to compensate for low encounter rates due to their inability to move. 
+It is with the evolution of endothermy in the therapsid-mammal lineage \citep{clarke2010temperature} that terrestrial vertebrates would have gained the ability to range more widely, a vital component in seeking out carrion.
+
+Modern endothermic mammals can sustain longer periods of energetically expensive activity \citep{bennett1979endothermy} resulting in larger foraging ranges. 
 To quantify this effect with a simple example we can turn to some allometric relationships relating sustainable travelling speed to body mass.
@@
 
-Today, terrestrial scavenging in the mammals is probably best known in an African context where hyenas, jackals and lions all take sizable proportions of carrion in their diet.
+Today, terrestrial scavenging in the mammals is probably best known in an African context where hyenas, jackals and lions all take sizeable proportions of carrion in their diet.
 In the spotted hyena (\textit{Crocuta crocuta}), striped hyena (\textit{Hyaena hyaena}) and brown hyena (\textit{Hyaena brunnea}) it can be over 90\% \citep{jones2015african}.
-And although no contemporary terrestrial vertebrate exists as an obligate scavenger, most, if not all, %TG: or maybe no brackets... Up to you Adam.
- are facultative to some extent \citep{beasley2015vertebrates}.
+And although no contemporary terrestrial vertebrate exists as an obligate scavenger, most, if not all, are facultative to some extent \citep{beasley2015vertebrates}.
 The particular reliance of hyenas on carrion means we can use them as examples of efficient terrestrial scavengers to compare with other forms. 
@@
 %TG: missing some transition here. Maybe something short like "Further back in time" or "Previous to that"...
+The evolution of bipedal movement freed up the forelimbs for other purposes e.g. wings, tool-use etc. 
+In terms of its cost \cite{alexander2004bipedal} shows that, in the case of humans, we are more economical than predicted while walking and less so while running.
 
+%  KH Something more a bipedal movement.
 Unsurprisingly, given their enduring appeal, the prevalence of scavenging has been explored in the carnivorous, theropod dinosaurs.
@@
 
-Of course, the importance of terrestrial tetrapods predates the evolution of the dinosaurs.
-%TG: I will down that "dominance" term.
-Maybe change the sentence to "Of course the importance of terrestrial tetrapods in ecosystems predates the evolution of the dinosaurs." AK Done
-It is during the Permian (298.9 - 252.17 Mya) that we have the earliest evidence of vertebrate scavenging where a temnospondyl amphibian fed on the carcass of \textit{Varanops}, a predatory synapsid of the time \citep{reisz2006articulated}.
-And it is with the evolution of endothermy in the therapsid-mammal lineage \citep{clarke2010temperature} that terrestrial vertebrates would have gained the ability to range more widely, a vital component in seeking out carrion.
-
-Scavenging behaviour may have evolved on land as soon as the first terrestrial tetrapods emerged.
-In fact, some of the earlier tetrapods tracks dating back to the early Middle Devonian (393.3 - 387.7 Mya) were found in intertidal environments \citep{Niedzwiedzki2009}.
-These environments are isolated from marine systems twice a day leaving potential carrion unexploited by marine vertebrates.
-\cite{Niedzwiedzki2009} suggest that these environments ``would thus have allowed marine ancestors of tetrapods gradually to acquire terrestrial competence while accessing a new and essentially untouched resource.''
-
-%TG: I think it's missing a transition here: maybe something along "But moving on land as it's limitations..."
-But it is in the air that we find scavengers \textit{par excellence}. %TG: not sure how reviewers will go with French but I like it!
-Certainly, scavenging should be particularly attractive to avian predators compared to mammals. % Not sure about this paragraph 
-The latter can kill prey up to the same body mass as themselves and sometimes an order of magnitude heavier \citep[e.g. socially hunting lions; ][]{owen2008predator}.
-In contrast, birds of prey tend to kill prey smaller than themselves \citep{slagsvold2007prey} because of the greater cost of injury and need to carry off their food.
-% This is likely due to their need to kill animals that they can fly away with, as well as the risk of injury being higher (which carries a higher mortality risk) for a bird than a mammal.
-Scavenging provides a means for birds to exploit species that would otherwise be too big for them to kill.
-% Thus, for birds, scavenging means they can exploit species that would otherwise be too big for them to kill.
-
-Flight is a cheaper means of locomotion than either walking or running \citep{tucker1975energetic}. 
-We know that many extant birds exist as facultative scavengers because storks, raptors and corvids all take substantial quantities of carrion in their diet \citep{kendall2013alternative}.
-The advantage of flight can be extended further in larger species that engage in soaring instead of flapping flight, which is even cheaper still for searching the environment \citep[approximately twice the basal metabolic rate; ][]{hedenstrom1993migration,spivey2014analysing}.
-The benefits this confers are clear from the information we have on the enormous foraging ranges of many vultures \citep{spiegel2013factors} and seabirds \citep{thaxter2012seabird}.
-The difficulty inherent in observing marine birds foraging at sea means our knowledge of their feeding habits is limited. 
-But it does seem that albatrosses, who can range many hundreds of kilometres, take a substantial amount of carrion in their diet \citep{croxall1994dead}.
-This is typically in the form of squid, which float on the surface, allowing the birds to readily pluck their remains out of the water \citep{croxall1994dead}. 
-In vultures, who are more easily observed on land, we have the best known scavengers on Earth. 
-These birds consist of two convergent groups, from the old and the new world where they represent the only example of obligate vertebrate scavengers to date.
-%TG: Or "obligate vertebrate scavengers known to date." For a more emphasizing turn.
-
-%TG: I thought the paragraph was a bit long so here's a cut for "flying scavengers in the past"
-The two families from which modern vultures arose, the Accipitridae and Cathartidae, appear during the Palaeocene \citep[66 - 56 Mya; ][]{Jetz2012, Jarvis2014}.
-Yet, avian flight is far older than this and originates in the Late Jurassic (163.5-145 Mya), conincident with the fossils of \textit{Archaeopteryx lithographica}, so many of the benefits of flight would have been realised from that point on for carnivorous birds.
-Vertebrate flight is much older still, where pterosaurs predate bird origins by a considerable margin in the Late Triassic (235-201.3 Mya).
-Scavenging in this diverse group has been hypothesied many times before \citep{witton2008reappraisal}.
-Certain clades of these animals could reach enormous sizes \citep[e.g.
-Azhdarchids with wingspans of 11 metres; ][]{witton2010size} and, notably, appear to have engaged in soaring flight \citep{witton2010size}.
-%TG: this paragraph's a killer! Love it!
-
-The only other vertebrate group capable of powered flight are the bats where scavenging has not been recorded to our knowledge.
-However, some fossil from the Eocene (56 to 33.9 Mya) suggest the presence of possible scavenging bats.
-A large bat from the middle to late Eocene (56 to 33.9 Mya) with a robust cranio-mandibular morphology, \textit{Necromatis} (``death-eater''), could have intake carrion as a significant proportion of its diet \citep{Weithofer_Necromantis_1887,Hand_Necromantis_2012}.
-Although, this example might be peculiar to this exceptional fossil given most other remains represent animals less suited to the lifestyle \citep{eiting2009global}.
-
-Similar to aerial species, aquatic scavengers have a locomotory benefit because water is a medium that is conducive to low-cost movement \citep{tucker1975energetic}. 
-In fact, the cost of swimming is lower than either running or flying \citep{williams1999evolution}. 
-This has led some researchers to argue for the feasibility of an obligate scavenging fish \citep{ruxton2004energetic,ruxton2005searching}. 
-As in aerial and terrestrial environments we have evidence of facultative scavenging among extinct aquatic species.
-For example, shark teeth have been found in the remains of dinosaurs, mososaurs and 4-million-year-old mysticete whales \citep{schwimmer1997scavenging,ehret2009caught}.  
-%the remains of a mosasaur and a terrestrial hadrosaur were discovered with embedded teeth from a Cretaceous shark, \textit{Squalicorax} \citep{schwimmer1997scavenging}.
-%As well as a likely instance of scavenging between a 4-million-year-old white shark (\textit{Carcharodon}) and  whale from Peru \citep{ehret2009caught}.
-Extant White sharks (\textit{Carcharodon carcharias}) too are known to feed on whale carcasses \citep{fallows2013white}.
-
-We might expect then that by combining an aquatic environment and an endothermic metabolism that marine mammals would prosper as scavengers. 
-Fossil pinnipeds and cetaceans from 60 Mya have transitional features indicative of their trajectory to fully aquatic species \citep{williams1999evolution}.  
-But despite their movement away from land their energetic savings were negligible because the \textit{total} cost incurred by a swimming marine mammal is high \citep{williams1999evolution}. 
-Indeed, their total energetic cost is similar to an equivalent terrestrial mammal \citep{williams1999evolution}.
-This underscores the trade offs between the benefits of endothermy in terms of activity periods and the costs of maintaining such an expensive system.
-
-%\cite{williams1999evolution} notes, ``free-ranging animals must contend with the total energetic expenditure associated with supporting basic biological functions as well as with moving the body and appendages through the environment."
+% KH Something about humans (also the carnivours kangaroo which might be the most efficient mover.)
 
 \subsection*{Detection}
-It would be pointless to have incredible ranging abilites and not have the sensory architecture to benefit from it.
-If we came at this from a position of complete ignorance we would predict scavengers to have well-developed senses and indeed, this is what we find. 
+It would be pointless to have incredible ranging abilities and not have the sensory architecture to benefit from it.
+%If we came at this from a position of complete ignorance we would predict scavengers to have well-developed senses and indeed, this is what we find. %TG: I don't really like this sentence. How about:
+As predicted by the necessity of an increased encounter rate, scavengers have well-developed detection senses.
 A simplification of terrestrial, vertebrate scavengers in sensory terms is one of them existing in a two-dimensional plane while foraging for carrion directly.
@@
 %could resolve a 2 metre target at 1 km distance. 
-If we compare this to the energetics approach by \cite{ruxton2004obligate}, they calculated a terrestrial scavenger needs to be able to detect carrion at 500 m in order to survive, which is clearly within the ability of hyenas.
+If we compare this to the energetics approach by \cite{ruxton2004obligate}, they calculated a terrestrial scavenger needs to be able to detect carrion at 500 meter in order to survive, which is clearly within the ability of hyenas.
 % While considering prehistoric habitats \cite{ruxton2004obligate} calculated that ``a 1 tonne mammal or reptile, in an ecosystem yielding carrion at densities similar to the current Serengeti, could have met its energy requirements if it could detect carrion over a distance of the order of 400–500 m".
-Moreover, the senses of many extant (and in all probabilty extinct) carnivores meet this required distance, making scavenging feasible for terrestrial species \citep{farlow1994speculations,mech2010wolves}.
+Moreover, the senses of many extant (and in all probability extinct) carnivores meet this required distance, making scavenging feasible for terrestrial species \citep{farlow1994speculations,mech2010wolves}.
 
@@
 Such an ability has obvious benefits in detecting carrion.
-Vultures are known to have impressive visual acuity, with one estimate indicating Lappet-faced Vultures (\textit{Torgos tracheliotus}) are capable of detecting a 2 metre carcass over 10 km away \citep{spiegel2013factors}.
+Vultures are known to have impressive visual acuity, with one estimate indicating lappet-faced vultures %TG: We need a standardisation for english species names, I'll go for all lower case (similar as what you'd do for "pig" or other english common species names).
+ (\textit{Torgos tracheliotus}) are capable of detecting a 2 metre carcass over 10 km away \citep{spiegel2013factors}.
 Eagles too are known to have highly developed visual abilities \citep{reymond1985spatial}.
 It follows from this that the evolution of flight allowed aerial animals to detect far more carrion than their terrestrial counterparts \citep{AR:AR22815}.
-We can contrast this ability with bats, whose visual acuity is famously poor. 
-It also appears that echolocation would not lend itself to discovering immobile carrion.
-Their small size and poor terrestrial ability would also count against them at a carcass \citep{riskin2006terrestrial}.
+%We can contrast this ability with bats, whose visual acuity is famously poor.
+%TG: Humph!
+That's not nice!
+I'm wouldn't go with all the "this is known" thingies here.
+It really lowers down the quality of our claim and is totally un-necessary (and actually probably wrong: lot's of fruit bats (like flying foxes) do not echolocate and thus have to find their fruits among the flipping tropical folliage using their eyes and noses only)...
+We can contrast that with bats for which echolocation would not lend itself to discover immobile carrion.
+Additionally, their small size and poor terrestrial ability would also count against them when feeding at a carcass \citep{riskin2006terrestrial}.
+
+%TG: missing some transition here.
 
@@
 
-%TG: I think it's missing some connection here.
-The "returning" makes it a bit jumpy.
-Maybe it could be better to include that directly in the paragraph above since it is making the point: "Some other volant groups, however, have not developed all these things".
-
-%TG: Same, I feel it's missing a transition here.
-
 % Depending on the species, a carcass in water either floats or descends to the sea floor \citep{Whitehead415}.
-In contrast to the air, aquatic species have to contend with the low-light environment of water where visual detection distances are far lower (< 100 m) than they would be in the air.
-As such, animals here detect resources through chemo- and mechanoreception more so than through vision \citep{ruxton2004energetic}.
-This is particularly relevant to extant sharks and aquatic snakes who are deemed as having the most suitable physiology for scavenging.
+In contrast to the air, aquatic species have to contend with the low-light environment where visual detection distances are far lower (< 100 m) than they would be in the air.
+As such, aquatic animals detect resources through chemo- and mechanoreception more so than through vision \citep{ruxton2004energetic}.
+This is particularly relevant to sharks and aquatic snakes who are deemed as having the most suitable physiology for scavenging.
 A hypothesis put forth by \cite{sazima1990necrofagia} argued that chemical gradients in water would allow for a relatively easier detection of carrion by snakes.
@@
 Smell seems to be the primary means of carcass detection in sharks as well. 
-\cite{fallows2013white} found that wind speed determined the number of sharks feeding at whale carcasses, indicating they were dependent on detecting the odours from the decaying whales.
+\cite{fallows2013white} found that wind speed %TG: wind speed? Or current speed?
+ determined the number of sharks feeding at whale carcasses, indicating they were dependent on detecting the odours from the decaying whales.
 
@@
 \cite{devault2002scavenging} suggest this is an avenue for scavenging in snakes because they ``exhibit  exceedingly  low  maintenance  metabolisms,  and most  can  survive  on  a  few  scant  feedings per year.
-It  is, therefore, possible for snakes to rely largely  on  infrequent,  less  energy-rich  meals." In the same review the authors found occurrences of scavenging spread across five families of snakes and stated that this behaviour is ``far more common than currently acknowledged."\citep{devault2002scavenging}.
+It  is, therefore, possible for snakes to rely largely  on  infrequent,  less  energy-rich  meals."
+In the same review the authors found occurrences of scavenging spread across five families of snakes and stated that this behaviour is ``far more common than currently acknowledged."\citep{devault2002scavenging}.
 The same reasoning can be applied to crocodlies and their allies \citep{forrest2003evidence}. 

### 2016-04-07 natural history of scavenging.tex
Edits to Handling Time part
Prey Availability to come soon!

@@
 Being able to extract nutrients from remnants gives a scavenger a great advantage.
-Thus, the bone crushing ability of hyenas and others reveals another useful scavenger trait.
-Osteophagy is known across a range of terrestrial carnivores and given some fat-rich mammalian bones have an energy density (6.7 kJ/g) comparable with that of muscle tissue, it makes skeletal remains an enticing resource \citep{brown1989study}.
+%Thus, the bone crushing ability of hyenas and others reveals another useful scavenger trait. %TG: how about the following for a smoother transition:
+Thus, some phenotypic adaptation such as the ability to crush bone (like in hyenas) reveals another useful scavenger trait.
+Osteophagy is known across a range of terrestrial carnivores and given that some fat-rich mammalian bones have an energy density (6.7 kJ/g) comparable with that of muscle tissue, it makes skeletal remains an enticing resource \citep{brown1989study}.
 This ability reached its zenith among hyenas with the evolution of the 110 kg \textit{Pachycrocuta brevirostris} during the Pliocene \citep[3.6 - 2.58 Mya; ][]{palmqvist2011giant}.
 Indeed, their extinction has been blamed on the decline of sabretooth cats (Machairodontinae), the unique skull morphology of the latter meant they would leave a large amount of food on a carcass for would-be scavengers \citep{palmqvist2011giant}.
-Earlier in the evolution of mammals, the bone-crushing dogs that evolved during the Oligocene (subfamily Borophaginae; 33.9 - 23.03 Mya) have been compared to hyenas in terms of their feeding ecology as well \citep{van2003chapter,martin2016pursuit}.
+Earlier in the evolution of mammals, the bone-crushing dogs that evolved during the Oligocene (Borophaginae; 33.9 - 23.03 Mya) have been compared to hyenas in terms of their feeding ecology as well \citep{van2003chapter,martin2016pursuit}.
 
@@
 % These traits have been used to infer scavenging lifestyles in extinct mammals. 
-
 %%%%%%%%%%this is moved from the locomotion section as it didn't seem to fit.
@@
 
+%TG: This part above makes sense in the adaptations parts (i.e.
+in the Handling Time part) but we can see when we clean down the whole paper if we keep it (then we need to integrate it) or if we have enough.
+I actually fits way better in the new narrative by the way!
+
 Interestingly, such comparisons have given insight into the feeding ecology of early hominins who, for instance, had the ability to craft tools for breaking open bones \citep{ARCM:ARCM12084}.
 The question of where our ancestors placed on the hunter-scavenger axis during the Plio-Pleistocene has been a matter of debate for years \citep{dominguez2002hunting}.
-A recent study investigating potential scavenging opportunities for hominins in Kenya found that, even when discounting bone material, there is a substantial amount of scavengeable meat left on predated remains; sufficient to sustain the requirements of an adult male \textit{Homo erectus} \citep{pobiner2015new}.
+A recent study investigating potential scavenging opportunities for hominins in Kenya found that, even when discounting bone material, there is a substantial amount of scavengeable %TG: is that an adjective?
+meat left on predated remains; sufficient to sustain the requirements of an adult male \textit{Homo erectus} \citep{pobiner2015new}.
 In some historical hominin-inhabited areas there were a greater number of felids than hyenids.
 Again, this is significant because hyenas are likely to have left far less flesh on a carcass than a felid such as a sabretooth enabling contemperaneous hominins to benefit \citep{pobiner2015new}.
-The intelligence, resultant tool-use and cooperative nature of hominins meant they could likely adapt to take on more or less carrion depending on their environment \citep{moleon2014humans}.
-%TG: or "could"? This is hypothetical behaviour stuff! Maybe past hyenas where exclusive sushi eaters!
-%TG: Not sure about the intelligence thing: the sentence as the same meaning without and will be less anthropocentric sounding.
+The use of tools and the cooperative nature of hominins meant they could likely take a certain part of their energetic requirements from carrions depending on their environment \citep{moleon2014humans}.
+%TG: I've removed the intelligence bit from the sentence.
 
-In Mesozoic systems some extremely large theropod dinosaurs had a morphology indicative of an ability to process bone e.g. the robust skull and dentition of \textit{T. rex} \citep{hone2010feeding}.
+In Mesozoic systems some extremely large theropod dinosaurs had a morphology indicative of an ability to process bone e.g.
+the robust skull and dentition of \textit{Tyrannosaurus rex} \citep{hone2010feeding}.
 There is direct evidence that \textit{T.
@@
 The animal also had an enormous bite force, with one estimate putting it at 57000 Newtons \citep{bates2012estimating}.
+%TG: A comment about the future cleaning up phase.
+I think mentions like the 570000 Newtons are not essential and could go to some supplementary files.
+Same applies for most of the physiques in the locomotion part, it is not essential in the narrative to have the number (yet important to link to them!).
 This is noted as being powerful enough to break open skeletal material \citep{rayfield2001cranial}.
 
-Further, much work has focused on the existence of scavenging in dinosaurs by using simple energetics approaches that typically focused on a single species namely \textit{Tyrannosaurus rex} \citep{ruxton2003could,carbone2011intra} but a recent modelling study investigated the likely prevelance of scavenging across a range of body sizes.
+Further, much work has focused on the existence of scavenging in dinosaurs by using simple energetics approaches that typically focused on a single species namely \textit{T.
+rex} \citep{ruxton2003could,carbone2011intra} but a recent modelling study investigated the likely prevalence of scavenging across a range of body sizes.
 In it the authors demonstrated that species of intermediate body masses (approximatively 500 kg) would have gained the most benefit from scavenging.
@@
 In the same review the authors found occurrences of scavenging spread across five families of snakes and stated that this behaviour is ``far more common than currently acknowledged."\citep{devault2002scavenging}.
-The same reasoning can be applied to crocodlies and their allies \citep{forrest2003evidence}. 
+The same reasoning can be applied to crocodiles and their allies \citep{forrest2003evidence}. 
 A sit and wait strategy is viable for an ectotherm. 
 This low existence cost is also realised in many sharks who have coupled low locomotory costs with an ectothermic metabolism. 
-The upshot is that 30 kg of blubber can sustain a White shark for over six weeks \citep{carey1982temperature}.
+The upshot is that 30 kg of blubber can sustain a white shark for over six weeks \citep{carey1982temperature}.
 
@@
 Thus, we would expect scavengers to have this trait selected for even in the case of weight-constrained fliers.
-Wandering Albatrosses (\textit{Diomedea exulans}), Cinereous Vultures (\textit{Aegypius monachus}) and condors (\textit{Vultur gryphus}, \textit{Gymnogyps californianus}) all have body masses that can exceed 10 kg and represent some of the heaviest bird species capable of flight \citep{weimerskirch1992reproductive,ferguson2001raptors,donazar2002effects}.
+Wandering Albatrosses (\textit{Diomedea exulans}), cinereous vultures (\textit{Aegypius monachus}) and condors (\textit{Vultur gryphus}, \textit{Gymnogyps californianus}) all have body masses that can exceed 10 kg and represent some of the heaviest bird species capable of flight \citep{weimerskirch1992reproductive,ferguson2001raptors,donazar2002effects}.
 
 And as we have noted the Azhdarchid pterosaurs were far bigger again, with estimated body masses of over 200 kg \citep{witton2010size}.
-Although \cite{witton2008reappraisal} argued that neck inflexibility and straight, rather than hooked jaw morphology points against pterosaurs existing as \textit{obligate} scavengers, Azhdarchid terrestrial proficency indicates they would have been comfortable foraging on the ground.
-Indeed, extant Marabou Storks (\textit{Leptoptilos crumenifer}) 
-have a comparable morphology and are noted facultative scavengers \citep{monadjem2012survival} so it is reasonable to believe that certain pterosaurs behaved similarly.
+Although \cite{witton2008reappraisal} argued that neck inflexibility and straight, rather than hooked jaw morphology points against pterosaurs existing as \textit{obligate} scavengers, Azhdarchid terrestrial proficiency indicates they would have been comfortable foraging on the ground.
+Indeed, extant Marabou Storks (\textit{Leptoptilos crumenifer}) have a comparable morphology and are noted facultative scavengers \citep{monadjem2012survival} so it is reasonable to believe that certain pterosaurs behaved similarly.
 
-The competitive ability of even the largest bird is radically diminished in their interactions with mammalian competitors however, and as such they tend to consume carrion rapidly. 
+The competitive ability of even the largest living bird is radically diminished in their interactions with mammalian competitors however, and as such they tend to consume carrion rapidly.
+%TG: I've added "living" bird here.
+In the Paleocene you have these giant terror birds that were still top predators (like during the Cretaceous...
+How weird!
+;)).
 \cite{houston1974role} observed a group of \textit{Gyps} vultures consuming all of the soft tissue from a 50 kg Grant’s gazelle (\textit{Nanger granti}) in eight minutes. 
@@
 Avian scavengers have evolved incredibly acidic stomachs that allow them to consume and process putrefied flesh with no ill effects \citep{houston1975digestive,roggenbuck2014microbiome}. 
-This adapation is not restricted to vultures though, \cite{gremillet2012vultures} showed wandering albatrosses (\textit{Diomedea exulans}; so-called 'vultures of the seas') had an average pH of 1.5, which enables them to consume fisheries discards and squid carcasses.
-There is also evidence of selection for 'toxification minimizers' beyond birds among the ectotherms.
+This adaptation is not restricted to vultures though, \cite{gremillet2012vultures} showed wandering albatrosses (\textit{Diomedea exulans}; so-called ``vultures of the seas'') had an average pH of 1.5, which enables them to consume fisheries discards and squid carcasses.
+There is also evidence of selection for ``toxification minimizers'' beyond birds among the ectotherms.
 From our earlier arguments we know that ecthotherms are limited in their ability to find carrion as quickly as endotherms. 
@@
 \cite{shivik2006vultures} suggests that ``specialized oral structures in snakes may have evolved under pressures associated with scavenging."
-Moreover, some authorities have charted an evolutionary course from basal fossorial snakes to modern terrestrial species by way of an obligate scavenger intermediate \citep{bauchot2006snakes}.
+Moreover, some suggested an evolutionary course from basal fossorial snakes to modern terrestrial species by way of an obligate scavenger intermediate \citep{bauchot2006snakes}.
 
@@
 %TG: Two points here: one speculation for the non-scavenging behaviour could be because of diet.
-It is in the ability to process carrion that bats suffer.
+Conversely, some other entire clades might lack of such phenotypic adaptations. %TG: I've tried some transition here
+It is the case, for example in living bats that lack the ability to process carrion.
 Big bats (which are better suited for scavenging, following our previous argument) are typically frugivores and therefore lack the adaptations for digesting meat.
@@
 % Maybe they scavenge on dead insects but here we're talking about scavenging on real meat right?
-%TG: Second point: bats fossil record is of course not so good but it's actually way better than birds!
-Most of the data we have is on teeth of course but it's probably enough for having a rough body mass and diet estimate!
-And that's much better than birds where you're stuck with just some fragments of long bones.
 
@@
 As we discussed earlier, this ability is often extremely beneficial to a scavenger.
+
+%TG: This bit can be squeezed somewhere here:
+% Scavenging behaviour may have evolved on land as soon as the first terrestrial tetrapods emerged.
+% In fact, some of the earlier tetrapods tracks dating back to the early Middle Devonian (393.3 - 387.7 Mya) were found in intertidal environments \citep{Niedzwiedzki2009}.
+% These environments are isolated from marine systems twice a day leaving potential carrion unexploited by marine vertebrates.
+% \cite{Niedzwiedzki2009} suggest that these environments ``would thus have allowed marine ancestors of tetrapods gradually to acquire terrestrial competence while accessing a new and essentially untouched resource.''
 

### 2016-04-07 natural history of scavenging.tex
Edits to Prey Availability section

@@
 \section*{Prey Availability}
-Both the biotic and abiotic environment a would-be scavenger finds itself in can influence to degree to which it can depend on carrion. 
-\cite{ruxton2004obligate} suggest a system similar to the Serengeti in productivity could have supported a mammalian or reptilian terrestrial scavenger \citep{ruxton2004obligate}.
-Indeed, in systems that were dominated by large ectothermic or mesothermic vertebrates, 
-the same primary productivity would have supported a greater biomass \citep{mcnab2009resources}.
-The upshot of this is there was a higher biomass of herbivores dying and offering scavenging opportunities.
+%Both the biotic and abiotic environment a would-be scavenger finds itself in can influence to degree to which it can depend on carrion. %TG: This sentence is convoluted: maybe go with the following:
+The position of species on the ``scavenging scale'' can also be influenced, to some degree by the availability of carrions in the environment which is in turn depending from biotic and abiotic factors.
+%TG: It's actually not better...
+Feel free to change again!
+\cite{ruxton2004obligate} suggest a system with a productivity similar to the Serengeti could have supported a mammalian or reptilian terrestrial scavenger \citep{ruxton2004obligate}.
+Indeed, in systems that were dominated by large ectothermic or mesothermic vertebrates, the same primary productivity would have supported a greater biomass \citep{mcnab2009resources}.
+%TG: We're now talking about Late Mesozoic systems right? Needs to be mentioned somewhere.
+The upshot of this is that there was a higher biomass of herbivores dying and offering scavenging opportunities.
 Predators were large-bodied too compared to extant mammalian predators \citep{mcnab2009resources}, and so, especially if they were ectothermic, could last longer between meals, rendering scavenging a more attractive behaviour relative to predation.
@@
 
-%TG: This bit can be squeezed somewhere here:
-% Scavenging behaviour may have evolved on land as soon as the first terrestrial tetrapods emerged.
-% In fact, some of the earlier tetrapods tracks dating back to the early Middle Devonian (393.3 - 387.7 Mya) were found in intertidal environments \citep{Niedzwiedzki2009}.
-% These environments are isolated from marine systems twice a day leaving potential carrion unexploited by marine vertebrates.
-% \cite{Niedzwiedzki2009} suggest that these environments ``would thus have allowed marine ancestors of tetrapods gradually to acquire terrestrial competence while accessing a new and essentially untouched resource.''
-
 Vultures and eagles tend to soar using thermals and if these air pockets don't form, say on a cloudy day, the bird is grounded \citep{mundy1992vultures}.
-In many habitats (e.g. the arctic) it is simply not possible for sufficiently powerful thermals to form and as a consequence large-bodied vultures cannot exist.
+In many habitats (e.g. the Arctic) it is simply not possible for sufficiently powerful thermals to form and as a consequence large-bodied vultures cannot exist.
 The upshot of this is that terrestrial carnivores like bears and wolves take more carrion \citep{devault2003scavenging}.
 Certainly, a major difficulty for terrestrial scavengers is competition with vultures.
-Noctural behaviour in the hyaenidae in general has been put forth as an adaptation to reduce competition with these exclusively diurnal birds \citep{gittleman2013carnivore}.
+Nocturnal behaviour in the hyaenidae in general has been put forth as an adaptation to reduce competition with these exclusively diurnal birds \citep{gittleman2013carnivore}.
 If we apply this line of reasoning over evolutionary time-scales, the absence of flying vertebrates in the Palaeozoic may have permitted terrestrial forms to take in a higher proportion of carrion in their diet.
+
+%TG: Added the early tetrapod bit here:
+This absence or reduced competition in before the Jurassic might have allowed some species to exploit these scavenger niches.
+In fact, scavenging behaviour may have evolved on land as soon as the first terrestrial tetrapods emerged because.
+Some of the earlier tetrapods tracks dating back to the early Middle Devonian (393.3 - 387.7 Mya) were found in intertidal environments \citep{Niedzwiedzki2009}.
+These environments are isolated from marine systems twice a day leaving potential carrion unexploited by marine vertebrates.
+\cite{Niedzwiedzki2009} suggest that these environments ``would thus have allowed marine ancestors of tetrapods gradually to acquire terrestrial competence while accessing a new and essentially untouched resource.''
 
@@
 
+%TG: maybe change transition here.
 Staying in the aquatic setting, the phenomenon of occasional bounties of carrion in the form of whale falls has led some researchers to investigate if a scavenger could survive by seeking out these remains exclusively.
@@
 Although no aquatic species have ever exceeded the size of whales, some enormous animals have evolved in this environment before the evolution of whales, including \textit{Leedsichthys}, a bony fish from the Middle Jurassic (174.1-163.5 Mya), that weighed in excess of 20 tonnes.
-Thus, the energetic feasiblity of a marine scavenger that specialises on large carcasses has a long history.
+Thus, the energetic feasibility of a marine scavenger that specialises on large carcasses has a long history.
 One point of interest is that of the whaling industry, which provided a bonanza of floating carcasses especially during the 20th century \citep{Whitehead415}.
-This meant Killer Whales (\textit{Orcinus orca}) could switch from hunting to scavenging, a switch made that much easier by the noise of the whaling vessels that would effectively ring the ``dinner-bells" \citep{Whitehead415}.
-Early whales such as \textit{Basilosaurus} seem to fit into the same niche as Killer Whales and we have some evidence for scavenging in this group as well \citep{fahlke2012bite}.
+This meant killer whales (\textit{Orcinus orca}) could switch from hunting to scavenging, a switch made that much easier by the noise of the whaling vessels that would effectively ring the ``dinner-bells" \citep{Whitehead415}.
+Early whales such as \textit{Basilosaurus} seem to fit into the same niche as killer whales and we have some evidence for scavenging in this group as well \citep{fahlke2012bite}.
 
@@
 This will have had a significant bearing on the availability and persistence of carrion.
-To illustrate the point, a 10$^{\circ}$C increase in ambient temperature can double carcass decomposition rates \citep{parmenter2009carrion} and geological evidence indicates that the Mesozoic Earth was at least 6 $^{\circ}$C warmer than now \citep{sellwood2006mesozoic}.
+To illustrate the point, a 10$^{\circ}$C increase in ambient temperature can double carcass decomposition rates \citep{parmenter2009carrion} and geological evidence indicates that the Mesozoic Earth was on average at least 6 $^{\circ}$C warmer than now \citep{sellwood2006mesozoic}.
 In terms of specific habitats, it has been shown that decomposition is greater in warm and moist areas versus more xeric ones \citep{beasley2015vertebrates}.
@@
 
-%\section*{The Role of Climate} %TG: Use it as a opening at the end of the paper (i.e. conclusion + finishing with this bit)
+%TG: Really nice to finish on this paragraph + good for transition to conclusion!
+
 \section*{Conclusion} 

### 2016-04-07 natural history of scavenging.tex
More edits!

@@
 
-%\affiliations\{
-%\item A. Kane (\url{adam.kane@ucc.ie}), University College Cork, Cooperage Building, School of Biological Earth and Environmental Sciences, Cork, Ireland.
-%\item K. Healy and A. Jackson, Trinity College Dublin, Department of Zoology, School of Natural Sciences, Dublin Ireland.
-%\item T. Guillerme, Imperial College London, Silwood Park Campus, Department of Life Sciences, Buckhurst Road, Ascot SL5 7PY, United Kingdom.
-%\item G. Ruxton, School of Biology, Sir Harold Mitchell Building, Greenside Place, St Andrews, KY16 9TH, United Kingdom.
-%}
+\affiliations{
+\item A. Kane (\url{adam.kane@ucc.ie}), University College Cork, Cooperage Building, School of Biological Earth and Environmental Sciences, Cork, Ireland.
+\item K. Healy and A. Jackson, Trinity College Dublin, Department of Zoology, School of Natural Sciences, Dublin Ireland.
+\item T. Guillerme, Imperial College London, Silwood Park Campus, Department of Life Sciences, Buckhurst Road, Ascot SL5 7PY, United Kingdom.
+\item G. Ruxton, School of Biology, Sir Harold Mitchell Building, Greenside Place, St Andrews, KY16 9TH, United Kingdom.
+}
 
@@
 
-% \maketitle
+ \maketitle
 
@@
 The chief hurdle of scavenging is finding a resource that is often difficult to predict in space and time.
-Through chance alone many species will avail of some level of opportunistic scavenging chances. %TG: Two times "opportunities", replaced by "chances"
+Through chance alone many species will avail of some opportunistic scavenging. %TG: Two times "opportunities", replaced by "chances"
 However, species that rely on scavenging to sustain substantial portions of their diets must increase the probability of encountering a sufficient amount of carrion in order to meet their energetic demands.
@@
 %TG: make the following sentence more sense?
-Once found, the scavenger must be able to out-compete any potential competitors and process the carrion that is undergoing decay and thus contains toxins and diseases produced by micro-organisms \citep{ruxton2014fruit}.
-Finally, the level of scavenging opportunities available will also depend on the density, size, and quality of carcases produced, all of which are dependent on complex ecosystem dynamics.
-These facets which are required for scavenging are essentially the key parameters found in the species functional response curves of optimal foraging theory, namely encounter rate, handling time and prey availability \citep{jeschke2002predator}.
+Once found, the scavenger must be able to out-compete any potential competitors and process the typically decaying carcass which contains microorganism derived toxins  \citep{ruxton2014fruit}.
+Finally, the potential for scavenging will also depend on the density, size, and quality of carcases produced, all of which are affected by complex ecosystem dynamics.
+All of these facets are essentially the key parameters found in functional response curves, namely encounter rate, handling time and prey availability \citep{jeschke2002predator}.
 %TG: Jaysus, this sentence doesn't read well!
 "Functional response curves of optimal foraging theory" can probably be phrased in a nicer/easier way...
-We compare the ability of scavenging across vertebrates focusing on each of these facets and their relationship with scavenging.
-By considering scavenging in a framework of optimal foraging and identifying the prerequisite attributes and processes required for scavenging we propose a ``scale of scavenging'' whereupon we can place any vertebrate species, past or present, and assess the importance of scavenging in its diet.
+%With this framework we compare the ability of scavenging across vertebrates. % focusing on each of these facets and their relationship with scavenging.
+By considering scavenging in this context of optimal foraging we can identify the prerequisite attributes and processes required for the behaviour. 
+This has enabled us to propose a scale of scavenging whereupon we can place any vertebrate species, past or present, and assess the importance of carrion in its diet.
 
@@
 Vertebrates do more than simply bump into resources like gas molecules though, because they can actively detect them through their sensory abilities. 
-Therefore, scavenging will also depend on search rates, which are dependent on both species physiology and the dimensionality of the environment \citep{pawar2012dimensionality}.
+As such, scavenging potential is strongly affected by search rates, which are determined by both species' physiology and the dimensionality of the environment \citep{pawar2012dimensionality}.
 
 \subsection*{Locomotion}
-In contrast to the importance of locomotion in predation, locomotion relating to scavenging depends more on the ability to efficiently move over large areas. 
-This ability generally requires an efficient transfer of metabolic energy into movement which relies on both physiology (i.e.
+Scavenging depends more on the ability to efficiently move over large areas than does predation because of the inherent unpredictability of carrion. 
+This generally requires an efficient transfer of metabolic energy into movement which relies on both physiology (i.e.
 metabolism) and the medium of the environment in which the animal is moving (i.e.
@@
 Despite the energetic costs of flight, the only known vertebrate obligate scavengers are the old and the new world vultures. 
-This is because, although powered flight is energetically expensive, species like vultures can exploit air currents using large wingspans which allows them to soar while only doubling their metabolic rate \citep{hedenstrom1993migration,spivey2014analysing}.
+And, although powered flight is energetically expensive, species like vultures can exploit air currents using their large wingspans which allows them to soar at a cost of only twice their metabolic rate \citep{hedenstrom1993migration,spivey2014analysing}.
 %TG: I didn't get the meaning of this sentence, is soaring doubling the metabolic rate?
@@
 By depending on thermal air flows these species can forage over vast ranges \citep{spiegel2013factors}. 
-This mode of locomotion is also exploited by sea birds, which use strong ocean winds to search large areas of the oceans \citep{norberg2012vertebrate,thaxter2012seabird}. 
+An analagous mode of locomotion is also exploited by sea birds, which use strong ocean winds to search large areas of the oceans \citep{norberg2012vertebrate,thaxter2012seabird}. 
 While many species of seabird are likely primarily predators, it seems that albatrosses, who can range many hundreds of kilometres, take a substantial amount of carrion in their diet \citep{croxall1994dead}.
@@
 Azhdarchids with wingspans of 11 metres; ][]{witton2010size} and, notably, appear to have engaged in soaring flight \citep{witton2010size}.
-Thus, the use of soaring in extinct species as a means for scavenging is also likely.
+It seems that the use of soaring in extinct species as a means for scavenging is likely.
 
-While soaring is perhaps the only viable means of locomotion that allows for an obligate, scavenging life-style \citep{ruxton2004obligate} powered flight is still an efficient means of locomotion. 
-Avian flight is a cheaper means of locomotion than either walking or running \citep{tucker1975energetic} 
-Even taking account of maintenance costs this still bears out, where total cost of movement (J kg\textsuperscript{-1} m\textsuperscript{-1}) scales according to 5.2 $\times$ body mass (kg) \textsuperscript{-0.23} for fliers and 10.7 $\times$ body mass (kg) \textsuperscript{-0.32} for runners \citep{williams1999evolution}.
+While soaring is perhaps the only viable means of locomotion that allows for an obligate, scavenging life-style \citep{ruxton2004obligate}, powered flight is still an efficient means of locomotion. 
+Avian flight is cheaper than either walking or running \citep{tucker1975energetic}. 
+Even taking account of maintenance costs, this still bears out, where total cost of movement (J kg\textsuperscript{-1} m\textsuperscript{-1}) scales according to 5.2 $\times$ body mass (kg) \textsuperscript{-0.23} for fliers and 10.7 $\times$ body mass (kg) \textsuperscript{-0.32} for runners \citep{williams1999evolution}.
 
@@
 Many shark species show very large ranges (e.g.
-great white sharks (\textit{Carcharodon carcharias})) and have the lowest cost of swimming.%TG: I guess the following is a comment: (I know the Greenland shark is the slowest swimmer).
+great white sharks (\textit{Carcharodon carcharias})) and have the lowest cost of swimming.
+%TG: I guess the following is a comment: (I know the Greenland shark is the slowest swimmer).
 Facultative scavenging is seen in many marine groups including species of extant sharks such as great white sharks \citep[known to feed on whale carcasses;][]{fallows2013white}, Greenland sharks \citep[feeding on seals;][]{watanabe2012slowest}, and sixgill sharks \citep{anderson2016impact}.
@@
 % KH endotermy - why it doesn't seem to work
-We might expect then that by combining an aquatic environment and an endothermic metabolism marine mammals would prosper as scavengers.
+We might expect then that by combining an aquatic environment and an endothermic metabolism marine mammals would especially prosper as scavengers.
 Fossil pinnipeds and cetaceans from 60 Mya %TG: that early???
@@
 This underscores the trade offs between the benefits of endothermy in terms of activity periods and the costs of maintaining such an expensive system. 
-
+That said, aquatic endotherms have and do scavenge. 
+For instance, early whales such as \textit{Basilosaurus} seem to fit into the same niche as killer whales \textit{Orcinus orca} and we have some evidence for scavenging in both \citep{fahlke2012bite,Whitehead415}.
 %\cite{williams1999evolution} notes, ``free-ranging animals must contend with the total energetic expenditure associated with supporting basic biological functions as well as with moving the body and appendages through the environment."
@@
 Terrestrial environments are the most energetically costly in which to move, because an animal must trade off the cost of supporting its body off the ground while propelling itself forward. 
-The idea that the evolutionary transition in posture from the sprawling gait of reptiles to the erect posture of mammals conferred a huge advantage to the latter is often mentioned \citep[e.g.
-benefits in terms of speed, efficiency, muscle effort and manoeuvrability; ]{sullivan2015posture}.
+The idea that the evolutionary transition in posture from the sprawling gait of reptiles to the erect posture of mammals conferred a huge advantage to the latter is often mentioned.
+The purported advantages include benefits in terms of speed, efficiency, muscle effort and manoeuvrability \citep{sullivan2015posture}. 
 Yet, as \cite{sullivan2015posture} states, most of the hypotheses in favour of this idea remain to be tested in the context of archosaur evolution. 
-Although, metabolic rate unquestionably impacts terrestrial species whereby ectotherms such as many modern reptiles, cannot move for sustained periods \citep{bennett1979endothermy}. 
+Metabolic rate however, unquestionably impacts terrestrial species whereby ectotherms such as many modern reptiles, cannot move for sustained periods \citep{bennett1979endothermy}. 
 % Species which do not support an upright position such as modern reptiles limit their ability to move for sustained periods \citep{bennett1979endothermy}. 
@@
 Modern endothermic mammals can sustain longer periods of energetically expensive activity \citep{bennett1979endothermy} resulting in larger foraging ranges. 
-To quantify this effect with a simple example we can turn to some allometric relationships relating sustainable travelling speed to body mass.
-In the case of mammals and reptiles these are 1.15 $\times$ body mass (kg) \textsuperscript{0.12} and 0.23 $\times$ body mass (kg) \textsuperscript{0.12} respectively \citep{ruxton2004obligate}.
-If we insert these into a foraging radius model $\sfrac{\frac{\text{duration} \times \text{speed}}{2}}{1000}$ for a 12 hour foraging day which shows that while a 10 kg reptile can range 6.5 km, an equally sized mammal can range nearly 33 km \citep{Enstipp2006Energetics}.
+To quantify this effect with a simple example we can turn to some allometric relationships relating sustainable travelling speed to body mass \citep{ruxton2004obligate}.
+If we insert these into a foraging radius model \citep{Enstipp2006Energetics} for a 12 hour foraging day it shows that while a 10 kg reptile can range 6.5 km, an equally sized mammal can range nearly 33 km .
 For a foraging scavenger, this ability translates into a greater area searched for food.
@@
 Such an ability has obvious benefits in detecting carrion.
-Vultures are known to have impressive visual acuity, with one estimate indicating lappet-faced vultures %TG: We need a standardisation for english species names, I'll go for all lower case (similar as what you'd do for "pig" or other english common species names).
+Certainly, vultures are known to have impressive visual acuity, with one estimate indicating lappet-faced vultures %TG: We need a standardisation for english species names, I'll go for all lower case (similar as what you'd do for "pig" or other english common species names).
  (\textit{Torgos tracheliotus}) are capable of detecting a 2 metre carcass over 10 km away \citep{spiegel2013factors}.
-Eagles too are known to have highly developed visual abilities \citep{reymond1985spatial}.
-It follows from this that the evolution of flight allowed aerial animals to detect far more carrion than their terrestrial counterparts \citep{AR:AR22815}.
+Eagles too are known to have highly developed vision \citep{reymond1985spatial}.
+It follows that the evolution of flight allowed aerial animals to detect far more carrion than their terrestrial counterparts \citep{AR:AR22815}.
 %We can contrast this ability with bats, whose visual acuity is famously poor.
@@
 It really lowers down the quality of our claim and is totally un-necessary (and actually probably wrong: lot's of fruit bats (like flying foxes) do not echolocate and thus have to find their fruits among the flipping tropical folliage using their eyes and noses only)...
-We can contrast that with bats for which echolocation would not lend itself to discover immobile carrion.
-Additionally, their small size and poor terrestrial ability would also count against them when feeding at a carcass \citep{riskin2006terrestrial}.
+We can contrast this with many bats whose reliance on echolocation would not lend itself to discovering immobile carrion.
+%
 
@@
 
-Having a panoramic view means being able to gather a wealth of information from other foragers, be they conspecifics or otherwise \citep{jackson2008effect}.
+Having a panoramic view also means being able to gather a wealth of information from other foragers, be they conspecifics or otherwise \citep{jackson2008effect}.
 Again, returning to vultures, the genus \textit{Gyps} consists of highly social and colonially nesting species \citep{fernandez2015density}.
@@
 %Thus, the bone crushing ability of hyenas and others reveals another useful scavenger trait. %TG: how about the following for a smoother transition:
-Thus, some phenotypic adaptation such as the ability to crush bone (like in hyenas) reveals another useful scavenger trait.
+Thus, some phenotypic adaptations such as the ability to crush bone (as seen in hyenas) reveals another useful scavenger trait.
 Osteophagy is known across a range of terrestrial carnivores and given that some fat-rich mammalian bones have an energy density (6.7 kJ/g) comparable with that of muscle tissue, it makes skeletal remains an enticing resource \citep{brown1989study}.
@@
 Indeed, their extinction has been blamed on the decline of sabretooth cats (Machairodontinae), the unique skull morphology of the latter meant they would leave a large amount of food on a carcass for would-be scavengers \citep{palmqvist2011giant}.
-Earlier in the evolution of mammals, the bone-crushing dogs that evolved during the Oligocene (Borophaginae; 33.9 - 23.03 Mya) have been compared to hyenas in terms of their feeding ecology as well \citep{van2003chapter,martin2016pursuit}.
+Earlier in the evolution of mammals, the bone-crushing dogs that evolved during the Oligocene (Borophaginae; 33.9 - 23.03 Mya) have also been compared to hyenas in terms of their feeding ecology \citep{van2003chapter,martin2016pursuit}.
 
@@
 % These traits have been used to infer scavenging lifestyles in extinct mammals.
-%%%%%%%%%%this is moved from the locomotion section as it didn't seem to fit.
-Certainly, scavenging should be particularly attractive to avian predators compared to mammals. % Not sure about this paragraph 
-The latter can kill prey up to the same body mass as themselves and sometimes an order of magnitude heavier \citep[e.g. socially hunting lions; ][]{owen2008predator}.
-In contrast, birds of prey tend to kill prey smaller than themselves \citep{slagsvold2007prey} because of the greater cost of injury and need to carry off their food.
-% This is likely due to their need to kill animals that they can fly away with, as well as the risk of injury being higher (which carries a higher mortality risk) for a bird than a mammal.
-Scavenging provides a means for birds to exploit species that would otherwise be too big for them to kill.
-% Thus, for birds, scavenging means they can exploit species that would otherwise be too big for them to kill.
-%%%%%%%%%%this is moved from the locomotion section as it didn't seem to fit.
-
-%TG: This part above makes sense in the adaptations parts (i.e.
-in the Handling Time part) but we can see when we clean down the whole paper if we keep it (then we need to integrate it) or if we have enough.
-I actually fits way better in the new narrative by the way!
 
@@
 Wandering Albatrosses (\textit{Diomedea exulans}), cinereous vultures (\textit{Aegypius monachus}) and condors (\textit{Vultur gryphus}, \textit{Gymnogyps californianus}) all have body masses that can exceed 10 kg and represent some of the heaviest bird species capable of flight \citep{weimerskirch1992reproductive,ferguson2001raptors,donazar2002effects}.
-
 And as we have noted the Azhdarchid pterosaurs were far bigger again, with estimated body masses of over 200 kg \citep{witton2010size}.
@@
 
-The competitive ability of even the largest living bird is radically diminished in their interactions with mammalian competitors however, and as such they tend to consume carrion rapidly.
+%%%%%%%%%%  AK This annoying paragraph's new home
+Certainly, scavenging should be particularly attractive to flying species compared to mammals. % Not sure about this paragraph 
+The latter can kill prey up to the same body mass as themselves and sometimes an order of magnitude heavier \citep[e.g. socially hunting lions; ][]{owen2008predator}.
+In contrast, birds of prey tend to kill prey smaller than themselves \citep{slagsvold2007prey} because of the greater cost of injury and the need to carry off their food.
+% This is likely due to their need to kill animals that they can fly away with, as well as the risk of injury being higher (which carries a higher mortality risk) for a bird than a mammal.
+Scavenging provides a means for birds to exploit species that would otherwise be too big for them to kill.
+% Thus, for birds, scavenging means they can exploit species that would otherwise be too big for them to kill.
+%%%%%%%%%%this is moved from the locomotion section as it didn't seem to fit.
+
+%TG: This part above makes sense in the adaptations parts (i.e.
+in the Handling Time part) but we can see when we clean down the whole paper if we keep it (then we need to integrate it) or if we have enough.
+I actually fits way better in the new narrative by the way!
+On the ground, the competitive ability of even the largest flying bird is radically diminished in their interactions with mammalian competitors however, and as such they tend to consume carrion rapidly.
 %TG: I've added "living" bird here.
@@
 ;)).
+% AK How about flying bird because that's the key thing here 
 \cite{houston1974role} observed a group of \textit{Gyps} vultures consuming all of the soft tissue from a 50 kg Grant’s gazelle (\textit{Nanger granti}) in eight minutes. 
 Their serrated tongues and hooked bills enabling them to achieve this feat \citep{houston1975digestive}. 
-Outside of raptors such as vultures, the specialised beaks of many modern bird lineages hinders their ability to eat meat which is in contrast to the first lineages that did not have this feature \citep{martyniuk2012field} .
+Outside of raptors such as vultures, the specialised beaks of many modern bird lineages hinders their ability to eat meat which is in contrast to the first lineages that did not have this feature \citep{martyniuk2012field}.
 As \cite{martyniuk2012field} notes these early birds would thus have been predominantly carnivorous, which implies that scavenging would have been a live opportunity cf. their descendants.
 
-Although the observation of \cite{shivik2006vultures} that ``evolutionary pressures favor detection maximizers relative to toxification minimizers in competitive interactions for carcasses." appears sound, the fact remains that overcoming microorganism toxins is still a beneficial adaptation to any scavenger.
+Although the findings of \cite{shivik2006vultures} that ``evolutionary pressures favor detection maximizers relative to toxification minimizers in competitive interactions for carcasses." appears sound, the fact remains that overcoming microorganism toxins is still a beneficial adaptation to any scavenger.
 Avian scavengers have evolved incredibly acidic stomachs that allow them to consume and process putrefied flesh with no ill effects \citep{houston1975digestive,roggenbuck2014microbiome}. 
@@
 \cite{shivik2006vultures} suggests that ``specialized oral structures in snakes may have evolved under pressures associated with scavenging."
-Moreover, some suggested an evolutionary course from basal fossorial snakes to modern terrestrial species by way of an obligate scavenger intermediate \citep{bauchot2006snakes}.
+Moreover, some researchers have suggested an evolutionary course from basal fossorial snakes to modern terrestrial species by way of an obligate scavenger intermediate \citep{bauchot2006snakes}.
 
@@
 %TG: Two points here: one speculation for the non-scavenging behaviour could be because of diet.
-Conversely, some other entire clades might lack of such phenotypic adaptations. %TG: I've tried some transition here
-It is the case, for example in living bats that lack the ability to process carrion.
-Big bats (which are better suited for scavenging, following our previous argument) are typically frugivores and therefore lack the adaptations for digesting meat.
-While carnivorous bats are mainly found in the microbats which are insectivorous \citep{aguirre2003implications}.
+Conversely, entire clades appear to lack many, if not all, of these phenotypic adaptations. %TG: I've tried some transition here
+For example, the extant bats appear to lack most of the features we have identified as important in reducing handling time.
+The larger forms (which are better suited for scavenging, following our previous arguments) are typically frugivores and therefore lack the adaptations for digesting meat.
+While the smaller carnivorous bats are mainly found in the microbats which are insectivorous \citep{aguirre2003implications}.
+Additionally, their poor terrestrial ability would also count against them when feeding at a carcass \citep{riskin2006terrestrial}.
 % Maybe they scavenge on dead insects but here we're talking about scavenging on real meat right?
@@
 %Both the biotic and abiotic environment a would-be scavenger finds itself in can influence to degree to which it can depend on carrion. %TG: This sentence is convoluted: maybe go with the following:
-The position of species on the ``scavenging scale'' can also be influenced, to some degree by the availability of carrions in the environment which is in turn depending from biotic and abiotic factors.
+The position of a species on the scavenging scale can also be influenced by the availability of carrion in the environment, which is dependent on biotic and abiotic factors.
 %TG: It's actually not better...
 Feel free to change again!
-\cite{ruxton2004obligate} suggest a system with a productivity similar to the Serengeti could have supported a mammalian or reptilian terrestrial scavenger \citep{ruxton2004obligate}.
+Aspects including, primary productivity, relief, temperature and competition will all greatly affect scavenging tendency. 
+\cite{ruxton2004obligate} suggest a system with a productivity similar to the Serengeti could have supported a mammalian or reptilian terrestrial scavenger.
 Indeed, in systems that were dominated by large ectothermic or mesothermic vertebrates, the same primary productivity would have supported a greater biomass \citep{mcnab2009resources}.
@@
 Predators were large-bodied too compared to extant mammalian predators \citep{mcnab2009resources}, and so, especially if they were ectothermic, could last longer between meals, rendering scavenging a more attractive behaviour relative to predation.
-Osteophagy may have been even more viable during the Mesozoic era because of this skewed body mass distribution of herbviores towards larger sizes \citep{10.1371/journal.pone.0051925}.
+Osteophagy may have been even more viable during the Mesozoic era as well because of this skewed body mass distribution of herbviores towards larger sizes \citep{10.1371/journal.pone.0051925}.
 When we couple this with the fact that skeletal mass scales greater than linearly with body mass \citep{prange1979scaling} there would have been a lot of bone material to consume in the environment provided an animal had the biology to process it \citep{chure1997one}.
-As we discussed earlier, this ability is often extremely beneficial to a scavenger.
+% As we discussed earlier, this ability is often extremely beneficial to a scavenger.
 
-Vultures and eagles tend to soar using thermals and if these air pockets don't form, say on a cloudy day, the bird is grounded \citep{mundy1992vultures}.
+Frequently, the interplay between abiotic and biotic factors can impact the ability of an animal to scavenge. 
+We know vultures and eagles tend to soar using thermals and if these air pockets don't form, say on a cloudy day, the bird is grounded \citep{mundy1992vultures}.
 In many habitats (e.g. the Arctic) it is simply not possible for sufficiently powerful thermals to form and as a consequence large-bodied vultures cannot exist.
-The upshot of this is that terrestrial carnivores like bears and wolves take more carrion \citep{devault2003scavenging}.
+One result of this is that terrestrial carnivores like bears and wolves take more carrion \citep{devault2003scavenging}.
 Certainly, a major difficulty for terrestrial scavengers is competition with vultures.
@@
 %TG: Added the early tetrapod bit here:
-This absence or reduced competition in before the Jurassic might have allowed some species to exploit these scavenger niches.
-In fact, scavenging behaviour may have evolved on land as soon as the first terrestrial tetrapods emerged because.
+% This absence or reduced competition in before the Jurassic might have allowed some species to exploit these scavenger niches.
+In fact, scavenging behaviour may have evolved on land as soon as the first terrestrial tetrapods emerged.
 Some of the earlier tetrapods tracks dating back to the early Middle Devonian (393.3 - 387.7 Mya) were found in intertidal environments \citep{Niedzwiedzki2009}.
@@
 One point of interest is that of the whaling industry, which provided a bonanza of floating carcasses especially during the 20th century \citep{Whitehead415}.
-This meant killer whales (\textit{Orcinus orca}) could switch from hunting to scavenging, a switch made that much easier by the noise of the whaling vessels that would effectively ring the ``dinner-bells" \citep{Whitehead415}.
-Early whales such as \textit{Basilosaurus} seem to fit into the same niche as killer whales and we have some evidence for scavenging in this group as well \citep{fahlke2012bite}.
+This meant killer whales could switch from hunting to scavenging, a switch made that much easier by the noise of the whaling vessels that would effectively ring the ``dinner-bells" \citep{Whitehead415}.
 
@@
 \cite{beasley2015vertebrates} who point to a series of studies showing how microbes and invertebrates benefit at higher temperatures to the detriment of vertebrate scavengers such that ``above 20$^{\circ}$C vertebrates were able to detect and consume only 19\% of small-mammal carcasses, whereas at temperatures below 18$^{\circ}$C, vertebrates consumed 49\% of carcasses".
-
+This is a sobering thought given the impact we humans have had on the Earth's climate. 
 %TG: Really nice to finish on this paragraph + good for transition to conclusion!
@@
 % Add bits on climate change/dectection rate change, etc...
+
+\section*{Appendix}
+Scaling relationships for sustainable travel speed are 1.15 $\times$ body mass (kg) \textsuperscript{0.12} and 0.23 $\times$ body mass (kg) \textsuperscript{0.12} for mammals and reptiles respectively \citep{ruxton2004obligate}.
+These are fed into the foraging model $\sfrac{\frac{\text{duration} \times \text{speed}}{2}}{1000}$ \citep{Enstipp2006Energetics}.
 

### 2016-04-08 natural history of scavenging.tex
bipedalism and swimming locomotion

@@
 \section*{Encounter Rate}
-All foraging processes depend on the encounter rate between the consumer and resources. 
+All foraging processes depend on the encounter rate between consumer and resource. 
 In the simplest case, this rate can be thought of in terms of a gas diffusion model where the movement of two agents depends only on their relative speed.
@@
 By depending on thermal air flows these species can forage over vast ranges \citep{spiegel2013factors}. 
-An analagous mode of locomotion is also exploited by sea birds, which use strong ocean winds to search large areas of the oceans \citep{norberg2012vertebrate,thaxter2012seabird}. 
+An analagous mode of locomotion is also exploited by seabirds, who use strong ocean winds to search large areas of the oceans \citep{norberg2012vertebrate,thaxter2012seabird}. 
 While many species of seabird are likely primarily predators, it seems that albatrosses, who can range many hundreds of kilometres, take a substantial amount of carrion in their diet \citep{croxall1994dead}.
@@
 This has led some researchers to argue for the likelihood of an obligate scavenging fish \citep{ruxton2004energetic,ruxton2005searching}. 
-Sharks perhaps best resemble the large soaring fliers as they depend on large pectoral fins in order to maintain lift as they swim. 
-Many shark species show very large ranges (e.g.
-great white sharks (\textit{Carcharodon carcharias})) and have the lowest cost of swimming.
+Interestingly, style of swimmming in fish does not significantly affect the cost of movement \citep{williams1999evolution}. 
+Though sharks perhaps best resemble the large soaring fliers as they depend on large pectoral fins in order to maintain lift as they swim. 
+Many shark species show large ranges (e.g.
+great white sharks (\textit{Carcharodon carcharias})) \citep{bruce2006movements} and it seems reasonable that they would use oceangraphic currents to further reduce movement costs  \citep{ruxton2004energetic}.
+%and are able to maintain a type of endothermy from thier movements \citep{bernal2015sharks}.
 %TG: I guess the following is a comment: (I know the Greenland shark is the slowest swimmer).
-Facultative scavenging is seen in many marine groups including species of extant sharks such as great white sharks \citep[known to feed on whale carcasses;][]{fallows2013white}, Greenland sharks \citep[feeding on seals;][]{watanabe2012slowest}, and sixgill sharks \citep{anderson2016impact}.
-There is evidence too of scavening in extinct species, where shark teeth have been found in the remains of dinosaurs, mosasaurs and mysticete whales from the Pliocene \citep[5.3 to 3.6 Mya;]{schwimmer1997scavenging,ehret2009caught}.
+Facultative scavenging is seen in many marine groups, including species of extant sharks like white sharks \citep[known to feed on whale carcasses;][]{fallows2013white}, Greenland sharks \citep[feeding on seals;][]{watanabe2012slowest}, and sixgill sharks \citep{anderson2016impact}.
+There is evidence too of scavening in extinct species, where shark teeth have been found in the remains of dinosaurs, mosasaurs and Pliocene mysticete whales \citep[5.3 to 3.6 Mya;]{schwimmer1997scavenging,ehret2009caught}.
 
@@
 %% So I think this should stay pitched as an endo/ ectothermy thing rather than posture 
-Terrestrial environments are the most energetically costly in which to move, because an animal must trade off the cost of supporting its body off the ground while propelling itself forward. 
+Terrestrial environments are the most energetically costly in which to move, which may be due to the low muscular efficiency of running \citep{tucker1975energetic} as well as the relative inefficiency of gas exchange in mammals cf.
+birds and fish \citep{williams1999evolution}.
 The idea that the evolutionary transition in posture from the sprawling gait of reptiles to the erect posture of mammals conferred a huge advantage to the latter is often mentioned.
 The purported advantages include benefits in terms of speed, efficiency, muscle effort and manoeuvrability \citep{sullivan2015posture}. 
-Yet, as \cite{sullivan2015posture} states, most of the hypotheses in favour of this idea remain to be tested in the context of archosaur evolution. 
+Despite being intuitive, as \cite{sullivan2015posture} states, most of the hypotheses in favour of this idea remain to be tested in the context of archosaur evolution. 
 Metabolic rate however, unquestionably impacts terrestrial species whereby ectotherms such as many modern reptiles, cannot move for sustained periods \citep{bennett1979endothermy}. 
@@
 To quantify this effect with a simple example we can turn to some allometric relationships relating sustainable travelling speed to body mass \citep{ruxton2004obligate}.
-If we insert these into a foraging radius model \citep{Enstipp2006Energetics} for a 12 hour foraging day it shows that while a 10 kg reptile can range 6.5 km, an equally sized mammal can range nearly 33 km .
+If we insert these into a foraging radius model \citep{Enstipp2006Energetics} for a 12 hour foraging day it shows that while a 10 kg reptile can range 6.5 km, an equally sized mammal can range nearly 33 km (See appendix).
 For a foraging scavenger, this ability translates into a greater area searched for food.
@@
 %TG: missing some transition here. Maybe something short like "Further back in time" or "Previous to that"...
-The evolution of bipedal movement freed up the forelimbs for other purposes e.g. wings, tool-use etc. 
-In terms of its cost \cite{alexander2004bipedal} shows that, in the case of humans, we are more economical than predicted while walking and less so while running.
+Although the evolution of bipedal movement was significant in that it freed up the forelimbs for other purposes e.g.
+climbing, tool-use, wing development etc.
+it does not differ radically in cost from quadrupedal locomotion \citep[and references therein]{williams1999evolution}.
+For instance, \cite{alexander2004bipedal} shows that, in the case of humans, we are more economical than predicted while walking and less so while running according to predicted costs of terrestrial movement.
 
-%  KH Something more a bipedal movement.
-Unsurprisingly, given their enduring appeal, the prevalence of scavenging has been explored in the carnivorous, theropod dinosaurs.
-They were the dominant terrestrial carnivores for most of the Mesozoic Era (252.17 - 66 Mya) and ranged from the chicken-sized to the whale-sized, all of which were bipedal.
+The best-known terrestrial bipeds are the dinosaurs and unsurprisingly, given their enduring appeal, the prevalence of scavenging has been explored in the carnivorous theropods.
+These were the dominant terrestrial carnivores for most of the Mesozoic Era (252.17 - 66 Mya) and ranged from the chicken-sized to the whale-sized, all of which were bipedal.
 They are quite alien to anything we know today which restricts our ability to understand their ecology far more so than extinct mammals \citep{weishampel2004dinosauria}.
@@
 %If we came at this from a position of complete ignorance we would predict scavengers to have well-developed senses and indeed, this is what we find. %TG: I don't really like this sentence. How about:
-As predicted by the necessity of an increased encounter rate, scavengers have well-developed detection senses.
+As predicted by the necessity of an increased encounter rate, scavengers have well-developed senses.
 A simplification of terrestrial, vertebrate scavengers in sensory terms is one of them existing in a two-dimensional plane while foraging for carrion directly.
@@
 %could resolve a 2 metre target at 1 km distance. 
-If we compare this to the energetics approach by \cite{ruxton2004obligate}, they calculated a terrestrial scavenger needs to be able to detect carrion at 500 meter in order to survive, which is clearly within the ability of hyenas.
+We can compare this to the energetics approach of \cite{ruxton2004obligate}, who calculated a terrestrial scavenger needs to be able to detect carrion at 500 meters in order to survive, which is clearly within the ability of hyenas.
 % While considering prehistoric habitats \cite{ruxton2004obligate} calculated that ``a 1 tonne mammal or reptile, in an ecosystem yielding carrion at densities similar to the current Serengeti, could have met its energy requirements if it could detect carrion over a distance of the order of 400–500 m".
@@
 These behaviours allow them forage far more efficiently because one bird can scrounge information on the location of food from another successful forager \citep{KaneVul}.
+This efficiency has been exploited by mammals such as hyenas who are known to follow groups of vultures \citep{jones2015african}.
 
@@
 %Thus, the bone crushing ability of hyenas and others reveals another useful scavenger trait. %TG: how about the following for a smoother transition:
-Thus, some phenotypic adaptations such as the ability to crush bone (as seen in hyenas) reveals another useful scavenger trait.
+Thus, some phenotypic adaptations such as the ability to crush bone (as seen in hyenas) reveal another useful scavenger trait.
 Osteophagy is known across a range of terrestrial carnivores and given that some fat-rich mammalian bones have an energy density (6.7 kJ/g) comparable with that of muscle tissue, it makes skeletal remains an enticing resource \citep{brown1989study}.
@@
 Again, this is significant because hyenas are likely to have left far less flesh on a carcass than a felid such as a sabretooth enabling contemperaneous hominins to benefit \citep{pobiner2015new}.
-The use of tools and the cooperative nature of hominins meant they could likely take a certain part of their energetic requirements from carrions depending on their environment \citep{moleon2014humans}.
+The use of tools and the cooperative nature of hominins meant they could likely get a substantial part of their energetic requirements through scavenging depending on their environment \citep{moleon2014humans}.
 %TG: I've removed the intelligence bit from the sentence.
@@
 
-Further, much work has focused on the existence of scavenging in dinosaurs by using simple energetics approaches that typically focused on a single species namely \textit{T.
-rex} \citep{ruxton2003could,carbone2011intra} but a recent modelling study investigated the likely prevalence of scavenging across a range of body sizes.
-In it the authors demonstrated that species of intermediate body masses (approximatively 500 kg) would have gained the most benefit from scavenging.
+We know that large body size confers substantial dominance and starvation-resistance benefits \citep{ruxton2004obligate}. 
+As such, theropod dinosaurs, who could get up to 15 tonnes, would seem likely candidates for scavenging. 
+Much work has focused on the existence of scavenging in dinosaurs by using simple energetics approaches that typically focused on a single species namely \textit{T.
+rex} \citep{ruxton2003could,carbone2011intra} but a recent modelling study investigated its prevalence across a range of body sizes.
+
+In their work, the authors demonstrated that species of \textit{intermediate} body masses (approximatively 500 kg) would have gained the most benefit from scavenging.
 This was the result of gut capacity limitations and the effects of competition at the carcass.
@@
 
-In addition to reducing locomotory costs we would expect adaptations that reduce energetic costs of maintenance to be selected for in scavengers because it would maximise the benefit derived from such a sporadic food source.
-Extant reptiles possess an advantage here, in that over the course of a year their food requirements can be 30 times lower than an endotherm of equal size \citep{Nagy1621}.
-\cite{devault2002scavenging} suggest this is an avenue for scavenging in snakes because they ``exhibit exceedingly low maintenance metabolisms, and most can survive on  a few scant feedings per year.
-It is, therefore, possible for snakes to rely largely on infrequent, less energy-rich meals."
-In the same review the authors found occurrences of scavenging spread across five families of snakes and stated that this behaviour is ``far more common than currently acknowledged."\citep{devault2002scavenging}.
-The same reasoning can be applied to crocodiles and their allies \citep{forrest2003evidence}. 
-A sit and wait strategy is viable for an ectotherm. 
-This low existence cost is also realised in many sharks who have coupled low locomotory costs with an ectothermic metabolism. 
-The upshot is that 30 kg of blubber can sustain a white shark for over six weeks \citep{carey1982temperature}.
+The support of water allows for many aquatic species to reach large sizes thus granting the benefits. 
+For example, white sharks, which grow up to 6 metres long, can be sustained by 30 kg of blubber for over six weeks \citep{carey1982temperature}.
 
-Large body size confers substantial dominance and starvation-resistance benefits \citep{ruxton2004obligate}.
-Thus, we would expect scavengers to have this trait selected for even in the case of weight-constrained fliers.
+Given the advantages of size, we would expect this trait to be selected for even in the case of weight-constrained scavenging fliers.
 Wandering Albatrosses (\textit{Diomedea exulans}), cinereous vultures (\textit{Aegypius monachus}) and condors (\textit{Vultur gryphus}, \textit{Gymnogyps californianus}) all have body masses that can exceed 10 kg and represent some of the heaviest bird species capable of flight \citep{weimerskirch1992reproductive,ferguson2001raptors,donazar2002effects}.
@@
 
+In addition to reducing locomotory costs we would expect adaptations that reduce energetic costs of maintenance to be selected for in scavengers because it would maximise the benefit derived from such a sporadic food source.
+Extant reptiles possess an advantage here, in that over the course of a year their food requirements can be 30 times lower than an endotherm of equal size \citep{Nagy1621}.
+\cite{devault2002scavenging} suggest this is an avenue for scavenging in snakes because they ``exhibit exceedingly low maintenance metabolisms, and most can survive on  a few scant feedings per year.
+It is, therefore, possible for snakes to rely largely on infrequent, less energy-rich meals."
+In the same review the authors found occurrences of scavenging spread across five families of snakes and stated that this behaviour is ``far more common than currently acknowledged."\citep{devault2002scavenging}.
+The same reasoning can be applied to crocodiles and their allies \citep{forrest2003evidence} because a sit and wait strategy is viable for an ectotherm. 
+%This low existence cost is also realised in many sharks who have coupled low locomotory costs with an ectothermic metabolism. 
+%
+
 Although the findings of \cite{shivik2006vultures} that ``evolutionary pressures favor detection maximizers relative to toxification minimizers in competitive interactions for carcasses." appears sound, the fact remains that overcoming microorganism toxins is still a beneficial adaptation to any scavenger.
@@
 While the smaller carnivorous bats are mainly found in the microbats which are insectivorous \citep{aguirre2003implications}.
-Additionally, their poor terrestrial ability would also count against them when feeding at a carcass \citep{riskin2006terrestrial}.
+Additionally, their poor terrestrial ability and cost of movement on the ground would also count against them when feeding at a carcass \citep{riskin2006terrestrial,voigt2012terrestrial}.
 % Maybe they scavenge on dead insects but here we're talking about scavenging on real meat right?
@@
 Although no aquatic species have ever exceeded the size of whales, some enormous animals have evolved in this environment before the evolution of whales, including \textit{Leedsichthys}, a bony fish from the Middle Jurassic (174.1-163.5 Mya), that weighed in excess of 20 tonnes.
-Thus, the energetic feasibility of a marine scavenger that specialises on large carcasses has a long history.
+So, despite being unlikely, the energetic feasibility of a marine scavenger that specialises on large carcasses has a long history.
 One point of interest is that of the whaling industry, which provided a bonanza of floating carcasses especially during the 20th century \citep{Whitehead415}.
@@
 \cite{beasley2015vertebrates} who point to a series of studies showing how microbes and invertebrates benefit at higher temperatures to the detriment of vertebrate scavengers such that ``above 20$^{\circ}$C vertebrates were able to detect and consume only 19\% of small-mammal carcasses, whereas at temperatures below 18$^{\circ}$C, vertebrates consumed 49\% of carcasses".
-This is a sobering thought given the impact we humans have had on the Earth's climate. 
+This is a sobering thought given the impact we humans are having on the Earth's climate. 
 %TG: Really nice to finish on this paragraph + good for transition to conclusion!

### 2016-04-15 natural history of scavenging.tex
added figure and cleaned up handling time

@@
 Yet, there are certain intrinsic and environmental features of a species that push it towards a scavenging lifestyle. 
-Chief among these are low-cost locomotion, high detection distances, effective carcass processing and a carrion-producing habitat.
-We use these four components to highlight the morphologies and environments that would have been conducive to scavenging over geological time by focussing on dominant vertebrate groups of the land, sea and air.
+These can be thought of as some of the principal parameters in optimal foraging theory namely, encounter rate, handling time and prey availability. 
+% Chief among these are low-cost locomotion, high detection distances, effective carcass processing and a carrion-producing habitat. 
+We use these components to highlight the morphologies and environments that would have been conducive to scavenging over geological time by focussing on the dominant vertebrate groups of the land, sea and air.
 The result is a document on the natural history of scavenging, the first to our knowledge. 
@@
 \subsection*{The Difficulty of Scavenging}
-The chief hurdle of scavenging is finding a resource that is often difficult to predict in space and time.
+The chief hurdle to scavenging is finding a resource that is often difficult to predict in space and time.
 Through chance alone many species will avail of some opportunistic scavenging. %TG: Two times "opportunities", replaced by "chances"
@@
 %TG: make the following sentence more sense?
-Once found, the scavenger must be able to out-compete any potential competitors and process the typically decaying carcass which contains microorganism derived toxins  \citep{ruxton2014fruit}.
+Once found, the scavenger must be able to out-compete any potential competitors and process the, typically decaying, carcass replete with microorganism derived toxins  \citep{ruxton2014fruit}.
 Finally, the potential for scavenging will also depend on the density, size, and quality of carcases produced, all of which are affected by complex ecosystem dynamics.
@@
 Azhdarchids with wingspans of 11 metres; ][]{witton2010size} and, notably, appear to have engaged in soaring flight \citep{witton2010size}.
-It seems that the use of soaring in extinct species as a means for scavenging is likely.
+It seems probable that extinct species used soaring as a means for scavenging.
 
@@
 Similarly we would expect that extinct species would also scavenge in a similar fashion depending on the efficiency of their flight. 
-For example, early birds such as \textit{Archaeopteryx} are predicted to have been poor, relatively inefficient fliers \citep{nudds2010narrow}. 
+For example, early birds such as \textit{Archaeopteryx} are predicted to have been poor, relatively inefficient fliers \citep{nudds2010narrow} and so ill-suited to finding carrion. 
 %Similarly something about the efficiency of pterosaurs flight.
@@
 Similar to aerial species, aquatic scavengers have a locomotory benefit because water is a medium that is conducive to low-cost movement \citep{tucker1975energetic}. 
-In fact, the total cost of movement (again in J kg\textsuperscript{-1} m\textsuperscript{-1}) in fish is lower than either running or flying where it scales according to 2.15 $\times$ body mass (kg) \textsuperscript{-0.25} \citep{williams1999evolution} with only \textit{soaring} flight likely to surpass it.
+In fact, the total cost of movement (again in J kg\textsuperscript{-1} m\textsuperscript{-1}) in (salmonid) fish is lower than either running or flying where it scales according to 2.15 $\times$ body mass (kg) \textsuperscript{-0.25} \citep{williams1999evolution} with only \textit{soaring} flight likely to surpass it.
 %KH(I assume the reference is for powered flight, again this needs to be more precise).  
@@
 Though sharks perhaps best resemble the large soaring fliers as they depend on large pectoral fins in order to maintain lift as they swim. 
-Many shark species show large ranges (e.g.
-great white sharks (\textit{Carcharodon carcharias})) \citep{bruce2006movements} and it seems reasonable that they would use oceangraphic currents to further reduce movement costs  \citep{ruxton2004energetic}.
+Many shark species show large ranges, e.g.
+great white sharks (\textit{Carcharodon carcharias}) \citep{bruce2006movements}, and it seems reasonable that they would use oceangraphic currents to further reduce movement costs  \citep{ruxton2004energetic}.
 %and are able to maintain a type of endothermy from thier movements \citep{bernal2015sharks}.
@@
 Facultative scavenging is seen in many marine groups, including species of extant sharks like white sharks \citep[known to feed on whale carcasses;][]{fallows2013white}, Greenland sharks \citep[feeding on seals;][]{watanabe2012slowest}, and sixgill sharks \citep{anderson2016impact}.
-There is evidence too of scavening in extinct species, where shark teeth have been found in the remains of dinosaurs, mosasaurs and Pliocene mysticete whales \citep[5.3 to 3.6 Mya;]{schwimmer1997scavenging,ehret2009caught}.
+The former which grow up to 6 metres long, can be sustained by 30 kg of whale blubber for over six weeks \citep{carey1982temperature}. 
+There is evidence too of scavening in extinct species, where shark teeth have been found in the remains of dinosaurs, mosasaurs and Pliocene (5.3 to 3.6 Mya) mysticete whales \citep{schwimmer1997scavenging,ehret2009caught}.
 
@@
 That said, aquatic endotherms have and do scavenge. 
-For instance, early whales such as \textit{Basilosaurus} seem to fit into the same niche as killer whales \textit{Orcinus orca} and we have some evidence for scavenging in both \citep{fahlke2012bite,Whitehead415}.
+For instance, early whales such as \textit{Basilosaurus} (38 to 36.5 Mya) seem to fit into the same niche as killer whales (\textit{Orcinus orca}) and we have some evidence for scavenging in both \citep{fahlke2012bite,Whitehead415}.
 %\cite{williams1999evolution} notes, ``free-ranging animals must contend with the total energetic expenditure associated with supporting basic biological functions as well as with moving the body and appendages through the environment."
@@
 % These constraints are likely to mean that many ectotermic terrestrial species, which cannot maintain an upright posture would depend on a relatively low level of scavenging (I'm not sure this is true though).
-This would also be true of extinct species. 
-It is during the Permian (298.9 - 252.17 Mya) that we have the earliest evidence of vertebrate scavenging where a temnospondyl amphibian fed on the carcass of \textit{Varanops}, a predatory synapsid of the time \citep{reisz2006articulated}.
+This would also have been true of extinct species with the same physiology. 
 %TG: I guess this is a comment too: Snakes are a bit different are more effecent at moving I think - lots of scavanging. 
-Both these groups are more likely to depend on low energetic costs in order to compensate for low encounter rates due to their inability to move. 
+%Both these groups are more likely to depend on low energetic costs in order to compensate for low encounter rates due to their inability to move. 
 It is with the evolution of endothermy in the therapsid-mammal lineage \citep{clarke2010temperature} that terrestrial vertebrates would have gained the ability to range more widely, a vital component in seeking out carrion.
+Although the earliest evidence of vertebrate scavenging comes from the Permian (298.9 - 252.17 Mya) where a temnospondyl amphibian fed on the carcass of \textit{Varanops}, a predatory synapsid of the time \citep{reisz2006articulated}.
 
@@
 This allows us to make a broad distinction between the ambush strategies of cats and the pursuit/ pounce strategies of dogs, the latter being more suited to scavenging \citep{janis2014forelimb}. 
-We can (and have) use(d) these insights to compare extant terrestrial species to their prehistoric forebears given the dominance of mammalian carnviores since the Eocene (56-33.9 Million years ago; Mya) where the order split into the Caniforma and Feliforma \citep{van1987skeletal}.
-To take one example, \cite{anyonge1996locomotor} found that \textit{Nimravides}, a genus of sabretooth cat from the Miocence (10.3 to 5.3 Mya) were likely to have been ambush predators which would argue against them taking a lot of carrion.
+These insights allow us to compare extant terrestrial species to their prehistoric forebears given the dominance of mammalian carnviores since the Eocene (56-33.9 Million years ago; Mya) where the order split into the Caniforma and Feliforma \citep{van1987skeletal}.
+To take one example, \cite{anyonge1996locomotor} found that \textit{Nimravides}, a genus of sabretooth cat from the Miocence (10.3 to 5.3 Mya), were likely to have been ambush predators which would argue against them taking a lot of carrion.
 %The order Carnivora saw its origins in the Middle Eocene  %TG: a phylo side note: the correct sentence would be the "The stem Carnivora saw its origins ...
@@
 
-The best-known terrestrial bipeds are the dinosaurs and unsurprisingly, given their enduring appeal, the prevalence of scavenging has been explored in the carnivorous theropods.
+Aside from humans and our allies, the best-known terrestrial bipeds are the dinosaurs and unsurprisingly, given their enduring appeal, the prevalence of scavenging has been explored in the carnivorous theropods.
 These were the dominant terrestrial carnivores for most of the Mesozoic Era (252.17 - 66 Mya) and ranged from the chicken-sized to the whale-sized, all of which were bipedal.
@@
 Among them are the Turkey Vultures (\textit{Cathartes aura}) which were able to locate 90\% of baits set out in a tropical forest \citep{houston1986olfaction}.
-An atuned sense of smell is obviously useful in detecting decaying carrion from the air.
+An atuned sense of smell is obviously useful in detecting decaying carrion from the air over a heavily forested habitat.
 % This would be impossible for the visually reliant old world species.
@@
 % Depending on the species, a carcass in water either floats or descends to the sea floor \citep{Whitehead415}.
-In contrast to the air, aquatic species have to contend with the low-light environment where visual detection distances are far lower (< 100 m) than they would be in the air.
+In contrast to the air, aquatic species have to contend with a low-light environment where visual detection distances are far lower (< 100 m) than they would be in the air.
 As such, aquatic animals detect resources through chemo- and mechanoreception more so than through vision \citep{ruxton2004energetic}.
@@
 \section*{Handling Time}
-Since carrion is not dispatched directly, often the most easily accessible and choicest components of the carcass will be missing or, if present, will be fought over.
-Being able to extract nutrients from remnants gives a scavenger a great advantage.
+Since the food a scavenger depends on is not dispatched directly, often the most easily accessible and choicest components of the carcass will be missing or, if present, will be subject to decay as well as competition.
+So being able to overcome competitors, maximse the nutrient gain from the remnants, and survive long enough between meals are all essential parts of carcass handling time. 
+% Being able to extract nutrients from remnants gives a scavenger a great advantage.
+
 %Thus, the bone crushing ability of hyenas and others reveals another useful scavenger trait. %TG: how about the following for a smoother transition:
-Thus, some phenotypic adaptations such as the ability to crush bone (as seen in hyenas) reveal another useful scavenger trait.
+In the ability to eat bone scavengers have arrived at a way to feed on a resource that is typcially too hard for many predators to process. 
+% Thus, some phenotypic adaptations such as the ability to crush bone reveal another useful scavenger trait.
 Osteophagy is known across a range of terrestrial carnivores and given that some fat-rich mammalian bones have an energy density (6.7 kJ/g) comparable with that of muscle tissue, it makes skeletal remains an enticing resource \citep{brown1989study}.
@@
 In some historical hominin-inhabited areas there were a greater number of felids than hyenids.
-Again, this is significant because hyenas are likely to have left far less flesh on a carcass than a felid such as a sabretooth enabling contemperaneous hominins to benefit \citep{pobiner2015new}.
+Again, this is significant because hyenas are likely to have left far less flesh on a carcass than a felid such as a sabretooth, enabling contemperaneous hominins to benefit \citep{pobiner2015new}.
 The use of tools and the cooperative nature of hominins meant they could likely get a substantial part of their energetic requirements through scavenging depending on their environment \citep{moleon2014humans}.
@@
 Same applies for most of the physiques in the locomotion part, it is not essential in the narrative to have the number (yet important to link to them!).
-This is noted as being powerful enough to break open skeletal material \citep{rayfield2001cranial}.
+\cite{rayfield2001cranial} note this as being powerful enough to break open skeletal material \citep{rayfield2001cranial}.
 
@@
 
-The support of water allows for many aquatic species to reach large sizes thus granting the benefits. 
-For example, white sharks, which grow up to 6 metres long, can be sustained by 30 kg of blubber for over six weeks \citep{carey1982temperature}.
+The support of water allows for many aquatic species to reach large sizes thus granting its benefits. 
+\cite{collins2005trends} found "contrasting relationships between size (body mass) and depth in the scavenging and predatory demersal ichthyofauna". 
+Predatory species saw a reduction in body mass with depth whereas the reverse trend was true for scavengers. 
+This, the authors pointed out, is because randomly distributed carrion is better exploited by fish with larger body sizes owing to starvation resistance.
 
 Given the advantages of size, we would expect this trait to be selected for even in the case of weight-constrained scavenging fliers.
-Wandering Albatrosses (\textit{Diomedea exulans}), cinereous vultures (\textit{Aegypius monachus}) and condors (\textit{Vultur gryphus}, \textit{Gymnogyps californianus}) all have body masses that can exceed 10 kg and represent some of the heaviest bird species capable of flight \citep{weimerskirch1992reproductive,ferguson2001raptors,donazar2002effects}.
+This is true for Wandering Albatrosses (\textit{Diomedea exulans}), cinereous vultures (\textit{Aegypius monachus}) and condors (\textit{Vultur gryphus}, \textit{Gymnogyps californianus}) who all have body masses that can exceed 10 kg and represent some of the heaviest bird species capable of flight \citep{weimerskirch1992reproductive,ferguson2001raptors,donazar2002effects}.
 And as we have noted the Azhdarchid pterosaurs were far bigger again, with estimated body masses of over 200 kg \citep{witton2010size}.
@@
 
-In addition to reducing locomotory costs we would expect adaptations that reduce energetic costs of maintenance to be selected for in scavengers because it would maximise the benefit derived from such a sporadic food source.
+Because of the random nature of carrion we would expect adaptations that reduce energetic costs of maintenance to be selected for in scavengers as it would maximise the benefit derived from such a sporadic food source.
 Extant reptiles possess an advantage here, in that over the course of a year their food requirements can be 30 times lower than an endotherm of equal size \citep{Nagy1621}.
@@
 Any animal that could find such whale carcasses is unlikely to have ignored other types of carrion.
-Although no aquatic species have ever exceeded the size of whales, some enormous animals have evolved in this environment before the evolution of whales, including \textit{Leedsichthys}, a bony fish from the Middle Jurassic (174.1-163.5 Mya), that weighed in excess of 20 tonnes.
+Although no aquatic species have ever exceeded the size of whales, some enormous animals have evolved in this environment before the evolution of cetaceans, including \textit{Leedsichthys}, a bony fish from the Middle Jurassic (174.1-163.5 Mya) and the aquatic Mesozoic reptiles, the plesiosaurs, pliosaurs and ichtyosaurs, that could all exceed 15 metres in length \citep{ruxton2011zoology}.
 So, despite being unlikely, the energetic feasibility of a marine scavenger that specialises on large carcasses has a long history.
@@
 The animals of today, while often different (sometimes radically so) to their ancestors, can be used to make informed comparisons to extinct species. 
-We have used this technique to give insight into the drivers of scavenging across terrestrial vertebrates through time.
+We have used this technique to give insight into the drivers of scavenging across vertebrates through time.
 In common with any other forager be they grazer, browser or predator, scavengers past and present have had to balance their energetic costs with the gains of food. 

### 2016-04-17 natural history of scavenging.tex
Some (minor) edits

@@
 
-\subsection*{The Difficulty of Scavenging}
+\subsection*{The challenges of Scavenging} %TG: challenges or difficulty? up to you.
 The chief hurdle to scavenging is finding a resource that is often difficult to predict in space and time.
@@
 All foraging processes depend on the encounter rate between consumer and resource. 
-In the simplest case, this rate can be thought of in terms of a gas diffusion model where the movement of two agents depends only on their relative speed.
+In the simplest case, this rate can be thought of in terms of a gas molecules diffusion model where the movement of two agents (i.e. predators and preys) depends only on their relative speed.
+Vertebrates do more than simply bump into resources like gas molecules though, because predators can actively detect preys through their sensory abilities.
+%TG: I like the gas diffusion analogy but it might be adding too much text and be not essential.
+I've added some precisions though.
 As carcasses are stationary, the relative speed between a scavenger and carrion is only dependent on the movement of the scavenger.
-Vertebrates do more than simply bump into resources like gas molecules though, because they can actively detect them through their sensory abilities. 
 As such, scavenging potential is strongly affected by search rates, which are determined by both species' physiology and the dimensionality of the environment \citep{pawar2012dimensionality}.
@@
 \subsection*{Locomotion}
-Scavenging depends more on the ability to efficiently move over large areas than does predation because of the inherent unpredictability of carrion. 
+Because of the inherent unpredictability of carrion, scavenging depends more on the ability to efficiently move over large areas than predation.
 This generally requires an efficient transfer of metabolic energy into movement which relies on both physiology (i.e.
@@
 And, although powered flight is energetically expensive, species like vultures can exploit air currents using their large wingspans which allows them to soar at a cost of only twice their metabolic rate \citep{hedenstrom1993migration,spivey2014analysing}.
-%TG: I didn't get the meaning of this sentence, is soaring doubling the metabolic rate?
-or is it a word misplacement ("while only doubling their metabolic rate")?
-and if so, how does that make the point?
 By depending on thermal air flows these species can forage over vast ranges \citep{spiegel2013factors}. 
-An analagous mode of locomotion is also exploited by seabirds, who use strong ocean winds to search large areas of the oceans \citep{norberg2012vertebrate,thaxter2012seabird}. 
+An analogous mode of locomotion is also exploited by seabirds, who use strong ocean winds to search large areas of the oceans \citep{norberg2012vertebrate,thaxter2012seabird}. 
 While many species of seabird are likely primarily predators, it seems that albatrosses, who can range many hundreds of kilometres, take a substantial amount of carrion in their diet \citep{croxall1994dead}.
-This is typically in the form of squid, which float on the surface, allowing the birds to readily pluck their remains out of the water \citep{croxall1994dead}.
+This is typically in the form of squid carcases, which float on the surface, allowing the birds to readily pluck their remains out of the water \citep{croxall1994dead}.
 
-The two families from which modern vultures arose, the Accipitridae and Cathartidae, appear during the Palaeocene \citep[66 - 56 Million years ago (Mya); ][]{Jetz2012, Jarvis2014} % KH and are likley to have being soaring fliers as demonstrated by their large wingspans (Is this true?).
+The two families from which modern vultures arose, the Accipitridae and Cathartidae, appear during the Palaeocene \citep[66 - 56 Million years ago (Mya); ][]{Jetz2012, Jarvis2014}.
+% KH and are likley to have being soaring fliers as demonstrated by their large wingspans (Is this true?).
 (Is there any other form of soaring?).
@@
 Avian flight is cheaper than either walking or running \citep{tucker1975energetic}.
-Even taking account of maintenance costs, this still bears out, where total cost of movement (J kg\textsuperscript{-1} m\textsuperscript{-1}) scales according to 5.2 $\times$ body mass (kg) \textsuperscript{-0.23} for fliers and 10.7 $\times$ body mass (kg) \textsuperscript{-0.32} for runners \citep{williams1999evolution}.
-
+%TG: I'd rather put the following part as an appendix or foot note. Reader can take our (Tucker's) word for it and then check the math to see that even with maintenance and blablabla flying wins.
+%Even taking account of maintenance costs, this still bears out, where total cost of movement (J kg\textsuperscript{-1} m\textsuperscript{-1}) scales according to 5.2 $\times$ body mass (kg) \textsuperscript{-0.23} for fliers and 10.7 $\times$ body mass (kg) \textsuperscript{-0.32} for runners \citep{williams1999evolution}.
+%TG: This also allows to link directly to the next argument in the same paragraph: facultative active flying scavengers.
 We know that many extant birds exist as facultative scavengers because storks, raptors and corvids all take substantial quantities of carrion in their diet \citep{kendall2013alternative}. 
@@
 By the way, about pterosaurs (though it might not be necessary to mention it), most of them where tiny little feckers like Archaeopteryx or even smaller.
+
 The importance of efficient flying over large areas may explain the lack of scavenging behaviour in bats as they are generally nocturnal, a time when they would receive no aid from convective air currents \citep{norberg2012vertebrate}.
 %smaller and less efficient in comparison to larger birds (speculating here but I wouldnt be surprised), %AK Don't think it is the case that bats are less efficient than birds in terms of flight 
-That said, \textit{Necromantis} (``death-eater''), a large bat from the middle to late Eocene (56 to 33.9 Mya) had a robust cranio-mandibular morphology, and is a likely candidate for scavenging behaviour \citep{Weithofer_Necromantis_1887,Hand_Necromantis_2012}.% amongst the x amount of known bat species \citep{Weithofer_Necromantis_1887,Hand_Necromantis_2012}.
+That said, \textit{Necromantis} (``death-eater''), a large bat from the middle to late Eocene (56-33.9 Mya) had a robust cranio-mandibular morphology, and is a likely candidate for scavenging behaviour \citep{Weithofer_Necromantis_1887,Hand_Necromantis_2012}.% amongst the x amount of known bat species \citep{Weithofer_Necromantis_1887,Hand_Necromantis_2012}.
 %TG: we can cut this number of bat species to avoid getting into stamp collecting debate.
@@
 Similar to aerial species, aquatic scavengers have a locomotory benefit because water is a medium that is conducive to low-cost movement \citep{tucker1975energetic}.
-In fact, the total cost of movement (again in J kg\textsuperscript{-1} m\textsuperscript{-1}) in (salmonid) fish is lower than either running or flying where it scales according to 2.15 $\times$ body mass (kg) \textsuperscript{-0.25} \citep{williams1999evolution} with only \textit{soaring} flight likely to surpass it.
+%TG: Same as comment before, maybe put the following as a footnote or as a supplementary.
+% In fact, the total cost of movement (again in J kg\textsuperscript{-1} m\textsuperscript{-1}) in salmonids is lower than either running or flying where it scales according to 2.15 $\times$ body mass (kg) \textsuperscript{-0.25} \citep{williams1999evolution} with only \textit{soaring} flight likely to surpass it.
 %KH(I assume the reference is for powered flight, again this needs to be more precise).  
@@
 Though sharks perhaps best resemble the large soaring fliers as they depend on large pectoral fins in order to maintain lift as they swim. 
-Many shark species show large ranges, e.g.
-great white sharks (\textit{Carcharodon carcharias}) \citep{bruce2006movements}, and it seems reasonable that they would use oceangraphic currents to further reduce movement costs  \citep{ruxton2004energetic}.
+Many shark species have large foraging ranges \citep[e.g.
+the great white sharks \textit{Carcharodon carcharias};][]{bruce2006movements} and it seems reasonable that they would use oceangraphic currents to further reduce movement costs \citep{ruxton2004energetic}.
 %and are able to maintain a type of endothermy from thier movements \citep{bernal2015sharks}.
 %TG: I guess the following is a comment: (I know the Greenland shark is the slowest swimmer).
-Facultative scavenging is seen in many marine groups, including species of extant sharks like white sharks \citep[known to feed on whale carcasses;][]{fallows2013white}, Greenland sharks \citep[feeding on seals;][]{watanabe2012slowest}, and sixgill sharks \citep{anderson2016impact}.
+In fact, facultative scavenging is seen in many selachian groups, including species of extant sharks like white sharks \citep[known to feed on whale carcasses;][]{fallows2013white}, Greenland sharks \citep[feeding on seals;][]{watanabe2012slowest}, and sixgill sharks \citep{anderson2016impact}.
 The former which grow up to 6 metres long, can be sustained by 30 kg of whale blubber for over six weeks \citep{carey1982temperature}. 
-There is evidence too of scavening in extinct species, where shark teeth have been found in the remains of dinosaurs, mosasaurs and Pliocene (5.3 to 3.6 Mya) mysticete whales \citep{schwimmer1997scavenging,ehret2009caught}.
+There is evidence too of scavening in extinct species, where shark teeth have been found in the remains of dinosaurs, mosasaurs and Pliocene mysticete whales \citep[5.3-3.6 Mya; ][]{schwimmer1997scavenging,ehret2009caught}.
 
@@
 That said, aquatic endotherms have and do scavenge. 
-For instance, early whales such as \textit{Basilosaurus} (38 to 36.5 Mya) seem to fit into the same niche as killer whales (\textit{Orcinus orca}) and we have some evidence for scavenging in both \citep{fahlke2012bite,Whitehead415}.
+For instance, early whales such as \textit{Basilosaurus} (38-36.5 Mya) seem to have fit into the same niche as killer whales (\textit{Orcinus orca}) and we have some evidence for scavenging in both \citep{fahlke2012bite,Whitehead415}.
 %\cite{williams1999evolution} notes, ``free-ranging animals must contend with the total energetic expenditure associated with supporting basic biological functions as well as with moving the body and appendages through the environment."
@@
 %% So I think this should stay pitched as an endo/ ectothermy thing rather than posture 
-Terrestrial environments are the most energetically costly in which to move, which may be due to the low muscular efficiency of running \citep{tucker1975energetic} as well as the relative inefficiency of gas exchange in mammals cf.
-birds and fish \citep{williams1999evolution}.
+Terrestrial environments are the most energetically costly in which to move, which may be due to the low muscular efficiency of running \citep{tucker1975energetic} as well as the relative inefficiency of gas exchange in mammals \citep[cf.
+birds and fish ][]{williams1999evolution}.
+%TG: Hmm I get the point of gas exchange efficiency but (1) it should be more developed if left like that (i.e.
+what the flip is meant by gas efficient exchanges) and (2) that works for mammals, but in terms of terrestrial that ignores birds, squamates, crocodiles and amphibians.
+%
+%TG: It's missing a transition here, we go from gas exchanges to posture.
+Maybe we need something like "Among terrestrial scavengers, most are mammals.
+This might be due to the posture blablabalba." But then it doesn't connect logically with the sentence before: "mammals have are inefficient...".
+Maybe one way is to just remove the gas exchange and start with something like:
+%Terrestrial environments are the most energetically costly in which to move, which may be due to the low muscular efficiency of running \citep{tucker1975energetic}.
+%Most terrestrial species that have been observed (or are suspected to be) scavenging are mammals.
+%TG: and then link to the posture hypothesis (with the "meh, we're not sure" at the end).
+%
 The idea that the evolutionary transition in posture from the sprawling gait of reptiles to the erect posture of mammals conferred a huge advantage to the latter is often mentioned.
@@
 Although the earliest evidence of vertebrate scavenging comes from the Permian (298.9 - 252.17 Mya) where a temnospondyl amphibian fed on the carcass of \textit{Varanops}, a predatory synapsid of the time \citep{reisz2006articulated}.
+%TG: I do like the way all our paragraphs end up with a "But wait, there's an exception!"
 
 Modern endothermic mammals can sustain longer periods of energetically expensive activity \citep{bennett1979endothermy} resulting in larger foraging ranges. 
-To quantify this effect with a simple example we can turn to some allometric relationships relating sustainable travelling speed to body mass \citep{ruxton2004obligate}.
-If we insert these into a foraging radius model \citep{Enstipp2006Energetics} for a 12 hour foraging day it shows that while a 10 kg reptile can range 6.5 km, an equally sized mammal can range nearly 33 km (See appendix).
-For a foraging scavenger, this ability translates into a greater area searched for food.
-
+%TG: Same comment as for the other math, I'd rather put that as a footnote or in the appendix (the whole yoke).
+Again, it seems a bit disproportionate that for some example, we just dumb a line (or even an example in brackets) and that for some others, we explain the example in details.
+% To quantify this effect with a simple example we can turn to some allometric relationships relating sustainable travelling speed to body mass \citep{ruxton2004obligate}.
+% If we insert these into a foraging radius model \citep{Enstipp2006Energetics} for a 12 hour foraging day it shows that while a 10 kg reptile can range 6.5 km, an equally sized mammal can range nearly 33 km (See appendix).
+% For a foraging scavenger, this ability translates into a greater area searched for food.
 Today, terrestrial scavenging in the mammals is probably best known in an African context where hyenas, jackals and lions all take sizeable proportions of carrion in their diet.
@@
 In contrast, big cats like leopards (\textit{Panthera pardus}) rely on ambush \citep{pennycuick1995radius}. 
-This allows us to make a broad distinction between the ambush strategies of cats and the pursuit/ pounce strategies of dogs, the latter being more suited to scavenging \citep{janis2014forelimb}. 
-These insights allow us to compare extant terrestrial species to their prehistoric forebears given the dominance of mammalian carnviores since the Eocene (56-33.9 Million years ago; Mya) where the order split into the Caniforma and Feliforma \citep{van1987skeletal}.
+This allows us to make a broad distinction between the ambush strategies of cats-like and the pursuit/ pounce strategies of dogs-like%TG: I've added the "-like" to underline that we're not talking about the phylogeny here (i.e.
+not Felidae (incl.
+hyena) vs Canidae)
+, the latter being more suited to scavenging \citep{janis2014forelimb}. 
+These insights allow us to compare extant terrestrial species to their prehistoric forebears given the dominance of mammalian carnviores since the Eocene (56-33.9 Mya) where the order split into the Caniforma and Feliforma \citep{van1987skeletal}.
 To take one example, \cite{anyonge1996locomotor} found that \textit{Nimravides}, a genus of sabretooth cat from the Miocence (10.3 to 5.3 Mya), were likely to have been ambush predators which would argue against them taking a lot of carrion.
@@
 %TG: missing some transition here. Maybe something short like "Further back in time" or "Previous to that"...
-Although the evolution of bipedal movement was significant in that it freed up the forelimbs for other purposes e.g.
-climbing, tool-use, wing development etc.
-it does not differ radically in cost from quadrupedal locomotion \citep[and references therein]{williams1999evolution}.
+Although the evolution of bipedal movement was significant in that it freed up the forelimbs for other purposes (e.g.
+climbing, tool-use, wing development etc.) it does not differ radically in cost from quadrupedal locomotion \citep[][and references therein]{williams1999evolution}.
 For instance, \cite{alexander2004bipedal} shows that, in the case of humans, we are more economical than predicted while walking and less so while running according to predicted costs of terrestrial movement.
@@
 They are quite alien to anything we know today which restricts our ability to understand their ecology far more so than extinct mammals \citep{weishampel2004dinosauria}.
-Of relevance, are the questions that still persist about their metabolism, with the latest evidence suggesting they were mesothermic i.e.
-intermediate to ecto- and endotherms \citep{grady2014evidence}.
+Of relevance, are the questions that still persist about their metabolism, with the latest evidence suggesting they were mesothermic \citep[i.e.
+intermediate to ecto- and endotherms;][]{grady2014evidence}.
 We do know that they walked with the erect gait of mammals or birds rather than the sprawling gait of lizards and that they were most likely facultative scavengers \citep{weishampel2004dinosauria,depalma2013physical}.
-%TG: and cite "Kane et al in press" no?
-I think these people did some kind of work on that matter.
 Taken together, this implies dinosaurs had a foraging range that fell in between that of modern terrestrial mammals and reptiles.
+
+%TG: Dudes! Cite your godamn Am Nat paper in this paragraph! It's totally more relevant than Weishampel's book!
 
@@
 
-Species capable of flight have effectively added an extra spatial dimension, i.e. the vertical component, to their sensory environment over land animals.
+Species capable of flight have effectively added an extra spatial dimension (i.e. the vertical component) to their sensory environment over land animals.
 This allows them to look down on a landscape where they are unencumbered by obstacles that would obstruct the view of a terrestrial scavenger.
@@
 It really lowers down the quality of our claim and is totally un-necessary (and actually probably wrong: lot's of fruit bats (like flying foxes) do not echolocate and thus have to find their fruits among the flipping tropical folliage using their eyes and noses only)...
-We can contrast this with many bats whose reliance on echolocation would not lend itself to discovering immobile carrion.
+We can contrast this with other flying vertebrates such as many bats whose reliance on echolocation would not lend itself to discovering immobile carrion.
 %
@@
 Again, returning to vultures, the genus \textit{Gyps} consists of highly social and colonially nesting species \citep{fernandez2015density}.
-These behaviours allow them forage far more efficiently because one bird can scrounge information on the location of food from another successful forager \citep{KaneVul}.
+These behaviours allow them to forage far more efficiently because one bird can scrounge information on the location of food from another successful forager \citep{KaneVul}.
 This efficiency has been exploited by mammals such as hyenas who are known to follow groups of vultures \citep{jones2015african}.
@@
 % three species of vultures within the new world family Cathartidae, (genus \textit{Cathartes}), have well developed olfactory systems \citep{AR:AR22815}.
-Among them are the Turkey Vultures (\textit{Cathartes aura}) which were able to locate 90\% of baits set out in a tropical forest \citep{houston1986olfaction}.
+Among them are the turkey vultures (\textit{Cathartes aura}) which were able to locate 90\% of baits set out in a tropical forest \citep{houston1986olfaction}.
 An atuned sense of smell is obviously useful in detecting decaying carrion from the air over a heavily forested habitat.
@@
 % Depending on the species, a carcass in water either floats or descends to the sea floor \citep{Whitehead415}.
-In contrast to the air, aquatic species have to contend with a low-light environment where visual detection distances are far lower (< 100 m) than they would be in the air.
+In contrast to the air, aquatic species have to contend with a low-light environment where visual detection distances are far lower ($<$ 100 m) than they would be in the air.
 As such, aquatic animals detect resources through chemo- and mechanoreception more so than through vision \citep{ruxton2004energetic}.
@@
 Smell seems to be the primary means of carcass detection in sharks as well. 
-\cite{fallows2013white} found that wind speed %TG: wind speed? Or current speed?
- determined the number of sharks feeding at whale carcasses, indicating they were dependent on detecting the odours from the decaying whales.
+\cite{fallows2013white} found that wind speed determined the number of sharks feeding at whale carcasses, indicating they were dependent on detecting the odours from the decaying whales.
 
@@
 %Thus, the bone crushing ability of hyenas and others reveals another useful scavenger trait. %TG: how about the following for a smoother transition:
-In the ability to eat bone scavengers have arrived at a way to feed on a resource that is typcially too hard for many predators to process. 
+In the ability to eat bone scavengers have arrived at a way to feed on a resource that is typically too hard for many predators to process. 
 % Thus, some phenotypic adaptations such as the ability to crush bone reveal another useful scavenger trait.
 Osteophagy is known across a range of terrestrial carnivores and given that some fat-rich mammalian bones have an energy density (6.7 kJ/g) comparable with that of muscle tissue, it makes skeletal remains an enticing resource \citep{brown1989study}.
-This ability reached its zenith among hyenas with the evolution of the 110 kg \textit{Pachycrocuta brevirostris} during the Pliocene \citep[3.6 - 2.58 Mya; ][]{palmqvist2011giant}.
+This ability reached its zenith among hyenas with the evolution of the estimated 110 kg \textit{Pachycrocuta brevirostris} during the Pliocene \citep[3.6 - 2.58 Mya; ][]{palmqvist2011giant}.
 Indeed, their extinction has been blamed on the decline of sabretooth cats (Machairodontinae), the unique skull morphology of the latter meant they would leave a large amount of food on a carcass for would-be scavengers \citep{palmqvist2011giant}.
@@
 The question of where our ancestors placed on the hunter-scavenger axis during the Plio-Pleistocene has been a matter of debate for years \citep{dominguez2002hunting}.
-A recent study investigating potential scavenging opportunities for hominins in Kenya found that, even when discounting bone material, there is a substantial amount of scavengeable %TG: is that an adjective?
-meat left on predated remains; sufficient to sustain the requirements of an adult male \textit{Homo erectus} \citep{pobiner2015new}.
+A recent study investigating potential scavenging opportunities for hominins in Kenya found that, even when discounting bone material, there is a substantial amount of scavengeable meat left on predated remains; sufficient to sustain the requirements of an adult male \textit{Homo erectus} \citep{pobiner2015new}.
 In some historical hominin-inhabited areas there were a greater number of felids than hyenids.
@@
 The use of tools and the cooperative nature of hominins meant they could likely get a substantial part of their energetic requirements through scavenging depending on their environment \citep{moleon2014humans}.
-%TG: I've removed the intelligence bit from the sentence.
 
-In Mesozoic systems some extremely large theropod dinosaurs had a morphology indicative of an ability to process bone e.g.
-the robust skull and dentition of \textit{Tyrannosaurus rex} \citep{hone2010feeding}.
+In Mesozoic systems some extremely large theropod dinosaurs had a morphology indicative of an ability to process bone \citep[e.g.
+the robust skull and dentition of \textit{Tyrannosaurus rex}][]{hone2010feeding}.
 There is direct evidence that \textit{T.
 rex} did this in the form of distinctive wear marks on its tooth apices \citep{farlow1994wear,schubert2005wear} and the presence of bone fragments in its coprolites \citep{chin1998king}.
-The animal also had an enormous bite force, with one estimate putting it at 57000 Newtons \citep{bates2012estimating}.
-%TG: A comment about the future cleaning up phase.
-I think mentions like the 570000 Newtons are not essential and could go to some supplementary files.
-Same applies for most of the physiques in the locomotion part, it is not essential in the narrative to have the number (yet important to link to them!).
-\cite{rayfield2001cranial} note this as being powerful enough to break open skeletal material \citep{rayfield2001cranial}.
+The animal also had an enormous bite force, with one estimate putting it at 57000 Newtons \citep{bates2012estimating} which would have been powerful enough to break open skeletal material \citep{rayfield2001cranial}.
 
@@
 Much work has focused on the existence of scavenging in dinosaurs by using simple energetics approaches that typically focused on a single species namely \textit{T.
-rex} \citep{ruxton2003could,carbone2011intra} but a recent modelling study investigated its prevalence across a range of body sizes.
+rex} \citep{ruxton2003could,carbone2011intra} but a recent modelling study investigated its prevalence across a range of body sizes.%TG: missing THE cite here!
 
@@
 Given the advantages of size, we would expect this trait to be selected for even in the case of weight-constrained scavenging fliers.
-This is true for Wandering Albatrosses (\textit{Diomedea exulans}), cinereous vultures (\textit{Aegypius monachus}) and condors (\textit{Vultur gryphus}, \textit{Gymnogyps californianus}) who all have body masses that can exceed 10 kg and represent some of the heaviest bird species capable of flight \citep{weimerskirch1992reproductive,ferguson2001raptors,donazar2002effects}.
-And as we have noted the Azhdarchid pterosaurs were far bigger again, with estimated body masses of over 200 kg \citep{witton2010size}.
+This is true for wandering albatrosses (\textit{Diomedea exulans}), cinereous vultures (\textit{Aegypius monachus}) and condors (\textit{Vultur gryphus}, \textit{Gymnogyps californianus}) who all have body masses that can exceed 10 kg and represent some of the heaviest bird species capable of flight \citep{weimerskirch1992reproductive,ferguson2001raptors,donazar2002effects}.
+Additionally, as we have noted the Azhdarchid pterosaurs were far bigger again, with estimated body masses of over 200 kg \citep{witton2010size}.
 Although \cite{witton2008reappraisal} argued that neck inflexibility and straight, rather than hooked jaw morphology points against pterosaurs existing as \textit{obligate} scavengers, Azhdarchid terrestrial proficiency indicates they would have been comfortable foraging on the ground.
@@
 
-%%%%%%%%%%  AK This annoying paragraph's new home
+%%%%%%%%%%  AK This annoying paragraph's new home TG: well done finding him a home!
 Certainly, scavenging should be particularly attractive to flying species compared to mammals. % Not sure about this paragraph 
@@
 ;)).
-% AK How about flying bird because that's the key thing here 
+% AK How about flying bird because that's the key thing here TG: yep! Agree!
 \cite{houston1974role} observed a group of \textit{Gyps} vultures consuming all of the soft tissue from a 50 kg Grant’s gazelle (\textit{Nanger granti}) in eight minutes. 
@@
 Additionally, their poor terrestrial ability and cost of movement on the ground would also count against them when feeding at a carcass \citep{riskin2006terrestrial,voigt2012terrestrial}.
-% Maybe they scavenge on dead insects but here we're talking about scavenging on real meat right?
 
@@
 The position of a species on the scavenging scale can also be influenced by the availability of carrion in the environment, which is dependent on biotic and abiotic factors.
-%TG: It's actually not better...
-Feel free to change again!
 Aspects including, primary productivity, relief, temperature and competition will all greatly affect scavenging tendency. 

### 2016-04-18 natural history of scavenging.tex
thomas corrections and new figure

@@
 % Chief among these are low-cost locomotion, high detection distances, effective carcass processing and a carrion-producing habitat. 
-We use these components to highlight the morphologies and environments that would have been conducive to scavenging over geological time by focussing on the dominant vertebrate groups of the land, sea and air.
+We use these components to highlight the morphologies and environments that would have been conducive to scavenging over geological time by focusing on the dominant vertebrate groups of the land, sea and air.
 The result is a document on the natural history of scavenging, the first to our knowledge. 
@@
 Some simple heuristics can inform us, for instance, in cases where the prey item was simply too large to have been killed by the ostensible predator \citep{pobiner2008paleoecological}.
-But clearly, a scavenger doesn’t only feed on animals too big for it to have hunted.
+But clearly, a scavenger doesn't only feed on animals too big for it to have hunted.
 The obvious lack of direct behavioural data compounds the difficulty of discerning scavenging among extinct forms.
@@
 As a result, we are now beginning to realise the extent of this behaviour such that, ``in some ecosystems, vertebrates have been documented to assimilate as much as 90\% of the available carrion" \citep{beasley2015vertebrates}.
-This has profound implications for the trophic ecology of these systems and particuarly our models of them.
+This has profound implications for the trophic ecology of these systems and particularly our models of them.
 Even Tenney’s noble big cats are now known to take in a significant portion of carrion in their diet where some lion populations get over 50\% of their meat from carcasses \citep{jones2015african}.
@@
 
-\subsection*{The challenges of Scavenging} %TG: challenges or difficulty? up to you.
+\subsection*{The Challenges of Scavenging} %TG: challenges or difficulty? up to you.
 The chief hurdle to scavenging is finding a resource that is often difficult to predict in space and time.
@@
    \includegraphics[width=1\textwidth]{Summary_figure/Summary_figure_Landscape.pdf}
-\caption{Factors influencing the proportion of scavenging in a vertebrates' diet. Blue lines indicates a reduction in the factor and red lines indicates an increase.}
+\caption{Factors influencing the proportion of scavenging in a vertebrates' diet. Blue lines indicate a reduction in the factor and red lines indicate an increase.}
 %TG: Probably needs some more explanations...
@@
 All foraging processes depend on the encounter rate between consumer and resource. 
-In the simplest case, this rate can be thought of in terms of a gas molecules diffusion model where the movement of two agents (i.e. predators and preys) depends only on their relative speed.
-Vertebrates do more than simply bump into resources like gas molecules though, because predators can actively detect preys through their sensory abilities.
+In the simplest case, this rate can be thought of in terms of a gas diffusion model where the movement of two agents (i.e. predator and prey) depends only on their relative speed.
+Vertebrates do more than simply bump into resources like gas molecules though, because predators can actively detect prey through their sensory abilities.
 %TG: I like the gas diffusion analogy but it might be adding too much text and be not essential.
@@
 
-The two families from which modern vultures arose, the Accipitridae and Cathartidae, appear during the Palaeocene \citep[66 - 56 Million years ago (Mya); ][]{Jetz2012, Jarvis2014}.
+The groups from which modern vultures and seabirds arose, appear during the Palaeocene \citep[66 - 56 Million years ago (Mya); ][]{Jetz2012, Jarvis2014} and Cretaceous \citep{chiappe2006early} respectively.
 % KH and are likley to have being soaring fliers as demonstrated by their large wingspans (Is this true?).
@@
 Azhdarchids with wingspans of 11 metres; ][]{witton2010size} and, notably, appear to have engaged in soaring flight \citep{witton2010size}.
-It seems probable that extinct species used soaring as a means for scavenging.
+It seems probable that extinct species used soaring as a means for scavenging \citep{witton2013pterosaurs}.
 
 While soaring is perhaps the only viable means of locomotion that allows for an obligate, scavenging life-style \citep{ruxton2004obligate}, powered flight is still an efficient means of locomotion. 
-Avian flight is cheaper than either walking or running \citep{tucker1975energetic}.
+Certainly, avian flight is cheaper than either walking or running \citep{tucker1975energetic}.
 %TG: I'd rather put the following part as an appendix or foot note. Reader can take our (Tucker's) word for it and then check the math to see that even with maintenance and blablabla flying wins.
@@
 
-Similar to aerial species, aquatic scavengers have a locomotory benefit because water is a medium that is conducive to low-cost movement \citep{tucker1975energetic}.
+Similar to aerial species, aquatic scavengers have a locomotory benefit because water is a medium that is conducive to low-cost movement \citep{tucker1975energetic,williams1999evolution}.
 %TG: Same as comment before, maybe put the following as a footnote or as a supplementary.
@@
 This has led some researchers to argue for the likelihood of an obligate scavenging fish \citep{ruxton2004energetic,ruxton2005searching}. 
-Interestingly, style of swimmming in fish does not significantly affect the cost of movement \citep{williams1999evolution}. 
+Interestingly, style of swimming in fish does not significantly affect the cost of movement \citep{williams1999evolution}. 
 Though sharks perhaps best resemble the large soaring fliers as they depend on large pectoral fins in order to maintain lift as they swim. 
@@
 The former which grow up to 6 metres long, can be sustained by 30 kg of whale blubber for over six weeks \citep{carey1982temperature}. 
-There is evidence too of scavening in extinct species, where shark teeth have been found in the remains of dinosaurs, mosasaurs and Pliocene mysticete whales \citep[5.3-3.6 Mya; ][]{schwimmer1997scavenging,ehret2009caught}.
+There is evidence too of scavenging in extinct species, where shark teeth have been found in the remains of dinosaurs, mosasaurs and Pliocene mysticete whales \citep[5.3-3.6 Mya; ][]{schwimmer1997scavenging,ehret2009caught}.
 
@@
 %% So I think this should stay pitched as an endo/ ectothermy thing rather than posture 
-Terrestrial environments are the most energetically costly in which to move, which may be due to the low muscular efficiency of running \citep{tucker1975energetic} as well as the relative inefficiency of gas exchange in mammals \citep[cf.
+Terrestrial environments are the most energetically costly in which to move, which may be due to the low muscular efficiency of running \citep{tucker1975energetic} as well as the relative inefficiency of gas exchange in the case of terrestrial mammals \citep[cf.
 birds and fish ][]{williams1999evolution}.
@@
 %
-The idea that the evolutionary transition in posture from the sprawling gait of reptiles to the erect posture of mammals conferred a huge advantage to the latter is often mentioned.
+However, the evolutionary transition in posture from the sprawling gait of reptiles to the erect posture of mammals has often been supposed as conferring a huge advantage to the latter.
 The purported advantages include benefits in terms of speed, efficiency, muscle effort and manoeuvrability \citep{sullivan2015posture}. 
-Despite being intuitive, as \cite{sullivan2015posture} states, most of the hypotheses in favour of this idea remain to be tested in the context of archosaur evolution. 
+Despite being intuitive, \cite{sullivan2015posture} states most of the hypotheses in favour of this idea remain to be tested in the context of archosaur evolution. 
 Metabolic rate however, unquestionably impacts terrestrial species whereby ectotherms such as many modern reptiles, cannot move for sustained periods \citep{bennett1979endothermy}. 
@@
 In contrast, big cats like leopards (\textit{Panthera pardus}) rely on ambush \citep{pennycuick1995radius}. 
-This allows us to make a broad distinction between the ambush strategies of cats-like and the pursuit/ pounce strategies of dogs-like%TG: I've added the "-like" to underline that we're not talking about the phylogeny here (i.e.
+This allows us to make a broad distinction between the ambush strategies of cats-like forms and the pursuit/ pounce strategies of more dog-like forms%TG: I've added the "-like" to underline that we're not talking about the phylogeny here (i.e.
 not Felidae (incl.
@@
 
+Of course, terrestrial animals can also move bipedally. 
 %TG: missing some transition here. Maybe something short like "Further back in time" or "Previous to that"...
@@
 A hypothesis put forth by \cite{sazima1990necrofagia} argued that chemical gradients in water would allow for a relatively easier detection of carrion by snakes.
-This gained some support from \cite{devault2002scavenging}, who found a preponderence of aquatic snake species in their review of this behaviour.
+This gained some support from \cite{devault2002scavenging}, who found a preponderance of aquatic snake species in their review of this behaviour.
 Smell seems to be the primary means of carcass detection in sharks as well. 
@@
 Since the food a scavenger depends on is not dispatched directly, often the most easily accessible and choicest components of the carcass will be missing or, if present, will be subject to decay as well as competition.
-So being able to overcome competitors, maximse the nutrient gain from the remnants, and survive long enough between meals are all essential parts of carcass handling time. 
+So being able to overcome competitors, maximise the nutrient gain from the remnants, and survive long enough between meals are all essential parts of carcass handling time. 
 % Being able to extract nutrients from remnants gives a scavenger a great advantage.
@@
 In some historical hominin-inhabited areas there were a greater number of felids than hyenids.
-Again, this is significant because hyenas are likely to have left far less flesh on a carcass than a felid such as a sabretooth, enabling contemperaneous hominins to benefit \citep{pobiner2015new}.
+Again, this is significant because hyenas are likely to have left far less flesh on a carcass than a felid such as a sabretooth, enabling contemporaneous hominins to benefit \citep{pobiner2015new}.
 The use of tools and the cooperative nature of hominins meant they could likely get a substantial part of their energetic requirements through scavenging depending on their environment \citep{moleon2014humans}.
@@
 Much work has focused on the existence of scavenging in dinosaurs by using simple energetics approaches that typically focused on a single species namely \textit{T.
-rex} \citep{ruxton2003could,carbone2011intra} but a recent modelling study investigated its prevalence across a range of body sizes.%TG: missing THE cite here!
+rex} \citep{ruxton2003could,carbone2011intra} but a recent modelling study investigated its prevalence across a range of body sizes \citep{kane2016body}.
 
-In their work, the authors demonstrated that species of \textit{intermediate} body masses (approximatively 500 kg) would have gained the most benefit from scavenging.
+In their work, the authors demonstrated that species of \textit{intermediate} body masses would have gained the most benefit from scavenging \citep{kane2016body}.
 This was the result of gut capacity limitations and the effects of competition at the carcass.
@@
 Additionally, as we have noted the Azhdarchid pterosaurs were far bigger again, with estimated body masses of over 200 kg \citep{witton2010size}.
-Although \cite{witton2008reappraisal} argued that neck inflexibility and straight, rather than hooked jaw morphology points against pterosaurs existing as \textit{obligate} scavengers, Azhdarchid terrestrial proficiency indicates they would have been comfortable foraging on the ground.
-Indeed, extant Marabou Storks (\textit{Leptoptilos crumenifer}) have a comparable morphology and are noted facultative scavengers \citep{monadjem2012survival} so it is reasonable to believe that certain pterosaurs behaved similarly.
+Although \cite{witton2008reappraisal} argued that neck inflexibility and straight, rather than hooked jaw morphology points against Azhdarchids existing as \textit{obligate} scavengers, their terrestrial proficiency indicates they would have been comfortable foraging on the ground.
+Indeed, extant Marabou Storks (\textit{Leptoptilos crumenifer}) have a comparable morphology and are noted facultative scavengers \citep{monadjem2012survival} so it is reasonable to believe that these pterosaurs behaved similarly.
 
@@
 As \cite{martyniuk2012field} notes these early birds would thus have been predominantly carnivorous, which implies that scavenging would have been a live opportunity cf. their descendants. 
+Among the pterosaurs, \cite{witton2013pterosaurs} makes the case that the istiodactyl pterosaurs were the most likely scavengers of this group based on their potential handling time. 
+The mix of strong and weak features in their skull morphology is indicative of animals that were suited to removing large amounts of flesh from an immobile foodstuff \citep{witton2013pterosaurs}.
 
@@
 Predators were large-bodied too compared to extant mammalian predators \citep{mcnab2009resources}, and so, especially if they were ectothermic, could last longer between meals, rendering scavenging a more attractive behaviour relative to predation.
-Osteophagy may have been even more viable during the Mesozoic era as well because of this skewed body mass distribution of herbviores towards larger sizes \citep{10.1371/journal.pone.0051925}.
+Osteophagy may have been even more viable during the Mesozoic era as well because of this skewed body mass distribution of herbivores towards larger sizes \citep{10.1371/journal.pone.0051925}.
 When we couple this with the fact that skeletal mass scales greater than linearly with body mass \citep{prange1979scaling} there would have been a lot of bone material to consume in the environment provided an animal had the biology to process it \citep{chure1997one}.

### 2016-04-19 natural history of scavenging.tex
figure changes
still not past the locomotion section yet

@@
 All foraging processes depend on the encounter rate between consumer and resource. 
-In the simplest case, this rate can be thought of in terms of a gas diffusion model where the movement of two agents (i.e. predator and prey) depends only on their relative speed.
-Vertebrates do more than simply bump into resources like gas molecules though, because predators can actively detect prey through their sensory abilities.
+In the simplest case, this rate can be thought of in terms of a gas diffusion model where the movement of two agents (i.e.
+predator and prey) depends only on as they can actively detect resources through their sensory abilities.
 %TG: I like the gas diffusion analogy but it might be adding too much text and be not essential.
 I've added some precisions though.
+%KH I'v shortened it some more.
 As carcasses are stationary, the relative speed between a scavenger and carrion is only dependent on the movement of the scavenger.
-As such, scavenging potential is strongly affected by search rates, which are determined by both species' physiology and the dimensionality of the environment \citep{pawar2012dimensionality}.
+%KH There are too many definite statements throughout from my liking.
+For example, not all carcasses are stationary in fact the sinking of carcasses in marine systems is what drives whale fall fauna as they essentially let the carcass come to them.
+I know it might make the text more clunky but I would play it safe and "As carcasses are usually stationary" etc.
+As such, scavenging potential is strongly affected by search rates, which are determined by both species' physiology and what type of environment the scavenger is foraging in.
+%KH is including dimensionality making this too complex as the only search environment that is 3d is for semi-buoyant marine carcasses.
+Also the dimensionality talked about in the vision section is different to this dimensionality (the Dim of the search space for both terrestrial and aerial are both 2D).
+Might be a bit confusing so I changed it to just the physiology and environmental
 
 \subsection*{Locomotion}
-Because of the inherent unpredictability of carrion, scavenging depends more on the ability to efficiently move over large areas than predation.
+Because of the inherent unpredictability of carrion, scavenging heavily depends on the ability to efficiently move over large areas than predation.
+%KH changed "more" to "heavily"
 This generally requires an efficient transfer of metabolic energy into movement which relies on both physiology (i.e.
@@
 And, although powered flight is energetically expensive, species like vultures can exploit air currents using their large wingspans which allows them to soar at a cost of only twice their metabolic rate \citep{hedenstrom1993migration,spivey2014analysing}.
+%KH This and seems out of place a bit, maybe replace with "While powered flight..."
 By depending on thermal air flows these species can forage over vast ranges \citep{spiegel2013factors}. 
@@
 
-The groups from which modern vultures and seabirds arose, appear during the Palaeocene \citep[66 - 56 Million years ago (Mya); ][]{Jetz2012, Jarvis2014} and Cretaceous \citep{chiappe2006early} respectively.
-% KH and are likley to have being soaring fliers as demonstrated by their large wingspans (Is this true?).
+The groups from which these modern soaring birds arose, appear during the Palaeocene \citep[66 - 56 Million years ago (Mya); ][]{Jetz2012, Jarvis2014} and Cretaceous \citep{chiappe2006early} respectively.
+% KH and are likely to have being soaring fliers as demonstrated by their large wingspans (Is this true?).
 (Is there any other form of soaring?).
 AK I don't think this is true from what I've seen
+%KH OK I changed it to these modern soaring birds to emphasize we are still on the tred of low cost flying. However if you don't think its clear enough what this group is feel free to change it back.
 However, soaring flight is likely to be far older than this with avian flight originating in the Late Jurassic (163.5-145 Mya) and vertebrate flight in the Late Triassic (235-201.3 Mya) coincident with the pterosaurs.
@@
 Azhdarchids with wingspans of 11 metres; ][]{witton2010size} and, notably, appear to have engaged in soaring flight \citep{witton2010size}.
-It seems probable that extinct species used soaring as a means for scavenging \citep{witton2013pterosaurs}.
-
+It seems probable that as least some of these extinct species used soaring as a means for scavenging \citep{witton2013pterosaurs}. 
+%KH added "at least some of these" as it was not clear what species used soaring as a means for scavenging.
 While soaring is perhaps the only viable means of locomotion that allows for an obligate, scavenging life-style \citep{ruxton2004obligate}, powered flight is still an efficient means of locomotion. 
@@
 %TG: This also allows to link directly to the next argument in the same paragraph: facultative active flying scavengers.
+%KHI agree avoid as much maths as possible (If only someone would propose some sort of drop down thing for maths in papers.)
 We know that many extant birds exist as facultative scavengers because storks, raptors and corvids all take substantial quantities of carrion in their diet \citep{kendall2013alternative}. 
@@
 %KH(I assume the reference is for powered flight, again this needs to be more precise). 
+%KH again I think its good  to keep away from the maths 
 This has led some researchers to argue for the likelihood of an obligate scavenging fish \citep{ruxton2004energetic,ruxton2005searching}.
-Interestingly, style of swimming in fish does not significantly affect the cost of movement \citep{williams1999evolution}. 
-Though sharks perhaps best resemble the large soaring fliers as they depend on large pectoral fins in order to maintain lift as they swim. 
+
+Sharks are one likely candidate for general scavenging behaviors as their locomotion, which depends on large pectoral fins in order to maintain lift as they swim, resemble large soaring fliers.
+%changed this sentence up a bit 
 Many shark species have large foraging ranges \citep[e.g.
-the great white sharks \textit{Carcharodon carcharias};][]{bruce2006movements} and it seems reasonable that they would use oceangraphic currents to further reduce movement costs \citep{ruxton2004energetic}.
+the great white sharks \textit{Carcharodon carcharias};][]{bruce2006movements} and it seems reasonable that they would use oceanographic currents to further reduce movement costs \citep{ruxton2004energetic}.
 %and are able to maintain a type of endothermy from thier movements \citep{bernal2015sharks}.
@@
 The former which grow up to 6 metres long, can be sustained by 30 kg of whale blubber for over six weeks \citep{carey1982temperature}. 
+%A personal thing but I hate the use of former or latter as I have to break my reading and go back and see which species that is.
+I did'nt change it but I would go with naming the species there and then for clarity.
 There is evidence too of scavenging in extinct species, where shark teeth have been found in the remains of dinosaurs, mosasaurs and Pliocene mysticete whales \citep[5.3-3.6 Mya; ][]{schwimmer1997scavenging,ehret2009caught}.
 
-% KH Other types of swimming in fish, other species propelled by fins (turtles and plesiosaurs etc)
-
-% KH snakes - ie undulating
+Interestingly, style of swimming in fish does not significantly affect the cost of movement \citep{williams1999evolution}, hence it is likely that aquatic species with large ranges may also encounter scavenging opportunities.
+It may be that jaw morphology is more likely to define which species can avail of carcasses.
+%KH I moved this down here as it kinda throws our shark soaring thing out the window (which I like as a linking function).
+However I think this kinda solves our problems in the complexity of swimming by saying apart from large ranges ability to eat is probably more important.
+%KH also is fish an ok term to use, will people get pissy or anything like with reptiles?
 
 % KH endotermy - why it doesn't seem to work
-We might expect then that by combining an aquatic environment and an endothermic metabolism marine mammals would especially prosper as scavengers.
+We might expect then that by combining an aquatic environment and an endothermic metabolism marine mammals would especially prosper as scavengers, especially as this group has perhaps the largest home range of any animals (whales and shit).
 Fossil pinnipeds and cetaceans from 60 Mya %TG: that early???
@@
 Indeed, the total energetic cost is similar to an equivalent terrestrial or aerial mammal \citep{williams1999evolution}.
-This underscores the trade offs between the benefits of endothermy in terms of activity periods and the costs of maintaining such an expensive system. 
+This underscores the trade offs between the benefits of endothermy in terms of activity periods and the costs of maintaining such an energetically costly system. 
 That said, aquatic endotherms have and do scavenge. 
 For instance, early whales such as \textit{Basilosaurus} (38-36.5 Mya) seem to have fit into the same niche as killer whales (\textit{Orcinus orca}) and we have some evidence for scavenging in both \citep{fahlke2012bite,Whitehead415}.
+
+%%KH Is there anything we should add about other extinct groups such as plesiosaurs and https://en.wikipedia.org/wiki/Plesiosauria#Gait
+
 %\cite{williams1999evolution} notes, ``free-ranging animals must contend with the total energetic expenditure associated with supporting basic biological functions as well as with moving the body and appendages through the environment."
@@
 
+\begin{figure}[!htbp]
+\centering
+   \includegraphics[width=1\textwidth]{timeline_figure/timeLine.pdf}
+\caption{The diversity of scavengers through time. Each species has either direct evidence for it being a scavenger or would be positioned high up on our scavenging scale.}
+\label{Timeline}
+\end{figure}
+
 \section*{Conclusion} 

### 2016-04-20 natural history of scavenging.tex
Sorry about the hold up

@@
 
-%%KH Is there anything we should add about other extinct groups such as plesiosaurs and https://en.wikipedia.org/wiki/Plesiosauria#Gait
+%%KH Is there anything we should add about other extinct groups such as plesiosaurs and https://en.wikipedia.org/wiki/Plesiosauria#Gait maybe an air breathing non-airbreathing devide would be good.
 
@@
 %% So I think this should stay pitched as an endo/ ectothermy thing rather than posture 
-Terrestrial environments are the most energetically costly in which to move, which may be due to the low muscular efficiency of running \citep{tucker1975energetic} as well as the relative inefficiency of gas exchange in the case of terrestrial mammals \citep[cf.
-birds and fish ][]{williams1999evolution}.
+Terrestrial environments are the most energetically costly in which to move \citep{tucker1975energetic}.
+Unlike aerial and aquatic environments support must be provided through the animals posture leading to trade offs between low cost of maintain and the efficiency of transport.
+For example, the sprawling posture of many ectotherm groups (except frogs!) would be expected to allow for a lower maintenance cost relating to body support in comparison to groups with the limbs position underneath the body, such as mammals, aves and dinosaurs.
+%KH: I agree with TG and have removed the gas exchange thing as its a physiology thing only applicable to certain groups and not got to do with terrestrial environments really. I also removed the 
+%Also that reference is from 1975, is this up to date?
+
 %TG: Hmm I get the point of gas exchange efficiency but (1) it should be more developed if left like that (i.e.
 what the flip is meant by gas efficient exchanges) and (2) that works for mammals, but in terms of terrestrial that ignores birds, squamates, crocodiles and amphibians.
+
 %
@@
 %
-However, the evolutionary transition in posture from the sprawling gait of reptiles to the erect posture of mammals has often been supposed as conferring a huge advantage to the latter.
-The purported advantages include benefits in terms of speed, efficiency, muscle effort and manoeuvrability \citep{sullivan2015posture}. 
+While the low maintenance costs of a sprawling like posture may reduce the need to scavenge frequently, the ability to search large areas is several restricted in these groups hence likely reducing such opportunities (ref on home range of terrestrial ectoterms).
+For example, a sprawling gait can results in the phenomenon known as Carrier's constraint such that the animal can't move and breathe at the same time because the lateral movements impedes its lungs \citep{carrier1987evolution}.
+On the other hand the more erect body positions and often associated higher basal metabolic rates, allow for sustained locomotion over long periods of time and hence allow such species to search large areas.
+
+The early transition from a sprawling gait, such as seen in Tiktaalik, to the more erect posture of Synapsids and later dinosaurs and mammals, has often been supposed as conferring a huge advantage to the latter.
+%KH Even after changing it I dont like this transition type narrative.
+%KH also needs to be more clear on what the advantage to scavenging is for these groups.
+%The purported advantages include benefits in terms of speed, efficiency, muscle effort and manoeuvrability \citep{sullivan2015posture}. 
+%none of these advantages refer very clearly to scavenging
 Despite being intuitive, \cite{sullivan2015posture} states most of the hypotheses in favour of this idea remain to be tested in the context of archosaur evolution. 
-Metabolic rate however, unquestionably impacts terrestrial species whereby ectotherms such as many modern reptiles, cannot move for sustained periods \citep{bennett1979endothermy}. 
+%KH I'm always wary of deeming something as intuitive,
+
+%Metabolic rate however, unquestionably impacts terrestrial species whereby ectotherms such as many modern reptiles, cannot move for sustained periods \citep{bennett1979endothermy}. 
 % Species which do not support an upright position such as modern reptiles limit their ability to move for sustained periods \citep{bennett1979endothermy}.
-This is exacerbated by their sprawling gait which results in the phenomenon known as Carrier's constraint such that the animal can't move and breathe at the same time because the lateral movements impedes its lungs \citep{carrier1987evolution}.
+
+%KH I think as ectothermy is almost completely associated with sprawling gaits I would tie it up together to avoid repeating something similar here.
+
+%This is exacerbated by their sprawling gait which results in the phenomenon known as Carrier's constraint such that the animal can't move and breathe at the same time because the lateral movements impedes its lungs \citep{carrier1987evolution}.
 % These constraints are likely to mean that many ectotermic terrestrial species, which cannot maintain an upright posture would depend on a relatively low level of scavenging (I'm not sure this is true though).
-This would also have been true of extinct species with the same physiology. 
+%This would also have been true of extinct species with the same physiology. 
 %TG: I guess this is a comment too: Snakes are a bit different are more effecent at moving I think - lots of scavanging. 
 %Both these groups are more likely to depend on low energetic costs in order to compensate for low encounter rates due to their inability to move. 
-It is with the evolution of endothermy in the therapsid-mammal lineage \citep{clarke2010temperature} that terrestrial vertebrates would have gained the ability to range more widely, a vital component in seeking out carrion.
+Whatever the case, it is with the evolution of endothermy in the therapsid-mammal lineage \citep{clarke2010temperature} that terrestrial vertebrates would have gained the ability to range more widely, a vital component in seeking out carrion.
 Although the earliest evidence of vertebrate scavenging comes from the Permian (298.9 - 252.17 Mya) where a temnospondyl amphibian fed on the carcass of \textit{Varanops}, a predatory synapsid of the time \citep{reisz2006articulated}.
@@
 
-Of course, terrestrial animals can also move bipedally. 
+While we can use modern analogues of quadrupedal species, the absence of scavenging in bipedal animals makes this more difficult.
+However, as a groups bipedal locomotion may have been a particularly useful strategy for  searching large areas.
+For instance, \cite{alexander2004bipedal} shows that, in the case of humans, we are more economical than predicted while walking according to predicted costs of terrestrial movement.
+In fact the ability of humans to engage in hunting strategies that involve traversing long ranges, such as persistence hunting, has led to several researches highlighting the possible role of scavenging in early hominids.
+Indeed the low energetics of human walking at least allows for a substantial amount of scavenging in our hominid ancestors (Graemes papers on scavenging).
+%%KH I know this is excluding other hominids so feel free to write a more inclusive line.
+
+%Of course, terrestrial animals can also move bipedally. 
 %TG: missing some transition here. Maybe something short like "Further back in time" or "Previous to that"...
-Although the evolution of bipedal movement was significant in that it freed up the forelimbs for other purposes (e.g.
+%I agree, 
+%Although the evolution of bipedal movement was significant in that it freed up the forelimbs for other purposes (e.g.
 climbing, tool-use, wing development etc.) it does not differ radically in cost from quadrupedal locomotion \citep[][and references therein]{williams1999evolution}.
-For instance, \cite{alexander2004bipedal} shows that, in the case of humans, we are more economical than predicted while walking and less so while running according to predicted costs of terrestrial movement.
+%For instance, \cite{alexander2004bipedal} shows that, in the case of humans, we are more economical than predicted while walking and less so while running according to predicted costs of terrestrial movement.
+%the second sentence does not really follow from the first -
+Aside from humans and our allies, the best-known terrestrial bipeds are the dinosaurs and unsurprisingly, given their enduring appeal, the prevalence of scavenging has been extensively explored in the carnivorous theropods.
+%KH you could reference all the shit on theropod scavenging here if you wanted.
+These were the dominant terrestrial carnivores for most of the Mesozoic Era (252.17 - 66 Mya) and ranged from the chicken-sized to the whale-sized, all of which were bipedal.
+While the locomotory ability of theropods has been debate since their first inception, more recent studies have reconstructed them as relatively mobile animals (John Hutchinsons papers). 
+%See that Pontzer 
+Despite some suggestions that the large size of some forms may have allowed some advantage in scavenging, partially due to the abilty to search large areas (Graemes paper etc), more recent studies have shown that the energetic demands of such large forms meant that scavenging was likely to be more prevalent in mid sized theropods of approximately half and thon (Us and Carbones paper).
 
-Aside from humans and our allies, the best-known terrestrial bipeds are the dinosaurs and unsurprisingly, given their enduring appeal, the prevalence of scavenging has been explored in the carnivorous theropods.
-These were the dominant terrestrial carnivores for most of the Mesozoic Era (252.17 - 66 Mya) and ranged from the chicken-sized to the whale-sized, all of which were bipedal.
-They are quite alien to anything we know today which restricts our ability to understand their ecology far more so than extinct mammals \citep{weishampel2004dinosauria}.
-Of relevance, are the questions that still persist about their metabolism, with the latest evidence suggesting they were mesothermic \citep[i.e.
+%They are quite alien to anything we know today which restricts our ability to understand their ecology far more so than extinct mammals \citep{weishampel2004dinosauria}.
+%Of relevance, are the questions that still persist about their metabolism, with the latest evidence suggesting they were mesothermic \citep[i.e.
 intermediate to ecto- and endotherms;][]{grady2014evidence}.
-We do know that they walked with the erect gait of mammals or birds rather than the sprawling gait of lizards and that they were most likely facultative scavengers \citep{weishampel2004dinosauria,depalma2013physical}.
-Taken together, this implies dinosaurs had a foraging range that fell in between that of modern terrestrial mammals and reptiles.
+%KH Do we need to get into this?
+%We do know that they walked with the erect gait of mammals or birds rather than the sprawling gait of lizards and that they were most likely facultative scavengers \citep{weishampel2004dinosauria,depalma2013physical}.
+%KH Is this not a bit trivial almost?
+%Taken together, this implies dinosaurs had a foraging range that fell in between that of modern terrestrial mammals and reptiles. 
+%I don't think we can say that, we have no idea on thier range.
+For example using body size they should be larger than modern mammals, however the importance of metabolic rate is still not clear with home range (body size explains most.).
+This might be useful http://www.jstor.org/stable/10.1086/682070?seq=1#page_scan_tab_contents
 
 %TG: Dudes! Cite your godamn Am Nat paper in this paragraph! It's totally more relevant than Weishampel's book!
-
-% KH Something about humans (also the carnivours kangaroo which might be the most efficient mover.)
+%KH I think the terrestrial section needs a little more work still, it doesn't hang together quite right yet although it is not far off it.
+While I have wrote it up this way I think ending with humans would be good as it gives a returning to the present type feel which wraps the whole section up without being too explicit about it.
+Maybe using large bird locomotion would allow this.
 
 \subsection*{Detection}
-It would be pointless to have incredible ranging abilities and not have the sensory architecture to benefit from it.
+
+%KH Just my opinion but I found it really difficult to understand the flow of this section.
+I starts off with a type of 2d- 3d thing but then kinda bounces around different ideas with everything talked about in relation to aerial viewing.
+Its also missing how understanding detection ranges can help us understand past scavenging behavior.
+I would include in the terrestrial section, the use of looking at the portion of the brain attributed to olfactory sensing (like the way we did the scaling of detection range in our paper) as a possible indicator of scavenging.
+I would also throw into the social communication section that these behaviors can show how groups can be good scavengers through active/passive communication and we might expect that in other extinct social groups (the vulture thing was almost impossible for pterodactyls to avoid I would say and probably common in many soaring flying species).
+Also Graeme has a paper on human scavenging about how hominies may have exploited vultures and also probably searched in a social sense.
+%I might be too grumpy at the end of the day, but I think a bit of a restructure might help this section.
+
+%It would be pointless to have incredible ranging abilities and not have the sensory architecture to benefit from it.
+%KH I also don't like this sentence as it wouldn't be pointless, 
 %If we came at this from a position of complete ignorance we would predict scavengers to have well-developed senses and indeed, this is what we find. %TG: I don't really like this sentence. How about:
-As predicted by the necessity of an increased encounter rate, scavengers have well-developed senses.
-A simplification of terrestrial, vertebrate scavengers in sensory terms is one of them existing in a two-dimensional plane while foraging for carrion directly.
-They can detect carcasses at a range that is defined by the radius of their sensory organs. %, usually the visual and olfactory senses.
-As a consequence, they have a much more restricted view of the landscape than do aerial foragers.
-Hyenas make up for this in their ability to smell a rotting carcass 4 km away and to hear the vocalisations of conspecifics at a distance of 10 km \citep{mills1989comparative}. 
+As predicted by the necessity of an increased encounter rate, scavengers have well-developed senses, with the visual and olfactory sensory systems most often associated with scavenging behavior.
+This is perhaps no surprise as the nature of carrion means that sensory systems that rely on detecting signals associated with a living animal, such as audioceptive, electroreception, thermoreception and ecololocation are not likely to be associated with scavenging.
+For example, bats that rely on echolocation do not lend themselves to discovering immobile carrion.
+%
+Apart from the basic capacity of these senses to detect carrion, how they function in different environments is also important.
+In the most simple case the search space is a two dimensional plane (Pawar et al 2012).
+If the scavenger itself is searching on the plane, such as the case for terrestrial scavengers, the detection range is simply defined by the radius of their sensory organs.
+%KH I changed it to make it less confusion with regards to dimension.
+
+%A simplification of terrestrial, vertebrate scavengers in sensory terms is one of them existing in a two-dimensional plane while foraging for carrion directly.
+%They can detect carcasses at a range that is defined by the radius of their sensory organs. %, usually the visual and olfactory senses.
+For vision in such a scenario, the ability to detect carrion can be seriously restricted. Such species can make up for this though using olfactory sensory abilities which are less restricted.
+%KH this kinda jumps ahead about the aerial foraging thing which really dominates the section.
+For example, Hyenas make up for this in their ability to smell a rotting carcass 4 km away and to hear the vocalisations of conspecifics at a distance of 10 km \citep{mills1989comparative}. 
 % Using the approach of \cite{spiegel2013factors} we estimate a spotted hyena %TG: do we need a (\textit{Crocuta crocuta}) here?
@@
 We can compare this to the energetics approach of \cite{ruxton2004obligate}, who calculated a terrestrial scavenger needs to be able to detect carrion at 500 meters in order to survive, which is clearly within the ability of hyenas.
+
+%%KH I'm not sure of the argument here and I'm kinda finding it hard to grasp the tread of this entire section to be honest.
+
 % While considering prehistoric habitats \cite{ruxton2004obligate} calculated that ``a 1 tonne mammal or reptile, in an ecosystem yielding carrion at densities similar to the current Serengeti, could have met its energy requirements if it could detect carrion over a distance of the order of 400–500 m".
@@
 Species capable of flight have effectively added an extra spatial dimension (i.e. the vertical component) to their sensory environment over land animals.
+%I would 
 This allows them to look down on a landscape where they are unencumbered by obstacles that would obstruct the view of a terrestrial scavenger.
@@
 It really lowers down the quality of our claim and is totally un-necessary (and actually probably wrong: lot's of fruit bats (like flying foxes) do not echolocate and thus have to find their fruits among the flipping tropical folliage using their eyes and noses only)...
-We can contrast this with other flying vertebrates such as many bats whose reliance on echolocation would not lend itself to discovering immobile carrion.
-%
 
@@
 %Thus, the bone crushing ability of hyenas and others reveals another useful scavenger trait. %TG: how about the following for a smoother transition:
-In the ability to eat bone scavengers have arrived at a way to feed on a resource that is typically too hard for many predators to process. 
+
+%KH still not sure about this transition.
+Perhaps the most extreme example of utilizing such late carcass stage material is the ability to eat bone, a resource that is typically too hard for many predators to process. 
 % Thus, some phenotypic adaptations such as the ability to crush bone reveal another useful scavenger trait.
@@
 The question of where our ancestors placed on the hunter-scavenger axis during the Plio-Pleistocene has been a matter of debate for years \citep{dominguez2002hunting}.
+%%KH I think this shouldn't be the first place we mention humans as they do appear and fit well in other sections.
 A recent study investigating potential scavenging opportunities for hominins in Kenya found that, even when discounting bone material, there is a substantial amount of scavengeable meat left on predated remains; sufficient to sustain the requirements of an adult male \textit{Homo erectus} \citep{pobiner2015new}.
@@
 The use of tools and the cooperative nature of hominins meant they could likely get a substantial part of their energetic requirements through scavenging depending on their environment \citep{moleon2014humans}.
+%KH should this go in the prey availability section?
 
@@
 The animal also had an enormous bite force, with one estimate putting it at 57000 Newtons \citep{bates2012estimating} which would have been powerful enough to break open skeletal material \citep{rayfield2001cranial}.
+%KH I would mention that however, our paper shows that such species that would have been able to have such bites were probably too large for this to be an effective stratagy.
+Then you can segue into the body size thing.
 

### 2016-04-21 natural history of scavenging.tex
More updates

@@
 
+% KH maybe my issue is that the aquatic, terrestrial, aerial structure, which works well for the locomotion, struggles in the the other sections.
+It kinda leads to sentences that look like they were dropped in and from the discussion to prematurely end on one point before its fully developed.
+% KH an alternative structure would be process type structure that would follow the order of events at the carcass. Something like competition, processing the food and then digestion?
+
 \section*{Handling Time}
@@
 %KH still not sure about this transition.
-Perhaps the most extreme example of utilizing such late carcass stage material is the ability to eat bone, a resource that is typically too hard for many predators to process. 
+One adaptation is to utilize late carcass stage material that is less subject to decomposition and unavailable to other competitors, for example bone.
 % Thus, some phenotypic adaptations such as the ability to crush bone reveal another useful scavenger trait.
@@
 %%KH I think this shouldn't be the first place we mention humans as they do appear and fit well in other sections.
+%KH Here and in other section there is a lack of mentioning fossil/remains etc. For example I'm sure there is something about the tools or the marks on the bones that would fit in well here.
+
 A recent study investigating potential scavenging opportunities for hominins in Kenya found that, even when discounting bone material, there is a substantial amount of scavengeable meat left on predated remains; sufficient to sustain the requirements of an adult male \textit{Homo erectus} \citep{pobiner2015new}.
@@
 The animal also had an enormous bite force, with one estimate putting it at 57000 Newtons \citep{bates2012estimating} which would have been powerful enough to break open skeletal material \citep{rayfield2001cranial}.
-%KH I would mention that however, our paper shows that such species that would have been able to have such bites were probably too large for this to be an effective stratagy.
+%KH something on the robust morphology and size required to generate such forces however essentially restricts flying species from having such a strategy, unless they can use tools (dosent some bird fly up and then drop bones.)
+%KH I would also mention that however, our paper shows that such species that would have been able to have such bites were probably too large for this to be an effective stratagy.
 Then you can segue into the body size thing.
+%%An alternative structure would be to go straight into morphologies that crack bones, i.e large heads, then follow such morphology restricts some groups for example flying species or smaller animals like humans.
+Then you could move into tool use and similar type behavior in humans and birds.
 
@@
 This was the result of gut capacity limitations and the effects of competition at the carcass.
+% KH I would mention that competition really only affects smaller individuals in this scenario.
+That is without competition in the model big things did about just as well but small ones did much better.
+Maybe take the shift in optimum between the models as an example of the affect.
 At the larger extreme this owes to the fact that gut capacity doesn't scale isometrically with body mass so the benefits of greater mass level off; there's only so much food an individual can consume at a single sitting \citep{calder1996size}.
+%% KH This sound specific to theropods while if I remember correctly its for mammals.
+I think we may need to say this in a more general aspect (again changing structure may help.
+(i.e a stomach section would allow snakes to be included and other extremes))
 For the smaller species, larger competitors would have prevented their access to carrion.
@@
 
+%KH Just as a though, people bring this up a lot that body size allows for long period of starvation being specific for scavenging but I never seen anything past a verbal argument.
+Is there anything that really supports it that say locomotion cant explain (i.e large search area thing)?
+
 %%%%%%%%%%  AK This annoying paragraph's new home TG: well done finding him a home!
+%% KH Does it need any home? 
 Certainly, scavenging should be particularly attractive to flying species compared to mammals. % Not sure about this paragraph 
@@
 Aspects including, primary productivity, relief, temperature and competition will all greatly affect scavenging tendency. 
-\cite{ruxton2004obligate} suggest a system with a productivity similar to the Serengeti could have supported a mammalian or reptilian terrestrial scavenger.
-Indeed, in systems that were dominated by large ectothermic or mesothermic vertebrates, the same primary productivity would have supported a greater biomass \citep{mcnab2009resources}.
+\cite{ruxton2004obligate} suggest a system with a productivity similar to the Serengeti could have supported an obligate mammalian or reptilian terrestrial scavenger.
+%% KH Be careful about slipping between scavenger and obligate scavenger.
+Indeed, in systems that were dominated by large ectothermic or mesothermic herbivore vertebrates, the same primary productivity would have supported a greater biomass, due to the scaling of mass with metabolic rate \citep{mcnab2009resources}.
+%KH Need to make it clear why here. Only works for larger herbivores as well
 %TG: We're now talking about Late Mesozoic systems right? Needs to be mentioned somewhere.
 The upshot of this is that there was a higher biomass of herbivores dying and offering scavenging opportunities.
+%KH  I have a feeling the population dynamics to this are actually pretty complex.
+Large animals may live longer and represent a standing biomass similar to tropical rainforest's were all the nutrients are locked up in big trees.
+The life-history of these animals would hence be what really determines the production of carcasses however that is way outside the scope of this paper and would require a paper all on its own.
+Maybe we could drop in that point as I have not seen anyone ever mention this point.
 Predators were large-bodied too compared to extant mammalian predators \citep{mcnab2009resources}, and so, especially if they were ectothermic, could last longer between meals, rendering scavenging a more attractive behaviour relative to predation.
+%%KH Not sure this sentence belongs here, already covered int he handling section.
 Osteophagy may have been even more viable during the Mesozoic era as well because of this skewed body mass distribution of herbivores towards larger sizes \citep{10.1371/journal.pone.0051925}.
@@
 In many habitats (e.g. the Arctic) it is simply not possible for sufficiently powerful thermals to form and as a consequence large-bodied vultures cannot exist.
+%% KH I dont think this belong in this section, we probably have covered this enough in the locomotion part.
 One result of this is that terrestrial carnivores like bears and wolves take more carrion \citep{devault2003scavenging}.
@@
 If we apply this line of reasoning over evolutionary time-scales, the absence of flying vertebrates in the Palaeozoic may have permitted terrestrial forms to take in a higher proportion of carrion in their diet.
+%%KH I think the focus needs to be more on the actual environment and opportunities available.
+I would suggest that compititon either stays in the handling section (my preference) or is moved down to here (which also works really).
+%%KH You could then split the structure into production in terrestial/marine which focuses on primary productivity and species composition and how they affect what available then the next part being competition in both section.
 
@@
 %TG: maybe change transition here.
-Staying in the aquatic setting, the phenomenon of occasional bounties of carrion in the form of whale falls has led some researchers to investigate if a scavenger could survive by seeking out these remains exclusively.
+Staying in the aquatic setting, the phenomenon of occasional bounties of carrion in the form of whale falls has led some researchers to investigate if an obligate scavenger could survive by seeking out these remains exclusively.
 \cite{ruxton2005searching} argued that although this is energetically feasible it's ecologically unlikely.
 Any animal that could find such whale carcasses is unlikely to have ignored other types of carrion.
-Although no aquatic species have ever exceeded the size of whales, some enormous animals have evolved in this environment before the evolution of cetaceans, including \textit{Leedsichthys}, a bony fish from the Middle Jurassic (174.1-163.5 Mya) and the aquatic Mesozoic reptiles, the plesiosaurs, pliosaurs and ichtyosaurs, that could all exceed 15 metres in length \citep{ruxton2011zoology}.
+Although no aquatic species have ever exceeded the size of whales, some enormous animals have evolved in this environment before the evolution of cetaceans, including \textit{Leedsichthys}, a bony fish from the Middle Jurassic (174.1-163.5 Mya) and the aquatic Mesozoic reptiles, the plesiosaurs, pliosaurs and ichtyosaurs, that could all exceed 15 metres in length \citep{ruxton2011zoology} (reptile fall ref here).
+
+%% KH check out this paper on Mesozoic reptile falls http://www.nature.com/ncomms/2014/140910/ncomms5789/full/ncomms5789.html
 So, despite being unlikely, the energetic feasibility of a marine scavenger that specialises on large carcasses has a long history.
@@
 % Add bits on climate change/dectection rate change, etc...
+
+% KH Maybe we need to more strongly reaffirm that the framework that we give really is useful and should be applied to all sorts of things. etc
 

### 2016-04-22 natural history of scavenging.tex
minor edits to the intro

@@
 \item A. Kane (\url{adam.kane@ucc.ie}), University College Cork, Cooperage Building, School of Biological Earth and Environmental Sciences, Cork, Ireland.
-\item K. Healy and A. Jackson, Trinity College Dublin, Department of Zoology, School of Natural Sciences, Dublin Ireland.
+\item K.
+Healy and A.L.
+Jackson, Trinity College Dublin, Department of Zoology, School of Natural Sciences, Dublin 2, Ireland; Trinity Centre for Biodiversity Research, Trinity College Dublin, Dublin 2, Ireland.
 \item T. Guillerme, Imperial College London, Silwood Park Campus, Department of Life Sciences, Buckhurst Road, Ascot SL5 7PY, United Kingdom.
@@
 Some simple heuristics can inform us, for instance, in cases where the prey item was simply too large to have been killed by the ostensible predator \citep{pobiner2008paleoecological}.
-But clearly, a scavenger doesn't only feed on animals too big for it to have hunted.
-The obvious lack of direct behavioural data compounds the difficulty of discerning scavenging among extinct forms.
+But clearly, a scavenger does not only feed on animals too big for it to have hunted.
+The obvious lack of direct behavioural data compounds the difficulty of discerning scavenging from predation among extinct forms.
 Indeed, a single species of dinosaur notwithstanding \citep{carbone2011intra}, a synthesis describing the natural history of scavengers is absent from the literature.
 %TG: I don't understand the dinosaur part of this sentence. Might be my non-native english but I thought it's worth noting for other non-english folk out there (even though I know they aren't many!).
-Fortunately, research on scavenging is on the rise \citep{manga2006vulture}.
-As a result, we are now beginning to realise the extent of this behaviour such that, ``in some ecosystems, vertebrates have been documented to assimilate as much as 90\% of the available carrion" \citep{beasley2015vertebrates}.
+With research on scavenging on the rise \citep{manga2006vulture} we are now beginning to realise the extent of this behaviour such that, ``in some ecosystems, vertebrates have been documented to assimilate as much as 90\% of the available carrion" \citep{beasley2015vertebrates}.
 This has profound implications for the trophic ecology of these systems and particularly our models of them.
-Even Tenney’s noble big cats are now known to take in a significant portion of carrion in their diet where some lion populations get over 50\% of their meat from carcasses \citep{jones2015african}.
-By recognising the difficulty in directly observing scavenging, other methods have been turned to to discern the most suitable morphologies, physiologies and environments for a scavenging lifestyle to prosper.
+Even Tenney’s noble big cats are now known to take in a significant portion of carrion in their diet where some lion populations acquire over 50\% of their meat from carcasses \citep{jones2015african}.
+While recognising the difficulty in directly observing scavenging, it is possible to turn to other methods in order to discern the most suitable morphologies, physiologies and environments for a scavenging lifestyle to prosper.
 Here we chart the natural history of scavenging by assessing the potential for the behaviour in dominant vertebrate groups given their ecology and functional traits.
@@
 Through chance alone many species will avail of some opportunistic scavenging. %TG: Two times "opportunities", replaced by "chances"
-However, species that rely on scavenging to sustain substantial portions of their diets must increase the probability of encountering a sufficient amount of carrion in order to meet their energetic demands.
+However, species that rely on scavenging to sustain substantial portions of their diets must encounter a sufficient amount of carrion in order to meet their energetic demands.
 %Once found, the scavenger must be able to out-compete any potential competitors and process a product under the processes of decay from toxin producing and disease causing microorganisms \citep{ruxton2014fruit}.
@@
 Once found, the scavenger must be able to out-compete any potential  competitors and process the, typically decaying, carcass replete with microorganism derived toxins  \citep{ruxton2014fruit}.
-Finally, the potential for scavenging will also depend on the density, size, and quality of carcases produced, all of which are affected by complex ecosystem dynamics.
-All of these facets are essentially the key parameters found in functional response curves, namely encounter rate, handling time and prey availability \citep{jeschke2002predator}.
+Both these characteristics can be assumed to be under evolutionary selection pressures for traits that increase carrion discovery and monopoly.
+Finally, the potential for scavenging will also depend on the density, size, and quality of carcases produced, all of which are affected by complex ecosystem dynamics and are outside the selection pressures on the scavenger.
+Each of these facets are essentially the backbone of fundamental ecological theory and are the key parameters defined in functional response curves, namely encounter rate, handling time and prey availability \citep{jeschke2002predator}.
 %TG: Jaysus, this sentence doesn't read well!

### 2016-04-25 natural history of scavenging.tex
trying to get this pull request to go through with pointless edit
reordered sections, fixed up detection and handling time

@@
 \begin{abstract} 
-% Here, we chart the natural history of scavenging. 
 Despite its prevalence, scavenging is a difficult behaviour to observe in modern day carnivores and impossible to study directly in extinct species. 
@@
 These can be thought of as some of the principal parameters in optimal foraging theory namely, encounter rate, handling time and prey availability. 
-% Chief among these are low-cost locomotion, high detection distances, effective carcass processing and a carrion-producing habitat. 
 We use these components to highlight the morphologies and environments that would have been conducive to scavenging over geological time by focusing on the dominant vertebrate groups of the land, sea and air.
@@
 Indeed, a single species of dinosaur notwithstanding \citep{carbone2011intra}, a synthesis describing the natural history of scavengers is absent from the literature.
-%TG: I don't understand the dinosaur part of this sentence. Might be my non-native english but I thought it's worth noting for other non-english folk out there (even though I know they aren't many!).
 With research on scavenging on the rise \citep{manga2006vulture} we are now beginning to realise the extent of this behaviour such that, ``in some ecosystems, vertebrates have been documented to assimilate as much as 90\% of the available carrion" \citep{beasley2015vertebrates}.
@@
 
-\subsection*{The Challenges of Scavenging} %TG: challenges or difficulty? up to you.
+\subsection*{The Challenges of Scavenging} 
 The chief hurdle to scavenging is finding a resource that is often difficult to predict in space and time.
-Through chance alone many species will avail of some opportunistic scavenging. %TG: Two times "opportunities", replaced by "chances"
+Through chance alone many species will avail of some opportunistic scavenging. 
 However, species that rely on scavenging to sustain substantial portions of their diets must encounter a sufficient amount of carrion in order to meet their energetic demands.
-%Once found, the scavenger must be able to out-compete any potential competitors and process a product under the processes of decay from toxin producing and disease causing microorganisms \citep{ruxton2014fruit}.
-%TG: make the following sentence more sense?
 Once found, the scavenger must be able to out-compete any potential competitors and process the, typically decaying, carcass replete with microorganism derived toxins  \citep{ruxton2014fruit}. 
-Both these characteristics can be assumed to be under evolutionary selection pressures for traits that increase carrion discovery and monopoly.
+Both of these characteristics can be assumed to be under evolutionary selection pressures for traits that increase carrion discovery and monopoly.
 Finally, the potential for scavenging will also depend on the density, size, and quality of carcases produced, all of which are affected by complex ecosystem dynamics and are outside the selection pressures on the scavenger.
 Each of these facets are essentially the backbone of fundamental ecological theory and are the key parameters defined in functional response curves, namely encounter rate, handling time and prey availability \citep{jeschke2002predator}.
-%TG: Jaysus, this sentence doesn't read well!
-"Functional response curves of optimal foraging theory" can probably be phrased in a nicer/easier way...
-%With this framework we compare the ability of scavenging across vertebrates. % focusing on each of these facets and their relationship with scavenging.
 By considering scavenging in this context of optimal foraging we can identify the prerequisite attributes and processes required for the behaviour. 
@@
 All foraging processes depend on the encounter rate between consumer and resource. 
-In the simplest case, this rate can be thought of in terms of a gas diffusion model where the movement of two agents (i.e.
-predator and prey) depends only on as they can actively detect resources through their sensory abilities.
-%TG: I like the gas diffusion analogy but it might be adding too much text and be not essential.
-I've added some precisions though.
-%KH I'v shortened it some more.
-As carcasses are stationary, the relative speed between a scavenger and carrion is only dependent on the movement of the scavenger.
-%KH There are too many definite statements throughout from my liking.
-For example, not all carcasses are stationary in fact the sinking of carcasses in marine systems is what drives whale fall fauna as they essentially let the carcass come to them.
-I know it might make the text more clunky but I would play it safe and "As carcasses are usually stationary" etc.
-As such, scavenging potential is strongly affected by search rates, which are determined by both species' physiology and what type of environment the scavenger is foraging in.
-%KH is including dimensionality making this too complex as the only search environment that is 3d is for semi-buoyant marine carcasses.
-Also the dimensionality talked about in the vision section is different to this dimensionality (the Dim of the search space for both terrestrial and aerial are both 2D).
-Might be a bit confusing so I changed it to just the physiology and environmental
+Locomotory speed, foraging time and detection radius all increase the encounter rate between a scavenger and the carcasses it is searching for. 
+We would thus expect selection pressures to act on the various traits that govern these parameters.  
+% However, encounter rate is also determined by the productivity of the environment, something which the selective forces acting on the scavenging have no control over.
+
+%As such, scavenging potential is strongly affected by search rates, which are determined by anatomy, physiology and behaviour which are all under selective pressure, as well as the type of environment the scavenger is foraging in.
 
 \subsection*{Locomotion}
-Because of the inherent unpredictability of carrion, scavenging heavily depends on the ability to efficiently move over large areas than predation.
-%KH changed "more" to "heavily"
-This generally requires an efficient transfer of metabolic energy into movement which relies on both physiology (i.e.
-metabolism) and the medium of the environment in which the animal is moving (i.e.
+Because of the inherent unpredictability of carrion, scavenging depends more on the ability to efficiently move over large areas than predation.
+This generally requires an efficient transfer of metabolic energy into movement which relies on the animal's anatomy and physiology as well as the medium of the environment in which the animal is moving (i.e.
 aerial, aquatic or terrestrial).
@@
 Despite the energetic costs of flight, the only known vertebrate obligate scavengers are the old and the new world vultures. 
-And, although powered flight is energetically expensive, species like vultures can exploit air currents using their large wingspans which allows them to soar at a cost of only twice their metabolic rate \citep{hedenstrom1993migration,spivey2014analysing}.
-%KH This and seems out of place a bit, maybe replace with "While powered flight..."
+While powered flight is energetically expensive, species like vultures have evolved behavioural and anatomical features to exploit air currents using their large wingspans, allowing them to soar at a cost of only twice their metabolic rate \citep{hedenstrom1993migration,spivey2014analysing}.
 By depending on thermal air flows these species can forage over vast ranges \citep{spiegel2013factors}. 
@@
 The groups from which these modern soaring birds arose, appear during the Palaeocene \citep[66 - 56 Million years ago (Mya); ][]{Jetz2012, Jarvis2014} and Cretaceous \citep{chiappe2006early} respectively.
-% KH and are likely to have being soaring fliers as demonstrated by their large wingspans (Is this true?).
-(Is there any other form of soaring?).
-AK I don't think this is true from what I've seen
-%KH OK I changed it to these modern soaring birds to emphasize we are still on the tred of low cost flying. However if you don't think its clear enough what this group is feel free to change it back.
 However, soaring flight is likely to be far older than this with avian flight originating in the Late Jurassic (163.5-145 Mya) and vertebrate flight in the Late Triassic (235-201.3 Mya) coincident with the pterosaurs.
@@
 Certainly, avian flight is cheaper than either walking or running \citep{tucker1975energetic}.
-%TG: I'd rather put the following part as an appendix or foot note. Reader can take our (Tucker's) word for it and then check the math to see that even with maintenance and blablabla flying wins.
-%Even taking account of maintenance costs, this still bears out, where total cost of movement (J kg\textsuperscript{-1} m\textsuperscript{-1}) scales according to 5.2 $\times$ body mass (kg) \textsuperscript{-0.23} for fliers and 10.7 $\times$ body mass (kg) \textsuperscript{-0.32} for runners \citep{williams1999evolution}.
-%TG: This also allows to link directly to the next argument in the same paragraph: facultative active flying scavengers.
-%KHI agree avoid as much maths as possible (If only someone would propose some sort of drop down thing for maths in papers.)
+
 We know that many extant birds exist as facultative scavengers because storks, raptors and corvids all take substantial quantities of carrion in their diet \citep{kendall2013alternative}. 
@@
 For example, early birds such as \textit{Archaeopteryx} are predicted to have been poor, relatively inefficient fliers \citep{nudds2010narrow} and so ill-suited to finding carrion. 
-%Similarly something about the efficiency of pterosaurs flight.
-TG: Was this a comment?
-By the way, about pterosaurs (though it might not be necessary to mention it), most of them where tiny little feckers like Archaeopteryx or even smaller.
-
 The importance of efficient flying over large areas may explain the lack of scavenging behaviour in bats as they are generally nocturnal, a time when they would receive no aid from convective air currents \citep{norberg2012vertebrate}.
-%smaller and less efficient in comparison to larger birds (speculating here but I wouldnt be surprised), %AK Don't think it is the case that bats are less efficient than birds in terms of flight 
-That said, \textit{Necromantis} (``death-eater''), a large bat from the middle to late Eocene (56-33.9 Mya) had a robust cranio-mandibular morphology, and is a likely candidate for scavenging behaviour \citep{Weithofer_Necromantis_1887,Hand_Necromantis_2012}.% amongst the x amount of known bat species \citep{Weithofer_Necromantis_1887,Hand_Necromantis_2012}.
-%TG: we can cut this number of bat species to avoid getting into stamp collecting debate.
-+ there are effectively a two shit load units of them which makes the Necromantis example not really convincing.
 
 Similar to aerial species, aquatic scavengers have a locomotory benefit because water is a medium that is conducive to low-cost movement \citep{tucker1975energetic,williams1999evolution}.
-%TG: Same as comment before, maybe put the following as a footnote or as a supplementary.
-% In fact, the total cost of movement (again in J kg\textsuperscript{-1} m\textsuperscript{-1}) in salmonids is lower than either running or flying where it scales according to 2.15 $\times$ body mass (kg) \textsuperscript{-0.25} \citep{williams1999evolution} with only \textit{soaring} flight likely to surpass it.
-%KH(I assume the reference is for powered flight, again this needs to be more precise). 
-%KH again I think its good  to keep away from the maths 
 This has led some researchers to argue for the likelihood of an obligate scavenging fish \citep{ruxton2004energetic,ruxton2005searching}.
 
-Sharks are one likely candidate for general scavenging behaviors as their locomotion, which depends on large pectoral fins in order to maintain lift as they swim, resemble large soaring fliers.
+Sharks are one likely candidate for general scavenging behaviors as their locomotion, which depends on large pectoral fins to generate lift as they swim, resembles that of the large soaring fliers.
 %changed this sentence up a bit 
@@
 In fact, facultative scavenging is seen in many selachian groups, including species of extant sharks like white sharks \citep[known to feed on whale carcasses;][]{fallows2013white}, Greenland sharks \citep[feeding on seals;][]{watanabe2012slowest}, and sixgill sharks \citep{anderson2016impact}.
-The former which grow up to 6 metres long, can be sustained by 30 kg of whale blubber for over six weeks \citep{carey1982temperature}. 
-%A personal thing but I hate the use of former or latter as I have to break my reading and go back and see which species that is.
-I did'nt change it but I would go with naming the species there and then for clarity.
 There is evidence too of scavenging in extinct species, where shark teeth have been found in the remains of dinosaurs, mosasaurs and Pliocene mysticete whales \citep[5.3-3.6 Mya; ][]{schwimmer1997scavenging,ehret2009caught}.
 
-Interestingly, style of swimming in fish does not significantly affect the cost of movement \citep{williams1999evolution}, hence it is likely that aquatic species with large ranges may also encounter scavenging opportunities.
-It may be that jaw morphology is more likely to define which species can avail of carcasses.
+Interestingly, style of swimming in fish does not significantly affect the cost of movement \citep{williams1999evolution}, hence it is likely that many aquatic species with large ranges will encounter scavenging opportunities.
+It is perhaps jaw morphology that is more likely to define which species can avail of carcasses.
 %KH I moved this down here as it kinda throws our shark soaring thing out the window (which I like as a linking function).
@@
 % KH endotermy - why it doesn't seem to work
-We might expect then that by combining an aquatic environment and an endothermic metabolism marine mammals would especially prosper as scavengers, especially as this group has perhaps the largest home range of any animals (whales and shit).
-Fossil pinnipeds and cetaceans from 60 Mya %TG: that early???
-have transitional features indicative of their trajectory to fully aquatic species \citep{williams1999evolution}.  
+We might expect then that by combining an aquatic environment and an endothermic metabolism marine mammals would especially prosper as scavengers.
+%, especially as this group has perhaps the largest home range of any animals (whales and shit).
+Fossil pinnipeds and cetaceans from 60 Mya have transitional features indicative of their evolutionary trajectory to fully aquatic species \citep{williams1999evolution}.  
 But despite their movement away from land their energetic savings were negligible because the \textit{total} cost incurred by a swimming marine mammal is high \citep{williams1999evolution}. 
 Indeed, the total energetic cost is similar to an equivalent terrestrial or aerial mammal \citep{williams1999evolution}.
-This underscores the trade offs between the benefits of endothermy in terms of activity periods and the costs of maintaining such an energetically costly system. 
+This underscores the trade offs between the benefits of endothermy in terms of activity periods and the costs of maintaining such an energetically expensive system. 
 That said, aquatic endotherms have and do scavenge. 
@@
 %%KH Is there anything we should add about other extinct groups such as plesiosaurs and https://en.wikipedia.org/wiki/Plesiosauria#Gait maybe an air breathing non-airbreathing devide would be good.
-
-%\cite{williams1999evolution} notes, ``free-ranging animals must contend with the total energetic expenditure associated with supporting basic biological functions as well as with moving the body and appendages through the environment."
 
@@
 %Also that reference is from 1975, is this up to date?
-
-%TG: Hmm I get the point of gas exchange efficiency but (1) it should be more developed if left like that (i.e.
-what the flip is meant by gas efficient exchanges) and (2) that works for mammals, but in terms of terrestrial that ignores birds, squamates, crocodiles and amphibians.
 
@@
 
-The early transition from a sprawling gait, such as seen in Tiktaalik, to the more erect posture of Synapsids and later dinosaurs and mammals, has often been supposed as conferring a huge advantage to the latter.
+The early transition from a sprawling gait, such as seen in \textit{Tiktaalik}, to the more erect posture of Synapsids and later dinosaurs and mammals, has often been supposed as conferring a huge advantage to the latter.
 %KH Even after changing it I dont like this transition type narrative.
@@
 Although the earliest evidence of vertebrate scavenging comes from the Permian (298.9 - 252.17 Mya) where a temnospondyl amphibian fed on the carcass of \textit{Varanops}, a predatory synapsid of the time \citep{reisz2006articulated}.
-%TG: I do like the way all our paragraphs end up with a "But wait, there's an exception!"
 
 Modern endothermic mammals can sustain longer periods of energetically expensive activity \citep{bennett1979endothermy} resulting in larger foraging ranges. 
-%TG: Same comment as for the other math, I'd rather put that as a footnote or in the appendix (the whole yoke).
-Again, it seems a bit disproportionate that for some example, we just dumb a line (or even an example in brackets) and that for some others, we explain the example in details.
-% To quantify this effect with a simple example we can turn to some allometric relationships relating sustainable travelling speed to body mass \citep{ruxton2004obligate}.
-% If we insert these into a foraging radius model \citep{Enstipp2006Energetics} for a 12 hour foraging day it shows that while a 10 kg reptile can range 6.5 km, an equally sized mammal can range nearly 33 km (See appendix).
-% For a foraging scavenger, this ability translates into a greater area searched for food.
 Today, terrestrial scavenging in the mammals is probably best known in an African context where hyenas, jackals and lions all take sizeable proportions of carrion in their diet.
@@
 In contrast, big cats like leopards (\textit{Panthera pardus}) rely on ambush \citep{pennycuick1995radius}. 
-This allows us to make a broad distinction between the ambush strategies of cats-like forms and the pursuit/ pounce strategies of more dog-like forms%TG: I've added the "-like" to underline that we're not talking about the phylogeny here (i.e.
-not Felidae (incl.
-hyena) vs Canidae)
-, the latter being more suited to scavenging \citep{janis2014forelimb}. 
+This allows us to make a broad distinction between the ambush strategies of cat-like forms and the pursuit/ pounce strategies of more dog-like forms, the latter being more suited to scavenging \citep{janis2014forelimb}.
 These insights allow us to compare extant terrestrial species to their prehistoric forebears given the dominance of mammalian carnviores since the Eocene (56-33.9 Mya) where the order split into the Caniforma and Feliforma \citep{van1987skeletal}.
 To take one example, \cite{anyonge1996locomotor} found that \textit{Nimravides}, a genus of sabretooth cat from the Miocence (10.3 to 5.3 Mya), were likely to have been ambush predators which would argue against them taking a lot of carrion.
-%The order Carnivora saw its origins in the Middle Eocene  %TG: a phylo side note: the correct sentence would be the "The stem Carnivora saw its origins ...
-when it split into Caniforma and Feliforma." But I think we can leave it out because, god, it sounds pedantic!
-% So we can trace efficient terrestrial movement by carnivores from this point on.
 
-While we can use modern analogues of quadrupedal species, the absence of scavenging in bipedal animals makes this more difficult.
-However, as a groups bipedal locomotion may have been a particularly useful strategy for  searching large areas.
+Of course, terrestrial animals can also move bipedally. 
+Although the evolution of bipedal  movement was significant in that it freed up the forelimbs for other purposes (e.g.
+climbing, tool-use, wing development etc.) it does not differ radically in cost from quadrupedal locomotion (\cite{williams1999evolution}, and references therein).
+For instance, \cite{alexander2004bipedal} shows  that, in the case of humans, we are more economical than predicted while walking and less so while running according to predicted costs of terrestrial movement.
+Our locomotory efficiency has fed into the question of where our ancestors placed on the hunter-scavenger axis during the Plio-Pleistocene, which has been a matter of debate for years \citep{dominguez2002hunting}.
+\cite{ruxton2013endurance} added to this debate with his argument that long distance endurance running, often used by humans for hunting, was not an important feature of hominin scavenging.
+He arrives at this conclusion on the basis of the high cost of running coupled with the fact that dangerous competitors will tend to arrive beforehand.
+
+% While we can use modern analogues of quadrupedal species, the lack of scavenging in bipeds makes this more difficult.
+However, bipedal locomotion may have been a particularly useful strategy for searching large areas.
 For instance, \cite{alexander2004bipedal} shows that, in the case of humans, we are more economical than predicted while walking according to predicted costs of terrestrial movement.
-In fact the ability of humans to engage in hunting strategies that involve traversing long ranges, such as persistence hunting, has led to several researches highlighting the possible role of scavenging in early hominids.
+In fact the ability of humans to engage in hunting strategies that involve traversing long distances, such as persistence hunting, has led to several researches highlighting the possible role of scavenging in early hominids.
 Indeed the low energetics of human walking at least allows for a substantial amount of scavenging in our hominid ancestors (Graemes papers on scavenging).
@@
 
-%Of course, terrestrial animals can also move bipedally. 
-%TG: missing some transition here. Maybe something short like "Further back in time" or "Previous to that"...
-%I agree, 
-%Although the evolution of bipedal movement was significant in that it freed up the forelimbs for other purposes (e.g.
-climbing, tool-use, wing development etc.) it does not differ radically in cost from quadrupedal locomotion \citep[][and references therein]{williams1999evolution}.
-%For instance, \cite{alexander2004bipedal} shows that, in the case of humans, we are more economical than predicted while walking and less so while running according to predicted costs of terrestrial movement.
-%the second sentence does not really follow from the first -
 Aside from humans and our allies, the best-known terrestrial bipeds are the dinosaurs and unsurprisingly, given their enduring appeal, the prevalence of scavenging has been extensively explored in the carnivorous theropods.
-%KH you could reference all the shit on theropod scavenging here if you wanted.
 These were the dominant terrestrial carnivores for most of the Mesozoic Era (252.17 - 66 Mya) and ranged from the chicken-sized to the whale-sized, all of which were bipedal.
-While the locomotory ability of theropods has been debate since their first inception, more recent studies have reconstructed them as relatively mobile animals (John Hutchinsons papers). 
-%See that Pontzer 
-Despite some suggestions that the large size of some forms may have allowed some advantage in scavenging, partially due to the abilty to search large areas (Graemes paper etc), more recent studies have shown that the energetic demands of such large forms meant that scavenging was likely to be more prevalent in mid sized theropods of approximately half and thon (Us and Carbones paper).
+While the locomotory ability of theropods has been debated since their first inception, more recent studies have reconstructed them as relatively mobile animals \citep{pontzer2009biomechanics}. 
+Despite some suggestions that larger species may have had some advantage in scavenging, partially due to the ability to search large areas \citep{ruxton2003could}, more recent studies have shown that the energetic demands of such large forms meant scavenging was likely more prevalent in mid-sized theropods of approximately half a tonne \citep{carbone2011intra,kane2016body}
 
@@
 
-%TG: Dudes! Cite your godamn Am Nat paper in this paragraph! It's totally more relevant than Weishampel's book!
 %KH I think the terrestrial section needs a little more work still, it doesn't hang together quite right yet although it is not far off it.
@@
 
-%It would be pointless to have incredible ranging abilities and not have the sensory architecture to benefit from it.
-%KH I also don't like this sentence as it wouldn't be pointless, 
-%If we came at this from a position of complete ignorance we would predict scavengers to have well-developed senses and indeed, this is what we find. %TG: I don't really like this sentence. How about:
-As predicted by the necessity of an increased encounter rate, scavengers have well-developed senses, with the visual and olfactory sensory systems most often associated with scavenging behavior.
-This is perhaps no surprise as the nature of carrion means that sensory systems that rely on detecting signals associated with a living animal, such as audioceptive, electroreception, thermoreception and ecololocation are not likely to be associated with scavenging.
-For example, bats that rely on echolocation do not lend themselves to discovering immobile carrion.
-%
+As predicted by the importance of an increased encounter rate, scavengers have well-developed senses, with the visual and olfactory sensory systems most often associated with scavenging behavior. 
+This is perhaps no surprise because sensory systems that rely on detecting signals associated with living animals, such as audioception, electroreception, thermoreception and echolocation will be limited in their ability to detect an already dead animal.
+
 Apart from the basic capacity of these senses to detect carrion, how they function in different environments is also important. 
-In the most simple case the search space is a two dimensional plane (Pawar et al 2012).
-If the scavenger itself is searching on the plane, such as the case for terrestrial scavengers, the detection range is simply defined by the radius of their sensory organs.
-%KH I changed it to make it less confusion with regards to dimension.
+In the simplest case, the search space is a two dimensional plane \citep{pawar2012dimensionality}. 
+If the scavenger itself is searching on the plane, as is so for terrestrial species, the detection range is simply defined by the radius of their sensory organs.
+Consequently, the ability to detect carrion can be seriously restricted for visually reliant, terrestrial species. 
+They may overcome this restriction however, by using olfaction, which is less affected by the relief of the land.
+For example, hyenas have the ability to smell a rotting carcass 4 km away \citep{mills1989comparative}, which exceeds the 500 m range deemed necessary by \cite{ruxton2004obligate} to be able to survive as a scavenger.
 
-%A simplification of terrestrial, vertebrate scavengers in sensory terms is one of them existing in a two-dimensional plane while foraging for carrion directly.
-%They can detect carcasses at a range that is defined by the radius of their sensory organs. %, usually the visual and olfactory senses.
-For vision in such a scenario, the ability to detect carrion can be seriously restricted. Such species can make up for this though using olfactory sensory abilities which are less restricted.
-%KH this kinda jumps ahead about the aerial foraging thing which really dominates the section.
-For example, Hyenas make up for this in their ability to smell a rotting carcass 4 km away and to hear the vocalisations of conspecifics at a distance of 10 km \citep{mills1989comparative}. 
-% Using the approach of \cite{spiegel2013factors} we estimate a spotted hyena %TG: do we need a (\textit{Crocuta crocuta}) here?
-%could resolve a 2 metre target at 1 km distance. 
-We can compare this to the energetics approach of \cite{ruxton2004obligate}, who calculated a terrestrial scavenger needs to be able to detect carrion at 500 meters in order to survive, which is clearly within the ability of hyenas.
-
-%%KH I'm not sure of the argument here and I'm kinda finding it hard to grasp the tread of this entire section to be honest.
-
-% While considering prehistoric habitats \cite{ruxton2004obligate} calculated that ``a 1 tonne mammal or reptile, in an ecosystem yielding carrion at densities similar to the current Serengeti, could have met its energy requirements if it could detect carrion over a distance of the order of 400–500 m".
-Moreover, the senses of many extant (and in all probability extinct) carnivores meet this required distance, making scavenging feasible for terrestrial species \citep{farlow1994speculations,mech2010wolves}.
+Indeed, the olfactory senses of many extant (and in all probability extinct) carnivores meet this required distance, making scavenging feasible for most terrestrial carnivores \citep{farlow1994speculations,mech2010wolves}.
+Among extinct species in particular, we can use the ratio of olfactory bulb to brain size to infer a preference for olfactory foraging \citep{zelenitsky2011evolution}.
+This approach was used by \cite{zelenitsky2011evolution} to hypothesise such a mode for the theropod dinosaur \textit{Bambiraptor} and by \cite{witmer2009new} for tyrannosaurs.
+The flying pterosaurs however, had tiny olfactory bulbs indicating this sense was not relied on \citep{witton2013pterosaurs}
 
 Species capable of flight have effectively added an extra spatial dimension (i.e. the vertical component) to their sensory environment over land animals.
-%I would 
 This allows them to look down on a landscape where they are unencumbered by obstacles that would obstruct the view of a terrestrial scavenger.
 Such an ability has obvious benefits in detecting carrion.
-Certainly, vultures are known to have impressive visual acuity, with one estimate indicating lappet-faced vultures %TG: We need a standardisation for english species names, I'll go for all lower case (similar as what you'd do for "pig" or other english common species names).
- (\textit{Torgos tracheliotus}) are capable of detecting a 2 metre carcass over 10 km away \citep{spiegel2013factors}.
+Certainly, vultures are known to have impressive visual acuity, with one estimate indicating lappet-faced vultures (\textit{Torgos tracheliotus}) are capable of detecting a 2 metre carcass over 10 km away \citep{spiegel2013factors}.
 Eagles too are known to have highly developed vision \citep{reymond1985spatial}.
-It follows that the evolution of flight allowed aerial animals to detect far more carrion than their terrestrial counterparts \citep{AR:AR22815}.
-%We can contrast this ability with bats, whose visual acuity is famously poor.
-%TG: Humph!
-That's not nice!
-I'm wouldn't go with all the "this is known" thingies here.
-It really lowers down the quality of our claim and is totally un-necessary (and actually probably wrong: lot's of fruit bats (like flying foxes) do not echolocate and thus have to find their fruits among the flipping tropical folliage using their eyes and noses only)...
+It follows that the evolution of flight allowed aerial animals to detect far more carrion than their terrestrial counterparts through vision \citep{AR:AR22815} as the pterosaurs convergently evolved large orbits and optic lobes \citep{witton2013pterosaurs}.
 
-%TG: missing some transition here.
-
-Having a panoramic view also means being able to gather a wealth of information from other foragers, be they conspecifics or otherwise \citep{jackson2008effect}.
+Existing in this '3D' environment also means being able to gather a wealth of information from other foragers, be they conspecifics or otherwise \citep{jackson2008effect}.
 Again, returning to vultures, the genus \textit{Gyps} consists of highly social and colonially nesting species \citep{fernandez2015density}.
 These behaviours allow them to forage far more efficiently because one bird can scrounge information on the location of food from another successful forager \citep{KaneVul}.
-This efficiency has been exploited by mammals such as hyenas who are known to follow groups of vultures \citep{jones2015african}.
+Information transfer of this kind is typically inadvertent and as a consequence no complex social interactions are required, simply the ability to recognise a successful forager.
+Thus, it seems probable that scrounging behaviours were seen in the flying pterosaurs as well.
 
-Aside from sight, many birds have well developed olfactory systems \citep{AR:AR22815} including three species of vultures within the new world family Cathartidae, (genus \textit{Cathartes}).
-% three species of vultures within the new world family Cathartidae, (genus \textit{Cathartes}), have well developed olfactory systems \citep{AR:AR22815}.
-Among them are the turkey vultures (\textit{Cathartes aura}) which were able to locate 90\% of baits set out in a tropical forest \citep{houston1986olfaction}.
-An atuned sense of smell is obviously useful in detecting decaying carrion from the air over a heavily forested habitat.
-% This would be impossible for the visually reliant old world species.
+The terrestrial-olfaction, aerial-visual divide is not total though.
+Terrestrial species like hyenas and hominins exploit the efficiency of birds by looking to the skies for groups of vultures to follow to carrion \citep{jones2015african,ruxton2013endurance}. 
+And many birds, e.g.
+turkey vultures (\textit{Cathartes aura}), have well-developed olfactory systems \citep{AR:AR22815} which they use to forage in heavily forested areas where vision is limited \citep{houston1986olfaction}.
 
-% Depending on the species, a carcass in water either floats or descends to the sea floor \citep{Whitehead415}.
-In contrast to the air, aquatic species have to contend with a low-light environment where visual detection distances are far lower ($<$ 100 m) than they would be in the air.
+Although aquatic species also have a vertical component because they forage in a water body, they must contend with the low-light environment where visual detection distances are far lower ($<$ 100 m) than they would be for air.
 As such, aquatic animals detect resources through chemo- and mechanoreception more so than through vision \citep{ruxton2004energetic}.
@@
 
+\section*{Prey Availability}
+The environmental influence on prey availability is an aspect that greatly affects encounter rate but is invisible to the selective forces acting on the scavenger. 
+Aspects including, primary productivity, relief, and temperature will all greatly affect scavenging tendency. 
+\cite{ruxton2004obligate} suggest a system with a productivity similar to the Serengeti could have supported an obligate mammalian or reptilian terrestrial scavenger.
+%% KH Be careful about slipping between scavenger and obligate scavenger.
+Indeed, in systems that were dominated by large ectothermic or mesothermic herbivore vertebrates, the same primary productivity would have supported a greater biomass, due to the scaling of mass with metabolic rate \citep{mcnab2009resources}.
+%KH Need to make it clear why here. Only works for larger herbivores as well
+%TG: We're now talking about Late Mesozoic systems right? Needs to be mentioned somewhere.
+The upshot of this is that there was a higher biomass of herbivores dying and offering scavenging opportunities.
+%KH  I have a feeling the population dynamics to this are actually pretty complex.
+Large animals may live longer and represent a standing biomass similar to tropical rainforest's were all the nutrients are locked up in big trees.
+The life-history of these animals would hence be what really determines the production of carcasses however that is way outside the scope of this paper and would require a paper all on its own.
+Maybe we could drop in that point as I have not seen anyone ever mention this point.
+Predators were large-bodied too compared to extant mammalian predators \citep{mcnab2009resources}, and so, especially if they were ectothermic, could last longer between meals, rendering scavenging a more attractive behaviour relative to predation.
+%%KH Not sure this sentence belongs here, already covered int he handling section.
+Osteophagy may have been even more viable during the Mesozoic era as well because of this skewed body mass distribution of herbivores towards larger sizes \citep{10.1371/journal.pone.0051925}.
+When we couple this with the fact that skeletal mass scales greater than linearly with body mass \citep{prange1979scaling} there would have been a lot of bone material to consume in the environment provided an animal had the biology to process it \citep{chure1997one}.
+
+Frequently, the interplay between abiotic and biotic factors can impact the ability of an animal to scavenge. 
+We know vultures and eagles tend to soar using thermals and if these air pockets don't form, say on a cloudy day, the bird is grounded \citep{mundy1992vultures}.
+In many habitats (e.g. the Arctic) it is simply not possible for sufficiently powerful thermals to form and as a consequence large-bodied vultures cannot exist.
+%% KH I dont think this belong in this section, we probably have covered this enough in the locomotion part.
+One result of this is that terrestrial carnivores like bears and wolves take more carrion \citep{devault2003scavenging}.
+Certainly, a major difficulty for terrestrial scavengers is competition with vultures.
+Nocturnal behaviour in the hyaenidae in general has been put forth as an adaptation to reduce competition with these exclusively diurnal birds \citep{gittleman2013carnivore}.
+If we apply this line of reasoning over evolutionary time-scales, the absence of flying vertebrates in the Palaeozoic may have permitted terrestrial forms to take in a higher proportion of carrion in their diet.
+%%KH I think the focus needs to be more on the actual environment and opportunities available.
+I would suggest that competition either stays in the handling section (my preference) or is moved down to here (which also works really).
+%%KH You could then split the structure into production in terrestrial/ marine which focuses on primary productivity and species composition and how they affect what available then the next part being competition in both section.
+
+In fact, scavenging behaviour may have evolved on land as soon as the first terrestrial tetrapods emerged.
+Some of the earlier tetrapods tracks dating back to the early Middle Devonian (393.3 - 387.7 Mya) were found in intertidal environments \citep{Niedzwiedzki2009}.
+These environments are isolated from marine systems twice a day leaving potential carrion unexploited by marine vertebrates.
+\cite{Niedzwiedzki2009} suggest that these environments ``would thus have allowed marine ancestors of tetrapods gradually to acquire terrestrial competence while accessing a new and essentially untouched resource.''
+
+Staying in the aquatic setting, the phenomenon of occasional bounties of carrion in the form of whale falls has led some researchers to investigate if an obligate scavenger could survive by seeking out these remains exclusively.
+\cite{ruxton2005searching} argued that although this is energetically feasible it's ecologically unlikely.
+Any animal that could find such whale carcasses is unlikely to have ignored other types of carrion.
+Although no aquatic species have ever exceeded the size of whales, some enormous animals have evolved in this environment before the evolution of cetaceans, including \textit{Leedsichthys}, a bony fish from the Middle Jurassic (174.1-163.5 Mya) and the aquatic Mesozoic reptiles, the plesiosaurs, pliosaurs and ichtyosaurs, that could all exceed 15 metres in length \citep{ruxton2011zoology,danise2014ecological}.
+
+So, despite being unlikely, the energetic feasibility of a marine scavenger that specialises on large carcasses has a long history.
+One point of interest is that of the whaling industry, which provided a bonanza of floating carcasses especially during the 20th century \citep{Whitehead415}.
+This meant killer whales could switch from hunting to scavenging, a switch made that much easier by the noise of the whaling vessels that would effectively ring the ``dinner-bells" \citep{Whitehead415}.
+
+Perhaps the greatest environmental driver of scavenging tendency is that of temperature. 
+The geological record shows the Earth has undergone radical fluctuations in temperature over time.
+This will have had a significant bearing on the availability and persistence of carrion.
+To illustrate the point, a 10$^{\circ}$C increase in ambient temperature can double carcass decomposition rates \citep{parmenter2009carrion} and geological evidence indicates that the Mesozoic Earth was on average at least 6 $^{\circ}$C warmer than now \citep{sellwood2006mesozoic}.
+In terms of specific habitats, it has been shown that decomposition is greater in warm and moist areas versus more xeric ones \citep{beasley2015vertebrates}.
+Moreover, oceanic productivity and habitat structure are all impacted by climactic conditions.
+The impacts these can have on scavengers have been empirically supported e.g.
+\cite{beasley2015vertebrates} who point to a series of studies showing how microbes and invertebrates benefit at higher temperatures to the detriment of vertebrate scavengers such that ``above 20$^{\circ}$C vertebrates were able to detect and consume only 19\% of small-mammal carcasses, whereas at temperatures below 18$^{\circ}$C, vertebrates consumed 49\% of carcasses".
+This is a sobering thought given the impact we humans are having on the Earth's climate.
+
 % KH maybe my issue is that the aquatic, terrestrial, aerial structure, which works well for the locomotion, struggles in the the other sections.
@@
 \section*{Handling Time}
-Since the food a scavenger depends on is not dispatched directly, often the most easily accessible and choicest components of the carcass will be missing or, if present, will be subject to decay as well as competition.
-So being able to overcome competitors, maximise the nutrient gain from the remnants, and survive long enough between meals are all essential parts of carcass handling time. 
-% Being able to extract nutrients from remnants gives a scavenger a great advantage.
+Since the food a scavenger depends upon is not dispatched directly, often the most easily accessible and choicest components of the carcass will be missing or, if present, will be subject to decay as well as competition.
+So being able to overcome competitors and maximise the nutrient gain from the remnants are all essential parts of carcass handling time. 
+We will first consider the adaptations to defeating competitors.
 
-%Thus, the bone crushing ability of hyenas and others reveals another useful scavenger trait. %TG: how about the following for a smoother transition:
+As well as conferring starvation resistance, large body size has substantial advantages in agonistic interactions \citep{ruxton2004obligate}. 
+For instance, lions can acquire much of their carrion through kleptoparasitism of hyena kills \citep{trinkel2005competitive}. 
+This line of reasoning suggests that some theropod dinosaurs, who could get up to 15 tonnes, would have had easily monopolised a carcass \citep{weishampel2004dinosauria} provided they could find them \citep{kane2016body}.
 
-%KH still not sure about this transition.
-One adaptation is to utilize late carcass stage material that is less subject to decomposition and unavailable to other competitors, for example bone.
-% Thus, some phenotypic adaptations such as the ability to crush bone reveal another useful scavenger trait.
+We would expect this trait to be selected for even in the case of weight-constrained, scavenging fliers.
+This is true for wandering albatrosses (\textit{Diomedea exulans}), cinereous vultures (\textit{Aegypius monachus}) and condors (\textit{Vultur gryphus}, \textit{Gymnogyps californianus}) who all have body masses that can exceed 10 kg and represent some of the heaviest bird species capable of flight \citep{weimerskirch1992reproductive,ferguson2001raptors,donazar2002effects}.
+Indeed, such is the competitive advantage held by vultures over other facultative scavenging birds that temporal niche partitioning at the carcass has evolved \citep{kendall2013alternative,KaneVul}. 
+Additionally, many pterosaurs were far bigger again, with estimated body masses of over 200 kg in the Azhdarchids \citep{witton2010size}.
+Although \cite{witton2008reappraisal} argued that neck inflexibility and straight, rather than hooked jaw morphology points against Azhdarchids existing as \textit{obligate} scavengers, their terrestrial proficiency indicates they would have been comfortable foraging on the ground.
+Extant Marabou Storks (\textit{Leptoptilos crumenifer}) have a comparable morphology and are noted facultative scavengers \citep{monadjem2012survival} so it is reasonable to believe that these pterosaurs behaved similarly.
+
+Smaller species can compensate for a lack of individual body size by weight of numbers in competitive interactions. 
+This is true for a host of notable scavengers, such as vultures, early hominins and hyenas, who can dominate larger competitors provided they substantially outnumber them \citep{KaneVul,trinkel2005competitive,ruxton2013endurance}.
+
+In addition to fending off other vertebrates, scavengers also have to contend with competition from microorganisms, which requires a specialised physiology. 
+Although the findings of \cite{shivik2006vultures} that ``evolutionary pressures favor detection maximizers relative to toxification minimizers in competitive interactions for carcasses." appear sound, the fact remains that overcoming microorganism toxins is still a beneficial adaptation to any scavenger.
+Avian scavengers have evolved incredibly acidic stomachs that allow them to consume and process putrefied flesh with no ill effects \citep{houston1975digestive,roggenbuck2014microbiome}. 
+This adaptation is not restricted to vultures though, \cite{gremillet2012vultures} showed wandering albatrosses (\textit{Diomedea exulans}; so-called ``vultures of the seas'') had an average pH of 1.5, which enables them to consume fisheries discards and squid carcasses.
+There is also evidence of selection for ``toxification minimizers'' beyond birds among the ectotherms.
+From our earlier arguments we know that ecthotherms are limited in their ability to find carrion as quickly as endotherms. 
+These later arrivers would thus benefit especially from well-developed detoxifying apparatus. 
+\cite{shivik2006vultures} suggests that ``specialized oral structures in snakes may have evolved under pressures associated with scavenging."
+Moreover, some researchers have charted an evolutionary course from basal fossorial snakes to modern terrestrial species by way of an obligate scavenger intermediate \citep{bauchot2006snakes}.
+
+Aside from coping with competitors, another vital component of carrion handling time is the ability to maximise the energy gain from the remains. 
+At whale carcases, white and blue sharks are known to preferentially feed on the blubber layer \citep{long1996sharks}. 
+Blubber is an energy rich portion of the carcass that can allow a shark to survive for 1.5 months on 20 kg of the material \citep{carey1982temperature}. 
+On land many scavengers utilize late-stage carcass material that is less subject to decomposition and may be unavailable to other competitors, for example bone.
 Osteophagy is known across a range of terrestrial carnivores and given that some fat-rich mammalian bones have an energy density (6.7 kJ/g) comparable with that of muscle tissue, it makes skeletal remains an enticing resource \citep{brown1989study}.
@@
 Earlier in the evolution of mammals, the bone-crushing dogs that evolved during the Oligocene (Borophaginae; 33.9 - 23.03 Mya) have also been compared to hyenas in terms of their feeding ecology \citep{van2003chapter,martin2016pursuit}.
+
+In Mesozoic systems some large theropod dinosaurs had a morphology indicative of an ability to process bone \citep[e.g.
+the robust skull and dentition of \textit{Tyrannosaurus rex}][]{hone2010feeding}.
+There is direct evidence that \textit{T.
+rex} did this in the form of distinctive wear marks on its tooth apices \citep{farlow1994wear,schubert2005wear} and the presence of bone fragments in its coprolites \citep{chin1998king}.
+The animal also had an enormous bite force, with one estimate putting it at 57000 Newtons \citep{bates2012estimating} which would have been powerful enough to break open skeletons \citep{rayfield2001cranial}.
+
+Despite not having the anatomical ability to break open bone, the bearded vulture (\textit{Gypaetus barbatus}) has evolved a technique whereby it drops long bones from a height, splintering them on the rocks below which allows them to feed \citep{margalida2008bearded}.
+Similarly, early hominins developed the ability to craft tools for breaking open bones \citep{ARCM:ARCM12084}.
+A recent study investigating potential scavenging opportunities for hominins in Kenya found that, in addition to skeletal material, there is a substantial amount of scavengeable meat left on predated remains; sufficient to sustain the requirements of an adult male \textit{Homo erectus} \citep{pobiner2015new}.
+In some historical hominin-inhabited areas there were a greater number of felids than hyenids.
+Again, this is significant because hyenas are likely to have left far less flesh on a carcass than a felid such as a sabretooth, enabling contemporaneous hominins to benefit \citep{pobiner2015new}.
+The use of tools and the cooperative nature of hominins meant they could likely get a substantial part of their energetic requirements through scavenging depending on their environment \citep{moleon2014humans}.
+
+On the ground, and despite the advantages of social resource defence, the competitive ability of even the largest flying bird is radically diminished in their interactions with mammalian competitors, and as such they tend to consume carrion rapidly.
+\cite{houston1974role} observed a group of \textit{Gyps} vultures consuming all of the soft tissue from a 50 kg Grant’s gazelle (\textit{Nanger granti}) in eight minutes. 
+Their serrated tongues and hooked bills enabling them to achieve this feat \citep{houston1975digestive}. 
+Aside from raptors, the specialised beaks of many modern bird lineages tends to hinder their ability to eat meat which is in contrast to the first lineages that did not have this feature \citep{martyniuk2012field}.
+As \cite{martyniuk2012field} notes these early birds would thus have been predominantly carnivorous, implying scavenging was a live opportunity cf. their descendants. 
+Among the pterosaurs, \cite{witton2013pterosaurs} makes the case that the istiodactyl pterosaurs were the most likely scavengers of this group based on their potential handling time. 
+The mix of strong and weak features in their skull morphology is indicative of animals that were suited to removing large amounts of flesh from an immobile foodstuff \citep{witton2013pterosaurs}.
+
+Some clades appear to lack many, if not all, of the phenotypic adaptations associated with an efficient carcass handling time. 
+For example, the extant bats seem to lack all of the features we have identified as important.
+The larger forms (which are better suited for scavenging, following our previous arguments) are typically frugivores and therefore lack the adaptations for digesting meat while the smaller carnivorous bats are mainly found in the microbats, which are insectivorous \citep{aguirre2003implications}.
+Additionally, their poor terrestrial ability and cost of movement on the ground would also count against them while attempting to fend off competitors at a carcass \citep{riskin2006terrestrial,voigt2012terrestrial}.
+That said, Necromantis (“death-eater”), a large bat from the middle to late Eocene (56-33.9 Mya) had a robust cranio-mandibular morphology, and is a likely candidate for an extinct scavenging bat \citep{Weithofer_Necromantis_1887,Hand_Necromantis_2012}
+
+%KH still not sure about this transition.
 
@@
 
-Interestingly, such comparisons have given insight into the feeding ecology of early hominins who, for instance, had the ability to craft tools for breaking open bones \citep{ARCM:ARCM12084}.
-The question of where our ancestors placed on the hunter-scavenger axis during the Plio-Pleistocene has been a matter of debate for years \citep{dominguez2002hunting}.
-%%KH I think this shouldn't be the first place we mention humans as they do appear and fit well in other sections.
-%KH Here and in other section there is a lack of mentioning fossil/remains etc. For example I'm sure there is something about the tools or the marks on the bones that would fit in well here.
-
-A recent study investigating potential scavenging opportunities for hominins in Kenya found that, even when discounting bone material, there is a substantial amount of scavengeable meat left on predated remains; sufficient to sustain the requirements of an adult male \textit{Homo erectus} \citep{pobiner2015new}.
-In some historical hominin-inhabited areas there were a greater number of felids than hyenids.
-Again, this is significant because hyenas are likely to have left far less flesh on a carcass than a felid such as a sabretooth, enabling contemporaneous hominins to benefit \citep{pobiner2015new}.
-The use of tools and the cooperative nature of hominins meant they could likely get a substantial part of their energetic requirements through scavenging depending on their environment \citep{moleon2014humans}.
-%KH should this go in the prey availability section?
-
-In Mesozoic systems some extremely large theropod dinosaurs had a morphology indicative of an ability to process bone \citep[e.g.
-the robust skull and dentition of \textit{Tyrannosaurus rex}][]{hone2010feeding}.
-There is direct evidence that \textit{T.
-rex} did this in the form of distinctive wear marks on its tooth apices \citep{farlow1994wear,schubert2005wear} and the presence of bone fragments in its coprolites \citep{chin1998king}.
-The animal also had an enormous bite force, with one estimate putting it at 57000 Newtons \citep{bates2012estimating} which would have been powerful enough to break open skeletal material \citep{rayfield2001cranial}.
 %KH something on the robust morphology and size required to generate such forces however essentially restricts flying species from having such a strategy, unless they can use tools (dosent some bird fly up and then drop bones.)
@@
 
-We know that large body size confers substantial dominance and starvation-resistance benefits \citep{ruxton2004obligate}. 
-As such, theropod dinosaurs, who could get up to 15 tonnes, would seem likely candidates for scavenging. 
-Much work has focused on the existence of scavenging in dinosaurs by using simple energetics approaches that typically focused on a single species namely \textit{T.
-rex} \citep{ruxton2003could,carbone2011intra} but a recent modelling study investigated its prevalence across a range of body sizes \citep{kane2016body}.
-
-In their work, the authors demonstrated that species of \textit{intermediate} body masses would have gained the most benefit from scavenging \citep{kane2016body}.
-This was the result of gut capacity limitations and the effects of competition at the carcass.
-% KH I would mention that competition really only affects smaller individuals in this scenario.
-That is without competition in the model big things did about just as well but small ones did much better.
-Maybe take the shift in optimum between the models as an example of the affect.
-At the larger extreme this owes to the fact that gut capacity doesn't scale isometrically with body mass so the benefits of greater mass level off; there's only so much food an individual can consume at a single sitting \citep{calder1996size}.
-%% KH This sound specific to theropods while if I remember correctly its for mammals.
-I think we may need to say this in a more general aspect (again changing structure may help.
-(i.e a stomach section would allow snakes to be included and other extremes))
-For the smaller species, larger competitors would have prevented their access to carrion.
-
  The support of water allows for many aquatic species to reach large sizes thus granting its benefits. 
@@
 This, the authors pointed out, is because randomly distributed carrion is better exploited by fish with larger body sizes owing to starvation resistance.
-
-Given the advantages of size, we would expect this trait to be selected for even in the case of weight-constrained scavenging fliers.
-This is true for wandering albatrosses (\textit{Diomedea exulans}), cinereous vultures (\textit{Aegypius monachus}) and condors (\textit{Vultur gryphus}, \textit{Gymnogyps californianus}) who all have body masses that can exceed 10 kg and represent some of the heaviest bird species capable of flight \citep{weimerskirch1992reproductive,ferguson2001raptors,donazar2002effects}.
-Additionally, as we have noted the Azhdarchid pterosaurs were far bigger again, with estimated body masses of over 200 kg \citep{witton2010size}.
-Although \cite{witton2008reappraisal} argued that neck inflexibility and straight, rather than hooked jaw morphology points against Azhdarchids existing as \textit{obligate} scavengers, their terrestrial proficiency indicates they would have been comfortable foraging on the ground.
-Indeed, extant Marabou Storks (\textit{Leptoptilos crumenifer}) have a comparable morphology and are noted facultative scavengers \citep{monadjem2012survival} so it is reasonable to believe that these pterosaurs behaved similarly.
 
@@
 
-%TG: This part above makes sense in the adaptations parts (i.e.
-in the Handling Time part) but we can see when we clean down the whole paper if we keep it (then we need to integrate it) or if we have enough.
-I actually fits way better in the new narrative by the way!
-On the ground, the competitive ability of even the largest flying bird is radically diminished in their interactions with mammalian competitors however, and as such they tend to consume carrion rapidly.
-%TG: I've added "living" bird here.
-In the Paleocene you have these giant terror birds that were still top predators (like during the Cretaceous...
-How weird!
-;)).
-% AK How about flying bird because that's the key thing here TG: yep! Agree!
-\cite{houston1974role} observed a group of \textit{Gyps} vultures consuming all of the soft tissue from a 50 kg Grant’s gazelle (\textit{Nanger granti}) in eight minutes. 
-Their serrated tongues and hooked bills enabling them to achieve this feat \citep{houston1975digestive}. 
-Outside of raptors such as vultures, the specialised beaks of many modern bird lineages hinders their ability to eat meat which is in contrast to the first lineages that did not have this feature \citep{martyniuk2012field}.
-As \cite{martyniuk2012field} notes these early birds would thus have been predominantly carnivorous, which implies that scavenging would have been a live opportunity cf. their descendants. 
-Among the pterosaurs, \cite{witton2013pterosaurs} makes the case that the istiodactyl pterosaurs were the most likely scavengers of this group based on their potential handling time. 
-The mix of strong and weak features in their skull morphology is indicative of animals that were suited to removing large amounts of flesh from an immobile foodstuff \citep{witton2013pterosaurs}.
-
 Because of the random nature of carrion we would expect adaptations that reduce energetic costs of maintenance to be selected for in scavengers as it would maximise the benefit derived from such a sporadic food source.
@@
 The same reasoning can be applied to crocodiles and their allies \citep{forrest2003evidence} because a sit and wait strategy is viable for an ectotherm.
-%This low existence cost is also realised in many sharks who have coupled low locomotory costs with an ectothermic metabolism. 
-%
-
-Although the findings of \cite{shivik2006vultures} that ``evolutionary pressures favor detection maximizers relative to toxification minimizers in competitive interactions for carcasses." appears sound, the fact remains that overcoming microorganism toxins is still a beneficial adaptation to any scavenger.
-Avian scavengers have evolved incredibly acidic stomachs that allow them to consume and process putrefied flesh with no ill effects \citep{houston1975digestive,roggenbuck2014microbiome}. 
-This adaptation is not restricted to vultures though, \cite{gremillet2012vultures} showed wandering albatrosses (\textit{Diomedea exulans}; so-called ``vultures of the seas'') had an average pH of 1.5, which enables them to consume fisheries discards and squid carcasses.
-There is also evidence of selection for ``toxification minimizers'' beyond birds among the ectotherms.
-From our earlier arguments we know that ecthotherms are limited in their ability to find carrion as quickly as endotherms. 
-This implies later arrivers would benefit especially from well-developed detoxifying apparatus. 
-\cite{shivik2006vultures} suggests that ``specialized oral structures in snakes may have evolved under pressures associated with scavenging."
-Moreover, some researchers have suggested an evolutionary course from basal fossorial snakes to modern terrestrial species by way of an obligate scavenger intermediate \citep{bauchot2006snakes}.
-
-%Being far less vagile than other vertebrate species,
-%snakes are expected to develop detoxification strategies to overcome
-%chemical defenses and make the best use of carrion.
-%There is additional evidence that  Evolutionary pressures are not limited to competition
-%for carcasses, of course, but evolutionarily, as snakes
-%developed from eyeless fossorial species and radiated into terrestrial
-%and arboreal predatory species (Rage 1994), an obligate
-%scavenger evolutionary intermediate was likely.
-
-%TG: Two points here: one speculation for the non-scavenging behaviour could be because of diet.
-Conversely, entire clades appear to lack many, if not all, of these phenotypic adaptations. %TG: I've tried some transition here
-For example, the extant bats appear to lack most of the features we have identified as important in reducing handling time.
-The larger forms (which are better suited for scavenging, following our previous arguments) are typically frugivores and therefore lack the adaptations for digesting meat.
-While the smaller carnivorous bats are mainly found in the microbats which are insectivorous \citep{aguirre2003implications}.
-Additionally, their poor terrestrial ability and cost of movement on the ground would also count against them when feeding at a carcass \citep{riskin2006terrestrial,voigt2012terrestrial}.
 
 % Durophagy amount aquatic species, turtles, sharks, trigger fish, crocs etc.
-
-\section*{Prey Availability}
-%Both the biotic and abiotic environment a would-be scavenger finds itself in can influence to degree to which it can depend on carrion. %TG: This sentence is convoluted: maybe go with the following:
-The position of a species on the scavenging scale can also be influenced by the availability of carrion in the environment, which is dependent on biotic and abiotic factors.
-Aspects including, primary productivity, relief, temperature and competition will all greatly affect scavenging tendency. 
-\cite{ruxton2004obligate} suggest a system with a productivity similar to the Serengeti could have supported an obligate mammalian or reptilian terrestrial scavenger.
-%% KH Be careful about slipping between scavenger and obligate scavenger.
-Indeed, in systems that were dominated by large ectothermic or mesothermic herbivore vertebrates, the same primary productivity would have supported a greater biomass, due to the scaling of mass with metabolic rate \citep{mcnab2009resources}.
-%KH Need to make it clear why here. Only works for larger herbivores as well
-%TG: We're now talking about Late Mesozoic systems right? Needs to be mentioned somewhere.
-The upshot of this is that there was a higher biomass of herbivores dying and offering scavenging opportunities.
-%KH  I have a feeling the population dynamics to this are actually pretty complex.
-Large animals may live longer and represent a standing biomass similar to tropical rainforest's were all the nutrients are locked up in big trees.
-The life-history of these animals would hence be what really determines the production of carcasses however that is way outside the scope of this paper and would require a paper all on its own.
-Maybe we could drop in that point as I have not seen anyone ever mention this point.
-Predators were large-bodied too compared to extant mammalian predators \citep{mcnab2009resources}, and so, especially if they were ectothermic, could last longer between meals, rendering scavenging a more attractive behaviour relative to predation.
-%%KH Not sure this sentence belongs here, already covered int he handling section.
-Osteophagy may have been even more viable during the Mesozoic era as well because of this skewed body mass distribution of herbivores towards larger sizes \citep{10.1371/journal.pone.0051925}.
-When we couple this with the fact that skeletal mass scales greater than linearly with body mass \citep{prange1979scaling} there would have been a lot of bone material to consume in the environment provided an animal had the biology to process it \citep{chure1997one}.
-% As we discussed earlier, this ability is often extremely beneficial to a scavenger.
-
-Frequently, the interplay between abiotic and biotic factors can impact the ability of an animal to scavenge. 
-We know vultures and eagles tend to soar using thermals and if these air pockets don't form, say on a cloudy day, the bird is grounded \citep{mundy1992vultures}.
-In many habitats (e.g. the Arctic) it is simply not possible for sufficiently powerful thermals to form and as a consequence large-bodied vultures cannot exist.
-%% KH I dont think this belong in this section, we probably have covered this enough in the locomotion part.
-One result of this is that terrestrial carnivores like bears and wolves take more carrion \citep{devault2003scavenging}.
-Certainly, a major difficulty for terrestrial scavengers is competition with vultures.
-Nocturnal behaviour in the hyaenidae in general has been put forth as an adaptation to reduce competition with these exclusively diurnal birds \citep{gittleman2013carnivore}.
-If we apply this line of reasoning over evolutionary time-scales, the absence of flying vertebrates in the Palaeozoic may have permitted terrestrial forms to take in a higher proportion of carrion in their diet.
-%%KH I think the focus needs to be more on the actual environment and opportunities available.
-I would suggest that compititon either stays in the handling section (my preference) or is moved down to here (which also works really).
-%%KH You could then split the structure into production in terrestial/marine which focuses on primary productivity and species composition and how they affect what available then the next part being competition in both section.
-
-%TG: Added the early tetrapod bit here:
-% This absence or reduced competition in before the Jurassic might have allowed some species to exploit these scavenger niches.
-In fact, scavenging behaviour may have evolved on land as soon as the first terrestrial tetrapods emerged.
-Some of the earlier tetrapods tracks dating back to the early Middle Devonian (393.3 - 387.7 Mya) were found in intertidal environments \citep{Niedzwiedzki2009}.
-These environments are isolated from marine systems twice a day leaving potential carrion unexploited by marine vertebrates.
-\cite{Niedzwiedzki2009} suggest that these environments ``would thus have allowed marine ancestors of tetrapods gradually to acquire terrestrial competence while accessing a new and essentially untouched resource.''
-
-% The use of different sensory systems also illustrates the impact of the environment. 
-% The relatively open savanna systems of Africa are well suited to a visually dependent vulture whereas more forested areas would select for species that have a well-developed olfactory system  \citep{houston1986olfaction}.
-% Again, a similar line of reasoning can be applied to aquatic species depending whether they forage near the well-lit surface or the dark benthos. 
-% Although, as with detection ability, the environment has a role to play here.
-
-%TG: maybe change transition here.
-Staying in the aquatic setting, the phenomenon of occasional bounties of carrion in the form of whale falls has led some researchers to investigate if an obligate scavenger could survive by seeking out these remains exclusively.
-\cite{ruxton2005searching} argued that although this is energetically feasible it's ecologically unlikely.
-Any animal that could find such whale carcasses is unlikely to have ignored other types of carrion.
-Although no aquatic species have ever exceeded the size of whales, some enormous animals have evolved in this environment before the evolution of cetaceans, including \textit{Leedsichthys}, a bony fish from the Middle Jurassic (174.1-163.5 Mya) and the aquatic Mesozoic reptiles, the plesiosaurs, pliosaurs and ichtyosaurs, that could all exceed 15 metres in length \citep{ruxton2011zoology} (reptile fall ref here).
-
-%% KH check out this paper on Mesozoic reptile falls http://www.nature.com/ncomms/2014/140910/ncomms5789/full/ncomms5789.html
-So, despite being unlikely, the energetic feasibility of a marine scavenger that specialises on large carcasses has a long history.
-One point of interest is that of the whaling industry, which provided a bonanza of floating carcasses especially during the 20th century \citep{Whitehead415}.
-This meant killer whales could switch from hunting to scavenging, a switch made that much easier by the noise of the whaling vessels that would effectively ring the ``dinner-bells" \citep{Whitehead415}.
-
-Perhaps the greatest environmental driver of scavenging tendency is that of temperature. 
-The geological record shows the Earth has undergone radical fluctuations in temperature over time.
-This will have had a significant bearing on the availability and persistence of carrion.
-To illustrate the point, a 10$^{\circ}$C increase in ambient temperature can double carcass decomposition rates \citep{parmenter2009carrion} and geological evidence indicates that the Mesozoic Earth was on average at least 6 $^{\circ}$C warmer than now \citep{sellwood2006mesozoic}.
-In terms of specific habitats, it has been shown that decomposition is greater in warm and moist areas versus more xeric ones \citep{beasley2015vertebrates}.
-Moreover, oceanic productivity and habitat structure are all impacted by climactic conditions.
-The impacts these can have on scavengers have been empirically supported e.g.
-\cite{beasley2015vertebrates} who point to a series of studies showing how microbes and invertebrates benefit at higher temperatures to the detriment of vertebrate scavengers such that ``above 20$^{\circ}$C vertebrates were able to detect and consume only 19\% of small-mammal carcasses, whereas at temperatures below 18$^{\circ}$C, vertebrates consumed 49\% of carcasses".
-This is a sobering thought given the impact we humans are having on the Earth's climate. 
-%TG: Really nice to finish on this paragraph + good for transition to conclusion!
 

### 2016-04-26 natural history of scavenging.tex
moved around sections
paper is effectively encounter rate and handling time

@@
 However, species that rely on scavenging to sustain substantial portions of their diets must encounter a sufficient amount of carrion in order to meet their energetic demands.
-Once found, the scavenger must be able to out-compete any potential competitors and process the, typically decaying, carcass replete with microorganism derived toxins  \citep{ruxton2014fruit}. 
-Both of these characteristics can be assumed to be under evolutionary selection pressures for traits that increase carrion discovery and monopoly.
-Finally, the potential for scavenging will also depend on the density, size, and quality of carcases produced, all of which are affected by complex ecosystem dynamics and are outside the selection pressures on the scavenger.
+Once found, the scavenger must be able to out-compete any potential competitors and process the, typically decaying, carcass replete with micro-organism derived toxins  \citep{ruxton2014fruit}. 
+These characteristics can be assumed to be under evolutionary selection pressures for traits that increase carrion discovery and monopoly.
+Finally, the potential for scavenging will also depend on the density, size, and quality of carcases produced, all of which are affected by complex ecosystem dynamics but are outside the selection pressures on the scavenger.
 Each of these facets are essentially the backbone of fundamental ecological theory and are the key parameters defined in functional response curves, namely encounter rate, handling time and prey availability \citep{jeschke2002predator}.
 By considering scavenging in this context of optimal foraging we can identify the prerequisite attributes and processes required for the behaviour. 
-This has enabled us to propose a scale of scavenging whereupon we can place any vertebrate species, past or present, and assess the importance of carrion in its diet.
+This has enabled us to propose a scale of scavenging whereupon we can place any vertebrate species, past or present, and assess the likely importance of carrion in its diet.
 
@@
 Locomotory speed, foraging time and detection radius all increase the encounter rate between a scavenger and the carcasses it is searching for. 
+Alternatively, a species can reduce its metabolic requirements so that it can survive long periods between meals. 
 We would thus expect selection pressures to act on the various traits that govern these parameters.
-% However, encounter rate is also determined by the productivity of the environment, something which the selective forces acting on the scavenging have no control over.
+
+However, as we noted above, encounter rate is also determined by the productivity of the environment, something which the selective forces acting on the scavenging have no control over.
 
@@
 
+\subsection*{Metabolism}
+Because of the sporadic nature of carrion we would expect adaptations that reduce energetic costs of maintenance to be selected for in scavengers as it would allow for longer inter-feeding periods. 
+Extant reptiles possess an advantage here, in that over the course of a year their food requirements can be 30 times lower than an endotherm of equal size \citep{Nagy1621}.
+\cite{devault2002scavenging} suggest this is an avenue for scavenging in snakes because they ``exhibit exceedingly low maintenance metabolisms, and most can survive on a few scant feedings per year.
+It is, therefore, possible for snakes to rely largely on infrequent, less energy-rich meals."
+In the same review the authors found occurrences of scavenging spread across five families of snakes and stated that this behaviour is ``far more common than currently acknowledged."\citep{devault2002scavenging}.
+The same reasoning can be applied to crocodiles and their allies \citep{forrest2003evidence} because a sit and wait strategy is viable for an ectotherm \citep{forrest2003evidence,moleon2015carcass}. 
+\cite{carey1982temperature} found ectothermic sharks have the ability to exist weeks between meals because they focus on the energy-rich sections of carcasses (see Handling Time). 
+Endotherms have also evolved physiological mechanisms that allow them to depress their otherwise high metabolic rates at certain times e.g.
+vultures who do so while resting at their roost and during periods of food deprivation \citep{bahat1998nocturnal}.
+
 \subsection*{Locomotion}
-Because of the inherent unpredictability of carrion, scavenging depends more on the ability to efficiently move over large areas than predation.
+Because of the inherent unpredictability of carrion, scavenging depends more on the ability to efficiently move over large areas than does predation.
 This generally requires an efficient transfer of metabolic energy into movement which relies on the animal's anatomy and physiology as well as the medium of the environment in which the animal is moving (i.e.
@@
 Azhdarchids with wingspans of 11 metres; ][]{witton2010size} and, notably, appear to have engaged in soaring flight \citep{witton2010size}.
-It seems probable that as least some of these extinct species used soaring as a means for scavenging \citep{witton2013pterosaurs}. 
+It seems probable that at least some of these extinct species used soaring as a means for scavenging \citep{witton2013pterosaurs}. 
 %KH added "at least some of these" as it was not clear what species used soaring as a means for scavenging.
@@
 Sharks are one likely candidate for general scavenging behaviors as their locomotion, which depends on large pectoral fins to generate lift as they swim, resembles that of the large soaring fliers.
-%changed this sentence up a bit 
 Many shark species have large foraging ranges \citep[e.g.
 the great white sharks \textit{Carcharodon carcharias};][]{bruce2006movements} and it seems reasonable that they would use oceanographic currents to further reduce movement costs \citep{ruxton2004energetic}.
-%and are able to maintain a type of endothermy from thier movements \citep{bernal2015sharks}.
-%TG: I guess the following is a comment: (I know the Greenland shark is the slowest swimmer).
 In fact, facultative scavenging is seen in many selachian groups, including species of extant sharks like white sharks \citep[known to feed on whale carcasses;][]{fallows2013white}, Greenland sharks \citep[feeding on seals;][]{watanabe2012slowest}, and sixgill sharks \citep{anderson2016impact}.
@@
 However I think this kinda solves our problems in the complexity of swimming by saying apart from large ranges ability to eat is probably more important.
-%KH also is fish an ok term to use, will people get pissy or anything like with reptiles?
 
-% KH endotermy - why it doesn't seem to work
 We might expect then that by combining an aquatic environment and an endothermic metabolism marine mammals would especially prosper as scavengers.
@@
 Terrestrial environments are the most energetically costly in which to move \citep{tucker1975energetic}. 
-Unlike aerial and aquatic environments support must be provided through the animals posture leading to trade offs between low cost of maintain and the efficiency of transport.
-For example, the sprawling posture of many ectotherm groups (except frogs!) would be expected to allow for a lower maintenance cost relating to body support in comparison to groups with the limbs position underneath the body, such as mammals, aves and dinosaurs.
+Unlike aerial and aquatic environments support must be provided through the animal's posture. % leading to trade offs between low cost of maintain and the efficiency of transport. 
+The early transition from a sprawling gait, seen in early tetrapods, to the more erect posture of synapsids and later dinosaurs and mammals, has often been supposed as conferring a huge advantage to the latter groups \citep{sullivan2015posture}.
+The purported advantages include benefits in terms of speed, efficiency, muscle effort and manoeuvrability \citep{sullivan2015posture}.
+Clearly, for a scavenger, an ability to efficiently cover an area at a high speed would increase the encounter rate with carrion. 
+Despite being intuitive, \cite{sullivan2015posture} states most of the hypotheses in favour of this idea remain to be tested in the context of archosaur evolution. 
+One noted consequence of a sprawling gait is the phenomenon known as Carrier's constraint such that the animal can't move and undergo costal ventilation at the same time because the lateral movements impedes its lungs \citep{carrier1987evolution}.
+The evolution of an upright posture has been offered as one of the primary mechanisms that allowed early archosaurs to overcome this constraint \citep{uriona2008recruitment}.
+
+% For example, the sprawling posture of many ectotherm groups (except frogs!) would be expected to allow for a lower maintenance cost relating to body support in comparison to groups with the limbs position underneath the body, such as mammals, aves and dinosaurs.
 %KH: I agree with TG and have removed the gas exchange thing as its a physiology thing only applicable to certain groups and not got to do with terrestrial environments really. I also removed the 
@@
 
-%
-%TG: It's missing a transition here, we go from gas exchanges to posture.
-Maybe we need something like "Among terrestrial scavengers, most are mammals.
-This might be due to the posture blablabalba." But then it doesn't connect logically with the sentence before: "mammals have are inefficient...".
-Maybe one way is to just remove the gas exchange and start with something like:
-%Terrestrial environments are the most energetically costly in which to move, which may be due to the low muscular efficiency of running \citep{tucker1975energetic}.
-%Most terrestrial species that have been observed (or are suspected to be) scavenging are mammals.
-%TG: and then link to the posture hypothesis (with the "meh, we're not sure" at the end).
-%
-While the low maintenance costs of a sprawling like posture may reduce the need to scavenge frequently, the ability to search large areas is several restricted in these groups hence likely reducing such opportunities (ref on home range of terrestrial ectoterms).
-For example, a sprawling gait can results in the phenomenon known as Carrier's constraint such that the animal can't move and breathe at the same time because the lateral movements impedes its lungs \citep{carrier1987evolution}.
-On the other hand the more erect body positions and often associated higher basal metabolic rates, allow for sustained locomotion over long periods of time and hence allow such species to search large areas.
+% While the low maintenance costs of a sprawling-like posture may reduce the need to scavenge frequently, the ability to search large areas quickly is severely restricted in these groups reducing scavenging opportunities (ref on home range of terrestrial ectoterms).
+% For example, a sprawling gait results in 
+% On the other hand the more erect body positions and often associated higher basal metabolic rates, allow for sustained locomotion over long periods of time allowing such species to search large areas.
 
-The early transition from a sprawling gait, such as seen in \textit{Tiktaalik}, to the more erect posture of Synapsids and later dinosaurs and mammals, has often been supposed as conferring a huge advantage to the latter.
 %KH Even after changing it I dont like this transition type narrative.
 %KH also needs to be more clear on what the advantage to scavenging is for these groups.
-%The purported advantages include benefits in terms of speed, efficiency, muscle effort and manoeuvrability \citep{sullivan2015posture}. 
+%
 %none of these advantages refer very clearly to scavenging
-Despite being intuitive, \cite{sullivan2015posture} states most of the hypotheses in favour of this idea remain to be tested in the context of archosaur evolution. 
 %KH I'm always wary of deeming something as intuitive,
@@
 %Both these groups are more likely to depend on low energetic costs in order to compensate for low encounter rates due to their inability to move.
+
+% Although the earliest evidence of vertebrate scavenging comes from the Permian (298.9 - 252.17 Mya) where a temnospondyl amphibian fed on the carcass of \textit{Varanops}, a predatory synapsid of the time \citep{reisz2006articulated}.
 Whatever the case, it is with the evolution of endothermy in the therapsid-mammal lineage \citep{clarke2010temperature} that terrestrial vertebrates would have gained the ability to range more widely, a vital component in seeking out carrion.
-Although the earliest evidence of vertebrate scavenging comes from the Permian (298.9 - 252.17 Mya) where a temnospondyl amphibian fed on the carcass of \textit{Varanops}, a predatory synapsid of the time \citep{reisz2006articulated}.
-
 Modern endothermic mammals can sustain longer periods of energetically expensive activity \citep{bennett1979endothermy} resulting in larger foraging ranges. 
@@
 Our locomotory efficiency has fed into the question of where our ancestors placed on the hunter-scavenger axis during the Plio-Pleistocene, which has been a matter of debate for years \citep{dominguez2002hunting}.
-\cite{ruxton2013endurance} added to this debate with his argument that long distance endurance running, often used by humans for hunting, was not an important feature of hominin scavenging.
-He arrives at this conclusion on the basis of the high cost of running coupled with the fact that dangerous competitors will tend to arrive beforehand.
+\cite{ruxton2013endurance} added to this debate with their argument that long distance endurance running, often used by humans for hunting, was not an important feature of hominin scavenging.
+They arrive at this conclusion on the basis of the high cost of running coupled with the fact that dangerous competitors will tend to arrive beforehand.
 
@@
 While the locomotory ability of theropods has been debated since their first inception, more recent studies have reconstructed them as relatively mobile animals \citep{pontzer2009biomechanics}. 
-Despite some suggestions that larger species may have had some advantage in scavenging, partially due to the ability to search large areas \citep{ruxton2003could}, more recent studies have shown that the energetic demands of such large forms meant scavenging was likely more prevalent in mid-sized theropods of approximately half a tonne \citep{carbone2011intra,kane2016body}
+Despite some suggestions that larger species may have had some advantage in scavenging, partially due to the ability to search large areas \citep{ruxton2003could}, more recent studies have shown that the energetic demands of such large forms meant scavenging was likely more prevalent in mid-sized theropods of approximately half a tonne \citep{carbone2011intra,kane2016body}.
 
@@
 
-As predicted by the importance of an increased encounter rate, scavengers have well-developed senses, with the visual and olfactory sensory systems most often associated with scavenging behavior. 
+As predicted by the importance of an increased encounter rate, known scavengers have evolved well-developed senses, with the visual and olfactory sensory systems most often associated with scavenging behavior.
 This is perhaps no surprise because sensory systems that rely on detecting signals associated with living animals, such as audioception, electroreception, thermoreception and echolocation will be limited in their ability to detect an already dead animal.
@@
 This approach was used by \cite{zelenitsky2011evolution} to hypothesise such a mode for the theropod dinosaur \textit{Bambiraptor} and by \cite{witmer2009new} for tyrannosaurs.
-The flying pterosaurs however, had tiny olfactory bulbs indicating this sense was not relied on \citep{witton2013pterosaurs}
+The flying pterosaurs however, had tiny olfactory bulbs indicating this sense was not relied on \citep{witton2013pterosaurs}.
 
@@
 Eagles too are known to have highly developed vision \citep{reymond1985spatial}.
-It follows that the evolution of flight allowed aerial animals to detect far more carrion than their terrestrial counterparts through vision \citep{AR:AR22815} as the pterosaurs convergently evolved large orbits and optic lobes \citep{witton2013pterosaurs}.
+The flying pterosaurs also convergently evolved large orbits and optic lobes \citep{witton2013pterosaurs}. 
+It follows that the evolution of flight allowed aerial animals to detect far more carrion than their terrestrial counterparts through vision \citep{AR:AR22815}.
 
@@
 
-Although aquatic species also have a vertical component because they forage in a water body, they must contend with the low-light environment where visual detection distances are far lower ($<$ 100 m) than they would be for air.
+Although aquatic species also have a vertical component to their environment, they must contend with low-light levels where visual detection distances are far lower ($<$ 100 m) than they would be for air.
 As such, aquatic animals detect resources through chemo- and mechanoreception more so than through vision \citep{ruxton2004energetic}.
@@
 
-\section*{Prey Availability}
+\subsection*{Prey Availability}
 The environmental influence on prey availability is an aspect that greatly affects encounter rate but is invisible to the selective forces acting on the scavenger. 
 Aspects including, primary productivity, relief, and temperature will all greatly affect scavenging tendency. 
-\cite{ruxton2004obligate} suggest a system with a productivity similar to the Serengeti could have supported an obligate mammalian or reptilian terrestrial scavenger.
-%% KH Be careful about slipping between scavenger and obligate scavenger.
+\cite{ruxton2004obligate} suggest an historic ecosystem with a productivity similar to the Serengeti could have supported an \textit{obligate} mammalian or reptilian terrestrial scavenger.
 Indeed, in systems that were dominated by large ectothermic or mesothermic herbivore vertebrates, the same primary productivity would have supported a greater biomass, due to the scaling of mass with metabolic rate \citep{mcnab2009resources}.
-%KH Need to make it clear why here. Only works for larger herbivores as well
-%TG: We're now talking about Late Mesozoic systems right? Needs to be mentioned somewhere.
-The upshot of this is that there was a higher biomass of herbivores dying and offering scavenging opportunities.
-%KH  I have a feeling the population dynamics to this are actually pretty complex.
-Large animals may live longer and represent a standing biomass similar to tropical rainforest's were all the nutrients are locked up in big trees.
-The life-history of these animals would hence be what really determines the production of carcasses however that is way outside the scope of this paper and would require a paper all on its own.
-Maybe we could drop in that point as I have not seen anyone ever mention this point.
-Predators were large-bodied too compared to extant mammalian predators \citep{mcnab2009resources}, and so, especially if they were ectothermic, could last longer between meals, rendering scavenging a more attractive behaviour relative to predation.
-%%KH Not sure this sentence belongs here, already covered int he handling section.
-Osteophagy may have been even more viable during the Mesozoic era as well because of this skewed body mass distribution of herbivores towards larger sizes \citep{10.1371/journal.pone.0051925}.
-When we couple this with the fact that skeletal mass scales greater than linearly with body mass \citep{prange1979scaling} there would have been a lot of bone material to consume in the environment provided an animal had the biology to process it \citep{chure1997one}.
-
-Frequently, the interplay between abiotic and biotic factors can impact the ability of an animal to scavenge. 
-We know vultures and eagles tend to soar using thermals and if these air pockets don't form, say on a cloudy day, the bird is grounded \citep{mundy1992vultures}.
-In many habitats (e.g. the Arctic) it is simply not possible for sufficiently powerful thermals to form and as a consequence large-bodied vultures cannot exist.
-%% KH I dont think this belong in this section, we probably have covered this enough in the locomotion part.
-One result of this is that terrestrial carnivores like bears and wolves take more carrion \citep{devault2003scavenging}.
-Certainly, a major difficulty for terrestrial scavengers is competition with vultures.
-Nocturnal behaviour in the hyaenidae in general has been put forth as an adaptation to reduce competition with these exclusively diurnal birds \citep{gittleman2013carnivore}.
-If we apply this line of reasoning over evolutionary time-scales, the absence of flying vertebrates in the Palaeozoic may have permitted terrestrial forms to take in a higher proportion of carrion in their diet.
-%%KH I think the focus needs to be more on the actual environment and opportunities available.
-I would suggest that competition either stays in the handling section (my preference) or is moved down to here (which also works really).
-%%KH You could then split the structure into production in terrestrial/ marine which focuses on primary productivity and species composition and how they affect what available then the next part being competition in both section.
+The upshot of this may have been a higher biomass of herbivores dying and offering scavenging opportunities (although these larger species may have also lived longer).
 
@@
 
-% KH maybe my issue is that the aquatic, terrestrial, aerial structure, which works well for the locomotion, struggles in the the other sections.
-It kinda leads to sentences that look like they were dropped in and from the discussion to prematurely end on one point before its fully developed.
-% KH an alternative structure would be process type structure that would follow the order of events at the carcass. Something like competition, processing the food and then digestion?
+%Predators were large-bodied too compared to extant mammalian predators \citep{mcnab2009resources}, and so, especially if they were ectothermic, could last longer between meals, rendering scavenging a more attractive behaviour relative to predation.
+%%KH Not sure this sentence belongs here, already covered int he handling section.
+
+%Frequently, the interplay between abiotic and biotic factors can impact the ability of an animal to scavenge. 
+%We know vultures and eagles tend to soar using thermals and if these air pockets don't form, say on a cloudy day, the bird is grounded \citep{mundy1992vultures}.
+%In many habitats (e.g. the Arctic) it is simply not possible for sufficiently powerful thermals to form and as a consequence large-bodied vultures cannot exist.
+%% KH I dont think this belong in this section, we probably have covered this enough in the locomotion part.
+%One result of this is that t.
+%Certainly, a major difficulty for terrestrial scavengers is competition with vultures.
+%Nocturnal behaviour in the hyaenidae in general has been put forth as an adaptation to reduce competition with these exclusively diurnal birds \citep{gittleman2013carnivore}.
+%%KH I think the focus needs to be more on the actual environment and opportunities available.
+I would suggest that competition either stays in the handling section (my preference) or is moved down to here (which also works really).
+%%KH You could then split the structure into production in terrestrial/ marine which focuses on primary productivity and species composition and how they affect what available then the next part being competition in both section.
 
@@
 So being able to overcome competitors and maximise the nutrient gain from the remnants are all essential parts of carcass handling time.
-We will first consider the adaptations to defeating competitors.
 
-As well as conferring starvation resistance, large body size has substantial advantages in agonistic interactions \citep{ruxton2004obligate}. 
+\subsection*{Competition}
+Large body size has substantial advantages in agonistic interactions \citep{ruxton2004obligate}. 
 For instance, lions can acquire much of their carrion through kleptoparasitism of hyena kills \citep{trinkel2005competitive}. 
-This line of reasoning suggests that some theropod dinosaurs, who could get up to 15 tonnes, would have had easily monopolised a carcass \citep{weishampel2004dinosauria} provided they could find them \citep{kane2016body}.
+This line of reasoning suggests that some theropod dinosaurs, who could get up to 15 tonnes, would have had easily monopolised a carcass \citep{weishampel2004dinosauria} provided they could find them efficiently \citep{kane2016body}.
 
@@
 
+By contrast, extant bats seem poorly equipped to deal with competitors. 
+Their poor terrestrial ability, smalls size and cost of movement on the ground would count against them while attempting to fend off other species at a carcass \citep{riskin2006terrestrial,voigt2012terrestrial}.
+
 Smaller species can compensate for a lack of individual body size by weight of numbers in competitive interactions. 
@@
 
-In addition to fending off other vertebrates, scavengers also have to contend with competition from microorganisms, which requires a specialised physiology. 
-Although the findings of \cite{shivik2006vultures} that ``evolutionary pressures favor detection maximizers relative to toxification minimizers in competitive interactions for carcasses." appear sound, the fact remains that overcoming microorganism toxins is still a beneficial adaptation to any scavenger.
+Direct confrontation can be circumvented by certain behavioural adaptations. 
+The evolution of nocturnal behaviour in some mammals, for instance, has been put forth as an adaptation to reduce competition with the exclusively diurnal vultures \citep{gittleman2013carnivore}. 
+In areas absent of vultures such as the Arctic, terrestrial carnivores like bears and wolves take more carrion \citep{devault2003scavenging}
+Thus, in the Palaeozoic, the absence of flying vertebrate competitors may have permitted terrestrial forms to take in a higher proportion of carrion in their diet.
+
+In addition to fending off other vertebrates, scavengers also have to contend with competition from micro-organisms, which requires a specialised physiology. 
+Although the findings of \cite{shivik2006vultures} that ``evolutionary pressures favor detection maximizers relative to toxification minimizers in competitive interactions for carcasses." appear sound, the fact remains that overcoming micro-organism toxins is still a beneficial adaptation to any scavenger.
 Avian scavengers have evolved incredibly acidic stomachs that allow them to consume and process putrefied flesh with no ill effects \citep{houston1975digestive,roggenbuck2014microbiome}. 
@@
 
+\subsection*{Food Processing}
 Aside from coping with competitors, another vital component of carrion handling time is the ability to maximise the energy gain from the remains. 
 At whale carcases, white and blue sharks are known to preferentially feed on the blubber layer \citep{long1996sharks}. 
-Blubber is an energy rich portion of the carcass that can allow a shark to survive for 1.5 months on 20 kg of the material \citep{carey1982temperature}. 
+Blubber is an energy rich portion of the carcass that can allow a shark to survive for 1.5 months on 30 kg of the material \citep{carey1982temperature}. 
 On land many scavengers utilize late-stage carcass material that is less subject to decomposition and may be unavailable to other competitors, for example bone.
@@
 In Mesozoic systems some large theropod dinosaurs had a morphology indicative of an ability to process bone \citep[e.g.
-the robust skull and dentition of \textit{Tyrannosaurus rex}][]{hone2010feeding}.
+the robust skull and dentition of \textit{Tyrannosaurus rex;}][]{hone2010feeding}.
 There is direct evidence that \textit{T.
@@
 The animal also had an enormous bite force, with one estimate putting it at 57000 Newtons \citep{bates2012estimating} which would have been powerful enough to break open skeletons \citep{rayfield2001cranial}.
+Osteophagy may have been even more viable during the Mesozoic era as well because of this skewed body mass distribution of herbivores towards larger sizes \citep{10.1371/journal.pone.0051925}.
+When we couple this with the fact that skeletal mass scales greater than linearly with body mass \citep{prange1979scaling} there would have been a lot of bone material to consume in the environment provided an animal had the biology to process it \citep{chure1997one}.
 
@@
 
-Some clades appear to lack many, if not all, of the phenotypic adaptations associated with an efficient carcass handling time. 
-For example, the extant bats seem to lack all of the features we have identified as important.
-The larger forms (which are better suited for scavenging, following our previous arguments) are typically frugivores and therefore lack the adaptations for digesting meat while the smaller carnivorous bats are mainly found in the microbats, which are insectivorous \citep{aguirre2003implications}.
-Additionally, their poor terrestrial ability and cost of movement on the ground would also count against them while attempting to fend off competitors at a carcass \citep{riskin2006terrestrial,voigt2012terrestrial}.
-That said, Necromantis (“death-eater”), a large bat from the middle to late Eocene (56-33.9 Mya) had a robust cranio-mandibular morphology, and is a likely candidate for an extinct scavenging bat \citep{Weithofer_Necromantis_1887,Hand_Necromantis_2012}
+Again, we can draw a comparison with clades that are lacking in these features. 
+As we mentioned earlier the skull morphology of extinct sabretooths meant they would have left a large amount of flesh on a carcass and were unable to feed on bones \citep{palmqvist2011giant}. 
+Extant bats are also poorly equipped when it comes to feeding on carrion; the larger forms are typically frugivores and therefore lack the adaptations for digesting meat, while the smaller carnivorous bats are mainly found in the microbats, which are insectivorous \citep{aguirre2003implications}.
+That said, Necromantis (“death-eater”), a large bat from the middle to late Eocene (56-33.9 Mya) had a robust cranio-mandibular morphology, and is a likely candidate for an extinct scavenging bat \citep{Weithofer_Necromantis_1887,Hand_Necromantis_2012}.
+
+% Some clades appear to lack many, if not all, of the phenotypic adaptations associated with an efficient carcass handling time. 
+% The extant bats seem to lack all of the features we have identified as important.
 
@@
 
- The support of water allows for many aquatic species to reach large sizes thus granting its benefits. 
-\cite{collins2005trends} found "contrasting relationships between size (body mass) and depth in the scavenging and predatory demersal ichthyofauna". 
-Predatory species saw a reduction in body mass with depth whereas the reverse trend was true for scavengers. 
-This, the authors pointed out, is because randomly distributed carrion is better exploited by fish with larger body sizes owing to starvation resistance.
+ % The support of water allows for many aquatic species to reach large sizes thus granting its benefits. 
+% \cite{collins2005trends} found "contrasting relationships between size (body mass) and depth in the scavenging and predatory demersal ichthyofauna". 
+% Predatory species saw a reduction in body mass with depth whereas the reverse trend was true for scavengers. 
+% This, the authors pointed out, is because randomly distributed carrion is better exploited by fish with larger body sizes owing to starvation resistance.
 
@@
 %% KH Does it need any home? 
-Certainly, scavenging should be particularly attractive to flying species compared to mammals. % Not sure about this paragraph 
-The latter can kill prey up to the same body mass as themselves and sometimes an order of magnitude heavier \citep[e.g. socially hunting lions; ][]{owen2008predator}.
-In contrast, birds of prey tend to kill prey smaller than themselves \citep{slagsvold2007prey} because of the greater cost of injury and the need to carry off their food.
+% Certainly, scavenging should be particularly attractive to flying species compared to mammals. % Not sure about this paragraph 
+% The latter can kill prey up to the same body mass as themselves and sometimes an order of magnitude heavier \citep[e.g. socially hunting lions; ][]{owen2008predator}.
+% In contrast, birds of prey tend to kill prey smaller than themselves \citep{slagsvold2007prey} because of the greater cost of injury and the need to carry off their food.
 % This is likely due to their need to kill animals that they can fly away with, as well as the risk of injury being higher (which carries a higher mortality risk) for a bird than a mammal.
-Scavenging provides a means for birds to exploit species that would otherwise be too big for them to kill.
+% Scavenging provides a means for birds to exploit species that would otherwise be too big for them to kill.
 % Thus, for birds, scavenging means they can exploit species that would otherwise be too big for them to kill.
 %%%%%%%%%%this is moved from the locomotion section as it didn't seem to fit.
-
-Because of the random nature of carrion we would expect adaptations that reduce energetic costs of maintenance to be selected for in scavengers as it would maximise the benefit derived from such a sporadic food source.
-Extant reptiles possess an advantage here, in that over the course of a year their food requirements can be 30 times lower than an endotherm of equal size \citep{Nagy1621}.
-\cite{devault2002scavenging} suggest this is an avenue for scavenging in snakes because they ``exhibit exceedingly low maintenance metabolisms, and most can survive on  a few scant feedings per year.
-It is, therefore, possible for snakes to rely largely on infrequent, less energy-rich meals."
-In the same review the authors found occurrences of scavenging spread across five families of snakes and stated that this behaviour is ``far more common than currently acknowledged."\citep{devault2002scavenging}.
-The same reasoning can be applied to crocodiles and their allies \citep{forrest2003evidence} because a sit and wait strategy is viable for an ectotherm.
 

### 2016-04-26 natural history of scavenging.tex
section edits

@@
 In the same review the authors found occurrences of scavenging spread across five families of snakes and stated that this behaviour is ``far more common than currently acknowledged."\citep{devault2002scavenging}.
-The same reasoning can be applied to crocodiles and their allies \citep{forrest2003evidence} because a sit and wait strategy is viable for an ectotherm \citep{forrest2003evidence,moleon2015carcass}. 
+The same reasoning can be applied to crocodiles and their allies because a sit and wait strategy is viable for an ectotherm \citep{forrest2003evidence,moleon2015carcass}. 
 \cite{carey1982temperature} found ectothermic sharks have the ability to exist weeks between meals because they focus on the energy-rich sections of carcasses (see Handling Time). 
@@
 \subsection*{Locomotion}
-Because of the inherent unpredictability of carrion, scavenging depends more on the ability to efficiently move over large areas than does predation.
+The inherent unpredictability of carrion also means scavenging depends more on the ability to efficiently move over large areas than does predation.
 This generally requires an efficient transfer of metabolic energy into movement which relies on the animal's anatomy and physiology as well as the medium of the environment in which the animal is moving (i.e.

### 2016-04-27 natural history of scavenging.tex
updated figure
Figures legends update

@@
 \centering
-   \includegraphics[width=1\textwidth]{Summary_figure/Summary_figure_Landscape.pdf}
-\caption{Factors influencing the proportion of scavenging in a vertebrates' diet. Blue lines indicate a reduction in the factor and red lines indicate an increase.}
-%TG: Probably needs some more explanations...
+   \includegraphics[width=1\textwidth]{Summary_figure/Summary_figure.pdf}
+\caption{Factors influencing the proportion of scavenging in a vertebrates' diet.
+The black and white gradient bar on the top represent the proportion of scavenging in the diet, ranging from low (- -) on the left hand side to high (+ +) on the right hand side.
+Blue lines indicate that a reduction in the factor will lead to more or less proportion of scavenging in the diet and red lines indicate an increase.}
+%Silhouettes from left to right on each line: \textit{Somelizard ofsomesort}, \textit{Felis silvestris}, \textit{Somelizard ofsomesort}, Diomedeidae, \textit{Antennarius striatus}, \textit{Gyps}, Actinopterigii skeleton, \textit{Stegosaurus} skeleton, \textit{Gyps}, \textit{Crocuta crocuta}, \textit{Rousettus aegyptiacus}, \textit{Crocuta crocuta}.
 \label{Summary_figure}
@@
    \includegraphics[width=1\textwidth]{timeline_figure/timeLine.pdf}
-\caption{The diversity of scavengers through time. Each species has either direct evidence for it being a scavenger or would be positioned high up on our scavenging scale.}
+\caption{The diversity of scavengers through time.
+Each species has either direct evidence for it being a scavenger or would be positioned high up on our scavenging scale.
+The three lines represent the three different environments (from top to bottom: aerial, terrestrial and aquatic).
+Time is represented on the horizontal axis with each third representing respectively the Paleozoic (541 to 252 Mya), the Mesozoic (252 to 66 Mya) and the Cenozoic (66 Mya to the present).}
+%Silhouettes from left to right in each different environment: Aerial: \textit{Some pterosaur}, Azhdarchid (which one?), Diomedeidae, \textit{Gyps}, \textit{Some preybird}; Terrestrial: \textit{Some mammaliaform?}, \textit{Some temnospondyl?}, \textit{Allosaurus}?, \textit{Dilophosaurus}?, \textit{Canis lupus?}, \textit{Crocuta crocuta}, \textit{Some shneakyshnake}, \textit{Homo sapiens}; Aquatic: \textit{Tiktaalik}, \textit{Some antiarchplacoderm}, Ichtyosauridae, \textit{Some metriorhynchidae?}, Mosasauridae, \textit{Basilosaurus}, \textit{Carcharodon carcharias}, \textit{Orcinus orca}.
 \label{Timeline}

### 2016-04-27 natural history of scavenging.tex

@@
 All foraging processes depend on the encounter rate between consumer and resource. 
-Locomotory speed, foraging time and detection radius all increase the encounter rate between a scavenger and the carcasses it is searching for. 
-Alternatively, a species can reduce its metabolic requirements so that it can survive long periods between meals. 
-We would thus expect selection pressures to act on the various traits that govern these parameters.
-
-However, as we noted above, encounter rate is also determined by the productivity of the environment, something which the selective forces acting on the scavenging have no control over.
+Locomotory speed, foraging time and detection radius all determine the encounter rate between a scavenger and the carcasses it is searching for and we would thus 
+expect selection pressures to act on the various traits that govern these parameters.
+Alternatively, selection might drive a species to reduce its metabolic requirements so that it can survive long periods between meals, or expand a species ability to 
+process tissues or decaying flesh that might ordinarily be discarded.  However, as we noted above, 
+encounter rate is also determined by the productivity of the environment, and ultimately the rate at which carcasses are produced, 
+an aspect which the selective forces acting on the scavenger have no influence.
 
@@
 \subsection*{Locomotion}
-The inherent unpredictability of carrion also means scavenging depends more on the ability to efficiently move over large areas than does predation.
+The inherent unpredictability of carrion also means scavenging depends more on the ability to efficiently move over larger areas than does predation.
 This generally requires an efficient transfer of metabolic energy into movement which relies on the animal's anatomy and physiology as well as the medium of the environment in which the animal is moving (i.e.

### 2016-04-28 natural history of scavenging.tex
Andrew's edits

--- 2016-04-26 natural history of scavenging.tex
@@
 \usepackage{xfrac}
+\usepackage{lineno}
 
  \renewcommand{\familydefault}{\sfdefault}
+
+\nolinenumbers
 
@@
 We use these components to highlight the morphologies and environments that would have been conducive to scavenging over geological time by focusing on the dominant vertebrate groups of the land, sea and air.
-The result is a document on the natural history of scavenging, the first to our knowledge. 
+The result is a synthesis on the natural history of scavenging, the first to our knowledge. 
 Our idea of a scale of scavenging can be applied to any species at any time to judge the importance of this behaviour in its diet. 
@@
 However, species that rely on scavenging to sustain substantial portions of their diets must encounter a sufficient amount of carrion in order to meet their energetic demands.
-Once found, the scavenger must be able to out-compete any potential competitors and process the, typically decaying, carcass replete with micro-organism derived toxins  \citep{ruxton2014fruit}. 
+Once found, the scavenger must be able to out-compete any potential competitors and efficiently process the, typically decaying, carcass replete with micro-organism derived toxins  \citep{ruxton2014fruit}.
 These characteristics can be assumed to be under evolutionary selection pressures for traits that increase carrion discovery and monopoly.
@@
 \centering
-   \includegraphics[width=1\textwidth]{Summary_figure/Summary_figure_Landscape.pdf}
+   \includegraphics[width=1\textwidth]{Summary_figure/Summary_figure.pdf}
 \caption{Factors influencing the proportion of scavenging in a vertebrates' diet. Blue lines indicate a reduction in the factor and red lines indicate an increase.}
@@
 All foraging processes depend on the encounter rate between consumer and resource. 
-Locomotory speed, foraging time and detection radius all increase the encounter rate between a scavenger and the carcasses it is searching for. 
+Locomotory speed, foraging time and detection radius all determine the encounter rate between a scavenger and the carcasses it is searching for. 
 Alternatively, a species can reduce its metabolic requirements so that it can survive long periods between meals. 
@@
 In the same review the authors found occurrences of scavenging spread across five families of snakes and stated that this behaviour is ``far more common than currently acknowledged."\citep{devault2002scavenging}.
-The same reasoning can be applied to crocodiles and their allies because a sit and wait strategy is viable for an ectotherm \citep{forrest2003evidence,moleon2015carcass}. 
-\cite{carey1982temperature} found ectothermic sharks have the ability to exist weeks between meals because they focus on the energy-rich sections of carcasses (see Handling Time). 
+The same reasoning can be applied to crocodiles and their allies \citep{forrest2003evidence,moleon2015carcass}. 
+\cite{carey1982temperature} found ectothermic sharks have the ability to go weeks between meals because they focus on the energy-rich sections of carcasses (see Handling Time). 
 Endotherms have also evolved physiological mechanisms that allow them to depress their otherwise high metabolic rates at certain times e.g.
@@
 climbing, tool-use, wing development etc.) it does not differ radically in cost from quadrupedal locomotion (\cite{williams1999evolution}, and references therein).
-For instance, \cite{alexander2004bipedal} shows  that, in the case of humans, we are more economical than predicted while walking and less so while running according to predicted costs of terrestrial movement.
+For instance, \cite{alexander2004bipedal} shows  that, in the case of humans, we are more economical than predicted while walking and less so while running according to predicted costs of terrestrial movement calculated by allometric scaling relationships.
 Our locomotory efficiency has fed into the question of where our ancestors placed on the hunter-scavenger axis during the Plio-Pleistocene, which has been a matter of debate for years \citep{dominguez2002hunting}.
@@
 While the locomotory ability of theropods has been debated since their first inception, more recent studies have reconstructed them as relatively mobile animals \citep{pontzer2009biomechanics}. 
-Despite some suggestions that larger species may have had some advantage in scavenging, partially due to the ability to search large areas \citep{ruxton2003could}, more recent studies have shown that the energetic demands of such large forms meant scavenging was likely more prevalent in mid-sized theropods of approximately half a tonne \citep{carbone2011intra,kane2016body}.
+Despite some suggestions that larger species may have had some advantage in scavenging, partially due to the ability to search large areas \citep{ruxton2003could}, more recent work has shown that the energetic demands of such large forms meant scavenging was likely more prevalent in mid-sized theropods of approximately half a tonne \citep{kane2016body}.
 
@@
 
-Species capable of flight have effectively added an extra spatial dimension (i.e. the vertical component) to their sensory environment over land animals.
+Species capable of flight have added an extra spatial dimension (i.e. the vertical component) to their sensory environment over land animals.
 This allows them to look down on a landscape where they are unencumbered by obstacles that would obstruct the view of a terrestrial scavenger.
-Such an ability has obvious benefits in detecting carrion.
+In this way they are effectively cheating the 2D system by gaining a bird's eye view which has has obvious benefits in detecting carrion.
 Certainly, vultures are known to have impressive visual acuity, with one estimate indicating lappet-faced vultures (\textit{Torgos tracheliotus}) are capable of detecting a 2 metre carcass over 10 km away \citep{spiegel2013factors}.
@@
 
-Existing in this '3D' environment also means being able to gather a wealth of information from other foragers, be they conspecifics or otherwise \citep{jackson2008effect}.
+Existing in this environment also means being able to gather a wealth of information from other foragers, be they conspecifics or otherwise \citep{jackson2008effect}.
+These interactions are properly three dimensional in the sense of \cite{pawar2012dimensionality} because both producer and scrounger are in the air \citep{dall2005information}. 
 Again, returning to vultures, the genus \textit{Gyps} consists of highly social and colonially nesting species \citep{fernandez2015density}.
@@
 \cite{beasley2015vertebrates} who point to a series of studies showing how microbes and invertebrates benefit at higher temperatures to the detriment of vertebrate scavengers such that ``above 20$^{\circ}$C vertebrates were able to detect and consume only 19\% of small-mammal carcasses, whereas at temperatures below 18$^{\circ}$C, vertebrates consumed 49\% of carcasses".
-This is a sobering thought given the impact we humans are having on the Earth's climate.
 
@@
 \section*{Handling Time}
-Since the food a scavenger depends upon is not dispatched directly, often the most easily accessible and choicest components of the carcass will be missing or, if present, will be subject to decay as well as competition.
+Since the food a scavenger depends upon is not dispatched directly, often the most easily accessible and choicest components of the carcass will be missing owing to the activity of predators and other scavengers, or, if present, will be subject to decay as well as competition.
 So being able to overcome competitors and maximise the nutrient gain from the remnants are all essential parts of carcass handling time.
@@
 \subsection*{Food Processing}
-Aside from coping with competitors, another vital component of carrion handling time is the ability to maximise the energy gain from the remains. 
+Aside from coping with competitors, another vital component of carrion handling time is the ability to maximise the energy gain from the remains while reducing the energetics of doing so. 
 At whale carcases, white and blue sharks are known to preferentially feed on the blubber layer \citep{long1996sharks}. 
@@
    \includegraphics[width=1\textwidth]{timeline_figure/timeLine.pdf}
-\caption{The diversity of scavengers through time. Each species has either direct evidence for it being a scavenger or would be positioned high up on our scavenging scale.}
+\caption{The diversity of scavengers through time across air, land and sea.
+Each species has either direct evidence for it being a scavenger or would be positioned high up on our scavenging scale.
+The ticks on the x axis represent the transitions for the Palaeozoic-Mesozoic (252.17 Mya) and Mesozoic-Cenozoic (66Mya) boundaries}
 \label{Timeline}
@@
 \section*{Acknowledgments}
-
-A lot of people are to thank here.
+Thanks to Natalie Cooper for highlighting the potential for this review, and to Ara Monadjem and Deirdre McClean for their comments on the manuscript.
 

### 2016-04-28 natural history of scavenging.tex
Fixed weird merge conflict

@@
    \includegraphics[width=1\textwidth]{Summary_figure/Summary_figure.pdf}
-\caption{Factors influencing the proportion of scavenging in a vertebrates' diet. Blue lines indicate a reduction in the factor and red lines indicate an increase.}
-%TG: Probably needs some more explanations...
+\caption{Factors influencing the proportion of scavenging in a vertebrates' diet.
+The black and white gradient bar on the top represent the proportion of scavenging in the diet, ranging from low (- -) on the left hand side to high (+ +) on the right hand side.
+Blue lines indicate that a reduction in the factor will lead to more or less proportion of scavenging in the diet and red lines indicate an increase.}
+%Silhouettes from left to right on each line: \textit{Somelizard ofsomesort}, \textit{Felis silvestris}, \textit{Somelizard ofsomesort}, Diomedeidae, \textit{Antennarius striatus}, \textit{Gyps}, Actinopterigii skeleton, \textit{Stegosaurus} skeleton, \textit{Gyps}, \textit{Crocuta crocuta}, \textit{Rousettus aegyptiacus}, \textit{Crocuta crocuta}.
 \label{Summary_figure}
@@
 All foraging processes depend on the encounter rate between consumer and resource. 
-Locomotory speed, foraging time and detection radius all determine the encounter rate between a scavenger and the carcasses it is searching for. 
-Alternatively, a species can reduce its metabolic requirements so that it can survive long periods between meals. 
-We would thus expect selection pressures to act on the various traits that govern these parameters.
-
-However, as we noted above, encounter rate is also determined by the productivity of the environment, something which the selective forces acting on the scavenging have no control over.
+Locomotory speed, foraging time and detection radius all determine the encounter rate between a scavenger and the carcasses it is searching for and we would thus 
+expect selection pressures to act on the various traits that govern these parameters.
+Alternatively, selection might drive a species to reduce its metabolic requirements so that it can survive long periods between meals, or expand a species ability to 
+process tissues or decaying flesh that might ordinarily be discarded.  However, as we noted above, 
+encounter rate is also determined by the productivity of the environment, and ultimately the rate at which carcasses are produced, 
+an aspect which the selective forces acting on the scavenger have no influence.
 
@@
    \includegraphics[width=1\textwidth]{timeline_figure/timeLine.pdf}
-\caption{The diversity of scavengers through time across air, land and sea.
+\caption{The diversity of scavengers through time.
 Each species has either direct evidence for it being a scavenger or would be positioned high up on our scavenging scale.
-The ticks on the x axis represent the transitions for the Palaeozoic-Mesozoic (252.17 Mya) and Mesozoic-Cenozoic (66Mya) boundaries}
+The three lines represent the three different environments (from top to bottom: aerial, terrestrial and aquatic).
+Time is represented on the horizontal axis with each third representing respectively the Paleozoic (541 to 252 Mya), the Mesozoic (252 to 66 Mya) and the Cenozoic (66 Mya to the present).}
+%Silhouettes from left to right in each different environment: Aerial: \textit{Some pterosaur}, Azhdarchid (which one?), Diomedeidae, \textit{Gyps}, \textit{Some preybird}; Terrestrial: \textit{Some mammaliaform?}, \textit{Some temnospondyl?}, \textit{Allosaurus}?, \textit{Dilophosaurus}?, \textit{Canis lupus?}, \textit{Crocuta crocuta}, \textit{Some shneakyshnake}, \textit{Homo sapiens}; Aquatic: \textit{Tiktaalik}, \textit{Some antiarchplacoderm}, Ichtyosauridae, \textit{Some metriorhynchidae?}, Mosasauridae, \textit{Basilosaurus}, \textit{Carcharodon carcharias}, \textit{Orcinus orca}.
 \label{Timeline}

### 2016-04-28 natural history of scavenging.tex
Minor edits

@@
 All foraging processes depend on the encounter rate between consumer and resource. 
-Locomotory speed, foraging time and detection radius all determine the encounter rate between a scavenger and the carcasses it is searching for and we would thus 
-expect selection pressures to act on the various traits that govern these parameters.
-Alternatively, selection might drive a species to reduce its metabolic requirements so that it can survive long periods between meals, or expand a species ability to 
-process tissues or decaying flesh that might ordinarily be discarded.  However, as we noted above, 
-encounter rate is also determined by the productivity of the environment, and ultimately the rate at which carcasses are produced, 
-an aspect which the selective forces acting on the scavenger have no influence.
+Locomotory speed, foraging time and detection radius all determine the encounter rate between a scavenger and the carcasses it is searching for and we would thus  expect selection pressures to act on the various traits that govern these parameters.
+Alternatively, selection might drive a species to reduce its metabolic requirements so that it can survive long periods between meals, or expand a species ability to  process tissues or decaying flesh that might ordinarily be discarded.
+However, as we noted above, encounter rate is also determined by the productivity of the environment, and ultimately the rate at which carcasses are produced,  an aspect which the selective forces acting on the scavenger have no influence.
 
@@
 The same reasoning can be applied to crocodiles and their allies \citep{forrest2003evidence,moleon2015carcass}. 
-\cite{carey1982temperature} found ectothermic sharks have the ability to go weeks between meals because they focus on the energy-rich sections of carcasses (see Handling Time). 
+\cite{carey1982temperature} found that sharks, as ectothermics, have the ability to go weeks between meals because they focus on the energy-rich sections of carcasses (see Handling Time \ref{handlingTime}).
 Endotherms have also evolved physiological mechanisms that allow them to depress their otherwise high metabolic rates at certain times e.g.
@@
 The groups from which these modern soaring birds arose, appear during the Palaeocene \citep[66 - 56 Million years ago (Mya); ][]{Jetz2012, Jarvis2014} and Cretaceous \citep{chiappe2006early} respectively.
-However, soaring flight is likely to be far older than this with avian flight originating in the Late Jurassic (163.5-145 Mya) and vertebrate flight in the Late Triassic (235-201.3 Mya) coincident with the pterosaurs.
+However, soaring flight is likely to be far older than this with avian flight originating in the Late Jurassic (163.5 - 145 Mya) and vertebrate flight in the Late Triassic (235 - 201.3 Mya) coincident with the pterosaurs.
 Indeed, scavenging among pterosaurs has been hypothesised many times before \citep{witton2008reappraisal}. 
@@
 In fact, facultative scavenging is seen in many selachian groups, including species of extant sharks like white sharks \citep[known to feed on whale carcasses;][]{fallows2013white}, Greenland sharks \citep[feeding on seals;][]{watanabe2012slowest}, and sixgill sharks \citep{anderson2016impact}.
-There is evidence too of scavenging in extinct species, where shark teeth have been found in the remains of dinosaurs, mosasaurs and Pliocene mysticete whales \citep[5.3-3.6 Mya; ][]{schwimmer1997scavenging,ehret2009caught}.
+There is evidence too of scavenging in extinct species, where shark teeth have been found in the remains of dinosaurs, mosasaurs and Pliocene mysticete whales \citep[5.3 - 3.6 Mya; ][]{schwimmer1997scavenging,ehret2009caught}.
 
-Interestingly, style of swimming in fish does not significantly affect the cost of movement \citep{williams1999evolution}, hence it is likely that many aquatic species with large ranges will encounter scavenging opportunities.
+Interestingly, however, style of swimming in fishes does not significantly affect the cost of movement \citep{williams1999evolution}, hence it is likely that many aquatic species with large ranges will encounter scavenging opportunities.
 It is perhaps jaw morphology that is more likely to define which species can avail of carcasses.
@@
 That said, aquatic endotherms have and do scavenge. 
-For instance, early whales such as \textit{Basilosaurus} (38-36.5 Mya) seem to have fit into the same niche as killer whales (\textit{Orcinus orca}) and we have some evidence for scavenging in both \citep{fahlke2012bite,Whitehead415}.
+For instance, early whales such as \textit{Basilosaurus} (38 - 36.5 Mya) seem to have fit into the same niche as killer whales (\textit{Orcinus orca}) and we have some evidence for scavenging in both \citep{fahlke2012bite,Whitehead415}.
 
@@
 Despite being intuitive, \cite{sullivan2015posture} states most of the hypotheses in favour of this idea remain to be tested in the context of archosaur evolution. 
-One noted consequence of a sprawling gait is the phenomenon known as Carrier's constraint such that the animal can't move and undergo costal ventilation at the same time because the lateral movements impedes its lungs \citep{carrier1987evolution}.
+One noted consequence of a sprawling gait is the phenomenon known as Carrier's constraint such that the animal can't move and undergo coastal ventilation at the same time because the lateral movements impedes its lungs \citep{carrier1987evolution}.
 The evolution of an upright posture has been offered as one of the primary mechanisms that allowed early archosaurs to overcome this constraint \citep{uriona2008recruitment}.
@@
 These insights allow us to compare extant terrestrial species to their prehistoric forebears given the dominance of mammalian carnviores since the Eocene (56-33.9 Mya) where the order split into the Caniforma and Feliforma \citep{van1987skeletal}.
-To take one example, \cite{anyonge1996locomotor} found that \textit{Nimravides}, a genus of sabretooth cat from the Miocence (10.3 to 5.3 Mya), were likely to have been ambush predators which would argue against them taking a lot of carrion.
+To take one example, \cite{anyonge1996locomotor} found that \textit{Nimravides}, a genus of sabretooth cat from the Miocence (10.3 - 5.3 Mya), were likely to have been ambush predators which would argue against them taking a lot of carrion.
 
@@
 Any animal that could find such whale carcasses is unlikely to have ignored other types of carrion.
-Although no aquatic species have ever exceeded the size of whales, some enormous animals have evolved in this environment before the evolution of cetaceans, including \textit{Leedsichthys}, a bony fish from the Middle Jurassic (174.1-163.5 Mya) and the aquatic Mesozoic reptiles, the plesiosaurs, pliosaurs and ichtyosaurs, that could all exceed 15 metres in length \citep{ruxton2011zoology,danise2014ecological}.
+Although no aquatic species have ever exceeded the size of whales, some enormous animals have evolved in this environment before the evolution of cetaceans, including \textit{Leedsichthys}, a bony fish from the Middle Jurassic (174.1 - 163.5 Mya) and the aquatic Mesozoic reptiles, the plesiosaurs, pliosaurs and ichtyosaurs, that could all exceed 15 metres in length \citep{ruxton2011zoology,danise2014ecological}.
 
@@
 \section*{Handling Time}
+\label{handlingTime}
 Since the food a scavenger depends upon is not dispatched directly, often the most easily accessible and choicest components of the carcass will be missing owing to the activity of predators and other scavengers, or, if present, will be subject to decay as well as competition.
@@
 Extant bats are also poorly equipped when it comes to feeding on carrion; the larger forms are typically frugivores and therefore lack the adaptations for digesting meat, while the smaller carnivorous bats are mainly found in the microbats, which are insectivorous \citep{aguirre2003implications}.
-That said, Necromantis (“death-eater”), a large bat from the middle to late Eocene (56-33.9 Mya) had a robust cranio-mandibular morphology, and is a likely candidate for an extinct scavenging bat \citep{Weithofer_Necromantis_1887,Hand_Necromantis_2012}.
+That said, Necromantis (“death-eater”), a large bat from the middle to late Eocene (56 - 33.9 Mya) had a robust cranio-mandibular morphology, and is a likely candidate for an extinct scavenging bat \citep{Weithofer_Necromantis_1887,Hand_Necromantis_2012}.
 
@@
 The three lines represent the three different environments (from top to bottom: aerial, terrestrial and aquatic).
-Time is represented on the horizontal axis with each third representing respectively the Paleozoic (541 to 252 Mya), the Mesozoic (252 to 66 Mya) and the Cenozoic (66 Mya to the present).}
+Time is represented on the horizontal axis with each third representing respectively the Paleozoic (541 - 252 Mya), the Mesozoic (252 - 66 Mya) and the Cenozoic (66 Mya to the present).}
 %Silhouettes from left to right in each different environment: Aerial: \textit{Some pterosaur}, Azhdarchid (which one?), Diomedeidae, \textit{Gyps}, \textit{Some preybird}; Terrestrial: \textit{Some mammaliaform?}, \textit{Some temnospondyl?}, \textit{Allosaurus}?, \textit{Dilophosaurus}?, \textit{Canis lupus?}, \textit{Crocuta crocuta}, \textit{Some shneakyshnake}, \textit{Homo sapiens}; Aquatic: \textit{Tiktaalik}, \textit{Some antiarchplacoderm}, Ichtyosauridae, \textit{Some metriorhynchidae?}, Mosasauridae, \textit{Basilosaurus}, \textit{Carcharodon carcharias}, \textit{Orcinus orca}.
@@
 \section*{Conclusion}
+%TG: not sure where to put that in the conclusion but it'll be cool to mention that our scavenger scale can not only be used for the past but also for species hard to observe in the present (or even - if going wild - for predicted behavioral changes in the future!)
 As is often the case in science, the present provides the key to the past.

### 2016-07-12 response_letter.tex
start of response

\documentclass{letter}
%\signature{Your name}
%\address{Street \\ City \\ Country}
\begin{document}
\begin{letter}{Company name \\ Street\\ City\\ Country}
\opening{To the editor}

We would like to thank the editor, subject editor and reviewers for their comments. By bringing attention to elements of the manuscript that require more careful communication we believe that you have helped us to significantly improve our manuscript. In this submission we have clarified these areas and added a body of work to bolster the conclusions drawn from our work. Please see below where we consider each of your points in turn.

Reviewer: 3

Comments to the Author
This is a well-written, enjoyable manuscript that assesses the functional traits associated to past and, mostly, current major vertebrate scavengers. As a result, the authors provide a sort of guide to identify the extent of scavenging behaviour among carnivores in both current and past ecosystems. I have several suggestions that should be considered before publication (please see below).

Major comments:

1. Invertebrates are ignored completely throughout the ms. Please specify in the Introduction that this review focuses on vertebrate scavengers. In this line, I recommend to include “in vertebrates” at the end of the title to be more specific. Also, in the “Competition” section you mention competition with vertebrates and micro-organisms, but not with invertebrates. You should consider them in this section (and in “Facilitation”; see next point).

2. I found difficult to follow the main point of the paper because the ms is not well organized. Following a hierarchical order, I’d define four/five main sections: “Introduction”, “The Challenges of Scavenging” (this could be a subsection within the previous section), “Encounter Rate” (with four different subsections: “Metabolism”, “Locomotion”, “Sensory Detection” and “Carcass Availability”), “Handling Time” (which would include “Food Processing”) and “Conclusion”. \\

Moreover, “Competition” and “Facilitation” (I strongly encourage you to consider not only competitive, but also facilitative processes, which are mostly neglected in your review) should be treated as transversal factors that can modulate each of the abovementioned parameters. Thus, I’d include them in a different section, or mention the competitive and facilitative processes related to each parameter within each subsection. Fig. 1 should be re-organized accordingly.

3. The use of the scientific literature on scavenging is deficient, especially for a review paper like this. A number of key references are missing, e.g., Selva \& Fortuna (2007 Proc. R Soc. B 274:1101-1108), Wilson \& Wolkovich (2011 TREE 23:129-135), Beasley et al. (2012 Oikos 121:1021-1026), Cortés-Avizanda et al. (2012 Ecology 93:2570-2579), Cortés-Avizanda et al. (2014 Ecology 95:1799-1808), Moleón et al. (2014 Biol. Rev. 89:1042-1054), Pereira et al. (2014 Mammal Rev. 44:44-55), Périquet et al. (2015 Biol. Rev. 90:1197-1214), Mateo-Tomás et al. (2015 Divers. Distrib. 21:913-924), Moleón \& Sánchez-Zapata (2015 BioScience 65:1003-1010), Moreno-Opo et al. (2016 Behav. Ecol.), Sebastián-González et al. (2016 Ecology 97:95-105). Your literature synthesis might benefit from the reading of these articles, for instance regarding competitive and facilitative interactions among scavengers. Also, the review would be more informative and useful for readers. Next I mention some particular comments regarding the use of the references (I mention others in Minor comments): \\

Page 2, 1st sentence: but see Moleón \& Sánchez-Zapata (2015); many of the most charismatic vertebrates of the world are scavengers. Rather than “scavengers”, what is not charismatic is probably “scavenging”.

We now refer to the behaviour rather than the species as lacking charisma.

 Page 2, 13th line: Moleón \& Sánchez-Zapata (2015) may be more appropriate than Koenig (2006) because the former is more general and not restricted to vultures.
 Page 2, lines 17th-18th: see Pereira et al. (2014) and Périquet et al. (2015) for reviews.

These references have been added.

 Page 3, bottom paragraph, bottom line: this is not true, as scavenging may indirectly affect herbivore populations and thus carrion availability (see Moleón et al. 2014 Biol. Rev.).

The potential for scavengers to affect carrion availability has been acknowledged.

 Page 5, bottom paragraph, 1st sentence: Mateo-Tomás et al. (2015) is more general than Kendall (2013).
 Page 8, 1st and 2nd sentences: see Pereira et al. (2014) and Périquet et al. (2015).
 Page 8, 3rd sentence: you may also cite DeVault et al. (2003) and Pereira et al. (2014).

These references have been added.

 Page 8, 8th-17th lines: please reword; wild dogs hardly scavenge, while leopards and lions are frequent scavengers (see Pereira et al. 2014).

We have reworded this section to make the more general point that ambush predators can rely more on hunting than can cursorial species, a point explicitly stated in the Pereira et al. (2014) article, which we now cite.

 Page 11, 1st sentence: see Kane et al. (2014) and Moleón et al. (2014 Biol. Rev.) for inter-specific interactions.
 Page 11, 4th sentence: Kane et al. (2014) is not appropriate here because it is about inter-specific interactions; use Cortés-Avizanda et al. (2014) instead.
 Page 14, 2nd paragraph, 1st sentence: see also Moleón et al. (2014 Biol. Rev.) and Pereira et al. (2014).
 Page 14, 2nd paragraph, 2nd sentence: see better Pereira et al. (2014) and Périquet et al. (2015). 
 Page 14, bottom paragraph, 3rd sentence: you may also consider Moreno-Opo et al. (2016).
 Page 15, 2nd paragraph, 2nd sentence: see also Moleón et al. (2014 Biol. Rev.) and Pereira et al. (2014).

All of these references have been added.

4. You should include more discussion about the importance of carcass size throughout the ms, as it has a strong influence on scavenging patterns and interactions among vertebrates and between vertebrates and smaller carrion consumers (e.g., see Moleón et al. 2015).

Minor comments:

GENERAL

1. Please provide line numbers to facilitate the reviewing process.

Line numbers have now been included.

ABSTRACT

2. You could remove “, the first to our knowledge”.

This has been removed.

3. Also, you could mention some relevant specific findings of your synthesis.

THE CHALLENGES OF SCAVENGING

4. Page 3, 1st sentence: “often difficult to predict” compared to what? Please specify. There are many examples in which carrion is highly predictable, e.g., during salmon spawning (see also Pereira et al. 2014).

ENCOUNTER RATE

5. Page 3, bottom paragraph, 3rd line: change “Alternatively” to “Also”?

This has been changed.

METABOLISM

6. Page 4, 1st sentence: “the sporadic nature of carrion”; please explain.

LOCOMOTION

7. Page 4, bottom paragraph, 1st sentence: change “inherent” to “relative”.

This has been changed.

8. Page 4, bottom paragraph, 3rd sentence: remove “, paradoxically,”.

This has been deleted.

9. Page 5, 2nd paragraph, 1st sentence: change “appear” to “appeared”. Also, specify Mya for the Cretaceous period.

This has been changed.

10. Page 9, top paragraph: please be more cautious in your assertion, as cooperation skills and tools of early hominins probably allowed them to defend and usurp carcasses from more powerful scavengers.

We have changed this paragraph and take into account the authors' reference to tool use and sociality in hominins. The point is that even with these advantages, running to a carcass over long distances still wasn't energetically sensible.

DETECTION

11. Page 11, top and 2nd paragraph, 1st sentence: this is about facilitation. As said above, facilitative processes may affect other parameters as well, not only the ability to detect carrion.

12. Page 11, last sentence: explain why wind speed affects the olfactory capacity of an aquatic species.

PREY AVAILABILITY

13. I’d change this title to “Carcass Availability”, which is more specific.

This has been changed.

14. Among the different components of carcass availability, e.g., abundance, duration, predictability and accessibility, in this section you only talk about abundance and duration. You should also discuss about carcass predictability (e.g., see Cortés-Avizanda et al. 2012) and carcass accessibility (e.g., see DeVault \& Krochmal 2002).

15. Page 12, 2nd paragraph, 1st sentence: as said before, this is not true, given that scavenging may indirectly affect herbivore populations and thus carrion availability (see Moleón et al. 2014 Biol. Rev.).

We have removed the part of the sentence where we mistakenly suggest that scavengers cannot affect carrion availability.

16. Page 13, 3rd and 4th sentences: whaling has exerted also a huge negative impact on scavengers by depleting whale populations. If you are going to talk about human impacts, maybe is better to do it at the end of the ms, perhaps in a new section in which you explore which scavenging traits could be more successful in a human dominated world.

This has been removed.

17. You should also take into account facilitative processes. For instance, some scavengers are able to access the interior of carcasses only once other, more powerful species, has opened the skin (see Moleón et al. 2014 Biol. Rev.).

18. Page 14, bottom paragraph, 2nd sentence: I’d change it to “This is true for many vultures and other major avian scavengers such as albatrosses who all have […]”.

We have reworded the sentence as suggested.

FOOD PROCESSING

19. Page 16, 2nd paragraph, 6th sentence: insert “Within mammals,” before “This ability”.

This has been added.

20. Page 17, 2nd paragraph, 4th sentence: do you refer to species or individuals? Higher species diversity doesn’t mean more individuals.

The reference of Pobiner (2015) states that there were more individuals and we have changed the sentence to reflect this.

21. Page 18, 1st sentence: change “enabling” to “enable”.

This has been changed.

22. Page 18, 3rd sentence: “thus”? I don’t understand this sentence; could you please reword it?

We have reworded this sentence. The idea is that early birds didn't have beaks and so could more easily eat meat. This is in contrast to modern birds where a beak may be a hindrance.

23. Page 18, 5th sentence: “The mix of strong and weak features in their skull morphology”; please explain it.

CONCLUSION

24. Page 18, bottom paragraph, 3rd sentence: change “technique” to “approach”?

This has been changed.

FIGURES

25. Fig. 2: you may insert the name and Mya range of each period in the X axis rather than in the figure legend. Also, I’d indicate “aerial”, “terrestrial” and “aquatic” in the Y axis rather than in the legend.

Hope this helps,
Marcos Moleón

\bigskip
Yours Sincerely,

Adam Kane and co-authors 
%\closing{Yours Faithfully,}
%\ps{P.S. Here goes your ps.}
%\encl{Enclosures.}
\end{letter}
\end{document}

### 2016-07-12 natural history of scavenging.tex
start of response

@@
 
-\title{A Recipe for Scavenging - the natural history of a behaviour}
-%TG: nice title! I'd replace the hyphen with a comma though
+\title{A Recipe for Scavenging in Vertebrates - the natural history of a behaviour}
+
 \running{Scavenging in vertebrates}
@@
 We use these components to highlight the morphologies and environments that would have been conducive to scavenging over geological time by focusing on the dominant vertebrate groups of the land, sea and air.
-The result is a synthesis on the natural history of scavenging, the first to our knowledge. 
-Our idea of a scale of scavenging can be applied to any species at any time to judge the importance of this behaviour in its diet. 
+The result is a synthesis on the natural history of scavenging. 
+Our idea of a scale of scavenging can be applied to any species at any time to help judge the importance of this behaviour in its diet. 
 \end{abstract}
@@
 
-\section*{Introduction}
+\section{Introduction}
 
-Historically, scavengers have not been viewed as the most charismatic of animals.
-This may go some way to explaining the gap in our knowledge of the prevalence of this behaviour \citep{devault2003scavenging}.
-Professor Sanborn Tenney writing in 1877 for The American Naturalist had this to say about one well known group, ``prominent among the mammalian scavengers are the hyenas, the ugliest in their general appearance of all the flesh eaters."
-He contrasts these with ``nobler kinds" of carnivores such as lions and tigers \citep{tenney1877few}.
-Even aside from our own subjective biases, scavenging is a difficult behaviour to detect after the fact.
-Without catching a carnivore in the act of killing we are left to infer how the prey was killed.
-Some simple heuristics can inform us, for instance, in cases where the prey item was simply too large to have been killed by the ostensible predator \citep{pobiner2008paleoecological}.
+Historically, food webs have classified most acts of carnivory as predation events \citep{wilson2011scavenging}. 
+This is clearly a mistaken view because of the existence of scavenging, a behaviour displayed by almost all carnivorous vertebrates \citep{devault2003scavenging}. 
+Recent research has begun to redress this imbalance by recognising the prevalence of scavenging and detailing its ecological consequences across modern ecosystems where we are beginning to appreciate the implications for the trophic ecology of these systems \citep{pereira2014facultative,periquet2015lion}.
+Indeed, there is no discrete divide between predators and scavengers but rather a continuous gradient \citep{pereira2014facultative}.
+Unfortunately, the lack of direct behavioural data makes it difficult to discern scavenging from predation among extinct forms where it was surely exhibited to a similar extent among prehistoric vertebrate carnivores.
+Some simple heuristics can inform us whether or not the prey was scavenged, for instance, in cases where it was simply too large to have been killed by the ostensible predator \citep{pobiner2008paleoecological}.
 But clearly, a scavenger does not only feed on animals too big for it to have hunted.
-The obvious lack of direct behavioural data compounds the difficulty of discerning scavenging from predation among extinct forms.
-Indeed, a single species of dinosaur notwithstanding \citep{carbone2011intra}, a synthesis describing the natural history of scavengers is absent from the literature.
-With research on scavenging on the rise \citep{manga2006vulture} we are now beginning to realise the extent of this behaviour such that, ``in some ecosystems, vertebrates have been documented to assimilate as much as 90\% of the available carrion" \citep{beasley2015vertebrates}.
-This has profound implications for the trophic ecology of these systems and particularly our models of them.
-Even Tenney’s noble big cats are now known to take in a significant portion of carrion in their diet where some lion populations acquire over 50\% of their meat from carcasses \citep{jones2015african}.
-While recognising the difficulty in directly observing scavenging, it is possible to turn to other methods in order to discern the most suitable morphologies, physiologies and environments for a scavenging lifestyle to prosper.
-Here we chart the natural history of scavenging by assessing the potential for the behaviour in dominant vertebrate groups given their ecology and functional traits.
-%It is our aim in this review to employ these methods to gain an understanding of scavengers past and present.
+There are other methods however, that allow us to discern the most suitable morphologies, physiologies and environments for a scavenging lifestyle to prosper e.g.
+energetics models, comparative anatomy, palaeontology etc.
+\citep{shipman1986scavenging,ruxton2003could}.
+Yet, a synthesis describing the natural history of scavengers is absent from the literature.
+Here we rectify this shortcoming in charting the natural history of scavenging by assessing its potential among extinct vertebrate groups given their ecology, functional traits and the enviornmental context.
 
-%the intro needs to lead towards a aim a bit more strongly
+\subsection{The Challenges of Scavenging}
 
-\subsection*{The Challenges of Scavenging} 
-The chief hurdle to scavenging is finding a resource that is often difficult to predict in space and time.
-Through chance alone many species will avail of some opportunistic scavenging. 
-However, species that rely on scavenging to sustain substantial portions of their diets must encounter a sufficient amount of carrion in order to meet their energetic demands.
+Species that rely on scavenging to sustain substantial portions of their diets must encounter a sufficient amount of carrion in order to meet their energetic demands.
 Once found, the scavenger must be able to out-compete any potential competitors and efficiently process the, typically decaying, carcass replete with micro-organism derived toxins \citep{ruxton2014fruit}.
 These characteristics can be assumed to be under evolutionary selection pressures for traits that increase carrion discovery and monopoly.
-Finally, the potential for scavenging will also depend on the density, size, and quality of carcases produced, all of which are affected by complex ecosystem dynamics but are outside the selection pressures on the scavenger.
+Finally, the potential for scavenging will also depend on the density, size, and quality of carcases produced, all of which are affected by complex ecosystem dynamics \citep{moleon2014inter}. 
+Even these aspects can be indirectly impacted by the behaviour of the scavenger \citep{moleon2014inter}.
 Each of these facets are essentially the backbone of fundamental ecological theory and are the key parameters defined in functional response curves, namely encounter rate, handling time and prey availability \citep{jeschke2002predator}.
 By considering scavenging in this context of optimal foraging we can identify the prerequisite attributes and processes required for the behaviour. 
-This has enabled us to propose a scale of scavenging whereupon we can place any vertebrate species, past or present, and assess the likely importance of carrion in its diet.
+This has enabled us to propose a scale of scavenging whereupon we can place an extinct vertebrate carnivore and assess the likely importance of carrion in its diet.
 
@@
 Blue lines indicate that a reduction in the factor will lead to more or less proportion of scavenging in the diet and red lines indicate an increase.}
-%Silhouettes from left to right on each line: \textit{Somelizard ofsomesort}, \textit{Felis silvestris}, \textit{Somelizard ofsomesort}, Diomedeidae, \textit{Antennarius striatus}, \textit{Gyps}, Actinopterigii skeleton, \textit{Stegosaurus} skeleton, \textit{Gyps}, \textit{Crocuta crocuta}, \textit{Rousettus aegyptiacus}, \textit{Crocuta crocuta}.
 \label{Summary_figure}
@@
 
-%KH: need paragraph for how do we know something is (was) a scavenger
-\section*{Encounter Rate}
+\section{Encounter Rate}
 All foraging processes depend on the encounter rate between consumer and resource. 
 Locomotory speed, foraging time and detection radius all determine the encounter rate between a scavenger and the carcasses it is searching for and we would thus expect selection pressures to act on the various traits that govern these parameters.
-Alternatively, selection might drive a species to reduce its metabolic requirements so that it can survive long periods between meals, or expand a species ability to  process tissues or decaying flesh that might ordinarily be discarded.
-However, as we noted above, encounter rate is also determined by the productivity of the environment, and ultimately the rate at which carcasses are produced,  an aspect which the selective forces acting on the scavenger have no influence.
+Selection might also drive a species to reduce its metabolic requirements so that it can survive long periods between meals, or expand a species ability to process tissues or decaying flesh that might ordinarily be discarded.
+However, as we noted above, encounter rate is also determined by the productivity of the environment, and ultimately the rate at which carcasses are produced. 
+Scavengers can drive this process indirectly, for instance, by consuming the majority of carrion, obligate scavengers deny a resource to their facultative counterparts obliging the latter to hunt more and thus produce more carrion \citep{moleon2014inter}.
 
-%As such, scavenging potential is strongly affected by search rates, which are determined by anatomy, physiology and behaviour which are all under selective pressure, as well as the type of environment the scavenger is foraging in.
-
-\subsection*{Metabolism}
+\subsection{Metabolism}
 Because of the sporadic nature of carrion we would expect adaptations that reduce energetic costs of maintenance to be selected for in scavengers as it would allow for longer inter-feeding periods. 
@@
 
-\subsection*{Locomotion}
-The inherent unpredictability of carrion also means scavenging depends more on the ability to efficiently move over larger areas than does predation.
+\subsection{Locomotion}
+The relative unpredictability of carrion also means scavenging depends more on the ability to efficiently move over larger areas than does predation.
 This generally requires an efficient transfer of metabolic energy into movement which relies on the animal's anatomy and physiology as well as the medium of the environment in which the animal is moving (i.e.
 aerial, aquatic or terrestrial).
-Perhaps the most efficient form of locomotion in vertebrates is, paradoxically, found in flying species. 
+Perhaps the most efficient form of locomotion in vertebrates is found in flying species. 
 Despite the energetic costs of flight, the only known vertebrate obligate scavengers are the old and the new world vultures. 
@@
 
-The groups from which these modern soaring birds arose, appear during the Palaeocene \citep[66 - 56 Million years ago (Mya); ][]{Jetz2012, Jarvis2014} and Cretaceous \citep{chiappe2006early} respectively.
+The groups from which these modern soaring birds arose, appeared during the Palaeocene \citep[66 - 56 Million years ago (Mya); ][]{Jetz2012, Jarvis2014} and Cretaceous \citep[145.5 - 65.5 Mya; ][]{chiappe2006early} respectively.
 However, soaring flight is likely to be far older than this with avian flight originating in the Late Jurassic (163.5 - 145 Mya) and vertebrate flight in the Late Triassic (235 - 201.3 Mya) coincident with the pterosaurs.
@@
 It seems probable that at least some of these extinct species used soaring as a means for scavenging \citep{witton2013pterosaurs}.
-%KH added "at least some of these" as it was not clear what species used soaring as a means for scavenging.
+
 While soaring is perhaps the only viable means of locomotion that allows for an obligate, scavenging life-style \citep{ruxton2004obligate}, powered flight is still an efficient means of locomotion. 
@@
 
-We know that many extant birds exist as facultative scavengers because storks, raptors and corvids all take substantial quantities of carrion in their diet \citep{kendall2013alternative}. 
+We know that many extant birds exist as facultative scavengers because storks, raptors and corvids all take substantial quantities of carrion in their diet \citep{mateo2015regional}. 
 Similarly we would expect that extinct species would also scavenge in a similar fashion depending on the efficiency of their flight. 
@@
 It is perhaps jaw morphology that is more likely to define which species can avail of carcasses.
-%KH I moved this down here as it kinda throws our shark soaring thing out the window (which I like as a linking function).
-However I think this kinda solves our problems in the complexity of swimming by saying apart from large ranges ability to eat is probably more important.
 
 We might expect then that by combining an aquatic environment and an endothermic metabolism marine mammals would especially prosper as scavengers. 
-%, especially as this group has perhaps the largest home range of any animals (whales and shit).
 Fossil pinnipeds and cetaceans from 60 Mya have transitional features indicative of their evolutionary trajectory to fully aquatic species \citep{williams1999evolution}.  
@@
 
-%%KH Is there anything we should add about other extinct groups such as plesiosaurs and https://en.wikipedia.org/wiki/Plesiosauria#Gait maybe an air breathing non-airbreathing devide would be good.
-
-%% AK:  Need to be careful here because this whole idea of sprawling posture being worse off than erect posture is not that well supported in the literature. 
-%% Great Transformations in Vertebrate Evolution By Kenneth P. Dial, Neil Shubin, Elizabeth L. Brainerd talks about it at length 
-%% So I think this should stay pitched as an endo/ ectothermy thing rather than posture 
 Terrestrial environments are the most energetically costly in which to move \citep{tucker1975energetic}. 
-Unlike aerial and aquatic environments support must be provided through the animal's posture. % leading to trade offs between low cost of maintain and the efficiency of transport. 
+Unlike aerial and aquatic environments support must be provided through the animal's posture. 
 The early transition from a sprawling gait, seen in early tetrapods, to the more erect posture of synapsids and later dinosaurs and mammals, has often been supposed as conferring a huge advantage to the latter groups \citep{sullivan2015posture}.
@@
 
-% For example, the sprawling posture of many ectotherm groups (except frogs!) would be expected to allow for a lower maintenance cost relating to body support in comparison to groups with the limbs position underneath the body, such as mammals, aves and dinosaurs.
-%KH: I agree with TG and have removed the gas exchange thing as its a physiology thing only applicable to certain groups and not got to do with terrestrial environments really. I also removed the 
-%Also that reference is from 1975, is this up to date?
-
-% While the low maintenance costs of a sprawling-like posture may reduce the need to scavenge frequently, the ability to search large areas quickly is severely restricted in these groups reducing scavenging opportunities (ref on home range of terrestrial ectoterms).
-% For example, a sprawling gait results in 
-% On the other hand the more erect body positions and often associated higher basal metabolic rates, allow for sustained locomotion over long periods of time allowing such species to search large areas.
-
-%KH Even after changing it I dont like this transition type narrative.
-%KH also needs to be more clear on what the advantage to scavenging is for these groups.
-%
-%none of these advantages refer very clearly to scavenging
-%KH I'm always wary of deeming something as intuitive,
-
-%Metabolic rate however, unquestionably impacts terrestrial species whereby ectotherms such as many modern reptiles, cannot move for sustained periods \citep{bennett1979endothermy}. 
-% Species which do not support an upright position such as modern reptiles limit their ability to move for sustained periods \citep{bennett1979endothermy}.
-
-%KH I think as ectothermy is almost completely associated with sprawling gaits I would tie it up together to avoid repeating something similar here.
-
-%This is exacerbated by their sprawling gait which results in the phenomenon known as Carrier's constraint such that the animal can't move and breathe at the same time because the lateral movements impedes its lungs \citep{carrier1987evolution}.
-% These constraints are likely to mean that many ectotermic terrestrial species, which cannot maintain an upright posture would depend on a relatively low level of scavenging (I'm not sure this is true though).
-%This would also have been true of extinct species with the same physiology. 
-%TG: I guess this is a comment too: Snakes are a bit different are more effecent at moving I think - lots of scavanging. 
-%Both these groups are more likely to depend on low energetic costs in order to compensate for low encounter rates due to their inability to move.
-
-% Although the earliest evidence of vertebrate scavenging comes from the Permian (298.9 - 252.17 Mya) where a temnospondyl amphibian fed on the carcass of \textit{Varanops}, a predatory synapsid of the time \citep{reisz2006articulated}.
 Whatever the case, it is with the evolution of endothermy in the therapsid-mammal lineage \citep{clarke2010temperature} that terrestrial vertebrates would have gained the ability to range more widely, a vital component in seeking out carrion.
 Modern endothermic mammals can sustain longer periods of energetically expensive activity \citep{bennett1979endothermy} resulting in larger foraging ranges. 
-Today, terrestrial scavenging in the mammals is probably best known in an African context where hyenas, jackals and lions all take sizeable proportions of carrion in their diet.
+Today, terrestrial scavenging in the mammals is probably best known in the species rich African context where hyenas, jackals and lions all take sizeable proportions of carrion in their diet \citep{pereira2014facultative,periquet2015lion}.
 In the spotted hyena (\textit{Crocuta crocuta}), striped hyena (\textit{Hyaena hyaena}) and brown hyena (\textit{Hyaena brunnea}) it can be over 90\% \citep{jones2015african}.
-And although no contemporary terrestrial vertebrate exists as an obligate scavenger, most, if not all, are facultative to some extent \citep{beasley2015vertebrates}.
-The particular reliance of hyenas on carrion means we can use them as examples of efficient terrestrial scavengers to compare with other forms. 
+And although no contemporary terrestrial vertebrate exists as an obligate scavenger, most, if not all, are facultative to some extent \citep{devault2003scavenging,beasley2015vertebrates,pereira2014facultative}.
+The particular reliance of hyenas on carrion means we can use them as examples of efficient terrestrial scavengers to compare with other forms \citep{periquet2015lion}. 
 In terms of locomotion, they employ a characteristic ``rocking horse gait"  which allows them to cover great distances efficiently, loping at 10 km/hr \citep{mills1989comparative,jones2015african}. 
-Such long-distance travel is apparent in African wild dogs (\textit{Lycaon pictus}) and many other canids \citep{pennycuick1995radius,janis2014forelimb}. 
-In contrast, big cats like leopards (\textit{Panthera pardus}) rely on ambush \citep{pennycuick1995radius}. 
-This allows us to make a broad distinction between the ambush strategies of cat-like forms and the pursuit/ pounce strategies of more dog-like forms, the latter being more suited to scavenging \citep{janis2014forelimb}.
+Such long-distance travel is apparent many other cursorial predators \citep{pennycuick1995radius,janis2014forelimb,pereira2014facultative}. 
+In contrast, big cats rely on ambush \citep{pennycuick1995radius,pereira2014facultative}. 
+This difference means ambush predators tend to be more obligate hunters \citep{pereira2014facultative} 
 These insights allow us to compare extant terrestrial species to their prehistoric forebears given the dominance of mammalian carnviores since the Eocene (56-33.9 Mya) where the order split into the Caniforma and Feliforma \citep{van1987skeletal}.
@@
 Our locomotory efficiency has fed into the question of where our ancestors placed on the hunter-scavenger axis during the Plio-Pleistocene, which has been a matter of debate for years \citep{dominguez2002hunting}.
-\cite{ruxton2013endurance} added to this debate with their argument that long distance endurance running, often used by humans for hunting, was not an important feature of hominin scavenging.
-They arrive at this conclusion on the basis of the high cost of running coupled with the fact that dangerous competitors will tend to arrive beforehand.
-
-% While we can use modern analogues of quadrupedal species, the lack of scavenging in bipeds makes this more difficult.
-However, bipedal locomotion may have been a particularly useful strategy for searching large areas.
-For instance, \cite{alexander2004bipedal} shows that, in the case of humans, we are more economical than predicted while walking according to predicted costs of terrestrial movement.
-In fact the ability of humans to engage in hunting strategies that involve traversing long distances, such as persistence hunting, has led to several researches highlighting the possible role of scavenging in early hominids.
-Indeed the low energetics of human walking at least allows for a substantial amount of scavenging in our hominid ancestors (Graemes papers on scavenging).
-%%KH I know this is excluding other hominids so feel free to write a more inclusive line.
+\cite{ruxton2013endurance} added to this debate with their argument that long distance endurance running was not an important feature of hominin scavenging but was instead used by humans for hunting.
+Even when considering the use of weapons and social foraging, the authors argue that the high costs of running long distances would render it an unattractive foraging strategy for scavenging.
 
@@
 
-%They are quite alien to anything we know today which restricts our ability to understand their ecology far more so than extinct mammals \citep{weishampel2004dinosauria}.
-%Of relevance, are the questions that still persist about their metabolism, with the latest evidence suggesting they were mesothermic \citep[i.e.
-intermediate to ecto- and endotherms;][]{grady2014evidence}.
-%KH Do we need to get into this?
-%We do know that they walked with the erect gait of mammals or birds rather than the sprawling gait of lizards and that they were most likely facultative scavengers \citep{weishampel2004dinosauria,depalma2013physical}.
-%KH Is this not a bit trivial almost?
-%Taken together, this implies dinosaurs had a foraging range that fell in between that of modern terrestrial mammals and reptiles. 
-%I don't think we can say that, we have no idea on thier range.
-For example using body size they should be larger than modern mammals, however the importance of metabolic rate is still not clear with home range (body size explains most.).
-This might be useful http://www.jstor.org/stable/10.1086/682070?seq=1#page_scan_tab_contents
-
-%KH I think the terrestrial section needs a little more work still, it doesn't hang together quite right yet although it is not far off it.
-While I have wrote it up this way I think ending with humans would be good as it gives a returning to the present type feel which wraps the whole section up without being too explicit about it.
-Maybe using large bird locomotion would allow this.
-
-\subsection*{Detection}
-
-%KH Just my opinion but I found it really difficult to understand the flow of this section.
-I starts off with a type of 2d- 3d thing but then kinda bounces around different ideas with everything talked about in relation to aerial viewing.
-Its also missing how understanding detection ranges can help us understand past scavenging behavior.
-I would include in the terrestrial section, the use of looking at the portion of the brain attributed to olfactory sensing (like the way we did the scaling of detection range in our paper) as a possible indicator of scavenging.
-I would also throw into the social communication section that these behaviors can show how groups can be good scavengers through active/passive communication and we might expect that in other extinct social groups (the vulture thing was almost impossible for pterodactyls to avoid I would say and probably common in many soaring flying species).
-Also Graeme has a paper on human scavenging about how hominies may have exploited vultures and also probably searched in a social sense.
-%I might be too grumpy at the end of the day, but I think a bit of a restructure might help this section.
+\subsection{Sensory Detection}
 
@@
 
-Existing in this environment also means being able to gather a wealth of information from other foragers, be they conspecifics or otherwise \citep{jackson2008effect}.
+Existing in this environment also means being able to gather a wealth of information from other foragers, be they conspecifics or otherwise \citep{jackson2008effect,KaneVul,moleon2014inter}.
 These interactions are properly three dimensional in the sense of \cite{pawar2012dimensionality} because both producer and scrounger are in the air \citep{dall2005information}. 
 Again, returning to vultures, the genus \textit{Gyps} consists of highly social and colonially nesting species \citep{fernandez2015density}.
-These behaviours allow them to forage far more efficiently because one bird can scrounge information on the location of food from another successful forager \citep{KaneVul}.
+These behaviours allow them to forage far more efficiently because one bird can scrounge information on the location of food from another successful forager \citep{cortes2014bird}.
 Information transfer of this kind is typically inadvertent and as a consequence no complex social interactions are required, simply the ability to recognise a successful forager.
@@
 
-\subsection*{Prey Availability}
-The environmental influence on prey availability is an aspect that greatly affects encounter rate but is invisible to the selective forces acting on the scavenger. 
-Aspects including, primary productivity, relief, and temperature will all greatly affect scavenging tendency. 
+\subsection{Carcass Availability}
+The environmental influence on carcass availability is an aspect that greatly affects encounter rate. %but is invisible to the selective forces acting on the scavenger. 
+Aspects including, primary productivity, relief, and temperature will all affect scavenging tendency. 
 \cite{ruxton2004obligate} suggest an historic ecosystem with a productivity similar to the Serengeti could have supported an \textit{obligate} mammalian or reptilian terrestrial scavenger.
@@
 
-Staying in the aquatic setting, the phenomenon of occasional bounties of carrion in the form of whale falls has led some researchers to investigate if an obligate scavenger could survive by seeking out these remains exclusively.
+The physical differences between water and air mean carcass availability is radically different between these environments \citep{beasley2012carrion}. 
+For one, carcasses get moved around by water which results in a more diffuse signal being produced for would-be scavengers.
+Carcasses also tend to sink in water where they are no longer accessible to pelagic scavengers \citep{beasley2012carrion}. 
+Research has shown that an animal need only travel between 16 to 36 km to encounter a fresh whale carcass \citep{moleon2015living}. 
+The phenomenon of occasional bounties of carrion in the form of these whale falls has led some researchers to investigate if an obligate scavenger could survive by seeking out these remains exclusively.
 \cite{ruxton2005searching} argued that although this is energetically feasible it's ecologically unlikely.
@@
 Although no aquatic species have ever exceeded the size of whales, some enormous animals have evolved in this environment before the evolution of cetaceans, including \textit{Leedsichthys}, a bony fish from the Middle Jurassic (174.1 - 163.5 Mya) and the aquatic Mesozoic reptiles, the plesiosaurs, pliosaurs and ichtyosaurs, that could all exceed 15 metres in length \citep{ruxton2011zoology,danise2014ecological}.
-
 So, despite being unlikely, the energetic feasibility of a marine scavenger that specialises on large carcasses has a long history.
-One point of interest is that of the whaling industry, which provided a bonanza of floating carcasses especially during the 20th century \citep{Whitehead415}.
-This meant killer whales could switch from hunting to scavenging, a switch made that much easier by the noise of the whaling vessels that would effectively ring the ``dinner-bells" \citep{Whitehead415}.
 
@@
 
-%Predators were large-bodied too compared to extant mammalian predators \citep{mcnab2009resources}, and so, especially if they were ectothermic, could last longer between meals, rendering scavenging a more attractive behaviour relative to predation.
-%%KH Not sure this sentence belongs here, already covered int he handling section.
-
-%Frequently, the interplay between abiotic and biotic factors can impact the ability of an animal to scavenge. 
-%We know vultures and eagles tend to soar using thermals and if these air pockets don't form, say on a cloudy day, the bird is grounded \citep{mundy1992vultures}.
-%In many habitats (e.g. the Arctic) it is simply not possible for sufficiently powerful thermals to form and as a consequence large-bodied vultures cannot exist.
-%% KH I dont think this belong in this section, we probably have covered this enough in the locomotion part.
-%One result of this is that t.
-%Certainly, a major difficulty for terrestrial scavengers is competition with vultures.
-%Nocturnal behaviour in the hyaenidae in general has been put forth as an adaptation to reduce competition with these exclusively diurnal birds \citep{gittleman2013carnivore}.
-%%KH I think the focus needs to be more on the actual environment and opportunities available.
-I would suggest that competition either stays in the handling section (my preference) or is moved down to here (which also works really).
-%%KH You could then split the structure into production in terrestrial/ marine which focuses on primary productivity and species composition and how they affect what available then the next part being competition in both section.
-
-\section*{Handling Time}
+\section{Handling Time}
 \label{handlingTime}
@@
 
-\subsection*{Competition}
-Large body size has substantial advantages in agonistic interactions \citep{ruxton2004obligate}. 
-For instance, lions can acquire much of their carrion through kleptoparasitism of hyena kills \citep{trinkel2005competitive}. 
+\subsection{Competition}
+Large body size has substantial advantages in agonistic interactions \citep{ruxton2004obligate,moleon2014inter,pereira2014facultative}. 
+For instance, lions can acquire much of their carrion through kleptoparasitism of hyena kills \citep{trinkel2005competitive,pereira2014facultative,periquet2015lion}. 
 This line of reasoning suggests that some theropod dinosaurs, who could get up to 15 tonnes, would have had easily monopolised a carcass \citep{weishampel2004dinosauria} provided they could find them efficiently \citep{kane2016body}.
@@
 We would expect this trait to be selected for even in the case of weight-constrained, scavenging fliers.
-This is true for wandering albatrosses (\textit{Diomedea exulans}), cinereous vultures (\textit{Aegypius monachus}) and condors (\textit{Vultur gryphus}, \textit{Gymnogyps californianus}) who all have body masses that can exceed 10 kg and represent some of the heaviest bird species capable of flight \citep{weimerskirch1992reproductive,ferguson2001raptors,donazar2002effects}.
-Indeed, such is the competitive advantage held by vultures over other facultative scavenging birds that temporal niche partitioning at the carcass has evolved \citep{kendall2013alternative,KaneVul}. 
+This is true for many vultures and other major avian scavengers such as albatrosses who all have who can have body masses in excess of 10 kg and represent some of the heaviest bird species capable of flight \citep{weimerskirch1992reproductive,ferguson2001raptors,donazar2002effects}.
+Indeed, such is the competitive advantage held by vultures over other facultative scavenging birds that temporal niche partitioning at the carcass has evolved \citep{kendall2013alternative,KaneVul,moreno2016behavioral}.
 Additionally, many pterosaurs were far bigger again, with estimated body masses of over 200 kg in the Azhdarchids \citep{witton2010size}.
@@
 Direct confrontation can be circumvented by certain behavioural adaptations. 
-The evolution of nocturnal behaviour in some mammals, for instance, has been put forth as an adaptation to reduce competition with the exclusively diurnal vultures \citep{gittleman2013carnivore}. 
+The evolution of nocturnal behaviour in some mammals, for instance, has been put forth as an adaptation to reduce competition with the exclusively diurnal vultures as well as with other larger predators \citep{gittleman2013carnivore,moleon2014inter,pereira2014facultative}.
 In areas absent of vultures such as the Arctic, terrestrial carnivores like bears and wolves take more carrion \citep{devault2003scavenging}
@@
 Moreover, some researchers have charted an evolutionary course from basal fossorial snakes to modern terrestrial species by way of an obligate scavenger intermediate \citep{bauchot2006snakes}. 
+In water, by contrast, competition with micro-organims is significantly reduced because of the high pressures and low temperatures \citep{beasley2012carrion}.
 
-\subsection*{Food Processing}
+\subsection{Food Processing}
 Aside from coping with competitors, another vital component of carrion handling time is the ability to maximise the energy gain from the remains while reducing the energetics of doing so. 
@@
 Osteophagy is known across a range of terrestrial carnivores and given that some fat-rich mammalian bones have an energy density (6.7 kJ/g) comparable with that of muscle tissue, it makes skeletal remains an enticing resource \citep{brown1989study}.
-This ability reached its zenith among hyenas with the evolution of the estimated 110 kg \textit{Pachycrocuta brevirostris} during the Pliocene \citep[3.6 - 2.58 Mya; ][]{palmqvist2011giant}.
-Indeed, their extinction has been blamed on the decline of sabretooth cats (Machairodontinae), the unique skull morphology of the latter meant they would leave a large amount of food on a carcass for would-be scavengers \citep{palmqvist2011giant}.
+Within mammals, this ability reached its zenith among hyenas with the evolution of the estimated 110 kg \textit{Pachycrocuta brevirostris} during the Pliocene \citep[3.6 - 2.58 Mya; ][]{palmqvist2011giant}.
+Indeed, their extinction has been blamed on the decline of sabretooth cats (Machairodontinae), because the unique skull morphology of the latter meant they would leave a large amount of food on a carcass for would-be scavengers \citep{palmqvist2011giant} (Note, however that the inability of these cats to deal with bone may be overstated, \citep{binder2010comparison}).
 Earlier in the evolution of mammals, the bone-crushing dogs that evolved during the Oligocene (Borophaginae; 33.9 - 23.03 Mya) have also been compared to hyenas in terms of their feeding ecology \citep{van2003chapter,martin2016pursuit}.
@@
 A recent study investigating potential scavenging opportunities for hominins in Kenya found that, in addition to skeletal material, there is a substantial amount of scavengeable meat left on predated remains; sufficient to sustain the requirements of an adult male \textit{Homo erectus} \citep{pobiner2015new}.
-In some historical hominin-inhabited areas there were a greater number of felids than hyenids.
+In some historical hominin-inhabited areas there were higher populations of felids compared to hyenids.
 Again, this is significant because hyenas are likely to have left far less flesh on a carcass than a felid such as a sabretooth, enabling contemporaneous hominins to benefit \citep{pobiner2015new}.
@@
 \cite{houston1974role} observed a group of \textit{Gyps} vultures consuming all of the soft tissue from a 50 kg Grant’s gazelle (\textit{Nanger granti}) in eight minutes. 
-Their serrated tongues and hooked bills enabling them to achieve this feat \citep{houston1975digestive}. 
+Their serrated tongues and hooked bills enable them to achieve this feat \citep{houston1975digestive}. 
 Aside from raptors, the specialised beaks of many modern bird lineages tends to hinder their ability to eat meat which is in contrast to the first lineages that did not have this feature \citep{martyniuk2012field}.
-As \cite{martyniuk2012field} notes these early birds would thus have been predominantly carnivorous, implying scavenging was a live opportunity cf. their descendants. 
+As \cite{martyniuk2012field} notes, this skull morphology of early birds indicates they were predominantly carnivorous, implying scavenging was a live opportunity cf. their descendants. 
 Among the pterosaurs, \cite{witton2013pterosaurs} makes the case that the istiodactyl pterosaurs were the most likely scavengers of this group based on their potential handling time. 
-The mix of strong and weak features in their skull morphology is indicative of animals that were suited to removing large amounts of flesh from an immobile foodstuff \citep{witton2013pterosaurs}.
+Their skull morphologies are indicative of animals that were suited to removing large amounts of flesh from an immobile foodstuff \citep{witton2013pterosaurs}.
 
 Again, we can draw a comparison with clades that are lacking in these features. 
-As we mentioned earlier the skull morphology of extinct sabretooths meant they would have left a large amount of flesh on a carcass and were unable to feed on bones \citep{palmqvist2011giant}. 
-Extant bats are also poorly equipped when it comes to feeding on carrion; the larger forms are typically frugivores and therefore lack the adaptations for digesting meat, while the smaller carnivorous bats are mainly found in the microbats, which are insectivorous \citep{aguirre2003implications}.
+Extant bats are poorly equipped when it comes to feeding on carrion; the larger forms are typically frugivores and therefore lack the adaptations for digesting meat, while the smaller carnivorous bats are mainly found in the microbats, which are insectivorous \citep{aguirre2003implications}.
 That said, Necromantis (“death-eater”), a large bat from the middle to late Eocene (56 - 33.9 Mya) had a robust cranio-mandibular morphology, and is a likely candidate for an extinct scavenging bat \citep{Weithofer_Necromantis_1887,Hand_Necromantis_2012}.
-
-% Some clades appear to lack many, if not all, of the phenotypic adaptations associated with an efficient carcass handling time. 
-% The extant bats seem to lack all of the features we have identified as important.
-
-%KH still not sure about this transition.
-
-% Some work on extinct sabretooths suggests they may have left a large amount of food for would-be scavengers because of their unique skull morphology.
-% As a result, the decline of Machairodontinae sabretooths has been offered as an explanation for the extinction of \textit{P brevirostris} \citep{palmqvist2011giant}.
-% And many of the aforesaid adapations for scavenging are found in these other major terrestrial mammalian carnviores.
-% Though the specific mix of features realised in hyenas suggest this is the model organism for terrestrial scavenging among mammals in the past.
-% These traits have been used to infer scavenging lifestyles in extinct mammals.
-
-%KH something on the robust morphology and size required to generate such forces however essentially restricts flying species from having such a strategy, unless they can use tools (dosent some bird fly up and then drop bones.)
-%KH I would also mention that however, our paper shows that such species that would have been able to have such bites were probably too large for this to be an effective stratagy.
-Then you can segue into the body size thing.
-%%An alternative structure would be to go straight into morphologies that crack bones, i.e large heads, then follow such morphology restricts some groups for example flying species or smaller animals like humans.
-Then you could move into tool use and similar type behavior in humans and birds.
-
- % The support of water allows for many aquatic species to reach large sizes thus granting its benefits. 
-% \cite{collins2005trends} found "contrasting relationships between size (body mass) and depth in the scavenging and predatory demersal ichthyofauna". 
-% Predatory species saw a reduction in body mass with depth whereas the reverse trend was true for scavengers. 
-% This, the authors pointed out, is because randomly distributed carrion is better exploited by fish with larger body sizes owing to starvation resistance.
-
-%KH Just as a though, people bring this up a lot that body size allows for long period of starvation being specific for scavenging but I never seen anything past a verbal argument.
-Is there anything that really supports it that say locomotion cant explain (i.e large search area thing)?
-
-%%%%%%%%%%  AK This annoying paragraph's new home TG: well done finding him a home!
-%% KH Does it need any home? 
-% Certainly, scavenging should be particularly attractive to flying species compared to mammals. % Not sure about this paragraph 
-% The latter can kill prey up to the same body mass as themselves and sometimes an order of magnitude heavier \citep[e.g. socially hunting lions; ][]{owen2008predator}.
-% In contrast, birds of prey tend to kill prey smaller than themselves \citep{slagsvold2007prey} because of the greater cost of injury and the need to carry off their food.
-% This is likely due to their need to kill animals that they can fly away with, as well as the risk of injury being higher (which carries a higher mortality risk) for a bird than a mammal.
-% Scavenging provides a means for birds to exploit species that would otherwise be too big for them to kill.
-% Thus, for birds, scavenging means they can exploit species that would otherwise be too big for them to kill.
-%%%%%%%%%%this is moved from the locomotion section as it didn't seem to fit.
-
-% Durophagy amount aquatic species, turtles, sharks, trigger fish, crocs etc.
 
@@
 Time is represented on the horizontal axis with each third representing respectively the Paleozoic (541 - 252 Mya), the Mesozoic (252 - 66 Mya) and the Cenozoic (66 Mya to the present).}
-%Silhouettes from left to right in each different environment: Aerial: \textit{Some pterosaur}, Azhdarchid (which one?), Diomedeidae, \textit{Gyps}, \textit{Some preybird}; Terrestrial: \textit{Some mammaliaform?}, \textit{Some temnospondyl?}, \textit{Allosaurus}?, \textit{Dilophosaurus}?, \textit{Canis lupus?}, \textit{Crocuta crocuta}, \textit{Some shneakyshnake}, \textit{Homo sapiens}; Aquatic: \textit{Tiktaalik}, \textit{Some antiarchplacoderm}, Ichtyosauridae, \textit{Some metriorhynchidae?}, Mosasauridae, \textit{Basilosaurus}, \textit{Carcharodon carcharias}, \textit{Orcinus orca}.
 \label{Timeline}
@@
 \section*{Conclusion}
-%TG: not sure where to put that in the conclusion but it'll be cool to mention that our scavenger scale can not only be used for the past but also for species hard to observe in the present (or even - if going wild - for predicted behavioral changes in the future!)
 As is often the case in science, the present provides the key to the past.
 The animals of today, while often different (sometimes radically so) to their ancestors, can be used to make informed comparisons to extinct species. 
-We have used this technique to give insight into the drivers of scavenging across vertebrates through time.
+We have used this approach to give insight into the drivers of scavenging across vertebrates through time.
 In common with any other forager be they grazer, browser or predator, scavengers past and present have had to balance their energetic costs with the gains of food. 
@@
 We hope this approach will be useful in the effort to explore this most understudied of feeding ecologies.
-% Add bits on climate change/dectection rate change, etc...
-
-% KH Maybe we need to more strongly reaffirm that the framework that we give really is useful and should be applied to all sorts of things. etc
 
@@
 \end{document}
-
-%TG: what do you mean by their position? Phylogenetic? Position on the "scavenging scale"?
-
-% This point illustrates how the environment can impact search efficiency depending on the sensory system that's used.
-
-% \section*{Terrestrial Scavengers}
-
-%\cite{ruxton2004obligate} offer a reason for this in that the traits that allow for vultures to exist as scavengers undermined their ability to hunt but that the same forces have not prevented mammals from doing so.
-
-%\subsection*{Environment}
-
-%TG: To be modified and inserted somewhere:
-
-%TG: orginal sentence "The intertidal environment provides a ready food source of stranded marine animals on a twice-daily basis, in the immediate vicinity of the sea, and would thus have allowed marine ancestors of tetrapods gradually to acquire terrestrial competence while accessing a new and essentially untouched resource."
-
-% and sensory perception \citep{farlow1994speculations}.
-
-%\section*{Aquatic Scavengers} 
-%An aquatic environment presents challenges for direct observational studies and so, similar to the approaches involving extinct species, much work has approached the question of scavenging propensity from an energetics perspective.
-%But it is certainly known to occur in many aquatic vertebrates.
-
-%A point to note is that vertebrates are relatively rarer in aquatic environments, because even large animals can get support from the buoyancy of the water without needing a backbone.
-
-%\subsection*{Locomotion}
-
-%\subsection*{Detection}
-%The existence of an obligate scavenger in a marine setting is uncertain \citep{britton1994marine,smith2003ecology,ruxton2004energetic,ruxton2005searching}.
-%Depending on the species, a carcass in this environment either floats or descends to the sea floor \citep{Whitehead415}.
-%In the latter low-light environment, visual detection distances are far lower (< 100 m) than they would be in the air.
-%As such, animals detect resources through chemo- and mechanoreception more so than through vision \citep{ruxton2004energetic}.
-\cite{beasley2015vertebrates} do note that ``some benthic scavengers (e.g., hagfish: family Myxinidae) rely on necrophagy for a large portion of their diet and may indeed be obligate scavengers".
-
-%\subsection*{Processing}
-
-%\subsection*{Environment}
-
-%Primary productivity is lower in almost all aquatic systems than terrestrial systems (except deserts) so as we go up the food chain the density of carcasses worth scavenging is going to be lower.
-
-% Maybe drop this bit
-% \section*{Ecological Role}
-% It is recognised that scavengers keep energy flows at a higher trophic level in food webs than decomposers because they consume relatively more carrion \citep{devault2003scavenging}.
-%They are also hugely important for the dispersal of nutrients \citep{beasley2015vertebrates}.
-%Consider the diversity of animals that can end up feeding at the carcass of an elephant.
-%Here we have an incredibly dense and nutrient rich patch that ends up being distributed widely.
-%In the absence of vertebrate scavengers, invertebrates and microorganisms would consume the carcass in-situ or at least distribute the constituent nutrients over a much shorter range.
-%This effect has been magnified as vertebrates evolved certain key traits that allowed them to range farther, namely an upright gait, an endothermic metabolism and of course, flight.
-%To quantify this effect with a simple example we can turn to some allometric relationships relating sustainable travelling speed to body mass.
-%In the case of mammals and reptiles these are 1.15 * body mass (kg) \textsuperscript{0.12} and 0.23 * body mass (kg) \textsuperscript{0.12}
-% respectively \citep{ruxton2004obligate}.
-%We can insert these into a foraging radius model ((duration * speed)/2)/1000 for a 12 hour foraging day which shows that while a 10 kg reptile can range 6.5 km an equally sized mammal can range nearly 33 km \citep{Enstipp2006Energetics}.
-%Thus, in an ecological context, the evolution of these steps coupled with the ability to scavenge resulted in a world with a far more widely distributed nutrient landscape.
-
-% %TG: LaTeX format template for the equations in the R file
-% \begin{equation}
-%   \text{Travelling speed} = \frac{duration \times speed}{2} \times 10^3
-% \end{equation}
-% where \textit{C} is a constant of $1.15$ for mammals and $0.23$ for squamates, %TG: if you meant squamates from reptiles
-% and speed being:
-% \begin{equation}
-%   speed = C \times M^{0.12}
-% \end{equation}
-
-%Scavenging is a widespread behaviour among vertebrates where most if not all carnivores act as facultative scavengers.
-%TG: or something more along the lines as:
-%Most if not all carnivorous vertebrates are facultative scavenging behaviour to some extant.
-%It is recognised that scavengers have an important role in keeping energy flows at a higher trophic level in food webs than decomposers because they consume relatively more carrion \citep{devault2003scavenging}.
-%Scavengers also provide useful ecosystem services by acting as barriers to the spread of disease by quickly consuming rotting carcasses which have often died from illness \citep{ogada2012dropping}.
-%(Since we are intrested in scavanging in th paleo record ecosystem services might not be that relavent, although modulators of disease is still relavent)
-
-%The limitations in studying extant scavenging behaviour is much larger in extinct species and systems with the obvious lack of observational data available.
-%TG: or more like this? 
-% The limitations in studying scavenging in extant species are even bigger in extinct species and past systems since the obvious lack of available direct behavioural data (CITE) %TG: bet you there's some paper for that
-% This means indirect observations in the fossil record and other approaches such as energetics must be used to infer these behaviours. %TG: I'll squiz your Am Nat paper here!
-% One avenue to infer scavenging from palaeontological data can be achieved by determining if a prey item was simply too big for the carnivore to have tackled in cases where tooth marks are found \citep{pobiner2008paleoecological}.
-% Comparative analysis can also allow us to ascertain which morphologies and physiologies are likely to have been found in scavenging species in the past \citep{ruxton2004obligate}.
-% The development of indirect measures of scavenging in palaeontology can in turn be applied to current scavenging systems that also suffer from a lack of observational data.
-
-%§4
-% In this review we collate methods (could this be another way of structuring it, just an idea) and research form palaeontology relating to scavenging behaviour and show that ignore this literature would be a missed opportunity for understanding extant scavenging.
-%TG: totally agree.
-I think we need to clearly know what this review is about: scavenging through time (like a story line of scavenging past and present - a bit boring if I can speak my mind) or how to study scavenging (through time or any other aspect - probably more interesting to more people I guess).
-%TG: additionally I find the divisions land/air/sea and Ceno/Meso/Paleozoic are a bit scholar.
-It might be better to just discuss the different techniques for investigating scavenging and include land/air/sea and Ceno/Meso/Paleo in there no?
-For example:
-%\section{Method 1: direct observation}
-%This can be totally doable in extant system for scavengers everywhere.
-%It is advantageous because it's direct and reliable observations.
-%But it has some limitations such as sending Adam to sit in the sun for ours and is hard to apply in deep sea environements or in any past ecosystems...
-
-%remember the journal is an ecological one at heart so I would always keep in mind about making it useful for ecologists over paelo people.
-In particular from the website "ECOGRAPHY publishes papers focused on broad spatial and temporal patterns, particularly studies of population and community ecology, macroecology, biogeography, and ecological conservation.
-Studies in ecological genetics and historical ecology are welcomed in the context of explaining contemporary ecological patterns".
-
-%"However, in general, vertebrate scavenging represents the widest dispersal of nutrients and energy from carcasses as vertebrate movement scales away to the broader landscape" \citep{benbow2015introduction}
-
- %"Increased kill rates by top predators may represent a little acknowledged marginal cost to the vertebrate community, directly attributable to scavenging activity.
-Given the impor -tance of top-down effects in many ecosystems, even a minor alteration to predation rates as driven by vertebrate scavengers may cause a significant flux in community composition."\citep{benbow2015introduction}
-
- %"Cortés-Avizanda et al.
-(2009) found that the abundance of prey species (i.e., hares—Lepus spp.
-and squirrels—Sciurus spp.
-in this case) decreased in sectors containing a carcass based on evidence from tracks in snow.
-An interesting hypothesis emerged, in which the scavengers that are recruited to a carcass may have temporarily played the dual role of increasing predator abundance near each carcass"\citep{benbow2015introduction}
-
-%"Historically, the prevalence of scavenging activities has been greatly underestimated.
-However, upon recognition that (1) in most ecosystems, a large number of animals die from causes other than predation and thus become available to scavengers; (2) most carcasses are scavenged by vertebrates before they are completely decomposed by arthropods and bacteria; and (3) almost all carnivorous animals are facultative scavengers, the importance of scavenging in food webs seems unsurprising" \citep{benbow2015introduction}
-
-%"This dispersion of carrion biomass by vertebrates is especially evident when carrion is initially concentrated spatially.
-For example, carcasses produced from fishing by-catch (Furness 2003), salmon (Salmonidae) die-offs (Hewson 1995), forest fires (Blanchard and Knight 1990), and single large carcasses (e.g., whales—Cetacea; Smith and Baco 2003) are often visited by multiple scavengers that range widely and therefore transport the nutrients from those carcasses over large distances."\citep{benbow2015introduction}
-
-%"Cross-habitat nutrient transport can produce a variety of important outcomes in recipient systems (e.g., Polis et al.
-2004), and scavengers can play a significant role in moving ``ecologi -cal subsidies" between habitats.
-For example, the use of ocean-derived carrion by terrestrial mammals (Rose and Polis 1998) and birds (Schlacher et al.
-2013) is extensive and may strongly influence dynamics of coastal food webs."\citep{benbow2015introduction}
-
-% "Markandya et al.
-(2008) estimated that the total costs to human health (including rabies cases from dog bites) that resulted from severe vulture declines totaled over $34 billion from 1993 to 2006.
-Also, Ogada et al.
-(2012) determined that the exclusion of vultures from large animal carcasses in Kenya resulted in a tripling of carcass decomposition times."\citep{benbow2015introduction}
-
-%"Of all the mammalian carnivores in Africa, brown, striped, and spotted hyenas (hereafter referred to simply as hyenas) derive the largest portion of their diet from scavenging.
-This is not surprising, as they show many unique adaptations specifically for scavenging.
-They have a unique body posture and a ``rocking-horse" gait (Eloff 1964; Tilson and Henschel 1986; Hofer and East 1993a; Frank 1996) that allows them to cover long distances in search of carrion and prey."\citep{benbow2015introduction}
-
-%\textit{Allosaurus} tooth marks on a hadrosaur in the Late Jurassic. 
-% Late Triassic scavenging on a prosauropod by basal carnivorous archosaurs \citep{hungerbuhler1998taphonomy}.
-%A likely instance of scavenging between a 4-million-year-old white shark (\textit{Carcharodon}) and mysticete whale from Peru \citep{ehret2009caught}.
-%Bite marks on early Holocene Tursiops truncatus fossils from the North Sea indicate scavenging by rays (Chondrichthyes, Rajidae) \citep{van2009bite}. 
-%Possible scavenging on a juvenile fur seal from the Late Neogene \citep{boessenecker2011mammalian}.

### 2016-07-13 response_letter.tex
added all reviewer comments to response

@@
 Please see below where we consider each of your points in turn.
+
+Reviewer: 1
+
+Comments to the Author
+This paper attempts to synthesize literature on traits that enable scavenging behavior in extant and extinct vertebrates.
+The paper is mostly well written, and most (but not all) pertinent literature sources are included.
+I do worry about the lack of novelty of the manuscript, and what I perceived as haphazard organization.
+Here are some specific issues that should be addressed in my opinion:
+
+1.      Abstract:  Scavenging can be difficult to observe, but probably no more so than predation.  Also, using experimental carcasses, scavenging is actually quite easy to observe.
+
+2.      Abstract:  You should point out that your ‘scale of scavenging’ is only qualitative; you made no attempt to quantify such traits.
+
+3.      Introduction:  Much of this section seems to follow DeVault’s (2003) review paper on scavenging—in fact, there is almost nothing new here that is not covered in that earlier review.
+
+4.      Introduction:  You mention that you will use “other methods in order to discern the most suitable morphologies…”  but you don’t say what they are.  What other methods are you talking about?
+
+5.
+I think that early on in the paper you need to define obligate and facultative scavengers and then throughout the paper, specify which you are referring to.
+For example, in the first sentence of the Detection section, you talk about ‘known scavengers’, but it is unclear if you really mean obligate scavengers.
+Nearly all carnivorous vertebrates are scavengers at least to some extent, so you need to elaborate on what you mean when you talk about ‘scavengers’.
+
+6.      Last sentence of Introduction:  Again, you propose a ‘scale of scavenging’, but without quantification, it is unclear how useful this ‘scale’ is.
+
+7.      Page 6, 12:  In your discussion of aquatic scavengers, you might want to refer to Beasley et al. (2012; Oikos 121:1021-1026).
+
+8.      Page 6:  Why is jaw morphology likely to define scavengers?  Please elaborate.
+
+9.
+Page 9:  It is true that carrion can comprise a large percentage of a hyena’s diet, but your text seems misleading.
+Most studies have shown that most of a hyena’s diet is from predation.
+For example, see Cooper et al.
+(1999; Afr.
+J.
+Ecol.
+37:149–160) and Gasaway et al.
+(1991; Afr.
+J.
+Ecol.
+29: 64–75).
+Reviewer: 1
+
+Comments to the Author
+This paper attempts to synthesize literature on traits that enable scavenging behavior in extant and extinct vertebrates.
+The paper is mostly well written, and most (but not all) pertinent literature sources are included.
+I do worry about the lack of novelty of the manuscript, and what I perceived as haphazard organization.
+Here are some specific issues that should be addressed in my opinion:
+
+1.      Abstract:  Scavenging can be difficult to observe, but probably no more so than predation.  Also, using experimental carcasses, scavenging is actually quite easy to observe.
+2.      Abstract:  You should point out that your ‘scale of scavenging’ is only qualitative; you made no attempt to quantify such traits.
+3.      Introduction:  Much of this section seems to follow DeVault’s (2003) review paper on scavenging—in fact, there is almost nothing new here that is not covered in that earlier review.
+4.      Introduction:  You mention that you will use “other methods in order to discern the most suitable morphologies…”  but you don’t say what they are.  What other methods are you talking about?
+5.
+I think that early on in the paper you need to define obligate and facultative scavengers and then throughout the paper, specify which you are referring to.
+For example, in the first sentence of the Detection section, you talk about ‘known scavengers’, but it is unclear if you really mean obligate scavengers.
+Nearly all carnivorous vertebrates are scavengers at least to some extent, so you need to elaborate on what you mean when you talk about ‘scavengers’.
+6.      Last sentence of Introduction:  Again, you propose a ‘scale of scavenging’, but without quantification, it is unclear how useful this ‘scale’ is.
+7.      Page 6, 12:  In your discussion of aquatic scavengers, you might want to refer to Beasley et al. (2012; Oikos 121:1021-1026).
+8.      Page 6:  Why is jaw morphology likely to define scavengers?  Please elaborate.
+9.
+Page 9:  It is true that carrion can comprise a large percentage of a hyena’s diet, but your text seems misleading.
+Most studies have shown that most of a hyena’s diet is from predation.
+For example, see Cooper et al.
+(1999; Afr.
+J.
+Ecol.
+37:149–160) and Gasaway et al.
+(1991; Afr.
+J.
+Ecol.
+29: 64–75).
+10.     Your section on Handling Time is so short that I think it should be cut or combined with another section.
+
+Reviewer: 2
+
+Comments to the Author
+This article is a review of the literature what the authors term a "natural history" of scavenging.
+They have approached the topic from the perspective of ecological variables (and other variables that condition them) in optimal foraging theory including encounter rate and handling time, but also metabolism, locomotion (e.g soaring and swimming), detection, prey availability, competition, and food processing.
+They describe many useful examples of the behavior and ecology of obligate and facultative scavengers.
+But I think they often go to far in their assumptions in trying to weave these examples together in a synthetic way.
+Some examples are inferring scrounging behaviors in flying pterosaurs (pace 11), assuming that hominins used vultures to identify scavenging opportunities (page 11), surmising that scavenging evolved with the earliest tetrapods based on "potential carrion unexploited by marine vertebrates" (page 12).
+There is a lot of "these researchers have suggested X" and then taking that as a given fact.
+
+While the authors distinguish between obligate and facultative scavenging, they tend to lump them together when it comes to constructing their arguments.
+In the end, is the argument about the former, or the latter?
+Is the real interest in whether species scavenge at all, what proportion of their resources are obtained by scavenging, or what the conditions are that produce facultative or obligate scavengers?
+They also seem to ignore flexibility in species' behaviors.
+For instance, while spotted hyenas are on their highly scavenging side of their Figures 1 and 2, in some ecosystems they hunt significant proportion of their prey.
+Similarly, lions can take a large proportion of their prey as scavenged, as the authors note.
+But to suggest, as they do in their abstract, that they can apply their scale of scavenging to any species at any time to judge how important scavenging was in its diet is frankly completely ignoring behavioral variability over time and space.
+It is substantially more difficult with fossil species for which we may know far fewer of the variables they list in Figure 1, in a nuanced way.
+Figure 2 does not display the diversity of scavengers through time, it displays examples of species that display observed or inferred scavenging behaviors.
+Diversity has specific ecological definition(s) that are being ignored here.
+
+If the authors chose to resubmit, I would suggest they organize their manuscript based on known information about modern species, known (e.g.
+bone processing by theropods) and inferred (e.g.
+scavenging in hominins) information about fossil species, and then try to draw broader conclusions.
+They might look to articles about other modern scavenging opportunities such as Capaldo and Peter's study of wildebeest drownings http://www.sciencedirect.com/science/article/pii/S0305440385700398, and suggestions that sabertooth cats which they claim were unable to feed on bones may have consumed more bone than penecontemporaneous dire wolves https://www.jstor.org/stable/20627167?seq=1#page_scan_tab_contents.
+
+In sum, while I think the aim of the article is a very worthwhile one, I think the authors cannot claim to have produced an all-encompassing model that can predict the presence or level of scavenging in any organism - because I think this simply cannot be done.
+
+Finally, some of the perspectives of this manuscript overlap substantially with parts of Beasley et al. 2015 (pages 108-112), which is cited numerous times in the manuscript.
+10.     Your section on Handling Time is so short that I think it should be cut or combined with another section.
+
+Reviewer: 2
+
+Comments to the Author
+This article is a review of the literature what the authors term a "natural history" of scavenging.
+They have approached the topic from the perspective of ecological variables (and other variables that condition them) in optimal foraging theory including encounter rate and handling time, but also metabolism, locomotion (e.g soaring and swimming), detection, prey availability, competition, and food processing.
+They describe many useful examples of the behavior and ecology of obligate and facultative scavengers.
+But I think they often go to far in their assumptions in trying to weave these examples together in a synthetic way.
+Some examples are inferring scrounging behaviors in flying pterosaurs (pace 11), assuming that hominins used vultures to identify scavenging opportunities (page 11), surmising that scavenging evolved with the earliest tetrapods based on "potential carrion unexploited by marine vertebrates" (page 12).
+There is a lot of "these researchers have suggested X" and then taking that as a given fact.
+
+While the authors distinguish between obligate and facultative scavenging, they tend to lump them together when it comes to constructing their arguments.
+In the end, is the argument about the former, or the latter?
+Is the real interest in whether species scavenge at all, what proportion of their resources are obtained by scavenging, or what the conditions are that produce facultative or obligate scavengers?
+They also seem to ignore flexibility in species' behaviors.
+For instance, while spotted hyenas are on their highly scavenging side of their Figures 1 and 2, in some ecosystems they hunt significant proportion of their prey.
+Similarly, lions can take a large proportion of their prey as scavenged, as the authors note.
+But to suggest, as they do in their abstract, that they can apply their scale of scavenging to any species at any time to judge how important scavenging was in its diet is frankly completely ignoring behavioral variability over time and space.
+It is substantially more difficult with fossil species for which we may know far fewer of the variables they list in Figure 1, in a nuanced way.
+Figure 2 does not display the diversity of scavengers through time, it displays examples of species that display observed or inferred scavenging behaviors.
+Diversity has specific ecological definition(s) that are being ignored here.
+
+If the authors chose to resubmit, I would suggest they organize their manuscript based on known information about modern species, known (e.g.
+bone processing by theropods) and inferred (e.g.
+scavenging in hominins) information about fossil species, and then try to draw broader conclusions.
+They might look to articles about other modern scavenging opportunities such as Capaldo and Peter's study of wildebeest drownings http://www.sciencedirect.com/science/article/pii/S0305440385700398, and suggestions that sabertooth cats which they claim were unable to feed on bones may have consumed more bone than penecontemporaneous dire wolves
+
+In sum, while I think the aim of the article is a very worthwhile one, I think the authors cannot claim to have produced an all-encompassing model that can predict the presence or level of scavenging in any organism - because I think this simply cannot be done.
+
+Finally, some of the perspectives of this manuscript overlap substantially with parts of Beasley et al. 2015 (pages 108-112), which is cited numerous times in the manuscript.
 

### 2016-07-13 response_letter.tex
Response and manuscript formatted for response

--- 2016-07-12 response_letter.tex
@@
-\documentclass{letter}
+\documentclass[12pt,letterpaper]{article}
+\usepackage{url}
+\usepackage{color}
 %\signature{Your name}
@@
 \begin{document}
-\begin{letter}{Company name \\ Street\\ City\\ Country}
-\opening{To the editor}
+%\begin{letter}{Company name \\ Street\\ City\\ Country}
+To the editor
+
+\bigskip
 
@@
 Please see below where we consider each of your points in turn.
+For improving clarity in this document and in the manuscript, we highlighted our response in blue.
 
-Reviewer: 3
+\section{Reviewer: 1}
+This paper attempts to synthesize literature on traits that enable scavenging behavior in extant and extinct vertebrates.
+The paper is mostly well written, and most (but not all) pertinent literature sources are included.
+I do worry about the lack of novelty of the manuscript, and what I perceived as haphazard organization.
+Here are some specific issues that should be addressed in my opinion:
 
-Comments to the Author
+\begin{enumerate}
+\item{Abstract:  Scavenging can be difficult to observe, but probably no more so than predation.  Also, using experimental carcasses, scavenging is actually quite easy to observe.}
+\item{Abstract:  You should point out that your ``scale of scavenging'' is only qualitative; you made no attempt to quantify such traits.}
+\item{Introduction:  Much of this section seems to follow DeVault's (2003) review paper on scavenging—in fact, there is almost nothing new here that is not covered in that earlier review.}
+\item{Introduction:  You mention that you will use ``other methods in order to discern the most suitable morphologies…''  but you don''t say what they are.  What other methods are you talking about?}
+\item{I think that early on in the paper you need to define obligate and facultative scavengers and then throughout the paper, specify which you are referring to.
+For example, in the first sentence of the Detection section, you talk about ``known scavengers'', but it is unclear if you really mean obligate scavengers.
+Nearly all carnivorous vertebrates are scavengers at least to some extent, so you need to elaborate on what you mean when you talk about ``scavengers''.}
+\item{Last sentence of Introduction:  Again, you propose a ``scale of scavenging'', but without quantification, it is unclear how useful this ``scale'' is.}
+\item{Page 6, 12:  In your discussion of aquatic scavengers, you might want to refer to Beasley et al. (2012; Oikos 121:1021-1026).}
+\item{Page 6:  Why is jaw morphology likely to define scavengers?  Please elaborate.}
+\item{Page 9:  It is true that carrion can comprise a large percentage of a hyena's diet, but your text seems misleading.
+Most studies have shown that most of a hyena's diet is from predation.
+For example, see Cooper et al.
+(1999; Afr.
+J.
+Ecol.
+37:149–160) and Gasaway et al.
+(1991; Afr.
+J.
+Ecol.
+29: 64–75).}
+\item{Your section on Handling Time is so short that I think it should be cut or combined with another section.}
+\end{enumerate}
+
+\section{Reviewer: 2}
+This article is a review of the literature what the authors term a ``natural history'' of scavenging.
+They have approached the topic from the perspective of ecological variables (and other variables that condition them) in optimal foraging theory including encounter rate and handling time, but also metabolism, locomotion (e.g soaring and swimming), detection, prey availability, competition, and food processing.
+They describe many useful examples of the behavior and ecology of obligate and facultative scavengers.
+But I think they often go to far in their assumptions in trying to weave these examples together in a synthetic way.
+Some examples are inferring scrounging behaviors in flying pterosaurs (pace 11), assuming that hominins used vultures to identify scavenging opportunities (page 11), surmising that scavenging evolved with the earliest tetrapods based on ``potential carrion unexploited by marine vertebrates'' (page 12).
+There is a lot of ``these researchers have suggested X'' and then taking that as a given fact.
+
+While the authors distinguish between obligate and facultative scavenging, they tend to lump them together when it comes to constructing their arguments.
+In the end, is the argument about the former, or the latter?
+Is the real interest in whether species scavenge at all, what proportion of their resources are obtained by scavenging, or what the conditions are that produce facultative or obligate scavengers?
+They also seem to ignore flexibility in species' behaviors.
+For instance, while spotted hyenas are on their highly scavenging side of their Figures 1 and 2, in some ecosystems they hunt significant proportion of their prey.
+Similarly, lions can take a large proportion of their prey as scavenged, as the authors note.
+But to suggest, as they do in their abstract, that they can apply their scale of scavenging to any species at any time to judge how important scavenging was in its diet is frankly completely ignoring behavioral variability over time and space.
+It is substantially more difficult with fossil species for which we may know far fewer of the variables they list in Figure 1, in a nuanced way.
+Figure 2 does not display the diversity of scavengers through time, it displays examples of species that display observed or inferred scavenging behaviors.
+Diversity has specific ecological definition(s) that are being ignored here.
+
+If the authors chose to resubmit, I would suggest they organize their manuscript based on known information about modern species, known (e.g.
+bone processing by theropods) and inferred (e.g.
+scavenging in hominins) information about fossil species, and then try to draw broader conclusions.
+They might look to articles about other modern scavenging opportunities such as Capaldo and Peter's study of wildebeest drownings \url{http://www.sciencedirect.com/science/article/pii/S0305440385700398}, and suggestions that sabertooth cats which they claim were unable to feed on bones may have consumed more bone than penecontemporaneous dire wolves \url{https://www.jstor.org/stable/20627167?seq=1#page_scan_tab_contents}.
+
+In sum, while I think the aim of the article is a very worthwhile one, I think the authors cannot claim to have produced an all-encompassing model that can predict the presence or level of scavenging in any organism - because I think this simply cannot be done.
+
+Finally, some of the perspectives of this manuscript overlap substantially with parts of Beasley et al. 2015 (pages 108-112), which is cited numerous times in the manuscript.
+
+\section{Reviewer: 3}
 This is a well-written, enjoyable manuscript that assesses the functional traits associated to past and, mostly, current major vertebrate scavengers.
@@
 
-Major comments:
+\subsection{Major comments:}
 
-1.
-Invertebrates are ignored completely throughout the ms.
+\begin{enumerate}
+\item{Invertebrates are ignored completely throughout the ms.
 Please specify in the Introduction that this review focuses on vertebrate scavengers.
-In this line, I recommend to include “in vertebrates” at the end of the title to be more specific.
-Also, in the “Competition” section you mention competition with vertebrates and micro-organisms, but not with invertebrates.
-You should consider them in this section (and in “Facilitation”; see next point).
+In this line, I recommend to include ``in vertebrates'' at the end of the title to be more specific.
+Also, in the ``Competition'' section you mention competition with vertebrates and micro-organisms, but not with invertebrates.
+You should consider them in this section (and in ``Facilitation''; see next point).}
 
-2.
-I found difficult to follow the main point of the paper because the ms is not well organized.
-Following a hierarchical order, I’d define four/five main sections: “Introduction”, “The Challenges of Scavenging” (this could be a subsection within the previous section), “Encounter Rate” (with four different subsections: “Metabolism”, “Locomotion”, “Sensory Detection” and “Carcass Availability”), “Handling Time” (which would include “Food Processing”) and “Conclusion”.
+\item{I found difficult to follow the main point of the paper because the ms is not well organized.
+Following a hierarchical order, I'd define four/five main sections: ``Introduction'', ``The Challenges of Scavenging'' (this could be a subsection within the previous section), ``Encounter Rate'' (with four different subsections: ``Metabolism'', ``Locomotion'', ``Sensory Detection'' and ``Carcass Availability''), ``Handling Time'' (which would include ``Food Processing'') and ``Conclusion''.
 \\
 
-Moreover, “Competition” and “Facilitation” (I strongly encourage you to consider not only competitive, but also facilitative processes, which are mostly neglected in your review) should be treated as transversal factors that can modulate each of the abovementioned parameters.
-Thus, I’d include them in a different section, or mention the competitive and facilitative processes related to each parameter within each subsection.
+Moreover, ``Competition'' and ``Facilitation'' (I strongly encourage you to consider not only competitive, but also facilitative processes, which are mostly neglected in your review) should be treated as transversal factors that can modulate each of the abovementioned parameters.
+Thus, I'd include them in a different section, or mention the competitive and facilitative processes related to each parameter within each subsection.
 Fig.
-1 should be re-organized accordingly.
+1 should be re-organized accordingly.}
 
-3.
-The use of the scientific literature on scavenging is deficient, especially for a review paper like this.
+\item{The use of the scientific literature on scavenging is deficient, especially for a review paper like this.
 A number of key references are missing, e.g., Selva \& Fortuna (2007 Proc.
@@
 B 274:1101-1108), Wilson \& Wolkovich (2011 TREE 23:129-135), Beasley et al.
-(2012 Oikos 121:1021-1026), Cortés-Avizanda et al.
-(2012 Ecology 93:2570-2579), Cortés-Avizanda et al.
-(2014 Ecology 95:1799-1808), Moleón et al.
+(2012 Oikos 121:1021-1026), Cort\'{e}s-Avizanda et al.
+(2012 Ecology 93:2570-2579), Cort\'{e}s-Avizanda et al.
+(2014 Ecology 95:1799-1808), Mole\'{o}n et al.
 (2014 Biol.
@@
 (2014 Mammal Rev.
-44:44-55), Périquet et al.
+44:44-55), P\'{e}riquet et al.
 (2015 Biol.
 Rev.
-90:1197-1214), Mateo-Tomás et al.
+90:1197-1214), Mateo-Tom\'{a}s et al.
 (2015 Divers.
 Distrib.
-21:913-924), Moleón \& Sánchez-Zapata (2015 BioScience 65:1003-1010), Moreno-Opo et al.
+21:913-924), Mole\'{o}n \& S\'{a}nchez-Zapata (2015 BioScience 65:1003-1010), Moreno-Opo et al.
 (2016 Behav.
-Ecol.), Sebastián-González et al.
+Ecol.), Sebasti\'{a}n-Gonz\'{a}lez et al.
 (2016 Ecology 97:95-105).
@@
 Also, the review would be more informative and useful for readers.
-Next I mention some particular comments regarding the use of the references (I mention others in Minor comments): \\
+Next I mention some particular comments regarding the use of the references (I mention others in Minor comments):}\\
 
-Page 2, 1st sentence: but see Moleón \& Sánchez-Zapata (2015); many of the most charismatic vertebrates of the world are scavengers.
-Rather than “scavengers”, what is not charismatic is probably “scavenging”.
+Page 2, 1st sentence: but see Mole\'{o}n \& S\'{a}nchez-Zapata (2015); many of the most charismatic vertebrates of the world are scavengers.
+Rather than ``scavengers'', what is not charismatic is probably ``scavenging''.
 
-We now refer to the behaviour rather than the species as lacking charisma.
+\textcolor{blue}{We now refer to the behaviour rather than the species as lacking charisma.}
 
- Page 2, 13th line: Moleón \& Sánchez-Zapata (2015) may be more appropriate than Koenig (2006) because the former is more general and not restricted to vultures.
- Page 2, lines 17th-18th: see Pereira et al. (2014) and Périquet et al. (2015) for reviews.
+ Page 2, 13th line: Mole\'{o}n \& S\'{a}nchez-Zapata (2015) may be more appropriate than Koenig (2006) because the former is more general and not restricted to vultures.\\
+ Page 2, lines 17th-18th: see Pereira et al. (2014) and P\'{e}riquet et al. (2015) for reviews.
 
-These references have been added.
+\textcolor{blue}{These references have been added.}
 
- Page 3, bottom paragraph, bottom line: this is not true, as scavenging may indirectly affect herbivore populations and thus carrion availability (see Moleón et al. 2014 Biol. Rev.).
+ Page 3, bottom paragraph, bottom line: this is not true, as scavenging may indirectly affect herbivore populations and thus carrion availability (see Mole\'{o}n et al. 2014 Biol. Rev.).
 
-The potential for scavengers to affect carrion availability has been acknowledged.
+\textcolor{blue}{The potential for scavengers to affect carrion availability has been acknowledged.}
 
- Page 5, bottom paragraph, 1st sentence: Mateo-Tomás et al. (2015) is more general than Kendall (2013).
- Page 8, 1st and 2nd sentences: see Pereira et al. (2014) and Périquet et al. (2015).
+ Page 5, bottom paragraph, 1st sentence: Mateo-Tom\'{a}s et al. (2015) is more general than Kendall (2013).\\
+ Page 8, 1st and 2nd sentences: see Pereira et al. (2014) and P\'{e}riquet et al. (2015).\\
  Page 8, 3rd sentence: you may also cite DeVault et al. (2003) and Pereira et al. (2014).
 
-These references have been added.
+\textcolor{blue}{These references have been added.}
 
@@
 
-We have reworded this section to make the more general point that ambush predators can rely more on hunting than can cursorial species, a point explicitly stated in the Pereira et al.
-(2014) article, which we now cite.
+\textcolor{blue}{We have reworded this section to make the more general point that ambush predators can rely more on hunting than can cursorial species, a point explicitly stated in the Pereira et al.
+(2014) article, which we now cite.}
 
- Page 11, 1st sentence: see Kane et al. (2014) and Moleón et al. (2014 Biol. Rev.) for inter-specific interactions.
- Page 11, 4th sentence: Kane et al. (2014) is not appropriate here because it is about inter-specific interactions; use Cortés-Avizanda et al. (2014) instead.
- Page 14, 2nd paragraph, 1st sentence: see also Moleón et al. (2014 Biol. Rev.) and Pereira et al. (2014).
- Page 14, 2nd paragraph, 2nd sentence: see better Pereira et al. (2014) and Périquet et al. (2015). 
- Page 14, bottom paragraph, 3rd sentence: you may also consider Moreno-Opo et al. (2016).
- Page 15, 2nd paragraph, 2nd sentence: see also Moleón et al. (2014 Biol. Rev.) and Pereira et al. (2014).
+ Page 11, 1st sentence: see Kane et al. (2014) and Mole\'{o}n et al. (2014 Biol. Rev.) for inter-specific interactions.\\
+ Page 11, 4th sentence: Kane et al. (2014) is not appropriate here because it is about inter-specific interactions; use Cort\'{e}s-Avizanda et al. (2014) instead.\\
+ Page 14, 2nd paragraph, 1st sentence: see also Mole\'{o}n et al. (2014 Biol. Rev.) and Pereira et al. (2014).\\
+ Page 14, 2nd paragraph, 2nd sentence: see better Pereira et al. (2014) and P\'{e}riquet et al. (2015). \\
+ Page 14, bottom paragraph, 3rd sentence: you may also consider Moreno-Opo et al. (2016).\\
+ Page 15, 2nd paragraph, 2nd sentence: see also Mole\'{o}n et al. (2014 Biol. Rev.) and Pereira et al. (2014).
 
-All of these references have been added.
+\textcolor{blue}{All of these references have been added.}
 
-4.
-You should include more discussion about the importance of carcass size throughout the ms, as it has a strong influence on scavenging patterns and interactions among vertebrates and between vertebrates and smaller carrion consumers (e.g., see Moleón et al.
-2015).
+\item{You should include more discussion about the importance of carcass size throughout the ms, as it has a strong influence on scavenging patterns and interactions among vertebrates and between vertebrates and smaller carrion consumers (e.g., see Mole\'{o}n et al.
+2015).}
+\end{enumerate}
 
-Minor comments:
+\subsection{Minor comments:}
 
-GENERAL
+\begin{enumerate}
 
-1. Please provide line numbers to facilitate the reviewing process.
+\item{GENERAL:} Please provide line numbers to facilitate the reviewing process.
 
-Line numbers have now been included.
+\textcolor{blue}{Line numbers have now been included. }
 
-ABSTRACT
+\item{ABSTRACT:} You could remove ``, the first to our knowledge''.
 
-2. You could remove “, the first to our knowledge”.
+\textcolor{blue}{This has been removed.}
 
-This has been removed.
+\item{ABSTRACT:} Also, you could mention some relevant specific findings of your synthesis.
 
-3. Also, you could mention some relevant specific findings of your synthesis.
-
-THE CHALLENGES OF SCAVENGING
-
-4.
-Page 3, 1st sentence: “often difficult to predict” compared to what?
+\item{THE CHALLENGES OF SCAVENGING:} Page 3, 1st sentence: ``often difficult to predict'' compared to what?
 Please specify.
@@
 
-ENCOUNTER RATE
+\item{ENCOUNTER RATE:} Page 3, bottom paragraph, 3rd line: change ``Alternatively'' to ``Also''?
 
-5. Page 3, bottom paragraph, 3rd line: change “Alternatively” to “Also”?
+\textcolor{blue}{This has been changed.}
 
-This has been changed.
+\item{METABOLISM:} Page 4, 1st sentence: ``the sporadic nature of carrion''; please explain.
 
-METABOLISM
+\item{LOCOMOTION:} Page 4, bottom paragraph, 1st sentence: change ``inherent'' to ``relative''.
 
-6. Page 4, 1st sentence: “the sporadic nature of carrion”; please explain.
+\textcolor{blue}{This has been changed.}
 
-LOCOMOTION
+\item{LOCOMOTION:} Page 4, bottom paragraph, 3rd sentence: remove ``, paradoxically,''.
 
-7. Page 4, bottom paragraph, 1st sentence: change “inherent” to “relative”.
+\textcolor{blue}{This has been deleted.}
 
-This has been changed.
+\item{LOCOMOTION:} Page 5, 2nd paragraph, 1st sentence: change ``appear'' to ``appeared''. Also, specify Mya for the Cretaceous period.
 
-8. Page 4, bottom paragraph, 3rd sentence: remove “, paradoxically,”.
+\textcolor{blue}{This has been changed.}
 
-This has been deleted.
+\item{LOCOMOTION:} Page 9, top paragraph: please be more cautious in your assertion, as cooperation skills and tools of early hominins probably allowed them to defend and usurp carcasses from more powerful scavengers.
 
-9. Page 5, 2nd paragraph, 1st sentence: change “appear” to “appeared”. Also, specify Mya for the Cretaceous period.
+\textcolor{blue}{We have changed this paragraph and take into account the authors' reference to tool use and sociality in hominins.
+The point is that even with these advantages, running to a carcass over long distances still wasn't energetically sensible.}
 
-This has been changed.
+\item{DETECTION:} Page 11, top and 2nd paragraph, 1st sentence: this is about facilitation.
+As said above, facilitative processes may affect other parameters as well, not only the ability to detect carrion.
 
-10.
-Page 9, top paragraph: please be more cautious in your assertion, as cooperation skills and tools of early hominins probably allowed them to defend and usurp carcasses from more powerful scavengers.
+\item{DETECTION:} Page 11, last sentence: explain why wind speed affects the olfactory capacity of an aquatic species.
 
-We have changed this paragraph and take into account the authors' reference to tool use and sociality in hominins.
-The point is that even with these advantages, running to a carcass over long distances still wasn't energetically sensible.
+\item{PREY AVAILABILITY:} I'd change this title to ``Carcass Availability'', which is more specific.
 
-DETECTION
+\textcolor{blue}{This has been changed.}
 
-11. Page 11, top and 2nd paragraph, 1st sentence: this is about facilitation. As said above, facilitative processes may affect other parameters as well, not only the ability to detect carrion.
-
-12. Page 11, last sentence: explain why wind speed affects the olfactory capacity of an aquatic species.
-
-PREY AVAILABILITY
-
-13. I’d change this title to “Carcass Availability”, which is more specific.
-
-This has been changed.
-
-14.
-Among the different components of carcass availability, e.g., abundance, duration, predictability and accessibility, in this section you only talk about abundance and duration.
-You should also discuss about carcass predictability (e.g., see Cortés-Avizanda et al.
+\item{PREY AVAILABILITY:} Among the different components of carcass availability, e.g., abundance, duration, predictability and accessibility, in this section you only talk about abundance and duration.
+You should also discuss about carcass predictability (e.g., see Cort\'{e}s-Avizanda et al.
 2012) and carcass accessibility (e.g., see DeVault \& Krochmal 2002).
 
-15.
-Page 12, 2nd paragraph, 1st sentence: as said before, this is not true, given that scavenging may indirectly affect herbivore populations and thus carrion availability (see Moleón et al.
+\item{PREY AVAILABILITY:} Page 12, 2nd paragraph, 1st sentence: as said before, this is not true, given that scavenging may indirectly affect herbivore populations and thus carrion availability (see Mole\'{o}n et al.
 2014 Biol.
@@
 
-We have removed the part of the sentence where we mistakenly suggest that scavengers cannot affect carrion availability.
+\textcolor{blue}{We have removed the part of the sentence where we mistakenly suggest that scavengers cannot affect carrion availability.}
 
-16.
-Page 13, 3rd and 4th sentences: whaling has exerted also a huge negative impact on scavengers by depleting whale populations.
+\item{PREY AVAILABILITY:} Page 13, 3rd and 4th sentences: whaling has exerted also a huge negative impact on scavengers by depleting whale populations.
 If you are going to talk about human impacts, maybe is better to do it at the end of the ms, perhaps in a new section in which you explore which scavenging traits could be more successful in a human dominated world.
 
-This has been removed.
+\textcolor{blue}{This has been removed.}
 
-17.
-You should also take into account facilitative processes.
-For instance, some scavengers are able to access the interior of carcasses only once other, more powerful species, has opened the skin (see Moleón et al.
+\item{PREY AVAILABILITY:} You should also take into account facilitative processes.
+For instance, some scavengers are able to access the interior of carcasses only once other, more powerful species, has opened the skin (see Mole\'{o}n et al.
 2014 Biol.
@@
 
-18. Page 14, bottom paragraph, 2nd sentence: I’d change it to “This is true for many vultures and other major avian scavengers such as albatrosses who all have […]”.
+\item{PREY AVAILABILITY:} Page 14, bottom paragraph, 2nd sentence: I'd change it to ``This is true for many vultures and other major avian scavengers such as albatrosses who all have […]''.
 
-We have reworded the sentence as suggested.
+\textcolor{blue}{We have reworded the sentence as suggested.}
 
-FOOD PROCESSING
+\item{FOOD PROCESSING:} Page 16, 2nd paragraph, 6th sentence: insert ``Within mammals,'' before ``This ability''.
 
-19. Page 16, 2nd paragraph, 6th sentence: insert “Within mammals,” before “This ability”.
+\textcolor{blue}{This has been added.}
 
-This has been added.
+\item{FOOD PROCESSING:} Page 17, 2nd paragraph, 4th sentence: do you refer to species or individuals? Higher species diversity doesn''t mean more individuals.
 
-20. Page 17, 2nd paragraph, 4th sentence: do you refer to species or individuals? Higher species diversity doesn’t mean more individuals.
+\textcolor{blue}{The reference of Pobiner (2015) states that there were more individuals and we have changed the sentence to reflect this.}
 
-The reference of Pobiner (2015) states that there were more individuals and we have changed the sentence to reflect this.
+\item{FOOD PROCESSING:} Page 18, 1st sentence: change ``enabling'' to ``enable''.
 
-21. Page 18, 1st sentence: change “enabling” to “enable”.
+\textcolor{blue}{This has been changed.}
 
-This has been changed.
+\item{FOOD PROCESSING:} Page 18, 3rd sentence: ``thus''? I don''t understand this sentence; could you please reword it?
 
-22. Page 18, 3rd sentence: “thus”? I don’t understand this sentence; could you please reword it?
+\textcolor{blue}{We have reworded this sentence. The idea is that early birds didn't have beaks and so could more easily eat meat. This is in contrast to modern birds where a beak may be a hindrance.}
 
-We have reworded this sentence. The idea is that early birds didn't have beaks and so could more easily eat meat. This is in contrast to modern birds where a beak may be a hindrance.
+\item{FOOD PROCESSING:} Page 18, 5th sentence: ``The mix of strong and weak features in their skull morphology''; please explain it.
 
-23. Page 18, 5th sentence: “The mix of strong and weak features in their skull morphology”; please explain it.
+\item{CONCLUSION:} Page 18, bottom paragraph, 3rd sentence: change ``technique'' to ``approach''?
 
-CONCLUSION
+\textcolor{blue}{This has been changed.}
 
-24. Page 18, bottom paragraph, 3rd sentence: change “technique” to “approach”?
+\item{FIGURES:} Fig.
+2: you may insert the name and Mya range of each period in the X axis rather than in the figure legend.
+Also, I'd indicate ``aerial'', ``terrestrial'' and ``aquatic'' in the Y axis rather than in the legend.
 
-This has been changed.
-
-FIGURES
-
-25.
-Fig.
-2: you may insert the name and Mya range of each period in the X axis rather than in the figure legend.
-Also, I’d indicate “aerial”, “terrestrial” and “aquatic” in the Y axis rather than in the legend.
+\end{enumerate}
 
 Hope this helps,
-Marcos Moleón
+Marcos Mole\'{o}n
 
@@
 %\encl{Enclosures.}
-\end{letter}
+%\end{letter}
 \end{document}

### 2016-07-13 natural history of scavenging.tex
Response and manuscript formatted for response

@@
 \usepackage{lineno}
+\usepackage{color}
 
@@
 
-\title{A Recipe for Scavenging in Vertebrates - the natural history of a behaviour}
+\title{A Recipe for Scavenging \textcolor{blue}{in Vertebrates} - the natural history of a behaviour}
 
@@
 The result is a synthesis on the natural history of scavenging. 
-Our idea of a scale of scavenging can be applied to any species at any time to help judge the importance of this behaviour in its diet. 
+Our idea of a scale of scavenging can be applied to any species at any time to help \textcolor{blue}{judge} the importance of this behaviour in its diet. 
 \end{abstract}
@@
 
-Historically, food webs have classified most acts of carnivory as predation events \citep{wilson2011scavenging}. 
+Historically, \textcolor{blue}{food webs have classified most acts of carnivory as predation events \citep{wilson2011scavenging}. 
 This is clearly a mistaken view because of the existence of scavenging, a behaviour displayed by almost all carnivorous vertebrates \citep{devault2003scavenging}. 
@@
 Unfortunately, the lack of direct behavioural data makes it difficult to discern scavenging from predation among extinct forms where it was surely exhibited to a similar extent among prehistoric vertebrate carnivores.
-Some simple heuristics can inform us whether or not the prey was scavenged, for instance, in cases where it was simply too large to have been killed by the ostensible predator \citep{pobiner2008paleoecological}.
+Some simple heuristics can inform us whether or not the prey was scavenged, for instance, in cases where it was simply too large to have been killed by the ostensible predator \citep{pobiner2008paleoecological}.}
 But clearly, a scavenger does not only feed on animals too big for it to have hunted.
-There are other methods however, that allow us to discern the most suitable morphologies, physiologies and environments for a scavenging lifestyle to prosper e.g.
+\textcolor{blue}{There are other methods however, that allow us to discern the most suitable morphologies, physiologies and environments for a scavenging lifestyle to prosper e.g.
 energetics models, comparative anatomy, palaeontology etc.
@@
 Yet, a synthesis describing the natural history of scavengers is absent from the literature.
-Here we rectify this shortcoming in charting the natural history of scavenging by assessing its potential among extinct vertebrate groups given their ecology, functional traits and the enviornmental context.
+Here we rectify this shortcoming in charting the natural history of scavenging by assessing its potential among extinct vertebrate groups given their ecology, functional traits and the environmental context.}
 
@@
 
-Species that rely on scavenging to sustain substantial portions of their diets must encounter a sufficient amount of carrion in order to meet their energetic demands.
-Once found, the scavenger must be able to out-compete any potential competitors and efficiently process the, typically decaying, carcass replete with micro-organism derived toxins \citep{ruxton2014fruit}.
+\textcolor{blue}{Species that rely on scavenging to sustain substantial portions of their diets must encounter a sufficient amount of carrion in order to meet their energetic demands.
+Once found, the scavenger must be able to out-compete any potential competitors and efficiently process the, typically decaying, carcass replete with micro-organism derived toxins \citep{ruxton2014fruit}.}
 These characteristics can be assumed to be under evolutionary selection pressures for traits that increase carrion discovery and monopoly.
-Finally, the potential for scavenging will also depend on the density, size, and quality of carcases produced, all of which are affected by complex ecosystem dynamics \citep{moleon2014inter}. 
-Even these aspects can be indirectly impacted by the behaviour of the scavenger \citep{moleon2014inter}.
+Finally, the potential for scavenging will also depend on the density, size, and quality of carcases produced, all of which are affected by complex ecosystem dynamics \textcolor{blue}{\citep{moleon2014inter}.
+Even these aspects can be indirectly impacted by the behaviour of the scavenger \citep{moleon2014inter}.}
 Each of these facets are essentially the backbone of fundamental ecological theory and are the key parameters defined in functional response curves, namely encounter rate, handling time and prey availability \citep{jeschke2002predator}.
 By considering scavenging in this context of optimal foraging we can identify the prerequisite attributes and processes required for the behaviour. 
-This has enabled us to propose a scale of scavenging whereupon we can place an extinct vertebrate carnivore and assess the likely importance of carrion in its diet.
+This has enabled us to propose a scale of scavenging whereupon we can place \textcolor{blue}{an extinct vertebrate carnivore} and assess the likely importance of carrion in its diet.
 
@@
 However, as we noted above, encounter rate is also determined by the productivity of the environment, and ultimately the rate at which carcasses are produced. 
-Scavengers can drive this process indirectly, for instance, by consuming the majority of carrion, obligate scavengers deny a resource to their facultative counterparts obliging the latter to hunt more and thus produce more carrion \citep{moleon2014inter}.
+\textcolor{blue}{Scavengers can drive this process indirectly, for instance, by consuming the majority of carrion, obligate scavengers deny a resource to their facultative counterparts obliging the latter to hunt more and thus produce more carrion \citep{moleon2014inter}.}
 
@@
 \subsection{Locomotion}
-The relative unpredictability of carrion also means scavenging depends more on the ability to efficiently move over larger areas than does predation.
+The \textcolor{blue}{relative} unpredictability of carrion also means scavenging depends more on the ability to efficiently move over larger areas than does predation.
 This generally requires an efficient transfer of metabolic energy into movement which relies on the animal's anatomy and physiology as well as the medium of the environment in which the animal is moving (i.e.
@@
 
-We know that many extant birds exist as facultative scavengers because storks, raptors and corvids all take substantial quantities of carrion in their diet \citep{mateo2015regional}. 
+We know that many extant birds exist as facultative scavengers because storks, raptors and corvids all take substantial quantities of carrion in their diet \textcolor{blue}{\citep{mateo2015regional}}.
 Similarly we would expect that extinct species would also scavenge in a similar fashion depending on the efficiency of their flight. 
@@
 Modern endothermic mammals can sustain longer periods of energetically expensive activity \citep{bennett1979endothermy} resulting in larger foraging ranges. 
-Today, terrestrial scavenging in the mammals is probably best known in the species rich African context where hyenas, jackals and lions all take sizeable proportions of carrion in their diet \citep{pereira2014facultative,periquet2015lion}.
+Today, terrestrial scavenging in mammals is probably best known in the African Savannah context where hyenas, jackals and lions all take sizeable proportions of carrion in their diet \textcolor{blue}{\citep{pereira2014facultative,periquet2015lion}.}
 In the spotted hyena (\textit{Crocuta crocuta}), striped hyena (\textit{Hyaena hyaena}) and brown hyena (\textit{Hyaena brunnea}) it can be over 90\% \citep{jones2015african}.
-And although no contemporary terrestrial vertebrate exists as an obligate scavenger, most, if not all, are facultative to some extent \citep{devault2003scavenging,beasley2015vertebrates,pereira2014facultative}.
-The particular reliance of hyenas on carrion means we can use them as examples of efficient terrestrial scavengers to compare with other forms \citep{periquet2015lion}. 
-In terms of locomotion, they employ a characteristic ``rocking horse gait"  which allows them to cover great distances efficiently, loping at 10 km/hr \citep{mills1989comparative,jones2015african}. 
-Such long-distance travel is apparent many other cursorial predators \citep{pennycuick1995radius,janis2014forelimb,pereira2014facultative}. 
+And although no contemporary terrestrial vertebrate exists as an obligate scavenger, most, if not all, are facultative to some extent \textcolor{blue}{\citep{devault2003scavenging,beasley2015vertebrates,pereira2014facultative}}.
+The particular reliance of hyenas on carrion means we can use them as examples of efficient terrestrial scavengers to compare with other forms \textcolor{blue}{\citep{periquet2015lion}}. 
+In terms of locomotion, they employ a characteristic ``rocking horse gait''  which allows them to cover great distances efficiently, loping at 10 km/hr \citep{mills1989comparative,jones2015african}. 
+Such long-distance travel is apparent \textcolor{blue}{in many other cursorial predators \citep{pennycuick1995radius,janis2014forelimb,pereira2014facultative}}. 
 In contrast, big cats rely on ambush \citep{pennycuick1995radius,pereira2014facultative}. 
@@
 Our locomotory efficiency has fed into the question of where our ancestors placed on the hunter-scavenger axis during the Plio-Pleistocene, which has been a matter of debate for years \citep{dominguez2002hunting}.
-\cite{ruxton2013endurance} added to this debate with their argument that long distance endurance running was not an important feature of hominin scavenging but was instead used by humans for hunting.
-Even when considering the use of weapons and social foraging, the authors argue that the high costs of running long distances would render it an unattractive foraging strategy for scavenging.
+\cite{ruxton2013endurance} added to this debate with their argument that long distance endurance running \textcolor{blue}{was not an important feature of hominin scavenging but was instead used by humans for hunting.
+Even when considering the use of weapons and social foraging, the authors argue that the high costs of running long distances would render it an unattractive foraging strategy for scavenging.}
 
@@
 
-\subsection{Sensory Detection}
+\subsection{\textcolor{blue}{Sensory} Detection}
 
@@
 
-Existing in this environment also means being able to gather a wealth of information from other foragers, be they conspecifics or otherwise \citep{jackson2008effect,KaneVul,moleon2014inter}.
+Existing in this environment also means being able to gather a wealth of information from other foragers, be they conspecifics or otherwise \textcolor{blue}{\citep{jackson2008effect,KaneVul,moleon2014inter}}.
 These interactions are properly three dimensional in the sense of \cite{pawar2012dimensionality} because both producer and scrounger are in the air \citep{dall2005information}. 
 Again, returning to vultures, the genus \textit{Gyps} consists of highly social and colonially nesting species \citep{fernandez2015density}.
-These behaviours allow them to forage far more efficiently because one bird can scrounge information on the location of food from another successful forager \citep{cortes2014bird}.
+These behaviours allow them to forage far more efficiently because one bird can scrounge information on the location of food from another successful forager \textcolor{blue}{\citep{cortes2014bird}}.
 Information transfer of this kind is typically inadvertent and as a consequence no complex social interactions are required, simply the ability to recognise a successful forager.
@@
 
-\subsection{Carcass Availability}
-The environmental influence on carcass availability is an aspect that greatly affects encounter rate. %but is invisible to the selective forces acting on the scavenger. 
+\subsection{\textcolor{blue}{Carcass} Availability}
+The environmental influence on \textcolor{blue}{carcass} availability is an aspect that greatly affects encounter rate. %but is invisible to the selective forces acting on the scavenger. 
 Aspects including, primary productivity, relief, and temperature will all affect scavenging tendency. 
@@
 
-The physical differences between water and air mean carcass availability is radically different between these environments \citep{beasley2012carrion}. 
+\textcolor{blue}{The physical differences between water and air mean carcass availability is radically different between these environments \citep{beasley2012carrion}. 
 For one, carcasses get moved around by water which results in a more diffuse signal being produced for would-be scavengers.
@@
 Research has shown that an animal need only travel between 16 to 36 km to encounter a fresh whale carcass \citep{moleon2015living}. 
-The phenomenon of occasional bounties of carrion in the form of these whale falls has led some researchers to investigate if an obligate scavenger could survive by seeking out these remains exclusively.
+The phenomenon of occasional bounties of carrion in the form of these whale falls has led some researchers to investigate if an obligate scavenger could survive by seeking out these remains exclusively.}
 \cite{ruxton2005searching} argued that although this is energetically feasible it's ecologically unlikely.
@@
 \subsection{Competition}
-Large body size has substantial advantages in agonistic interactions \citep{ruxton2004obligate,moleon2014inter,pereira2014facultative}. 
-For instance, lions can acquire much of their carrion through kleptoparasitism of hyena kills \citep{trinkel2005competitive,pereira2014facultative,periquet2015lion}. 
+Large body size has substantial advantages in agonistic interactions \textcolor{blue}{\citep{ruxton2004obligate,moleon2014inter,pereira2014facultative}}. 
+For instance, lions can acquire much of their carrion through kleptoparasitism of hyena kills \textcolor{blue}{\citep{trinkel2005competitive,pereira2014facultative,periquet2015lion}}. 
 This line of reasoning suggests that some theropod dinosaurs, who could get up to 15 tonnes, would have had easily monopolised a carcass \citep{weishampel2004dinosauria} provided they could find them efficiently \citep{kane2016body}.
@@
 We would expect this trait to be selected for even in the case of weight-constrained, scavenging fliers.
-This is true for many vultures and other major avian scavengers such as albatrosses who all have who can have body masses in excess of 10 kg and represent some of the heaviest bird species capable of flight \citep{weimerskirch1992reproductive,ferguson2001raptors,donazar2002effects}.
-Indeed, such is the competitive advantage held by vultures over other facultative scavenging birds that temporal niche partitioning at the carcass has evolved \citep{kendall2013alternative,KaneVul,moreno2016behavioral}.
+This is true for \textcolor{blue}{many vultures and other major avian scavengers such as albatrosses who can have body masses in excess of} 10 kg and represent some of the heaviest bird species capable of flight \citep{weimerskirch1992reproductive,ferguson2001raptors,donazar2002effects}.
+Indeed, such is the competitive advantage held by vultures over other facultative scavenging birds that temporal niche partitioning at the carcass has evolved \textcolor{blue}{\citep{kendall2013alternative,KaneVul,moreno2016behavioral}}.
 Additionally, many pterosaurs were far bigger again, with estimated body masses of over 200 kg in the Azhdarchids \citep{witton2010size}.
@@
 Moreover, some researchers have charted an evolutionary course from basal fossorial snakes to modern terrestrial species by way of an obligate scavenger intermediate \citep{bauchot2006snakes}. 
-In water, by contrast, competition with micro-organims is significantly reduced because of the high pressures and low temperatures \citep{beasley2012carrion}.
+\textcolor{blue}{In water, by contrast, competition with micro-organims is significantly reduced because of the high pressures and low temperatures \citep{beasley2012carrion}.}
 
@@
 Osteophagy is known across a range of terrestrial carnivores and given that some fat-rich mammalian bones have an energy density (6.7 kJ/g) comparable with that of muscle tissue, it makes skeletal remains an enticing resource \citep{brown1989study}.
-Within mammals, this ability reached its zenith among hyenas with the evolution of the estimated 110 kg \textit{Pachycrocuta brevirostris} during the Pliocene \citep[3.6 - 2.58 Mya; ][]{palmqvist2011giant}.
-Indeed, their extinction has been blamed on the decline of sabretooth cats (Machairodontinae), because the unique skull morphology of the latter meant they would leave a large amount of food on a carcass for would-be scavengers \citep{palmqvist2011giant} (Note, however that the inability of these cats to deal with bone may be overstated, \citep{binder2010comparison}).
+\textcolor{blue}{Within mammals,} this ability reached its zenith among hyenas with the evolution of the estimated 110 kg \textit{Pachycrocuta brevirostris} during the Pliocene \citep[3.6 - 2.58 Mya; ][]{palmqvist2011giant}.
+Indeed, their extinction has been blamed on the decline of sabretooth cats (Machairodontinae), \textcolor{blue}{because} the unique skull morphology of the latter meant they would leave a large amount of food on a carcass for would-be scavengers (\citealt{palmqvist2011giant}; \textcolor{blue}{note, however that the inability of these cats to deal with bone may be overstated; \citealt{binder2010comparison}}).
 Earlier in the evolution of mammals, the bone-crushing dogs that evolved during the Oligocene (Borophaginae; 33.9 - 23.03 Mya) have also been compared to hyenas in terms of their feeding ecology \citep{van2003chapter,martin2016pursuit}.
@@
 A recent study investigating potential scavenging opportunities for hominins in Kenya found that, in addition to skeletal material, there is a substantial amount of scavengeable meat left on predated remains; sufficient to sustain the requirements of an adult male \textit{Homo erectus} \citep{pobiner2015new}.
-In some historical hominin-inhabited areas there were higher populations of felids compared to hyenids.
+In some historical hominin-inhabited areas there were \textcolor{blue}{higher populations of felids compared to hyenids.}
 Again, this is significant because hyenas are likely to have left far less flesh on a carcass than a felid such as a sabretooth, enabling contemporaneous hominins to benefit \citep{pobiner2015new}.
@@
 Aside from raptors, the specialised beaks of many modern bird lineages tends to hinder their ability to eat meat which is in contrast to the first lineages that did not have this feature \citep{martyniuk2012field}.
-As \cite{martyniuk2012field} notes, this skull morphology of early birds indicates they were predominantly carnivorous, implying scavenging was a live opportunity cf. their descendants. 
+As \cite{martyniuk2012field} notes, \textcolor{blue}{this skull morphology of early birds indicates} they were predominantly carnivorous, implying scavenging was a live opportunity cf.
+their descendants.
 Among the pterosaurs, \cite{witton2013pterosaurs} makes the case that the istiodactyl pterosaurs were the most likely scavengers of this group based on their potential handling time. 
@@
 The animals of today, while often different (sometimes radically so) to their ancestors, can be used to make informed comparisons to extinct species. 
-We have used this approach to give insight into the drivers of scavenging across vertebrates through time.
+We have used this \textcolor{blue}{approach} to give insight into the drivers of scavenging across vertebrates through time.
 In common with any other forager be they grazer, browser or predator, scavengers past and present have had to balance their energetic costs with the gains of food. 

### 2016-07-13 response_letter.tex
Comments on reviewers 1 suggestions

@@
 \item{Introduction: Much of this section seems to follow DeVault's (2003) review paper on scavenging—in fact, there is almost nothing new here that is not covered in that earlier review.}
-\item{Introduction:  You mention that you will use ``other methods in order to discern the most suitable morphologies…''  but you don''t say what they are.  What other methods are you talking about?}
+
+\textcolor{blue}{We rewrote the introduction to reflect more accurately the aim of this review.} %TG: i.e. the scavenging gradient allows us to infer scavenging in the fossil record?
+
+\item{Introduction: You mention that you will use ``other methods in order to discern the most suitable morphologies...''  but you don't say what they are. What other methods are you talking about?}
 \item{I think that early on in the paper you need to define obligate and facultative scavengers and then throughout the paper, specify which you are referring to.
@@
 Nearly all carnivorous vertebrates are scavengers at least to some extent, so you need to elaborate on what you mean when you talk about ``scavengers''.}
+
+%TG: maybe say something along the lines:
+\textcolor{blue}{We do not consider a binary response to scavenging behavior (i.e.
+facultative/obligate) and argue that all carnivorous vertebrates are placed along a gradient of scavenging.
+We state this more clearly in the introduction: \textit{``there is no discrete divide between predators and scavengers but rather a continuous gradient (Pereira et al.
+2014)''}.}
+
 \item{Last sentence of Introduction: Again, you propose a ``scale of scavenging'', but without quantification, it is unclear how useful this ``scale'' is.}
+
 \item{Page 6, 12: In your discussion of aquatic scavengers, you might want to refer to Beasley et al. (2012; Oikos 121:1021-1026).}
+
+%TG: yep we did that
+
 \item{Page 6: Why is jaw morphology likely to define scavengers?  Please elaborate.}
+
+%TG: something along the lines: "The jaw is the principal tool for food ingestion and it's morphology will determine and the range of possible food items the organism is able to ingest"
+
 \item{Page 9: It is true that carrion can comprise a large percentage of a hyena's diet, but your text seems misleading.
@@
 \item{Your section on Handling Time is so short that I think it should be cut or combined with another section.}
+
+\textcolor{blue}{We rewrote the introduction to reflect more accurately the aim of this review.}
+
 \end{enumerate}
@@
 
+\textcolor{blue}{We changed the title of the manuscript to ``A Recipe for Scavenging in Vertebrates - the natural history of a behaviour''.}
+%TG: Also maybe mention invertebrates in the competition part of the manuscript.
++ For a response to the reviewer, here is probably a good point for stating clearly that we wanted this manuscript to be focusing more on vertebrates because of the palaeo aspect.
+Both micro-organisms and invertebrates (I guess mainly arthropods) are unlikely to display preserved morphological clues for suggesting a scavenging behaviour because of their poor preservation anyway.
+
 \item{I found difficult to follow the main point of the paper because the ms is not well organized.
@@
 1 should be re-organized accordingly.}
+
+\textcolor{blue}{We clarified the structure of the manuscript by introducing clearer hierarchical subsections.}
+%TG: + more blabalal on his ``Facilitation'' part
 
@@
 
-\item{FOOD PROCESSING:} Page 17, 2nd paragraph, 4th sentence: do you refer to species or individuals? Higher species diversity doesn''t mean more individuals.
+\item{FOOD PROCESSING:} Page 17, 2nd paragraph, 4th sentence: do you refer to species or individuals? Higher species diversity doesn't mean more individuals.
 
@@
 
-\item{FOOD PROCESSING:} Page 18, 3rd sentence: ``thus''? I don''t understand this sentence; could you please reword it?
+\item{FOOD PROCESSING:} Page 18, 3rd sentence: ``thus''? I don't understand this sentence; could you please reword it?
 

### 2016-07-13 response_letter.tex
added facilitation sections

@@
 Please see below where we consider each of your points in turn.
-For improving clarity in this document and in the manuscript, we highlighted our response in blue.
+For improved clarity in this document and in the manuscript, we have highlighted our responses in \textcolor{blue}{blue}.
+
+\section{Recommendation by the Subject Editor (anonymous):}
+Three experienced reviews have evaluated your manuscript.
+Two of them are extremely critical of several aspects related to to the general handcraft and novelly of the review.
+The third reviewer is more positive, while making substantial recommendations for improvement of the manuscript.
+Should you be able to address all of the reviewers comments in a much revised version of the manuscript I would happy to reconsider the manuscript for consideration in Ecography and the E4 Award.
+This is not a provisional acceptance, since the manuscript would need to be revised again.
+
+\smallskip
+
+\textcolor{blue}{As we note below, we feel our earlier manuscript did not capture the hierarchy we had in mind.
+This updated version is more explicit in terms of the structure where we now highlight the sections and subsections.
+Regarding the novelty of our manuscript, we have changed our introduction which was regarded as too similar to deValult et al.
+(2003) and argue our case for what we consider inevitable overlap with Beasley et al.
+(2015).}
 
@@
 
-\textcolor{blue}{We rewrote the introduction to reflect more accurately the aim of this review.} %TG: i.e. the scavenging gradient allows us to infer scavenging in the fossil record?
+\textcolor{blue}{We rewrote the introduction to more accurately reflect the aim of this review.} %TG: i.e. the scavenging gradient allows us to infer scavenging in the fossil record?
 
 \item{Introduction: You mention that you will use ``other methods in order to discern the most suitable morphologies...''  but you don't say what they are. What other methods are you talking about?}
+
+\textcolor{blue}{We now specify the approaches we have in mind:  \textit{``e.g. energetics models, comparative anatomy, palaeontology etc.'' }}
+
 \item{I think that early on in the paper you need to define obligate and facultative scavengers and then throughout the paper, specify which you are referring to.
@@
 
-%TG: yep we did that
+\textcolor{blue}{We have included this reference and explicitly discuss how differences in the physical properties of water Vs air can affect scavenging: \textit{``The physical differences between water and air mean carcass availability is radically different between these environments (Beasley et al.
+2012).
+For one, carcasses get moved around by water which results in a more diffuse signal being produced for would-be scavengers.
+Carcasses also tend to sink in water where they are no longer accessible to pelagic scavengers (Beasley et al.
+2012).
+Research has shown that an animal need only travel between 16 to 36 km to encounter a fresh whale carcass  (Moleón and Sánchez-Zapata 2015).
+The phenomenon of occasional bounties of carrion in the form of these whale falls has led some researchers to investigate if an obligate scavenger could survive by seeking out these remains exclusively.''}}
 
@@
 29: 64–75).}
+
+\textcolor{blue}{We have clarified this point about hyenas to say it is their ability to take large proportions of carrion under certain circumstances that makes them a useful focal species: \textit{``The particular ability of hyenas to subsist on high proportions of carrion means we can use them as examples of efficient terrestrial scavengers to compare with other forms.''}}
+
 \item{Your section on Handling Time is so short that I think it should be cut or combined with another section.}
 
-\textcolor{blue}{We rewrote the introduction to reflect more accurately the aim of this review.}
+\textcolor{blue}{The hierarchy of our paper wasn't clear in the first version but we have now clarified that this is the lead in to a larger section which has subsections e.g. food processing}
 
@@
 
+\smallskip
+\textcolor{blue}{One of the central points of our paper is to compare the biology of extant scavengers with extinct species in order to determine if they were suited to scavenging.
+All of the examples we use are backed up by references from previous work.
+Rather than taking them as fact we offer these as possible behaviours that the extinct species engaged in.
+Because we can't observe their behaviour directly we are left to make assumptions about how likely they were to have enganged in them.
+But this is hardly unique to our work.
+We have added some extra references in support of our claims: \textit {``Thus, given pterosaurs seem to have cohabited in large numbers (Witton 2013), and the known benefits this can have for social foraging in birds (Dermody et al.
+2011), it seems probable that scrounging behaviours were seen in the flying pterosaurs as well.''}}
+\smallskip
+
 While the authors distinguish between obligate and facultative scavenging, they tend to lump them together when it comes to constructing their arguments.
@@
 Is the real interest in whether species scavenge at all, what proportion of their resources are obtained by scavenging, or what the conditions are that produce facultative or obligate scavengers?
+
 They also seem to ignore flexibility in species' behaviors.
@@
 But to suggest, as they do in their abstract, that they can apply their scale of scavenging to any species at any time to judge how important scavenging was in its diet is frankly completely ignoring behavioral variability over time and space.
+\smallskip
+
+\textcolor{blue}{It is certainly true that facultative scavengers can be variable in the proportion of carrion they consume but we would argue that species like hyenas have certain characteristics that make this possible.
+In contrast, cheetahs do not have a biology that lends itself to scavenging in the wild, due to poor competitive ability etc.
+We appreciate that there is a continuum across and among species which we specifically note in this revised version.}
+\smallskip
+
 It is substantially more difficult with fossil species for which we may know far fewer of the variables they list in Figure 1, in a nuanced way.
@@
 Finally, some of the perspectives of this manuscript overlap substantially with parts of Beasley et al. 2015 (pages 108-112), which is cited numerous times in the manuscript.
+\smallskip
+
+\textcolor{blue}{The sections of Beasley et al.
+(2015) are the evolution of scavengers, competition between scavengers and environmental effects on scavenger ecology.
+These are very broad areas so in a review of this sort some overlap is inevitable.
+We feel our emphasis on extinct species and systems as well as the frame of optimal foraging theory gives this manuscript its novelty.}
+\smallskip
 
@@
 
-\textcolor{blue}{This has been removed.}
+\textcolor{blue}{We have decided to remove this section because we felt human impacts go beyond the scope of our paper.}
 

### 2016-07-13 natural history of scavenging.tex
added facilitation sections

@@
 And although no contemporary terrestrial vertebrate exists as an obligate scavenger, most, if not all, are facultative to some extent \textcolor{blue}{\citep{devault2003scavenging,beasley2015vertebrates,pereira2014facultative}}.
-The particular reliance of hyenas on carrion means we can use them as examples of efficient terrestrial scavengers to compare with other forms \textcolor{blue}{\citep{periquet2015lion}}. 
+\textcolor{blue}{The particular ability of hyenas to subsist on high proportions of carrion means we can use them as examples of efficient terrestrial scavengers to compare with other forms \citep{periquet2015lion}}.
 In terms of locomotion, they employ a characteristic ``rocking horse gait''  which allows them to cover great distances efficiently, loping at 10 km/hr \citep{mills1989comparative,jones2015african}. 
-Such long-distance travel is apparent \textcolor{blue}{in many other cursorial predators \citep{pennycuick1995radius,janis2014forelimb,pereira2014facultative}}. 
+Such long-distance travel is apparent \textcolor{blue}{in many other cursorial predators \citep{pennycuick1995radius,janis2014forelimb,pereira2014facultative}. 
 In contrast, big cats rely on ambush \citep{pennycuick1995radius,pereira2014facultative}. 
-This difference means ambush predators tend to be more obligate hunters \citep{pereira2014facultative} 
+This difference means ambush predators tend to be more obligate hunters \citep{pereira2014facultative}.} 
 These insights allow us to compare extant terrestrial species to their prehistoric forebears given the dominance of mammalian carnviores since the Eocene (56-33.9 Mya) where the order split into the Caniforma and Feliforma \citep{van1987skeletal}.
@@
 Information transfer of this kind is typically inadvertent and as a consequence no complex social interactions are required, simply the ability to recognise a successful forager.
-Thus, it seems probable that scrounging behaviours were seen in the flying pterosaurs as well.
+\textcolor{blue}{Thus, given pterosaurs seem to have cohabited in large numbers \citep{witton2013pterosaurs}, and the known benefits this can have for social foraging in birds \citep{jackson2011evolutionary}, it seems probable that scrounging behaviours were seen in the flying pterosaurs as well.}
 
@@
 
-\subsection{\textcolor{blue}{Carcass} Availability}
+\subsection{\textcolor{blue}{Abiotic effects on Carcass} Availability}
 The environmental influence on \textcolor{blue}{carcass} availability is an aspect that greatly affects encounter rate. %but is invisible to the selective forces acting on the scavenger. 
 Aspects including, primary productivity, relief, and temperature will all affect scavenging tendency.
+
 \cite{ruxton2004obligate} suggest an historic ecosystem with a productivity similar to the Serengeti could have supported an \textit{obligate} mammalian or reptilian terrestrial scavenger.
@@
 Perhaps the greatest environmental driver of scavenging tendency is that of temperature. 
-The geological record shows the Earth has undergone radical fluctuations in temperature over time.
-This will have had a significant bearing on the availability and persistence of carrion.
+his has a significant bearing on the availability and persistence of carrion.
+We know from the geological record shows that the Earth has undergone radical fluctuations in temperature over time.
+\textcolor{blue}{TOn land, there are a wide variety of vegetation types, from thickly forested areas to open grassland and the extent of this vegetation changes with season.}
+All of these factors influence the availabiltiy of carrion.  
+Moreover, oceanic productivity is also impacted by climactic conditions.
+
 To illustrate the point, a 10$^{\circ}$C increase in ambient temperature can double carcass decomposition rates \citep{parmenter2009carrion} and geological evidence indicates that the Mesozoic Earth was on average at least 6 $^{\circ}$C warmer than now \citep{sellwood2006mesozoic}.
 In terms of specific habitats, it has been shown that decomposition is greater in warm and moist areas versus more xeric ones \citep{beasley2015vertebrates}.
-Moreover, oceanic productivity and habitat structure are all impacted by climactic conditions.
 The impacts these can have on scavengers have been empirically supported e.g.
 \cite{beasley2015vertebrates} who point to a series of studies showing how microbes and invertebrates benefit at higher temperatures to the detriment of vertebrate scavengers such that ``above 20$^{\circ}$C vertebrates were able to detect and consume only 19\% of small-mammal carcasses, whereas at temperatures below 18$^{\circ}$C, vertebrates consumed 49\% of carcasses".
+\textcolor{blue}{In savannah, dense vegetation can prevent vultures from landing to a carcass because they will be unable to take off again \citep{bamford2009effect}. 
+The Earth has also undergone a series of ice ages of various spatio-temporal extent \citep{diedrich2012cave} and modern terrestrial settings that experience sub-zero conditions can act as a microcosm to show the effect extreme cold can have on scavenging.
+Notably, freezing carcasses can become too hard to consume by most vertebrate carnviores \citep{selva2003scavenging}.}
+
+\subsection{\textcolor{blue}{Biotic effects on Carcass} Availability}
+Through facilitation e.g. local enhancement, adept scavengers may increase the amount of carrion mammalian carnivores feed on thus increasing their population size \citep{moleon2014inter}. 
+This higher population of predators will take more prey and so produce more carcasses.
+Conversely, by feeding on carcasses, predators may hunt less because they are sated by carrion, ultimately reducing predation risk on their prey base \citep{moleon2014inter}.
 
@@
 Moreover, some researchers have charted an evolutionary course from basal fossorial snakes to modern terrestrial species by way of an obligate scavenger intermediate \citep{bauchot2006snakes}. 
-\textcolor{blue}{In water, by contrast, competition with micro-organims is significantly reduced because of the high pressures and low temperatures \citep{beasley2012carrion}.}
+\textcolor{blue}{In water, by contrast, competition with micro-organisms is significantly reduced because of the high pressures and low temperatures \citep{beasley2012carrion}.}
+
+\subsection{Facilitation}
+As well as competitive interactions, there are also facilitatory processes at play between vertebrate scavengers that can reduce handling time for the facilitated species. 
+Some examples include vultures, hyenas and wolves opening the tough hides of ungulate carcasses that would otherwise be inaccessible to corvids and smaller mammalian carnivores \citep{selva2003scavenging,moleon2015carcass}.
 
@@
 
-Again, we can draw a comparison with clades that are lacking in these features. 
+Again, we can draw a comparison with species that are lacking in these features. 
+\textcolor{blue}{Despite readily eating carrion in captivity, cheetahs (\textit{Acinonyx jubatus}) and African Wild Dogs (\textit{Lycaon pictus}) rarely do so in the wild because they are subordinate to many of the mammalian competitors they coexist with \citep{pereira2014facultative}.}
 Extant bats are poorly equipped when it comes to feeding on carrion; the larger forms are typically frugivores and therefore lack the adaptations for digesting meat, while the smaller carnivorous bats are mainly found in the microbats, which are insectivorous \citep{aguirre2003implications}.
@@
 \end{document}
+
+% Modern scavenging assemblages can be inflenced by carcass size \citep{moleon2015carcass}. 
+% Larger carcasses tend to last longer and also present a more conspicuous target for a foraging scavenger which results in more species attending them \citep{moleon2015carcass}.
+% This will have had implications for extinct assemblages because body size distribtions vary across different environments but also across time. 
+% \cite{10.1371/journal.pone.0051925} showed Mesozoic faunal distributions may have been skewed towards larger species even when fossil biases are taken into account. 
+% If this pattern is real then the scavenger assemblages during this era would have been particularly diverse.
+
+% As well as competitive interactions, there are also facilitatory processes at play between vertebrate scavengers that can reduce handling time. 
+% Rather than there being a random assortment of species at a carcass, scavenging assemblages tend to be nested \citep{selva2007nested}. 
+% This means the species that feed on the majority of carcasses, for example vultures, are a subset of a larger community which comprises hyenas, jackals, raptors etc. \citep{sebastian2016nested}. 
+% In extant systems these interactions can reduce handling times for the facilitated species. 
+% Some examples include vultures, hyenas and wolves opening the tough hides of ungulate carcasses that would otherwise be inaccessible to corvids and smaller mammalian carnivores \citep{selva2003scavenging,moleon2015carcass}.
+% The availability of carcasses can be indirectly influenced by the action of scavengers. 
+% Through facilitation e.g. local enhancement, vultures may increase the amount of carrion mammalian carnivores feed on thus increasing their population size \citep{moleon2014inter}. 
+% This higher population of predators will take more prey and so produce more carcasses.
+Conversely, by feeding on carcasses, predators may hunt less because they are sated by carrion, ultimately reducing predation risk on their prey base \citep{moleon2014inter}.

### 2016-07-14 response_letter.tex
Sub sectioning carcasse availability + adding the commented bits to Facilitation part

@@
 
-\textcolor{blue}{As we note below, we feel our earlier manuscript did not capture the hierarchy we had in mind.
-This updated version is more explicit in terms of the structure where we now highlight the sections and subsections.
+\textcolor{blue}{As we note below, we feel our earlier manuscript did not capture the hierarchy we had in mind.This updated version is more explicit in terms of the structure where we now highlight the sections and subsections.
 Regarding the novelty of our manuscript, we have changed our introduction which was regarded as too similar to deValult et al.

### 2016-07-14 natural history of scavenging.tex
Sub sectioning carcasse availability + adding the commented bits to Facilitation part

@@
 
-\subsection{\textcolor{blue}{Abiotic effects on Carcass} Availability}
+\subsection{Carcass Availability}
+\subsubsection{\textcolor{blue}{Abiotic effects}}
 The environmental influence on \textcolor{blue}{carcass} availability is an aspect that greatly affects encounter rate. %but is invisible to the selective forces acting on the scavenger. 
@@
 We know from the geological record shows that the Earth has undergone radical fluctuations in temperature over time.
-\textcolor{blue}{TOn land, there are a wide variety of vegetation types, from thickly forested areas to open grassland and the extent of this vegetation changes with season.}
+\textcolor{blue}{On land, there are a wide variety of vegetation types, from thickly forested areas to open grassland and the extent of this vegetation changes with season.}
 All of these factors influence the availabiltiy of carrion.  
@@
 
-\subsection{\textcolor{blue}{Biotic effects on Carcass} Availability}
+\subsubsection{\textcolor{blue}{Biotic effects}}
 Through facilitation e.g. local enhancement, adept scavengers may increase the amount of carrion mammalian carnivores feed on thus increasing their population size \citep{moleon2014inter}. 
@@
 Conversely, by feeding on carcasses, predators may hunt less because they are sated by carrion, ultimately reducing predation risk on their prey base \citep{moleon2014inter}.
+
+\textcolor{blue}{Modern scavenging assemblages can be influenced by carcass size \citep{moleon2015carcass}. 
+Larger carcasses tend to last longer and also present a more conspicuous target for a foraging scavenger which results in more species attending them \citep{moleon2015carcass}.
+This will have had implications for extinct assemblages because body size distributions vary across different environments but also across time. 
+\cite{10.1371/journal.pone.0051925} showed Mesozoic faunal distributions may have been skewed towards larger species even when fossil biases are taken into account. 
+If this pattern is real then the scavenger assemblages during this era would have been particularly diverse.}
 
@@
 
-\subsection{Facilitation}
+\textcolor{blue}{\subsection{Facilitation}
 As well as competitive interactions, there are also facilitatory processes at play between vertebrate scavengers that can reduce handling time for the facilitated species. 
+Rather than there being a random assortment of species at a carcass, scavenging assemblages tend to be nested \citep{selva2007nested}. 
+This means the species that feed on the majority of carcasses, for example vultures, are a subset of a larger community which comprises hyenas, jackals, raptors etc. \citep{sebastian2016nested}. 
+In extant systems these interactions can reduce handling times for the facilitated species. 
 Some examples include vultures, hyenas and wolves opening the tough hides of ungulate carcasses that would otherwise be inaccessible to corvids and smaller mammalian carnivores \citep{selva2003scavenging,moleon2015carcass}.
+The availability of carcasses can be indirectly influenced by the action of scavengers. 
+Through facilitation e.g. local enhancement, vultures may increase the amount of carrion mammalian carnivores feed on thus increasing their population size \citep{moleon2014inter}. 
+This higher population of predators will take more prey and so produce more carcasses.
+Conversely, by feeding on carcasses, predators may hunt less because they are sated by carrion, ultimately reducing predation risk on their prey base \citep{moleon2014inter}.}
 
@@
 
-% Modern scavenging assemblages can be inflenced by carcass size \citep{moleon2015carcass}. 
+% Modern scavenging assemblages can be influenced by carcass size \citep{moleon2015carcass}. 
 % Larger carcasses tend to last longer and also present a more conspicuous target for a foraging scavenger which results in more species attending them \citep{moleon2015carcass}.
-% This will have had implications for extinct assemblages because body size distribtions vary across different environments but also across time. 
+% This will have had implications for extinct assemblages because body size distributions vary across different environments but also across time. 
 % \cite{10.1371/journal.pone.0051925} showed Mesozoic faunal distributions may have been skewed towards larger species even when fossil biases are taken into account. 

### 2016-07-15 natural history of scavenging.tex
whale reference added

@@
 Carcasses also tend to sink in water where they are no longer accessible to pelagic scavengers \citep{beasley2012carrion}. 
-Research has shown that an animal need only travel between 16 to 36 km to encounter a fresh whale carcass \citep{moleon2015living}. 
+\textcolor{blue}{Research has shown that an animal need only travel 36 km to encounter a fresh whale carcass \citep{smith2003ecology}.} 
 The phenomenon of occasional bounties of carrion in the form of these whale falls has led some researchers to investigate if an obligate scavenger could survive by seeking out these remains exclusively.}
@@
 Perhaps the greatest environmental driver of scavenging tendency is that of temperature. 
-his has a significant bearing on the availability and persistence of carrion.
+This has a significant bearing on the availability and persistence of carrion.
 We know from the geological record shows that the Earth has undergone radical fluctuations in temperature over time.
@@
 Conversely, by feeding on carcasses, predators may hunt less because they are sated by carrion, ultimately reducing predation risk on their prey base \citep{moleon2014inter}.
+
+%Carcass size shapes the structure and functioning of an African scavenging assemblage
+%Moleon et al
+%We found strong evidence indicating that carcass size is a major factor governing the associated scavenger assemblage.
+Scavenger species richness per carcass and carcass consumption time and rate increased with carcass size, while carcass detection time and percentage of carrion biomass consumed were negatively related to carcass size.
+%During the dry season, less vegetation cover might also favor the detection of carrion by scavengers, especially mammals (Ruzicka and Conover 2012).
+%The mean number of scavenger species exploiting a particular carcass increased with carcass size, probably because the longer temporal availability and higher conspicuousness of larger carcasses (DeVault et al.
+2004) allowed more species to aggregate.
+%This suggests that hyenas were involved in some interactions that enhance community structure, such as carcass signaling and tearing open of thick-skinned animals (Moleón et al. 2014a).

### 2016-07-19 response_letter.tex
updates on ms and letter

@@
 \item{Abstract: Scavenging can be difficult to observe, but probably no more so than predation. Also, using experimental carcasses, scavenging is actually quite easy to observe.}
+
+\textcolor{blue}{We emphasise that food webs have tended to lump all carnivory into predation and it is with extinct species that we have no observational evidence.}
+
 \item{Abstract: You should point out that your ``scale of scavenging'' is only qualitative; you made no attempt to quantify such traits.}
@@
 
+\textcolor{blue} {We make the following point in our introduction: \textit{Indeed, there is no discrete divide between predators and scavengers but rather a continuous gradient.
+Even vultures, the canonical example of obligate scavengers, can hunt.} We have tried to be clearer throughout the paper as well.}
+
 %TG: maybe say something along the lines:
-\textcolor{blue}{We do not consider a binary response to scavenging behavior (i.e.
-facultative/obligate) and argue that all carnivorous vertebrates are placed along a gradient of scavenging.
-We state this more clearly in the introduction: \textit{``there is no discrete divide between predators and scavengers but rather a continuous gradient (Pereira et al.
-2014)''}.}
+%\textcolor{blue}{As with the above comment we argue that all carnivorous vertebrates are placed along a gradient of scavenging.}
 
@@
 
+\textcolor{blue} {This was also raised by reviewer 1.
+We make the following point in our introduction: \textit{Indeed, there is no discrete divide between predators and scavengers but rather a continuous gradient.
+Even vultures, the canonical example of obligate scavengers, can hunt.} We have tried to be clearer throughout the paper as well.
+}
+
 They also seem to ignore flexibility in species' behaviors.
@@
 
-\textcolor{blue}{It is certainly true that facultative scavengers can be variable in the proportion of carrion they consume but we would argue that species like hyenas have certain characteristics that make this possible.
+\textcolor{blue}{It is certainly true that scavengers can be variable in the proportion of carrion they consume but we would argue that species like hyenas have certain characteristics that make this possible.
 In contrast, cheetahs do not have a biology that lends itself to scavenging in the wild, due to poor competitive ability etc.
-We appreciate that there is a continuum across and among species which we specifically note in this revised version.}
+We appreciate that there is a continuum across and among species which we specifically note in this revised version.
+Moreover, researchers working in palaeobiology often try to reconstruct the behaviours of extinct species; we are following this approach when we make inferences about ancient systems.}
 \smallskip
@@
 They might look to articles about other modern scavenging opportunities such as Capaldo and Peter's study of wildebeest drownings \url{http://www.sciencedirect.com/science/article/pii/S0305440385700398}, and suggestions that sabertooth cats which they claim were unable to feed on bones may have consumed more bone than penecontemporaneous dire wolves \url{https://www.jstor.org/stable/20627167?seq=1#page_scan_tab_contents}.
-
+\\
+\smallskip
+\textcolor{blue}{The issue of the structure of our paper was noted by all reviewers who all had suggestions to improve it. 
+We have made an effort to delineate all of the main sections and subsections so that our frame of optimal foraging is more apparent.}
+\smallskip
+\\
 In sum, while I think the aim of the article is a very worthwhile one, I think the authors cannot claim to have produced an all-encompassing model that can predict the presence or level of scavenging in any organism - because I think this simply cannot be done.
@@
 \textcolor{blue}{The sections of Beasley et al.
-(2015) are the evolution of scavengers, competition between scavengers and environmental effects on scavenger ecology.
+(2015) are:  1.
+the evolution of scavengers, 2.
+competition between scavengers and 3.
+environmental effects on scavenger ecology.
 These are very broad areas so in a review of this sort some overlap is inevitable.
-We feel our emphasis on extinct species and systems as well as the frame of optimal foraging theory gives this manuscript its novelty.}
+We feel our inclusion of extinct species and systems as well as the frame of optimal foraging theory gives this manuscript its novelty.}
 \smallskip

### 2016-07-19 natural history of scavenging.tex
updates on ms and letter

@@
 The result is a synthesis on the natural history of scavenging. 
-Our idea of a scale of scavenging can be applied to any species at any time to help \textcolor{blue}{judge} the importance of this behaviour in its diet. 
+These features that make up our scale of scavenging can be applied to any given species and allow us to judge the likely importance of this foraging behaviour. 
 \end{abstract}
@@
 This is clearly a mistaken view because of the existence of scavenging, a behaviour displayed by almost all carnivorous vertebrates \citep{devault2003scavenging}. 
-Recent research has begun to redress this imbalance by recognising the prevalence of scavenging and detailing its ecological consequences across modern ecosystems where we are beginning to appreciate the implications for the trophic ecology of these systems \citep{pereira2014facultative,periquet2015lion}.
-Indeed, there is no discrete divide between predators and scavengers but rather a continuous gradient \citep{pereira2014facultative}.
-Unfortunately, the lack of direct behavioural data makes it difficult to discern scavenging from predation among extinct forms where it was surely exhibited to a similar extent among prehistoric vertebrate carnivores.
+Recent research has begun to redress this imbalance by recognising the prevalence of scavenging and detailing its ecological consequences across modern ecosystems such that are beginning to appreciate the implications for trophic ecology \citep{pereira2014facultative,periquet2015lion}.
+Indeed, there is no discrete divide between predators and scavengers but rather a continuous gradient in terms of the proportion of carrion in the diet \citep{pereira2014facultative}.
+Even vultures, the canonical example of obligate scavengers, can hunt \citep{margalida2011scavenger}. 
+Unfortunately, the lack of direct behavioural data makes it difficult to discern scavenging from predation among extinct forms where it was surely exhibited by prehistoric vertebrate carnivores.
 Some simple heuristics can inform us whether or not the prey was scavenged, for instance, in cases where it was simply too large to have been killed by the ostensible predator \citep{pobiner2008paleoecological}.}
 But clearly, a scavenger does not only feed on animals too big for it to have hunted.
-\textcolor{blue}{There are other methods however, that allow us to discern the most suitable morphologies, physiologies and environments for a scavenging lifestyle to prosper e.g.
+
+\textcolor{blue}{There are other methods however, that can allow us to discern the most suitable morphologies, physiologies and environments for a scavenging lifestyle to prosper e.g.
 energetics models, comparative anatomy, palaeontology etc.
@@
 Yet, a synthesis describing the natural history of scavengers is absent from the literature.
-Here we rectify this shortcoming in charting the natural history of scavenging by assessing its potential among extinct vertebrate groups given their ecology, functional traits and the environmental context.}
+Here we rectify this shortcoming in charting the natural history of scavenging by assessing its potential among vertebrate groups past and present given their ecology, functional traits and the environmental context.}
 
@@
 By considering scavenging in this context of optimal foraging we can identify the prerequisite attributes and processes required for the behaviour. 
-This has enabled us to propose a scale of scavenging whereupon we can place \textcolor{blue}{an extinct vertebrate carnivore} and assess the likely importance of carrion in its diet.
+This has enabled us to propose a scale of scavenging whereupon we can place \textcolor{blue}{a given vertebrate carnivore} and assess the likely importance of carrion in its diet.
 
@@
 Locomotory speed, foraging time and detection radius all determine the encounter rate between a scavenger and the carcasses it is searching for and we would thus expect selection pressures to act on the various traits that govern these parameters.
-Selection might also drive a species to reduce its metabolic requirements so that it can survive long periods between meals, or expand a species ability to process tissues or decaying flesh that might ordinarily be discarded.
+Selection might also drive a species to reduce its metabolic requirements so that it can survive long periods between meals, or expand a species' ability to process tissues or decaying flesh that might ordinarily be discarded.
 However, as we noted above, encounter rate is also determined by the productivity of the environment, and ultimately the rate at which carcasses are produced. 
-\textcolor{blue}{Scavengers can drive this process indirectly, for instance, by consuming the majority of carrion, obligate scavengers deny a resource to their facultative counterparts obliging the latter to hunt more and thus produce more carrion \citep{moleon2014inter}.}
+\textcolor{blue}{Scavengers can drive this process indirectly, for instance, by consuming the majority of carrion, more obligate scavengers deny a resource to their facultative counterparts obliging the latter to hunt more, and thus, increase the amount of carrion \citep{moleon2014inter}.}
 
 \subsection{Metabolism}
-Because of the sporadic nature of carrion we would expect adaptations that reduce energetic costs of maintenance to be selected for in scavengers as it would allow for longer inter-feeding periods. 
+Because of the ephemeral nature of carrion \citep{devault2003scavenging,ruxton2004obligate} we would expect adaptations that reduce energetic costs of maintenance to be selected for in scavengers as it would allow for longer inter-feeding periods.
 Extant reptiles possess an advantage here, in that over the course of a year their food requirements can be 30 times lower than an endotherm of equal size \citep{Nagy1621}.
@@
 \subsection{Locomotion}
-The \textcolor{blue}{relative} unpredictability of carrion also means scavenging depends more on the ability to efficiently move over larger areas than does predation.
+Carrion can often be unpredictable, such as in cases where animals succumb to disease \citep{devault2003scavenging}, which means scavenging depends more on the ability to efficiently move over larger areas than does predation.
 This generally requires an efficient transfer of metabolic energy into movement which relies on the animal's anatomy and physiology as well as the medium of the environment in which the animal is moving (i.e.
@@
 Perhaps the most efficient form of locomotion in vertebrates is found in flying species. 
-Despite the energetic costs of flight, the only known vertebrate obligate scavengers are the old and the new world vultures. 
-While powered flight is energetically expensive, species like vultures have evolved behavioural and anatomical features to exploit air currents using their large wingspans, allowing them to soar at a cost of only twice their metabolic rate \citep{hedenstrom1993migration,spivey2014analysing}.
+Despite the energetic costs of flight, the only known vertebrates that can be obligate scavengers are the old and the new world vultures. 
+While powered flight is energetically expensive, species like vultures have evolved behavioural and anatomical features to exploit air currents using their large wingspans, allowing them to soar at a cost of only around 1.5 times their metabolic rate \citep{hedenstrom1993migration,duriez2014cheap}.
 By depending on thermal air flows these species can forage over vast ranges \citep{spiegel2013factors}. 
@@
 
-Existing in this environment also means being able to gather a wealth of information from other foragers, be they conspecifics or otherwise \textcolor{blue}{\citep{jackson2008effect,KaneVul,moleon2014inter}}.
-These interactions are properly three dimensional in the sense of \cite{pawar2012dimensionality} because both producer and scrounger are in the air \citep{dall2005information}. 
-Again, returning to vultures, the genus \textit{Gyps} consists of highly social and colonially nesting species \citep{fernandez2015density}.
-These behaviours allow them to forage far more efficiently because one bird can scrounge information on the location of food from another successful forager \textcolor{blue}{\citep{cortes2014bird}}.
-Information transfer of this kind is typically inadvertent and as a consequence no complex social interactions are required, simply the ability to recognise a successful forager.
-\textcolor{blue}{Thus, given pterosaurs seem to have cohabited in large numbers \citep{witton2013pterosaurs}, and the known benefits this can have for social foraging in birds \citep{jackson2011evolutionary}, it seems probable that scrounging behaviours were seen in the flying pterosaurs as well.}
-
 The terrestrial-olfaction, aerial-visual divide is not total though.
@@
 \subsubsection{\textcolor{blue}{Biotic effects}}
-Through facilitation e.g. local enhancement, adept scavengers may increase the amount of carrion mammalian carnivores feed on thus increasing their population size \citep{moleon2014inter}. 
-This higher population of predators will take more prey and so produce more carcasses.
+Foragers do not exist in isolation and we know from field observations that scavengers can scrounge on the discoveries of other carnivores. 
+This sort of facilitation occurs in both obligate and facultative scavengers, in the air and on the ground \citep{KaneVul,jones2015african}.
+
+In flight, birds are able to gather a wealth of information from other foragers, be they conspecifics or otherwise \textcolor{blue}{\citep{jackson2008effect,KaneVul,moleon2014inter}}.
+% These interactions are properly three dimensional in the sense of \cite{pawar2012dimensionality} because both producer and scrounger are in the air \citep{dall2005information}. 
+Again, returning to vultures, the genus \textit{Gyps} consists of highly social and colonially nesting species \citep{fernandez2015density}.
+These behaviours allow them to forage far more efficiently because one bird can scrounge information on the location of food from another successful forager \textcolor{blue}{\citep{cortes2014bird}}.
+Information transfer of this kind is typically inadvertent and as a consequence no complex social interactions are required, simply the ability to recognise a successful forager.
+\textcolor{blue}{Thus, given pterosaurs seem to have cohabited in large numbers \citep{witton2013pterosaurs}, and the known benefits this can have for social foraging in birds \citep{jackson2011evolutionary}, it seems probable that scrounging behaviours were seen in the flying pterosaurs as well.}
+
+This type of facilitation then increases the encounter rate of the facilitated species and can increase the population size of the latter \citep{moleon2014inter}. 
+This higher population of predators may take more prey and so produce more carcasses.
 Conversely, by feeding on carcasses, predators may hunt less because they are sated by carrion, ultimately reducing predation risk on their prey base \citep{moleon2014inter}.
 
-\textcolor{blue}{Modern scavenging assemblages can be influenced by carcass size \citep{moleon2015carcass}. 
+\textcolor{blue}{Modern scavenging assemblages are known to be influenced by carcass size \citep{moleon2015carcass}. 
 Larger carcasses tend to last longer and also present a more conspicuous target for a foraging scavenger which results in more species attending them \citep{moleon2015carcass}.
@@
 \cite{10.1371/journal.pone.0051925} showed Mesozoic faunal distributions may have been skewed towards larger species even when fossil biases are taken into account. 
-If this pattern is real then the scavenger assemblages during this era would have been particularly diverse.}
+Similarly, the megafauna of the Pleistocene \citep{doughty2013legacy} would have produced large carcasses. 
+As a result, the scavenger assemblages during this era would have been particularly diverse.}
 
@@
 \textcolor{blue}{\subsection{Facilitation}
-As well as competitive interactions, there are also facilitatory processes at play between vertebrate scavengers that can reduce handling time for the facilitated species. 
+In contrast to competitive interactions, there are facilitatory processes at play between vertebrate scavengers that can benefit the facilitated species. 
 Rather than there being a random assortment of species at a carcass, scavenging assemblages tend to be nested \citep{selva2007nested}. 
 This means the species that feed on the majority of carcasses, for example vultures, are a subset of a larger community which comprises hyenas, jackals, raptors etc. \citep{sebastian2016nested}. 
-In extant systems these interactions can reduce handling times for the facilitated species. 
-Some examples include vultures, hyenas and wolves opening the tough hides of ungulate carcasses that would otherwise be inaccessible to corvids and smaller mammalian carnivores \citep{selva2003scavenging,moleon2015carcass}.
-The availability of carcasses can be indirectly influenced by the action of scavengers. 
-Through facilitation e.g. local enhancement, vultures may increase the amount of carrion mammalian carnivores feed on thus increasing their population size \citep{moleon2014inter}. 
-This higher population of predators will take more prey and so produce more carcasses.
-Conversely, by feeding on carcasses, predators may hunt less because they are sated by carrion, ultimately reducing predation risk on their prey base \citep{moleon2014inter}.}
+We noted above that this often results in an increased encounter rate but it can also reduce handling time for the species that follow. 
+% In extant systems these interactions can reduce handling times for the facilitated species. 
+Some examples of reduced handling time include vultures, hyenas and wolves opening the tough hides of ungulate carcasses that would otherwise be inaccessible to corvids and smaller mammalian carnivores \citep{selva2003scavenging,moleon2015carcass}.}
 
 \subsection{Food Processing}
-Aside from coping with competitors, another vital component of carrion handling time is the ability to maximise the energy gain from the remains while reducing the energetics of doing so. 
+Another vital component of carrion handling time is the ability to maximise the energy gain from the remains while reducing the energetics of doing so. 
 At whale carcases, white and blue sharks are known to preferentially feed on the blubber layer \citep{long1996sharks}. 

### 2016-07-20 natural history of scavenging.tex
Kevin updates
Kevin updates

@@
 This is clearly a mistaken view because of the existence of scavenging, a behaviour displayed by almost all carnivorous vertebrates \citep{devault2003scavenging}. 
-Recent research has begun to redress this imbalance by recognising the prevalence of scavenging and detailing its ecological consequences across modern ecosystems such that are beginning to appreciate the implications for trophic ecology \citep{pereira2014facultative,periquet2015lion}.
+Recent research has begun to redress this imbalance by recognising the prevalence of scavenging and its implications for trophic ecology through detailing the ecological consequences of scavanging across modern ecosystems \citep{pereira2014facultative,periquet2015lion}.
 Indeed, there is no discrete divide between predators and scavengers but rather a continuous gradient in terms of the proportion of carrion in the diet \citep{pereira2014facultative}.
 Even vultures, the canonical example of obligate scavengers, can hunt \citep{margalida2011scavenger}.
-Unfortunately, the lack of direct behavioural data makes it difficult to discern scavenging from predation among extinct forms where it was surely exhibited by prehistoric vertebrate carnivores.
+While it is relatively easy to determine whether a species engages in scavenging through observation, determining proportions requires extensive behavioral data (I put this in to placate the reviewer who said it was easy).
+Unfortunately, the lack of direct behavioural data makes it even more difficult to discern scavenging from predation among extinct forms where it was surely exhibited by prehistoric vertebrate carnivores.
 Some simple heuristics can inform us whether or not the prey was scavenged, for instance, in cases where it was simply too large to have been killed by the ostensible predator \citep{pobiner2008paleoecological}.}
@@
 However, as we noted above, encounter rate is also determined by the productivity of the environment, and ultimately the rate at which carcasses are produced. 
-\textcolor{blue}{Scavengers can drive this process indirectly, for instance, by consuming the majority of carrion, more obligate scavengers deny a resource to their facultative counterparts obliging the latter to hunt more, and thus, increase the amount of carrion \citep{moleon2014inter}.}
+\textcolor{blue}{Carcass production itself depends on factors such as predation, disease and scavenging itself.
+For instance, by consuming the majority of carrion, more obligate scavengers deny a resource to their facultative counterparts obliging the latter to hunt more, and thus, increase the amount of carrion \citep{moleon2014inter}.}
 
@@
 vultures who do so while resting at their roost and during periods of food deprivation \citep{bahat1998nocturnal}.
+
+\textcolor{blue}{While it is difficult to infer the metabolisms of extinct species, similarities can still also be drawn.
+Mainly ectothermic groups such as sharks and bony fish were likely to benefit from the ability to bridge long time-spans between meals such as seen in snakes.
+In contrast, forms such as dinosaurs \citep{grady2014evidence}, pterosuars \citep{wellnhofer1991illustrated}, some marine reptiles \citep{bernard2010regulation} and some early mammals \citep{lovegrove2016phenology} have been proposed to have at least partial endothermy resulting in requiring a higher intake of material in comparison to their ectothermic equivalents.
+However, such higher metabolic rates would also increased the ability of such species to actively search for carrion.} \textcolor{red}{I put this section in as there is no reference in how we can use the above paragraph to infer scavenging in extinct species}
 
@@
 Although the evolution of bipedal movement was significant in that it freed up the forelimbs for other purposes (e.g.
-climbing, tool-use, wing development etc.) it does not differ radically in cost from quadrupedal locomotion (\cite{williams1999evolution}, and references therein).
+climbing, tool-use, wing development etc.) it does not differ radically in cost from quadrupedal locomotion (\cite{williams1999evolution}, \textcolor{blue}{and references therein}).
 For instance, \cite{alexander2004bipedal} shows that, in the case of humans, we are more economical than predicted while walking and less so while running according to predicted costs of terrestrial movement calculated by allometric scaling relationships.
@@
 While the locomotory ability of theropods has been debated since their first inception, more recent studies have reconstructed them as relatively mobile animals \citep{pontzer2009biomechanics}. 
-Despite some suggestions that larger species may have had some advantage in scavenging, partially due to the ability to search large areas \citep{ruxton2003could}, more recent work has shown that the energetic demands of such large forms meant scavenging was likely more prevalent in mid-sized theropods of approximately half a tonne \citep{kane2016body}.
+Despite some suggestions that larger species may have had some advantage in scavenging, partially due to the ability to search large areas \citep{ruxton2003could}, more recent work has shown that the energetic demands of locomotion in such large forms meant scavenging was likely more prevalent in mid-sized theropods of approximately half a tonne \citep{kane2016body}.
 
@@
 Although no aquatic species have ever exceeded the size of whales, some enormous animals have evolved in this environment before the evolution of cetaceans, including \textit{Leedsichthys}, a bony fish from the Middle Jurassic (174.1 - 163.5 Mya) and the aquatic Mesozoic reptiles, the plesiosaurs, pliosaurs and ichtyosaurs, that could all exceed 15 metres in length \citep{ruxton2011zoology,danise2014ecological}.
-So, despite being unlikely, the energetic feasibility of a marine scavenger that specialises on large carcasses has a long history.
+So, despite being unlikely \textcolor{red}{Is there an estimate of how much of the fishes diet can be made up from this type of scavenging?}, the energetic feasibility of a marine scavenger that specialises on large carcasses has a long history.
 
@@
 This has a significant bearing on the availability and persistence of carrion.
-We know from the geological record shows that the Earth has undergone radical fluctuations in temperature over time.
+We know from the geological record that the Earth has undergone radical fluctuations in temperature over time.
 \textcolor{blue}{On land, there are a wide variety of vegetation types, from thickly forested areas to open grassland and the extent of this vegetation changes with season.}
 All of these factors influence the availabiltiy of carrion.  
-Moreover, oceanic productivity is also impacted by climactic conditions.
+Moreover, oceanic productivity is also impacted by climactic conditions. \textcolor{red}{I would delete the last two lines of this paragraph and merge it with the next one}
 
@@
 \cite{beasley2015vertebrates} who point to a series of studies showing how microbes and invertebrates benefit at higher temperatures to the detriment of vertebrate scavengers such that ``above 20$^{\circ}$C vertebrates were able to detect and consume only 19\% of small-mammal carcasses, whereas at temperatures below 18$^{\circ}$C, vertebrates consumed 49\% of carcasses".
-\textcolor{blue}{In savannah, dense vegetation can prevent vultures from landing to a carcass because they will be unable to take off again \citep{bamford2009effect}. 
+\textcolor{blue}{In savannah, dense vegetation can prevent vultures from landing to a carcass because they will be unable to take off again \citep{bamford2009effect} \textcolor{red}{This is a biotic interaction so I would either remove it or move it somewhere else}.
 The Earth has also undergone a series of ice ages of various spatio-temporal extent \citep{diedrich2012cave} and modern terrestrial settings that experience sub-zero conditions can act as a microcosm to show the effect extreme cold can have on scavenging.
@@
 Moreover, some researchers have charted an evolutionary course from basal fossorial snakes to modern terrestrial species by way of an obligate scavenger intermediate \citep{bauchot2006snakes}.
+However, snake venom is primary associated with predation in extant species suggesting that if venom played such a role in extinct species it is no longer its main function \citep{casewell2013complex}.
 \textcolor{blue}{In water, by contrast, competition with micro-organisms is significantly reduced because of the high pressures and low temperatures \citep{beasley2012carrion}.}

### 2016-07-29 natural history of scavenging.tex
Commented out some of Kev's red comments
They’re now using the %KH: format

@@
 In contrast, forms such as dinosaurs \citep{grady2014evidence}, pterosuars \citep{wellnhofer1991illustrated}, some marine reptiles \citep{bernard2010regulation} and some early mammals \citep{lovegrove2016phenology} have been proposed to have at least partial endothermy resulting in requiring a higher intake of material in comparison to their ectothermic equivalents.
-However, such higher metabolic rates would also increased the ability of such species to actively search for carrion.} \textcolor{red}{I put this section in as there is no reference in how we can use the above paragraph to infer scavenging in extinct species}
+However, such higher metabolic rates would also increased the ability of such species to actively search for carrion.}
+%KH: I put this section in as there is no reference in how we can use the above paragraph to infer scavenging in extinct species
 
@@
 Although no aquatic species have ever exceeded the size of whales, some enormous animals have evolved in this environment before the evolution of cetaceans, including \textit{Leedsichthys}, a bony fish from the Middle Jurassic (174.1 - 163.5 Mya) and the aquatic Mesozoic reptiles, the plesiosaurs, pliosaurs and ichtyosaurs, that could all exceed 15 metres in length \citep{ruxton2011zoology,danise2014ecological}.
-So, despite being unlikely \textcolor{red}{Is there an estimate of how much of the fishes diet can be made up from this type of scavenging?}, the energetic feasibility of a marine scavenger that specialises on large carcasses has a long history.
+So, despite being unlikely 
+% KH: Is there an estimate of how much of the fishes diet can be made up from this type of scavenging?
+, the energetic feasibility of a marine scavenger that specialises on large carcasses has a long history.
 
@@
 All of these factors influence the availabiltiy of carrion.  
-Moreover, oceanic productivity is also impacted by climactic conditions. \textcolor{red}{I would delete the last two lines of this paragraph and merge it with the next one}
+Moreover, oceanic productivity is also impacted by climactic conditions. 
+%KH: I would delete the last two lines of this paragraph and merge it with the next one
 
@@
 \cite{beasley2015vertebrates} who point to a series of studies showing how microbes and invertebrates benefit at higher temperatures to the detriment of vertebrate scavengers such that ``above 20$^{\circ}$C vertebrates were able to detect and consume only 19\% of small-mammal carcasses, whereas at temperatures below 18$^{\circ}$C, vertebrates consumed 49\% of carcasses".
-\textcolor{blue}{In savannah, dense vegetation can prevent vultures from landing to a carcass because they will be unable to take off again \citep{bamford2009effect} \textcolor{red}{This is a biotic interaction so I would either remove it or move it somewhere else}.
+\textcolor{blue}{In savannah, dense vegetation can prevent vultures from landing to a carcass because they will be unable to take off again \citep{bamford2009effect}.
+%KH: This is a biotic interaction so I would either remove it or move it somewhere else. 
 The Earth has also undergone a series of ice ages of various spatio-temporal extent \citep{diedrich2012cave} and modern terrestrial settings that experience sub-zero conditions can act as a microcosm to show the effect extreme cold can have on scavenging.

### 2016-07-29 response_letter.tex
Dealt with a couple of comments from reviewer 1

@@
 %\begin{letter}{Company name \\ Street\\ City\\ Country}
-To the editor
+Dear Miguel Ara\'{u}jo
 
@@
 
-\textcolor{blue}{As we note below, we feel our earlier manuscript did not capture the hierarchy we had in mind.This updated version is more explicit in terms of the structure where we now highlight the sections and subsections.
+\textcolor{blue}{As we note below, we feel our earlier manuscript did not capture the hierarchy we had in mind.
+This updated version is more explicit in terms of the structure where we now highlight the sections and subsections.
 Regarding the novelty of our manuscript, we have changed our introduction which was regarded as too similar to deValult et al.
@@
 The paper is mostly well written, and most (but not all) pertinent literature sources are included.
-I do worry about the lack of novelty of the manuscript, and what I perceived as haphazard organization.
-Here are some specific issues that should be addressed in my opinion:
+I do worry about the lack of novelty of the manuscript, and what I perceived as haphazard organization. Here are some specific issues that should be addressed in my opinion:
 
 \begin{enumerate}
-\item{Abstract: Scavenging can be difficult to observe, but probably no more so than predation. Also, using experimental carcasses, scavenging is actually quite easy to observe.}
+\item{Abstract: Scavenging can be difficult to observe, but probably no more so than predation.
+Also, using experimental carcasses, scavenging is actually quite easy to observe.}
 
@@
 \item{Abstract: You should point out that your ``scale of scavenging'' is only qualitative; you made no attempt to quantify such traits.}
+
+\textcolor{blue}{We now clearly stated that the scale is qualitative.}
+
 \item{Introduction: Much of this section seems to follow DeVault's (2003) review paper on scavenging—in fact, there is almost nothing new here that is not covered in that earlier review.}
@@
 
-\textcolor{blue} {We make the following point in our introduction: \textit{Indeed, there is no discrete divide between predators and scavengers but rather a continuous gradient.
-Even vultures, the canonical example of obligate scavengers, can hunt.} We have tried to be clearer throughout the paper as well.}
-
-%TG: maybe say something along the lines:
-%\textcolor{blue}{As with the above comment we argue that all carnivorous vertebrates are placed along a gradient of scavenging.}
+\textcolor{blue} {We make the following point in our introduction: \textit{``Indeed, there is no discrete divide between predators and scavengers but rather a continuous gradient.
+Even vultures, the canonical example of obligate scavengers, can hunt.''} We have tried to be clearer throughout the paper as well.}
 
 \item{Last sentence of Introduction: Again, you propose a ``scale of scavenging'', but without quantification, it is unclear how useful this ``scale'' is.}
+
+\textcolor{blue}{We now clarified one of the advantages of such a scale at the end of the introduction: \textit{``We argue that such a scale could be valuable when behaviour can not be observed directly, e.g.
+for extinct vertebrates.''}}
 
@@
 2012).
-Research has shown that an animal need only travel between 16 to 36 km to encounter a fresh whale carcass  (Moleón and Sánchez-Zapata 2015).
+Research has shown that an animal need only travel between 16 to 36 km to encounter a fresh whale carcass  (Mole\'{o}n and Sánchez-Zapata 2015).
 The phenomenon of occasional bounties of carrion in the form of these whale falls has led some researchers to investigate if an obligate scavenger could survive by seeking out these remains exclusively.''}}
@@
 
-%TG: something along the lines: "The jaw is the principal tool for food ingestion and it's morphology will determine and the range of possible food items the organism is able to ingest"
+\textcolor{The jaw is the principal tool for food ingestion and it's morphology will determine and the range of possible food items the organism is able to ingest.}
+%TG: Note that there's no change in the text here. Just a lecture on jaw morphologies. We might want to elaborate a bit in the manuscript.
 
@@
 
-\textcolor{blue}{The hierarchy of our paper wasn't clear in the first version but we have now clarified that this is the lead in to a larger section which has subsections e.g. food processing}
+\textcolor{blue}{The hierarchy of our paper wasn't clear in the first version but we have now clarified that this is the lead in to a larger section which has subsections e.g.
+food processing (see also updated figure 1).}
 

### 2016-07-29 natural history of scavenging.tex
Dealt with a couple of comments from reviewer 1

@@
 The result is a synthesis on the natural history of scavenging. 
-These features that make up our scale of scavenging can be applied to any given species and allow us to judge the likely importance of this foraging behaviour. 
+These features that make up our \textcolor{blue}{qualitative} scale of scavenging can be applied to any given species and allow us to judge the likely importance of this foraging behaviour. 
 \end{abstract}
@@
 By considering scavenging in this context of optimal foraging we can identify the prerequisite attributes and processes required for the behaviour. 
-This has enabled us to propose a scale of scavenging whereupon we can place \textcolor{blue}{a given vertebrate carnivore} and assess the likely importance of carrion in its diet.
+This has enabled us to propose a scale of scavenging whereupon we can place \textcolor{blue}{a given carnivore vertebrate} and assess the likely importance of carrion in its diet.
+\textcolor{blue}{We argue that such a scale could be valuable when behaviour can not be observed directly, e.g. for extinct vertebrates.}
 

### 2016-07-29 response_letter.tex
Comments for reviewer 2

@@
 
-\textcolor{The jaw is the principal tool for food ingestion and it's morphology will determine and the range of possible food items the organism is able to ingest.}
+\textcolor{blue}{The jaw is the principal tool for food ingestion and it's morphology will determine and the range of possible food items the organism is able to ingest.}
 %TG: Note that there's no change in the text here. Just a lecture on jaw morphologies. We might want to elaborate a bit in the manuscript.
@@
 
+%TG: They make a point here (although their diversity definition rant is completely out of topic). Should we just drop figure 2 and use it as an abstract/press release picture?
+
 If the authors chose to resubmit, I would suggest they organize their manuscript based on known information about modern species, known (e.g.
@@
 We have made an effort to delineate all of the main sections and subsections so that our frame of optimal foraging is more apparent.}
+
 \smallskip
-\\
 In sum, while I think the aim of the article is a very worthwhile one, I think the authors cannot claim to have produced an all-encompassing model that can predict the presence or level of scavenging in any organism - because I think this simply cannot be done.
+
+%TG: State somewhere here as a response to the reviewer that we don't develop any model in this paper?
 
@@
 
-\textcolor{blue}{The sections of Beasley et al.
-(2015) are:  1.
-the evolution of scavengers, 2.
-competition between scavengers and 3.
-environmental effects on scavenger ecology.
+\textcolor{blue}{The sections of Beasley et al. (2015) are: 1. the evolution of scavengers, 2. competition between scavengers and 3. environmental effects on scavenger ecology.
 These are very broad areas so in a review of this sort some overlap is inevitable.

### 2016-08-08 response_letter.tex
andrew comments added

@@
 %\signature{Your name}
-%\address{Street \\ City \\ Country}
+%\address{Cooperage Building \\ North Mall Campus \\ University College Cork \\ Cork \\ Ireland}
 \begin{document}
 %\begin{letter}{Company name \\ Street\\ City\\ Country}
-Dear Miguel Ara\'{u}jo
+Dear Miguel Ara\'{u}jo,
 
@@
 
-\textcolor{blue}{We now clearly stated that the scale is qualitative.}
+\textcolor{blue}{We now clearly state that the scale is qualitative.}
 
@@
 
-\textcolor{blue}{We rewrote the introduction to more accurately reflect the aim of this review.} %TG: i.e. the scavenging gradient allows us to infer scavenging in the fossil record?
+\textcolor{blue}{We have rewritten the introduction to more accurately reflect the aim of this review i.e.
+that the scavenging gradient allows us to infer scavenging in the fossil record as well as in modern species.} %TG:
 
@@
 
-\textcolor{blue}{We now clarified one of the advantages of such a scale at the end of the introduction: \textit{``We argue that such a scale could be valuable when behaviour can not be observed directly, e.g.
+\textcolor{blue}{See the point above which highlights that the concept of a scale, and not a false caterogisation, is central to our paper.
+We have now clarified one of the advantages of such a scale at the end of the introduction: \textit{``We argue that such a scale could be valuable when behaviour can not be observed directly, e.g.
 for extinct vertebrates.''}}
@@
 
-\textcolor{blue}{The jaw is the principal tool for food ingestion and it's morphology will determine and the range of possible food items the organism is able to ingest.}
+\textcolor{blue}{We've removed the reference to jaw morphology here because this section is about locomotion.}
 %TG: Note that there's no change in the text here. Just a lecture on jaw morphologies. We might want to elaborate a bit in the manuscript.
@@
 
-\textcolor{blue}{The hierarchy of our paper wasn't clear in the first version but we have now clarified that this is the lead in to a larger section which has subsections e.g.
+\textcolor{blue}{The hierarchy of our paper wasn't clear as we intended in the first version but thanks to the review process we have now clarified that this is the lead in to a larger section which has subsections e.g.
 food processing (see also updated figure 1).}
@@
 \textcolor{blue}{One of the central points of our paper is to compare the biology of extant scavengers with extinct species in order to determine if they were suited to scavenging.
-All of the examples we use are backed up by references from previous work.
+And as we discuss in our introduction, because we can't observe the behaviour of extinct species directly we are left to make assumptions about their behaviour. 
 Rather than taking them as fact we offer these as possible behaviours that the extinct species engaged in.
-Because we can't observe their behaviour directly we are left to make assumptions about how likely they were to have enganged in them.
-But this is hardly unique to our work.
-We have added some extra references in support of our claims: \textit {``Thus, given pterosaurs seem to have cohabited in large numbers (Witton 2013), and the known benefits this can have for social foraging in birds (Dermody et al.
+We have added some extra references in support of our claims e.g.
+\textit {``Thus, given pterosaurs seem to have cohabited in large numbers (Witton 2013), and the theoretical benefits this can have for social foraging in birds (Dermody et al.
 2011), it seems probable that scrounging behaviours were seen in the flying pterosaurs as well.''}}
+
+% All of the examples we use are backed up by references from previous work.
+% But this is hardly unique to our work.
+
 \smallskip
@@
 Even vultures, the canonical example of obligate scavengers, can hunt.} We have tried to be clearer throughout the paper as well.
+Indeed, this is a key point of our review paper: to change our view of scavenging from the classic dichotomy of obligate and facultative scavengers, to a sliding scale depending on the species and the ecosystem.
 }
@@
 We appreciate that there is a continuum across and among species which we specifically note in this revised version. 
+As we note in the above comment, a combination of a species traits, and the state of the ecosystem they find themselves in will determine their scale of scavenging. 
+Search rates can be altered by environmmetal or ecological conditions as well as intrinsic traits of the forager. 
+So, far from invalidating our idea it reinforces it.
 Moreover, researchers working in palaeobiology often try to reconstruct the behaviours of extinct species; we are following this approach when we make inferences about ancient systems.}
@@
 
-%TG: They make a point here (although their diversity definition rant is completely out of topic). Should we just drop figure 2 and use it as an abstract/press release picture?
+\textcolor{blue}{Figure 2 has been removed and figure 1 has been expanded.}
+
+%TG: They make a point here (although their diversity definition rant is completely out of topic). Should we just drop figure 2 and use it as an abstract/press release picture? AK Agreed
 
@@
 These are very broad areas so in a review of this sort some overlap is inevitable.
-We feel our inclusion of extinct species and systems as well as the frame of optimal foraging theory gives this manuscript its novelty.} 
+We feel our inclusion of extinct species and systems as well as the novel framing under optimal foraging theory gives this manuscript its novelty.} 
 \smallskip
@@
 2015).}
+
+\textcolor{blue}{We have included a new section detailing the influence of carcass size on scavenger ecology \textit{``Modern scavenging assemblages are known to be influenced by carcass size (Moleon et al.
+2015).
+Larger carcasses tend to last longer and also present a more conspicuous target for a foraging scavenger which results in more species attending them (Moleon et al. 2015).
+This will have had implications for extinct assemblages because body size distributions vary across different environments but also across time. 
+O'Gorman \& Hone (2012) showed Mesozoic faunal distributions may have been skewed towards larger species even when fossil biases are taken into account. 
+Similarly, the megafauna of the Pleistocene (Doughty et al. 2013) would have produced large carcasses. 
+As a result, the scavenger assemblages during this era would have been particularly diverse.''}}
+
 \end{enumerate}
@@
 
+\textcolor{blue}{We have added a final line to our abstract giving some specific findings: \textit{``Our approach offers, among others, an explanation as to why extinct pterosaurs likely scavenged and why extant bats don't.''}}
+
 \item{THE CHALLENGES OF SCAVENGING:} Page 3, 1st sentence: ``often difficult to predict'' compared to what?
@@
 2014).
+
+\textcolor{blue}{This passage has been removed from the section and later on in the paper we note that: \textit{``Carrion can often be unpredictable, such as in cases where animals succumb to disease''}}
 
@@
 \item{METABOLISM:} Page 4, 1st sentence: ``the sporadic nature of carrion''; please explain.
+
+\textcolor{blue}{This has been changed to ephemeral.}
 
@@
 
+\textcolor{blue}{Facilitation is now considered throughout the manuscript.}
+
 \item{DETECTION:} Page 11, last sentence: explain why wind speed affects the olfactory capacity of an aquatic species.
+
+\textcolor{blue}{We have added a section explaining why this is so: \textit{``wind speed determined the number of sharks feeding at whale carcasses due to chemical stimului from the carcasses being propogated through the water by the wind, indicating they were dependent on detecting the odours from the decaying whales.''}}
 
@@
 2012) and carcass accessibility (e.g., see DeVault \& Krochmal 2002).
+
+\textcolor{blue}{We now discuss all 4 aspects e.g.
+\textit{``Notably, freezing carcasses can become too hard to consume by most vertebrate carnviores (Selva et al 2003)''} and \textit{``Spikes in temperature can result in severe droughts causing mass mortality events which result in relatively predictable peaks in carrion availability (Kendall et al.
+2014).''}}
 
@@
 Rev.).
+
+\textcolor{blue}{Our paper now references the literature on the facilitative processes involved in scavenging.}
 
@@
 
+\textcolor{blue}{We now note that their \textit{``skull morphologies and neck musculature are indicative of animals that were suited to removing large amounts of flesh from an immobile foodstuff (Witton 2013)''.}}
+
 \item{CONCLUSION:} Page 18, bottom paragraph, 3rd sentence: change ``technique'' to ``approach''?
@@
 
+\textcolor{blue}{This figure has been removed}
+
 \end{enumerate}
@@
 \bigskip
+\bigskip
+\bigskip
+\bigskip
+\bigskip
+\bigskip
+\bigskip
+\bigskip
+
 Yours Sincerely,
+\bigskip
 
 Adam Kane and co-authors
+
 %\closing{Yours Faithfully,}

### 2016-08-08 natural history of scavenging.tex
andrew comments added

@@
 
-\nolinenumbers
+%\nolinenumbers
 
@@
 
-\author{Adam Kane, Kevin Healy, Thomas Guillerme, Graeme Ruxton, \& Andrew Jackson.}
+\author{Adam Kane, Kevin Healy, Thomas Guillerme, Graeme D. Ruxton, \\ \& Andrew L. Jackson.}
 
@@
 These features that make up our \textcolor{blue}{qualitative} scale of scavenging can be applied to any given species and allow us to judge the likely importance of this foraging behaviour. 
+\textcolor{blue}{Our approach offers, among others, an explanation as to why extinct pterosaurs likely scavenged and why extant bats don't.}    
 \end{abstract}
@@
 
-Historically, \textcolor{blue}{food webs have classified most acts of carnivory as predation events \citep{wilson2011scavenging}. 
+\textcolor{blue}{Historically, food webs have classified most acts of carnivory as predation events \citep{wilson2011scavenging}. 
 This is clearly a mistaken view because of the existence of scavenging, a behaviour displayed by almost all carnivorous vertebrates \citep{devault2003scavenging}. 
-Recent research has begun to redress this imbalance by recognising the prevalence of scavenging and its implications for trophic ecology through detailing the ecological consequences of scavanging across modern ecosystems \citep{pereira2014facultative,periquet2015lion}.
+Recent research has begun to redress this imbalance by recognising the prevalence of scavenging and its implications for trophic ecology across modern ecosystems \citep{pereira2014facultative,periquet2015lion}.
+By increasing multichannel feeding and the number of food web links scavenging can give stability to an ecosystem \citep{moleon2015living}. 
 Indeed, there is no discrete divide between predators and scavengers but rather a continuous gradient in terms of the proportion of carrion in the diet \citep{pereira2014facultative}.
 Even vultures, the canonical example of obligate scavengers, can hunt \citep{margalida2011scavenger}.
-While it is relatively easy to determine whether a species engages in scavenging through observation, determining proportions requires extensive behavioral data (I put this in to placate the reviewer who said it was easy).
+While it is relatively easy to determine whether a species engages in scavenging through observation, determining the proportion of carrion in the diet requires extensive behavioral data.
+% (I put this in to placate the reviewer who said it was easy).
 Unfortunately, the lack of direct behavioural data makes it even more difficult to discern scavenging from predation among extinct forms where it was surely exhibited by prehistoric vertebrate carnivores.
@@
 These characteristics can be assumed to be under evolutionary selection pressures for traits that increase carrion discovery and monopoly.
-Finally, the potential for scavenging will also depend on the density, size, and quality of carcases produced, all of which are affected by complex ecosystem dynamics \textcolor{blue}{\citep{moleon2014inter}.
+Finally, the potential for scavenging, in either a species as a whole, or for individuals within a population, will also depend on the density, size, and quality of carcases produced, all of which are affected by complex ecosystem dynamics \textcolor{blue}{\citep{moleon2014inter}.
 Even these aspects can be indirectly impacted by the behaviour of the scavenger \citep{moleon2014inter}.}
@@
 \centering
-   \includegraphics[width=1\textwidth]{Summary_figure/Summary_figure.pdf}
+   \includegraphics[width=1\textwidth]{Composite figure/Compositefig4.pdf}
 \caption{Factors influencing the proportion of scavenging in a vertebrates' diet.
-The black and white gradient bar on the top represent the proportion of scavenging in the diet, ranging from low (- -) on the left hand side to high (+ +) on the right hand side.
-Blue lines indicate that a reduction in the factor will lead to more or less proportion of scavenging in the diet and red lines indicate an increase.}
+Each of the traits/ factors ranges from low on the left to high on the right.
+A high value for a given trait can either increase scavenging propensity (+ +) or reduce it (- -), the same is true for a low value of said trait.}
 \label{Summary_figure}
@@
 \textcolor{blue}{Carcass production itself depends on factors such as predation, disease and scavenging itself.
-For instance, by consuming the majority of carrion, more obligate scavengers deny a resource to their facultative counterparts obliging the latter to hunt more, and thus, increase the amount of carrion \citep{moleon2014inter}.}
+For instance, by consuming the majority of carrion, more prolific scavengers may deny a resource to their less prolific counterparts obliging the latter to hunt more, and thus, increase the amount of carrion \citep{moleon2014inter}.}
 
 \subsection{Metabolism}
-Because of the ephemeral nature of carrion \citep{devault2003scavenging,ruxton2004obligate} we would expect adaptations that reduce energetic costs of maintenance to be selected for in scavengers as it would allow for longer inter-feeding periods.
+Because of the ephemeral nature of carrion \citep{devault2003scavenging,ruxton2004obligate} we expect adaptations that reduce energetic costs of maintenance to be selected for in scavengers as it would allow for longer inter-feeding periods.
 Extant reptiles possess an advantage here, in that over the course of a year their food requirements can be 30 times lower than an endotherm of equal size \citep{Nagy1621}.
@@
 The same reasoning can be applied to crocodiles and their allies \citep{forrest2003evidence,moleon2015carcass}. 
-\cite{carey1982temperature} found that sharks, as ectothermics, have the ability to go weeks between meals because they focus on the energy-rich sections of carcasses (see Handling Time \ref{handlingTime}).
+\cite{carey1982temperature} found that sharks, as ectotherms, have the ability to go weeks between meals because they focus on the energy-rich sections of carcasses (see Handling Time \ref{handlingTime}).
 Endotherms have also evolved physiological mechanisms that allow them to depress their otherwise high metabolic rates at certain times e.g.
@@
 \textcolor{blue}{While it is difficult to infer the metabolisms of extinct species, similarities can still also be drawn.
-Mainly ectothermic groups such as sharks and bony fish were likely to benefit from the ability to bridge long time-spans between meals such as seen in snakes.
-In contrast, forms such as dinosaurs \citep{grady2014evidence}, pterosuars \citep{wellnhofer1991illustrated}, some marine reptiles \citep{bernard2010regulation} and some early mammals \citep{lovegrove2016phenology} have been proposed to have at least partial endothermy resulting in requiring a higher intake of material in comparison to their ectothermic equivalents.
+Many ectothermic groups such as sharks and bony fish were likely to benefit from the ability to bridge long time-spans between meals such as seen in snakes.
+In contrast, forms such as dinosaurs \citep{grady2014evidence}, pterosuars \citep{wellnhofer1991illustrated}, some marine reptiles \citep{bernard2010regulation} and some early mammals \citep{lovegrove2016phenology} have been proposed to have possessed at least partial endothermy potentially requiring a higher intake of material in comparison to their ectothermic equivalents.
 However, such higher metabolic rates would also increased the ability of such species to actively search for carrion.}
@@
 \subsection{Locomotion}
-Carrion can often be unpredictable, such as in cases where animals succumb to disease \citep{devault2003scavenging}, which means scavenging depends more on the ability to efficiently move over larger areas than does predation.
+\textcolor{blue}{Carrion can often be unpredictable, such as in cases where animals succumb to disease \citep{devault2003scavenging}, which means scavenging depends more on the ability to efficiently move over larger areas than does predation.}
 This generally requires an efficient transfer of metabolic energy into movement which relies on the animal's anatomy and physiology as well as the medium of the environment in which the animal is moving (i.e.
@@
 Perhaps the most efficient form of locomotion in vertebrates is found in flying species. 
-Despite the energetic costs of flight, the only known vertebrates that can be obligate scavengers are the old and the new world vultures. 
+\textcolor{blue}{Despite the energetic costs of flight, the most prolific modern, vertebrate scavengers are the old and the new world vultures.} 
 While powered flight is energetically expensive, species like vultures have evolved behavioural and anatomical features to exploit air currents using their large wingspans, allowing them to soar at a cost of only around 1.5 times their metabolic rate \citep{hedenstrom1993migration,duriez2014cheap}.
@@
 
-While soaring is perhaps the only viable means of locomotion that allows for an obligate, scavenging life-style \citep{ruxton2004obligate}, powered flight is still an efficient means of locomotion. 
+While soaring is perhaps the only viable means of locomotion that allows for scavenger to rely entirely on carrion \citep{ruxton2004obligate}, powered flight is still an efficient means of locomotion.
 Certainly, avian flight is cheaper than either walking or running \citep{tucker1975energetic}.
@@
 
-Interestingly, however, style of swimming in fishes does not significantly affect the cost of movement \citep{williams1999evolution}, hence it is likely that many aquatic species with large ranges will encounter scavenging opportunities.
-It is perhaps jaw morphology that is more likely to define which species can avail of carcasses.
+Interestingly, however, style of swimming in fishes does not significantly affect the cost of movement \citep{williams1999evolution}.
+Hence, it is likely that many aquatic species with large ranges will encounter scavenging opportunities.
+% It is perhaps jaw morphology that is more likely to define which species can avail of carcasses.
 
@@
 Terrestrial environments are the most energetically costly in which to move \citep{tucker1975energetic}. 
-Unlike aerial and aquatic environments support must be provided through the animal's posture. 
+Unlike aerial and aquatic environments, support must be provided through the animal's posture. 
 The early transition from a sprawling gait, seen in early tetrapods, to the more erect posture of synapsids and later dinosaurs and mammals, has often been supposed as conferring a huge advantage to the latter groups \citep{sullivan2015posture}.
@@
 Modern endothermic mammals can sustain longer periods of energetically expensive activity \citep{bennett1979endothermy} resulting in larger foraging ranges. 
-Today, terrestrial scavenging in mammals is probably best known in the African Savannah context where hyenas, jackals and lions all take sizeable proportions of carrion in their diet \textcolor{blue}{\citep{pereira2014facultative,periquet2015lion}.}
+Today, terrestrial scavenging in mammals is probably best known in the \textcolor{blue}{African Savannah context where hyenas and lions can all take sizeable proportions of carrion in their diet \citep{pereira2014facultative,periquet2015lion}.}
 In the spotted hyena (\textit{Crocuta crocuta}), striped hyena (\textit{Hyaena hyaena}) and brown hyena (\textit{Hyaena brunnea}) it can be over 90\% \citep{jones2015african}.
-And although no contemporary terrestrial vertebrate exists as an obligate scavenger, most, if not all, are facultative to some extent \textcolor{blue}{\citep{devault2003scavenging,beasley2015vertebrates,pereira2014facultative}}.
+\textcolor{blue}{Most, if not all, vertebrate terrestrial carnivores take carrion to some extent \citep{devault2003scavenging,beasley2015vertebrates,pereira2014facultative}}.
 \textcolor{blue}{The particular ability of hyenas to subsist on high proportions of carrion means we can use them as examples of efficient terrestrial scavengers to compare with other forms \citep{periquet2015lion}}.
@@
 Our locomotory efficiency has fed into the question of where our ancestors placed on the hunter-scavenger axis during the Plio-Pleistocene, which has been a matter of debate for years \citep{dominguez2002hunting}.
-\cite{ruxton2013endurance} added to this debate with their argument that long distance endurance running \textcolor{blue}{was not an important feature of hominin scavenging but was instead used by humans for hunting.
-Even when considering the use of weapons and social foraging, the authors argue that the high costs of running long distances would render it an unattractive foraging strategy for scavenging.}
+\cite{ruxton2013endurance} added to this debate with their argument that long distance endurance running \textcolor{blue}{was not an important feature of hominin scavenging but was instead used by humans for hunting prey by chasing them to exhaustion over large distances.
+Even when considering the use of weapons and social foraging, the authors argue that the high cost of running long distances would render it an unattractive foraging strategy for scavenging.}
 
@@
 Smell seems to be the primary means of carcass detection in sharks as well. 
-\cite{fallows2013white} found that wind speed determined the number of sharks feeding at whale carcasses, indicating they were dependent on detecting the odours from the decaying whales.
+\cite{fallows2013white} found that wind speed determined the number of sharks feeding at whale carcasses due to chemical stimului from the carcasses being propogated through the water by the wind, indicating they were dependent on detecting the odours from the decaying whales.
 
@@
 \textcolor{blue}{Research has shown that an animal need only travel 36 km to encounter a fresh whale carcass \citep{smith2003ecology}.} 
-The phenomenon of occasional bounties of carrion in the form of these whale falls has led some researchers to investigate if an obligate scavenger could survive by seeking out these remains exclusively.}
+The phenomenon of occasional bounties of carrion in the form of these whale falls has led some researchers to investigate if a scavenger could survive by seeking out these remains exclusively.}
 \cite{ruxton2005searching} argued that although this is energetically feasible it's ecologically unlikely.
-Any animal that could find such whale carcasses is unlikely to have ignored other types of carrion.
+Any animal that could find such whale carcasses is unlikely to have ignored other types of carrion. \\
 Although no aquatic species have ever exceeded the size of whales, some enormous animals have evolved in this environment before the evolution of cetaceans, including \textit{Leedsichthys}, a bony fish from the Middle Jurassic (174.1 - 163.5 Mya) and the aquatic Mesozoic reptiles, the plesiosaurs, pliosaurs and ichtyosaurs, that could all exceed 15 metres in length \citep{ruxton2011zoology,danise2014ecological}.
-So, despite being unlikely 
+So, despite being unlikely, the energetic feasibility of a marine scavenger that specialises on large carcasses has a long history.
+
 % KH: Is there an estimate of how much of the fishes diet can be made up from this type of scavenging?
-, the energetic feasibility of a marine scavenger that specialises on large carcasses has a long history.
 
-Perhaps the greatest environmental driver of scavenging tendency is that of temperature. 
-This has a significant bearing on the availability and persistence of carrion.
+Perhaps the greatest environmental driver of scavenging tendency is that of temperature which has a significant bearing on the availability, predictability and persistence of carrion.
 We know from the geological record that the Earth has undergone radical fluctuations in temperature over time.
 \textcolor{blue}{On land, there are a wide variety of vegetation types, from thickly forested areas to open grassland and the extent of this vegetation changes with season.}
-All of these factors influence the availabiltiy of carrion.  
-Moreover, oceanic productivity is also impacted by climactic conditions. 
+%All of these factors influence the availabiltiy of carrion.  
+%Moreover, oceanic productivity is also impacted by climactic conditions. 
 %KH: I would delete the last two lines of this paragraph and merge it with the next one
-
 To illustrate the point, a 10$^{\circ}$C increase in ambient temperature can double carcass decomposition rates \citep{parmenter2009carrion} and geological evidence indicates that the Mesozoic Earth was on average at least 6 $^{\circ}$C warmer than now \citep{sellwood2006mesozoic}.
@@
 \cite{beasley2015vertebrates} who point to a series of studies showing how microbes and invertebrates benefit at higher temperatures to the detriment of vertebrate scavengers such that ``above 20$^{\circ}$C vertebrates were able to detect and consume only 19\% of small-mammal carcasses, whereas at temperatures below 18$^{\circ}$C, vertebrates consumed 49\% of carcasses".
-\textcolor{blue}{In savannah, dense vegetation can prevent vultures from landing to a carcass because they will be unable to take off again \citep{bamford2009effect}.
+\textcolor{blue}{
+Spikes in temperature can result in severe droughts causing mass mortality events which result in relatively predictable peaks in carrion availability \citep{kendall2014african}.%In savannah, dense vegetation can prevent vultures from landing to a carcass because they will be unable to take off again \citep{bamford2009effect}.
 %KH: This is a biotic interaction so I would either remove it or move it somewhere else. 
@@
 Foragers do not exist in isolation and we know from field observations that scavengers can scrounge on the discoveries of other carnivores. 
-This sort of facilitation occurs in both obligate and facultative scavengers, in the air and on the ground \citep{KaneVul,jones2015african}.
+This sort of facilitation occurs across a range of scavengers, in the air and on the ground \citep{KaneVul,jones2015african}.
 
@@
 Information transfer of this kind is typically inadvertent and as a consequence no complex social interactions are required, simply the ability to recognise a successful forager.
-\textcolor{blue}{Thus, given pterosaurs seem to have cohabited in large numbers \citep{witton2013pterosaurs}, and the known benefits this can have for social foraging in birds \citep{jackson2011evolutionary}, it seems probable that scrounging behaviours were seen in the flying pterosaurs as well.}
+\textcolor{blue}{Thus, given pterosaurs seem to have cohabited in large numbers \citep{witton2013pterosaurs}, and the theoretical benefits this can have for social foraging in birds \citep{jackson2011evolutionary}, it seems probable that scrounging behaviours were seen in the flying pterosaurs as well.}
 
-This type of facilitation then increases the encounter rate of the facilitated species and can increase the population size of the latter \citep{moleon2014inter}. 
+\textcolor{blue}{This type of facilitation then increases the encounter rate of the facilitated species and can increase the population size of the latter \citep{moleon2014inter}. 
 This higher population of predators may take more prey and so produce more carcasses.
-Conversely, by feeding on carcasses, predators may hunt less because they are sated by carrion, ultimately reducing predation risk on their prey base \citep{moleon2014inter}.
+Conversely, by feeding on carcasses, predators may hunt less because they are sated by carrion, ultimately reducing predation risk on their prey base \citep{moleon2014inter}.}
 
@@
 For instance, lions can acquire much of their carrion through kleptoparasitism of hyena kills \textcolor{blue}{\citep{trinkel2005competitive,pereira2014facultative,periquet2015lion}}. 
-This line of reasoning suggests that some theropod dinosaurs, who could get up to 15 tonnes, would have had easily monopolised a carcass \citep{weishampel2004dinosauria} provided they could find them efficiently \citep{kane2016body}.
+This line of reasoning suggests that some theropod dinosaurs, who could get up to 15 tonnes, would have easily monopolised a carcass \citep{weishampel2004dinosauria} provided they could find them efficiently \citep{kane2016body}.
 
@@
 Additionally, many pterosaurs were far bigger again, with estimated body masses of over 200 kg in the Azhdarchids \citep{witton2010size}.
-Although \cite{witton2008reappraisal} argued that neck inflexibility and straight, rather than hooked jaw morphology points against Azhdarchids existing as \textit{obligate} scavengers, their terrestrial proficiency indicates they would have been comfortable foraging on the ground.
-Extant Marabou Storks (\textit{Leptoptilos crumenifer}) have a comparable morphology and are noted facultative scavengers \citep{monadjem2012survival} so it is reasonable to believe that these pterosaurs behaved similarly.
+Although \cite{witton2008reappraisal} argued that neck inflexibility and straight, rather than hooked jaw morphology points against Azhdarchids being \textcolor{blue}{as well adapated as vultures} to scavenging, their terrestrial proficiency indicates they would have been comfortable foraging on the ground.
+Extant Marabou Storks (\textit{Leptoptilos crumeniferus}) have a comparable morphology and are noted facultative scavengers \citep{monadjem2012survival} so it is reasonable to believe that these pterosaurs behaved similarly.
 
 By contrast, extant bats seem poorly equipped to deal with competitors. 
-Their poor terrestrial ability, smalls size and cost of movement on the ground would count against them while attempting to fend off other species at a carcass \citep{riskin2006terrestrial,voigt2012terrestrial}.
+Their poor terrestrial ability, small size and cost of movement on the ground would count against them while attempting to fend off other species at a carcass \citep{riskin2006terrestrial,voigt2012terrestrial}.
 
 Smaller species can compensate for a lack of individual body size by weight of numbers in competitive interactions. 
-This is true for a host of notable scavengers, such as vultures, early hominins and hyenas, who can dominate larger competitors provided they substantially outnumber them \citep{KaneVul,trinkel2005competitive,ruxton2013endurance}.
+This is true for a host of notable scavengers, such as vultures, early hominins and hyenas, who can and could dominate larger competitors provided they substantially outnumber(ed) them \citep{KaneVul,trinkel2005competitive,ruxton2013endurance}.
 
@@
 \cite{shivik2006vultures} suggests that ``specialized oral structures in snakes may have evolved under pressures associated with scavenging."
-Moreover, some researchers have charted an evolutionary course from basal fossorial snakes to modern terrestrial species by way of an obligate scavenger intermediate \citep{bauchot2006snakes}.
+Moreover, some researchers have charted an evolutionary course from basal fossorial snakes to modern terrestrial species by way of a scavenging intermediate \citep{bauchot2006snakes}.
 However, snake venom is primary associated with predation in extant species suggesting that if venom played such a role in extinct species it is no longer its main function \citep{casewell2013complex}.
@@
 % In extant systems these interactions can reduce handling times for the facilitated species. 
-Some examples of reduced handling time include vultures, hyenas and wolves opening the tough hides of ungulate carcasses that would otherwise be inaccessible to corvids and smaller mammalian carnivores \citep{selva2003scavenging,moleon2015carcass}.}
+Some examples of reduced handling time include vultures, hyenas and wolves opening the tough hides of ungulate carcasses that would otherwise be inaccessible to corvids and smaller mammalian carnivores \citep{selva2003scavenging,moleon2015carcass}.
+Unfortunately these interactions are particularly difficult to detect in extinct species.}
 

### 2016-08-09 response_letter.tex
Graeme's comments addressed

@@
 
-\textcolor{blue}{We now clearly state that the scale is qualitative.}
+\textcolor{blue}{We now emphasise that our scale is qualitative here and throughout the manuscript.}
 
@@
 \textcolor{blue}{We have rewritten the introduction to more accurately reflect the aim of this review i.e.
-that the scavenging gradient allows us to infer scavenging in the fossil record as well as in modern species.} %TG:
+that the scavenging gradient allows us to infer scavenging in the fossil record as well as in modern species.
+In making this revision we have reduced overlap with De Vault (2003) considerably.} %TG:
 
@@
 \textcolor{blue} {We make the following point in our introduction: \textit{``Indeed, there is no discrete divide between predators and scavengers but rather a continuous gradient.
-Even vultures, the canonical example of obligate scavengers, can hunt.''} We have tried to be clearer throughout the paper as well.}
+Even vultures, the canonical example of obligate scavengers, can hunt.''} We have tried to be clearer throughout the paper as well by removing references to this division.}
 
@@
 
-\textcolor{blue}{We've removed the reference to jaw morphology here because this section is about locomotion.}
+\textcolor{blue}{We now avoid discussion of jaw morphology at this early point in the manuscript, and later develop fully our ideas on how feeding anatomy can be related to propensity to exploit scavenging opportunities.}
 %TG: Note that there's no change in the text here. Just a lecture on jaw morphologies. We might want to elaborate a bit in the manuscript.
@@
 
-\smallskip
+\bigskip
 \textcolor{blue}{One of the central points of our paper is to compare the biology of extant scavengers with extinct species in order to determine if they were suited to scavenging.
@@
 
-\smallskip
+\bigskip
 
@@
 
+\bigskip
+
 \textcolor{blue} {This was also raised by reviewer 1.
-We make the following point in our introduction: \textit{Indeed, there is no discrete divide between predators and scavengers but rather a continuous gradient.
+We make the following point in our introduction: \textit{``Indeed, there is no discrete divide between predators and scavengers but rather a continuous gradient.
 Even vultures, the canonical example of obligate scavengers, can hunt.} We have tried to be clearer throughout the paper as well.
-Indeed, this is a key point of our review paper: to change our view of scavenging from the classic dichotomy of obligate and facultative scavengers, to a sliding scale depending on the species and the ecosystem.
-}
+Indeed, this is a key point of our review paper: to change our view of scavenging from the classic dichotomy of obligate and facultative scavengers, to a sliding scale depending on the species and the ecosystem.''}
+
+\bigskip
 
@@
 But to suggest, as they do in their abstract, that they can apply their scale of scavenging to any species at any time to judge how important scavenging was in its diet is frankly completely ignoring behavioral variability over time and space.
-\smallskip
+\bigskip
 
-\textcolor{blue}{It is certainly true that scavengers can be variable in the proportion of carrion they consume but we would argue that species like hyenas have certain characteristics that make this possible.
+\textcolor{blue}{It is undoubtedly true that scavengers can be variable in the proportion of carrion they consume but we would argue that species like hyenas have certain characteristics that make this possible.
 In contrast, cheetahs do not have a biology that lends itself to scavenging in the wild, due to poor competitive ability etc.
 We appreciate that there is a continuum across and among species which we specifically note in this revised version. 
-As we note in the above comment, a combination of a species traits, and the state of the ecosystem they find themselves in will determine their scale of scavenging. 
+As we note in the above comment, a combination of a species' traits, and the state of the ecosystem they find themselves in will determine their scale of scavenging. 
 Search rates can be altered by environmmetal or ecological conditions as well as intrinsic traits of the forager. 
@@
 Moreover, researchers working in palaeobiology often try to reconstruct the behaviours of extinct species; we are following this approach when we make inferences about ancient systems.}
-\smallskip
+\bigskip
 
@@
 Diversity has specific ecological definition(s) that are being ignored here.
+\bigskip
 
@@
 %TG: They make a point here (although their diversity definition rant is completely out of topic). Should we just drop figure 2 and use it as an abstract/press release picture? AK Agreed 
+\bigskip
 
@@
 They might look to articles about other modern scavenging opportunities such as Capaldo and Peter's study of wildebeest drownings \url{http://www.sciencedirect.com/science/article/pii/S0305440385700398}, and suggestions that sabertooth cats which they claim were unable to feed on bones may have consumed more bone than penecontemporaneous dire wolves \url{https://www.jstor.org/stable/20627167?seq=1#page_scan_tab_contents}.
-\\
-\smallskip
+
+\bigskip
+
 \textcolor{blue}{The issue of the structure of our paper was noted by all reviewers who all had suggestions to improve it. 
-We have made an effort to delineate all of the main sections and subsections so that our frame of optimal foraging is more apparent.}
+We have made an effort to delineate all of the main sections and subsections so that our frame of optimal foraging is more apparent.
+We have added the two suggested references to our paper.}
 
-\smallskip
+\bigskip
 In sum, while I think the aim of the article is a very worthwhile one, I think the authors cannot claim to have produced an all-encompassing model that can predict the presence or level of scavenging in any organism - because I think this simply cannot be done.
@@
 Finally, some of the perspectives of this manuscript overlap substantially with parts of Beasley et al. 2015 (pages 108-112), which is cited numerous times in the manuscript.
-\smallskip
+\bigskip
 
@@
 We feel our inclusion of extinct species and systems as well as the novel framing under optimal foraging theory gives this manuscript its novelty.} 
-\smallskip
+\bigskip
 
@@
 
-\textcolor{blue}{We changed the title of the manuscript to ``A Recipe for Scavenging in Vertebrates - the natural history of a behaviour''.}
+\textcolor{blue}{We changed the title of the manuscript to ``A Recipe for Scavenging in Vertebrates - the natural history of a behaviour''.
+We have added two mentions of inverebrates as competitors to the manuscript but given our focus these are not expanded upon.}
 %TG: Also maybe mention invertebrates in the competition part of the manuscript.
@@
 
-\textcolor{blue}{We clarified the structure of the manuscript by introducing clearer hierarchical subsections.}
+\textcolor{blue}{We clarified the structure of the manuscript by introducing clearer hierarchical subsections, an additional section on facilitation and a revised figure 1.}
 %TG: + more blabalal on his ``Facilitation'' part
@@
 
+\textcolor{blue}{Thanks very much for these very useful references, all of which are now included and used to bolster some sections of the paper, particularly with respect to facilitation, see below for further details.}
+
 Page 2, 1st sentence: but see Mole\'{o}n \& S\'{a}nchez-Zapata (2015); many of the most charismatic vertebrates of the world are scavengers.
@@
 
-\textcolor{blue}{We now refer to the behaviour rather than the species as lacking charisma.}
+\textcolor{blue}{Our introduction has been rewritten and this passage no longer appears.}
 
@@
 
-\textcolor{blue}{We have added a final line to our abstract giving some specific findings: \textit{``Our approach offers, among others, an explanation as to why extinct pterosaurs likely scavenged and why extant bats don't.''}}
+\textcolor{blue}{We cannot provide details of specific predictions in sufficient detail to be intelligible within the work limit of the abstract, but have modified the structure of the manuscript to make these stand out more clearly.}
 

### 2016-08-09 natural history of scavenging.tex
Graeme's comments addressed

@@
 These features that make up our \textcolor{blue}{qualitative} scale of scavenging can be applied to any given species and allow us to judge the likely importance of this foraging behaviour.   
-\textcolor{blue}{Our approach offers, among others, an explanation as to why extinct pterosaurs likely scavenged and why extant bats don't.}    
 \end{abstract}
@@
 Recent research has begun to redress this imbalance by recognising the prevalence of scavenging and its implications for trophic ecology across modern ecosystems \citep{pereira2014facultative,periquet2015lion}.
-By increasing multichannel feeding and the number of food web links scavenging can give stability to an ecosystem \citep{moleon2015living}. 
+By increasing multichannel feeding and the number of food web links, scavenging can confer stability to an ecosystem \citep{moleon2015living}. 
 Indeed, there is no discrete divide between predators and scavengers but rather a continuous gradient in terms of the proportion of carrion in the diet \citep{pereira2014facultative}.
@@
 Some simple heuristics can inform us whether or not the prey was scavenged, for instance, in cases where it was simply too large to have been killed by the ostensible predator \citep{pobiner2008paleoecological}.}
-But clearly, a scavenger does not only feed on animals too big for it to have hunted.
+But clearly, a scavenger does not only feed on dead animals too big for it to have hunted.
 
@@
 \textcolor{blue}{Species that rely on scavenging to sustain substantial portions of their diets must encounter a sufficient amount of carrion in order to meet their energetic demands.
-Once found, the scavenger must be able to out-compete any potential competitors and efficiently process the, typically decaying, carcass replete with micro-organism derived toxins \citep{ruxton2014fruit}.}
+Once found, the scavenger must be able to out-compete any potential competitors and efficiently process the, typically decaying, carcass replete with textcolor{blue}{invertebrates} and micro-organism derived toxins \citep{ruxton2014fruit}.}
 These characteristics can be assumed to be under evolutionary selection pressures for traits that increase carrion discovery and monopoly.
@@
 By considering scavenging in this context of optimal foraging we can identify the prerequisite attributes and processes required for the behaviour. 
-This has enabled us to propose a scale of scavenging whereupon we can place \textcolor{blue}{a given carnivore vertebrate} and assess the likely importance of carrion in its diet.
+This has enabled us to propose a \textcolor{blue}{qualitative} scale of scavenging whereupon we can place \textcolor{blue}{a given carnivore vertebrate} and assess the likely importance of carrion in its diet.
 \textcolor{blue}{We argue that such a scale could be valuable when behaviour can not be observed directly, e.g. for extinct vertebrates.}
@@
 In contrast, forms such as dinosaurs \citep{grady2014evidence}, pterosuars \citep{wellnhofer1991illustrated}, some marine reptiles \citep{bernard2010regulation} and some early mammals \citep{lovegrove2016phenology} have been proposed to have possessed at least partial endothermy potentially requiring a higher intake of material in comparison to their ectothermic equivalents.
-However, such higher metabolic rates would also increased the ability of such species to actively search for carrion.}
+However, such higher metabolic rates would also have increased the ability of such species to actively search for carrion.}
 %KH: I put this section in as there is no reference in how we can use the above paragraph to infer scavenging in extinct species
@@
 While soaring is perhaps the only viable means of locomotion that allows for scavenger to rely entirely on carrion \citep{ruxton2004obligate}, powered flight is still an efficient means of locomotion.
-Certainly, avian flight is cheaper than either walking or running \citep{tucker1975energetic}.
+Certainly, avian flight is cheaper per unit distance than either walking or running \citep{tucker1975energetic}.
 
@@
 For instance, \cite{alexander2004bipedal} shows that, in the case of humans, we are more economical than predicted while walking and less so while running according to predicted costs of terrestrial movement calculated by allometric scaling relationships.
-Our locomotory efficiency has fed into the question of where our ancestors placed on the hunter-scavenger axis during the Plio-Pleistocene, which has been a matter of debate for years \citep{dominguez2002hunting}.
+Our locomotory efficiency has fed into the question of where our ancestors placed on the hunter-scavenger axis during the Plio-Pleistocene, which has been a matter of debate for years \citep{capaldo1995skeletal,dominguez2002hunting}.
 \cite{ruxton2013endurance} added to this debate with their argument that long distance endurance running \textcolor{blue}{was not an important feature of hominin scavenging but was instead used by humans for hunting prey by chasing them to exhaustion over large distances.
@@
 
-In addition to fending off other vertebrates, scavengers also have to contend with competition from micro-organisms, which requires a specialised physiology. 
+In addition to fending off other vertebrates, scavengers also have to contend with competition from \textcolor{blue}{invertebrates and micro-organisms, the latter of which may require a specialised physiology.}
 Although the findings of \cite{shivik2006vultures} that ``evolutionary pressures favor detection maximizers relative to toxification minimizers in competitive interactions for carcasses." appear sound, the fact remains that overcoming micro-organism toxins is still a beneficial adaptation to any scavenger.
 Avian scavengers have evolved incredibly acidic stomachs that allow them to consume and process putrefied flesh with no ill effects \citep{houston1975digestive,roggenbuck2014microbiome}. 
-This adaptation is not restricted to vultures though, \cite{gremillet2012vultures} showed wandering albatrosses (\textit{Diomedea exulans}; so-called ``vultures of the seas'') had an average pH of 1.5, which enables them to consume fisheries discards and squid carcasses.
+This adaptation is not restricted to vultures though, \cite{gremillet2012vultures} showed wandering albatrosses (\textit{Diomedea exulans}; so-called ``vultures of the seas'') had an average stomach pH of 1.5, which enables them to consume fisheries discards and squid carcasses.
 There is also evidence of selection for ``toxification minimizers'' beyond birds among the ectotherms.
@@
 Moreover, some researchers have charted an evolutionary course from basal fossorial snakes to modern terrestrial species by way of a scavenging intermediate \citep{bauchot2006snakes}.
-However, snake venom is primary associated with predation in extant species suggesting that if venom played such a role in extinct species it is no longer its main function \citep{casewell2013complex}.
+However, snake venom is primary associated with predation in extant species suggesting that if venom played such a role aiding digestion of discovered carcasses in extinct species it is no longer its main function \citep{casewell2013complex}.
 \textcolor{blue}{In water, by contrast, competition with micro-organisms is significantly reduced because of the high pressures and low temperatures \citep{beasley2012carrion}.}
@@
 The animal also had an enormous bite force, with one estimate putting it at 57000 Newtons \citep{bates2012estimating} which would have been powerful enough to break open skeletons \citep{rayfield2001cranial}.
-Osteophagy may have been even more viable during the Mesozoic era as well because of this skewed body mass distribution of herbivores towards larger sizes \citep{10.1371/journal.pone.0051925}.
+Osteophagy may have been even more viable during the Mesozoic era as well because of the skewed body mass distribution of herbivores towards larger sizes \citep{10.1371/journal.pone.0051925}.
 When we couple this with the fact that skeletal mass scales greater than linearly with body mass \citep{prange1979scaling} there would have been a lot of bone material to consume in the environment provided an animal had the biology to process it \citep{chure1997one}.
@@
 Aside from raptors, the specialised beaks of many modern bird lineages tends to hinder their ability to eat meat which is in contrast to the first lineages that did not have this feature \citep{martyniuk2012field}.
-As \cite{martyniuk2012field} notes, \textcolor{blue}{this skull morphology of early birds indicates} they were predominantly carnivorous, implying scavenging was a live opportunity cf.
-their descendants.
+As \cite{martyniuk2012field} notes, \textcolor{blue}{this skull morphology of early birds indicates} they were predominantly carnivorous, implying scavenging was a live opportunity compared to many of their descendants.
 Among the pterosaurs, \cite{witton2013pterosaurs} makes the case that the istiodactyl pterosaurs were the most likely scavengers of this group based on their potential handling time. 
@@
 In common with any other forager be they grazer, browser or predator, scavengers past and present have had to balance their energetic costs with the gains of food. 
-The main factors we considered namely, encounter rate, handling time and prey availability can be used to create a scale of scavenging whereupon any species can be placed in order to establish the importance of carrion in it diet.
+The main factors we considered namely, encounter rate, handling time and prey availability can be used to create a scale of scavenging whereupon any species can be placed in order \textcolor{blue}{to predict the qualitative importance} of carrion in it diet.
 We hope this approach will be useful in the effort to explore this most understudied of feeding ecologies.
@@
 Thanks to Natalie Cooper for highlighting the potential for this review, and to Ara Monadjem and Deirdre McClean for their comments on the manuscript.
+We also thank Marcos Moleón and two other anonymous reviewers for very perceptive and helpful feedback on a previous version.
 

### 2016-08-10 response_letter.tex
comments on response letter

@@
 (2015).}
+%KH I feel this is a bit weak.
+While we aknoledge the envitable overlap with Beasely et al we more carefully popint towards how this builds on out qualitative frameworks of how our understanding of extant scavangign can be exxtended to extinct species.
 
@@
 The paper is mostly well written, and most (but not all) pertinent literature sources are included.
-I do worry about the lack of novelty of the manuscript, and what I perceived as haphazard organization. Here are some specific issues that should be addressed in my opinion:
+
+\textcolor{blue}{Thank you}
+
+I do worry about the lack of novelty of the manuscript, and what I perceived as haphazard organization.
+
+\textcolor{red}{We feel we have now more clearly outlined the structure of the paper by including an updated figure one explicitly showing the breakdown of the chapters and their link to foraging ecology.
+This also more clearly highlights the novel aspect of our review, of building a qualitative framework base on foraging ecology to identify the potential levels of scavenging in extinct species and systems and pointing towards possible methods to better focus on certain systems, such as energetic models.}
+
+Here are some specific issues that should be addressed in my opinion:
 
@@
 \textcolor{blue}{We emphasise that food webs have tended to lump all carnivory into predation and it is with extinct species that we have no observational evidence.}
+ \textcolor{red}{We now note measures often used in dietary analysis, such as stomach contents and stable isotope analysis, are unable to distinguish between scavenging and predation.
+We emphasis that the subsequent need for extensive behavioral data has led to food webs have tending to lump all carnivory into predation and this is particularly likely with extinct species.
+(lines 53-56)}
 
@@
 
-\textcolor{blue}{We now emphasise that our scale is qualitative here and throughout the manuscript.}
+\textcolor{blue}{We now emphasise that our scale is qualitative here and throughout the manuscript.\textcolor{red}{ in lines (----)}
 
@@
 that the scavenging gradient allows us to infer scavenging in the fossil record as well as in modern species.
-In making this revision we have reduced overlap with De Vault (2003) considerably.} %TG:
+In making this revision we have reduced the \textcolor{red}{inevitable} overlap with De Vault (2003) considerably.} %TG:
 
@@
 
-\textcolor{blue}{We now specify the approaches we have in mind:  \textit{``e.g. energetics models, comparative anatomy, palaeontology etc.'' }}
+\textcolor{blue}{We now specify the approaches we have in mind as \textit{``e.g.
+energetics models, comparative anatomy, palaeontology etc.''}} \textcolor{red}{on lines 22-24 with references referring to these methods.}
 
@@
 \textcolor{blue} {We make the following point in our introduction: \textit{``Indeed, there is no discrete divide between predators and scavengers but rather a continuous gradient.
-Even vultures, the canonical example of obligate scavengers, can hunt.''} We have tried to be clearer throughout the paper as well by removing references to this division.}
+Even vultures, the canonical example of obligate scavengers, can hunt.''} We have also been clearer on this point throughout the paper by removing references to this division \textcolor{red}{and instead referring to the gradient of scavenging outlined in the introduction}.}
 
@@
 
-\textcolor{blue}{See the point above which highlights that the concept of a scale, and not a false caterogisation, is central to our paper.
+\textcolor{blue}{As mentioned in the point above the concept of a scale, and not a false caterogisation, is central to our paper.
 We have now clarified one of the advantages of such a scale at the end of the introduction: \textit{``We argue that such a scale could be valuable when behaviour can not be observed directly, e.g.
 for extinct vertebrates.''}}
+%KH I re-worded this sentance a little bit
 
@@
 
-\textcolor{blue}{We have included this reference and explicitly discuss how differences in the physical properties of water Vs air can affect scavenging: \textit{``The physical differences between water and air mean carcass availability is radically different between these environments (Beasley et al.
+\textcolor{blue}{We have included this reference and explicitly discuss how differences in the physical properties of water Vs air can affect scavenging \textcolor{blue}{through carcass availability on pages 14 and 15}: \textit{``The physical differences between water and air mean carcass availability is radically different between these environments (Beasley et al.
 2012).
@@
 
-\textcolor{blue}{We now avoid discussion of jaw morphology at this early point in the manuscript, and later develop fully our ideas on how feeding anatomy can be related to propensity to exploit scavenging opportunities.}
-%TG: Note that there's no change in the text here. Just a lecture on jaw morphologies. We might want to elaborate a bit in the manuscript.
+\textcolor{blue}{We now avoid discussion of jaw morphology at this early point in the manuscript, and later develop fully our ideas in section 3.3 on how feeding anatomy can be related to propensity to exploit scavenging opportunities}
+%TG: Note that there's no change in the text here. Just a lecture on jaw morphologies. We might want to elaborate a bit in the manuscript. Agree we never actually answer his question.
 
@@
 
-\textcolor{blue}{The hierarchy of our paper wasn't clear as we intended in the first version but thanks to the review process we have now clarified that this is the lead in to a larger section which has subsections e.g.
+\textcolor{blue}{The hierarchy of our paper wasn't clear as we intended in the first version but thanks to the review process we have now clarified that this is the lead in paragraph to a larger section which has subsections e.g.
 food processing (see also updated figure 1).}
@@
 Rather than taking them as fact we offer these as possible behaviours that the extinct species engaged in.
-We have added some extra references in support of our claims e.g.
+\textcolor{red}{Our paper proposes these sets of possible scenarios of scavenging to allow for more quantitative tools, such a energetic models, to rule out or support these each scenario leading to a more refined understanding of scavenging in extinct species.}
+We have also added extra references in support of our claims e.g.
 \textit {``Thus, given pterosaurs seem to have cohabited in large numbers (Witton 2013), and the theoretical benefits this can have for social foraging in birds (Dermody et al.
@@
 
-\textcolor{blue} {This was also raised by reviewer 1.
-We make the following point in our introduction: \textit{``Indeed, there is no discrete divide between predators and scavengers but rather a continuous gradient.
+\textcolor{blue} {This was also raised by reviewer 1. and we thank both reviewers for pointing out this aspect which was unclear in the original manuscript.
+We now make the following point in our introduction: \textit{``Indeed, there is no discrete divide between predators and scavengers but rather a continuous gradient.
 Even vultures, the canonical example of obligate scavengers, can hunt.} We have tried to be clearer throughout the paper as well.
-Indeed, this is a key point of our review paper: to change our view of scavenging from the classic dichotomy of obligate and facultative scavengers, to a sliding scale depending on the species and the ecosystem.''}
-
+Indeed, this is a key point of our review paper: to change our view of scavenging from the classic dichotomy of obligate and facultative scavengers, to a sliding scale depending on the species and the ecosystem.''}}
+ \textcolor{red}{We have also been clearer on this point throughout the paper by reducing references to this division between obligate and facultative scavengers were appropriate and instead referring to the gradient of scavenging outlined in the introduction}.
+%Is this point true
 \bigskip
@@
 
-\textcolor{blue}{It is undoubtedly true that scavengers can be variable in the proportion of carrion they consume but we would argue that species like hyenas have certain characteristics that make this possible.
+\textcolor{blue}{It is undoubtedly true that scavengers can be variable in the proportion of carrion they consume but we would argue that species like hyenas have certain characteristics that make this capacity to scavenge  a high proportion of their diet possible.
 In contrast, cheetahs do not have a biology that lends itself to scavenging in the wild, due to poor competitive ability etc.
-We appreciate that there is a continuum across and among species which we specifically note in this revised version. 
+We appreciate that there is a continuum across and among species which we specifically note in this revised version \textcolor{red}{(page lines xxx)}. 
 As we note in the above comment, a combination of a species' traits, and the state of the ecosystem they find themselves in will determine their scale of scavenging. 
@@
 So, far from invalidating our idea it reinforces it.
-Moreover, researchers working in palaeobiology often try to reconstruct the behaviours of extinct species; we are following this approach when we make inferences about ancient systems.}
+Moreover, researchers working in palaeobiology often try to reconstruct the behaviours of extinct species, \textcolor{red}{which themselves are likely to be extremely variable}; we are following this approach when we make inferences about ancient systems were we \textcolor{red}{are interested in determining whether a species has the capacity to scavenge like a Hyena or is more restricted in this ability such as a cheetah}}
 \bigskip
@@
 
+\textcolor{red}{We agree it is more difficult to infer behavior with lower quality data, however by mapping out the variables that are likely to driver high levels of scavenging, the absence of nuanced measures will highlight the extra caution needed when inferring the existence of scavenging in such a circumstance}
 \textcolor{blue}{Figure 2 has been removed and figure 1 has been expanded.}
@@
 %TG: They make a point here (although their diversity definition rant is completely out of topic). Should we just drop figure 2 and use it as an abstract/press release picture? AK Agreed
+
 \bigskip
@@
 \textcolor{blue}{The issue of the structure of our paper was noted by all reviewers who all had suggestions to improve it. 
-We have made an effort to delineate all of the main sections and subsections so that our frame of optimal foraging is more apparent.
-We have added the two suggested references to our paper.}
+We have made a significant effort to delineate all of the main sections and subsections so that our frame of optimal foraging is more apparent.
+We have also added the two suggested references to our paper.}
 
@@
 In sum, while I think the aim of the article is a very worthwhile one, I think the authors cannot claim to have produced an all-encompassing model that can predict the presence or level of scavenging in any organism - because I think this simply cannot be done.
+
+\textcolor{red}{We agree, and do not intend on our framework to act as an all encompassing model of scavenging but instead as a guide to determine a range of possible sceneries with different levels of scavenging.
+We highlight throughout the manuscript the requirement to use other methods such as energetic models and comparative anatomy in order to make accurate inferences on behaviour.}
 
@@
 These are very broad areas so in a review of this sort some overlap is inevitable.
-We feel our inclusion of extinct species and systems as well as the novel framing under optimal foraging theory gives this manuscript its novelty.} 
+We feel our inclusion of extinct species and systems as well as the novel framing under optimal foraging theory gives this manuscript its novelty.} \textcolor{red}{We have paid particular attention towards making the structure or our manuscript more clear and hence delineating it from Beasley et al.
+(2015).
 \bigskip
@@
 
+%KHwhat line are the mentions of invertebrates on. Do insects help facilitate through breaking open carcasses or anything?
+
 \item{I found difficult to follow the main point of the paper because the ms is not well organized.
@@
 %TG: + more blabalal on his ``Facilitation'' part
+
+%%KH I think a bit more needs to be added to this section.
+Something as simple "While such facilitation is difficult to detect in extinct species the presence of facilitatory species may indicate the possibility of the interaction.
+For example, early Hominds may have benefited from the presence of vultures, and in fact many  scavenging species now depend on human refuse as a means of scavenging in current systems.
+Any other examples?"
 

### 2016-08-10 natural history of scavenging.tex
Renamed composite figure folder without a space!
comments on response letter

@@
 Even vultures, the canonical example of obligate scavengers, can hunt \citep{margalida2011scavenger}.
-While it is relatively easy to determine whether a species engages in scavenging through observation, determining the proportion of carrion in the diet requires extensive behavioral data.
-% (I put this in to placate the reviewer who said it was easy).
+While it is relatively easy to determine whether a species engages in scavenging through observation, determining the proportion of carrion in the diet requires extensive behavioral data \textcolor{red}{as approaches such as stomach contents and stable isotope analysis are unable to distinguish between scavenging and predation.} % (I put this in to placate the reviewer who said it was easy).
 Unfortunately, the lack of direct behavioural data makes it even more difficult to discern scavenging from predation among extinct forms where it was surely exhibited by prehistoric vertebrate carnivores.
@@
 \centering
-   \includegraphics[width=1\textwidth]{Composite figure/Compositefig4.pdf}
+   \includegraphics[width=1\textwidth]{CompositeFigure/Compositefig4.pdf}
 \caption{Factors influencing the proportion of scavenging in a vertebrates' diet.

### 2016-08-11 response_letter.tex
Kev comments included

@@
 (2003) and argue our case for what we consider inevitable overlap with Beasley et al.
-(2015).}
+(2015).
+We now more carefully explain how our qualitative framework can be used to understand scavenging in extinct species.} 
 %KH I feel this is a bit weak.
@@
 
-\textcolor{red}{We feel we have now more clearly outlined the structure of the paper by including an updated figure one explicitly showing the breakdown of the chapters and their link to foraging ecology.
+\textcolor{blue}{We feel we have now more clearly outlined the structure of the paper by including an updated figure one explicitly showing the breakdown of the chapters and their link to foraging ecology.
 This also more clearly highlights the novel aspect of our review, of building a qualitative framework base on foraging ecology to identify the potential levels of scavenging in extinct species and systems and pointing towards possible methods to better focus on certain systems, such as energetic models.}
@@
 
-\textcolor{blue}{We emphasise that food webs have tended to lump all carnivory into predation and it is with extinct species that we have no observational evidence.}
- \textcolor{red}{We now note measures often used in dietary analysis, such as stomach contents and stable isotope analysis, are unable to distinguish between scavenging and predation.
+\textcolor{blue}{We emphasise that food webs have tended to lump all carnivory into predation and it is with extinct species that we have no observational evidence.
+We now note measures often used in dietary analysis, such as stomach contents and stable isotope analysis, are unable to distinguish between scavenging and predation.
 We emphasis that the subsequent need for extensive behavioral data has led to food webs have tending to lump all carnivory into predation and this is particularly likely with extinct species.
@@
 
-\textcolor{blue}{We now emphasise that our scale is qualitative here and throughout the manuscript.\textcolor{red}{ in lines (----)}
+\textcolor{blue}{We now emphasise that our scale is qualitative here and throughout the manuscript.}%\textcolor{red}{ in lines (----)}
 
@@
 that the scavenging gradient allows us to infer scavenging in the fossil record as well as in modern species.
-In making this revision we have reduced the \textcolor{red}{inevitable} overlap with De Vault (2003) considerably.} %TG:
+In making this revision we have considerably reduced the inevitable overlap with De Vault (2003).} %TG:
 
@@
 \textcolor{blue}{We now specify the approaches we have in mind as \textit{``e.g.
-energetics models, comparative anatomy, palaeontology etc.''}} \textcolor{red}{on lines 22-24 with references referring to these methods.}
+energetics models, comparative anatomy, palaeontology etc.''} on lines 22-24 with references referring to these methods.}
 
@@
 \textcolor{blue} {We make the following point in our introduction: \textit{``Indeed, there is no discrete divide between predators and scavengers but rather a continuous gradient.
-Even vultures, the canonical example of obligate scavengers, can hunt.''} We have also been clearer on this point throughout the paper by removing references to this division \textcolor{red}{and instead referring to the gradient of scavenging outlined in the introduction}.}
+Even vultures, the canonical example of obligate scavengers, can hunt.''} We have also been clearer on this point throughout the paper by removing references to this division.
+Instead we refer to the gradient of scavenging outlined in the introduction.}
 
@@
 Rather than taking them as fact we offer these as possible behaviours that the extinct species engaged in.
-\textcolor{red}{Our paper proposes these sets of possible scenarios of scavenging to allow for more quantitative tools, such a energetic models, to rule out or support these each scenario leading to a more refined understanding of scavenging in extinct species.}
+Our paper proposes these sets of possible scenarios of scavenging to allow for more quantitative tools, such as energetics models, to be used to rule out or support each scenario.
+This will help us develop a more refined understanding of scavenging in extinct species.
 We have also added extra references in support of our claims e.g.
@@
 
-\textcolor{blue} {This was also raised by reviewer 1. and we thank both reviewers for pointing out this aspect which was unclear in the original manuscript.
+\textcolor{blue} {This was also raised by reviewer 1 and we thank both reviewers for pointing out this aspect which was unclear in the original manuscript.
 We now make the following point in our introduction: \textit{``Indeed, there is no discrete divide between predators and scavengers but rather a continuous gradient.
-Even vultures, the canonical example of obligate scavengers, can hunt.} We have tried to be clearer throughout the paper as well.
-Indeed, this is a key point of our review paper: to change our view of scavenging from the classic dichotomy of obligate and facultative scavengers, to a sliding scale depending on the species and the ecosystem.''}}
- \textcolor{red}{We have also been clearer on this point throughout the paper by reducing references to this division between obligate and facultative scavengers were appropriate and instead referring to the gradient of scavenging outlined in the introduction}.
-%Is this point true
+Even vultures, the canonical example of obligate scavengers, can hunt.
+We have tried to be clearer throughout the paper as well.
+Indeed, this is a key point of our review paper: to change our view of scavenging from the classic dichotomy of obligate and facultative scavengers, to a sliding scale depending on the species and the ecosystem.''}
+We have also been clearer on this point throughout the paper by reducing references to this division between obligate and facultative scavengers where appropriate and again refer to the gradient of scavenging outlined in the introduction}.
+%Is this point true. AK Yep! 
 \bigskip
@@
 In contrast, cheetahs do not have a biology that lends itself to scavenging in the wild, due to poor competitive ability etc.
-We appreciate that there is a continuum across and among species which we specifically note in this revised version \textcolor{red}{(page lines xxx)}. 
+We appreciate that there is a continuum across and among species which we specifically note in this revised version. 
 As we note in the above comment, a combination of a species' traits, and the state of the ecosystem they find themselves in will determine their scale of scavenging. 
 Search rates can be altered by environmmetal or ecological conditions as well as intrinsic traits of the forager. 
-So, far from invalidating our idea it reinforces it.
-Moreover, researchers working in palaeobiology often try to reconstruct the behaviours of extinct species, \textcolor{red}{which themselves are likely to be extremely variable}; we are following this approach when we make inferences about ancient systems were we \textcolor{red}{are interested in determining whether a species has the capacity to scavenge like a Hyena or is more restricted in this ability such as a cheetah}}
+So, far from invalidating our idea, it reinforces it.
+Moreover, researchers working in palaeobiology often try to reconstruct the behaviours of extinct species, which themselves are likely to be extremely variable; we are following this approach when we make inferences about ancient systems.}
 \bigskip
@@
 
-\textcolor{red}{We agree it is more difficult to infer behavior with lower quality data, however by mapping out the variables that are likely to driver high levels of scavenging, the absence of nuanced measures will highlight the extra caution needed when inferring the existence of scavenging in such a circumstance}
+\textcolor{blue}{We agree it is more difficult to infer behavior with lower quality data, however by mapping out the variables that are likely to driver high levels of scavenging, the absence of nuanced measures will highlight the extra caution needed when inferring the existence of scavenging in such a circumstance}
 \textcolor{blue}{Figure 2 has been removed and figure 1 has been expanded.}
@@
 
-\textcolor{red}{We agree, and do not intend on our framework to act as an all encompassing model of scavenging but instead as a guide to determine a range of possible sceneries with different levels of scavenging.
+\textcolor{blue}{We agree, and do not intend on our framework to act as an all encompassing model of scavenging but instead as a guide to determine a range of possible sceneries with different levels of scavenging.
 We highlight throughout the manuscript the requirement to use other methods such as energetic models and comparative anatomy in order to make accurate inferences on behaviour.}
@@
 These are very broad areas so in a review of this sort some overlap is inevitable.
-We feel our inclusion of extinct species and systems as well as the novel framing under optimal foraging theory gives this manuscript its novelty.} \textcolor{red}{We have paid particular attention towards making the structure or our manuscript more clear and hence delineating it from Beasley et al.
-(2015).
+We feel our inclusion of extinct species and systems as well as the novel framing under optimal foraging theory gives this manuscript its novelty.
+We have paid particular attention towards making the structure of our manuscript more clear and hence differentiating it from Beasley et al.
+(2015).}
 \bigskip

### 2016-08-11 natural history of scavenging.tex
Kev comments included

--- 2016-08-09 natural history of scavenging.tex
@@
 While it is relatively easy to determine whether a species engages in scavenging through observation, determining the proportion of carrion in the diet requires extensive behavioral data.
-% (I put this in to placate the reviewer who said it was easy).
+This is because approaches such as stomach content and stable isotope analysis are unable to distinguish between scavenging and predation.} % (I put this in to placate the reviewer who said it was easy).
 Unfortunately, the lack of direct behavioural data makes it even more difficult to discern scavenging from predation among extinct forms where it was surely exhibited by prehistoric vertebrate carnivores.
-Some simple heuristics can inform us whether or not the prey was scavenged, for instance, in cases where it was simply too large to have been killed by the ostensible predator \citep{pobiner2008paleoecological}.}
+Some simple heuristics can inform us whether or not the prey was scavenged, for instance, in cases where it was simply too large to have been killed by the ostensible predator \citep{pobiner2008paleoecological}.
 But clearly, a scavenger does not only feed on dead animals too big for it to have hunted.
@@
 
-\begin{figure}[!htbp]
-\centering
-   \includegraphics[width=1\textwidth]{timeline_figure/timeLine.pdf}
-\caption{The diversity of scavengers through time.
+%\begin{figure}[!htbp]
+%\centering
+%   \includegraphics[width=1\textwidth]{timeline_figure/timeLine.pdf}
+%\caption{The diversity of scavengers through time.
 Each species has either direct evidence for it being a scavenger or would be positioned high up on our scavenging scale.
@@
 Time is represented on the horizontal axis with each third representing respectively the Paleozoic (541 - 252 Mya), the Mesozoic (252 - 66 Mya) and the Cenozoic (66 Mya to the present).}
-\label{Timeline}
-\end{figure}
+%\label{Timeline}
+%\end{figure}
 

### 2016-08-11 response_letter.tex
small changes to the response letter

@@
 We now note measures often used in dietary analysis, such as stomach contents and stable isotope analysis, are unable to distinguish between scavenging and predation.
-We emphasis that the subsequent need for extensive behavioral data has led to food webs have tending to lump all carnivory into predation and this is particularly likely with extinct species.
-(lines 53-56)}
+We emphasis that the subsequent need for extensive behavioral data has led to food webs tending to lump all carnivory into predation and this is particularly likely with extinct species.
+(page 1, lines 11-16)}
 
@@
 
-\textcolor{blue}{The hierarchy of our paper wasn't clear as we intended in the first version but thanks to the review process we have now clarified that this is the lead in paragraph to a larger section which has subsections e.g.
+\textcolor{blue}{The hierarchy of our paper wasn't clear as we intended in the first version but thanks to the review process we have now clarified that this is the opening paragraph to a larger section which has subsections e.g.
 food processing (see also updated figure 1).}
@@
 
-\textcolor{blue}{We agree it is more difficult to infer behavior with lower quality data, however by mapping out the variables that are likely to driver high levels of scavenging, the absence of nuanced measures will highlight the extra caution needed when inferring the existence of scavenging in such a circumstance}
+\textcolor{blue}{We agree it is more difficult to infer behavior with lower quality data.
+However by mapping out the variables that are likely to driver high levels of scavenging, such as in Figure 1, the absence of data for particular variblese, sfor example, will highlight the extra caution needed when inferring the existence of scavenging in such a particular circumstance.}
 \textcolor{blue}{Figure 2 has been removed and figure 1 has been expanded.}
